### 2015-12-31 netft.tex
Initial commit.

\documentclass[twocolumn]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{breqn}

\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}

\title{DRAFT: Robust Communication in Structured Networks}
\author{
    Edward L. Platt\\
    elplatt@umich.edu
    \and
    Daniel M. Romero\\
    drom@umich.edu
}
\date{December 2015}

\begin{document}

\maketitle

\section{Introduction}

\section{Trust Model}

\section{Fault Tolerance}

\section{Butterfly Network}

\section{Discussion}

\section{Conclusion}

\bibliographystyle{plain}
\bibliography{netft}

\end{document}

### 2015-12-31 netft.tex
Finish trust model.

@@
 
+We assume that a sender $A$ (or Alice) wants to send sensitive information
+to a receiver $B$ (or Bob).
+Alice and Bob are both nodes on a communication network represented by a graph
+$G = (V, E)$.
+Two nodes are connected when they are able to communicate directly with each
+other.
+We also assume the presence of an adversary $C$ (or Eve) who wishes to disrupt
+that communication by altering or delaying the transmission.
+Eve has compromised a subset of nodes and has full control over their behavior.
+Any transmission routed through a compromised node is considered faulty.
+Alice and Bob do not know which nodes have been compromised.
+
+In our model, links also represent trust.
+Specifically, two nodes are connected if each node believes that the other
+has not been compromised by its adversary.
+In other words, a communication link is severed if one node believes the other
+to be compromised.
+If trust were transitive, a connected path from Alice to Bob, would guarantee
+a trusted channel, but trust is not transitive \cite{christianson_why_1997}.
+Rather than entirely abandoning transisitivy, we assume that trust is
+partially transitive.
+Specifically each node has a {\em trusted neighborhood} of nodes up to $h$ hops
+away, which we assume have not been compromised by that node's adversaries.
+Our objective is now to construct a robust communication channel from Alice's
+trusted neighborhood to Bob's.
+
 \section{Fault Tolerance}

### 2015-12-31 netft.tex
Finish section Multipath Fault Tolerance.

@@
 \newtheorem{theorem}{Theorem}
+
+\newcommand{\beq}{\begin{eqnarray}}
+\newcommand{\eeq}{\end{eqnarray}}
 
@@
 
-\section{Fault Tolerance}
+\section{Multipath Fault Tolerance}
+
+We achieve robust communication between trusted neighborhoods through
+Concurrent Multipath Routing (CMR) \cite{}.
+If there are multiple {\em redundant paths} (i.e., paths sharing no nodes or
+links) between Alice and Bob's trusted neighborhoods, multiple copies of a
+message can be sent along different paths, enabling standard error detection and
+error correction techniques to be applied at the destination.
+If Alice randomly selects $k$ of $n$ total redundant paths, Eve's best strategy
+is to compromise at least one node on as many of the paths as possible.
+Because an error can be detected if even a single copy has been tampered with,
+Eve must compromise all $k$ paths in order to convince Bob to accept a compromised
+message.
+If Eve has the resources to compromise $l$ total nodes, the probability that
+Bob receives an undetectable error is:
+\beq
+\label{eq:pe}
+p_e &=& \frac{l!(n-k)!}{n!(l-k)!}.
+\eeq
+Letting $k=\alpha n$ and $l=\beta n$, then applying Stirling's
+approximation gives:
+\begin{eqnarray}
+\nonumber
+p_e &\approx&
+\frac{\sqrt{\beta(1-\alpha)}}{\sqrt{\beta-\alpha}} \\
+& &
+\times {}
+\label{eq:pe_approx}
+\left[
+    \left( \frac{\beta-\alpha}{1-\alpha} \right)^{\alpha}
+    \left( \frac{\beta}{\beta-\alpha} \right)^{\beta}
+    (1-\alpha)
+\right]^n
+\end{eqnarray}
+Note that Eq. (\ref{eq:pe_approx}) is exponential in the number of redundant
+paths $n$, which depends only on the network structure, and not on the
+resources available to individual nodes.
+If a network has a large number of redundant paths between trusted neighborhoods,
+a very low probability of undetected errors can be achieved, even if an adversary
+has the resources to compromise a very large fraction of the avialable paths.
 

### 2016-01-04 netft.tex
Add figure to section multipath fault tolerance.
Add figure showing butterfly network connection rule.
Finish routing subsection.

@@
 \usepackage{amsthm}
+\usepackage{amsfonts}
 \usepackage{graphicx}
@@
 
-\title{DRAFT: Robust Communication in Structured Networks}
+\DeclareMathOperator{\wbf}{W-BF}
+
+\title{DRAFT: Scalable, Robust Multipath Communication on the Butterfly Network}
 \author{
@@
 }
-\date{December 2015}
+\date{January 2016}
 
@@
 has the resources to compromise a very large fraction of the avialable paths.
+The value of $p_e$ for various $k$, $l$, and $n$ is shown in Figure
+\ref{fig:perror}.
+
+\begin{figure}
+\includegraphics{fig-perror.pdf}
+\caption{
+Probability $p_e$ of an undetected error as a function of the number of message
+copies $k$ and compromised paths $l$.
+\label{fig:perror}
+}
+\end{figure}
 
 \section{Butterfly Network}
+
+We now describe a Concurrent Multipath Routing scheme on a specific family
+of networks: the butterfly network \cite{}.
+Several variations on the butterfly network exist.
+Specifically, we utilize the wrap-around butterfly.
+We define the $m$-dimensional directed wrap-around butterfly $\wbf(m)$:
+\beq
+\wbf(m) &=& (V, E_\downarrow \cup E_\rightarrow) \\
+V &=& \mathbb{Z}_m \times \mathbb{Z}_{2^m} \\
+E_\downarrow &=& \{((l,z),(l+1 (\text{mod } m),z) \nonumber \\
+&& \; | \, l \in \mathbb{Z}_d, z \in \mathbb{Z}_{2^m}\} \\
+E_\rightarrow &=& \{(l,z),(l+1 (\text{mod } m), z \oplus 2^l) \nonumber \\
+&& \; | \, l \in \mathbb{Z}_d, z \in \mathbb{Z}_{2^m}\},
+\eeq
+where $\oplus$ represents the bitwise XOR operator.
+Each node is associated with a level $l$ and an $m$-bit integer $z$.
+There are two types of edges, shown in Figure \ref{fig:butterfly}.
+Down edges ($E_\downarrow$) connect nodes sharing the same $z$ value
+in a cycle of increasing level $l$.
+Down-right edges ($E_\rightarrow$) also link to a node of level $l + 1$,
+but one having the bitstring equal to $z$ with the $l$th bit flipped.
+
+The wrap-around butterfly network has a number of properties making it useful for multipath routing:
+\begin{description}
+\item[Vertex transitivity]
+The problem
+of finding a route between arbitrary nodes $v$ and $w$
+can be reduced to finding a route from node $(0,0)$ to some $\tilde{w}$.
+\item[Logarithmic diameter]
+For any two nodes, the length of the shortest path between them is logorithmic
+in the size of the network.
+\item[Constant degree]
+In practical applications, each communication link requires additional resources,
+such as physical infrastructure or entries in a routing table.
+A constant degree limits the number of such resources needed as the network
+grows in size.
+\end{description}
+
+\begin{figure}
+\begin{center}
+\includegraphics{fig-butterfly.pdf}
+\end{center}
+\caption{
+Schematic illustration of the two types of edges in a directed butterfly network.
+The node $(l,z)$ is shown as the bit string $z$ with a square around the $l$th bit.
+\label{fig:butterfly}
+}
+\end{figure}
+
+\subsection{Multipath Routing on the Butterfly Network}
+
+We now describe a routing scheme that provides $2^h$ redundant paths between
+the $h$-hop trusted neighborhoods of any two nodes in an $m$-bit
+wrap-around butterfly network.
+Utilizing vertex transitivity, we label the source node as $(0, 0)$ and
+denote the destination node as $w = (l_w, z_w)$.
+
+Let $s$ be an integer such that $0 \leq s < 2^h$.
+Let $v_s^{(t)} = (l^{(t)}, z^{(t)})$ be the $t$th node in the path labeled by $s$.
+For convenience, we will omit the subscript $s$.
+We define two partitionings of the integers in $\mathbb{Z}_{2^m}$:
+one having the lowest $h$ bits matching the
+bits of $s$, and one having the $h$ bits preceeding the destination level $l_w$
+matching $s$:
+\beq
+S_s &=& \{z \in \mathbb{Z}_{2^m} | \forall i \in \mathbb{Z}_h z_i = s_i \} \\
+R_s &=& \{z \in \mathbb{Z}_{2^m} | \forall i \in \mathbb{Z}_h z_{(l_w - h + i)} = s_i \}.
+\eeq
+Note that if $r \neq s$, then $S_s \cap S_r = R_s \cap R_r = \emptyset$.
+
+We now construct a path such that between trusted neighborhoods $z^{(t)}$ is always
+in $S_s$, $R_s$, or both, guaranteeing that the path does not overlap with the
+other paths $v_r^{(t)}$.
+Routing proceeds in stages, with the level $l$ increasing by 1 at each hop.
+In Stage 1 ($0 \leq t < h$), down or down-right edges
+are chosen such that the the $t$th bit of $z^{(t+1)}$ is equal to the $t$th bit
+of $s$. Throughout Stage 1, all nodes are within the sender's trusted neighborhood.
+At the end of Stage 1, $z^{(h)} \in S_s$, and $z^{(t)}$ will remain so until the level loops
+back to $0$ at $t = m$.
+
+In Stage 2 ($h \leq t < l_w - h$), edges are chosen to make the $t$th bit of
+$z^{(t+1)}$ match the $t$th bit of $z_w$.
+
+In Stage 3 ($l_w - h \leq t < l_w$), the bits of $z^{(t)}$ are chosen to match $s$,
+such that after the stage is complete, $z^{(l_w)} \in R_s$.
+
+In Stage 4 ($l_w \leq t < m$), as in stage 2,
+paths are chosen such that the $t$th bit of $z^{(t+1)}$ matches $z_w$.
+After Stage 4, all bits of $z^{(m)}$ are equal to those of $z_w$
+except for the first $h$ and the $h$ preceeding index $l_w$.
+$z^{(m)}$ is also in both $S_s$ and $R_s$.
+
+At this point, we define $\tau = t - m$.
+In Stage 5 ($0 \leq \tau < h$), the first $h$ bits of $z^{(t)}$ are set to
+match $z_w$, potentially removing $z^{(t)}$ from $S_s$.
+
+In Stage 6 ($h \leq \tau < l_w - h$), all down edges are chosen, incrementing
+the level without any effect on $z^{(t)}$.
+At the end of Stage 6, $z^{(m + l_w - h)}$ is still in $R_s$ and
+$v^{(m + l_w - h)}$ is now within the trusted neighborhood of $w$.
+
+In the seventh, and final stage ($l_w - h \leq \tau < l_w$), the $h$ bits of $z^{(t)}$
+preceeding index $l_w$ are set to match $z_w$.
+After this stage, $v^{(m + l_w)} = w$ and routing is complete.
 

### 2016-02-18 ec-2016/source/paper.tex
Add partial draft of EC16 writeup.

%\documentclass[prodmode]{acmsmall-ec16}
%\documentclass[prodmode,license]{acmsmall-ec16}
\documentclass[prodmode,permissions]{acmsmall-ec16}

% Package to generate and customize Algorithm as per ACM style
\usepackage[ruled]{algorithm2e}
\renewcommand{\algorithmcfname}{ALGORITHM}
\usepackage{amsmath,amsfonts,amssymb,bbm} 
\usepackage[numbers,sort&compress]{natbib} % for citet
\SetArgSty{textrm}  % for algorithm2e
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

%\conferenceinfo{EC'16,}{July 24--28, 2016, Maastricht, The Netherlands.}
%\CopyrightYear{2016}
%\crdata{978-1-4503-3936-0/16/07}
%\copyrighttext{Copyright is held by the owner/author(s). Publication rights licensed to ACM.}

\doi{XXXXXXX.XXXXXXX}% TeXSupport

\newcommand{\beq}{\begin{eqnarray}}
\newcommand{\eeq}{\end{eqnarray}}
\DeclareMathOperator{\wbf}{wBF}

% Document starts
\begin{document}

% Page heads
%\markboth{G. Zhou et al.}{A Multifrequency MAC Specially Designed for WSN Applications}

% Title portion
\title{
Attack-Tolerance in Structured Networks via Multipath Routing
\\ Draft: 2016-02-16
}
\author{EDWARD L. PLATT
\affil{University of Michigan}
DANIEL M. ROMERO
\affil{University of Michigan}}

\begin{abstract}
TODO
\end{abstract}

%ACM is moving forward with the 2012 Classification system: http://dl.acm.org/ccs_flat.cfm. Please generate the CCSXML tex code through the online interactive system and insert the code below.

\begin{CCSXML}
<ccs2012>
</ccs2012>  
\end{CCSXML}

\ccsdesc[500]{TODO~TODO}

%\terms{Design, Algorithms, Performance}

\keywords{
censorship,
decentralization,
fault tolerance,
multipath,
networks,
peer-to-peer,
trust
}

\begin{bottomstuff}
This work is supported by TODO.

Author’s addresses: E.L. Platt and D.M. Romero, School of Information; email: \{elplatt,drom\}@umich.edu.
\end{bottomstuff}

\maketitle

TODO: Fix bibtex encoding.

\section{Introduction}

Communication networks, expemplified by the Internet,
have become ubiquitous and critical infrastructural for
communities, organizations, and markets.
As with any critical infrastructure, the cost of a failure can be
immense, so methods for tolerating various kind of faults are an
important and ongoing area of research.
The key question is: what are the techniques and network structures that can
be used to create robust, fault-tolerant systems?

Many complex networks, including the Internet and World-Wide Web, exhibit
scale-free structure
\cite{barabasi_emergence_1999,barabasi_scale-free_2009}.
While scale-free networks tolerate random faults well,
they are highly susceptible to adversarial faults, i.e. attacks
\cite{albert_error_2000}.
For example, in 2008, YouTube suffered a worldwide outage for several hours
when a service provider in Pakistan advertised false routing information
\cite{hunter_pakistan_2008}.
The action (known as a {\em black hole attack}) was intended to censor YouTube
within Pakistan only, but resulted in a worldwide cascading failure.
Such vulnerabilities are not limited to any one system or protocol, but an
attribute of complex communication networks themselves.
General techniques for understanding and mitigating these vulnerabilities are
needed.

In this paper, we formally evaluate the effects of network structure on
attack-resistance. We also propose an efficient, fault-tolerant network architecture.
Our analysis focuses on {\em structured networks},
in which links between nodes are constrained to a particular architecture.
The proposed fault-tolerant architecture utilizes {\em multipath routing},
in which many possible paths exist betweeen two nodes on a communication network.
In our case, multiple paths allow multiple copies of a message to be sent,
which can be used for error-detection and error-correction by the receiver.

Our main contributions are:
\begin{itemize}
\item{We present a formal {\em partial trust model},
which makes weaker transitivity assumptions than previous web-of-trust models,
and use this model to quantify the effects of network structure on
fault tolerance;}
\item{We show that the probability of successfully detecting an
adversarial fault increases exponentially with the number of disjoint untrusted
paths between the neighborhoods of the sender and receiver, and that this
value depends on network structure;}
\item{We present a scalable, efficient, and attack-tolerant multipath
routing algorithm on the butterfly network topology.}
\end{itemize}

Paper organization. TODO

\section{Background and Related Work}

In {\em structured networks}, link structure is predetermined.
Such networks can be designed to have favorable structural and routing properties,
at the expense of complicating the addition or removal of nodes.
Structured networks have been a popular tool in parallel and distributed
computing architectures \cite{kshemkalyani_distributed_2008}.
More recently, peer-to-peer systems based on distributed hash tables have used
structured ``overlay'' networks to map table keys to local TCP/IP routes
\cite{lua_survey_2005}.

Many distributed consensus protocols (such as those used by crypto-currencies)
are designed to provide tolerance against arbitrary or adversarial faults.
Byzantine agreement protocols
\cite{lamport_byzantine_1982,castro_practical_1999}
provide tolerance against arbitrary faults (including attacks) under
some circumstances, but are limited to small networks due to poor scalability.
Proof-of-work \cite{dwork_pricing_1993,nakamoto_bitcoin:_2008}
and proof-of-stake \cite{king_ppcoin:_2012}
provide better scalability, but wasteful of computational and energy
resources.
Federated Byzantine Agreement (FBA) \cite{mazieres_stellar_2015}
is scalable and optimal, guaranteeing that the protocol only fails when
success is impossible, although it does not provide a way to evaluate
the probability of failure.
All of these protocols rely on the assumption that their cryptography
cannot be compromised.
While cryptography can be extremely resistant to technological attacks,
it can still be compromised through coercion (e.g., legal action).
In additon, the fault-tolerance of FBA depends on redundancy in the network structure
but leaves an unanswerd question: how should a network be structured to achieve
fault tolerance?
This paper addresses both of these issues by analyzing the role of network
structure on adversarial faults, without relying on assumptions of cryptographic
integrity.

{\em Multipath} routing uses multiple paths when routing a message through a
network, in contrsast to traditional {\em unipath} routing, which uses
a single path.
Multipath routing can have many benefits, including reduced congestion,
increased throughput, and more reliability
\cite{qadir_exploiting_2015}.
Many of these routing protocols offer better security 
\cite{zin_survey_2015}.
Some approaches utilize redundant paths as backups for increased
fault tolerance
\cite{alrajeh_secure_2013},
and some specificially protect against adversarial faults
\cite{kohno_improvement_2012, khalil_unmask:_2010, lou_h-spread:_2006}.
The method of Liu et al.
\cite{liu_secure_2012}
routes multiple messages first to random peers and then
to their final destination.
The butterfly algorithm we present takes a conceptually similar approach.
Most work on multipath routing has been motivated by applications related to
wireless sensor networks (WSNs),
and have thus focused on ad-hoc, unstructured networks, often having a central
base station.
The trust model presented in this paper provides a more general way to evaluate
the effect of network structure on adversarial fault-tolerance.

\section{Attack-Tolerant Networks Infrastructure}

In this section, we describe the functional proprties required of an
attack-tolerant network infrastructure and how those properties translate
into constraints on network structure.
We pay special attention to a property we call
{\em stabilizing asymmetry}.

\subsection{Functional Properties}

In most cases, infrastructure must remain functional as it grows;
it must be {\em scalable}.
Systems having single points of failure are less tolerant against faults at those
points, which both raises the likelihood of a failure and creates targets
where attackers might be able to focus their resources efficiently.
Attack-tolerant infrastructure must minimize single points of failure;
it must be {\em decentralized}.
In some, but not all, cases it is also desirable that an infrastructure
provides {\em secrecy},
protecting the contents of a message or the identity of the sender.

We add one additional property, which we call {\em stabilizing asymmetry}.
In the context of international conflict,
\cite{mack_why_1975}
observed that power imbalance usually determines the outcome of a conflict
(with the more powerful side winning)
except in the special case of {\em asymmetric conflicts}.
In asymmetric conflicts, the same level of resource expenditure yields different
results for different parties.
There are two possible cases:
the attacker's resources are either more or less effective than the defender's.
We call the latter case stabilizing asymmetry,
because it lowers the incentive to attack.
With this in mind, an attack-resistant infrastructure will benefit from a high
level of stabilizing asymmetry.

\subsection{Structural Properties}

Sparse

Low-diameter

Vertex-transitivity

Redundancy
\cite{baran_distributed_1964}

Routing

\subsection{Trust}

Scale and indirect communication

Source-to-network
\begin{figure}
\centerline{\includegraphics{fig-trust-source}}
\caption{TODO}
\label{fig:trust-source}
\end{figure}

Source-to-destination
\begin{figure}
\centerline{\includegraphics{fig-trust-source-destination}}
\caption{TODO}
\label{fig:trust-source-destionation}
\end{figure}

\section{Multipath Fault Tolerance}

Fault tolerance with partial trust and redundancy.

\subsection{Fault Tolerance}

Results

\begin{figure}
\centerline{\includegraphics{fig-perror}}
\caption{TODO}
\label{fig:one}
\end{figure}

\section{Multipath Routing on the Butterfly Topology}

We now describe a Concurrent Multipath Routing scheme on a specific family
of networks: the butterfly network \cite{}.
Several variations on the butterfly network exist.
Specifically, we utilize the wrap-around butterfly.
We denote the $m$-dimensional directed wrap-around butterfly as $\wbf(m)$:
\beq
\wbf(m) &=& (V, E_\downarrow \cup E_\rightarrow) \\
V &=& \mathbb{Z}_m \times \mathbb{Z}_{2^m} \\
E_\downarrow &=& \{((l,z),(l+1 (\text{mod } m),z) \nonumber \\
&& \; | \, l \in \mathbb{Z}_d, z \in \mathbb{Z}_{2^m}\} \\
E_\rightarrow &=& \{(l,z),(l+1 (\text{mod } m), z \oplus 2^l) \nonumber \\
&& \; | \, l \in \mathbb{Z}_d, z \in \mathbb{Z}_{2^m}\},
\eeq
where $\oplus$ represents the bitwise XOR operator.
Each node is associated with a level $l$ and an $m$-bit integer $z$.
There are two types of edges, shown in Figure \ref{fig:butterfly}.
Down edges ($E_\downarrow$) connect nodes sharing the same $z$ value
in a cycle of increasing level $l$.
Down-right edges ($E_\rightarrow$) also link to a node of level $l + 1$,
but one having the bitstring equal to $z$ with the $l$th bit flipped.

The wrap-around butterfly network has a number of properties making it useful for multipath routing:
\begin{description}
\item[Vertex transitivity]
The problem
of finding a route between arbitrary nodes $v$ and $w$
can be reduced to finding a route from node $(0,0)$ to some $\tilde{w}$.
\item[Logarithmic diameter]
For any two nodes, the length of the shortest path between them is logorithmic
in the size of the network.
\item[Constant degree]
In practical applications, each communication link requires additional resources,
such as physical infrastructure or entries in a routing table.
A constant degree limits the number of such resources needed as the network
grows in size.
\end{description}

\begin{figure}
\begin{center}
\includegraphics{fig-butterfly.pdf}
\end{center}
\caption{
Schematic illustration of the two types of edges in a directed butterfly network.
The node $(l,z)$ is shown as the bit string $z$ with a square around the $l$th bit.
\label{fig:butterfly}
}
\end{figure}

We now describe a routing scheme that provides $2^h$ redundant paths between
the $h$-hop trusted neighborhoods of any two nodes in an $m$-bit
wrap-around butterfly network.
Utilizing vertex transitivity, we label the source node as $(0, 0)$ and
denote the destination node as $w = (l_w, z_w)$.

Let $s$ be an integer such that $0 \leq s < 2^h$.
Let $v_s^{(t)} = (l^{(t)}, z^{(t)})$ be the $t$th node in the path labeled by $s$.
For convenience, we will omit the subscript $s$.
We define two partitionings of the integers in $\mathbb{Z}_{2^m}$:
one having the lowest $h$ bits matching the
bits of $s$, and one having the $h$ bits preceeding the destination level $l_w$
matching $s$:
\beq
S_s &=& \{z \in \mathbb{Z}_{2^m} | \forall i \in \mathbb{Z}_h z_i = s_i \} \\
R_s &=& \{z \in \mathbb{Z}_{2^m} | \forall i \in \mathbb{Z}_h z_{(l_w - h + i)} = s_i \}.
\eeq
Note that if $r \neq s$, then $S_s \cap S_r = R_s \cap R_r = \emptyset$.

We now construct a path such that between trusted neighborhoods $z^{(t)}$ is always
in $S_s$, $R_s$, or both, guaranteeing that the path does not overlap with the
other paths $v_r^{(t)}$.
Routing proceeds in stages, with the level $l$ increasing by 1 at each hop.
In Stage 1 ($0 \leq t < h$), down or down-right edges
are chosen such that the the $t$th bit of $z^{(t+1)}$ is equal to the $t$th bit
of $s$. Throughout Stage 1, all nodes are within the sender's trusted neighborhood.
At the end of Stage 1, $z^{(h)} \in S_s$, and $z^{(t)}$ will remain so until the level loops
back to $0$ at $t = m$.

In Stage 2 ($h \leq t < l_w - h$), edges are chosen to make the $t$th bit of
$z^{(t+1)}$ match the $t$th bit of $z_w$.

In Stage 3 ($l_w - h \leq t < l_w$), the bits of $z^{(t)}$ are chosen to match $s$,
such that after the stage is complete, $z^{(l_w)} \in R_s$.

In Stage 4 ($l_w \leq t < m$), as in stage 2,
paths are chosen such that the $t$th bit of $z^{(t+1)}$ matches $z_w$.
After Stage 4, all bits of $z^{(m)}$ are equal to those of $z_w$
except for the first $h$ and the $h$ preceeding index $l_w$.
$z^{(m)}$ is also in both $S_s$ and $R_s$.

At this point, we define $\tau = t - m$.
In Stage 5 ($0 \leq \tau < h$), the first $h$ bits of $z^{(t)}$ are set to
match $z_w$, potentially removing $z^{(t)}$ from $S_s$.

In Stage 6 ($h \leq \tau < l_w - h$), all down edges are chosen, incrementing
the level without any effect on $z^{(t)}$.
At the end of Stage 6, $z^{(m + l_w - h)}$ is still in $R_s$ and
$v^{(m + l_w - h)}$ is now within the trusted neighborhood of $w$.

In the seventh, and final stage ($l_w - h \leq \tau < l_w$), the $h$ bits of $z^{(t)}$
preceeding index $l_w$ are set to match $z_w$.
After this stage, $v^{(m + l_w)} = w$ and routing is complete.

\begin{figure}
\begin{center}
\includegraphics{fig-routing.pdf}
\end{center}
\caption{
\label{fig:routing}
}
\end{figure}

\section{Discussion}

Secrecy

Creation of the network.
Improves on web of trust.
Gives framework for determining where to build trust.

Applications
Distributed apps: storage, email, cryptocurrency, secure multiparty computation.
Wireless Sensor Networks

\section{Conclusion}

% Acknowledgments
\begin{acks}
The authors would like to thank TODO
\end{acks}

% Bibliography
\bibliographystyle{ACM-Reference-Format-Journals}
\bibliography{paper}

\end{document}

### 2016-02-19 ec-2016/source/paper.tex
Added most of Attack-Tolerant Network Infrastructure section.

@@
 
-Sparse
+To achieve scalability, networks must be {\em sparse} and have a
+{\em low diameter}.
+In practical settings, humans and devices have an upper limit on the number
+of connections they can maintain.
+In sparse networks, the number of links grows slowly as the network grows in
+size, allowing the network to scale without exceeding the nodes' capacity for
+links.
+Similarly, low-diameter guarantees that as a network grows, a short path will
+still exist between any pair of nodes.
+While low diameter guarantees a path exists,
+paths are only useful if an efficient {\em routing} algorithm exists
+to find them.
 
-Low-diameter
+Redundancy in a network can help reduce single points of failure as well as
+increase fault-tolerance.
+One measure of redundancy is given by the ratio of edges to nodes
+\cite{baran_distributed_1964}.
+A single points of failure occur when a node holds a uniquely
+central position,
+and can be eliminated by adding alternative
+paths around the node,
+which increases the network's redundancy.
+Similarly, redundancy in a network can enable fault-tolerance techniques
+that reduce the effectiveness of attacks and creating stabilizing asymmetry.
+When secrecy is desired, redundancy can help by allowing messages to be
+divided into parts and sent along different paths so even if one path is
+compromised, the whole message is not revealed.
 
-Vertex-transitivity
-
-Redundancy
-\cite{baran_distributed_1964}
-
-Routing
+Networks with {\em vertex transitivity} have the greatest protection agasint
+single points of failure.
+In vertex transitive networks, for any pair of nodes, an edge-preserving map
+always exists from one node to the other.
+In other words, all nodes occupy structurally indistinguishable positions
+in the network.
 

### 2016-02-23 ec-2016/source/paper.tex
Add date.
Revise section: introduction.

--- 2016-02-18 ec-2016/source/paper.tex
@@
 
+\usepackage{tikz}
+\usetikzlibrary{chains,decorations.pathreplacing}
 \newcommand{\beq}{\begin{eqnarray}}
@@
 Attack-Tolerance in Structured Networks via Multipath Routing
-\\ Draft: 2016-02-16
+\\ Draft: 2016-02-21
 }
@@
 
-Author’s addresses: E.L. Platt and D.M. Romero, School of Information; email: \{elplatt,drom\}@umich.edu.
+Author's addresses: E.L. Platt and D.M. Romero, School of Information; email: \{elplatt,drom\}@umich.edu.
 \end{bottomstuff}
@@
 
-Communication networks, expemplified by the Internet,
+Large-scale communication networks, exemplified by the Internet,
 have become ubiquitous and critical infrastructural for
@@
 immense, so methods for tolerating various kind of faults are an
-important and ongoing area of research.
+important and ongoing area of research
+\cite{zin_survey_2015,albert_error_2000,sterbenz_resilience_2010}.
 The key question is: what are the techniques and network structures that can
@@
 
-In this paper, we formally evaluate the effects of network structure on
-attack-resistance. We also propose an efficient, fault-tolerant network architecture.
-Our analysis focuses on {\em structured networks},
+We consider a setting in which a source node in a network attempts to route information
+to a target node by forwarding it through the links of the network. We 
+assume that some nodes in the network may be compromised by an attacker 
+and will forward the wrong information to their neighbors or
+block the information altogether.
+In order to evaluate network-based techniques for mitigating such attacks,
+we develop a decentralized {\em partial trust} model,
+similar to a web of trust
+\cite{zimmermann_official_1995,ferguson_practical_2003},
+but making weaker assumptions of trust transitivity.
+
+Given this framework, we develop and evaluate a \emph{multipath routing} technique to reduce 
+the likelihood that compromised nodes are undetected during the information 
+routing process by allowing multiple copies of a message to be sent through independent paths,
+which can be used for error-detection and error-correction by the receiver.
+
+The multipath routing technique relies on the existence and discoverability of
+multiple redundant paths between the source and the destination. Thus, network 
+structure and routing algorithms play a key role in the success of the technique.
+We formally evaluate the 
+effects of network structure on attack-resistance and show that the probability
+of undetected 
+errors decreases exponentially with the number of independent paths between 
+source and destination.
+
+The mutlipath routing technique is well-suted for {\em structured networks},
 in which links between nodes are constrained to a particular architecture.
-The proposed fault-tolerant architecture utilizes {\em multipath routing},
-in which many possible paths exist betweeen two nodes on a communication network.
-In our case, multiple paths allow multiple copies of a message to be sent,
-which can be used for error-detection and error-correction by the receiver.
+Specifically, we employ the butterfly network topology,
+used elsewhere in parallel computing \cite{kshemkalyani_distributed_2008}
+and peer-to-peer \cite{lua_survey_2005, korzun_structured_2013}
+applications.
+For this network topology,
+we propose an efficient, attack-tolerant multipath routing algorithm.
 
@@
 \begin{itemize}
-\item{We present a formal {\em partial trust model},
+\item{We develop a decentralized {\em partial trust model},
 which makes weaker transitivity assumptions than previous web-of-trust models,
@@
 fault tolerance;}
-\item{We show that the probability of successfully detecting an
-adversarial fault increases exponentially with the number of disjoint untrusted
-paths between the neighborhoods of the sender and receiver, and that this
-value depends on network structure;}
+\item{We show that as the number of disjoint paths between the neighborhoods
+of the sender and reciever grows,
+the probability of successfully detecting adversarial faults
+approaches 1 exponentially.
+Furthermore, we show that the number of disjoint paths is a function of
+network structure;}
 \item{We present a scalable, efficient, and attack-tolerant multipath
@@
 structured ``overlay'' networks to map table keys to local TCP/IP routes
-\cite{lua_survey_2005}.
+\cite{lua_survey_2005,korzun_structured_2013}.
 
@@
 
-In this section, we describe the functional proprties required of an
+In this section, we describe the functional properties required of an
 attack-tolerant network infrastructure and how those properties translate
@@
 
-In most cases, infrastructure must remain functional as it grows;
-it must be {\em scalable}.
+\begin{description}
+\item[Scalability]
+For large scale networks it is important that the infrastructure allows for the network to grow while remaining functional.
+In practice, people, devises, and connections, have limited capabilities and these limitations need to be considered as part of the design of the infrastructure.
+
+\item [Decentralization]
 Systems having single points of failure are less tolerant against faults at those
@@
 where attackers might be able to focus their resources efficiently.
-Attack-tolerant infrastructure must minimize single points of failure;
-it must be {\em decentralized}.
-In some, but not all, cases it is also desirable that an infrastructure
-provides {\em secrecy},
-protecting the contents of a message or the identity of the sender.
+Attack-tolerant infrastructure must be decentralized in order to minimize single points of failure.
 
-We add one additional property, which we call {\em stabilizing asymmetry}.
+\item[Stabilizing asymmetry]
 In the context of international conflict,
@@
 
+\item[Secrecy]
+In some, but not all cases, it is desirable that the infrastructure provides secrecy and allow a source and a target to exchange private information while still relying on intermediary nodes.
+Our proposed multipath routing technique can be adapted to provide secrecy but we do not focus on this aspect in this paper.
+
+\end{description}
 \subsection{Structural Properties}
 
-Sparse
+We now describe the structural properties of a network that allow it to achieve the functional properties described above. 
+\begin{description}
+\item [Sparsity and low diameter]
+To achieve scalability, networks must be {\em sparse} and have a
+{\em low diameter}.
+In practical settings, humans and devices have an upper limit on the number
+of connections they can maintain.
+In sparse networks, the number of links grows slowly as the network grows in
+size, allowing the network to scale without exceeding the nodes' capacity for
+links.
+Similarly, low-diameter guarantees that as a network grows, a short path will
+still exist between any pair of nodes.
+While low diameter guarantees a path exists,
+paths are only useful if an efficient {\em routing} algorithm exists
+to find them.
 
-Low-diameter
+\item [Vertex transitivity]
+In vertex transitive networks, for any pair of nodes, an edge-preserving map
+always exists from one node to the other.
+In other words, all nodes occupy structurally indistinguishable positions
+in the network. Networks with {\em vertex transitivity} are decentralized in the sense that they structure does no allow for  single points of failure.
 
-Vertex-transitivity
+\item [Redundancy]
+Redundancy in a network refers to the extent to which nodes are connected to each other by multiple independent paths. Redundancy can help reduce single points of failure as well as
+increase fault-tolerance.
+One measure of redundancy is given by the ratio of edges to nodes
+\cite{baran_distributed_1964}.
+A single point of failure occurs when a node holds a uniquely
+central position,
+and can be eliminated by adding alternative
+paths around the node,
+which increases the network's redundancy.
+Similarly, as we will later show, redundancy in a network can enable fault-tolerance techniques
+that reduce the effectiveness of attacks and creating stabilizing asymmetry.
+When secrecy is desired, redundancy can help by allowing messages to be
+divided into parts and sent along different paths so even if one path is
+compromised, the whole message is not revealed \cite{}.
 
-Redundancy
-\cite{baran_distributed_1964}
-
-Routing
+\end{description}
 
@@
 
-Scale and indirect communication
+The sparsity requirement has an important consequence for network infrastructure:
+as a network grows, the number of nodes will exceed the number of links a given
+node can support and many pairs of nodes will only be indirectly connected,
+with all paths between them passing through at least one intermediary node.
+These intermediary nodes are targets for potential attacks.
+More specifically, the threat model is as follows: Alice sends a message to
+Bob via one or more intermediaries, but some intermediaries may try to
+either block or alter the message.
+The ability to tolerate these types of adversarial attacks depends on
+Alice and Bob's ability to assume that some nodes have not been
+compromised---to {\em trust} them.
+
+Both centralized and decentralized approaches are commonly used to create
+trust infrastructures.
+Centralized approaches such as {\em public key infrastructure} (PKI)
+suffer from a number of vulnerabilities
+\cite{ellison_ten_2000},
+which stem largely from the single points of failure inherent to
+centralization.
+Alternatively, the decentralized {\em web of trust} model
+\cite{zimmermann_official_1995,ferguson_practical_2003}
+allows individuals
+to choose who they trust initially.
+Trust is then extended to new individuals if they are vouched for by a
+currently-trusted individual.
+However, the web of trust approach assumes that trust can be extended
+transitively,
+which is not generally true
+\cite{christianson_why_1997}.
+We note that while such approaches are decentralized in the sense that no central authority holds the keys,
+they do not necessarily incorporate redundancy and the network they form may not be vertex transitive. Thus they still may suffer from single
+points of failure.
+
+\section{Partial Trust}
+
+We now propose a decentralized trust model based on weaker transitivity
+assumptions than existing web of trust models.
+In following sections, we will use this model along with network redundancy to minimize vulnerabilities due to single points of failure.
+Our model is based on the assumption of partially-transitive trust.
+While this assumption is still idealized, it is a more realistic improvement
+on fully-transitive models.
+
+Our model (Figure \ref{fig:trust-source})
+is represented as an undirected graph,
+with nodes representing individual
+agents and edges representing mutual trust relationships.
+We define a {\em trust radius} $h$ and assume that trust is transitive
+up to $h$ hops away.
+In other words, the nodes $v$ and $w$ trust each other if the distance between $v$ and $w$, $d(u,v)$ is less than $h$.
+For a given node $v$, we call the set of nodes that $v$ trusts the {\em trusted neighborhood} $T_h(v)$,
+and all nodes at exactly distance $h$ from $v$ the {\em trust boundary} $B_h(v)$:
+\beq
+T_h(v) &=& \left\{ w \mid d(v,w) < h \right\} \\
+B_h(v) &=& \left\{ w \mid d(v,w) = h \right\}.
+\eeq
+We assume that $v$ is very familiar with nodes in its trusted neighbor and $v$ would know if any of these nodes were compromised by an attacher.
+Hence, an adversary of $v$ has no incentive to compromise a node in  $T_h(v)$.
+We thus assume that no nodes within the trusted neighborhood are compromised
+by an adversary of $v$. However, by compromising all nodes in the trust boundary, an adversary can
+ensure that all communications leaving the trusted neighborhood are compromised.
 
@@
 \begin{figure}
-\centerline{\includegraphics{fig-trust-source}}
-\caption{TODO}
+\centerline{\includegraphics[height =.4 \textwidth]{fig-alice_trusted_neigh2}}
+\caption{
+Illustration of the {\em partial trust} model.
+Edges represent direct, mutual trust relationships.
+Alice ($v$) trusts all nodes less than
+$h$ hops away---her {\em trusted neighborhood} $T_h(v)$.
+Beyond that, the nodes at distance $h$ form her {\em trust boundary} $B_h(v)$.
+We assume adversaries are unable to compromise nodes in the trusted neighborhood.
+By compromising all nodes in the trust boundary, an adversary can ensure that
+all communications leaving Alice's trusted neighborhood are compromised.
+}
 \label{fig:trust-source}
@@
 
-Source-to-destination
+Now we consider a source $v$ (Alice) and a destination $w$ (Bob),
+shown in Figure \ref{fig:trust-source-destionation}.
+We also consider an adversary who aims at disrupting all possible communication paths between Alice and Bob.
+Since the adversary cannot compromise any nodes in the trusted neighborhoods of Alicia or Bob, the only option is to compromise nodes outside of the trust neighborhoods.
+An adversary could, for example, compromise all possible paths between Alice and Bob by compromising either Alice or Bob's entire trust boundary.
+This would require compromising $\min(|T_h(v)|, |T_h(w)|)$ nodes.
+In a vertex transitive network, these two trust boundaries will be the same size.
+Alternatively, the adversary could try to compromise a small set of nodes such that all paths between Alice and Bob pass through one of these nodes.
+However, If disjoint paths exist connecting each node in Alice's trust boundary to
+a node in Bob's, then the cut size between the two can never be smaller than
+$|T_h(v)| = |T_h(v)|$, so an adversary who compromises fewer nodes cannot compromise all possible paths between Alice and Bob.
+
 \begin{figure}
-\centerline{\includegraphics{fig-trust-source-destination}}
-\caption{TODO}
+\centerline{\includegraphics[height =.4 \textwidth]{fig-bob_Alice_trusted_neigh2}}
+\caption{
+Partial trust model with sender (Alice) and receiver (Bob).
+If disjoint paths exist connecting each node in Alice's trust boundary to
+a node in Bob's,
+than an adversary needs to compromise at least $|T_h(v)|$ nodes to
+compromise all possible paths between Alice and Bob.
+}
 \label{fig:trust-source-destionation}
@@
 \section{Multipath Routing on the Butterfly Topology}
+
+\subsection{Butterfly Network Topology}
 
@@
 \item[Logarithmic diameter]
-For any two nodes, the length of the shortest path between them is logorithmic
+For any two nodes, the length of the shortest path between them is logarithmic
 in the size of the network.
@@
 grows in size.
+\item [Redundancy] 
+As we show in the next section. The number of independent paths between two nodes increases exponentially with the size of the trusted neighborhood. 
 \end{description}
@@
 \end{figure}
+
+\subsection{Routing Algorithm: Overview}
+
+This section gives an informal overview of the multipath routing algorithm.
+In a wraparound butterfly network with trusted radius $h$,
+the algorithm constructs $2^h$ disjoint paths between the trusted neighborhoods
+of any two nodes $v$ and $w$.
+Because of vertex transitivity, we can make the simplifying assumption that $v = (0,0)$ without loss of generality.
+Each path is parameterized by an $h$-bit string $s \in (0,2^h]$ (see Table \ref{tab:routing}).
+Each path cycles around the levels of the butterfly network at most twice.
+
+During the first cycle, the place-within-level $z$ of all nodes is set such that
+no two paths overlap outside the trusted neighborhoods.
+During the second cycle, the place-within-level is set to its destination value
+and the algorithm terminates when the destination level is reached.
+
+More specifically, consider any node, $(l,z)$, in the path corresponding to $s$. The algorithm can be understood by dividing the $m$ bits of
+$z$ into four segments: $A, B, C, D$ (see figure \ref{fig:route-overview}).
+$A$ is the first $h$ bits, followed by $B$ until reaching $C$, which is the last $h$ bits before
+the destination level $l_w$, and then finally $D$. Hence, $A, B, C, $ and $D$ occupy $h$, $l_w - 2h$, $h$, and $m - l_w$ bits of $z$, respectively. 
+Note that due to the construction of the network, as we walk through any path $\{p_1, p_2, \dots\}$, the level $l_{t+1}$ of node $p_{t+1}$ is always one higher (mod $m$) than level $l_t$ the previous node $p_t$.
+Also, $z_{t+1}$ is always the same as the $z_t$, except for possibly the $(l+1)^{th}$ bit of $z_{t+1}$, which can be either 0 or 1.
+Hence, as the routing algorithm progresses through the levels $l$, the segments of $z$ get updated in order.
+
+There are 7 stages of the algorithm to construct a path from $v$ to $w$.
+Stages (1) through (4) consist of the first cycle thought the levels of the network and stages (5) though (7) consists of the second (potentially partial) cycle.
+We start at the source node $v$ and begin constructing the path.
+During stage (1), $A$ is set to match $s$.
+During stage (2) the bits of $B$ are inverted.
+During stage (3) a cyclic permutation of $s$ is placed into $C$.
+After this, in stages (4--7), $D$, $A$, $B$, and $C$ are set to their destination values (see figure \ref{fig:route-overview}).
+
+We now argue that in the above routing algorithm, given $s, s' \in (0, 2^h]$ such that $s \neq s'$, the corresponding paths, $P_s$ and $P_{s'}$, are disjoint when they are both outside of the trusted neighborhoods of $v$ and $w$.
+We informally refer to different segments of nodes in $s$ and $s'$ as $A,B,C,D$ and $A',B',C',D'$, respectively.
+
+First note that any over lap between the two paths must occur either while they are at the same stage of the routing, or while one is at stage (1) and the other one at (5), one is at stage (2) and the other one at (6), or one is at stage (3) and the other one at (7).
+This is because in any other case, the nodes in the paths will be at different levels, $l$, and thus they cannot overlap.
+Furthermore, when a path is in stage (1) or (7), it is in the trusted neighborhoods of $v$ or $w$, hence we are not concerned with overlaps in stages (1) or (7).
+
+When both paths are in stages (2), (3), or (4), they do not overlap because segment $A$ is set to $s$ and segment $A'$ is set to $s'$.
+When both paths are in stages (5) or (6), they do not overlap because segment $C$ is set to a cyclic permutation of $s$ and segment $C'$ is set to a cyclic permutation of $s'$.
+When one path is in stage (1) and the other one is in stage (5), say $P_s$ is in (1) and $P_{s'}$ is in (5), they do not overlap because all bits in $B$ are set to the original bits in the source node and all bits in $B'$ is set to the inverted bits of the source node.
+When one path is in stage (2) and the other one is in stage (6), they do not overlap because at least one pair of bits in $B$ and $B'$ is still be mismatched while stage (6) is being processed.
+
+\begin{figure}
+\begin{center}
+\begin{tikzpicture}[
+node distance=0pt,
+ start chain = A going right,
+    X/.style = {rectangle, draw,% styles of nodes in string (chain)
+                minimum width=10ex, minimum height=3ex,
+                outer sep=0pt, on chain},
+    B/.style = {decorate,
+                decoration={brace, amplitude=5pt,
+                pre=moveto,pre length=1pt,post=moveto,post length=1pt,
+                raise=1mm,
+                            #1}, % for mirroring of brace, if necessary
+                thick},
+    B/.default=mirror, % by default braces are mirrored
+                        ]
+\foreach \i in {0\ldots,{\ldots}0\ldots,{\ldots}0\ldots,{\ldots}0}% <-- content of nodes
+    \node[X] {\i};
+\draw[B] ( A-1.north west) -- node[above=1.5mm] {$h$} ( A-1.north east);
+\draw[B] ( A-2.north west) -- node[above=1.5mm] {$l_w - 2h$} ( A-2.north east);
+\draw[B] ( A-3.north west) -- node[above=1.5mm] {$h$} ( A-3.north east);
+\draw[B] ( A-4.north west) -- node[above=1.5mm] {$m - l_w$} ( A-4.north east);
+\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {start} ( A-1.west);
+\node (B1) [inner sep=1pt,above=of A-1.north,above=5ex] {\underline{A}};
+\node (B2) [inner sep=1pt,above=of A-2.north,above=5ex] {\underline{B}};
+\node (B3) [inner sep=1pt,above=of A-3.north,above=5ex] {\underline{C}};
+\node (B4) [inner sep=1pt,above=of A-4.north,above=5ex] {\underline{D}};
+\end{tikzpicture}
+\\
+\begin{tikzpicture}[
+node distance=0pt,
+ start chain = A going right,
+    X/.style = {rectangle, draw,% styles of nodes in string (chain)
+                minimum width=10ex, minimum height=3ex,
+                outer sep=0pt, on chain},
+                        ]
+\foreach \i in {$s$,{\ldots}0\ldots,{\ldots}0\ldots,{\ldots}0}% <-- content of nodes
+    \node[X] {\i};
+\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {1.} ( A-1.west);
+\end{tikzpicture}
+\\
+\begin{tikzpicture}[
+node distance=0pt,
+ start chain = A going right,
+    X/.style = {rectangle, draw,% styles of nodes in string (chain)
+                minimum width=10ex, minimum height=3ex,
+                outer sep=0pt, on chain},
+                        ]
+\foreach \i in {$s$,{\ldots}1\ldots,{\ldots}0\ldots,{\ldots}0}% <-- content of nodes
+    \node[X] {\i};
+\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {2.} ( A-1.west);
+\end{tikzpicture}
+\\
+\begin{tikzpicture}[
+node distance=0pt,
+ start chain = A going right,
+    X/.style = {rectangle, draw,% styles of nodes in string (chain)
+                minimum width=10ex, minimum height=3ex,
+                outer sep=0pt, on chain},
+                        ]
+\foreach \i in {$s$,{\ldots}1\ldots,{\ldots}$\tilde{s}$\ldots,{\ldots}0}% <-- content of nodes
+    \node[X] {\i};
+\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {3.} ( A-1.west);
+\end{tikzpicture}
+\\
+\begin{tikzpicture}[
+node distance=0pt,
+ start chain = A going right,
+    X/.style = {anchor=base, rectangle, draw,% styles of nodes in string (chain)
+                minimum width=10ex, minimum height=3ex,
+                outer sep=0pt, on chain},
+                        ]
+\foreach \i in {$s$,{\ldots}1\ldots,{\ldots}$\tilde{s}$\ldots,$z_{w,D}$}% <-- content of nodes
+    \node[X] {\i};
+\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {4.} ( A-1.west);
+\end{tikzpicture}
+\\
+\begin{tikzpicture}[
+node distance=0pt,
+ start chain = A going right,
+    X/.style = {anchor=base, rectangle, draw,% styles of nodes in string (chain)
+                minimum width=10ex, minimum height=3ex,
+                outer sep=0pt, on chain},
+                        ]
+\foreach \i in {$z_{w,A}$,{\ldots}1\ldots,{\ldots}$\tilde{s}$\ldots,$z_{w,D}$}% <-- content of nodes
+    \node[X] {\i};
+\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {5.} ( A-1.west);
+\end{tikzpicture}
+\\
+\begin{tikzpicture}[
+node distance=0pt,
+ start chain = A going right,
+    X/.style = {anchor=base, rectangle, draw,% styles of nodes in string (chain)
+                minimum width=10ex, minimum height=3ex,
+                outer sep=0pt, on chain},
+                        ]
+\foreach \i in {$z_{w,A}$,$z_{w,B}$,{\ldots}$\tilde{s}$\ldots,$z_{w,D}$}% <-- content of nodes
+    \node[X] {\i};
+\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {6.} ( A-1.west);
+\end{tikzpicture}
+\\
+\begin{tikzpicture}[
+node distance=0pt,
+ start chain = A going right,
+    X/.style = {anchor=base, rectangle, draw,% styles of nodes in string (chain)
+                minimum width=10ex, minimum height=3ex,
+                outer sep=0pt, on chain},
+                        ]
+\foreach \i in {$z_{w,A}$,$z_{w,B}$,$z_{w,C}$,$z_{w,D}$}% <-- content of nodes
+    \node[X] {\i};
+\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {7.} ( A-1.west);
+\end{tikzpicture}
+\end{center}
+\caption{
+Progression of place-within-level $z$ as the multipath routing algorithm
+cycles through the levels of the butterfly network.
+}
+\label{fig:route-overview}
+\end{figure}
+
+\begin{table}%
+\tbl{Butterfly Multipath Routing Variables\label{tab:routing}}
+{
+\begin{tabular}{|l|l|}
+\hline
+NAME & VARIABLE \\\hline
+butterfly dimension & $m \in \mathbb{Z}_+$ \\\hline
+node level & $l \in \mathbb{Z} : 0 \leq l < m$ \\\hline
+node place within level & $z \in \mathbb \{0,1\}^m$ \\\hline
+trust radius & $h \in \mathbb{Z} : 1 \leq h \leq \lfloor m/2 \rfloor$ \\\hline
+path index & $s \in \{0,1\}^h$ \\\hline
+\end{tabular}
+}
+\end{table}%
+
+\subsection{Routing Algorithm: Proof}
 
@@
 In Stage 1 ($0 \leq t < h$), down or down-right edges
-are chosen such that the the $t$th bit of $z^{(t+1)}$ is equal to the $t$th bit
+are chosen such that the $t$th bit of $z^{(t+1)}$ is equal to the $t$th bit
 of $s$. Throughout Stage 1, all nodes are within the sender's trusted neighborhood.

### 2016-02-23 ec-2016/source/paper.tex
Update figure: overview.

@@
 \usepackage{tikz}
-\usetikzlibrary{chains,decorations.pathreplacing}
+\usetikzlibrary{arrows,chains,decorations.pathreplacing}
 \newcommand{\beq}{\begin{eqnarray}}
@@
                 outer sep=0pt, on chain},
-    B/.style = {decorate,
-                decoration={brace, amplitude=5pt,
-                pre=moveto,pre length=1pt,post=moveto,post length=1pt,
-                raise=1mm,
-                            #1}, % for mirroring of brace, if necessary
-                thick},
-    B/.default=mirror, % by default braces are mirrored
                         ]
@@
     \node[X] {\i};
-\draw[B] ( A-1.north west) -- node[above=1.5mm] {$h$} ( A-1.north east);
-\draw[B] ( A-2.north west) -- node[above=1.5mm] {$l_w - 2h$} ( A-2.north east);
-\draw[B] ( A-3.north west) -- node[above=1.5mm] {$h$} ( A-3.north east);
-\draw[B] ( A-4.north west) -- node[above=1.5mm] {$m - l_w$} ( A-4.north east);
+\draw[<->] ([yshift=1.5mm] A-1.north east) -- node[above=0.25mm] {$h$} ([yshift=1.5mm] A-1.north west);
+\draw[<->] ([yshift=1.5mm] A-2.north east) -- node[above=0.25mm] {$l_w - 2h$} ([yshift=1.5mm] A-2.north west);
+\draw[<->] ([yshift=1.5mm] A-3.north east) -- node[above=0.25mm] {$h$} ([yshift=1.5mm] A-3.north west);
+\draw[<->] ([yshift=1.5mm] A-4.north east) -- node[above=0.25mm] {$m - l_w$} ([yshift=1.5mm] A-4.north west);
 \draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {start} ( A-1.west);
@@
                 outer sep=0pt, on chain},
+    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
+                minimum width=10ex, minimum height=3ex,
+                outer sep=0pt, on chain, thick},
                         ]
-\foreach \i in {$s$,{\ldots}0\ldots,{\ldots}0\ldots,{\ldots}0}% <-- content of nodes
+\node[Y] {$s$};
+\foreach \i in {{\ldots}0\ldots,{\ldots}0\ldots,{\ldots}0}% <-- content of nodes
     \node[X] {\i};
@@
                 outer sep=0pt, on chain},
+    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
+                minimum width=10ex, minimum height=3ex,
+                outer sep=0pt, on chain, thick},
                         ]
-\foreach \i in {$s$,{\ldots}1\ldots,{\ldots}0\ldots,{\ldots}0}% <-- content of nodes
+\foreach \i in {$s$}% <-- content of nodes
+    \node[X] {\i};
+\node[Y] {{\ldots}1\ldots};
+\foreach \i in {{\ldots}0\ldots,{\ldots}0}% <-- content of nodes
     \node[X] {\i};
@@
                 outer sep=0pt, on chain},
+    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
+                minimum width=10ex, minimum height=3ex,
+                outer sep=0pt, on chain, thick},
                         ]
-\foreach \i in {$s$,{\ldots}1\ldots,{\ldots}$\tilde{s}$\ldots,{\ldots}0}% <-- content of nodes
+\foreach \i in {$s$,{\ldots}1\ldots}% <-- content of nodes
+    \node[X] {\i};
+\node[Y] {$\tilde{s}$};
+\foreach \i in {{\ldots}0}% <-- content of nodes
     \node[X] {\i};
@@
                 outer sep=0pt, on chain},
+    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
+                minimum width=10ex, minimum height=3ex,
+                outer sep=0pt, on chain, thick},
                         ]
-\foreach \i in {$s$,{\ldots}1\ldots,{\ldots}$\tilde{s}$\ldots,$z_{w,D}$}% <-- content of nodes
+\foreach \i in {$s$,{\ldots}1\ldots,$\tilde{s}$}% <-- content of nodes
     \node[X] {\i};
+\node[Y] {$z_{w,D}$};
 \draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {4.} ( A-1.west);
@@
                 outer sep=0pt, on chain},
+    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
+                minimum width=10ex, minimum height=3ex,
+                outer sep=0pt, on chain, thick},
                         ]
-\foreach \i in {$z_{w,A}$,{\ldots}1\ldots,{\ldots}$\tilde{s}$\ldots,$z_{w,D}$}% <-- content of nodes
+\node[Y] {$z_{w,A}$};
+\foreach \i in {{\ldots}1\ldots,$\tilde{s}$,$z_{w,D}$}% <-- content of nodes
     \node[X] {\i};
@@
                 outer sep=0pt, on chain},
+    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
+                minimum width=10ex, minimum height=3ex,
+                outer sep=0pt, on chain, thick},
                         ]
-\foreach \i in {$z_{w,A}$,$z_{w,B}$,{\ldots}$\tilde{s}$\ldots,$z_{w,D}$}% <-- content of nodes
+\foreach \i in {$z_{w,A}$}% <-- content of nodes
+    \node[X] {\i};
+\node[Y] {$z_{w,B}$};
+\foreach \i in {$\tilde{s}$,$z_{w,D}$}% <-- content of nodes
     \node[X] {\i};
@@
                 outer sep=0pt, on chain},
+    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
+                minimum width=10ex, minimum height=3ex,
+                outer sep=0pt, on chain, thick},
                         ]
-\foreach \i in {$z_{w,A}$,$z_{w,B}$,$z_{w,C}$,$z_{w,D}$}% <-- content of nodes
+\foreach \i in {$z_{w,A}$,$z_{w,B}$}% <-- content of nodes
+    \node[X] {\i};
+\node[Y] {$z_{w,C}$};
+\foreach \i in {$z_{w,D}$}% <-- content of nodes
     \node[X] {\i};

### 2016-02-23 ec-2016/source/paper.tex
Remove old notes.
Update todo list.
Add to section: multipath fault tolerance.

@@
 
-TODO: Fix bibtex encoding.
+TODO:\\
+Fix bibtex encoding\\
+Move secrecy section\\
+Add ref to unipath butterfly routing\\
+Update intro to make significance more clear\\
 
@@
 
-Source-to-network
 \begin{figure}
@@
 
-Fault tolerance with partial trust and redundancy.
+When the assumptions of the partial trust model hold, a network can achieve
+robust fault tolerance by utilizing the redundancy of multiple paths
+between the sender and receiver.
+Furthermore, this fault tolerance can remain robust even when faults are
+adversarially chosen by an attacker who has access to more resources
+than the sender and receiver.
+On a structured network, multipath fault tolerance reveals not just when
+an error has occured, but also which routes are faulty and potentially
+compromised.
 
-\subsection{Fault Tolerance}
+In the terminology of fault tolerance, a {\em fault} occurs when one component
+of a system behaves incorrectly (e.g., a routing node blocking or
+altering a message).
+The result of that fault (e.g., a recipient receiving conflicting messages)
+is an {\em error} state.
+If the error is undetected or incorrectly corrected, the system is
+said to have experienced a {\em failure} (e.g., an altered message is
+accepted as authentic).
+We are concerned in particular with {\em adversarial faults},
+which (as opposed to random faults)
+are chosen strategically (within the assumptions of the trust model)
+to maximize the liklihood of a failure.
+In this section, we focus on the simplest possible question:
+when using multipath fault tolerance,
+what is the probability that adversarial faults will cauase an undetectable
+error, and thus guarantee a failure?
 
-Results
+They key idea behind multipath fault tolerance is that several copies of
+a message are sent, along multiple routes,
+rather than using traditional unicast routing.
+If any of the message copies differ from one another or do not arrive,
+the receiver can conclude that an error has occurred,
+and possibly correct it, depending on the magnitude of the error
+and the desired fault tolerance properties.
+In the simplest such scheme, a sender broadcasts copies of a message
+along all possible routes to the receiver.
+Such an approach takes advantage of all possible redundancy in the network,
+and requires an adversary to compromise at least one node on each path
+in order to create an undetectable error.
+Realistically, the sender may only have the resources to send a limited number of
+copies, or the network may not be able to handle the congestion caused by
+sending all messages along all paths.
+However, we will now show that a network can provide a high level of
+fault tolerance with only a few copies of a message,
+provided the number of possible paths is large.
 

### 2016-02-23 ec-2016/source/paper.tex
Add to section: multipath fault tolerance.

@@
 and the desired fault tolerance properties.
-In the simplest such scheme, a sender broadcasts copies of a message
-along all possible routes to the receiver.
+
+In the simplest such scheme, a sender (Alice) broadcasts copies of a message
+along all possible routes to the receiver (Bob).
 Such an approach takes advantage of all possible redundancy in the network,
-and requires an adversary to compromise at least one node on each path
+and requires an adversary (Eve) to compromise at least one node on each path
 in order to create an undetectable error.
+But how many nodes is that?
+The answer depends on the smallest {\em vertex cut} of the untrusted
+portion of the network---the smallest set of nodes than can be removed
+to separate the network into two components.
+Any path between Alice and Bob must pass through all such sets,
+so to guarantee an undetectable error, Eve must compromise at least as many
+nodes as are in the smallest vertex cut.
+Alice and Bob's trust boundaries are both vertex cuts,
+so the smaller of the two is an upper bound.
+It is not difficult to construct networks in which the trust boundary is exactly
+the smallest vertex cut,
+and we will show this to be true for the networks we consider below.
+We denote the size of the smallest vertex cut---the
+maximum number of disjoint paths between Alice and Bob's trusted
+neighborhoods---as $B$.
+
 Realistically, the sender may only have the resources to send a limited number of
@@
 fault tolerance with only a few copies of a message,
-provided the number of possible paths is large.
+provided the number of disjoint paths $B$ is large.
+
+In a more sophisticated multipath fault tolerance scheme,
+Alice randomly chooses $k$ paths and sends a copy of
+her message on each.
+We further assume that Alice posesses a routing algorithm
+(such as described in the next section)
+that enables her to construct paths that
+do not intersect between trusted neighborhoods.
+So the maximum number of paths Alice could choose is $B$,
+otherwise two paths would have to intersect in the minimum vertex cut.
+Let us assume that Eve is capable of compromising up to $l$ nodes.
+Eve's best adversarial strategy is now to compromise as many nodes in the smallest
+vertex cut as possible,
+succeeding when each of Alice's paths include one of the compromised nodes.
+Since Alice chooses paths randomly,
+all nodes in the vertex cut are equally likely to contain a path,
+so Eve can do no better than also choosing randomly.
+If $k > l$, at least one message will get through uncompromised and all
+errors are detectable.
+Otherwise, the probability of Eve producing an undetectable error is
+the probability that all of Alice's paths contain compromised nodes:
+\beq
+TODO
+\eeq
 

### 2016-02-23 ec-2016/source/paper.tex
Update section: multipath routing on the butterlfy topology.

@@
 \beq
-TODO
+\label{eq:pf}
+p_f &=& \frac{l!(B-k)!}{B!(l-k)!}.
 \eeq
+Letting $k=\alpha B$ and $l=\beta B$, then applying Stirling's
+approximation gives:
+\begin{eqnarray}
+\label{eq:pf_approx}
+p_f &\approx&
+\frac{\sqrt{\beta(1-\alpha)}}{\sqrt{\beta-\alpha}}
+\left[
+    \left( \frac{\beta-\alpha}{1-\alpha} \right)^{\alpha}
+    \left( \frac{\beta}{\beta-\alpha} \right)^{\beta}
+    (1-\alpha)
+\right]^B.
+\end{eqnarray}
+
+Figure \ref{fig:pfail} shows the value of $p_f$
+as a function of $k$ and $l$ for various values of $B$.
+Equation (\ref{eq:pf_approx}) shows that while $p_f$
+depends on the fractions of
+redundant paths actually utilized $\alpha$ and compromised $\beta$,
+it decreases exponentially in the total number of {\em possible}
+redundant paths.
+This result is significant because $B$ depends only on the network structure,
+not on the amount of resources available to the sender or the adversary.
+In other words, senders can tolerate attacks from significantly more powerful
+adversaries, as long as the network structure provides large $B$.
+Formally we have shown that under the partial trust assumption,
+multipath fault tolerance can provide {\em asymptotic security}
+against adversarial faults if $B$ can be made arbitrarily large.
+Similarly, because the cost of an attack is prohibitively large
+compared to the cost of protecting against an attack,
+this architecture creates what a stabilizing asymmetry.
+By putting attackers at a disadvantage, attacks are not only tolerated
+but also disincentivized.
 
@@
 \centerline{\includegraphics{fig-perror}}
-\caption{TODO}
-\label{fig:one}
+\caption{
+The probability of an undetectable error as a function of the number of
+message copies and the number of adversarial faults,
+for various numbers of redundant paths.
+}
+\label{fig:pfail}
 \end{figure}
@@
 
-We now describe a Concurrent Multipath Routing scheme on a specific family
-of networks: the butterfly network \cite{}.
+We now present a decentralized, fault-tolerant, multipath routing algorithm
+on a specific topology known as the butterfly network topology
+\cite{kshemkalyani_distributed_2008}.
 Several variations on the butterfly network exist.
 Specifically, we utilize the wrap-around butterfly.
-We denote the $m$-dimensional directed wrap-around butterfly as $\wbf(m)$:
+We denote the $m$-dimensional, directed wrap-around butterfly as $\wbf(m)$:
 \beq
 \wbf(m) &=& (V, E_\downarrow \cup E_\rightarrow) \\
-V &=& \mathbb{Z}_m \times \mathbb{Z}_{2^m} \\
-E_\downarrow &=& \{((l,z),(l+1 (\text{mod } m),z) \nonumber \\
+V &=& \mathbb{Z}_{m} \times \mathbb{Z}_2^m \\
+E_\downarrow
+&=&
+\{((l,z),(l+1 (\text{mod } m),z) \nonumber \\
 && \; | \, l \in \mathbb{Z}_d, z \in \mathbb{Z}_{2^m}\} \\
-E_\rightarrow &=& \{(l,z),(l+1 (\text{mod } m), z \oplus 2^l) \nonumber \\
+E_\rightarrow
+&=&
+\{(l,z),(l+1 (\text{mod } m),
+(z_0, \ldots, z_{l-1},z_l \oplus 1, z_{l+1}, \ldots, z_{m-1}) \nonumber \\
 && \; | \, l \in \mathbb{Z}_d, z \in \mathbb{Z}_{2^m}\},
 \eeq
-where $\oplus$ represents the bitwise XOR operator.
-Each node is associated with a level $l$ and an $m$-bit integer $z$.
+where $\mathbb{Z}_m$ is the set of integers modulo $m$,
+and $\oplus$ represents the addition modulo 2.
+Each node is associated with a level $l$ and an $m$-bit string $z$
+known as {\em the place-within-level}.
 There are two types of edges, shown in Figure \ref{fig:butterfly}.
@@
 Down-right edges ($E_\rightarrow$) also link to a node of level $l + 1$,
-but one having the bitstring equal to $z$ with the $l$th bit flipped.
+but one having the bitstring equal to $z$ with the $l$th bit inverted.
 
-The wrap-around butterfly network has a number of properties making it useful for multipath routing:
+The wrap-around butterfly network has a number of properties desired for
+a scalable, decentralized communication network.
 \begin{description}
-\item[Vertex transitivity]
-The problem
-of finding a route between arbitrary nodes $v$ and $w$
-can be reduced to finding a route from node $(0,0)$ to some $\tilde{w}$.
-\item[Logarithmic diameter]
-For any two nodes, the length of the shortest path between them is logarithmic
-in the size of the network.
-\item[Constant degree]
-In practical applications, each communication link requires additional resources,
-such as physical infrastructure or entries in a routing table.
-A constant degree limits the number of such resources needed as the network
-grows in size.
-\item [Redundancy] 
-As we show in the next section. The number of independent paths between two nodes increases exponentially with the size of the trusted neighborhood. 
+\item[Vertex-transitive.]
+Vertex transitivity means that the network is maximally decentralized.
+We will also use this property throught the proof, noting that
+the problem of finding a route between arbitrary nodes $\tilde{v}$ and $\tilde{w}$
+can be reduced to finding a route from node $(0,0)$ to some node $w$.
+\item[Small-diameter.]
+For any two nodes, the length of the shortest path between them is
+$O(log N)$, where N is the number of nodes in the network.
+\item[Sparse.]
+With a constant degree of 4, the wraparound butterfly is extremely sparse,
+and can scale indefnitely without node degree becoming a limitation.
+\item [Redundant] 
+The number of independent paths between the $h$-trusted neighborhoods of two
+nodes in the wraparound butterfly increases exponentially in $h$.
+We prove this assertion below by giving an algorithm to explicitly construct
+these paths.
 \end{description}
@@
 The node $(l,z)$ is shown as the bit string $z$ with a square around the $l$th bit.
+"Down" edges increment $l$, leaving $z$ unchanged, while "down-right" edges
+increment $l$ and invert the $l$th bit of $z$.
 \label{fig:butterfly}

### 2016-02-23 ec-2016/source/paper.tex
Minor revisions to subsection: butterfly network.

@@
 &=&
-\{((l,z),(l+1 (\text{mod } m),z) \nonumber \\
-&& \; | \, l \in \mathbb{Z}_d, z \in \mathbb{Z}_{2^m}\} \\
+\{((l,z),(l+1 (\text{mod } m),z) \} \\
 E_\rightarrow
@@
 \{(l,z),(l+1 (\text{mod } m),
-(z_0, \ldots, z_{l-1},z_l \oplus 1, z_{l+1}, \ldots, z_{m-1}) \nonumber \\
-&& \; | \, l \in \mathbb{Z}_d, z \in \mathbb{Z}_{2^m}\},
+(z_0, \ldots, z_{l-1},z_l \oplus 1, z_{l+1}, \ldots, z_{m-1}) \},
 \eeq
@@
 \begin{description}
-\item[Vertex-transitive.]
+\item[Vertex-transitive]
 Vertex transitivity means that the network is maximally decentralized.
@@
 can be reduced to finding a route from node $(0,0)$ to some node $w$.
-\item[Small-diameter.]
+\item[Small-diameter]
 For any two nodes, the length of the shortest path between them is
-$O(log N)$, where N is the number of nodes in the network.
-\item[Sparse.]
+$O(\log N)$, where N is the number of nodes in the network.
+\item[Sparse]
 With a constant degree of 4, the wraparound butterfly is extremely sparse,

### 2016-02-24 ec-2016/source/paper.tex
Update to routing proof.

@@
 Update intro to make significance more clear\\
+Clarify identifying fault\\
 
@@
 Down-right edges ($E_\rightarrow$) also link to a node of level $l + 1$,
-but one having the bitstring equal to $z$ with the $l$th bit inverted.
+but one having the place-within-level equal to $z$ with the $l$th bit inverted.
 
-The wrap-around butterfly network has a number of properties desired for
+The wrap-around butterfly network has a number of properties desirable for
 a scalable, decentralized communication network.
@@
 \item[Vertex-transitive]
-Vertex transitivity means that the network is maximally decentralized.
+Because the wraparound butterfly is vertex transitive,
+it is maximally decentralized.
 We will also use this property throught the proof, noting that
@@
 the algorithm constructs $2^h$ disjoint paths between the trusted neighborhoods
-of any two nodes $v$ and $w$.
-Because of vertex transitivity, we can make the simplifying assumption that $v = (0,0)$ without loss of generality.
-Each path is parameterized by an $h$-bit string $s \in (0,2^h]$ (see Table \ref{tab:routing}).
+of any two nodes $v$ and $w = (l_w, z_w)$.
+Because of vertex transitivity,
+we can make the simplifying assumption that $v = (0,0)$ without loss of generality.
+Each path is parameterized by an $h$-bit string $s \in \mathbb{Z}_2^h$ (see Table \ref{tab:routing}).
 Each path cycles around the levels of the butterfly network at most twice.
@@
 $z$ into four segments: $A, B, C, D$ (see figure \ref{fig:route-overview}).
-$A$ is the first $h$ bits, followed by $B$ until reaching $C$, which is the last $h$ bits before
-the destination level $l_w$, and then finally $D$. Hence, $A, B, C, $ and $D$ occupy $h$, $l_w - 2h$, $h$, and $m - l_w$ bits of $z$, respectively. 
-Note that due to the construction of the network, as we walk through any path $\{p_1, p_2, \dots\}$, the level $l_{t+1}$ of node $p_{t+1}$ is always one higher (mod $m$) than level $l_t$ the previous node $p_t$.
-Also, $z_{t+1}$ is always the same as the $z_t$, except for possibly the $(l+1)^{th}$ bit of $z_{t+1}$, which can be either 0 or 1.
-Hence, as the routing algorithm progresses through the levels $l$, the segments of $z$ get updated in order.
+$A$ is the first $h$ bits,
+and $C$ is the $h$ bits preceeding index $l_w$.
+$B$ is formed by the remaining bits between $A$ and $C$,
+while $D$ refers to the remaining bits after $C$.
+Hence, $A, B, C, $ and $D$ occupy $h$, $l_w - 2h$, $h$, and $m - l_w$ bits of $z$, respectively
+(we leave the caase of overlapping $A$ and $C$ for the formal proof).
+Note that due to the construction of the network,
+as we walk through any path $\{p_1, p_2, \dots\}$,
+the level $l_{t+1}$ of node $p_{t+1}$ is always one higher (mod $m$)
+than level $l_t$ the previous node $p_t$.
+Also, $z_{t+1}$ is always the same as the $z_t$,
+except for possibly the $(l+1)^{th}$ bit of $z_{t+1}$,
+which can be either 0 or 1.
+Hence, as the routing algorithm progresses through the levels $l$,
+the segments of $z$ get updated in order.
 
 There are 7 stages of the algorithm to construct a path from $v$ to $w$.
-Stages (1) through (4) consist of the first cycle thought the levels of the network and stages (5) though (7) consists of the second (potentially partial) cycle.
-We start at the source node $v$ and begin constructing the path.
+Stages (1) through (4) consist of the first cycle thought the levels of
+the network and stages (5) though (7) consists of the second
+(potentially partial) cycle.
+We start at the source node $v = (0,0)$ and begin constructing the path.
 During stage (1), $A$ is set to match $s$.
 During stage (2) the bits of $B$ are inverted.
-During stage (3) a cyclic permutation of $s$ is placed into $C$.
-After this, in stages (4--7), $D$, $A$, $B$, and $C$ are set to their destination values (see figure \ref{fig:route-overview}).
+During stage (3) a cyclic permutation of $s$ is placed into $C$
+(the details of the permutation are necessary for the overlapping case,
+but not relevant for this overview).
+After this, in stages (4--7), $D$, $A$, $B$, and $C$ are set to their
+destination values (see figure \ref{fig:route-overview}).
 
-We now argue that in the above routing algorithm, given $s, s' \in (0, 2^h]$ such that $s \neq s'$, the corresponding paths, $P_s$ and $P_{s'}$, are disjoint when they are both outside of the trusted neighborhoods of $v$ and $w$.
-We informally refer to different segments of nodes in $s$ and $s'$ as $A,B,C,D$ and $A',B',C',D'$, respectively.
+We now argue that in the above routing algorithm,
+given $s, s' \in \mathbb{Z}_2^h$ such that $s \neq s'$,
+the corresponding paths, $P_s$ and $P_{s'}$,
+are disjoint when they are both outside of the trusted neighborhoods of
+$v$ and $w$.
+We informally refer to different segments of nodes in $s$ and $s'$ as
+$A,B,C,D$ and $A',B',C',D'$, respectively.
 
-First note that any over lap between the two paths must occur either while they are at the same stage of the routing, or while one is at stage (1) and the other one at (5), one is at stage (2) and the other one at (6), or one is at stage (3) and the other one at (7).
-This is because in any other case, the nodes in the paths will be at different levels, $l$, and thus they cannot overlap.
-Furthermore, when a path is in stage (1) or (7), it is in the trusted neighborhoods of $v$ or $w$, hence we are not concerned with overlaps in stages (1) or (7).
+First note that any overlap between the two paths must occur either while they
+are at the same stage of the routing, or while one is at stage (1) and the other one at (5), one is at stage (2) and the other one at (6), or one is at stage (3) and the other one at (7).
+This is because in any other case,
+the nodes in the paths will be at different levels, $l$,
+and thus they cannot overlap.
+Furthermore, when a path is in stage (1) or (7),
+it is in the trusted neighborhoods of $v$ or $w$,
+hence we are not concerned with overlaps within stages (1) or (7).
 
-When both paths are in stages (2), (3), or (4), they do not overlap because segment $A$ is set to $s$ and segment $A'$ is set to $s'$.
-When both paths are in stages (5) or (6), they do not overlap because segment $C$ is set to a cyclic permutation of $s$ and segment $C'$ is set to a cyclic permutation of $s'$.
-When one path is in stage (1) and the other one is in stage (5), say $P_s$ is in (1) and $P_{s'}$ is in (5), they do not overlap because all bits in $B$ are set to the original bits in the source node and all bits in $B'$ is set to the inverted bits of the source node.
-When one path is in stage (2) and the other one is in stage (6), they do not overlap because at least one pair of bits in $B$ and $B'$ is still be mismatched while stage (6) is being processed.
+When both paths are in stages (2), (3), or (4), they do not overlap because
+segment $A$ is set to $s$ and segment $A'$ is set to $s'$.
+When both paths are in stages (5) or (6),
+they do not overlap because segment $C$ is set to a cyclic permutation of $s$
+and segment $C'$ is the same cyclic permutation of $s'$.
+When one path is in stage (1) and the other one is in stage (5),
+say $P_s$ is in (1) and $P_{s'}$ is in (5),
+they do not overlap because all bits in $B$ are set to the original bits in the
+source node and all bits in $B'$ are set to the inverted bits of the source node.
+When one path is in stage (2) and the other one is in stage (6), they do not
+overlap because at least one pair of bits in $B$ and $B'$ are still inverted
+while stage (6) is being processed.
 
@@
 node level & $l \in \mathbb{Z} : 0 \leq l < m$ \\\hline
-node place within level & $z \in \mathbb \{0,1\}^m$ \\\hline
+node place within level & $z \in \mathbb{Z}_2^m$ \\\hline
 trust radius & $h \in \mathbb{Z} : 1 \leq h \leq \lfloor m/2 \rfloor$ \\\hline
-path index & $s \in \{0,1\}^h$ \\\hline
+path index & $s \in \mathbb{Z}_2^h$ \\\hline
 \end{tabular}
@@
 
-\subsection{Routing Algorithm: Proof}
+\subsection{Routing Algorithm: Specification and Proof}
 
-We now describe a routing scheme that provides $2^h$ redundant paths between
+We now specify the multipath routing scheme and prove
+that provides $2^h$ disjoint paths between
 the $h$-hop trusted neighborhoods of any two nodes in an $m$-bit
 wrap-around butterfly network.
-Utilizing vertex transitivity, we label the source node as $(0, 0)$ and
-denote the destination node as $w = (l_w, z_w)$.
+Utilizing vertex transitivity, we label the source node as
+$(l^{(0)}, z^{(0)}) = (0, 0)$ and denote the destination node as $w = (l_w, z_w)$,
+without loss of generality.
 
-Let $s$ be an integer such that $0 \leq s < 2^h$.
-Let $v_s^{(t)} = (l^{(t)}, z^{(t)})$ be the $t$th node in the path labeled by $s$.
-For convenience, we will omit the subscript $s$.
-We define two partitionings of the integers in $\mathbb{Z}_{2^m}$:
-one having the lowest $h$ bits matching the
-bits of $s$, and one having the $h$ bits preceeding the destination level $l_w$
-matching $s$:
+Let $s$ be an $h$-bit binary string with $s_i$ denoting the bit at index $i$.
+There are $2^h$ such strings.
+Let $v_s^{(t)} = (l^{(t)}, z^{(t)})$ be the node at position $t$
+in the path parameterized by $s$.
+For convenience,
+we will omit the subscript $s$ when it is obvious from context.
+We define three distinct partitions of $m$-bit binary strings.
+Let $Q_{v^{(0)}}$ ($\overline{Q_{v^{(0)}}}$) be the set of $m$-bit
+strings in which the bits at
+all indeces $h \leq i < l_w - h$ match (do not all match) those of $z^{(0)}$.
+Note that $Q_{v^{(0)}}$ is trivially all $m$-bit strings if $l_w < 2h$.
+Let $R_s$ ($\overline{R_s}$) be the set of $m$-bit strings with the lowest $h$
+bits all matching (not all matching) the bits of $s$.
+Let $S_s$ ($\overline{S_s}$) be the set of $m$-bit strings with the $h$ bits
+preceeding index $l_w$ all matching (not all matching) the bits of $\tilde{s}$,
+where $\tilde{s}$ is a cyclic permutation of $s$ defined below.
 \beq
-S_s &=& \{z \in \mathbb{Z}_{2^m} | \forall i \in \mathbb{Z}_h z_i = s_i \} \\
-R_s &=& \{z \in \mathbb{Z}_{2^m} | \forall i \in \mathbb{Z}_h z_{(l_w - h + i)} = s_i \}.
+Q_{v^{(0)}} &=&
+\{z \in \mathbb{Z}_2^m \mid \;
+\neg \exists i \in \mathbb{Z} : h \leq (i < l_w - h) \land z_i \neq z_i^{(0)} \}
+\\
+\overline{Q_{v^{(0)}}} &=& \mathbb{Z}_2^m \setminus Q_{v^{(0)}}
+\\
+R_s &=&
+\{z \in \mathbb{Z}_{2}^m \mid \;
+\forall i \in \mathbb{Z}_2^h z_i = s_i \}
+\\
+\overline{R_s} &=& \mathbb{Z}_2^m \setminus R_s
+\\
+S_s &=&
+\{z \in \mathbb{Z}_{2}^m \mid \;
+\forall i \in \mathbb{Z}_2^h z_{(l_w - h + i)} = \tilde{s}_i \}
+\\
+\overline{S_s} &=& \mathbb{Z}_2^m \setminus S_s
+\\
+\tilde{s}_i &=& s_{(i + l_w) \text{ mod } h}.
 \eeq
-Note that if $r \neq s$, then $S_s \cap S_r = R_s \cap R_r = \emptyset$.
+We will make use of the fact that:
+\beq
+s \neq s^\prime &\implies&
+S_s \cap S_{s^\prime} = R_s \cap R_{s^\prime} = \emptyset.
+\eeq
 
-We now construct a path such that between trusted neighborhoods $z^{(t)}$ is always
-in $S_s$, $R_s$, or both, guaranteeing that the path does not overlap with the
-other paths $v_r^{(t)}$.
-Routing proceeds in stages, with the level $l$ increasing by 1 at each hop.
-In Stage 1 ($0 \leq t < h$), down or down-right edges
+Routes are contructed in 7 stages.
+The network topology dictates that $l^{(t+1)} = l^{(t)} + 1$ (mod $m$),
+so we let $l = t$ (mod $m$).
+and that $z^{(t+1)}$ is equal to $z^{(t)}$ with or without the bit in index
+$l^{(t)}$ inverted, depending on whether the down or down-right edge was
+taken at step $t$.
+\begin{description}
+\item[Stage 1: ($0 \leq t < h$)]{
+Down or down-right edges
 are chosen such that the $t$th bit of $z^{(t+1)}$ is equal to the $t$th bit
-of $s$. Throughout Stage 1, all nodes are within the sender's trusted neighborhood.
-At the end of Stage 1, $z^{(h)} \in S_s$, and $z^{(t)}$ will remain so until the level loops
-back to $0$ at $t = m$.
-
-In Stage 2 ($h \leq t < l_w - h$), edges are chosen to make the $t$th bit of
-$z^{(t+1)}$ match the $t$th bit of $z_w$.
-
-In Stage 3 ($l_w - h \leq t < l_w$), the bits of $z^{(t)}$ are chosen to match $s$,
-such that after the stage is complete, $z^{(l_w)} \in R_s$.
-
-In Stage 4 ($l_w \leq t < m$), as in stage 2,
-paths are chosen such that the $t$th bit of $z^{(t+1)}$ matches $z_w$.
-After Stage 4, all bits of $z^{(m)}$ are equal to those of $z_w$
-except for the first $h$ and the $h$ preceeding index $l_w$.
-$z^{(m)}$ is also in both $S_s$ and $R_s$.
-
-At this point, we define $\tau = t - m$.
-In Stage 5 ($0 \leq \tau < h$), the first $h$ bits of $z^{(t)}$ are set to
-match $z_w$, potentially removing $z^{(t)}$ from $S_s$.
-
-In Stage 6 ($h \leq \tau < l_w - h$), all down edges are chosen, incrementing
-the level without any effect on $z^{(t)}$.
-At the end of Stage 6, $z^{(m + l_w - h)}$ is still in $R_s$ and
-$v^{(m + l_w - h)}$ is now within the trusted neighborhood of $w$.
-
-In the seventh, and final stage ($l_w - h \leq \tau < l_w$), the $h$ bits of $z^{(t)}$
+of $s$.
+Throughout Stage 1, all nodes are within the sender's trusted neighborhood.
+Throughout Stage 1, $z^{(t)} \in Q_{v^{(0)}}$.
+At the end of Stage 1, $z^{(h)} \in S_s$, and $z^{(t)}$ will remain so until the level cycles to $0$ at $t = m$.
+}
+\item[Stage 2: ($h \leq t < l_w - h$)]{
+Edges are chosen to make the $t$th bit of
+$z^{(t+1)}$ the inverse of the $t$th bit of $z^{(0)}$.
+Note that this stage does not occur when $l_w < 2h$.
+If this stage occurs, then $z^{(t)} \in \overline{Q_{v^{(0)}}}$ until these
+levels are reached again in stage 6.
+}
+\item[Stage 3: ($l_w - h \leq t < l_w$)]{
+The bits of $z^{(t)}$ are chosen to match $\tilde{s}$,
+such that after the stage is complete, $z^{(t)} \in R_s$.
+}
+\item[Stage 4: ($l_w \leq t < m$)]{
+Paths are chosen such that the $t$th bit of $z^{(t+1)}$ matches that of the
+destination node $z_w$.
+This stage will not occur if $l_w > m - h$.
+}
+\item[Stage 5: ($m \leq t < m + h$)]{
+There are two cases.
+If $2h < l_w < m - h$,
+then there is no overlap between the indeces defining $R_s$ and $S_s$.
+In this case, the first $h$ bits of $z^{(t)}$ are set to
+match $z_w$.
+Otherwise there is some overlap between the indeces defining $R_s$ and
+$S_s$.
+In this case, the each of the first $h$ bits of $z^{(t)}$ is either kept the
+same if $l_w - h \leq l < l_w$, or set to the corresponding bit of $z_w$
+otherwise.
+In this stage and after, $z^{(t)}$ is no longer guaranteed to be in $R_s$.
+However, $z^{(t)}$ remains in $S_s$ during and after this stage.
+}
+\item[Stage 6: ($m + h \leq t < m + l_w - h$)]{
+In this stage, edges are chosen to set the bits of $z^{(t)}$ to their
+corresponding value in $z_w$.
+$z^{(t)} \in \overline{Q_{v^{(0)}}}$ throughout this stage,
+but not afterwards.
+}
+\item[Stage 7: ($m + l_w - h \leq t < m + l_w$)] {
+The $h$ bits of $z^{(t)}$
 preceeding index $l_w$ are set to match $z_w$.
+All nodes in this stage are within $h$ hops of $w$ and thus in its trusted
+neighborhood.
 After this stage, $v^{(m + l_w)} = w$ and routing is complete.
+}
+\end{description}
 

### 2016-02-24 ec-2016/source/paper.tex
Update proof.

@@
 destination values (see figure \ref{fig:route-overview}).
+Figure \ref{fig:routing} shows an explicit example of the routing algorithm
+for one value of $s$.
 
@@
 \end{table}%
+
+\begin{figure}
+\begin{center}
+\includegraphics{fig-routing.pdf}
+\end{center}
+\caption{
+An example of one path as constructed by the proposed multipath
+routing algorithm.
+The path is shown for $s = 10_2$
+and $w = (6, 0110111_2)$.
+\label{fig:routing}
+}
+\end{figure}
 
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics{fig-routing.pdf}
-\end{center}
-\caption{
-\label{fig:routing}
-}
-\end{figure}
+For all $2^h$ of these paths, the components excluding nodes within $h$ hops
+of $v$ or $w$ (in their trusted neighborhoods) are pairwise disjoint,
+which we now prove.
+Nodes from two paths can only coincide if their levels are the same.
+Nodes which share a level must either be in the same stage, or 4 stages
+apart.
+Let ($a$,$a^\prime$) denote a pair of sub-paths corresponding to stage $a$ of
+one path and stage $a^\prime$ of another.
+Excluding paths that intersect in their trusted neighborhoods, (1,1) and (7,7),
+we have reduced the list of possible intersections to the following cases:
+(2,2), (3,3), (4,4), (5,5), (6,6), (1,5), (2,6), and (3,7).
+Nodes in stages 2--4 belong to $R_s$ so cannot overlap with any stage 2--4
+nodes from another path, eliminating (2,2), (3,3), and (4,4).
+Similarly, nodes in stages 4--6 belong to a unique $S_s$,
+eliminating (5,5) and (6,6).
+Nodes in stage 1 belong to $Q_{v^{(0)}}$ while those in stage 5 belong in
+its complement, eliminating (1,5).
+Similarly, for all $l$ in stage 2, $z^{(l)}$ is equal to $z^{(0)}$,
+while in stage 6, $z^{(l)}$ is the inverse, eliminating (2,6)
+This leaves only (3,7), a collision which can occur only for only one path
+(with $s$ matching the first $h$ bits of $z_w$), and which enters the trusted
+neighborhood in stage 3.
+For this single path, we can proceed directly from stage 2 to stage 7,
+eliminating the last possible collision.
+
+Thus, we have shown that assuming the partial trust model, with trust transitive
+for $h$ hops, we can construct $2^h$ paths on a wraparound butterfly topology
+which do not intersect outside the trusted neighborhoods of the source and
+destination.
+This is a lower bound on the value $B$ in Equation (\ref{eq:pf),
+showing that the decentralized, redundant, structured networks such as the
+butterfly can have a very low probability of failure when faced with
+adverarial faults, even from a very powerful attacker.
 

### 2016-02-24 ec-2016/source/paper.tex
Add abstract and CCS codes.

@@
 Attack-Tolerance in Structured Networks via Multipath Routing
-\\ Draft: 2016-02-21
 }
@@
 \begin{abstract}
-TODO
+Networks with single points of failure are particularly susceptible
+to adversarial faults,
+in which an attacker creates faults strategically at central points.
+Centralized network architectures necessarily exhibit this vulnerability,
+leading to a growing popularity in decentralized architectures.
+However, decentralized architectures do not guarantee the absence of single
+points of failure, and networks such as the Internet and World-Wide Web have
+been shown theoretically and historically to be highly susceptible to
+adversarial faults.
+We develop a partial trust model which makes weaker transitivity
+assumptions than common web of trust models.
+Using this model, we propose a general multipath fault tolerance technique for
+structured networks which
+provides asymptotic security against adversarial faults.
+We also provide a specific algorithm for implementing this technique on the
+butterfly network topology.
+When network topology can be dictated, these results can be used to create scalable, attack-tolerant infrastructures.
+More generally, our results provide a formalism for evaluating the effects
+of network structure on adversarial fault tolerance.
 \end{abstract}
@@
 <ccs2012>
+<concept>
+<concept_id>10002978.10002986</concept_id>
+<concept_desc>Security and privacy~Formal methods and theory of security</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10003033.10003034</concept_id>
+<concept_desc>Networks~Network architectures</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10011007.10010940.10010971.10010972.10010540</concept_id>
+<concept_desc>Software and its engineering~Peer-to-peer architectures</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
 </ccs2012>
@@
 
-\ccsdesc[500]{TODO~TODO}
+\ccsdesc[500]{Security and privacy~Formal methods and theory of security}
+\ccsdesc[500]{Networks~Network architectures}
+\ccsdesc[500]{Software and its engineering~Peer-to-peer architectures}
 
@@
 destination.
-This is a lower bound on the value $B$ in Equation (\ref{eq:pf),
+This is a lower bound on the value $B$ in Equation (\ref{eq:pf}),
 showing that the decentralized, redundant, structured networks such as the

### 2016-02-24 ec-2016/source/paper.tex
Add paper overview.

@@
 \begin{bottomstuff}
-This work is supported by TODO.
-
+%This work is supported by TODO.
 Author's addresses: E.L. Platt and D.M. Romero, School of Information; email: \{elplatt,drom\}@umich.edu.
@@
 Move secrecy section\\
-Add ref to unipath butterfly routing\\
 Update intro to make significance more clear\\
@@
 
-Paper organization. TODO
+This paper is organized as follows.
+Section 2. reviews background and related work.
+Section 3. discusses the desired properties of attack-tolerant network
+infrastructure.
+Section 4. proposes the partial trust model.
+Section 5. proposes a general technique for multipath fault tolerance on
+structured networks.
+Section 6. proposes a specific application of this technique to the butterfly
+network topology.
+Section 7. discusses these results.
+And Section 8. concludes.
 

### 2016-02-24 ec-2016/source/paper.tex
Frantic last minute changes.

@@
 
-TODO:\\
-Fix bibtex encoding\\
-Move secrecy section\\
-Update intro to make significance more clear\\
-Clarify identifying fault\\
+%TODO:\\
+%Fix bibtex encoding\\
+%Move secrecy section\\
+%Update intro to make significance more clear\\
+%Clarify identifying fault\\
 
@@
 but making weaker assumptions of trust transitivity.
-
-Given this framework, we develop and evaluate a \emph{multipath routing} technique to reduce 
+While privacy and secrecy are often concerns in secure communication,
+it is not the focus of this paper.
+Existing techniques for privacy and secrecy are relatively mature compared
+to those for tolerating blocked or corrupted messages.
+In fact, an attacker might use message disruption as a fallback when
+surveillance isn't feasible.
+However, we would like to stress that the technqiues presented in this paper
+are entirely compatible with, and in some cases could enhance, existing
+privacy and secrecy techniques.
+
+Given the above framework, we develop and evaluate a \emph{multipath routing} technique to reduce 
 the likelihood that compromised nodes are undetected during the information 
@@
 
-\item[Secrecy]
-In some, but not all cases, it is desirable that the infrastructure provides secrecy and allow a source and a target to exchange private information while still relying on intermediary nodes.
-Our proposed multipath routing technique can be adapted to provide secrecy but we do not focus on this aspect in this paper.
-
 \end{description}
@@
 On a structured network, multipath fault tolerance reveals not just when
-an error has occured, but also which routes are faulty and potentially
-compromised.
+an error has occured, but also provides information about which paths provide
+conflicting information,
+which could be used to help locate compromised nodes.
 
@@
 
-Secrecy
-
-Creation of the network.
-Improves on web of trust.
-Gives framework for determining where to build trust.
-
-Applications
-Distributed apps: storage, email, cryptocurrency, secure multiparty computation.
-Wireless Sensor Networks
+While decentralized architectures have recieved much attention,
+methods for classifying, and evaluating decentralization
+are in their infancy.
+The trust model and multipath fault tolerance technique proposed in this
+paper provide a practical fault tolerance architecture for applications when
+network structure can be imposed,
+as well as a more general method for evaluating the role of network structure
+on fault tolerance.
+
+Fault-tolerant network infrastructures have many direct applications.
+Areas such as cryptocurrency, secure multiparty computation, and wireless
+sensor networks have immediate need for scalable, fault-tolerant
+infrastructures.
+Many Internet services---email, social networks, cloud storage---are still
+highly centralized and vulnearble to technical and non-technical attacks.
+Decentralized fault tolerance is one approach to securing these important
+services and making networked communication safer.
+
+While the primary purpose of our proposed fault tolerance technique is the
+threat of message blocking or corruption, it can provide some improvements
+to secrecy and privacy.
+For example, if a message is broken into parts and each part is sent along
+a different random subset of routes, an attacker who succesfully compromises
+one part of the message will still be unable to compromise the rest.
+
+Our present contributions are limited in several ways.
+The partial trust model, while making fewer assumptions than web of trust
+approaches, still depends on some transitivity of trust,
+which may be a questionable assumption in many cases.
+Better trust models are still needed.
+
+The primary limitation of our structured network approach is that
+link structure is imposed rather than self-organized.
+This requirement may be appropriate for appliations such as
+wireless sensor networks,
+or slow-changing institutional infrastructure,
+in which the architect has considerable control over link structure.
+But in general, it remains an open question whether self-organized networks
+can be restructured, or buit from the ground up in such a way that
+structured multipath fault tolerance technqiues can be applied.
+Similarly, structured network topologies such as the butterfly are defined
+only for specific values.
+Although these values can be as large as desired,
+how to apply methods such as ours to networks with an intermediate number
+of nodes remains an open question.
 
 \section{Conclusion}
+
+We have developed a partial trust model which enables a formal analysis
+of the effect of decentralized network structure on fault tolerance properties.
+We have also proposed a general technique for achieving and evaluating adversarial
+fault tolerance in a structured network with partial trust.
+We found that the probability of an adversary causing an undetectable error
+decreases exponentially as the number of redundant paths between the
+neighborhoods of the sender and receiver grows.
+This property puts attackers at a disadvantage, creating a stabilizing
+asymmetry which disincentivizes attack.
+We also presented a specific implementation of the multipath fault tolerance
+technique on a wraparound butterfly network topology and proved that the
+number of redundant paths grows exponentially with number of network hops
+that can be trusted.
+Our results are directly applicable to architectures in which the link structure
+can be imposed by the architect,
+and provide a formalism that can be used more generally to evaluate the role
+of network structure and redundancy on the fault tolerance of decentralized
+networks.
 
@@
 \begin{acks}
-The authors would like to thank TODO
 \end{acks}

### 2016-08-02 wsdm-2017/source/paper.tex
Add WSDM 2017 manuscript.

--- 2016-02-24 ec-2016/source/paper.tex
@@
-%\documentclass[prodmode]{acmsmall-ec16}
-%\documentclass[prodmode,license]{acmsmall-ec16}
-\documentclass[prodmode,permissions]{acmsmall-ec16}
-
-% Package to generate and customize Algorithm as per ACM style
-\usepackage[ruled]{algorithm2e}
-\renewcommand{\algorithmcfname}{ALGORITHM}
-\usepackage{amsmath,amsfonts,amssymb,bbm} 
-\usepackage[numbers,sort&compress]{natbib} % for citet
-\SetArgSty{textrm}  % for algorithm2e
-\SetAlFnt{\small}
-\SetAlCapFnt{\small}
-\SetAlCapNameFnt{\small}
-\SetAlCapHSkip{0pt}
-\IncMargin{-\parindent}
-
-%\conferenceinfo{EC'16,}{July 24--28, 2016, Maastricht, The Netherlands.}
-%\CopyrightYear{2016}
-%\crdata{978-1-4503-3936-0/16/07}
-%\copyrighttext{Copyright is held by the owner/author(s). Publication rights licensed to ACM.}
-
-\doi{XXXXXXX.XXXXXXX}% TeXSupport
+\documentclass{sig-alternate-05-2015}
+
+%\usepackage[ruled]{algorithm2e}
+%\renewcommand{\algorithmcfname}{ALGORITHM}
+%\usepackage{amsmath,amsfonts,amssymb,bbm} 
+%\usepackage[numbers,sort&compress]{natbib} % for citet
+%\SetArgSty{textrm}  % for algorithm2e
+%\SetAlFnt{\small}
+%\SetAlCapFnt{\small}
+%\SetAlCapNameFnt{\small}
+%\SetAlCapHSkip{0pt}
 
@@
 \usetikzlibrary{arrows,chains,decorations.pathreplacing}
+%\usepackage{draftwatermark}
+%\SetWatermarkText{DRAFT}
 \newcommand{\beq}{\begin{eqnarray}}
@@
 
-% Document starts
 \begin{document}
 
-% Page heads
-%\markboth{G. Zhou et al.}{A Multifrequency MAC Specially Designed for WSN Applications}
-
-% Title portion
-\title{
-Attack-Tolerance in Structured Networks via Multipath Routing
-}
-\author{EDWARD L. PLATT
-\affil{University of Michigan}
-DANIEL M. ROMERO
-\affil{University of Michigan}}
-
+% Copyright
+\setcopyright{acmcopyright}
+%\setcopyright{acmlicensed}
+%\setcopyright{rightsretained}
+%\setcopyright{usgov}
+%\setcopyright{usgovmixed}
+%\setcopyright{cagov}
+%\setcopyright{cagovmixed}
+
+% DOI
+%\doi{10.475/123_4}
+
+% ISBN
+%\isbn{123-4567-24-567/08/06}
+
+%Conference
+\conferenceinfo{WSDM '17}{February 6--10, 2017, Cambridge, UK, USA}
+
+\acmPrice{\$15.00}
+
+%
+% --- Author Metadata here ---
+%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
+%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
+%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
+% --- End of Author Metadata ---
+
+\title{Attack-Tolerant Networks: A Structural, Multipath Approach}
+%
+% You need the command \numberofauthors to handle the 'placement
+% and alignment' of the authors beneath the title.
+%
+% For aesthetic reasons, we recommend 'three authors at a time'
+% i.e. three 'name/affiliation blocks' be placed beneath the title.
+%
+% NOTE: You are NOT restricted in how many 'rows' of
+% "name/affiliations" may appear. We just ask that you restrict
+% the number of 'columns' to three.
+%
+% Because of the available 'opening page real-estate'
+% we ask you to refrain from putting more than six authors
+% (two rows with three columns) beneath the article title.
+% More than six makes the first-page appear very cluttered indeed.
+%
+% Use the \alignauthor commands to handle the names
+% and affiliations for an 'aesthetic maximum' of six authors.
+% Add names, affiliations, addresses for
+% the seventh etc. author(s) as the argument for the
+% \additionalauthors command.
+% These 'additional authors' will be output/set for you
+% without further effort on your part as the last section in
+% the body of your article BEFORE References or any Appendices.
+
+\numberofauthors{1} %  in this sample file, there are a *total*
+% of EIGHT authors. SIX appear on the 'first-page' (for formatting
+% reasons) and the remaining two appear in the \additionalauthors section.
+%
+\author{
+% You can go ahead and credit any number of authors here,
+% e.g. one 'row of three' or two rows (consisting of one row of three
+% and a second row of one, two or three).
+%
+% The command \alignauthor (no curly braces needed) should
+% precede each author name, affiliation/snail-mail address and
+% e-mail address. Additionally, tag each line of
+% affiliation/address with \affaddr, and tag the
+% e-mail address with \email.
+%
+% 1st. author
+\alignauthor
+Anonymized Authors
+}
+%       \affaddr{Institute for Clarity in Documentation}\\
+%       \affaddr{1932 Wallamaloo Lane}\\
+%       \affaddr{Wallamaloo, New Zealand}\\
+%       \email{trovato@corporation.com}
+% 2nd. author
+%\alignauthor
+%G.K.M. Tobin\titlenote{The secretary disavows
+%any knowledge of this author's actions.}\\
+%       \affaddr{Institute for Clarity in Documentation}\\
+%       \affaddr{P.O. Box 1212}\\
+%       \affaddr{Dublin, Ohio 43017-6221}\\
+%       \email{webmaster@marysville-ohio.com}
+% 3rd. author
+
+\maketitle
 \begin{abstract}
-Networks with single points of failure are particularly susceptible
-to adversarial faults,
-in which an attacker creates faults strategically at central points.
+Networks with single points of failure are particularly susceptible to
+adversarial faults, in which an attacker targets nodes strategically.
 Centralized network architectures necessarily exhibit this vulnerability,
-leading to a growing popularity in decentralized architectures.
-However, decentralized architectures do not guarantee the absence of single
-points of failure, and networks such as the Internet and World-Wide Web have
-been shown theoretically and historically to be highly susceptible to
-adversarial faults.
-We develop a partial trust model which makes weaker transitivity
-assumptions than common web of trust models.
-Using this model, we propose a general multipath fault tolerance technique for
-structured networks which
-provides asymptotic security against adversarial faults.
-We also provide a specific algorithm for implementing this technique on the
-butterfly network topology.
-When network topology can be dictated, these results can be used to create scalable, attack-tolerant infrastructures.
-More generally, our results provide a formalism for evaluating the effects
-of network structure on adversarial fault tolerance.
+leading to a growing popularity in decentralized architectures and protocols.
+However, centralized network structure can arise even when protocols are
+decentralized.
+While based on decentralized protocols, the Internet and World-Wide Web have
+been shown both theoretically and historically to be highly susceptible to
+adversarial faults, in part due to emergent, structural centralization.
+Existing network trust models, such as web of trust, fail to adequately address
+network structure. We develop a partial trust model that incorporates network
+structure and which makes weaker, more realistic transitivity assumptions than
+web of trust. We show that the partial trust model, when combined with
+concurrent multipath routing, allows standard fault tolerance techniques to be
+applied to communication across complex networks. We also provide a specific
+algorithm for implementing this scheme on the butterfly network topology.
+When network topology can be dictated, these results can be used to create
+scalable, attack-tolerant infrastructures. More generally, our results provide
+a formalism for evaluating the effects of network structure on adversarial
+fault tolerance.
 \end{abstract}
 
-%ACM is moving forward with the 2012 Classification system: http://dl.acm.org/ccs_flat.cfm. Please generate the CCSXML tex code through the online interactive system and insert the code below.
-
+%
+% The code below should be generated by the tool at
+% http://dl.acm.org/ccs.cfm
+% Please copy and paste the code instead of the example below. 
+%
 \begin{CCSXML}
@@
 
-%\terms{Design, Algorithms, Performance}
-
-\keywords{
-censorship,
-decentralization,
-fault tolerance,
-multipath,
-networks,
-peer-to-peer,
-trust
-}
-
-\begin{bottomstuff}
-%This work is supported by TODO.
-Author's addresses: E.L. Platt and D.M. Romero, School of Information; email: \{elplatt,drom\}@umich.edu.
-\end{bottomstuff}
-
-\maketitle
-
-%TODO:\\
-%Fix bibtex encoding\\
-%Move secrecy section\\
-%Update intro to make significance more clear\\
-%Clarify identifying fault\\
+%
+% End generated code
+%
+
+%
+%  Use this command to print the description
+%
+\printccsdesc
+
+% We no longer use \terms command
+%\terms{Theory}
+
+\keywords{networks; decentralization; fault tolerance; routing; multipath routing; adversarial faults; information security}
 
@@
 \begin{figure}
-\centerline{\includegraphics[height =.4 \textwidth]{fig-alice_trusted_neigh2}}
+\centerline{\includegraphics[width=2.12in,height=1.46in]{fig-alice_trusted_neigh2}}
 \caption{
@@
 \begin{figure}
-\centerline{\includegraphics[height =.4 \textwidth]{fig-bob_Alice_trusted_neigh2}}
+\centerline{\includegraphics[width=3.33in,height=1.46in]{fig-bob_Alice_trusted_neigh2}}
 \caption{
@@
 \begin{table}%
-\tbl{Butterfly Multipath Routing Variables\label{tab:routing}}
+\caption{Butterfly Multipath Routing Variables\label{tab:routing}}
 {
@@
 
-% Acknowledgments
-\begin{acks}
-\end{acks}
-
-% Bibliography
-\bibliographystyle{ACM-Reference-Format-Journals}
-\bibliography{paper}
-
+%ACKNOWLEDGMENTS are optional
+\section{Acknowledgments}
+Anonymized for blind review.
+
+%
+% The following two commands are all you need in the
+% initial runs of your .tex file to
+% produce the bibliography for the citations in your paper.
+\bibliographystyle{abbrv}
+\bibliography{paper}  % sigproc.bib is the name of the Bibliography in this case
+% You must have a proper ".bib" file
+%  and remember to run:
+% latex bibtex latex latex
+% to resolve all references
+%
+% ACM needs 'a single self-contained file'!
+%
+%
 \end{document}

### 2016-08-02 wsdm-2017/source/sig-alternate-sample.tex
Add WSDM 2017 manuscript.

\documentclass{sig-alternate-05-2015}

\usepackage[ruled]{algorithm2e}
\renewcommand{\algorithmcfname}{ALGORITHM}
\usepackage{amsmath,amsfonts,amssymb,bbm} 
%\usepackage[numbers,sort&compress]{natbib} % for citet
\SetArgSty{textrm}  % for algorithm2e
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}

\begin{document}

% Copyright
\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\doi{10.475/123_4}

% ISBN
\isbn{123-4567-24-567/08/06}

%Conference
\conferenceinfo{PLDI '13}{June 16--19, 2013, Seattle, WA, USA}

\acmPrice{\$15.00}

%
% --- Author Metadata here ---
\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Alternate {\ttlit ACM} SIG Proceedings Paper in LaTeX
Format\titlenote{(Produces the permission block, and
copyright information). For use with
SIG-ALTERNATE.CLS. Supported by ACM.}}
\subtitle{[Extended Abstract]
\titlenote{A full version of this paper is available as
\textit{Author's Guide to Preparing ACM SIG Proceedings Using
\LaTeX$2_\epsilon$\ and BibTeX} at
\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{8} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{1932 Wallamaloo Lane}\\
       \affaddr{Wallamaloo, New Zealand}\\
       \email{trovato@corporation.com}
% 2nd. author
\alignauthor
G.K.M. Tobin\titlenote{The secretary disavows
any knowledge of this author's actions.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{P.O. Box 1212}\\
       \affaddr{Dublin, Ohio 43017-6221}\\
       \email{webmaster@marysville-ohio.com}
% 3rd. author
\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
one who did all the really hard work.}\\
       \affaddr{The Th{\o}rv{\"a}ld Group}\\
       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
       \affaddr{Hekla, Iceland}\\
       \email{larst@affiliation.org}
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor Lawrence P. Leipuner\\
       \affaddr{Brookhaven Laboratories}\\
       \affaddr{Brookhaven National Lab}\\
       \affaddr{P.O. Box 5000}\\
       \email{lleipuner@researchlabs.org}
% 5th. author
\alignauthor Sean Fogarty\\
       \affaddr{NASA Ames Research Center}\\
       \affaddr{Moffett Field}\\
       \affaddr{California 94035}\\
       \email{fogartys@amesres.org}
% 6th. author
\alignauthor Charles Palmer\\
       \affaddr{Palmer Research Laboratories}\\
       \affaddr{8600 Datapoint Drive}\\
       \affaddr{San Antonio, Texas 78229}\\
       \email{cpalmer@prl.com}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
This paper provides a sample of a \LaTeX\ document which conforms,
somewhat loosely, to the formatting guidelines for
ACM SIG Proceedings. It is an {\em alternate} style which produces
a {\em tighter-looking} paper and was designed in response to
concerns expressed, by authors, over page-budgets.
It complements the document \textit{Author's (Alternate) Guide to
Preparing ACM SIG Proceedings Using \LaTeX$2_\epsilon$\ and Bib\TeX}.
This source file has been written with the intention of being
compiled under \LaTeX$2_\epsilon$\ and BibTeX.

The developers have tried to include every imaginable sort
of ``bells and whistles", such as a subtitle, footnotes on
title, subtitle and authors, as well as in the text, and
every optional component (e.g. Acknowledgments, Additional
Authors, Appendices), not to mention examples of
equations, theorems, tables and figures.

To make best use of this sample document, run it through \LaTeX\
and BibTeX, and compare this source code with the printed
output produced by the dvi file. A compiled PDF version
is available on the web page to help you with the
`look and feel'.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>  
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

%
% End generated code
%

%
%  Use this command to print the description
%
\printccsdesc

% We no longer use \terms command
%\terms{Theory}

\keywords{ACM proceedings; \LaTeX; text tagging}

\section{Introduction}
The \textit{proceedings} are the records of a conference.
ACM seeks to give these conference by-products a uniform,
high-quality appearance.  To do this, ACM has some rigid
requirements for the format of the proceedings documents: there
is a specified format (balanced  double columns), a specified
set of fonts (Arial or Helvetica and Times Roman) in
certain specified sizes (for instance, 9 point for body copy),
a specified live area (18 $\times$ 23.5 cm [7" $\times$ 9.25"]) centered on
the page, specified size of margins (1.9 cm [0.75"]) top, (2.54 cm [1"]) bottom
and (1.9 cm [.75"]) left and right; specified column width
(8.45 cm [3.33"]) and gutter size (.83 cm [.33"]).

The good news is, with only a handful of manual
settings\footnote{Two of these, the {\texttt{\char'134 numberofauthors}}
and {\texttt{\char'134 alignauthor}} commands, you have
already used; another, {\texttt{\char'134 balancecolumns}}, will
be used in your very last run of \LaTeX\ to ensure
balanced column heights on the last page.}, the \LaTeX\ document
class file handles all of this for you.

The remainder of this document is concerned with showing, in
the context of an ``actual'' document, the \LaTeX\ commands
specifically available for denoting the structure of a
proceedings paper, rather than with giving rigorous descriptions
or explanations of such commands.

\section{The {\secit Body} of The Paper}
Typically, the body of a paper is organized
into a hierarchical structure, with numbered or unnumbered
headings for sections, subsections, sub-subsections, and even
smaller sections.  The command \texttt{{\char'134}section} that
precedes this paragraph is part of such a
hierarchy.\footnote{This is the second footnote.  It
starts a series of three footnotes that add nothing
informational, but just give an idea of how footnotes work
and look. It is a wordy one, just so you see
how a longish one plays out.} \LaTeX\ handles the numbering
and placement of these headings for you, when you use
the appropriate heading commands around the titles
of the headings.  If you want a sub-subsection or
smaller part to be unnumbered in your output, simply append an
asterisk to the command name.  Examples of both
numbered and unnumbered headings will appear throughout the
balance of this sample document.

Because the entire article is contained in
the \textbf{document} environment, you can indicate the
start of a new paragraph with a blank line in your
input file; that is why this sentence forms a separate paragraph.

\subsection{Type Changes and {\subsecit Special} Characters}
We have already seen several typeface changes in this sample.  You
can indicate italicized words or phrases in your text with
the command \texttt{{\char'134}textit}; emboldening with the
command \texttt{{\char'134}textbf}
and typewriter-style (for instance, for computer code) with
\texttt{{\char'134}texttt}.  But remember, you do not
have to indicate typestyle changes when such changes are
part of the \textit{structural} elements of your
article; for instance, the heading of this subsection will
be in a sans serif\footnote{A third footnote, here.
Let's make this a rather short one to
see how it looks.} typeface, but that is handled by the
document class file. Take care with the use
of\footnote{A fourth, and last, footnote.}
the curly braces in typeface changes; they mark
the beginning and end of
the text that is to be in the different typeface.

You can use whatever symbols, accented characters, or
non-English characters you need anywhere in your document;
you can find a complete list of what is
available in the \textit{\LaTeX\
User's Guide}\cite{Lamport:LaTeX}.

\subsection{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of
the three are discussed in the next sections.

\subsubsection{Inline (In-text) Equations}
A formula that appears in the running text is called an
inline or in-text formula.  It is produced by the
\textbf{math} environment, which can be
invoked with the usual \texttt{{\char'134}begin. . .{\char'134}end}
construction or with the short form \texttt{\$. . .\$}. You
can use any of the symbols and structures,
from $\alpha$ to $\omega$, available in
\LaTeX\cite{Lamport:LaTeX}; this section will simply show a
few examples of in-text equations in context. Notice how
this equation: \begin{math}\lim_{n\rightarrow \infty}x=0\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsubsection{Display Equations}
A numbered display equation -- one set off by vertical space
from the text and centered horizontally -- is produced
by the \textbf{equation} environment. An unnumbered display
equation is produced by the \textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols
and structures available in \LaTeX; this section will just
give a couple of examples of display equations in context.
First, consider the equation, shown as an inline equation above:
\begin{equation}\lim_{n\rightarrow \infty}x=0\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}\sum_{i=0}^{\infty} x + 1\end{displaymath}
and follow it with another numbered equation:
\begin{equation}\sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\subsection{Citations}
Citations to articles \cite{bowman:reasoning,
clark:pct, braams:babel, herlihy:methodology},
conference proceedings \cite{clark:pct} or
books \cite{salas:calculus, Lamport:LaTeX} listed
in the Bibliography section of your
article will occur throughout the text of your article.
You should use BibTeX to automatically produce this bibliography;
you simply need to insert one of several citation commands with
a key of the item cited in the proper location in
the \texttt{.tex} file \cite{Lamport:LaTeX}.
The key is a short reference you invent to uniquely
identify each work; in this sample document, the key is
the first author's surname and a
word from the title.  This identifying key is included
with each item in the \texttt{.bib} file for your article.

The details of the construction of the \texttt{.bib} file
are beyond the scope of this sample document, but more
information can be found in the \textit{Author's Guide},
and exhaustive details in the \textit{\LaTeX\ User's
Guide}\cite{Lamport:LaTeX}.

This article shows only the plainest form
of the citation command, using \texttt{{\char'134}cite}.
This is what is stipulated in the SIGS style specifications.
No other citation format is endorsed or supported.

\subsection{Tables}
Because tables cannot be split across pages, the best
placement for them is typically the top of the page
nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and
the table caption.  The contents of the table itself must go
in the \textbf{tabular} environment, to
be aligned properly in rows and columns, with the desired
horizontal and vertical rules.  Again, detailed instructions
on \textbf{tabular} material
is found in the \textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table 1 is included in the input file; compare the
placement of the table here with the table in the printed
dvi output of this document.

\begin{table}
\centering
\caption{Frequency of Special Characters}
\begin{tabular}{|c|c|l|} \hline
Non-English or Math&Frequency&Comments\\ \hline
\O & 1 in 1,000& For Swedish names\\ \hline
$\pi$ & 1 in 5& Common in math\\ \hline
\$ & 4 in 5 & Used in business\\ \hline
$\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
\hline\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of
the page's live area, use the environment
\textbf{table*} to enclose the table's contents and
the table caption.  As with a single-column table, this wide
table will ``float" to a location deemed more desirable.
Immediately following this sentence is the point at which
Table 2 is included in the input file; again, it is
instructive to compare the placement of the
table here with the table in the printed dvi
output of this document.

\begin{table*}
\centering
\caption{Some Typical Commands}
\begin{tabular}{|c|c|l|} \hline
Command&A Number&Comments\\ \hline
\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
\texttt{{\char'134}table}& 300 & For tables\\ \hline
\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
\end{table*}
% end the environment with {table*}, NOTE not {table}!

\subsection{Figures}
Like tables, figures cannot be split across pages; the
best placement for them
is typically the top or the bottom of the page nearest
their initial cite.  To ensure this proper ``floating'' placement
of figures, use the environment
\textbf{figure} to enclose the figure and its caption.

This sample document contains examples of \textbf{.eps} files to be
displayable with \LaTeX.  If you work with pdf\LaTeX, use files in the
\textbf{.pdf} format.  Note that most modern \TeX\ system will convert
\textbf{.eps} to \textbf{.pdf} for you on the fly.  More details on
each of these is found in the \textit{Author's Guide}.

\begin{figure}
\centering
\includegraphics{fly}
\caption{A sample black and white graphic.}
\end{figure}

\begin{figure}
\centering
\includegraphics[height=1in, width=1in]{fly}
\caption{A sample black and white graphic
that has been resized with the \texttt{includegraphics} command.}
\end{figure}

As was the case with tables, you may want a figure
that spans two columns.  To do this, and still to
ensure proper ``floating'' placement of tables, use the environment
\textbf{figure*} to enclose the figure and its caption.
and don't forget to end the environment with
{figure*}, not {figure}!

\begin{figure*}
\centering
\includegraphics{flies}
\caption{A sample black and white graphic
that needs to span two columns of text.}
\end{figure*}

\begin{figure}
\centering
\includegraphics[height=1in, width=1in]{rosette}
\caption{A sample black and white graphic that has
been resized with the \texttt{includegraphics} command.}
\vskip -6pt
\end{figure}

\subsection{Theorem-like Constructs}
Other common constructs that may occur in your article are
the forms for logical constructs like theorems, axioms,
corollaries and proofs.  There are
two forms, one produced by the
command \texttt{{\char'134}newtheorem} and the
other by the command \texttt{{\char'134}newdef}; perhaps
the clearest and easiest way to distinguish them is
to compare the two in the output of this sample document:

This uses the \textbf{theorem} environment, created by
the\linebreak\texttt{{\char'134}newtheorem} command:
\newtheorem{theorem}{Theorem}
\begin{theorem}
Let $f$ be continuous on $[a,b]$.  If $G$ is
an antiderivative for $f$ on $[a,b]$, then
\begin{displaymath}\int^b_af(t)dt = G(b) - G(a).\end{displaymath}
\end{theorem}

The other uses the \textbf{definition} environment, created
by the \texttt{{\char'134}newdef} command:
\newdef{definition}{Definition}
\begin{definition}
If $z$ is irrational, then by $e^z$ we mean the
unique number which has
logarithm $z$: \begin{displaymath}{\log e^z = z}\end{displaymath}
\end{definition}

Two lists of constructs that use one of these
forms is given in the
\textit{Author's  Guidelines}.

There is one other similar construct environment, which is
already set up
for you; i.e. you must \textit{not} use
a \texttt{{\char'134}newdef} command to
create it: the \textbf{proof} environment.  Here
is a example of its use:
\begin{proof}
Suppose on the contrary there exists a real number $L$ such that
\begin{displaymath}
\lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
\end{displaymath}
Then
\begin{displaymath}
l=\lim_{x\rightarrow c} f(x)
= \lim_{x\rightarrow c}
\left[ g{x} \cdot \frac{f(x)}{g(x)} \right ]
= \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
\frac{f(x)}{g(x)} = 0\cdot L = 0,
\end{displaymath}
which contradicts our assumption that $l\neq 0$.
\end{proof}

Complete rules about using these environments and using the
two different creation commands are in the
\textit{Author's Guide}; please consult it for more
detailed instructions.  If you need to use another construct,
not listed therein, which you want to have the same
formatting as the Theorem
or the Definition\cite{salas:calculus} shown above,
use the \texttt{{\char'134}newtheorem} or the
\texttt{{\char'134}newdef} command,
respectively, to create it.

\subsection*{A {\secit Caveat} for the \TeX\ Expert}
Because you have just been given permission to
use the \texttt{{\char'134}newdef} command to create a
new form, you might think you can
use \TeX's \texttt{{\char'134}def} to create a
new command: \textit{Please refrain from doing this!}
Remember that your \LaTeX\ source code is primarily intended
to create camera-ready copy, but may be converted
to other forms -- e.g. HTML. If you inadvertently omit
some or all of the \texttt{{\char'134}def}s recompilation will
be, to say the least, problematic.

\section{Conclusions}
This paragraph will end the body of this sample document.
Remember that you might still have Acknowledgments or
Appendices; brief samples of these
follow.  There is still the Bibliography to deal with; and
we will make a disclaimer about that here: with the exception
of the reference to the \LaTeX\ book, the citations in
this paper are to articles which have nothing to
do with the present subject and are used as
examples only.
%\end{document}  % This is where a 'short' article might terminate

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
This section is optional; it is a location for you
to acknowledge grants, funding, editing assistance and
what have you.  In the present case, for example, the
authors would like to thank Gerald Murray of ACM for
his help in codifying this \textit{Author's Guide}
and the \textbf{.cls} and \textbf{.tex} files that it describes.

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
\appendix
%Appendix A
\section{Headings in Appendices}
The rules about hierarchical headings discussed above for
the body of the article are different in the appendices.
In the \textbf{appendix} environment, the command
\textbf{section} is used to
indicate the start of each Appendix, with alphabetic order
designation (i.e. the first is A, the second B, etc.) and
a title (if you include one).  So, if you need
hierarchical structure
\textit{within} an Appendix, start with \textbf{subsection} as the
highest level. Here is an outline of the body of this
document in Appendix-appropriate form:
\subsection{Introduction}
\subsection{The Body of the Paper}
\subsubsection{Type Changes and  Special Characters}
\subsubsection{Math Equations}
\paragraph{Inline (In-text) Equations}
\paragraph{Display Equations}
\subsubsection{Citations}
\subsubsection{Tables}
\subsubsection{Figures}
\subsubsection{Theorem-like Constructs}
\subsubsection*{A Caveat for the \TeX\ Expert}
\subsection{Conclusions}
\subsection{Acknowledgments}
\subsection{Additional Authors}
This section is inserted by \LaTeX; you do not insert it.
You just add the names and information in the
\texttt{{\char'134}additionalauthors} command at the start
of the document.
\subsection{References}
Generated by bibtex from your ~.bib file.  Run latex,
then bibtex, then latex twice (to resolve references)
to create the ~.bbl file.  Insert that ~.bbl file into
the .tex source file and comment out
the command \texttt{{\char'134}thebibliography}.
% This next section command marks the start of
% Appendix B, and does not continue the present hierarchy
\section{More Help for the Hardy}
The sig-alternate.cls file itself is chock-full of succinct
and helpful comments.  If you consider yourself a moderately
experienced to expert user of \LaTeX, you may find reading
it useful but please remember not to change it.
%\balancecolumns % GM June 2007
% That's all folks!
\end{document}

### 2016-08-03 wsdm-2017/source/paper.tex
Revise abstract and intro.

@@
 adversarial faults, in which an attacker targets nodes strategically.
-Centralized network architectures necessarily exhibit this vulnerability,
+In communication networks, these types of attacks leave users vulnerable to
+censorship and targeted surveillance.
+Centralized networks have single points of failure by definition,
 leading to a growing popularity in decentralized architectures and protocols.
@@
 adversarial faults, in part due to emergent, structural centralization.
-Existing network trust models, such as web of trust, fail to adequately address
+Existing network trust models, such as webs of trust, fail to adequately address
 network structure. We develop a partial trust model that incorporates network
 structure and which makes weaker, more realistic transitivity assumptions than
-web of trust. We show that the partial trust model, when combined with
+webs of trust. We show that the partial trust model, when combined with
 concurrent multipath routing, allows standard fault tolerance techniques to be
 applied to communication across complex networks. We also provide a specific
-algorithm for implementing this scheme on the butterfly network topology.
+routing algorithm for implementing this scheme on the butterfly network
+topology.
 When network topology can be dictated, these results can be used to create
@@
 Large-scale communication networks, exemplified by the Internet,
-have become ubiquitous and critical infrastructural for
-communities, organizations, and markets.
+have become ubiquitous, and are now crucial infrastructure for
+individuals, communities, and organizations around the world.
 As with any critical infrastructure, the cost of a failure can be
@@
 \cite{zin_survey_2015,albert_error_2000,sterbenz_resilience_2010}.
-The key question is: what are the techniques and network structures that can
-be used to create robust, fault-tolerant systems?
+In fact, the decentralized design of the Internet was intended to protect
+against targeted attacks, such as nuclear strikes
+\cite{baran_distributed_1964}.
+Such targeted attacks are formally described as {\em adversarial faults}
+and are some of the most difficult to protect against.
+Both theoretical results and recent events have demonstrated that
+the Internet is surprisingly vulnerable to adversarial faults.
+Decentralization remains a promising approach to building resilient networks,
+but further work is needed to understand the relationship between
+decentralized network structure and adversairal fault tolerance.
 

### 2016-08-03 wsdm-2017/source/paper.tex
Revise intro.

@@
 and are some of the most difficult to protect against.
-Both theoretical results and recent events have demonstrated that
+Both theoretical results and recent events (described below)
+have demonstrated that
 the Internet is surprisingly vulnerable to adversarial faults.
@@
 
-Many complex networks, including the Internet and World-Wide Web, exhibit
-scale-free structure
+Analysis of the Internet's router network has shown that while it is remarkably
+resilient against random faults,
+it is highly susceptible to adversarial faults \cite{albert_error_2000}.
+These results have been attributed to the scale-free structure of the Internet's
+router network
 \cite{barabasi_emergence_1999,barabasi_scale-free_2009}.
-While scale-free networks tolerate random faults well,
-they are highly susceptible to adversarial faults, i.e. attacks
-\cite{albert_error_2000}.
+In scale-free networks and other networks with heavy-tail degree distributions,
+random failures are highly likely to pick low-degree nodes, thus having
+little effect,
+while adversarial faults target the small number of high-degree nodes,
+thus removing a large number of edges with each fault.
+It may seem to be a contradiction that the Internet both be decentralized
+and exhibit a heavy-tailed degree distribution.
+This contradiction is resolved by observing that the Internet is decentralized
+in its {\em protocols}, but somewhat centralized in its {\em network structure}.
+In other words, the protocols of the Internet do not {\em require}
+centralization, but centralization may still emerge from the socio-technical
+processes that create its network structure.
+
 For example, in 2008, YouTube suffered a worldwide outage for several hours

### 2016-08-04 wsdm-2017/source/paper.tex
Revise intro and add references.

@@
 
+\usepackage[utf8]{inputenc}
 \usepackage{tikz}
@@
 
+The Internet's vulnerability to censorship and other targeted attacks
+has been confirmed by the success of several such attacks.
 For example, in 2008, YouTube suffered a worldwide outage for several hours
@@
 within Pakistan only, but resulted in a worldwide cascading failure.
-Such vulnerabilities are not limited to any one system or protocol, but an
-attribute of complex communication networks themselves.
-General techniques for understanding and mitigating these vulnerabilities are
-needed.
-
-We consider a setting in which a source node in a network attempts to route information
-to a target node by forwarding it through the links of the network. We 
-assume that some nodes in the network may be compromised by an attacker 
-and will forward the wrong information to their neighbors or
-block the information altogether.
+Such vulnerabilities are not limited to any one system or protocol,
+but result from the network structure itself.
+With strong theoretical and historical evidence that network structure
+can create vulnerabilities,
+methods for analyzing structural vulnerabilities and for designing
+fault tolerant networks are needed.
+This paper presents several contributions towards advancing those goals.
+
+We consider a setting in which a source node (Alice)
+in a network attempts to route a message
+to a target node (Bob) by forwarding it through the links of the network.
+We assume that some nodes in the network may be compromised by an attacker
+(Eve).
+We assume that Eve has full knowledge of the network structure, but has
+limited resources and thus can only compromise a fixed number of nodes.
+Compromised nodes may behave incorrectly by blocking the message,
+changing the message content, or routing the message incorrectly.
 In order to evaluate network-based techniques for mitigating such attacks,
 we develop a decentralized {\em partial trust} model,
-similar to a web of trust
+similar to the commonly used web of trust approach
 \cite{zimmermann_official_1995,ferguson_practical_2003},
-but making weaker assumptions of trust transitivity.
-While privacy and secrecy are often concerns in secure communication,
+but making weaker (more realistic) assumptions of trust transitivity,
+and incorporating network structure.
+The partial trust model can be used to design and analyze communication
+schemes that tolerate adversarial faults,
+even when no single path through the network is fully trusted.
+
+In addition we provide a novel concurrent multipath routing algorithm
+on the butterfly network topology and use the partial trust model to prove
+a high level of adversarial fault tolerance.
+While the ability to control network topology is only achievable in some
+applications, those applications are nonetheless significant,
+e.g. overlay networks \cite{lua_survey_2005, korzun_structured_2013},
+formal organizations \cite{mohr_explaining_1982},
+government-regulated cellular networks \cite{walker_mass_2012},
+and call tree notification systems \cite{nickerson_thinking_2010}.
+
+While {\em confidentiality} is a common concern in secure communication,
 it is not the focus of this paper.
-Existing techniques for privacy and secrecy are relatively mature compared
-to those for tolerating blocked or corrupted messages.
-In fact, an attacker might use message disruption as a fallback when
-surveillance isn't feasible.
-However, we would like to stress that the technqiues presented in this paper
+Existing cryptographic techniques for confidentiality are relatively
+mature compared to those for tolerating censorship.
+In fact, an attacker might use targeted attacks to achieve censorship
+as a fallback when surveillance isn't feasible.
+However, we stress that the technqiues presented in this paper
 are entirely compatible with, and in some cases could enhance, existing
 privacy and secrecy techniques.
+For example, the well-known {\em man-in-the-middle} attack exploits a privileged
+network position to attack otherwise secure cryptography,
+suggesting that structural approaches can complement cryptographic ones.
 

### 2016-08-04 wsdm-2017/source/paper.tex
Finish revisions to introduction.

@@
 
-In addition we provide a novel concurrent multipath routing algorithm
-on the butterfly network topology and use the partial trust model to prove
-a high level of adversarial fault tolerance.
-While the ability to control network topology is only achievable in some
-applications, those applications are nonetheless significant,
-e.g. overlay networks \cite{lua_survey_2005, korzun_structured_2013},
-formal organizations \cite{mohr_explaining_1982},
-government-regulated cellular networks \cite{walker_mass_2012},
-and call tree notification systems \cite{nickerson_thinking_2010}.
-
 While {\em confidentiality} is a common concern in secure communication,
@@
 
-Given the above framework, we develop and evaluate a \emph{multipath routing} technique to reduce 
-the likelihood that compromised nodes are undetected during the information 
-routing process by allowing multiple copies of a message to be sent through independent paths,
-which can be used for error-detection and error-correction by the receiver.
-
-The multipath routing technique relies on the existence and discoverability of
-multiple redundant paths between the source and the destination. Thus, network 
-structure and routing algorithms play a key role in the success of the technique.
-We formally evaluate the 
+Using the partial trust model, we are able to reduce complex networks to
+simply a number of redundant, independent paths.
+This reduction enables the application of standard error detection techniques,
+which rely on statistically independent errors.
+While adversarial faults along a single path are correlated,
+faults in randomly-chosen independent paths are not.
+Thus fault tolerance techniques can be applied,
+even in the presence of adversarial errors,
+when messages are sent in parallel,
+a technique known as {\em concurrent multipath routing}
+\cite{zin_survey_2015, qadir_exploiting_2015}.
+The receiver can then use the redundant messages to detect and/or correct
+errors.
+Using the partial trust model, we formally evaluate the 
 effects of network structure on attack-resistance and show that the probability
-of undetected 
-errors decreases exponentially with the number of independent paths between 
-source and destination.
-
-The mutlipath routing technique is well-suted for {\em structured networks},
+of undetected errors decreases exponentially with the number of
+independent paths between source and destination,
+even when no individual path is entirely trusted.
+
+Our specific multipath fault tolerance scheme relies on the existence
+and discoverability of multiple redundant paths between the source and
+the destination.
+Implementation thus requires a routing algorithm.
+While not a requirement, this technique is well-suited for
+{\em structured networks},
 in which links between nodes are constrained to a particular architecture.
-Specifically, we employ the butterfly network topology,
-used elsewhere in parallel computing \cite{kshemkalyani_distributed_2008}
-and peer-to-peer \cite{lua_survey_2005, korzun_structured_2013}
-applications.
-For this network topology,
-we propose an efficient, attack-tolerant multipath routing algorithm.
+
+In particular, we provide a novel concurrent multipath routing algorithm
+on the butterfly network topology and use the partial trust model to prove
+a high level of adversarial fault tolerance.
+The butterfly topology is popular in parallel processing
+\cite{kshemkalyani_distributed_2008} and
+peer-to-peer \cite{lua_survey_2005, korzun_structured_2013}
+applications, due to its regular structure and high connectivity.
+While the ability to control network topology is only achievable in some
+applications, those applications are nonetheless significant,
+e.g. overlay networks \cite{lua_survey_2005, korzun_structured_2013},
+formal organizations \cite{mohr_explaining_1982},
+government-regulated cellular networks \cite{walker_mass_2012},
+and call tree notification systems \cite{nickerson_thinking_2010}.
 
@@
 \item{We develop a decentralized {\em partial trust model},
-which makes weaker transitivity assumptions than previous web-of-trust models,
-and use this model to quantify the effects of network structure on
-fault tolerance;}
-\item{We show that as the number of disjoint paths between the neighborhoods
-of the sender and reciever grows,
-the probability of successfully detecting adversarial faults
-approaches 1 exponentially.
-Furthermore, we show that the number of disjoint paths is a function of
-network structure;}
-\item{We present a scalable, efficient, and attack-tolerant multipath
-routing algorithm on the butterfly network topology.}
+which makes weaker transitivity assumptions than web-of-trust models,
+and makes it possible to analyze the effect of network structure on
+adversarial fault tolerance;}
+\item{We show that the partial trust model enables standard fault
+tolerance techniques to be applied to adversarial faults in
+complex communication networks.
+We find that the probability of detecting adversarial faults
+approaches 1 exponentially as the number of independent paths between
+sender and receiver increases;}
+\item{We present a scalable, efficient, and attack-tolerant concurrent
+multipath routing algorithm on the butterfly network topology.}
 \end{itemize}
@@
 infrastructure.
-Section 4. proposes the partial trust model.
-Section 5. proposes a general technique for multipath fault tolerance on
+Section 4. describes the partial trust model.
+Section 5. describes a general technique for multipath fault tolerance on
 structured networks.
-Section 6. proposes a specific application of this technique to the butterfly
-network topology.
-Section 7. discusses these results.
+Section 6. describes a routing algorithm for multipath fault tolerance
+on the butterfly network topology.
+Section 7. discusses the our results.
 And Section 8. concludes.

### 2016-08-04 wsdm-2017/source/paper.tex
s/Eve/Mal.

@@
 We assume that some nodes in the network may be compromised by an attacker
-(Eve).
-We assume that Eve has full knowledge of the network structure, but has
+(Mal).
+We assume that Mal has full knowledge of the network structure, but has
 limited resources and thus can only compromise a fixed number of nodes.
@@
 Such an approach takes advantage of all possible redundancy in the network,
-and requires an adversary (Eve) to compromise at least one node on each path
+and requires an adversary (Mal) to compromise at least one node on each path
 in order to create an undetectable error.
@@
 Any path between Alice and Bob must pass through all such sets,
-so to guarantee an undetectable error, Eve must compromise at least as many
+so to guarantee an undetectable error, Mal must compromise at least as many
 nodes as are in the smallest vertex cut.
@@
 otherwise two paths would have to intersect in the minimum vertex cut.
-Let us assume that Eve is capable of compromising up to $l$ nodes.
-Eve's best adversarial strategy is now to compromise as many nodes in the smallest
+Let us assume that Mal is capable of compromising up to $l$ nodes.
+Mal's best adversarial strategy is now to compromise as many nodes in the smallest
 vertex cut as possible,
@@
 all nodes in the vertex cut are equally likely to contain a path,
-so Eve can do no better than also choosing randomly.
+so Mal can do no better than also choosing randomly.
 If $k > l$, at least one message will get through uncompromised and all
 errors are detectable.
-Otherwise, the probability of Eve producing an undetectable error is
+Otherwise, the probability of Mal producing an undetectable error is
 the probability that all of Alice's paths contain compromised nodes:

### 2016-08-04 wsdm-2017/source/paper.tex
Tweak section structure.
Revise background and related work.

@@
 
-In {\em structured networks}, link structure is predetermined.
-Such networks can be designed to have favorable structural and routing properties,
-at the expense of complicating the addition or removal of nodes.
-Structured networks have been a popular tool in parallel and distributed
-computing architectures \cite{kshemkalyani_distributed_2008}.
-More recently, peer-to-peer systems based on distributed hash tables have used
-structured ``overlay'' networks to map table keys to local TCP/IP routes
-\cite{lua_survey_2005,korzun_structured_2013}.
-
 Many distributed consensus protocols (such as those used by crypto-currencies)
@@
 and proof-of-stake \cite{king_ppcoin:_2012}
-provide better scalability, but wasteful of computational and energy
-resources.
+provide better scalability,
+but are wasteful of computational and energy resources.
 Federated Byzantine Agreement (FBA) \cite{mazieres_stellar_2015}
-is scalable and optimal, guaranteeing that the protocol only fails when
-success is impossible, although it does not provide a way to evaluate
-the probability of failure.
-All of these protocols rely on the assumption that their cryptography
-cannot be compromised.
-While cryptography can be extremely resistant to technological attacks,
-it can still be compromised through coercion (e.g., legal action).
-In additon, the fault-tolerance of FBA depends on redundancy in the network structure
-but leaves an unanswerd question: how should a network be structured to achieve
-fault tolerance?
-This paper addresses both of these issues by analyzing the role of network
-structure on adversarial faults, without relying on assumptions of cryptographic
-integrity.
-
-{\em Multipath} routing uses multiple paths when routing a message through a
-network, in contrsast to traditional {\em unipath} routing, which uses
+is scalable, allows for flexible trust,
+and guarantees that failure only occurs when the structure of the trust
+network makes success impossible.
+However, FBA does not provide a method for optimizing trust networks or
+evaluating the probability of failure.
+The mutlipath fault tolerance scheme we propose is the
+first example we're aware of that allows adversarial fault tolerance
+to be quantified as a function of network structure.
+
+Most existing adversarial fault tolerance schemes focus on end-to-end,
+protocol-based, cryptographic solutions
+\cite{ferguson_practical_2003}.
+There are however, a few notable exceptions that focus on network structure.
+Fiat and Saia described a scheme that combines the butterfly topology
+with expander graphs to create a highly censorship-resistant content-addressable
+network \cite{fiat_censorship_2002},
+although this scheme exhibits poor scaling due to a very high level of data
+replication.
+Perhaps the most mature structural solution is the Freenet collaboration
+\cite{clarke_freenet:_2001}.
+Freenet uses secret sharing and small-world routing
+\cite{zhang_using_2002,kleinberg_small-world_2000}
+to create a content-addressable network with a high level of both
+confidentiality and censorship resistance.
+
+{\em Multipath routing} protocols use multiple paths when
+routing a message through a network,
+in contrsast to traditional {\em unipath} routing, which uses
 a single path.
@@
 \cite{qadir_exploiting_2015}.
-Many of these routing protocols offer better security 
+Many of these routing protocols offer increased confidentiality
 \cite{zin_survey_2015}.
@@
 \cite{kohno_improvement_2012, khalil_unmask:_2010, lou_h-spread:_2006}.
+Most work on multipath routing has been motivated by applications related to
+wireless sensor networks (WSNs),
+and have thus focused on ad-hoc, unstructured networks, often having a central
+base station.
+The partial trust model presented in this paper provides a more general
+way to evaluate the effect of network structure on adversarial fault-tolerance.
 The method of Liu et al.
@@
 to their final destination.
-The butterfly algorithm we present takes a conceptually similar approach.
-Most work on multipath routing has been motivated by applications related to
-wireless sensor networks (WSNs),
-and have thus focused on ad-hoc, unstructured networks, often having a central
-base station.
-The trust model presented in this paper provides a more general way to evaluate
-the effect of network structure on adversarial fault-tolerance.
+The butterfly routing algorithm we present takes a conceptually similar
+approach,
+routing from the source to several intermediaries, and then to the destination.
+
+Our proposed routing algorithm makes use of a
+{\em structured network}, in which link structure is predetermined.
+Such networks can be designed to have favorable structural and routing
+properties,
+at the expense of complicating the addition or removal of nodes.
+Structured networks have been a popular tool in parallel processing
+architectures \cite{kshemkalyani_distributed_2008}.
+More recently, peer-to-peer systems based on distributed hash tables have used
+structured ``overlay'' networks to map table keys to local TCP/IP routes
+\cite{lua_survey_2005,korzun_structured_2013}.
 
@@
 
-\subsection{Trust}
+\section{Trust}
 
@@
 
-\section{Partial Trust}
+\subsection{Partial Trust Model}
 

### 2016-08-04 wsdm-2017/source/paper.tex
Revise attack-tolerant infrastructure section.

@@
 
-\section{Attack-Tolerant Networks Infrastructure}
+\section{Attack-Tolerant Infrastructure}
 
@@
 attack-tolerant network infrastructure and how those properties translate
-into constraints on network structure.
+into constraints on network structure and fault tolerance protocols.
 We pay special attention to a property we call
@@
 \begin{description}
-\item[Scalability]
-For large scale networks it is important that the infrastructure allows for the network to grow while remaining functional.
-In practice, people, devises, and connections, have limited capabilities and these limitations need to be considered as part of the design of the infrastructure.
-
-\item [Decentralization]
-Systems having single points of failure are less tolerant against faults at those
-points, which both raises the likelihood of a failure and creates targets
-where attackers might be able to focus their resources efficiently.
-Attack-tolerant infrastructure must be decentralized in order to minimize single points of failure.
-
-\item[Stabilizing asymmetry]
+\item[Scalability:]
+For large scale networks it is important that the infrastructure allows
+for the network to grow while remaining functional.
+In practice, people, devices, and connections, have limited capabilities
+and these limitations need to be considered as part of the design of the
+infrastructure.
+
+\item [Decentralization:]
+Systems having single points of failure are less tolerant against faults at
+those points.
+The existence of such points not only increases the likelihood that an attack
+will succeed,
+but also incentivises attack by presenting effective targets.
+Attack-tolerant infrastructure must be decentralized in order to
+minimize single points of failure.
+
+\item[Stabilizing asymmetry:]
 In the context of international conflict,
-\cite{mack_why_1975}
-observed that power imbalance usually determines the outcome of a conflict
-(with the more powerful side winning)
-except in the special case of {\em asymmetric conflicts}.
+{\em asymetric conflicts} are a special case that makes it possible for the
+less powerful party to have an advantage over the more powerful party
+\cite{mack_why_1975}.
 In asymmetric conflicts, the same level of resource expenditure yields different

### 2016-08-04 wsdm-2017/source/paper.tex
Revise Trust section intro.

@@
 \begin{description}
-\item [Sparsity and low diameter]
+\item [Sparsity and low diameter:]
 To achieve scalability, networks must be {\em sparse} and have a
@@
 In practical settings, humans and devices have an upper limit on the number
-of connections they can maintain.
+of connections they can maintain (e.g., Dunbar's number
+\cite{dunbar_neocortex_1992}).
 In sparse networks, the number of links grows slowly as the network grows in
@@
 
-\item [Vertex transitivity]
-In vertex transitive networks, for any pair of nodes, an edge-preserving map
-always exists from one node to the other.
+\item [Uniform centrality:]
+There are many ways to measure node centrality in networks
+\cite{freeman_centrality_1978}.
+The more uniform these measures are across nodes, the more decentralized
+a network is.
+Centrality is minimized in {\em vertex transitive} networks,
+for which an edge-preserving map
+always exists from any node to any other node.
 In other words, all nodes occupy structurally indistinguishable positions
-in the network. Networks with {\em vertex transitivity} are decentralized in the sense that they structure does no allow for  single points of failure.
-
-\item [Redundancy]
-Redundancy in a network refers to the extent to which nodes are connected to each other by multiple independent paths. Redundancy can help reduce single points of failure as well as
+in the network.
+
+\item [Redundancy:]
+In a network, redundancy refers to the existence of multiple non-overlapping
+paths between nodes or components.
+Redundancy can help reduce single points of failure as well as
 increase fault-tolerance.
@@
 A single point of failure occurs when a node holds a uniquely
-central position,
-and can be eliminated by adding alternative
-paths around the node,
-which increases the network's redundancy.
-Similarly, as we will later show, redundancy in a network can enable fault-tolerance techniques
+central position.
+When alterative paths are added to bypass central nodes,
+the network becomes more redundant.
+Similarly, as we will later show,
+redundancy in a network can enable fault-tolerance techniques
 that reduce the effectiveness of attacks and creating stabilizing asymmetry.
@@
 divided into parts and sent along different paths so even if one path is
-compromised, the whole message is not revealed \cite{}.
+compromised, the whole message is not revealed
+\cite{shamir_how_1979, blakley_safeguarding_1979}.
 
@@
 
-The sparsity requirement has an important consequence for network infrastructure:
-as a network grows, the number of nodes will exceed the number of links a given
-node can support and many pairs of nodes will only be indirectly connected,
-with all paths between them passing through at least one intermediary node.
+The requirement that networks remain sparse has important implications for
+resilience.
+As such networks grow, the number of nodes will exceed the number of links
+any single node can support.
+As a result, some pairs of nodes are only connected indirectly,
+through at least one intermediary node.
 These intermediary nodes are targets for potential attacks.
-More specifically, the threat model is as follows: Alice sends a message to
-Bob via one or more intermediaries, but some intermediaries may try to
-either block or alter the message.
+In communication networks, 
+the threat model is as follows: Alice sends a message to
+Bob via one or more intermediaries,
+but some intermediaries may block, change, or re-route the message.
 The ability to tolerate these types of adversarial attacks depends on
-Alice and Bob's ability to assume that some nodes have not been
-compromised---to {\em trust} them.
+Alice and Bob's ability to either find a path they {\em trust},
+or to communicate successfully across untrusted paths.
 
@@
 centralization.
-Alternatively, the decentralized {\em web of trust} model
+
+The well-known and widely-used {\em web of trust} model
 \cite{zimmermann_official_1995,ferguson_practical_2003}
-allows individuals
-to choose who they trust initially.
+is a decentralized alternative.
+In a web of trust,
+individuals choose who they trust initially.
 Trust is then extended to new individuals if they are vouched for by a
-currently-trusted individual.
+currently-trusted individual,
+making it possible to extend their web of trust to a large
+number of nodes.
 However, the web of trust approach assumes that trust can be extended
@@
 \cite{christianson_why_1997}.
-We note that while such approaches are decentralized in the sense that no central authority holds the keys,
-they do not necessarily incorporate redundancy and the network they form may not be vertex transitive. Thus they still may suffer from single
-points of failure.
+The web of trust model also fails to account for the structure
+of the trust network:
+trusted nodes are all trusted equally,
+regardless of how many nodes trust them.
+Similarly, while the process of building a web of trust is decentralized,
+the emergent structure of the trust network itself can be centralized.
+In order to analyze the influence of network structure on adversarial
+fault tolerance,
+a different trust model is needed.
 

### 2016-08-05 wsdm-2017/source/paper.tex
Finish revising trust section.

@@
 
-We now propose a decentralized trust model based on weaker transitivity
-assumptions than existing web of trust models.
-In following sections, we will use this model along with network redundancy to minimize vulnerabilities due to single points of failure.
-Our model is based on the assumption of partially-transitive trust.
-While this assumption is still idealized, it is a more realistic improvement
-on fully-transitive models.
-
-Our model (Figure \ref{fig:trust-source})
-is represented as an undirected graph,
-with nodes representing individual
-agents and edges representing mutual trust relationships.
+We now propose the {\em partial trust model}:
+a technique for modeling communication between any two nodes
+in an arbitrary network as communication across a number of
+independent, direct links,
+each of which might be compromised by an adversary.
+Two links are considered {\em independent} if they do not
+share a single point of failure.
+Rather than assuming that trust is infinitely transitive (as in webs of trust),
+this model assumes that trust is transitive up to a fixed number of hops.
+The level of trust transitivity enters the model as a parameter called the
+{\em trust radius}.
+Any paths longer than this radius will contain some untrusted links,
+hence {\em partial trust}.
+The benefit of the partial trust model is its ability to abstract
+away the details of the network structure,
+while quantifying the amount of structural redundancy that can be
+utilized by fault tolerance schemes.
+
+To apply the partial trust model to a network,
+we begin with an undirected graph $G = (V,E)$,
+with nodes representing individual agents
+and with edges representing mutual trust relationships
+between those agents.
 We define a {\em trust radius} $h$ and assume that trust is transitive
-up to $h$ hops away.
-In other words, the nodes $v$ and $w$ trust each other if the distance between $v$ and $w$, $d(u,v)$ is less than $h$.
-For a given node $v$, we call the set of nodes that $v$ trusts the {\em trusted neighborhood} $T_h(v)$,
-and all nodes at exactly distance $h$ from $v$ the {\em trust boundary} $B_h(v)$:
+up to $h - 1$ hops away.
+In other words, the nodes $v$ and $w$ trust each other if the distance between
+$v$ and $w$, $d(v,w)$ is less than $h$.
+For a given node $v$,
+we call the set of nodes that $v$ trusts the
+{\em trusted neighborhood} $T_h(v)$,
+and all nodes at exactly distance $h$ from $v$ the
+{\em trust boundary} $B_h(v)$:
 \beq
@@
 \eeq
-We assume that $v$ is very familiar with nodes in its trusted neighbor and $v$ would know if any of these nodes were compromised by an attacher.
-Hence, an adversary of $v$ has no incentive to compromise a node in  $T_h(v)$.
-We thus assume that no nodes within the trusted neighborhood are compromised
-by an adversary of $v$. However, by compromising all nodes in the trust boundary, an adversary can
-ensure that all communications leaving the trusted neighborhood are compromised.
+The trust boundary $B_h$ plays an important role because these nodes are not
+trusted by $v$,
+but if compromised can entirely isolate $v$ from the rest of the network.
+
+Using the above definitions, we now model communication between a
+source $v$ (Alice) and a destination $w$ (Bob).
+We assume that nodes trusted by {\em either} Alice or Bob will not
+compromise the message.
+In other words,
+faults are assumed to occur only in the part of the network lying
+between Alice and Bob's trust boundaries.
+We call this part of the network the {\em untrusted region}.
+Our goal is now to model the maximum number of independent paths a message
+can take through the untrusted region.
+This is exactly the problem of finding $\delta_{v,w,h}$: the max-flow/min-cut of
+the graph after each trusted neighborhood has been
+collapsed into a single vertex.
+Each trust boundary forms a cut of the network and places an upper bound on the
+min-cut:
+\beq
+\delta_{v,w,h} \leq \min\left( \mid B_h(v) \mid, \mid B_h(w) \mid \right).
+\eeq
+Equality holds when there are no bottlenecks within the untrusted region.
+For reasonably decentralized networks, equality should hold for all
+values of $h$.
+The redundancy of the entire graph can be characterized by the minimum over
+all vertex pairs:
+\beq
+\delta_h(V) \equiv \min_{v,w \in V} \delta_{v,w,h}.
+\eeq
+Thus, for any pair of nodes in the network, at least $\delta_h$ independent,
+redundant paths can be contstructed between them.
+While $\delta_h$ is a purely structural property of the network,
+it places an upper bound on the effectiveness of any
+redundancy-based fault tolerance scheme.
+We will now describe such a scheme.
 
@@
 \end{figure}
-
-Now we consider a source $v$ (Alice) and a destination $w$ (Bob),
-shown in Figure \ref{fig:trust-source-destionation}.
-We also consider an adversary who aims at disrupting all possible communication paths between Alice and Bob.
-Since the adversary cannot compromise any nodes in the trusted neighborhoods of Alicia or Bob, the only option is to compromise nodes outside of the trust neighborhoods.
-An adversary could, for example, compromise all possible paths between Alice and Bob by compromising either Alice or Bob's entire trust boundary.
-This would require compromising $\min(|T_h(v)|, |T_h(w)|)$ nodes.
-In a vertex transitive network, these two trust boundaries will be the same size.
-Alternatively, the adversary could try to compromise a small set of nodes such that all paths between Alice and Bob pass through one of these nodes.
-However, If disjoint paths exist connecting each node in Alice's trust boundary to
-a node in Bob's, then the cut size between the two can never be smaller than
-$|T_h(v)| = |T_h(v)|$, so an adversary who compromises fewer nodes cannot compromise all possible paths between Alice and Bob.
 

### 2016-08-06 wsdm-2017/source/paper.tex
Revise intro to partial trust model.

@@
 
-\section{Trust}
+\section{Trust and Fault Tolerance}
 
@@
 Similarly, while the process of building a web of trust is decentralized,
-the emergent structure of the trust network itself can be centralized.
+the emergent structure of the trust network itself might be centralized.
 In order to analyze the influence of network structure on adversarial
@@
 \subsection{Partial Trust Model}
+
+We can allow for more realistic transitivity assumptions and incorporate
+network structure into a trust model with one key insight:
+even if no single path between sender and receiver is fully trusted,
+multiple message copies can be sent along different paths and compared
+by the receiver to detect and correct errors.
+However, in the presence of an attacker, there is only benefit in sending an
+additional message along a path if that path does not share untrusted,
+single points of failure with any of the existing message paths
+(otherwise, the adversary can compromise both messages
+with the same effort required to compromise one of them).
+The maximum number of such paths represents the effective redundancy that
+can be utilized by any redundancy-based fault tolerance scheme.
+We now propose a model for quantifying the effective redundancy of a network
+as a function of its structure and the level of trust transitivity.
+We will then use this model to evaluate a specific fault-tolerance scheme.
 
@@
 
-\section{Multipath Fault Tolerance}
+\subsection{Multipath Fault Tolerance}
 

### 2016-08-06 wsdm-2017/source/paper.tex
Yo dawg, I know you like revisions so I revised your revisions.

--- 2016-08-05 wsdm-2017/source/paper.tex
@@
 
-We now propose the {\em partial trust model}:
-a technique for modeling communication between any two nodes
-in an arbitrary network as communication across a number of
-independent, direct links,
-each of which might be compromised by an adversary.
-Two links are considered {\em independent} if they do not
-share a single point of failure.
-Rather than assuming that trust is infinitely transitive (as in webs of trust),
-this model assumes that trust is transitive up to a fixed number of hops.
-The level of trust transitivity enters the model as a parameter called the
-{\em trust radius}.
-Any paths longer than this radius will contain some untrusted links,
-hence {\em partial trust}.
-The benefit of the partial trust model is its ability to abstract
-away the details of the network structure,
-while quantifying the amount of structural redundancy that can be
-utilized by fault tolerance schemes.
-
-To apply the partial trust model to a network,
-we begin with an undirected graph $G = (V,E)$,
-with nodes representing individual agents
-and with edges representing mutual trust relationships
-between those agents.
-We define a {\em trust radius} $h$ and assume that trust is transitive
-up to $h - 1$ hops away.
-In other words, the nodes $v$ and $w$ trust each other if the distance between
-$v$ and $w$, $d(v,w)$ is less than $h$.
-For a given node $v$,
-we call the set of nodes that $v$ trusts the
-{\em trusted neighborhood} $T_h(v)$,
-and all nodes at exactly distance $h$ from $v$ the
-{\em trust boundary} $B_h(v)$:
+We can allow for more realistic transitivity assumptions and incorporate
+network structure into a trust model with one key insight:
+even if no single path between a sender and receiver is fully trusted,
+multiple copies of a message can be sent along different paths and compared
+by the reciever to detect and correct errors.
+However, in the presence of an attacker, there is only benefit in sending an
+additional message along a path if that path does not share untrusted
+single points of failure with any of the existing message paths
+(otherwise, the adversary can compromise both messages
+by causing a single fault).
+We call paths that do not share any untrusted single points of failure
+{\em independent}.
+The maximum number of independent paths represents the effective redundancy that
+can be utilized by any redundancy-based fault-tolerance scheme.
+We now propose a method for quantifying the effective redundancy of a network
+as a function of its structure and level of trust transitivity.
+For the purpose of analyzing adversarial fault-tolerance,
+this method allows a communication network
+to be modeled as a set of virtual channels that directly
+connect each node pair,
+with each channel providing some level of redundancy.
+
+We now specify the {\em partial trust model}.
+We begin with an undirected graph $G = (V,E)$,
+with vertices representing agents,
+and with edges representing mutually trusted communication links.
+Let $v \in V$ be an arbitrary sender (Alice)
+and $w \in V$ be an arbitrary receiver (Bob).
+We assume the presence of an adversary (Mal) who knows the
+full structure of the network,
+and who can compromise a fixed number of agents,
+gaining complete control of their behavior,
+as long as those agents are not trusted by either Alice or Bob.
+We define a {\em trust radius} $h$ and that nodes $u$ and
+$u^\prime$ trust each other if their distance is less than $h$.
+For a given node $u$,
+we call the set of trusted nodes its
+{\em trusted neighborhood} $T_h(u)$,
+and all nodes at exactly distance $h$ the
+{\em trust boundary} $B_h(u)$:
 \beq
-T_h(v) &=& \left\{ w \mid d(v,w) < h \right\} \\
-B_h(v) &=& \left\{ w \mid d(v,w) = h \right\}.
+T_h(u) &=& \left\{ u \mid d(u,u^\prime) < h \right\} \\
+B_h(u) &=& \left\{ u \mid d(u,u^\prime) = h \right\}.
 \eeq
 The trust boundary $B_h$ plays an important role because these nodes are not
-trusted by $v$,
-but if compromised can entirely isolate $v$ from the rest of the network.
-
-Using the above definitions, we now model communication between a
-source $v$ (Alice) and a destination $w$ (Bob).
-We assume that nodes trusted by {\em either} Alice or Bob will not
-compromise the message.
-In other words,
-faults are assumed to occur only in the part of the network lying
-between Alice and Bob's trust boundaries.
-We call this part of the network the {\em untrusted region}.
+trusted by $u$,
+but if compromised can entirely isolate $u$ from the rest of the network.
+These trust assumptions imply that when Alice sends a message to Bob,
+Mal can only cause faults in the set of nodes outside both of their trusted
+neighborhoods: $V \setminus \left(T_h(v) \cup T_h(w)\right)$.
+We refer to this set of nodes as the {\em untrusted region}.
+
 Our goal is now to model the maximum number of independent paths a message

### 2016-08-06 wsdm-2017/source/paper.tex
Finish revising partial trust.

@@
 
-Our goal is now to model the maximum number of independent paths a message
-can take through the untrusted region.
-This is exactly the problem of finding $\delta_{v,w,h}$: the max-flow/min-cut of
+Having defined the assumptions of the partial trust model,
+we now quantify the effective redundancy between Alice and Bob
+when trust radius $h$ is assumed.
+This quantity, $\delta_{v,w,h}$ is exactly the max-flow/min-cut of
 the graph after each trusted neighborhood has been
@@
 \eeq
-Equality holds when there are no bottlenecks within the untrusted region.
-For reasonably decentralized networks, equality should hold for all
-values of $h$.
+Equality holds when there are no bottlenecks within the untrusted region,
+an indication that the network is decentralized.
 The redundancy of the entire graph can be characterized by the minimum over
@@
 redundant paths can be contstructed between them.
-While $\delta_h$ is a purely structural property of the network,
+While $\delta_h$ is a purely structural property of the graph and
+may have additional applications,
+we are interested in this quantity because
 it places an upper bound on the effectiveness of any
 redundancy-based fault tolerance scheme.
-We will now describe such a scheme.
+Furthermore, by applying the partial trust model,
+an adversarial fault tolerance scheme can treat each pair of nodes as being
+directly connected by a virtual channel with redundancy at least $\delta_h$.
+We will now describe such a scheme and evaluate its fault
+tolerance properties.
 

### 2016-08-07 wsdm-2017/source/paper.tex
Changes to trust and fault tolerance. Add intro to butterfly routing.

@@
 
-While {\em confidentiality} is a common concern in secure communication,
-it is not the focus of this paper.
-Existing cryptographic techniques for confidentiality are relatively
-mature compared to those for tolerating censorship.
-In fact, an attacker might use targeted attacks to achieve censorship
-as a fallback when surveillance isn't feasible.
-However, we stress that the technqiues presented in this paper
-are entirely compatible with, and in some cases could enhance, existing
-privacy and secrecy techniques.
-For example, the well-known {\em man-in-the-middle} attack exploits a privileged
-network position to attack otherwise secure cryptography,
-suggesting that structural approaches can complement cryptographic ones.
-
 Using the partial trust model, we are able to reduce complex networks to
 simply a number of redundant, independent paths.
-This reduction enables the application of standard error detection techniques,
-which rely on statistically independent errors.
+This reduction enables the application of standard error detection techniques
+\cite{avizienis_basic_2004, von_neumann_probabilistic_1956},
+which rely on independent errors.
 While adversarial faults along a single path are correlated,
-faults in randomly-chosen independent paths are not.
+faults in independent paths are not.
 Thus fault tolerance techniques can be applied,
@@
 
-{\em Multipath routing} protocols use multiple paths when
-routing a message through a network,
+{\em Multipath routing} protocols send redundant information
+over multiple paths when routing a message through a network,
 in contrsast to traditional {\em unipath} routing, which uses
@@
 The partial trust model presented in this paper provides a more general
-way to evaluate the effect of network structure on adversarial fault-tolerance.
+way to evaluate the effect of network structure on adversarial fault tolerance.
 The method of Liu et al.
@@
 Redundancy can help reduce single points of failure as well as
-increase fault-tolerance.
+increase fault tolerance.
 One measure of redundancy is given by the ratio of edges to nodes
@@
 Similarly, as we will later show,
-redundancy in a network can enable fault-tolerance techniques
+redundancy in a network can enable fault tolerance techniques
 that reduce the effectiveness of attacks and creating stabilizing asymmetry.
@@
 through at least one intermediary node.
-These intermediary nodes are targets for potential attacks.
+These intermediary nodes are targets for potential attacks,
+making network paths potentially unreliable.
+
+Within the field of {\em fault tolerance},
+many techniques have been developed for building reliable systems
+out of unreliable components
+\cite{avizienis_basic_2004, von_neumann_probabilistic_1956}.
+We will make use of standard fault tolerance terminology, summarized here.
+A {\em fault} occurs when one component
+of a system behaves incorrectly (e.g., a routing node blocking or
+altering a message).
+The result of that fault (e.g., a recipient receiving conflicting messages)
+is an {\em error} state.
+If the error is undetected or incorrectly corrected, the system is
+said to have experienced a {\em failure} (e.g., an altered message is
+accepted as authentic).
+We are concerned in particular with {\em adversarial faults},
+which (as opposed to random faults)
+are chosen strategically to maximize the liklihood of a failure.
+
 In communication networks, 
@@
 The maximum number of independent paths represents the effective redundancy that
-can be utilized by any redundancy-based fault-tolerance scheme.
+can be utilized by any redundancy-based fault tolerance scheme.
 We now propose a method for quantifying the effective redundancy of a network
 as a function of its structure and level of trust transitivity.
-For the purpose of analyzing adversarial fault-tolerance,
+For the purpose of analyzing adversarial fault tolerance,
 this method allows a communication network
-to be modeled as a set of virtual channels that directly
+to be modeled as a set of virtual links that directly
 connect each node pair,
-with each channel providing some level of redundancy.
+with each virtual link providing some level of redundancy.
 
@@
 redundancy-based fault tolerance scheme.
+As a function of $h$, $\delta_h$ describes a network's ability to create
+redundancy from trust.
+For attack-tolerant networks,
+it is thus desirable for $\delta_h$ to increase quickly as a function of $h$.
 Furthermore, by applying the partial trust model,
 an adversarial fault tolerance scheme can treat each pair of nodes as being
-directly connected by a virtual channel with redundancy at least $\delta_h$.
-We will now describe such a scheme and evaluate its fault
-tolerance properties.
+directly connected by a virtual link with least $\delta_h$ redundant
+channels.
+We will now describe such a scheme and evaluate its fault tolerance properties.
 
@@
 
-When the assumptions of the partial trust model hold, a network can achieve
-robust fault tolerance by utilizing the redundancy of multiple paths
-between the sender and receiver.
-Furthermore, this fault tolerance can remain robust even when faults are
-adversarially chosen by an attacker who has access to more resources
-than the sender and receiver.
-On a structured network, multipath fault tolerance reveals not just when
-an error has occured, but also provides information about which paths provide
-conflicting information,
-which could be used to help locate compromised nodes.
-
-In the terminology of fault tolerance, a {\em fault} occurs when one component
-of a system behaves incorrectly (e.g., a routing node blocking or
-altering a message).
-The result of that fault (e.g., a recipient receiving conflicting messages)
-is an {\em error} state.
-If the error is undetected or incorrectly corrected, the system is
-said to have experienced a {\em failure} (e.g., an altered message is
-accepted as authentic).
-We are concerned in particular with {\em adversarial faults},
-which (as opposed to random faults)
-are chosen strategically (within the assumptions of the trust model)
-to maximize the liklihood of a failure.
-In this section, we focus on the simplest possible question:
-when using multipath fault tolerance,
-what is the probability that adversarial faults will cauase an undetectable
-error, and thus guarantee a failure?
-
-They key idea behind multipath fault tolerance is that several copies of
-a message are sent, along multiple routes,
-rather than using traditional unicast routing.
-If any of the message copies differ from one another or do not arrive,
-the receiver can conclude that an error has occurred,
-and possibly correct it, depending on the magnitude of the error
-and the desired fault tolerance properties.
-
-In the simplest such scheme, a sender (Alice) broadcasts copies of a message
-along all possible routes to the receiver (Bob).
-Such an approach takes advantage of all possible redundancy in the network,
-and requires an adversary (Mal) to compromise at least one node on each path
-in order to create an undetectable error.
-But how many nodes is that?
-The answer depends on the smallest {\em vertex cut} of the untrusted
-portion of the network---the smallest set of nodes than can be removed
-to separate the network into two components.
-Any path between Alice and Bob must pass through all such sets,
-so to guarantee an undetectable error, Mal must compromise at least as many
-nodes as are in the smallest vertex cut.
-Alice and Bob's trust boundaries are both vertex cuts,
-so the smaller of the two is an upper bound.
-It is not difficult to construct networks in which the trust boundary is exactly
-the smallest vertex cut,
-and we will show this to be true for the networks we consider below.
-We denote the size of the smallest vertex cut---the
-maximum number of disjoint paths between Alice and Bob's trusted
-neighborhoods---as $B$.
-
-Realistically, the sender may only have the resources to send a limited number of
-copies, or the network may not be able to handle the congestion caused by
-sending all messages along all paths.
-However, we will now show that a network can provide a high level of
-fault tolerance with only a few copies of a message,
-provided the number of disjoint paths $B$ is large.
+Once we have determined a network's effective redundancy,
+we can apply redundancy-based fault tolerance techniques,
+by sending multiple copies of a message
+({\em concurrent multipath routing}).
+We model our sender (Alice) and receiver (Bob) as
+communicating over a direct link with $\delta_h$ redundant channels.
+The partial trust model allows us to make this simplifying assumption
+for analyzing the a fault tolerance scheme,
+but implementing such a scheme will require a method for constructing
+specific network paths.
+We will return to the question of constructing paths in the next section.
+For now, we concern ourselves with the question:
+given that the newtowrk provides $\delta_h$ redundant channels between
+Alice and Bob,
+what is the probability that an adversary (Mal) causes an undetectable
+error?
+
+Let us first assume that Alice sends a message copy over each available
+channel.
+When Bob recieves the messages, there are several possibilities.
+If some of the messages are missing
+(Alice can include the number of messages as part of the message)
+or if some of the messages disagree,
+Bob knows that some of the messages were either blocked or altered,
+and he has successfully detecetd an error.
+Bob has not accepted any compromised information,
+and can request retransmission, so no failure has occurred.
+If Bob receives all of the messages, and they all agree,
+he accepts the message.
+If the messages agree, there are two possible cases.
+The first case is that Mal has not compromised any of the messages,
+and Bob has correctly accepted them, so no failure has occurred.
+The second case is that Mal has compromised {\em all} of the messages,
+so Bob has accepted an erroneous message and a failure has occurred.
+In the present scenario,
+whether this occurs depends only on whether Mal has the resources to
+compromise all of the channels.
+In a more realistic, and more interesting, scenario,
+both Alice and Mal have limited resources and are not able to use or
+compromise all available channels.
 
 In a more sophisticated multipath fault tolerance scheme,
-Alice randomly chooses $k$ paths and sends a copy of
+Alice randomly chooses $k \leq \delta_h$ channels and sends a copy of
 her message on each.
-We further assume that Alice posesses a routing algorithm
-(such as described in the next section)
-that enables her to construct paths that
-do not intersect between trusted neighborhoods.
-So the maximum number of paths Alice could choose is $B$,
-otherwise two paths would have to intersect in the minimum vertex cut.
-Let us assume that Mal is capable of compromising up to $l$ nodes.
-Mal's best adversarial strategy is now to compromise as many nodes in the smallest
-vertex cut as possible,
-succeeding when each of Alice's paths include one of the compromised nodes.
-Since Alice chooses paths randomly,
-all nodes in the vertex cut are equally likely to contain a path,
-so Mal can do no better than also choosing randomly.
+We assume that Mal is capable of compromising $l \leq \delta_h$ channels.
+Having full knowledge of the network,
+Mal's best strategy is to identify a minimum node cut in the network
+and compromise nodes from that cut.
+With this strategy, each compromised node reduces effective redundancy by one,
+effectively compromising one of the virtual channels between Alice and Bob.
+Since Alice chooses channels randomly,
+all channels are equally likely to contain a message,
+so Mal can do no better than also chosing randomly.
 If $k > l$, at least one message will get through uncompromised and all
@@
 Otherwise, the probability of Mal producing an undetectable error is
-the probability that all of Alice's paths contain compromised nodes:
+the probability that all of Alice's chosen channels are compromised:
 \beq
 \label{eq:pf}
-p_f &=& \frac{l!(B-k)!}{B!(l-k)!}.
+p_f &=& \frac{l!(\delta_h-k)!}{\delta_h!(l-k)!}.
 \eeq
-Letting $k=\alpha B$ and $l=\beta B$, then applying Stirling's
+Letting $k=\alpha \delta_h$ and $l=\beta \delta_h$, then applying Stirling's
 approximation gives:
@@
     (1-\alpha)
-\right]^B.
+\right]^{\delta_h}.
 \end{eqnarray}
@@
 Figure \ref{fig:pfail} shows the value of $p_f$
-as a function of $k$ and $l$ for various values of $B$.
+as a function of $k$ and $l$.
 Equation (\ref{eq:pf_approx}) shows that while $p_f$
@@
 redundant paths actually utilized $\alpha$ and compromised $\beta$,
-it decreases exponentially in the total number of {\em possible}
-redundant paths.
-This result is significant because $B$ depends only on the network structure,
+it decreases exponentially with the effective redundancy $\delta_h$.
+This result is significant because $\delta_h$
+depends only on the network structure
+and the strength of trust transitivity,
 not on the amount of resources available to the sender or the adversary.
-In other words, senders can tolerate attacks from significantly more powerful
-adversaries, as long as the network structure provides large $B$.
-Formally we have shown that under the partial trust assumption,
-multipath fault tolerance can provide {\em asymptotic security}
-against adversarial faults if $B$ can be made arbitrarily large.
-Similarly, because the cost of an attack is prohibitively large
-compared to the cost of protecting against an attack,
-this architecture creates what a stabilizing asymmetry.
+In other words, this scheme exhibits a {\em stabilizing asymmetry},
+senders can tolerate attacks from significantly more powerful
+adversaries, as long as the network structure provides large $\delta_h$.
 By putting attackers at a disadvantage, attacks are not only tolerated
-but also disincentivized.
+but also disincentivized,
+further decreasing the likelihood of a successful attack.
+
+In order to derive the above results, we have assumed that Alice and all
+intermediary agents are able to identify specific,
+independent network paths that achieve the effective redundancy $\delta_h$.
+We now proceed to describe a routing algorithm for doing so in the special case
+of the butterfly network topology.
 
@@
 The probability of an undetectable error as a function of the number of
-message copies and the number of adversarial faults,
-for various numbers of redundant paths.
+message copies and the number of adversarial faults.
 }
@@
 \section{Multipath Routing on the Butterfly Topology}
+
+In previous sections, we showed that reliable communication across a network
+can be achieved even when any single message path might be compromised by
+an adversary,
+provided the network has sufficient redundancy,
+and provided the sender and intermediaries know how to route messages
+such that they follow independent paths.
+In this section, we address the latter requirement by proposing a novel routing
+algorithm for constructing independent paths on the butterfly network topology.
+This architecture and routing algorithm achieve an
+effective redundancy that increases exponentially with the trust radius,
+allowing a very high level of fault tolerance.
+
+The structure of the butterfly network is highly constrained,
+making it most suitable for applications where network structure can be
+designed or dictated.
+Examples of such networks include:
+overlay networks \cite{lua_survey_2005, korzun_structured_2013},
+formal organizations \cite{mohr_explaining_1982},
+government-regulated cellular networks \cite{walker_mass_2012},
+and call tree notification systems \cite{nickerson_thinking_2010}.
+For critical security applications in particular,
+the novel level of adversarial fault-tolerance offered by this scheme
+is highly desirable,
+even if it requires new approaches to designing decentralized communication
+and trust networks.
+Lastly, we note that the partial trust model and multipath fault tolerance
+schemes of the previous section do not rely on any particular network
+topology or routing algorithm,
+and are not necessarily limited to such applications.
 
@@
 
+While {\em confidentiality} is a common concern in secure communication,
+it is not the focus of this paper.
+Existing cryptographic techniques for confidentiality are relatively
+mature compared to those for tolerating censorship.
+In fact, an attacker might use targeted attacks to achieve censorship
+as a fallback when surveillance isn't feasible.
+However, we stress that the technqiues presented in this paper
+are entirely compatible with, and in some cases could enhance, existing
+privacy and secrecy techniques.
+For example, the well-known {\em man-in-the-middle} attack exploits a privileged
+network position to attack otherwise secure cryptography,
+suggesting that structural approaches can complement cryptographic ones.
+
 Our present contributions are limited in several ways.

### 2016-08-07 wsdm-2017/source/paper.tex
Revise butterfly topology section.

@@
 
-\section{Multipath Routing on the Butterfly Topology}
+\section{Multipath Butterfly Routing}
 
@@
 provided the network has sufficient redundancy,
-and provided the sender and intermediaries know how to route messages
-such that they follow independent paths.
+and provided the sender and intermediaries know how to route message
+copies along independent paths.
 In this section, we address the latter requirement by proposing a novel routing
@@
 effective redundancy that increases exponentially with the trust radius,
-allowing a very high level of fault tolerance.
+allowing a very high level of adversarial fault tolerance.
 
@@
 is highly desirable,
-even if it requires new approaches to designing decentralized communication
+even if it requires new approaches to designing communication
 and trust networks.
@@
 
-We now present a decentralized, fault-tolerant, multipath routing algorithm
-on a specific topology known as the butterfly network topology
-\cite{kshemkalyani_distributed_2008}.
-Several variations on the butterfly network exist.
-Specifically, we utilize the wrap-around butterfly.
-We denote the $m$-dimensional, directed wrap-around butterfly as $\wbf(m)$:
+We choose the butterfly topology
+\cite{kshemkalyani_distributed_2008}
+because of several desirable properties (described below)
+and because its strucutre allows for relatively straightforward
+design and analysis of routing algorithms.
+While several variations on the butterfly network exist,
+we utilize the wrap-around butterfly.
+We denote the $m$-dimensional, directed wrap-around butterfly as a
+graph $\wbf(m)$:
 \beq
@@
 &=&
-\{((l,z),(l+1 (\text{mod } m),z) \} \\
+\{((l,z),(l+1 \; (\text{mod } m),z) \} \\
 E_\rightarrow
 &=&
-\{(l,z),(l+1 (\text{mod } m),
-(z_0, \ldots, z_{l-1},z_l \oplus 1, z_{l+1}, \ldots, z_{m-1}) \},
+\{(l,z),(l+1 \; (\text{mod } m),
+z \oplus 1_l \},
 \eeq
 where $\mathbb{Z}_m$ is the set of integers modulo $m$,
-and $\oplus$ represents the addition modulo 2.
+$\oplus$ represents copmonent-wise addition modulo 2,
+and $1_l$ is a vector with a 1 in index $l$ and 0 elsewhere.
 Each node is associated with a level $l$ and an $m$-bit string $z$
 known as {\em the place-within-level}.
-There are two types of edges, shown in Figure \ref{fig:butterfly}.
+There are two types of edges (shown in Figure \ref{fig:butterfly}).
 Down edges ($E_\downarrow$) connect nodes sharing the same $z$ value
@@
 The wrap-around butterfly network has a number of properties desirable for
-a scalable, decentralized communication network.
+a scalable, decentralized communication network:
 \begin{description}
-\item[Vertex-transitive]
+\item[Vertex-transitivity:]
 Because the wraparound butterfly is vertex transitive,
-it is maximally decentralized.
-We will also use this property throught the proof, noting that
-the problem of finding a route between arbitrary nodes $\tilde{v}$ and $\tilde{w}$
-can be reduced to finding a route from node $(0,0)$ to some node $w$.
-\item[Small-diameter]
+it is maximally decentralized;
+\item[Small-diameter:]
 For any two nodes, the length of the shortest path between them is
-$O(\log N)$, where N is the number of nodes in the network.
-\item[Sparse]
+$O(\log N)$, where N is the number of nodes in the network;
+\item[Sparsity:]
 With a constant degree of 4, the wraparound butterfly is extremely sparse,
-and can scale indefnitely without node degree becoming a limitation.
-\item [Redundant] 
-The number of independent paths between the $h$-trusted neighborhoods of two
-nodes in the wraparound butterfly increases exponentially in $h$.
-We prove this assertion below by giving an algorithm to explicitly construct
-these paths.
+and can scale indefnitely without node degree becoming a limitation;
+\item [Redundancy:]
+The number of independent paths between two
+nodes in the wraparound butterfly increases exponentially in the trust
+radius $h$ (the routing algorithm below constructs thes explicitly).
 \end{description}
@@
 \caption{
-Schematic illustration of the two types of edges in a directed butterfly network.
-The node $(l,z)$ is shown as the bit string $z$ with a square around the $l$th bit.
-"Down" edges increment $l$, leaving $z$ unchanged, while "down-right" edges
-increment $l$ and invert the $l$th bit of $z$.
+Schematic illustration of the two types of edges in a directed butterfly
+network.
+The node $(l,z)$ is shown as the bit string $z$ with a square around the
+$l$th bit.
+``Down'' edges increment $l$, leaving $z$ unchanged,
+while ``down-right'' edges increment $l$ and invert the $l$th bit of $z$.
 \label{fig:butterfly}
@@
 
-\subsection{Routing Algorithm: Overview}
-
-This section gives an informal overview of the multipath routing algorithm.
-In a wraparound butterfly network with trusted radius $h$,
-the algorithm constructs $2^h$ disjoint paths between the trusted neighborhoods
-of any two nodes $v$ and $w = (l_w, z_w)$.
-Because of vertex transitivity,
-we can make the simplifying assumption that $v = (0,0)$ without loss of generality.
-Each path is parameterized by an $h$-bit string $s \in \mathbb{Z}_2^h$ (see Table \ref{tab:routing}).
-Each path cycles around the levels of the butterfly network at most twice.
-
-During the first cycle, the place-within-level $z$ of all nodes is set such that
-no two paths overlap outside the trusted neighborhoods.
-During the second cycle, the place-within-level is set to its destination value
-and the algorithm terminates when the destination level is reached.
-
-More specifically, consider any node, $(l,z)$, in the path corresponding to $s$. The algorithm can be understood by dividing the $m$ bits of
-$z$ into four segments: $A, B, C, D$ (see figure \ref{fig:route-overview}).
-$A$ is the first $h$ bits,
-and $C$ is the $h$ bits preceeding index $l_w$.
-$B$ is formed by the remaining bits between $A$ and $C$,
-while $D$ refers to the remaining bits after $C$.
-Hence, $A, B, C, $ and $D$ occupy $h$, $l_w - 2h$, $h$, and $m - l_w$ bits of $z$, respectively
-(we leave the caase of overlapping $A$ and $C$ for the formal proof).
-Note that due to the construction of the network,
-as we walk through any path $\{p_1, p_2, \dots\}$,
-the level $l_{t+1}$ of node $p_{t+1}$ is always one higher (mod $m$)
-than level $l_t$ the previous node $p_t$.
-Also, $z_{t+1}$ is always the same as the $z_t$,
-except for possibly the $(l+1)^{th}$ bit of $z_{t+1}$,
-which can be either 0 or 1.
-Hence, as the routing algorithm progresses through the levels $l$,
-the segments of $z$ get updated in order.
-
-There are 7 stages of the algorithm to construct a path from $v$ to $w$.
-Stages (1) through (4) consist of the first cycle thought the levels of
-the network and stages (5) though (7) consists of the second
-(potentially partial) cycle.
-We start at the source node $v = (0,0)$ and begin constructing the path.
-During stage (1), $A$ is set to match $s$.
-During stage (2) the bits of $B$ are inverted.
-During stage (3) a cyclic permutation of $s$ is placed into $C$
-(the details of the permutation are necessary for the overlapping case,
-but not relevant for this overview).
-After this, in stages (4--7), $D$, $A$, $B$, and $C$ are set to their
-destination values (see figure \ref{fig:route-overview}).
-Figure \ref{fig:routing} shows an explicit example of the routing algorithm
-for one value of $s$.
-
-We now argue that in the above routing algorithm,
-given $s, s' \in \mathbb{Z}_2^h$ such that $s \neq s'$,
-the corresponding paths, $P_s$ and $P_{s'}$,
-are disjoint when they are both outside of the trusted neighborhoods of
-$v$ and $w$.
-We informally refer to different segments of nodes in $s$ and $s'$ as
-$A,B,C,D$ and $A',B',C',D'$, respectively.
-
-First note that any overlap between the two paths must occur either while they
-are at the same stage of the routing, or while one is at stage (1) and the other one at (5), one is at stage (2) and the other one at (6), or one is at stage (3) and the other one at (7).
-This is because in any other case,
-the nodes in the paths will be at different levels, $l$,
-and thus they cannot overlap.
-Furthermore, when a path is in stage (1) or (7),
-it is in the trusted neighborhoods of $v$ or $w$,
-hence we are not concerned with overlaps within stages (1) or (7).
-
-When both paths are in stages (2), (3), or (4), they do not overlap because
-segment $A$ is set to $s$ and segment $A'$ is set to $s'$.
-When both paths are in stages (5) or (6),
-they do not overlap because segment $C$ is set to a cyclic permutation of $s$
-and segment $C'$ is the same cyclic permutation of $s'$.
-When one path is in stage (1) and the other one is in stage (5),
-say $P_s$ is in (1) and $P_{s'}$ is in (5),
-they do not overlap because all bits in $B$ are set to the original bits in the
-source node and all bits in $B'$ are set to the inverted bits of the source node.
-When one path is in stage (2) and the other one is in stage (6), they do not
-overlap because at least one pair of bits in $B$ and $B'$ are still inverted
-while stage (6) is being processed.
+The structure of the butterfly network lends itself to a well-known
+(unipath) routing algorithm,
+which we later extend to the multipath case.
+The unipath algorithm first follows a down or down-right edge at every step,
+increasing the level $l$ by 1 and cycling through the
+indeces of the place-within-level.
+If the current node's place-within-level matches the destination node's at
+index $l$,
+a down edge is chosen and the place-within-level does not change.
+Otherwise, a down-right edge is chosen and the $l$th component of the
+place-within-level is flipped,
+after which it matches the destination.
+After $m$ interations of this, all levels have been visited
+and the place-within-level matches that of the destination.
+Simply following down (or up) edges will then increment (decrement) the
+level until the destination node is reached.
+
+\subsection{Multipath Routing Algorithm}
 
@@
 \end{figure}
-
-\subsection{Routing Algorithm: Specification and Proof}
 

### 2016-08-07 wsdm-2017/source/paper.tex
Revise algorithm and discussion.

@@
 a technique known as {\em concurrent multipath routing}
-\cite{zin_survey_2015, qadir_exploiting_2015}.
+\cite{zin_survey_2015, qadir_exploiting_2015, khiani_comparative_2013}.
 The receiver can then use the redundant messages to detect and/or correct
@@
 
+We now present a routing algorithm to construct multiple independent paths
+between two nodes in a butterfly network.
+Independence requires that routes between two nodes $v$ and $w$ have no nodes
+in common, unless those nodes are less than distance $h$ from either $v$ or $w$.
+Informally, the algorithm achieves this by using the first $h$ hops to
+reach as many distinct nodes as possible,
+then routing from each of those to a distinct intermediary node,
+and finally routing from the intermediary to the destination.
+By ensuring that the place-within-level differs for all paths
+(outside of the trusted neighborhoods of $v$ and $w$)
+the algorithm guarnatees independence.
+As with the unipath routing algorithm,
+each of the multiple paths proceed using down and down-right edges,
+cycling through levels one at a time.
+However, we cycle through the levels twice, once to route from $v$ to a
+particular path's intermediary node,
+and again to route from the intermediary to $w$.
+In the first cycle, path independence is guaranteed by ensuring that
+all paths differ in the first $h$ bits of the place-within-level.
+Similarly, in the second cycle,
+independence is guaranteed by ensuring that all paths differ in the
+$h$ bits of the place-within-level preceeding the destination index
+(see Figure \ref{fig:route-overview}).
+
 \begin{figure}
@@
                 minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick},
+                outer sep=0pt, on chain, double},
                         ]
@@
 
-We now specify the multipath routing scheme and prove
+We now formally specify the multipath routing scheme and prove
 that provides $2^h$ disjoint paths between
@@
 preceeding index $l_w$ all matching (not all matching) the bits of $\tilde{s}$,
-where $\tilde{s}$ is a cyclic permutation of $s$ defined below.
+where $\tilde{s}$ is a cyclic permutation of $s$:
 \beq
-Q_{v^{(0)}} &=&
-\{z \in \mathbb{Z}_2^m \mid \;
-\neg \exists i \in \mathbb{Z} : h \leq (i < l_w - h) \land z_i \neq z_i^{(0)} \}
-\\
-\overline{Q_{v^{(0)}}} &=& \mathbb{Z}_2^m \setminus Q_{v^{(0)}}
-\\
-R_s &=&
-\{z \in \mathbb{Z}_{2}^m \mid \;
-\forall i \in \mathbb{Z}_2^h z_i = s_i \}
-\\
-\overline{R_s} &=& \mathbb{Z}_2^m \setminus R_s
-\\
-S_s &=&
-\{z \in \mathbb{Z}_{2}^m \mid \;
-\forall i \in \mathbb{Z}_2^h z_{(l_w - h + i)} = \tilde{s}_i \}
-\\
-\overline{S_s} &=& \mathbb{Z}_2^m \setminus S_s
-\\
 \tilde{s}_i &=& s_{(i + l_w) \text{ mod } h}.
@@
 
-While decentralized architectures have recieved much attention,
-methods for classifying, and evaluating decentralization
-are in their infancy.
-The trust model and multipath fault tolerance technique proposed in this
-paper provide a practical fault tolerance architecture for applications when
-network structure can be imposed,
-as well as a more general method for evaluating the role of network structure
-on fault tolerance.
+While decentralized protocols have received much attention for their potential
+fault tolerance applications,
+a better understanding of how network structure contributes to fault tolerance
+is necessary.
+We have proposed a number of contributions towards this understanding:
+a {\em partial trust model} for quantifying the effective redundancy
+of a network,
+a {\em multipath fault tolerance} scheme for using that redundancy to detect
+errors,
+and a {\em concurrent multipath routing} algorithm enabling this scheme to
+be implemented on the butterfly network.
 
 Fault-tolerant network infrastructures have many direct applications.
-Areas such as cryptocurrency, secure multiparty computation, and wireless
-sensor networks have immediate need for scalable, fault-tolerant
-infrastructures.
-Many Internet services---email, social networks, cloud storage---are still
-highly centralized and vulnearble to technical and non-technical attacks.
-Decentralized fault tolerance is one approach to securing these important
+Areas such as cryptocurrency
+\cite{mazieres_stellar_2015,nakamoto_bitcoin:_2008,king_ppcoin:_2012},
+secure multiparty computation
+\cite{yao_protocols_1982,chaum_multiparty_1988,goldreich_how_1987},
+and wireless sensor networks
+\cite{khiani_comparative_2013}
+have immediate need for scalable, fault-tolerant infrastructures.
+Many Internet services (e.g., email, social networks, cloud storage)
+are still highly centralized and vulnearble to technical and
+non-technical (i.e. coercive) attacks.
+Fault tolerance using {\em both} decentralized protocols and
+decentralized network structures
+is one approach to securing these important
 services and making networked communication safer.
 
-While the primary purpose of our proposed fault tolerance technique is the
-threat of message blocking or corruption, it can provide some improvements
-to secrecy and privacy.
-For example, if a message is broken into parts and each part is sent along
-a different random subset of routes, an attacker who succesfully compromises
-one part of the message will still be unable to compromise the rest.
-
-While {\em confidentiality} is a common concern in secure communication,
-it is not the focus of this paper.
+We have also focused specifically on adversarial faults that block or
+change messages (e.g., censorship).
+We have not specifically addressed {\em confidentiality}:
+how to keep those messages private.
 Existing cryptographic techniques for confidentiality are relatively
@@
 as a fallback when surveillance isn't feasible.
-However, we stress that the technqiues presented in this paper
+However, the technqiues presented in this paper
 are entirely compatible with, and in some cases could enhance, existing
-privacy and secrecy techniques.
+confidentiality techniques.
 For example, the well-known {\em man-in-the-middle} attack exploits a privileged
@@
 
-Our present contributions are limited in several ways.
-The partial trust model, while making fewer assumptions than web of trust
-approaches, still depends on some transitivity of trust,
-which may be a questionable assumption in many cases.
-Better trust models are still needed.
-
-The primary limitation of our structured network approach is that
-link structure is imposed rather than self-organized.
-This requirement may be appropriate for appliations such as
-wireless sensor networks,
-or slow-changing institutional infrastructure,
-in which the architect has considerable control over link structure.
-But in general, it remains an open question whether self-organized networks
-can be restructured, or buit from the ground up in such a way that
-structured multipath fault tolerance technqiues can be applied.
-Similarly, structured network topologies such as the butterfly are defined
-only for specific values.
-Although these values can be as large as desired,
-how to apply methods such as ours to networks with an intermediate number
-of nodes remains an open question.
+Our multipath fault tolerance scheme can be applied to any network
+that has both sufficient redundancy and a routing algorithm
+that maps local links to independent network paths.
+These requirements make the approach well-suited for structured networks,
+in which link structure can be dictated,
+such as the butterfy.
+While the ability to control network topology is only achievable in some
+applications, those applications are nonetheless significant
+(e.g. overlay networks \cite{lua_survey_2005, korzun_structured_2013},
+formal organizations \cite{mohr_explaining_1982},
+government-regulated cellular networks \cite{walker_mass_2012},
+and call tree notification systems \cite{nickerson_thinking_2010}).
+Furthermore, when high levels of adversarial fault tolerance are required,
+new socio-technical systems might choose to incorporate such networks.
+For example,
+a coallition of groups supporting free expression could use our results
+to construct a censorship-resistant communication network
+by cultivating the necessary trust relationships.
+In the general case,
+while the partial trust model can be applied to unstructured networks,
+additional work is needed to identify how to ensure redundancy and
+identify independent paths in unstructured networks.
 

### 2016-08-07 wsdm-2017/source/paper.tex
Revise conclusion.

@@
 We have developed a partial trust model which enables a formal analysis
-of the effect of decentralized network structure on fault tolerance properties.
-We have also proposed a general technique for achieving and evaluating adversarial
-fault tolerance in a structured network with partial trust.
-We found that the probability of an adversary causing an undetectable error
-decreases exponentially as the number of redundant paths between the
-neighborhoods of the sender and receiver grows.
-This property puts attackers at a disadvantage, creating a stabilizing
-asymmetry which disincentivizes attack.
-We also presented a specific implementation of the multipath fault tolerance
-technique on a wraparound butterfly network topology and proved that the
-number of redundant paths grows exponentially with number of network hops
-that can be trusted.
-Our results are directly applicable to architectures in which the link structure
-can be imposed by the architect,
-and provide a formalism that can be used more generally to evaluate the role
-of network structure and redundancy on the fault tolerance of decentralized
-networks.
+of the effect of network structure and trust transitivity on
+redundancy.
+We have also proposed a multipath fault tolerance scheme
+for using this redundancy to achieve adversarial
+fault tolerance.
+Under this scheme, the probability of an adversary causing an undetectable error
+decreases exponentially with the network's effective redundancy.
+Furthermore, a small increase in the number of messages sent can compensate
+for a large increase in the number of messages compromised by an adversary.
+This stabilizing asymetry puts attackers at a disadvantage,
+disincentivizing attack.
+We also presented a novel multipath routing algorithm allowing multipath fault
+tolerance to be used on the butterfly topology,
+and proved that the effective redundancy of the butterfly network increases
+exponentially with the trust radius.
+These results are directly applicable to architectures in which the link
+structure can be imposed by the architect,
+and more generally, provide a formalism that can be used more to evaluate
+the role of network structure, trust transitivity, and effective redundancy
+on adversarial fault tolerance.
 

### 2016-08-08 wsdm-2017/source/paper.tex
Revisions.

@@
 thus removing a large number of edges with each fault.
-It may seem to be a contradiction that the Internet both be decentralized
-and exhibit a heavy-tailed degree distribution.
-This contradiction is resolved by observing that the Internet is decentralized
-in its {\em protocols}, but somewhat centralized in its {\em network structure}.
+So while the {\em protocols} of the Internet are decentralized,
+the {\em network structure} is somewhat centralized. 
 In other words, the protocols of the Internet do not {\em require}
@@
 In order to evaluate network-based techniques for mitigating such attacks,
-we develop a decentralized {\em partial trust} model,
+we develop a {\em partial trust} model,
 similar to the commonly used web of trust approach
@@
 
-Using the partial trust model, we are able to reduce complex networks to
-simply a number of redundant, independent paths.
-This reduction enables the application of standard error detection techniques
+Using the partial trust model, we are able to model the nodes in complex
+networks as being directly linked to each other by a number of
+redundant channels.
+This models enables the application of standard error detection techniques
 \cite{avizienis_basic_2004, von_neumann_probabilistic_1956},
@@
 is scalable, allows for flexible trust,
-and guarantees that failure only occurs when the structure of the trust
-network makes success impossible.
-However, FBA does not provide a method for optimizing trust networks or
-evaluating the probability of failure.
+and optimal fault tolerance for a given network (under its assumptions).
+However, FBA does not provide a method for evaluating the
+fault tolerance properties of different network structures
+or calculating how the failure probability for a particular network.
 The mutlipath fault tolerance scheme we propose is the
@@
 \cite{ferguson_practical_2003}.
-There are however, a few notable exceptions that focus on network structure.
+A fully decentralized architecture must be decentalized both in protocol and
+in network structure.
+There are however, a few notable fault tolerance schemes
+that focus on network structure.
 Fiat and Saia described a scheme that combines the butterfly topology
@@
 \cite{clarke_freenet:_2001}.
-Freenet uses secret sharing and small-world routing
+Freenet uses secret sharing
+\cite{shamir_how_1979, blakley_safeguarding_1979}
+and small-world routing
 \cite{zhang_using_2002,kleinberg_small-world_2000}
@@
 confidentiality and censorship resistance.
+Freenet guarantees that data is stored redundantly,
+but still allows for centralized network structure and thus single points
+of failure as data travels from its origin to the redundant storage locations.
 
@@
 In this section, we describe the functional properties required of an
-attack-tolerant network infrastructure and how those properties translate
-into constraints on network structure and fault tolerance protocols.
+attack-tolerant network,
+which serve as guiding principles in later sections.
 We pay special attention to a property we call
@@
 but also incentivises attack by presenting effective targets.
-Attack-tolerant infrastructure must be decentralized in order to
-minimize single points of failure.
+In order to minimize single points of failure,
+attack tolerant infrastructures must use decentralized protocols and
+decentralized network structures.
 
@@
 In asymmetric conflicts, the same level of resource expenditure yields different
-results for different parties.
-There are two possible cases:
+results for different parties;
 the attacker's resources are either more or less effective than the defender's.
@@
 
-We now describe the structural properties of a network that allow it to achieve the functional properties described above. 
+We are specifically concerned with network structure-based approaches to
+fault tolerance.
+In networks, specific structural properties are required to achieve the
+functional properties described above.
 \begin{description}
@@
 paths between nodes or components.
-Redundancy can help reduce single points of failure as well as
-increase fault tolerance.
+Redundancy can help reduce single points of failure and decrease
+centralization.
 One measure of redundancy is given by the ratio of edges to nodes
@@
 the network becomes more redundant.
-Similarly, as we will later show,
-redundancy in a network can enable fault tolerance techniques
-that reduce the effectiveness of attacks and creating stabilizing asymmetry.
-When secrecy is desired, redundancy can help by allowing messages to be
-divided into parts and sent along different paths so even if one path is
-compromised, the whole message is not revealed
-\cite{shamir_how_1979, blakley_safeguarding_1979}.
 
@@
 is an {\em error} state.
-If the error is undetected or incorrectly corrected, the system is
+If the error is undetected or corrected to the wrong value, the system is
 said to have experienced a {\em failure} (e.g., an altered message is
@@
 transitively,
-which is not generally true
+which is unrealistic
 \cite{christianson_why_1997}.
@@
 We now specify the {\em partial trust model}.
-We begin with an undirected graph $G = (V,E)$,
-with vertices representing agents,
+We assume an undirected graph $G = (V,E)$,
+although the model can easily be extended to directed multigraphs.
+Vertices representing agents,
 and with edges representing mutually trusted communication links.
@@
 This quantity, $\delta_{v,w,h}$ is exactly the max-flow/min-cut of
-the graph after each trusted neighborhood has been
-collapsed into a single vertex.
+the graph after Alice's and Bob's trusted neighborhoods have been
+collapsed into single source/sink vertices.
 Each trust boundary forms a cut of the network and places an upper bound on the
@@
 redundancy from trust.
+Combined with a {\em routing algorithm} that tells nodes how to forward
+messages to guarantee they take independent paths,
+this redudnacy can be used to achieve fault tolerance.
 For attack-tolerant networks,
@@
 The partial trust model allows us to make this simplifying assumption
-for analyzing the a fault tolerance scheme,
+for analyzing a fault tolerance scheme,
 but implementing such a scheme will require a method for constructing
@@
 Bob knows that some of the messages were either blocked or altered,
-and he has successfully detecetd an error.
+and he has successfully detected an error.
 Bob has not accepted any compromised information,
@@
 copies along independent paths.
-In this section, we address the latter requirement by proposing a novel routing
+In this section, we address the both requirements by proposing a novel routing
 algorithm for constructing independent paths on the butterfly network topology.
@@
                 minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, double},
+                outer sep=0pt, on chain, thick},
                         ]
@@
 {
+\begin{center}
 \begin{tabular}{|l|l|}
@@
 \end{tabular}
+\end{center}
 }

### 2016-08-08 wsdm-2017/source/paper.tex
Revise partial trust figure. Use theorem/proof format.

@@
 \DeclareMathOperator{\wbf}{wBF}
+\newtheorem{theorem}{Theorem}
 
@@
 \begin{figure}
-\centerline{\includegraphics[width=2.12in,height=1.46in]{fig-alice_trusted_neigh2}}
+\centerline{\includegraphics[width=3in,height=2.08in]{fig-partial-trust}}
 \caption{
-Illustration of the {\em partial trust} model.
-Edges represent direct, mutual trust relationships.
-Alice ($v$) trusts all nodes less than
-$h$ hops away---her {\em trusted neighborhood} $T_h(v)$.
-Beyond that, the nodes at distance $h$ form her {\em trust boundary} $B_h(v)$.
-We assume adversaries are unable to compromise nodes in the trusted neighborhood.
-By compromising all nodes in the trust boundary, an adversary can ensure that
-all communications leaving Alice's trusted neighborhood are compromised.
+Illustration of a trusted communication network and the network properties
+used by the {\em partial trust model}.
+Edges represent mutually trusted communication links.
+The sender (Alice, $v$) and receiver (Bob, $w$) trust all nodes
+less than the {\em trust radius} $h$ hops away.
+These nodes form their {\em trusted neighborhoods} $T_h(v)$ and $T_h(w)$.
+We assume that all faults occur in the remaining nodes: the
+{\em untrusted region}.
+The untrusted nodes in contact with the trusted neighborhoods for the
+{\em trust boundaries} $B_h(v)$ and $B_h(w)$,
+which (in the absence of central bottlenecks) determine the
+{\em effective redundancy} $\delta_h$ provided by the network.
+Alice and Bob can be modelled as connected by a direct link with
+at least $\delta_h$ redundant channels.
 }
 \label{fig:trust-source}
-\end{figure}
-
-\begin{figure}
-\centerline{\includegraphics[width=3.33in,height=1.46in]{fig-bob_Alice_trusted_neigh2}}
-\caption{
-Partial trust model with sender (Alice) and receiver (Bob).
-If disjoint paths exist connecting each node in Alice's trust boundary to
-a node in Bob's,
-than an adversary needs to compromise at least $|T_h(v)|$ nodes to
-compromise all possible paths between Alice and Bob.
-}
-\label{fig:trust-source-destionation}
 \end{figure}
@@
 
-The wrap-around butterfly network has a number of properties desirable for
-a scalable, decentralized communication network:
+The wrap-around butterfly network is known to have several of the properties
+we desire for scalable, decentralized communication networks:
 \begin{description}
@@
 \item [Redundancy:]
-The number of independent paths between two
-nodes in the wraparound butterfly increases exponentially in the trust
-radius $h$ (the routing algorithm below constructs thes explicitly).
+Multiple paths exist between any two nodes.
+Specifically, we will prove below that the number of independent paths between two
+nodes increases exponentially with the trust
+radius $h$.
 \end{description}
@@
 
-We now formally specify the multipath routing scheme and prove
-that provides $2^h$ disjoint paths between
-the $h$-hop trusted neighborhoods of any two nodes in an $m$-bit
-wrap-around butterfly network.
+We now begin the formal specification of our multipath routing scheme for the
+wraparound butterfly network.
 Utilizing vertex transitivity, we label the source node as
@@
 
-For all $2^h$ of these paths, the components excluding nodes within $h$ hops
-of $v$ or $w$ (in their trusted neighborhoods) are pairwise disjoint,
-which we now prove.
+\begin{theorem}
+Given an $m$-bit wraparound ($m \geq 0$) butterfly network,
+and a trust radius $h \geq 1$,
+for all node pairs $(v, w)$,
+if there exists a node $u$ that belongs to
+two distinct routes $v^{(\cdot)}_s$ and $v^{(\cdot)}_{s^\prime}$
+($s \neq s^\prime$),
+it implies that $u \in T_h(v) \cup T_h(w)$.
+\end{theorem}
+\begin{proof}
 Nodes from two paths can only coincide if their levels are the same.
@@
 eliminating the last possible collision.
-
-Thus, we have shown that assuming the partial trust model, with trust transitive
+\end{proof}
+
+Thus, assuming the partial trust model with trust transitive
 for $h$ hops, we can construct $2^h$ paths on a wraparound butterfly topology
@@
 destination.
-This is a lower bound on the value $B$ in Equation (\ref{eq:pf}),
+Note that the node sequence $v_s^{(t)}$ can be calculated entirely
+from the source $v$, destination $w$, and path parameter $s$,
+meaning that with this informaiton nodes are able to determine which neighbor
+to route a given message copy to.
+Furthermore, the existence of $2^h$ paths places a lower bound on the
+effective redundancy $\delta_h$,
 showing that the decentralized, redundant, structured networks such as the

### 2016-08-08 wsdm-2017/source/paper.tex
Update routing example and formatting.
Spellcheck.

@@
 \usepackage[utf8]{inputenc}
+\usepackage{afterpage}
 \usepackage{tikz}
@@
 but further work is needed to understand the relationship between
-decentralized network structure and adversairal fault tolerance.
+decentralized network structure and adversarial fault tolerance.
 
@@
 In other words, the protocols of the Internet do not {\em require}
-centralization, but centralization may still emerge from the socio-technical
+centralization, but centralization may still emerge from the sociotechnical
 processes that create its network structure.
@@
 
-Many distributed consensus protocols (such as those used by crypto-currencies)
+Many distributed consensus protocols (such as those used by cryptocurrencies)
 are designed to provide tolerance against arbitrary or adversarial faults.
@@
 or calculating how the failure probability for a particular network.
-The mutlipath fault tolerance scheme we propose is the
+The multipath fault tolerance scheme we propose is the
 first example we're aware of that allows adversarial fault tolerance
@@
 \cite{ferguson_practical_2003}.
-A fully decentralized architecture must be decentalized both in protocol and
+A fully decentralized architecture must be decentralized both in protocol and
 in network structure.
@@
 over multiple paths when routing a message through a network,
-in contrsast to traditional {\em unipath} routing, which uses
+in contrast to traditional {\em unipath} routing, which uses
 a single path.
@@
 \cite{alrajeh_secure_2013},
-and some specificially protect against adversarial faults
+and some specifically protect against adversarial faults
 \cite{kohno_improvement_2012, khalil_unmask:_2010, lou_h-spread:_2006}.
@@
 wireless sensor networks (WSNs),
-and have thus focused on ad-hoc, unstructured networks, often having a central
+and have thus focused on ad hoc, unstructured networks, often having a central
 base station.
@@
 will succeed,
-but also incentivises attack by presenting effective targets.
+but also incentivizes attack by presenting effective targets.
 In order to minimize single points of failure,
@@
 In the context of international conflict,
-{\em asymetric conflicts} are a special case that makes it possible for the
+{\em asymmetric conflicts} are a special case that makes it possible for the
 less powerful party to have an advantage over the more powerful party
@@
 central position.
-When alterative paths are added to bypass central nodes,
+When alternative paths are added to bypass central nodes,
 the network becomes more redundant.
@@
 which (as opposed to random faults)
-are chosen strategically to maximize the liklihood of a failure.
+are chosen strategically to maximize the likelihood of a failure.
 
@@
 multiple copies of a message can be sent along different paths and compared
-by the reciever to detect and correct errors.
+by the receiver to detect and correct errors.
 However, in the presence of an attacker, there is only benefit in sending an
@@
 Thus, for any pair of nodes in the network, at least $\delta_h$ independent,
-redundant paths can be contstructed between them.
+redundant paths can be constructed between them.
 While $\delta_h$ is a purely structural property of the graph and
@@
 messages to guarantee they take independent paths,
-this redudnacy can be used to achieve fault tolerance.
+this redundancy can be used to achieve fault tolerance.
 For attack-tolerant networks,
@@
 For now, we concern ourselves with the question:
-given that the newtowrk provides $\delta_h$ redundant channels between
+given that the network provides $\delta_h$ redundant channels between
 Alice and Bob,
@@
 channel.
-When Bob recieves the messages, there are several possibilities.
+When Bob receives the messages, there are several possibilities.
 If some of the messages are missing
@@
 all channels are equally likely to contain a message,
-so Mal can do no better than also chosing randomly.
+so Mal can do no better than also choosing randomly.
 If $k > l$, at least one message will get through uncompromised and all
@@
 because of several desirable properties (described below)
-and because its strucutre allows for relatively straightforward
+and because its structure allows for relatively straightforward
 design and analysis of routing algorithms.
@@
 where $\mathbb{Z}_m$ is the set of integers modulo $m$,
-$\oplus$ represents copmonent-wise addition modulo 2,
+$\oplus$ represents componentwise addition modulo 2,
 and $1_l$ is a vector with a 1 in index $l$ and 0 elsewhere.
@@
 With a constant degree of 4, the wraparound butterfly is extremely sparse,
-and can scale indefnitely without node degree becoming a limitation;
+and can scale indefinitely without node degree becoming a limitation;
 \item [Redundancy:]
@@
 increasing the level $l$ by 1 and cycling through the
-indeces of the place-within-level.
+indices of the place-within-level.
 If the current node's place-within-level matches the destination node's at
@@
 after which it matches the destination.
-After $m$ interations of this, all levels have been visited
+After $m$ iterations of this, all levels have been visited
 and the place-within-level matches that of the destination.
@@
 (outside of the trusted neighborhoods of $v$ and $w$)
-the algorithm guarnatees independence.
+the algorithm guaranatees independence.
 As with the unipath routing algorithm,
@@
 independence is guaranteed by ensuring that all paths differ in the
-$h$ bits of the place-within-level preceeding the destination index
+$h$ bits of the place-within-level preceding the destination index
 (see Figure \ref{fig:route-overview}).
 
-\begin{figure}
+\afterpage{
+\begin{table}[h!]
+\caption{Butterfly Multipath Routing Variables\label{tab:routing}}
+{
+\begin{center}
+\begin{tabular}{|l|l|}
+\hline
+NAME & VARIABLE \\\hline
+butterfly dimension & $m \in \mathbb{Z}_+$ \\\hline
+node level & $l \in \mathbb{Z} : 0 \leq l < m$ \\\hline
+node place within level & $z \in \mathbb{Z}_2^m$ \\\hline
+trust radius & $h \in \mathbb{Z} : 1 \leq h \leq \lfloor m/2 \rfloor$ \\\hline
+path index & $s \in \mathbb{Z}_2^h$ \\\hline
+\end{tabular}
+\end{center}
+}
+\end{table}
+
+\begin{figure}[t!]
 \begin{center}
@@
 cycles through the levels of the butterfly network.
-}
 \label{fig:route-overview}
-\end{figure}
-
-\begin{table}%
-\caption{Butterfly Multipath Routing Variables\label{tab:routing}}
-{
-\begin{center}
-\begin{tabular}{|l|l|}
-\hline
-NAME & VARIABLE \\\hline
-butterfly dimension & $m \in \mathbb{Z}_+$ \\\hline
-node level & $l \in \mathbb{Z} : 0 \leq l < m$ \\\hline
-node place within level & $z \in \mathbb{Z}_2^m$ \\\hline
-trust radius & $h \in \mathbb{Z} : 1 \leq h \leq \lfloor m/2 \rfloor$ \\\hline
-path index & $s \in \mathbb{Z}_2^h$ \\\hline
-\end{tabular}
-\end{center}
-}
-\end{table}%
-
-\begin{figure}
+}
 \begin{center}
@@
 \end{figure}
+}
 
@@
 strings in which the bits at
-all indeces $h \leq i < l_w - h$ match (do not all match) those of $z^{(0)}$.
+all indices $h \leq i < l_w - h$ match (do not all match) those of $z^{(0)}$.
 Note that $Q_{v^{(0)}}$ is trivially all $m$-bit strings if $l_w < 2h$.
@@
 Let $S_s$ ($\overline{S_s}$) be the set of $m$-bit strings with the $h$ bits
-preceeding index $l_w$ all matching (not all matching) the bits of $\tilde{s}$,
+preceding index $l_w$ all matching (not all matching) the bits of $\tilde{s}$,
 where $\tilde{s}$ is a cyclic permutation of $s$:
@@
 
-Routes are contructed in 7 stages.
+Routes are constructed in 7 stages.
 The network topology dictates that $l^{(t+1)} = l^{(t)} + 1$ (mod $m$),
@@
 If $2h < l_w < m - h$,
-then there is no overlap between the indeces defining $R_s$ and $S_s$.
+then there is no overlap between the indices defining $R_s$ and $S_s$.
 In this case, the first $h$ bits of $z^{(t)}$ are set to
 match $z_w$.
-Otherwise there is some overlap between the indeces defining $R_s$ and
+Otherwise there is some overlap between the indices defining $R_s$ and
 $S_s$.
@@
 The $h$ bits of $z^{(t)}$
-preceeding index $l_w$ are set to match $z_w$.
+preceding index $l_w$ are set to match $z_w$.
 All nodes in this stage are within $h$ hops of $w$ and thus in its trusted
@@
 from the source $v$, destination $w$, and path parameter $s$,
-meaning that with this informaiton nodes are able to determine which neighbor
+meaning that with this information nodes are able to determine which neighbor
 to route a given message copy to.
@@
 butterfly can have a very low probability of failure when faced with
-adverarial faults, even from a very powerful attacker.
+adversarial faults, even from a very powerful attacker.
 
@@
 Many Internet services (e.g., email, social networks, cloud storage)
-are still highly centralized and vulnearble to technical and
+are still highly centralized and vulnerable to technical and
 non-technical (i.e. coercive) attacks.
@@
 as a fallback when surveillance isn't feasible.
-However, the technqiues presented in this paper
+However, the techniques presented in this paper
 are entirely compatible with, and in some cases could enhance, existing
@@
 in which link structure can be dictated,
-such as the butterfy.
+such as the butterfly.
 While the ability to control network topology is only achievable in some
@@
 Furthermore, when high levels of adversarial fault tolerance are required,
-new socio-technical systems might choose to incorporate such networks.
+new sociotechnical systems might choose to incorporate such networks.
 For example,
-a coallition of groups supporting free expression could use our results
+a coalition of groups supporting free expression could use our results
 to construct a censorship-resistant communication network
@@
 for a large increase in the number of messages compromised by an adversary.
-This stabilizing asymetry puts attackers at a disadvantage,
+This stabilizing asymmetry puts attackers at a disadvantage,
 disincentivizing attack.

### 2016-08-08 wsdm-2017/source/paper.tex
Remove pages from bib file.

@@
 \begin{theorem}
-Given an $m$-bit wraparound ($m \geq 0$) butterfly network,
+Given an $m$-bit wraparound butterfly network ($m \geq 0$),
 and a trust radius $h \geq 1$,

### 2016-10-21 www-2017/source/paper.tex
Add WWW17 version.

--- 2016-08-08 wsdm-2017/source/paper.tex
@@
 \documentclass{sig-alternate-05-2015}
-
-%\usepackage[ruled]{algorithm2e}
-%\renewcommand{\algorithmcfname}{ALGORITHM}
-%\usepackage{amsmath,amsfonts,amssymb,bbm} 
-%\usepackage[numbers,sort&compress]{natbib} % for citet
-%\SetArgSty{textrm}  % for algorithm2e
-%\SetAlFnt{\small}
-%\SetAlCapFnt{\small}
-%\SetAlCapNameFnt{\small}
-%\SetAlCapHSkip{0pt}
 
@@
 \usetikzlibrary{arrows,chains,decorations.pathreplacing}
-%\usepackage{draftwatermark}
-%\SetWatermarkText{DRAFT}
+
 \newcommand{\beq}{\begin{eqnarray}}
@@
 \newtheorem{theorem}{Theorem}
+
+%\usepackage{draftwatermark}
+%\SetWatermarkText{DRAFT}
 
@@
 %Conference
-\conferenceinfo{WSDM '17}{February 6--10, 2017, Cambridge, UK, USA}
+\conferenceinfo{WWW '17}{April 3--7, 2017, Perth, Australia}
 
@@
 
-\title{Attack-Tolerant Networks: A Structural, Multipath Approach}
+\title{Towards Attack-Tolerant Networks: Concurrent Multipath Routing and the Butterfly Network}
 %
@@
 
-\numberofauthors{1} %  in this sample file, there are a *total*
+\numberofauthors{2} %  in this sample file, there are a *total*
 % of EIGHT authors. SIX appear on the 'first-page' (for formatting
@@
 \alignauthor
-Anonymized Authors
+Edward L. Platt \\
+\affaddr{University of Michigan}\\
+\email{elplatt@umich.edu}
+\alignauthor
+Daniel M. Romero \\
+\affaddr{University of Michigan}\\
+\email{drom@umich.edu}
 }
@@
 Networks with single points of failure are particularly susceptible to
-adversarial faults, in which an attacker targets nodes strategically.
-In communication networks, these types of attacks leave users vulnerable to
-censorship and targeted surveillance.
+targeted attacks.
+In communication networks, these types of faults can leave users vulnerable to
+censorship and targeted surveillance,
+even when cryptography is utilized.
 Centralized networks have single points of failure by definition,
@@
 Existing network trust models, such as webs of trust, fail to adequately address
-network structure. We develop a partial trust model that incorporates network
-structure and which makes weaker, more realistic transitivity assumptions than
-webs of trust. We show that the partial trust model, when combined with
-concurrent multipath routing, allows standard fault tolerance techniques to be
-applied to communication across complex networks. We also provide a specific
-routing algorithm for implementing this scheme on the butterfly network
-topology.
+network structure.
+We describe a novel, adversarial fault-tolerant,
+concurrent multipath routing algorithm for the decentralized
+butterfly network topology.
+We also develop a partial trust model that makes it possible to
+quantify the adversarial fault-tolerance of a network,
+while also making more realistic transitivity assumptions than webs of trust.
 When network topology can be dictated, these results can be used to create
@@
 
-\keywords{networks; decentralization; fault tolerance; routing; multipath routing; adversarial faults; information security}
+\keywords{
+Butterfly Network;
+Fault Tolerance;
+Adversarial Faults;
+Multipath Routing;
+Censorship;
+Decentralization}
 

### 2016-10-21 www-2017/source/paper.tex
Update background and related.

@@
 \usepackage{tikz}
+\usepackage{soul}
+\usepackage[framemethod=tikz]{mdframed}
 \usetikzlibrary{arrows,chains,decorations.pathreplacing}
@@
 \newcommand{\eeq}{\end{eqnarray}}
+\newcommand{\bhlpar}{\begin{mdframed}[innerleftmargin=0,innerrightmargin=0,hidealllines=true,backgroundcolor=yellow!20]}
+\newcommand{\ehlpar}{\end{mdframed}}
 \DeclareMathOperator{\wbf}{wBF}
@@
 We also develop a partial trust model that makes it possible to
-quantify the adversarial fault-tolerance of a network,
+quantify the adversarial fault tolerance of a network,
 while also making more realistic transitivity assumptions than webs of trust.
@@
 \cite{zin_survey_2015,albert_error_2000,sterbenz_resilience_2010}.
-In fact, the decentralized design of the Internet was intended to protect
-against targeted attacks, such as nuclear strikes
+{\hl{
+In fact, the decentralized design of the Internet was motivated
+by the need to withstand nuclear strikes
+}}
 \cite{baran_distributed_1964}.
-Such targeted attacks are formally described as {\em adversarial faults}
-and are some of the most difficult to protect against.
+{\em Adversarial faults},
+{\hl{
+in which an adversary can strategically target attacks,
+are some of the most difficult to protect against.
+Such attacks are often used for censorship and targeted surveilance.
+}}
 Both theoretical results and recent events (described below)
@@
 In scale-free networks and other networks with heavy-tail degree distributions,
-random failures are highly likely to pick low-degree nodes, thus having
-little effect,
-while adversarial faults target the small number of high-degree nodes,
-thus removing a large number of edges with each fault.
+random failures are highly likely to
+{\hl{
+affect only low-degree nodes, thus having
+little effect.
+However, Adversarial faults target the few high-degree nodes,
+and therefore remove a large number of edges with each fault.
+}}
 So while the {\em protocols} of the Internet are decentralized,
@@
 
+\bhlpar
 We consider a setting in which a source node (Alice)
@@
 (Mal).
+Compromised nodes may behave incorrectly by blocking, altering,
+or incorrectly routing messages.
 We assume that Mal has full knowledge of the network structure, but has
 limited resources and thus can only compromise a fixed number of nodes.
-Compromised nodes may behave incorrectly by blocking the message,
-changing the message content, or routing the message incorrectly.
-In order to evaluate network-based techniques for mitigating such attacks,
-we develop a {\em partial trust} model,
-similar to the commonly used web of trust approach
-\cite{zimmermann_official_1995,ferguson_practical_2003},
-but making weaker (more realistic) assumptions of trust transitivity,
-and incorporating network structure.
-The partial trust model can be used to design and analyze communication
-schemes that tolerate adversarial faults,
-even when no single path through the network is fully trusted.
-
-Using the partial trust model, we are able to model the nodes in complex
-networks as being directly linked to each other by a number of
-redundant channels.
-This models enables the application of standard error detection techniques
+We also assume that nodes within a fixed network
+distance of either Alice or Bob are
+{\em trusted}
+and cannot be compromised.
+This assumption,
+which we call
+{\em partial trust transitivity},
+is a weaker (more realistic) form of the transitive
+trust assumption used by the popular 
+web of trust approach
+\cite{zimmermann_official_1995,ferguson_practical_2003}.
+
+Under the above assumptions,
+we show how to evaluate the influence of network structure
+on attack-tolerance,
+and propose a method for achieving a high-level of attack-tolerance on
+the butterfly network topology.
+The butterfly topology is popular in parallel processing
+\cite{kshemkalyani_distributed_2008} and
+peer-to-peer \cite{lua_survey_2005, korzun_structured_2013}
+applications, due to its regular structure and high connectivity.
+While it may seem questionable to dictate the topology of trust relationships,
+there are many applications where this occurs,
+e.g., overlay networks
+\cite{lua_survey_2005, korzun_structured_2013},
+formal organizations \cite{mohr_explaining_1982},
+government-regulated cellular networks \cite{walker_mass_2012},
+and call tree notification systems \cite{nickerson_thinking_2010}.
+Furthermore, we stress that trust is not static, it changes over time
+(ableit slowly).
+In this sense, we approach trust as deliberate infrastructure.
+Since centralized network structure is inherently
+vulnerable to attack, the ability to deliberately structure trust relationships
+is a necessary component of any attack-tolerance scheme.
+
+We first describe how fault tolerance techniques can be adapted and
+evaluated in a network setting with partial trust.
+Generally, faults in network paths can be correlated,
+preventing the appliation of standard fault tolerance techniques
 \cite{avizienis_basic_2004, von_neumann_probabilistic_1956},
-which rely on independent errors.
-While adversarial faults along a single path are correlated,
-faults in independent paths are not.
-Thus fault tolerance techniques can be applied,
-even in the presence of adversarial errors,
-when messages are sent in parallel,
+which assume independent faults.
+By using {\em independent paths},
+which have no untrusted nodes in common,
+we model communication across a complex network by a number of
+redundant and independent channels.
+Fault-tolerance schemes can then be applied when redundant messages
+are sent in parallel,
 a technique known as {\em concurrent multipath routing}
@@
 errors.
-Using the partial trust model, we formally evaluate the 
+We formally evaluate the
 effects of network structure on attack-resistance and show that the probability
@@
 
-Our specific multipath fault tolerance scheme relies on the existence
-and discoverability of multiple redundant paths between the source and
-the destination.
-Implementation thus requires a routing algorithm.
-While not a requirement, this technique is well-suited for
-{\em structured networks},
-in which links between nodes are constrained to a particular architecture.
-
-In particular, we provide a novel concurrent multipath routing algorithm
-on the butterfly network topology and use the partial trust model to prove
-a high level of adversarial fault tolerance.
-The butterfly topology is popular in parallel processing
-\cite{kshemkalyani_distributed_2008} and
-peer-to-peer \cite{lua_survey_2005, korzun_structured_2013}
-applications, due to its regular structure and high connectivity.
-While the ability to control network topology is only achievable in some
-applications, those applications are nonetheless significant,
-e.g. overlay networks \cite{lua_survey_2005, korzun_structured_2013},
-formal organizations \cite{mohr_explaining_1982},
-government-regulated cellular networks \cite{walker_mass_2012},
-and call tree notification systems \cite{nickerson_thinking_2010}.
+We also propose a novel concurrent multipath routing algorithm for the butterfly
+topology.
+The algorithm identifies independent paths,
+which when combined with the fault-tolerant concurrent multipath routing
+scheme above,
+achieves a high level of adversarial fault tolerance on the butterfly topology.
+\ehlpar
 
@@
 \begin{itemize}
-\item{We develop a decentralized {\em partial trust model},
-which makes weaker transitivity assumptions than web-of-trust models,
-and makes it possible to analyze the effect of network structure on
-adversarial fault tolerance;}
-\item{We show that the partial trust model enables standard fault
-tolerance techniques to be applied to adversarial faults in
+\item{
+\bhlpar
+We show that {\em independent paths} with {\em partial trust transitivity}
+enable standard fault tolerance techniques to be applied to adversarial faults in
 complex communication networks.
@@
 approaches 1 exponentially as the number of independent paths between
-sender and receiver increases;}
+sender and receiver increases;
+\ehlpar}
 \item{We present a scalable, efficient, and attack-tolerant concurrent
@@
 This paper is organized as follows.
-Section 2. reviews background and related work.
-Section 3. discusses the desired properties of attack-tolerant network
+Section \ref{sec-related}. reviews background and related work.
+Section \ref{sec-infrastructure}. discusses the desired properties of attack-tolerant network
 infrastructure.
-Section 4. describes the partial trust model.
-Section 5. describes a general technique for multipath fault tolerance on
+Section \ref{sec-ft}. describes adversarial fault tolerance on
 structured networks.
-Section 6. describes a routing algorithm for multipath fault tolerance
+Section \ref{sec-butterfly}. describes a routing algorithm for multipath fault tolerance
 on the butterfly network topology.
-Section 7. discusses the our results.
-And Section 8. concludes.
+Section \ref{sec-discussion}. discusses the our results.
+And Section \ref{sec-conclusion}. concludes.
 
 \section{Background and Related Work}
+\label{sec-related}
+
+There has been considerable work on trust in network security.
+Both centralized and decentralized approaches are commonly used to create
+trust infrastructures.
+Centralized approaches such as {\em public key infrastructure} (PKI)
+suffer from a number of vulnerabilities
+\cite{ellison_ten_2000},
+which stem largely from the single points of failure inherent to
+centralization.
+The well-known and widely-used {\em web of trust} approach
+\cite{zimmermann_official_1995,ferguson_practical_2003}
+is a decentralized alternative.
+In a web of trust,
+individuals choose who they trust initially.
+Trust is then extended to new individuals if they are vouched for by a
+currently-trusted individual,
+making it possible to extend their web of trust to a large
+number of nodes.
+However, the web of trust approach assumes that trust can be extended
+transitively,
+which is unrealistic
+\cite{christianson_why_1997}.
+\bhlpar
+Previous work on evaluating network structure has focused on authentication
+protocols, showing that independent paths can reduce an adversary's ability
+to impersonate a target
+\cite{levien_attack-resistant-2009}.
+Other work has shown that identifying independent paths in arbitrary networks
+is NP-hard and provided approximation algorithms
+\cite{reiter_resilient_1998}.
+Our work complements these by explicitly quantifying the effects of partial
+trust transitivity,
+and providing a routing algorithm to efficiently construct independent
+paths on a structured network.
+\ehlpar
 
@@
 Federated Byzantine Agreement (FBA) \cite{mazieres_stellar_2015}
+{\hl{
 is scalable, allows for flexible trust,
-and optimal fault tolerance for a given network (under its assumptions).
+and is highly fault-tolerant on networks meeting a set of requirements.
+}}
 However, FBA does not provide a method for evaluating the
@@
 or calculating how the failure probability for a particular network.
-The multipath fault tolerance scheme we propose is the
-first example we're aware of that allows adversarial fault tolerance
-to be quantified as a function of network structure.
 
@@
 protocol-based, cryptographic solutions
-\cite{ferguson_practical_2003}.
-A fully decentralized architecture must be decentralized both in protocol and
-in network structure.
+\cite{ferguson_practical_2003},
+without addressing single points of failure in-between
+(i.e., {\em man-in-the-middle attacks}).
 There are however, a few notable fault tolerance schemes
@@
 base station.
-The partial trust model presented in this paper provides a more general
-way to evaluate the effect of network structure on adversarial fault tolerance.
 The method of Liu et al.
 \cite{liu_secure_2012}
+{\hl{
 routes multiple messages first to random peers and then
-to their final destination.
+to a central base station,
+with the network edges constrained by sensors' physical location.
 The butterfly routing algorithm we present takes a conceptually similar
-approach,
-routing from the source to several intermediaries, and then to the destination.
+approach in a different network setting.
+}}
 
@@
 \section{Attack-Tolerant Infrastructure}
+\label{sec-infrastructure}
 
@@
 the attacker's resources are either more or less effective than the defender's.
+{\hl{
 We call the latter case stabilizing asymmetry,
-because it lowers the incentive to attack.
+because it reduces an attacker's power relative to their target.
+}}
 With this in mind, an attack-resistant infrastructure will benefit from a high
@@
 \section{Trust and Fault Tolerance}
-
-The requirement that networks remain sparse has important implications for
-resilience.
-As such networks grow, the number of nodes will exceed the number of links
-any single node can support.
-As a result, some pairs of nodes are only connected indirectly,
-through at least one intermediary node.
-These intermediary nodes are targets for potential attacks,
-making network paths potentially unreliable.
+\label{sec-ft}
 
@@
 are chosen strategically to maximize the likelihood of a failure.
-
-In communication networks, 
-the threat model is as follows: Alice sends a message to
-Bob via one or more intermediaries,
-but some intermediaries may block, change, or re-route the message.
-The ability to tolerate these types of adversarial attacks depends on
-Alice and Bob's ability to either find a path they {\em trust},
-or to communicate successfully across untrusted paths.
-
-Both centralized and decentralized approaches are commonly used to create
-trust infrastructures.
-Centralized approaches such as {\em public key infrastructure} (PKI)
-suffer from a number of vulnerabilities
-\cite{ellison_ten_2000},
-which stem largely from the single points of failure inherent to
-centralization.
-
-The well-known and widely-used {\em web of trust} model
-\cite{zimmermann_official_1995,ferguson_practical_2003}
-is a decentralized alternative.
-In a web of trust,
-individuals choose who they trust initially.
-Trust is then extended to new individuals if they are vouched for by a
-currently-trusted individual,
-making it possible to extend their web of trust to a large
-number of nodes.
-However, the web of trust approach assumes that trust can be extended
-transitively,
-which is unrealistic
-\cite{christianson_why_1997}.
-The web of trust model also fails to account for the structure
-of the trust network:
-trusted nodes are all trusted equally,
-regardless of how many nodes trust them.
-Similarly, while the process of building a web of trust is decentralized,
-the emergent structure of the trust network itself might be centralized.
-In order to analyze the influence of network structure on adversarial
-fault tolerance,
-a different trust model is needed.
 
@@
 \section{Multipath Butterfly Routing}
+\label{sec-butterfly}
 
@@
 For critical security applications in particular,
-the novel level of adversarial fault-tolerance offered by this scheme
+the novel level of adversarial fault tolerance offered by this scheme
 is highly desirable,
@@
 \section{Discussion}
+\label{sec-discussion}
 
@@
 \section{Conclusion}
+\label{sec-conclusion}
 

### 2016-10-24 www-2017/source/paper.tex
Add intro to trust/ft.

@@
 
+\bhlpar
+An important challenge in large-scale, secure communication is the ability to
+communicate reliably and securely with remote parties for whom no direct,
+trusted link exists.
+The commonly-used web of trust approach
+\cite{zimmermann_official_1995,ferguson_practical_2003}.
+extends trust infinitely transitively,
+to friends of friends, and friends of friends of friends, and so on.
+However, the assumption of infinitely transitive trust is unrealistic
+\cite{christianson_why_1997},
+and does not allow for the analysis of network structure.
+\ehlpar
+
 We can allow for more realistic transitivity assumptions and incorporate
@@
 by causing a single fault).
-We call paths that do not share any untrusted single points of failure
+Following \cite{reiter_resilient_1998},
+we call paths that do not share any untrusted single points of failure
 {\em independent}.
@@
 although the model can easily be extended to directed multigraphs.
-Vertices representing agents,
+Vertices representing commiunicating agents,
 and with edges representing mutually trusted communication links.

### 2016-10-24 www-2017/source/paper.tex
Update conclusion, remove highlights.

@@
 \cite{zin_survey_2015,albert_error_2000,sterbenz_resilience_2010}.
-{\hl{
 In fact, the decentralized design of the Internet was motivated
 by the need to withstand nuclear strikes
-}}
 \cite{baran_distributed_1964}.
 {\em Adversarial faults},
-{\hl{
 in which an adversary can strategically target attacks,
@@
 Such attacks are often used for censorship and targeted surveilance.
-}}
 Both theoretical results and recent events (described below)
@@
 random failures are highly likely to
-{\hl{
 affect only low-degree nodes, thus having
@@
 and therefore remove a large number of edges with each fault.
-}}
 So while the {\em protocols} of the Internet are decentralized,
@@
 
-\bhlpar
 We consider a setting in which a source node (Alice)
@@
 achieves a high level of adversarial fault tolerance on the butterfly topology.
-\ehlpar
 
@@
 \item{
-\bhlpar
 We show that {\em independent paths} with {\em partial trust transitivity}
@@
 sender and receiver increases;
-\ehlpar}
+}
 \item{We present a scalable, efficient, and attack-tolerant concurrent
@@
 \cite{christianson_why_1997}.
-\bhlpar
 Previous work on evaluating network structure has focused on authentication
@@
 to impersonate a target
-\cite{levien_attack-resistant-2009}.
+\cite{levien_attack-resistant_2009}.
 Other work has shown that identifying independent paths in arbitrary networks
@@
 paths on a structured network.
-\ehlpar
 
@@
 Federated Byzantine Agreement (FBA) \cite{mazieres_stellar_2015}
-{\hl{
 is scalable, allows for flexible trust,
 and is highly fault-tolerant on networks meeting a set of requirements.
-}}
 However, FBA does not provide a method for evaluating the
@@
 \cite{liu_secure_2012}
-{\hl{
 routes multiple messages first to random peers and then
@@
 approach in a different network setting.
-}}
 
@@
 the attacker's resources are either more or less effective than the defender's.
-{\hl{
 We call the latter case stabilizing asymmetry,
 because it reduces an attacker's power relative to their target.
-}}
 With this in mind, an attack-resistant infrastructure will benefit from a high
@@
 
-\bhlpar
 An important challenge in large-scale, secure communication is the ability to
@@
 and does not allow for the analysis of network structure.
-\ehlpar
 
@@
 is necessary.
-We have proposed a number of contributions towards this understanding:
-a {\em partial trust model} for quantifying the effective redundancy
-of a network,
-a {\em multipath fault tolerance} scheme for using that redundancy to detect
-errors,
-and a {\em concurrent multipath routing} algorithm enabling this scheme to
-be implemented on the butterfly network.
+We have proposed a network-based scheme for adversarial fault tolerance on the
+butterfly topology,
+utilizing a novel {\em concurrent multipath routing} algorithm.
+We have also demonstrated how {\em partial trust transitivity},
+in addition to being more realistic than infinite transitivity,
+enables the analysis of the relationship between trust, network structure,
+and fault tolerance.
 
@@
 non-technical (i.e. coercive) attacks.
-Fault tolerance using {\em both} decentralized protocols and
+Fault tolerance using {\em both}
+decentralized protocols and
 decentralized network structures
-is one approach to securing these important
-services and making networked communication safer.
-
-We have also focused specifically on adversarial faults that block or
+is one promising approach to securing these services.
+
+We have focused specifically on adversarial faults that block or
 change messages (e.g., censorship).
@@
 
-Our multipath fault tolerance scheme can be applied to any network
-that has both sufficient redundancy and a routing algorithm
-that maps local links to independent network paths.
-These requirements make the approach well-suited for structured networks,
-in which link structure can be dictated,
-such as the butterfly.
+While our present proposal is specific to the butterfly topology,
+the multipath fault tolerance scheme could
+be applied to any network that has both sufficient redundancy and a
+routing algorithm to discover independent paths.
+For general networks, finding all such paths is NP-hard,
+but efficient, suboptimal algorithms exist
+\cite{reiter_resilient_1998}.
+However, our approach is best suited for structured networks.
 While the ability to control network topology is only achievable in some
@@
 Furthermore, when high levels of adversarial fault tolerance are required,
-new sociotechnical systems might choose to incorporate such networks.
+designers of new sociotechnical systems might choose
+incorporate such networks.
 For example,
@@
 by cultivating the necessary trust relationships.
-In the general case,
-while the partial trust model can be applied to unstructured networks,
-additional work is needed to identify how to ensure redundancy and
-identify independent paths in unstructured networks.
+
+Our work suggests several directions for future work towards developing
+practical attack-tolerant communication infrastructure.
+The development of new multipath routing algorithms on other structured
+networks could achieve higher levels of redundancy.
+It is also desirable to identify dynamics that give
+rise to structured networks,
+and to evaluate whether our results can be generalized to unstructured or
+approximately structured networks.
+Finally, these results could be implemented to address specific
+applications, e.g., secure messaging, domain name resolution, or
+anonymous web browsing.
 
@@
 
-We have developed a partial trust model which enables a formal analysis
-of the effect of network structure and trust transitivity on
-redundancy.
-We have also proposed a multipath fault tolerance scheme
-for using this redundancy to achieve adversarial
-fault tolerance.
+We have presented a novel multipath routing algorithm for the butterfly
+topology,
+as well as a scheme for using this algorithm to achieve a high level of
+attack-tolerance in a communication network,
+even when no single path is entirely trusted.
 Under this scheme, the probability of an adversary causing an undetectable error
 decreases exponentially with the network's effective redundancy.
+The effective redundancy, in the case of the butterfly topology,
+grows exponentially with the trust radius.
 Furthermore, a small increase in the number of messages sent can compensate
 for a large increase in the number of messages compromised by an adversary.
-This stabilizing asymmetry puts attackers at a disadvantage,
-disincentivizing attack.
-We also presented a novel multipath routing algorithm allowing multipath fault
-tolerance to be used on the butterfly topology,
-and proved that the effective redundancy of the butterfly network increases
-exponentially with the trust radius.
-These results are directly applicable to architectures in which the link
-structure can be imposed by the architect,
+We have also demonstrated how the assumption of partial trust
+transitivity can enable a formal analysis of the effects of network structure
+on attack-tolerance.
+These results are directly applicable to systems in which the link
+structure can be imposed by the designer,
 and more generally, provide a formalism that can be used more to evaluate
@@
 \section{Acknowledgments}
-Anonymized for blind review.
 

### 2017-02-28 preprint/source/paper.tex
Add a bunch of stuff.

--- 2016-10-24 www-2017/source/paper.tex
@@
 
-%\usepackage{draftwatermark}
-%\SetWatermarkText{DRAFT}
-
 \begin{document}
@@
 % Copyright
-\setcopyright{acmcopyright}
+%\setcopyright{acmcopyright}
 %\setcopyright{acmlicensed}
@@
 % DOI
-%\doi{10.475/123_4}
+\doi{}
 
 % ISBN
-%\isbn{123-4567-24-567/08/06}
+\isbn{}
 
 %Conference
-\conferenceinfo{WWW '17}{April 3--7, 2017, Perth, Australia}
-
-\acmPrice{\$15.00}
+\conferenceinfo{Pre-print}{Currently under peer review.}
 
@@
 
-\title{Towards Attack-Tolerant Networks: Concurrent Multipath Routing and the Butterfly Network}
+\title{DRAFT: Towards Attack-Tolerant Networks: Concurrent Multipath Routing and the Butterfly Network}
 %
@@
 decentralized.
-While based on decentralized protocols, the Internet and World-Wide Web have
+Despite being based on decentralized protocols,
+the Internet and World-Wide Web have
 been shown both theoretically and historically to be highly susceptible to
-adversarial faults, in part due to emergent, structural centralization.
+adversarial faults, in part due to emergent structural centralization.
 Existing network trust models, such as webs of trust, fail to adequately address
@@
 quantify the adversarial fault tolerance of a network,
-while also making more realistic transitivity assumptions than webs of trust.
+and which makes more realistic transitivity assumptions than webs of trust.
 When network topology can be dictated, these results can be used to create
@@
 
+\vskip 3em
+
 \section{Introduction}
@@
 the Internet is surprisingly vulnerable to adversarial faults.
-Decentralization remains a promising approach to building resilient networks,
+Decentralization remains a promising approach towards
+building resilient networks,
 but further work is needed to understand the relationship between
 decentralized network structure and adversarial fault tolerance.
+
+The Internet's vulnerability to censorship and other targeted attacks
+has been confirmed by the success of several such attacks.
+For example, in 2008, YouTube suffered a worldwide outage for several hours
+when a service provider in Pakistan advertised false routing information
+\cite{hunter_pakistan_2008}.
+The action (known as a {\em black hole attack}) was intended to censor YouTube
+within Pakistan only, but resulted in a worldwide cascading failure.
+Such vulnerabilities are not limited to any one system or protocol,
+but result from the network structure itself.
 
@@
 
-The Internet's vulnerability to censorship and other targeted attacks
-has been confirmed by the success of several such attacks.
-For example, in 2008, YouTube suffered a worldwide outage for several hours
-when a service provider in Pakistan advertised false routing information
-\cite{hunter_pakistan_2008}.
-The action (known as a {\em black hole attack}) was intended to censor YouTube
-within Pakistan only, but resulted in a worldwide cascading failure.
-Such vulnerabilities are not limited to any one system or protocol,
-but result from the network structure itself.
 With strong theoretical and historical evidence that network structure
@@
 methods for analyzing structural vulnerabilities and for designing
-fault tolerant networks are needed.
+fault-tolerant networks are needed.
 This paper presents several contributions towards advancing those goals.
@@
 to a target node (Bob) by forwarding it through the links of the network.
+A ``link'' in this context could represent any kind of connection
+(e.g., physical cables, encrypted channels).
 We assume that some nodes in the network may be compromised by an attacker
@@
 and call tree notification systems \cite{nickerson_thinking_2010}.
-Furthermore, we stress that trust is not static, it changes over time
-(ableit slowly).
+Furthermore, we stress that trust is not static;
+it changes over time (ableit slowly).
 In this sense, we approach trust as deliberate infrastructure.
@@
 redundant and independent channels.
-Fault-tolerance schemes can then be applied when redundant messages
-are sent in parallel,
+Redundant messages can be sent across these channels in parallel,
 a technique known as {\em concurrent multipath routing}
-\cite{zin_survey_2015, qadir_exploiting_2015, khiani_comparative_2013}.
+\cite{zin_survey_2015, qadir_exploiting_2015, khiani_comparative_2013},
+and used for fault tolerance.
 The receiver can then use the redundant messages to detect and/or correct
@@
 We formally evaluate the
-effects of network structure on attack-resistance and show that the probability
-of undetected errors decreases exponentially with the number of
+effects of network structure on attack-tolerance and show that the probability
+of an undetected error decreases exponentially with the number of
 independent paths between source and destination,
@@
 The algorithm identifies independent paths,
-which when combined with the fault-tolerant concurrent multipath routing
-scheme above,
-achieves a high level of adversarial fault tolerance on the butterfly topology.
+which can combined with the fault-tolerant concurrent multipath routing
+scheme above to
+achieve a high level of adversarial fault tolerance on the butterfly topology.
 
@@
 structured networks.
-Section \ref{sec-butterfly}. describes a routing algorithm for multipath fault tolerance
-on the butterfly network topology.
+Section \ref{sec-butterfly}. describes the concurrent multipath routing
+algorithm for the butterfly network topology.
 Section \ref{sec-discussion}. discusses the our results.
@@
 currently-trusted individual,
-making it possible to extend their web of trust to a large
-number of nodes.
-However, the web of trust approach assumes that trust can be extended
-transitively,
+making it possible to quickly establish a large group of trusted nodes.
+However, the web of trust approach assumes that trust is infinitely
+transitive,
 which is unrealistic
 \cite{christianson_why_1997}.
-Previous work on evaluating network structure has focused on authentication
+
+Previous work on evaluating incorporating network structure into
+trust models has focused on authentication
 protocols, showing that independent paths can reduce an adversary's ability
@@
 Many distributed consensus protocols (such as those used by cryptocurrencies)
-are designed to provide tolerance against arbitrary or adversarial faults.
+are designed to tolerate arbitrary or adversarial faults.
 Byzantine agreement protocols
@@
 fault tolerance properties of different network structures
-or calculating how the failure probability for a particular network.
-
-Most existing adversarial fault tolerance schemes focus on end-to-end,
-protocol-based, cryptographic solutions
-\cite{ferguson_practical_2003},
-without addressing single points of failure in-between
-(i.e., {\em man-in-the-middle attacks}).
-There are however, a few notable fault tolerance schemes
-that focus on network structure.
+or for calculating the failure probabilities within a particular network.
+
+There are relatively few attack-tolerance schemes
+that focus on network structure,
+compared to more popular cryptographic approaches
+\cite{ferguson_practical_2003}.
 Fiat and Saia described a scheme that combines the butterfly topology
-with expander graphs to create a highly censorship-resistant content-addressable
-network \cite{fiat_censorship_2002},
+with expander graphs to create a highly censorship-resistant,
+content-addressable network \cite{fiat_censorship_2002},
 although this scheme exhibits poor scaling due to a very high level of data
@@
 Freenet guarantees that data is stored redundantly,
-but still allows for centralized network structure and thus single points
-of failure as data travels from its origin to the redundant storage locations.
+but still allows for centralized network structure,
+and thus single points of failure,
+as data travels from its origin to the redundant storage locations.
 
@@
 {\em structured network}, in which link structure is predetermined.
-Such networks can be designed to have favorable structural and routing
-properties,
-at the expense of complicating the addition or removal of nodes.
 Structured networks have been a popular tool in parallel processing
@@
 \cite{lua_survey_2005,korzun_structured_2013}.
+Such networks can be designed to have favorable structural and routing
+properties,
+which can be used to to guarantee attack-tolerance properties.
 
@@
 \item[Scalability:]
-For large scale networks it is important that the infrastructure allows
+For large-scale networks, it is important that the infrastructure allows
 for the network to grow while remaining functional.
@@
 In order to minimize single points of failure,
-attack tolerant infrastructures must use decentralized protocols and
+attack tolerant infrastructures must use {\em both} decentralized protocols and
 decentralized network structures.
@@
 In the context of international conflict,
-{\em asymmetric conflicts} are a special case that makes it possible for the
-less powerful party to have an advantage over the more powerful party
+{\em asymmetric conflicts} allow a
+less powerful party to have an advantage over a more powerful party
 \cite{mack_why_1975}.
@@
 links.
-Similarly, low-diameter guarantees that as a network grows, a short path will
+Similarly, a low diameter guarantees that as a network grows, a short path will
 still exist between any pair of nodes.
@@
 \item [Redundancy:]
-In a network, redundancy refers to the existence of multiple non-overlapping
+In a network, redundancy refers to the existence of multiple disjoint
 paths between nodes or components.
@@
 A {\em fault} occurs when one component
-of a system behaves incorrectly (e.g., a routing node blocking or
-altering a message).
+of a system behaves incorrectly (e.g., a routing node blocks or
+alters a message).
 The result of that fault (e.g., a recipient receiving conflicting messages)
@@
 We are concerned in particular with {\em adversarial faults},
-which (as opposed to random faults)
-are chosen strategically to maximize the likelihood of a failure.
+which are chosen strategically to maximize the likelihood of a failure.
 
@@
 The commonly-used web of trust approach
-\cite{zimmermann_official_1995,ferguson_practical_2003}.
-extends trust infinitely transitively,
+\cite{zimmermann_official_1995,ferguson_practical_2003}
+extends trust infinitely transitively:
 to friends of friends, and friends of friends of friends, and so on.
@@
 \cite{christianson_why_1997},
-and does not allow for the analysis of network structure.
+and does not allow for the analysis of the effects of network structure.
 
@@
 can be utilized by any redundancy-based fault tolerance scheme.
+
 We now propose a method for quantifying the effective redundancy of a network
@@
 this method allows a communication network
-to be modeled as a set of virtual links that directly
+to be modeled as a set of virtual channels that directly
 connect each node pair,
-with each virtual link providing some level of redundancy.
+with each virtual channel providing some level of redundancy.
 
@@
 full structure of the network,
-and who can compromise a fixed number of agents,
-gaining complete control of their behavior,
-as long as those agents are not trusted by either Alice or Bob.
-We define a {\em trust radius} $h$ and that nodes $u$ and
+and who can compromise a fixed number of nodes,
+gaining complete control of their behavior.
+We also assume that Mal is an adversary of Alice and/or Bob specifically,
+rather than the network as a whole.
+So Mal can compromise any node except for those trusted by Alice or Bob.
+We define a {\em trust radius} $h$ such that nodes $u$ and
 $u^\prime$ trust each other if their distance is less than $h$.
@@
 \beq
-T_h(u) &=& \left\{ u \mid d(u,u^\prime) < h \right\} \\
-B_h(u) &=& \left\{ u \mid d(u,u^\prime) = h \right\}.
+T_h(u) &=& \left\{ u^\prime \mid d(u,u^\prime) < h \right\} \\
+B_h(u) &=& \left\{ u^\prime \mid d(u,u^\prime) = h \right\}.
 \eeq
@@
 Having defined the assumptions of the partial trust model,
-we now quantify the effective redundancy between Alice and Bob
+we now quantify the {\em effective redundancy} between Alice and Bob
 when trust radius $h$ is assumed.
@@
 redundant paths can be constructed between them.
-While $\delta_h$ is a purely structural property of the graph and
-may have additional applications,
-we are interested in this quantity because
-it places an upper bound on the effectiveness of any
+$\delta_h$ is a purely structural network property,
+and places an upper bound on the effectiveness of any
 redundancy-based fault tolerance scheme.
-As a function of $h$, $\delta_h$ describes a network's ability to create
-redundancy from trust.
-Combined with a {\em routing algorithm} that tells nodes how to forward
-messages to guarantee they take independent paths,
-this redundancy can be used to achieve fault tolerance.
-For attack-tolerant networks,
-it is thus desirable for $\delta_h$ to increase quickly as a function of $h$.
-Furthermore, by applying the partial trust model,
-an adversarial fault tolerance scheme can treat each pair of nodes as being
-directly connected by a virtual link with least $\delta_h$ redundant
-channels.
-We will now describe such a scheme and evaluate its fault tolerance properties.
+The more quickly $\delta_h$ grows with $h$,
+the better a network is at creating redundancy from trust.
+Thus, the scaling of $\delta_h$ can be used to quantify a network's ability
+to withstand targeted attacks,
+even when the exact trust radius $h$ is unknown.
 
@@
 
-Let us first assume that Alice sends a message copy over each available
-channel.
+Let us first consider the case when
+Alice sends a message copy over each available channel.
 When Bob receives the messages, there are several possibilities.
@@
 In the present scenario,
-whether this occurs depends only on whether Mal has the resources to
+whether a failure occurs depends only on whether Mal has the resources to
 compromise all of the channels.
-In a more realistic, and more interesting, scenario,
+In a more realistic scenario,
 both Alice and Mal have limited resources and are not able to use or
@@
 depends only on the network structure
-and the strength of trust transitivity,
-not on the amount of resources available to the sender or the adversary.
-In other words, this scheme exhibits a {\em stabilizing asymmetry},
+and the strength of trust transitivity.
+Thus, the scheme can be effective, even when the number of copies
+sent $k$ is a small fraction of the effective redundancy.
+In other words, this scheme exhibits a {\em stabilizing asymmetry}:
 senders can tolerate attacks from significantly more powerful
-adversaries, as long as the network structure provides large $\delta_h$.
-By putting attackers at a disadvantage, attacks are not only tolerated
-but also disincentivized,
-further decreasing the likelihood of a successful attack.
+adversaries,
+as long as the network structure provides large $\delta_h$.
 
@@
 known as {\em the place-within-level}.
-There are two types of edges (shown in Figure \ref{fig:butterfly}).
+There are two types of edges: down, and down-right
+(shown in Figure \ref{fig:butterfly}).
 Down edges ($E_\downarrow$) connect nodes sharing the same $z$ value
@@
 We define three distinct partitions of $m$-bit binary strings.
-Let $Q_{v^{(0)}}$ ($\overline{Q_{v^{(0)}}}$) be the set of $m$-bit
+Let $Q_{v^{(0)}}$ be the set of $m$-bit
 strings in which the bits at
-all indices $h \leq i < l_w - h$ match (do not all match) those of $z^{(0)}$.
+all indices $h \leq i < l_w - h$ match those of $z^{(0)}$,
+and let $\overline{Q_{v^{(0)}}}$ be its complement.
 Note that $Q_{v^{(0)}}$ is trivially all $m$-bit strings if $l_w < 2h$.
-Let $R_s$ ($\overline{R_s}$) be the set of $m$-bit strings with the lowest $h$
-bits all matching (not all matching) the bits of $s$.
-Let $S_s$ ($\overline{S_s}$) be the set of $m$-bit strings with the $h$ bits
-preceding index $l_w$ all matching (not all matching) the bits of $\tilde{s}$,
+Let $R_s$ be the set of $m$-bit strings with the lowest $h$
+bits all matching the bits of $s$,
+and let $\overline{R_s}$ be its complement.
+Let $S_s$ be the set of $m$-bit strings with the $h$ bits
+preceding index $l_w$ all matching the bits of $\tilde{s}$,
 where $\tilde{s}$ is a cyclic permutation of $s$:
 \beq
-\tilde{s}_i &=& s_{(i + l_w) \text{ mod } h}.
+\tilde{s}_i &=& s_{(i + l_w) \text{ mod } h},
 \eeq
+and let $\overline{S_s}$ be its complement.
 We will make use of the fact that:
@@
 are still highly centralized and vulnerable to technical and
-non-technical (i.e. coercive) attacks.
+non-technical (i.e., coercive) attacks.
 Fault tolerance using {\em both}
@@
 confidentiality techniques.
-For example, the well-known {\em man-in-the-middle} attack exploits a privileged
+For example, {\em man-in-the-middle} attacks exploit a privileged
 network position to attack otherwise secure cryptography,
@@
 Furthermore, when high levels of adversarial fault tolerance are required,
-designers of new sociotechnical systems might choose
+designers of new sociotechnical systems might choose to
 incorporate such networks.
@@
 Our work suggests several directions for future work towards developing
-practical attack-tolerant communication infrastructure.
+practical, attack-tolerant communication infrastructure.
 The development of new multipath routing algorithms on other structured

### 2017-02-28 websci-2017/source/paper.tex
Add a bunch of stuff.

--- 2017-02-28 preprint/source/paper.tex
@@
 
+\usepackage{epigraph}
 \usepackage[utf8]{inputenc}
@@
 
+%\usepackage{draftwatermark}
+%\SetWatermarkText{DRAFT}
+
 \begin{document}
@@
 % Copyright
-%\setcopyright{acmcopyright}
+\setcopyright{acmcopyright}
 %\setcopyright{acmlicensed}
@@
 % DOI
-\doi{}
+%\doi{10.475/123_4}
 
 % ISBN
-\isbn{}
+%\isbn{123-4567-24-567/08/06}
 
 %Conference
-\conferenceinfo{Pre-print}{Currently under peer review.}
+\conferenceinfo{WWW '17}{April 3--7, 2017, Perth, Australia}
+
+\acmPrice{\$15.00}
 
@@
 
-\title{DRAFT: Towards Attack-Tolerant Networks: Concurrent Multipath Routing and the Butterfly Network}
+\title{Towards Attack-Tolerant Networks: Concurrent Multipath Routing and the Butterfly Network}
 %
@@
 
-Large-scale communication networks, exemplified by the Internet,
-have become ubiquitous, and are now crucial infrastructure for
-individuals, communities, and organizations around the world.
-As with any critical infrastructure, the cost of a failure can be
-immense, so methods for tolerating various kind of faults are an
+\epigraph{The Net interprets censorship as damage and routes around it.}{\textit{--John Gilmore}\cite{first_elmer-dewitt_1993}}
+
+Much of the world's infrastructure is networked: power grids,
+cellular networks, raods, and of course, the Internet.
+With any critical infrastructure, the cost of failures can be
+immense, so methods for tolerating various kind of faults within networks are an
 important and ongoing area of research
 \cite{zin_survey_2015,albert_error_2000,sterbenz_resilience_2010}.
-In fact, the decentralized design of the Internet was motivated
-by the need to withstand nuclear strikes
+{\em Adversarial faults},
+those in which an adversary can target attacks strategically,
+deserve special attention.
+Such attacks are both extremely difficult to guard against and 
+often have important social implications.
+In particular, censorship and surveillance are often achieved
+by targeting central network locations and either blocking or capturing
+the information flowing through them.
+The Internet's decentralized design was motivated
+by the need to withstand targeted attacks, such as nuclear strikes
 \cite{baran_distributed_1964}.
-{\em Adversarial faults},
-in which an adversary can strategically target attacks,
-are some of the most difficult to protect against.
-Such attacks are often used for censorship and targeted surveilance.
-Both theoretical results and recent events (described below)
+But despite longstanding common wisdom \cite{first_elmer-dewitt_1993},
+both theoretical results and recent events (described below)
 have demonstrated that
-the Internet is surprisingly vulnerable to adversarial faults.
+the Internet is surprisingly vulnerable to attack.
 Decentralization remains a promising approach towards
@@
 The Internet's vulnerability to censorship and other targeted attacks
-has been confirmed by the success of several such attacks.
-For example, in 2008, YouTube suffered a worldwide outage for several hours
+has been demonstrated by several recent events.
+In 2008, YouTube suffered a worldwide outage for several hours
 when a service provider in Pakistan advertised false routing information
@@
 within Pakistan only, but resulted in a worldwide cascading failure.
+The action was initiated by government order in Pakistan,
+and spread beyond Pakistan when a single misconfigured router allowed the false
+routing information to propagate globally.
+While the government order and router misconfiguration initiated the outage,
+it was the structure of the Internet's router network that allowed a fault in a
+single router to propagate globally.
+And while the action was not an intentional attack against the global Internet,
+the ability of an attacker to succeed without trying only strengthens our
+claim of the Intenrnet's vulnerability.
+
 Such vulnerabilities are not limited to any one system or protocol,

### 2017-02-28 websci-2017/source/paper.tex
Update text.

@@
 
-\epigraph{The Net interprets censorship as damage and routes around it.}{\textit{--John Gilmore}\cite{first_elmer-dewitt_1993}}
+\epigraph{The Net interprets censorship as damage and routes around it.}{\textit{--John Gilmore}\cite{elmer-dewitt_first_1993}}
 
@@
 \cite{baran_distributed_1964}.
-But despite longstanding common wisdom \cite{first_elmer-dewitt_1993},
+But despite longstanding common wisdom \cite{elmer-dewitt_first_1993},
 both theoretical results and recent events (described below)
-have demonstrated that
-the Internet is surprisingly vulnerable to attack.
+have demonstrated that the Internet can be surprisingly vulnerable to attack.
 Decentralization remains a promising approach towards
 building resilient networks,
-but further work is needed to understand the relationship between
+but there is a need to better understand the relationship between
 decentralized network structure and adversarial fault tolerance.
@@
 claim of the Intenrnet's vulnerability.
-
-Such vulnerabilities are not limited to any one system or protocol,
-but result from the network structure itself.
-
-Analysis of the Internet's router network has shown that while it is remarkably
-resilient against random faults,
+In 2013, the Texas-based email provider Lavabit was compelled by to turn
+their private SSL keys over to the FBI \cite{poulsen_edward_2013}.
+In order to protect their users from surveillance,
+Lavabit ceased operations.
+While originally intended as surveillance,
+this action effectively became an act of censorship.
+In this case, the attack was again successful due to a highly centralized
+architecture:
+SSL keys under control of a single entity, in a single legal jurisdiction.
+It is also important to note that there was no technological failure,
+the attack was achieved through coercion.
+So we see that such vulnerabilities are not limited to any one system or protocol,
+but result from centralized structure itself.
+
+Analysis of the Internet's router network has shown that while it
+is remarkably resilient against random faults,
 it is highly susceptible to adversarial faults \cite{albert_error_2000}.
@@
 
-With strong theoretical and historical evidence that network structure
-can create vulnerabilities,
+With strong theoretical and historical evidence that centralized
+network structure can create vulnerabilities,
 methods for analyzing structural vulnerabilities and for designing
@@
 (Mal).
+We also assume that Mal is an adversary of Alice specifically and targets
+nodes strategically with the goal of interfering with Alice's communications
+(rather than disrupting the network as a whole).
 Compromised nodes may behave incorrectly by blocking, altering,
@@
 limited resources and thus can only compromise a fixed number of nodes.
-We also assume that nodes within a fixed network
-distance of either Alice or Bob are
-{\em trusted}
-and cannot be compromised.
-This assumption,
-which we call
+
+We also assume that nodes {\em trust} their immediate neighbors,
+so Mal is unable to compromise Alice's node, or her direct neighbors.
+In the commonly used {\em web of trust} approach
+\cite{zimmermann_official_1995,ferguson_practical_2003},
+we would extend that trust transitively to the entire network.
+However, we make a weaker and more realistic assumption:
+that trust is extended transitively to nodes within a fixed
+distance.
+So only the beginning and end of a path between Alice and Bob is trusted
+against interference from Mal.
+We call this assumption
 {\em partial trust transitivity},
-is a weaker (more realistic) form of the transitive
-trust assumption used by the popular 
-web of trust approach
-\cite{zimmermann_official_1995,ferguson_practical_2003}.
+and refer to such paths as {\em partially trusted}.
 
@@
 on attack-tolerance,
+how to constructed trusted channels from redundant partially trusted paths,
 and propose a method for achieving a high-level of attack-tolerance on
@@
 peer-to-peer \cite{lua_survey_2005, korzun_structured_2013}
-applications, due to its regular structure and high connectivity.
-While it may seem questionable to dictate the topology of trust relationships,
-there are many applications where this occurs,
-e.g., overlay networks
+applications, due to its regular structure, low degree, and high connectivity.
+
+The reader may wonder whether it is realistic or useful to assume control over
+the network structure.
+We have already seen that whenever a single point of failure exists in the network,
+there is a potential to exploit it through coercion,
+without an attacker needing to compromise the technology.
+So, {\em attack-tolerance cannot be achieved without the ability to influence
+network structure}.
+Luckily, there are scenarios in which network topology can be dictated.
+Examples include overlay networks
 \cite{lua_survey_2005, korzun_structured_2013},
@@
 and call tree notification systems \cite{nickerson_thinking_2010}.
-Furthermore, we stress that trust is not static;
-it changes over time (ableit slowly).
-In this sense, we approach trust as deliberate infrastructure.
-Since centralized network structure is inherently
-vulnerable to attack, the ability to deliberately structure trust relationships
-is a necessary component of any attack-tolerance scheme.
+In general,
+{\em when the need for attack-tolerance is high enough to warrant investment
+in infrastructure, networks can be engineered and maintained as infrastructure}.
+It is also worth noting that attack-tolerant networks may be sub-components of
+larger, less-constrained systems.
+For example, a single server might be replaced by a distributed network of servers,
+each with different ownership, physical location, and legal jurisdiction,
+without placing any unrealistic constraints on the clients connecting to
+those servers.
 

### 2017-03-01 websci-2017/source/paper.tex
background

@@
 
-\epigraph{The Net interprets censorship as damage and routes around it.}{\textit{--John Gilmore}\cite{elmer-dewitt_first_1993}}
+\epigraph{The Net interprets censorship as damage and routes around it.}{\textit{--John Gilmore} \cite{elmer-dewitt_first_1993}}
 
@@
 cellular networks, raods, and of course, the Internet.
-With any critical infrastructure, the cost of failures can be
+As with any critical infrastructure, the cost of failures can be
 immense, so methods for tolerating various kind of faults within networks are an
@@
 And while the action was not an intentional attack against the global Internet,
-the ability of an attacker to succeed without trying only strengthens our
-claim of the Intenrnet's vulnerability.
+the ability of an attacker to succeed without even trying only highlights
+the Internet's vulnerability to adversarial faults.
+
 In 2013, the Texas-based email provider Lavabit was compelled by to turn
@@
 
-We also assume that nodes {\em trust} their immediate neighbors,
-so Mal is unable to compromise Alice's node, or her direct neighbors.
+We also assume that nodes {\em trust} their immediate neighbors.
+So Mal is unable to compromise Alice's node, or her direct neighbors.
 In the commonly used {\em web of trust} approach
@@
 that trust is extended transitively to nodes within a fixed
-distance.
+number of hops.
 So only the beginning and end of a path between Alice and Bob is trusted
@@
 We have already seen that whenever a single point of failure exists in the network,
-there is a potential to exploit it through coercion,
-without an attacker needing to compromise the technology.
+there is a potential for an attacker to exploit it through coercion,
+without needing to compromise the technology.
 So, {\em attack-tolerance cannot be achieved without the ability to influence
@@
 
-We first describe how fault tolerance techniques can be adapted and
-evaluated in a network setting with partial trust.
+We begin by describing how fault tolerance techniques can be adapted and
+evaluated in a network setting with partial trust transitivity and
+adversarial faults.
 Generally, faults in network paths can be correlated,
@@
 which assume independent faults.
-By using {\em independent paths},
+By constructing {\em independent paths},
 which have no untrusted nodes in common,
-we model communication across a complex network by a number of
-redundant and independent channels.
+we show how to model communication across a complex network in the presence
+of correlated adversarial faults as communication across redundant
+simple channels with random errors.
 Redundant messages can be sent across these channels in parallel,
@@
 topology.
-The algorithm identifies independent paths,
-which can combined with the fault-tolerant concurrent multipath routing
+The algorithm constructs independent paths,
+which can be combined with the fault-tolerant concurrent multipath routing
 scheme above to
@@
 \item{
-We show that {\em independent paths} with {\em partial trust transitivity}
-enable standard fault tolerance techniques to be applied to adversarial faults in
-complex communication networks.
-We find that the probability of detecting adversarial faults
-approaches 1 exponentially as the number of independent paths between
-sender and receiver increases;
+We propose a novel method for extending standard fault tolerance techniques to
+{\em adversarial} faults in {\em complex networks}.
+We do so by modelling redundant independent paths
+with partial trust transitivity as a single virtual channel,
+and show that the probability of detecting adversarial faults
+approaches 1 exponentially with the number of paths;
+}
+\item{We prove that the number of independent paths between two nodes
+in a wraparound butterfly network with partial trust transitivity
+increases exponentially with the trust radius;
 }
@@
 This paper is organized as follows.
-Section \ref{sec-related}. reviews background and related work.
-Section \ref{sec-infrastructure}. discusses the desired properties of attack-tolerant network
-infrastructure.
-Section \ref{sec-ft}. describes adversarial fault tolerance on
+Section \ref{sec-related} reviews background and related work.
+Section \ref{sec-ft} describes adversarial fault tolerance on
 structured networks.
-Section \ref{sec-butterfly}. describes the concurrent multipath routing
-algorithm for the butterfly network topology.
-Section \ref{sec-discussion}. discusses the our results.
-And Section \ref{sec-conclusion}. concludes.
+Section \ref{sec-butterfly} describes the concurrent multipath routing
+algorithm for the butterfly network topology
+Section \ref{sec-discussion} discusses the our results.
+And Section \ref{sec-conclusion} concludes.
 
@@
 making it possible to quickly establish a large group of trusted nodes.
-However, the web of trust approach assumes that trust is infinitely
-transitive,
-which is unrealistic
-\cite{christianson_why_1997}.
+However, web of trust's assumption of infinite transitivity is unrealistic
+\cite{christianson_why_1997},
+and does not distinguish between paths of different lengths.
+Our work addresses both of these limitations by incorporating a more realistic
+assumption of partial transitivity.
 
@@
 \cite{reiter_resilient_1998}.
-Our work complements these by explicitly quantifying the effects of partial
-trust transitivity,
-and providing a routing algorithm to efficiently construct independent
-paths on a structured network.
+Our work complements these by introducing the partial trust assumption and
+explicitly quantifying its relationship to network structure and adversarial
+fault tolerance.
+As an alternative to the NP-hard task of finding independent paths on arbitrary
+networks, we provide a routing algorithm to efficiently construct independent
+paths on a the butterfly topology.
 
@@
 \cite{ferguson_practical_2003}.
+All existing attack-tolerant networks we are aware of are content-addressable
+networks: data is routed to and from storage nodes rather than between sender
+and receiver.
 Fiat and Saia described a scheme that combines the butterfly topology
@@
 content-addressable network \cite{fiat_censorship_2002},
-although this scheme exhibits poor scaling due to a very high level of data
-replication.
+although this scheme does not scale well and is impractical due to a
+high level of data replication.
 Perhaps the most mature structural solution is the Freenet collaboration
@@
 as data travels from its origin to the redundant storage locations.
-
-{\em Multipath routing} protocols send redundant information
+Unlike the above content-addressable networks, our proposal is purely network based
+and does not require nodes to store data indefinitely.
+Our proposal also improves on the scalablity of Fiat and Saia's work,
+and does not rely on assumptions about existing social network structure.
+
+send redundant information
 over multiple paths when routing a message through a network,
+
+{\em Multipath routing} protocols identify multiple paths between
+source and destination
 in contrast to traditional {\em unipath} routing, which uses
 a single path.
-Multipath routing can have many benefits, including reduced congestion,
+The special case of {\em concurrent} multipath routing uses mutliple paths
+simultaneously.
+Multipath routing has many applications, including reduced congestion,
 increased throughput, and more reliability
@@
 with the network edges constrained by sensors' physical location.
-The butterfly routing algorithm we present takes a conceptually similar
-approach in a different network setting.
+We have found only few examples in the existing literature of applications of
+concurrent multipath routing to {\em adversarial} fault tolerance,
+and all have focused on ad-hoc wireless sensor networks, without attention
+to the role of network structure.
+The alogorithm we present for the butterfly topology complements existing work
+by addressing cases where links are not constrained by physical
+distance,
+and where network structure can be engineered for greater attack-tolerance.
 

### 2017-03-01 websci-2017/source/paper.tex
partial trust

@@
 
-In 2013, the Texas-based email provider Lavabit was compelled by to turn
-their private SSL keys over to the FBI \cite{poulsen_edward_2013}.
-In order to protect their users from surveillance,
-Lavabit ceased operations.
+In 2013, the Texas-based email provider Lavabit was ordered to disclose
+their private SSL keys to the FBI \cite{poulsen_edward_2013}.
+Rather than complying,
+Lavabit ceased operations
+in order to protect their users from surveillance.
 While originally intended as surveillance,
@@
 
-\section{Attack-Tolerant Infrastructure}
-\label{sec-infrastructure}
-
-In this section, we describe the functional properties required of an
-attack-tolerant network,
-which serve as guiding principles in later sections.
-We pay special attention to a property we call
-{\em stabilizing asymmetry}.
-
-\subsection{Functional Properties}
-
-\begin{description}
-\item[Scalability:]
-For large-scale networks, it is important that the infrastructure allows
-for the network to grow while remaining functional.
-In practice, people, devices, and connections, have limited capabilities
-and these limitations need to be considered as part of the design of the
-infrastructure.
-
-\item [Decentralization:]
-Systems having single points of failure are less tolerant against faults at
-those points.
-The existence of such points not only increases the likelihood that an attack
-will succeed,
-but also incentivizes attack by presenting effective targets.
-In order to minimize single points of failure,
-attack tolerant infrastructures must use {\em both} decentralized protocols and
-decentralized network structures.
-
-\item[Stabilizing asymmetry:]
-In the context of international conflict,
-{\em asymmetric conflicts} allow a
-less powerful party to have an advantage over a more powerful party
-\cite{mack_why_1975}.
-In asymmetric conflicts, the same level of resource expenditure yields different
-results for different parties;
-the attacker's resources are either more or less effective than the defender's.
-We call the latter case stabilizing asymmetry,
-because it reduces an attacker's power relative to their target.
-With this in mind, an attack-resistant infrastructure will benefit from a high
-level of stabilizing asymmetry.
-
-\end{description}
-\subsection{Structural Properties}
-
-We are specifically concerned with network structure-based approaches to
-fault tolerance.
-In networks, specific structural properties are required to achieve the
-functional properties described above.
-\begin{description}
-\item [Sparsity and low diameter:]
-To achieve scalability, networks must be {\em sparse} and have a
-{\em low diameter}.
-In practical settings, humans and devices have an upper limit on the number
-of connections they can maintain (e.g., Dunbar's number
-\cite{dunbar_neocortex_1992}).
-In sparse networks, the number of links grows slowly as the network grows in
-size, allowing the network to scale without exceeding the nodes' capacity for
-links.
-Similarly, a low diameter guarantees that as a network grows, a short path will
-still exist between any pair of nodes.
-While low diameter guarantees a path exists,
-paths are only useful if an efficient {\em routing} algorithm exists
-to find them.
-
-\item [Uniform centrality:]
-There are many ways to measure node centrality in networks
-\cite{freeman_centrality_1978}.
-The more uniform these measures are across nodes, the more decentralized
-a network is.
-Centrality is minimized in {\em vertex transitive} networks,
-for which an edge-preserving map
-always exists from any node to any other node.
-In other words, all nodes occupy structurally indistinguishable positions
-in the network.
-
-\item [Redundancy:]
-In a network, redundancy refers to the existence of multiple disjoint
-paths between nodes or components.
-Redundancy can help reduce single points of failure and decrease
-centralization.
-One measure of redundancy is given by the ratio of edges to nodes
-\cite{baran_distributed_1964}.
-A single point of failure occurs when a node holds a uniquely
-central position.
-When alternative paths are added to bypass central nodes,
-the network becomes more redundant.
-
-\end{description}
-
-\section{Trust and Fault Tolerance}
+\section{Trust Networks and Fault Tolerance}
 \label{sec-ft}
@@
 We will make use of standard fault tolerance terminology, summarized here.
-A {\em fault} occurs when one component
+A {\em fault} is said to occur when one component
 of a system behaves incorrectly (e.g., a routing node blocks or
@@
 The result of that fault (e.g., a recipient receiving conflicting messages)
-is an {\em error} state.
+is called an {\em error} state.
 If the error is undetected or corrected to the wrong value, the system is
@@
 accepted as authentic).
+Note that when an error is detected but cannot be corrected,
+the system has still tolerated the fault because it has not accepted an error
+state.
 We are concerned in particular with {\em adversarial faults},
@@
 
-An important challenge in large-scale, secure communication is the ability to
-communicate reliably and securely with remote parties for whom no direct,
-trusted link exists.
+A central question in large-scale, secure communication is this:
+how can two parties communicate reliably and securely
+when no direct trusted link exists between them?
 The commonly-used web of trust approach
@@
 
-We can allow for more realistic transitivity assumptions and incorporate
-network structure into a trust model with one key insight:
-even if no single path between a sender and receiver is fully trusted,
-multiple copies of a message can be sent along different paths and compared
-by the receiver to detect and correct errors.
-However, in the presence of an attacker, there is only benefit in sending an
-additional message along a path if that path does not share untrusted
-single points of failure with any of the existing message paths
-(otherwise, the adversary can compromise both messages
-by causing a single fault).
-Following \cite{reiter_resilient_1998},
-we call paths that do not share any untrusted single points of failure
-{\em independent}.
-The maximum number of independent paths represents the effective redundancy that
-can be utilized by any redundancy-based fault tolerance scheme.
-
-We now propose a method for quantifying the effective redundancy of a network
-as a function of its structure and level of trust transitivity.
-For the purpose of analyzing adversarial fault tolerance,
-this method allows a communication network
-to be modeled as a set of virtual channels that directly
-connect each node pair,
-with each virtual channel providing some level of redundancy.
-
-We now specify the {\em partial trust model}.
-We assume an undirected graph $G = (V,E)$,
+An alternative assumption might be that each hop away from Alice in
+in the network,
+reduces the probability that a node can resist compromise exponentially.
+Such a situation might occur because nodes more distant from Alice may be
+more favorably disposed to Alice's adversary, more likely to cooperate with that
+adversary, and less likely to take proactive security measures against that
+adversary.
+The above model can be further
+approximated by assuming that nodes up to some fixed number
+of hops cannot be compromised, and those beyond can.
+This simplified version is still more realistic than infinite transitivity
+and will be convenient for proving our results.
+We now proceed to define it formally.
+
+We define the {\em partial trust model} on
+an undirected graph $G = (V,E)$,
 although the model can easily be extended to directed multigraphs.
@@
 We refer to this set of nodes as the {\em untrusted region}.
+
+We can allow for more realistic transitivity assumptions and incorporate
+network structure into a trust model with one key insight:
+even if no single path between a sender and receiver is fully trusted,
+multiple copies of a message can be sent along different paths and compared
+by the receiver to detect and correct errors.
+However, in the presence of an attacker, there is only benefit in sending an
+additional message along a path if that path does not share untrusted
+single points of failure with any of the existing message paths
+(otherwise, the adversary can compromise both messages
+by causing a single fault).
+Following \cite{reiter_resilient_1998},
+we call paths that do not share any untrusted single points of failure
+{\em independent}.
+The maximum number of independent paths represents the effective redundancy that
+can be utilized by any redundancy-based fault tolerance scheme.
+
+We now propose a method for quantifying the effective redundancy of a network
+as a function of its structure and level of trust transitivity.
+For the purpose of analyzing adversarial fault tolerance,
+this method allows a communication network
+to be modeled as a set of virtual channels that directly
+connect each node pair,
+with each virtual channel providing some level of redundancy.
 

### 2017-03-01 websci-2017/source/paper.tex
more updates.

@@
 An alternative assumption might be that each hop away from Alice in
-in the network,
+in the network
 reduces the probability that a node can resist compromise exponentially.
-Such a situation might occur because nodes more distant from Alice may be
+Such a situation could occur if nodes more distant from Alice are
 more favorably disposed to Alice's adversary, more likely to cooperate with that
@@
 approximated by assuming that nodes up to some fixed number
-of hops cannot be compromised, and those beyond can.
+of hops cannot be compromised, and that those beyond can.
 This simplified version is still more realistic than infinite transitivity
@@
 
-We can allow for more realistic transitivity assumptions and incorporate
-network structure into a trust model with one key insight:
-even if no single path between a sender and receiver is fully trusted,
-multiple copies of a message can be sent along different paths and compared
-by the receiver to detect and correct errors.
-However, in the presence of an attacker, there is only benefit in sending an
-additional message along a path if that path does not share untrusted
-single points of failure with any of the existing message paths
-(otherwise, the adversary can compromise both messages
-by causing a single fault).
-Following \cite{reiter_resilient_1998},
-we call paths that do not share any untrusted single points of failure
-{\em independent}.
-The maximum number of independent paths represents the effective redundancy that
-can be utilized by any redundancy-based fault tolerance scheme.
+\subsection{Effective Redundancy}
 
@@
 For the purpose of analyzing adversarial fault tolerance,
-this method allows a communication network
-to be modeled as a set of virtual channels that directly
-connect each node pair,
-with each virtual channel providing some level of redundancy.
+this method allows us to model each pair of nodes
+in a complex communication network
+as being directly connected by a dedicated {\em virtual channel}
+that provides some level of redundancy.
 

### 2017-03-01 websci-2017/source/properties.tex
more updates.

--- 2017-02-28 preprint/source/paper.tex
@@
-\documentclass{sig-alternate-05-2015}
-
-\usepackage[utf8]{inputenc}
-\usepackage{afterpage}
-\usepackage{tikz}
-\usepackage{soul}
-\usepackage[framemethod=tikz]{mdframed}
-\usetikzlibrary{arrows,chains,decorations.pathreplacing}
-
-\newcommand{\beq}{\begin{eqnarray}}
-\newcommand{\eeq}{\end{eqnarray}}
-\newcommand{\bhlpar}{\begin{mdframed}[innerleftmargin=0,innerrightmargin=0,hidealllines=true,backgroundcolor=yellow!20]}
-\newcommand{\ehlpar}{\end{mdframed}}
-\DeclareMathOperator{\wbf}{wBF}
-\newtheorem{theorem}{Theorem}
-
-\begin{document}
-
-% Copyright
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\setcopyright{usgov}
-%\setcopyright{usgovmixed}
-%\setcopyright{cagov}
-%\setcopyright{cagovmixed}
-
-% DOI
-\doi{}
-
-% ISBN
-\isbn{}
-
-%Conference
-\conferenceinfo{Pre-print}{Currently under peer review.}
-
-%
-% --- Author Metadata here ---
-%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
-%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
-%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
-% --- End of Author Metadata ---
-
-\title{DRAFT: Towards Attack-Tolerant Networks: Concurrent Multipath Routing and the Butterfly Network}
-%
-% You need the command \numberofauthors to handle the 'placement
-% and alignment' of the authors beneath the title.
-%
-% For aesthetic reasons, we recommend 'three authors at a time'
-% i.e. three 'name/affiliation blocks' be placed beneath the title.
-%
-% NOTE: You are NOT restricted in how many 'rows' of
-% "name/affiliations" may appear. We just ask that you restrict
-% the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate'
-% we ask you to refrain from putting more than six authors
-% (two rows with three columns) beneath the article title.
-% More than six makes the first-page appear very cluttered indeed.
-%
-% Use the \alignauthor commands to handle the names
-% and affiliations for an 'aesthetic maximum' of six authors.
-% Add names, affiliations, addresses for
-% the seventh etc. author(s) as the argument for the
-% \additionalauthors command.
-% These 'additional authors' will be output/set for you
-% without further effort on your part as the last section in
-% the body of your article BEFORE References or any Appendices.
-
-\numberofauthors{2} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
-%
-\author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
-\alignauthor
-Edward L. Platt \\
-\affaddr{University of Michigan}\\
-\email{elplatt@umich.edu}
-\alignauthor
-Daniel M. Romero \\
-\affaddr{University of Michigan}\\
-\email{drom@umich.edu}
-}
-%       \affaddr{Institute for Clarity in Documentation}\\
-%       \affaddr{1932 Wallamaloo Lane}\\
-%       \affaddr{Wallamaloo, New Zealand}\\
-%       \email{trovato@corporation.com}
-% 2nd. author
-%\alignauthor
-%G.K.M. Tobin\titlenote{The secretary disavows
-%any knowledge of this author's actions.}\\
-%       \affaddr{Institute for Clarity in Documentation}\\
-%       \affaddr{P.O. Box 1212}\\
-%       \affaddr{Dublin, Ohio 43017-6221}\\
-%       \email{webmaster@marysville-ohio.com}
-% 3rd. author
-
-\maketitle
-\begin{abstract}
-Networks with single points of failure are particularly susceptible to
-targeted attacks.
-In communication networks, these types of faults can leave users vulnerable to
-censorship and targeted surveillance,
-even when cryptography is utilized.
-Centralized networks have single points of failure by definition,
-leading to a growing popularity in decentralized architectures and protocols.
-However, centralized network structure can arise even when protocols are
-decentralized.
-Despite being based on decentralized protocols,
-the Internet and World-Wide Web have
-been shown both theoretically and historically to be highly susceptible to
-adversarial faults, in part due to emergent structural centralization.
-Existing network trust models, such as webs of trust, fail to adequately address
-network structure.
-We describe a novel, adversarial fault-tolerant,
-concurrent multipath routing algorithm for the decentralized
-butterfly network topology.
-We also develop a partial trust model that makes it possible to
-quantify the adversarial fault tolerance of a network,
-and which makes more realistic transitivity assumptions than webs of trust.
-When network topology can be dictated, these results can be used to create
-scalable, attack-tolerant infrastructures. More generally, our results provide
-a formalism for evaluating the effects of network structure on adversarial
-fault tolerance.
-\end{abstract}
-
-%
-% The code below should be generated by the tool at
-% http://dl.acm.org/ccs.cfm
-% Please copy and paste the code instead of the example below. 
-%
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10002978.10002986</concept_id>
-<concept_desc>Security and privacy~Formal methods and theory of security</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003033.10003034</concept_id>
-<concept_desc>Networks~Network architectures</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10011007.10010940.10010971.10010972.10010540</concept_id>
-<concept_desc>Software and its engineering~Peer-to-peer architectures</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Security and privacy~Formal methods and theory of security}
-\ccsdesc[500]{Networks~Network architectures}
-\ccsdesc[500]{Software and its engineering~Peer-to-peer architectures}
-
-%
-% End generated code
-%
-
-%
-%  Use this command to print the description
-%
-\printccsdesc
-
-% We no longer use \terms command
-%\terms{Theory}
-
-\keywords{
-Butterfly Network;
-Fault Tolerance;
-Adversarial Faults;
-Multipath Routing;
-Censorship;
-Decentralization}
-
-\vskip 3em
-
-\section{Introduction}
-
-Large-scale communication networks, exemplified by the Internet,
-have become ubiquitous, and are now crucial infrastructure for
-individuals, communities, and organizations around the world.
-As with any critical infrastructure, the cost of a failure can be
-immense, so methods for tolerating various kind of faults are an
-important and ongoing area of research
-\cite{zin_survey_2015,albert_error_2000,sterbenz_resilience_2010}.
-In fact, the decentralized design of the Internet was motivated
-by the need to withstand nuclear strikes
-\cite{baran_distributed_1964}.
-{\em Adversarial faults},
-in which an adversary can strategically target attacks,
-are some of the most difficult to protect against.
-Such attacks are often used for censorship and targeted surveilance.
-Both theoretical results and recent events (described below)
-have demonstrated that
-the Internet is surprisingly vulnerable to adversarial faults.
-Decentralization remains a promising approach towards
-building resilient networks,
-but further work is needed to understand the relationship between
-decentralized network structure and adversarial fault tolerance.
-
-The Internet's vulnerability to censorship and other targeted attacks
-has been confirmed by the success of several such attacks.
-For example, in 2008, YouTube suffered a worldwide outage for several hours
-when a service provider in Pakistan advertised false routing information
-\cite{hunter_pakistan_2008}.
-The action (known as a {\em black hole attack}) was intended to censor YouTube
-within Pakistan only, but resulted in a worldwide cascading failure.
-Such vulnerabilities are not limited to any one system or protocol,
-but result from the network structure itself.
-
-Analysis of the Internet's router network has shown that while it is remarkably
-resilient against random faults,
-it is highly susceptible to adversarial faults \cite{albert_error_2000}.
-These results have been attributed to the scale-free structure of the Internet's
-router network
-\cite{barabasi_emergence_1999,barabasi_scale-free_2009}.
-In scale-free networks and other networks with heavy-tail degree distributions,
-random failures are highly likely to
-affect only low-degree nodes, thus having
-little effect.
-However, Adversarial faults target the few high-degree nodes,
-and therefore remove a large number of edges with each fault.
-So while the {\em protocols} of the Internet are decentralized,
-the {\em network structure} is somewhat centralized. 
-In other words, the protocols of the Internet do not {\em require}
-centralization, but centralization may still emerge from the sociotechnical
-processes that create its network structure.
-
-With strong theoretical and historical evidence that network structure
-can create vulnerabilities,
-methods for analyzing structural vulnerabilities and for designing
-fault-tolerant networks are needed.
-This paper presents several contributions towards advancing those goals.
-
-We consider a setting in which a source node (Alice)
-in a network attempts to route a message
-to a target node (Bob) by forwarding it through the links of the network.
-A ``link'' in this context could represent any kind of connection
-(e.g., physical cables, encrypted channels).
-We assume that some nodes in the network may be compromised by an attacker
-(Mal).
-Compromised nodes may behave incorrectly by blocking, altering,
-or incorrectly routing messages.
-We assume that Mal has full knowledge of the network structure, but has
-limited resources and thus can only compromise a fixed number of nodes.
-We also assume that nodes within a fixed network
-distance of either Alice or Bob are
-{\em trusted}
-and cannot be compromised.
-This assumption,
-which we call
-{\em partial trust transitivity},
-is a weaker (more realistic) form of the transitive
-trust assumption used by the popular 
-web of trust approach
-\cite{zimmermann_official_1995,ferguson_practical_2003}.
-
-Under the above assumptions,
-we show how to evaluate the influence of network structure
-on attack-tolerance,
-and propose a method for achieving a high-level of attack-tolerance on
-the butterfly network topology.
-The butterfly topology is popular in parallel processing
-\cite{kshemkalyani_distributed_2008} and
-peer-to-peer \cite{lua_survey_2005, korzun_structured_2013}
-applications, due to its regular structure and high connectivity.
-While it may seem questionable to dictate the topology of trust relationships,
-there are many applications where this occurs,
-e.g., overlay networks
-\cite{lua_survey_2005, korzun_structured_2013},
-formal organizations \cite{mohr_explaining_1982},
-government-regulated cellular networks \cite{walker_mass_2012},
-and call tree notification systems \cite{nickerson_thinking_2010}.
-Furthermore, we stress that trust is not static;
-it changes over time (ableit slowly).
-In this sense, we approach trust as deliberate infrastructure.
-Since centralized network structure is inherently
-vulnerable to attack, the ability to deliberately structure trust relationships
-is a necessary component of any attack-tolerance scheme.
-
-We first describe how fault tolerance techniques can be adapted and
-evaluated in a network setting with partial trust.
-Generally, faults in network paths can be correlated,
-preventing the appliation of standard fault tolerance techniques
-\cite{avizienis_basic_2004, von_neumann_probabilistic_1956},
-which assume independent faults.
-By using {\em independent paths},
-which have no untrusted nodes in common,
-we model communication across a complex network by a number of
-redundant and independent channels.
-Redundant messages can be sent across these channels in parallel,
-a technique known as {\em concurrent multipath routing}
-\cite{zin_survey_2015, qadir_exploiting_2015, khiani_comparative_2013},
-and used for fault tolerance.
-The receiver can then use the redundant messages to detect and/or correct
-errors.
-We formally evaluate the
-effects of network structure on attack-tolerance and show that the probability
-of an undetected error decreases exponentially with the number of
-independent paths between source and destination,
-even when no individual path is entirely trusted.
-
-We also propose a novel concurrent multipath routing algorithm for the butterfly
-topology.
-The algorithm identifies independent paths,
-which can combined with the fault-tolerant concurrent multipath routing
-scheme above to
-achieve a high level of adversarial fault tolerance on the butterfly topology.
-
-Our main contributions are:
-\begin{itemize}
-\item{
-We show that {\em independent paths} with {\em partial trust transitivity}
-enable standard fault tolerance techniques to be applied to adversarial faults in
-complex communication networks.
-We find that the probability of detecting adversarial faults
-approaches 1 exponentially as the number of independent paths between
-sender and receiver increases;
-}
-\item{We present a scalable, efficient, and attack-tolerant concurrent
-multipath routing algorithm on the butterfly network topology.}
-\end{itemize}
-
-This paper is organized as follows.
-Section \ref{sec-related}. reviews background and related work.
-Section \ref{sec-infrastructure}. discusses the desired properties of attack-tolerant network
-infrastructure.
-Section \ref{sec-ft}. describes adversarial fault tolerance on
-structured networks.
-Section \ref{sec-butterfly}. describes the concurrent multipath routing
-algorithm for the butterfly network topology.
-Section \ref{sec-discussion}. discusses the our results.
-And Section \ref{sec-conclusion}. concludes.
-
-\section{Background and Related Work}
-\label{sec-related}
-
-There has been considerable work on trust in network security.
-Both centralized and decentralized approaches are commonly used to create
-trust infrastructures.
-Centralized approaches such as {\em public key infrastructure} (PKI)
-suffer from a number of vulnerabilities
-\cite{ellison_ten_2000},
-which stem largely from the single points of failure inherent to
-centralization.
-The well-known and widely-used {\em web of trust} approach
-\cite{zimmermann_official_1995,ferguson_practical_2003}
-is a decentralized alternative.
-In a web of trust,
-individuals choose who they trust initially.
-Trust is then extended to new individuals if they are vouched for by a
-currently-trusted individual,
-making it possible to quickly establish a large group of trusted nodes.
-However, the web of trust approach assumes that trust is infinitely
-transitive,
-which is unrealistic
-\cite{christianson_why_1997}.
-
-Previous work on evaluating incorporating network structure into
-trust models has focused on authentication
-protocols, showing that independent paths can reduce an adversary's ability
-to impersonate a target
-\cite{levien_attack-resistant_2009}.
-Other work has shown that identifying independent paths in arbitrary networks
-is NP-hard and provided approximation algorithms
-\cite{reiter_resilient_1998}.
-Our work complements these by explicitly quantifying the effects of partial
-trust transitivity,
-and providing a routing algorithm to efficiently construct independent
-paths on a structured network.
-
-Many distributed consensus protocols (such as those used by cryptocurrencies)
-are designed to tolerate arbitrary or adversarial faults.
-Byzantine agreement protocols
-\cite{lamport_byzantine_1982,castro_practical_1999}
-provide tolerance against arbitrary faults (including attacks) under
-some circumstances, but are limited to small networks due to poor scalability.
-Proof-of-work \cite{dwork_pricing_1993,nakamoto_bitcoin:_2008}
-and proof-of-stake \cite{king_ppcoin:_2012}
-provide better scalability,
-but are wasteful of computational and energy resources.
-Federated Byzantine Agreement (FBA) \cite{mazieres_stellar_2015}
-is scalable, allows for flexible trust,
-and is highly fault-tolerant on networks meeting a set of requirements.
-However, FBA does not provide a method for evaluating the
-fault tolerance properties of different network structures
-or for calculating the failure probabilities within a particular network.
-
-There are relatively few attack-tolerance schemes
-that focus on network structure,
-compared to more popular cryptographic approaches
-\cite{ferguson_practical_2003}.
-Fiat and Saia described a scheme that combines the butterfly topology
-with expander graphs to create a highly censorship-resistant,
-content-addressable network \cite{fiat_censorship_2002},
-although this scheme exhibits poor scaling due to a very high level of data
-replication.
-Perhaps the most mature structural solution is the Freenet collaboration
-\cite{clarke_freenet:_2001}.
-Freenet uses secret sharing
-\cite{shamir_how_1979, blakley_safeguarding_1979}
-and small-world routing
-\cite{zhang_using_2002,kleinberg_small-world_2000}
-to create a content-addressable network with a high level of both
-confidentiality and censorship resistance.
-Freenet guarantees that data is stored redundantly,
-but still allows for centralized network structure,
-and thus single points of failure,
-as data travels from its origin to the redundant storage locations.
-
-{\em Multipath routing} protocols send redundant information
-over multiple paths when routing a message through a network,
-in contrast to traditional {\em unipath} routing, which uses
-a single path.
-Multipath routing can have many benefits, including reduced congestion,
-increased throughput, and more reliability
-\cite{qadir_exploiting_2015}.
-Many of these routing protocols offer increased confidentiality
-\cite{zin_survey_2015}.
-Some approaches utilize redundant paths as backups for increased
-fault tolerance
-\cite{alrajeh_secure_2013},
-and some specifically protect against adversarial faults
-\cite{kohno_improvement_2012, khalil_unmask:_2010, lou_h-spread:_2006}.
-Most work on multipath routing has been motivated by applications related to
-wireless sensor networks (WSNs),
-and have thus focused on ad hoc, unstructured networks, often having a central
-base station.
-The method of Liu et al.
-\cite{liu_secure_2012}
-routes multiple messages first to random peers and then
-to a central base station,
-with the network edges constrained by sensors' physical location.
-The butterfly routing algorithm we present takes a conceptually similar
-approach in a different network setting.
-
-Our proposed routing algorithm makes use of a
-{\em structured network}, in which link structure is predetermined.
-Structured networks have been a popular tool in parallel processing
-architectures \cite{kshemkalyani_distributed_2008}.
-More recently, peer-to-peer systems based on distributed hash tables have used
-structured ``overlay'' networks to map table keys to local TCP/IP routes
-\cite{lua_survey_2005,korzun_structured_2013}.
-Such networks can be designed to have favorable structural and routing
-properties,
-which can be used to to guarantee attack-tolerance properties.
-
-\section{Attack-Tolerant Infrastructure}
-\label{sec-infrastructure}
-
-In this section, we describe the functional properties required of an
-attack-tolerant network,
-which serve as guiding principles in later sections.
-We pay special attention to a property we call
-{\em stabilizing asymmetry}.
+We can allow for more realistic transitivity assumptions and incorporate
+network structure into a trust model with one key insight:
+even if no single path between a sender and receiver is fully trusted,
+multiple copies of a message can be sent along different paths and compared
+by the receiver to detect and correct errors.
+However, in the presence of an attacker, there is only benefit in sending an
+additional message along a path if that path does not share untrusted
+single points of failure with any of the existing message paths
+(otherwise, the adversary can compromise both messages
+by causing a single fault).
+Following \cite{reiter_resilient_1998},
+we call paths that do not share any untrusted single points of failure
+{\em independent}.
+The maximum number of independent paths represents the effective redundancy that
+can be utilized by any redundancy-based fault tolerance scheme.
 
@@
 \end{description}
-
-\section{Trust and Fault Tolerance}
-\label{sec-ft}
-
-Within the field of {\em fault tolerance},
-many techniques have been developed for building reliable systems
-out of unreliable components
-\cite{avizienis_basic_2004, von_neumann_probabilistic_1956}.
-We will make use of standard fault tolerance terminology, summarized here.
-A {\em fault} occurs when one component
-of a system behaves incorrectly (e.g., a routing node blocks or
-alters a message).
-The result of that fault (e.g., a recipient receiving conflicting messages)
-is an {\em error} state.
-If the error is undetected or corrected to the wrong value, the system is
-said to have experienced a {\em failure} (e.g., an altered message is
-accepted as authentic).
-We are concerned in particular with {\em adversarial faults},
-which are chosen strategically to maximize the likelihood of a failure.
-
-\subsection{Partial Trust Model}
-
-An important challenge in large-scale, secure communication is the ability to
-communicate reliably and securely with remote parties for whom no direct,
-trusted link exists.
-The commonly-used web of trust approach
-\cite{zimmermann_official_1995,ferguson_practical_2003}
-extends trust infinitely transitively:
-to friends of friends, and friends of friends of friends, and so on.
-However, the assumption of infinitely transitive trust is unrealistic
-\cite{christianson_why_1997},
-and does not allow for the analysis of the effects of network structure.
-
-We can allow for more realistic transitivity assumptions and incorporate
-network structure into a trust model with one key insight:
-even if no single path between a sender and receiver is fully trusted,
-multiple copies of a message can be sent along different paths and compared
-by the receiver to detect and correct errors.
-However, in the presence of an attacker, there is only benefit in sending an
-additional message along a path if that path does not share untrusted
-single points of failure with any of the existing message paths
-(otherwise, the adversary can compromise both messages
-by causing a single fault).
-Following \cite{reiter_resilient_1998},
-we call paths that do not share any untrusted single points of failure
-{\em independent}.
-The maximum number of independent paths represents the effective redundancy that
-can be utilized by any redundancy-based fault tolerance scheme.
-
-We now propose a method for quantifying the effective redundancy of a network
-as a function of its structure and level of trust transitivity.
-For the purpose of analyzing adversarial fault tolerance,
-this method allows a communication network
-to be modeled as a set of virtual channels that directly
-connect each node pair,
-with each virtual channel providing some level of redundancy.
-
-We now specify the {\em partial trust model}.
-We assume an undirected graph $G = (V,E)$,
-although the model can easily be extended to directed multigraphs.
-Vertices representing commiunicating agents,
-and with edges representing mutually trusted communication links.
-Let $v \in V$ be an arbitrary sender (Alice)
-and $w \in V$ be an arbitrary receiver (Bob).
-We assume the presence of an adversary (Mal) who knows the
-full structure of the network,
-and who can compromise a fixed number of nodes,
-gaining complete control of their behavior.
-We also assume that Mal is an adversary of Alice and/or Bob specifically,
-rather than the network as a whole.
-So Mal can compromise any node except for those trusted by Alice or Bob.
-We define a {\em trust radius} $h$ such that nodes $u$ and
-$u^\prime$ trust each other if their distance is less than $h$.
-For a given node $u$,
-we call the set of trusted nodes its
-{\em trusted neighborhood} $T_h(u)$,
-and all nodes at exactly distance $h$ the
-{\em trust boundary} $B_h(u)$:
-\beq
-T_h(u) &=& \left\{ u^\prime \mid d(u,u^\prime) < h \right\} \\
-B_h(u) &=& \left\{ u^\prime \mid d(u,u^\prime) = h \right\}.
-\eeq
-The trust boundary $B_h$ plays an important role because these nodes are not
-trusted by $u$,
-but if compromised can entirely isolate $u$ from the rest of the network.
-These trust assumptions imply that when Alice sends a message to Bob,
-Mal can only cause faults in the set of nodes outside both of their trusted
-neighborhoods: $V \setminus \left(T_h(v) \cup T_h(w)\right)$.
-We refer to this set of nodes as the {\em untrusted region}.
-
-Having defined the assumptions of the partial trust model,
-we now quantify the {\em effective redundancy} between Alice and Bob
-when trust radius $h$ is assumed.
-This quantity, $\delta_{v,w,h}$ is exactly the max-flow/min-cut of
-the graph after Alice's and Bob's trusted neighborhoods have been
-collapsed into single source/sink vertices.
-Each trust boundary forms a cut of the network and places an upper bound on the
-min-cut:
-\beq
-\delta_{v,w,h} \leq \min\left( \mid B_h(v) \mid, \mid B_h(w) \mid \right).
-\eeq
-Equality holds when there are no bottlenecks within the untrusted region,
-an indication that the network is decentralized.
-The redundancy of the entire graph can be characterized by the minimum over
-all vertex pairs:
-\beq
-\delta_h(V) \equiv \min_{v,w \in V} \delta_{v,w,h}.
-\eeq
-Thus, for any pair of nodes in the network, at least $\delta_h$ independent,
-redundant paths can be constructed between them.
-$\delta_h$ is a purely structural network property,
-and places an upper bound on the effectiveness of any
-redundancy-based fault tolerance scheme.
-The more quickly $\delta_h$ grows with $h$,
-the better a network is at creating redundancy from trust.
-Thus, the scaling of $\delta_h$ can be used to quantify a network's ability
-to withstand targeted attacks,
-even when the exact trust radius $h$ is unknown.
-
-\begin{figure}
-\centerline{\includegraphics[width=3in,height=2.08in]{fig-partial-trust}}
-\caption{
-Illustration of a trusted communication network and the network properties
-used by the {\em partial trust model}.
-Edges represent mutually trusted communication links.
-The sender (Alice, $v$) and receiver (Bob, $w$) trust all nodes
-less than the {\em trust radius} $h$ hops away.
-These nodes form their {\em trusted neighborhoods} $T_h(v)$ and $T_h(w)$.
-We assume that all faults occur in the remaining nodes: the
-{\em untrusted region}.
-The untrusted nodes in contact with the trusted neighborhoods for the
-{\em trust boundaries} $B_h(v)$ and $B_h(w)$,
-which (in the absence of central bottlenecks) determine the
-{\em effective redundancy} $\delta_h$ provided by the network.
-Alice and Bob can be modelled as connected by a direct link with
-at least $\delta_h$ redundant channels.
-}
-\label{fig:trust-source}
-\end{figure}
-
-\subsection{Multipath Fault Tolerance}
-
-Once we have determined a network's effective redundancy,
-we can apply redundancy-based fault tolerance techniques,
-by sending multiple copies of a message
-({\em concurrent multipath routing}).
-We model our sender (Alice) and receiver (Bob) as
-communicating over a direct link with $\delta_h$ redundant channels.
-The partial trust model allows us to make this simplifying assumption
-for analyzing a fault tolerance scheme,
-but implementing such a scheme will require a method for constructing
-specific network paths.
-We will return to the question of constructing paths in the next section.
-For now, we concern ourselves with the question:
-given that the network provides $\delta_h$ redundant channels between
-Alice and Bob,
-what is the probability that an adversary (Mal) causes an undetectable
-error?
-
-Let us first consider the case when
-Alice sends a message copy over each available channel.
-When Bob receives the messages, there are several possibilities.
-If some of the messages are missing
-(Alice can include the number of messages as part of the message)
-or if some of the messages disagree,
-Bob knows that some of the messages were either blocked or altered,
-and he has successfully detected an error.
-Bob has not accepted any compromised information,
-and can request retransmission, so no failure has occurred.
-If Bob receives all of the messages, and they all agree,
-he accepts the message.
-If the messages agree, there are two possible cases.
-The first case is that Mal has not compromised any of the messages,
-and Bob has correctly accepted them, so no failure has occurred.
-The second case is that Mal has compromised {\em all} of the messages,
-so Bob has accepted an erroneous message and a failure has occurred.
-In the present scenario,
-whether a failure occurs depends only on whether Mal has the resources to
-compromise all of the channels.
-In a more realistic scenario,
-both Alice and Mal have limited resources and are not able to use or
-compromise all available channels.
-
-In a more sophisticated multipath fault tolerance scheme,
-Alice randomly chooses $k \leq \delta_h$ channels and sends a copy of
-her message on each.
-We assume that Mal is capable of compromising $l \leq \delta_h$ channels.
-Having full knowledge of the network,
-Mal's best strategy is to identify a minimum node cut in the network
-and compromise nodes from that cut.
-With this strategy, each compromised node reduces effective redundancy by one,
-effectively compromising one of the virtual channels between Alice and Bob.
-Since Alice chooses channels randomly,
-all channels are equally likely to contain a message,
-so Mal can do no better than also choosing randomly.
-If $k > l$, at least one message will get through uncompromised and all
-errors are detectable.
-Otherwise, the probability of Mal producing an undetectable error is
-the probability that all of Alice's chosen channels are compromised:
-\beq
-\label{eq:pf}
-p_f &=& \frac{l!(\delta_h-k)!}{\delta_h!(l-k)!}.
-\eeq
-Letting $k=\alpha \delta_h$ and $l=\beta \delta_h$, then applying Stirling's
-approximation gives:
-\begin{eqnarray}
-\label{eq:pf_approx}
-p_f &\approx&
-\frac{\sqrt{\beta(1-\alpha)}}{\sqrt{\beta-\alpha}}
-\left[
-    \left( \frac{\beta-\alpha}{1-\alpha} \right)^{\alpha}
-    \left( \frac{\beta}{\beta-\alpha} \right)^{\beta}
-    (1-\alpha)
-\right]^{\delta_h}.
-\end{eqnarray}
-
-Figure \ref{fig:pfail} shows the value of $p_f$
-as a function of $k$ and $l$.
-Equation (\ref{eq:pf_approx}) shows that while $p_f$
-depends on the fractions of
-redundant paths actually utilized $\alpha$ and compromised $\beta$,
-it decreases exponentially with the effective redundancy $\delta_h$.
-This result is significant because $\delta_h$
-depends only on the network structure
-and the strength of trust transitivity.
-Thus, the scheme can be effective, even when the number of copies
-sent $k$ is a small fraction of the effective redundancy.
-In other words, this scheme exhibits a {\em stabilizing asymmetry}:
-senders can tolerate attacks from significantly more powerful
-adversaries,
-as long as the network structure provides large $\delta_h$.
-
-In order to derive the above results, we have assumed that Alice and all
-intermediary agents are able to identify specific,
-independent network paths that achieve the effective redundancy $\delta_h$.
-We now proceed to describe a routing algorithm for doing so in the special case
-of the butterfly network topology.
-
-\begin{figure}
-\centerline{\includegraphics{fig-perror}}
-\caption{
-The probability of an undetectable error as a function of the number of
-message copies and the number of adversarial faults.
-}
-\label{fig:pfail}
-\end{figure}
-
-\section{Multipath Butterfly Routing}
-\label{sec-butterfly}
-
-In previous sections, we showed that reliable communication across a network
-can be achieved even when any single message path might be compromised by
-an adversary,
-provided the network has sufficient redundancy,
-and provided the sender and intermediaries know how to route message
-copies along independent paths.
-In this section, we address the both requirements by proposing a novel routing
-algorithm for constructing independent paths on the butterfly network topology.
-This architecture and routing algorithm achieve an
-effective redundancy that increases exponentially with the trust radius,
-allowing a very high level of adversarial fault tolerance.
-
-The structure of the butterfly network is highly constrained,
-making it most suitable for applications where network structure can be
-designed or dictated.
-Examples of such networks include:
-overlay networks \cite{lua_survey_2005, korzun_structured_2013},
-formal organizations \cite{mohr_explaining_1982},
-government-regulated cellular networks \cite{walker_mass_2012},
-and call tree notification systems \cite{nickerson_thinking_2010}.
-For critical security applications in particular,
-the novel level of adversarial fault tolerance offered by this scheme
-is highly desirable,
-even if it requires new approaches to designing communication
-and trust networks.
-Lastly, we note that the partial trust model and multipath fault tolerance
-schemes of the previous section do not rely on any particular network
-topology or routing algorithm,
-and are not necessarily limited to such applications.
-
-\subsection{Butterfly Network Topology}
-
-We choose the butterfly topology
-\cite{kshemkalyani_distributed_2008}
-because of several desirable properties (described below)
-and because its structure allows for relatively straightforward
-design and analysis of routing algorithms.
-While several variations on the butterfly network exist,
-we utilize the wrap-around butterfly.
-We denote the $m$-dimensional, directed wrap-around butterfly as a
-graph $\wbf(m)$:
-\beq
-\wbf(m) &=& (V, E_\downarrow \cup E_\rightarrow) \\
-V &=& \mathbb{Z}_{m} \times \mathbb{Z}_2^m \\
-E_\downarrow
-&=&
-\{((l,z),(l+1 \; (\text{mod } m),z) \} \\
-E_\rightarrow
-&=&
-\{(l,z),(l+1 \; (\text{mod } m),
-z \oplus 1_l \},
-\eeq
-where $\mathbb{Z}_m$ is the set of integers modulo $m$,
-$\oplus$ represents componentwise addition modulo 2,
-and $1_l$ is a vector with a 1 in index $l$ and 0 elsewhere.
-Each node is associated with a level $l$ and an $m$-bit string $z$
-known as {\em the place-within-level}.
-There are two types of edges: down, and down-right
-(shown in Figure \ref{fig:butterfly}).
-Down edges ($E_\downarrow$) connect nodes sharing the same $z$ value
-in a cycle of increasing level $l$.
-Down-right edges ($E_\rightarrow$) also link to a node of level $l + 1$,
-but one having the place-within-level equal to $z$ with the $l$th bit inverted.
-
-The wrap-around butterfly network is known to have several of the properties
-we desire for scalable, decentralized communication networks:
-\begin{description}
-\item[Vertex-transitivity:]
-Because the wraparound butterfly is vertex transitive,
-it is maximally decentralized;
-\item[Small-diameter:]
-For any two nodes, the length of the shortest path between them is
-$O(\log N)$, where N is the number of nodes in the network;
-\item[Sparsity:]
-With a constant degree of 4, the wraparound butterfly is extremely sparse,
-and can scale indefinitely without node degree becoming a limitation;
-\item [Redundancy:]
-Multiple paths exist between any two nodes.
-Specifically, we will prove below that the number of independent paths between two
-nodes increases exponentially with the trust
-radius $h$.
-\end{description}
-
-\begin{figure}
-\begin{center}
-\includegraphics{fig-butterfly.pdf}
-\end{center}
-\caption{
-Schematic illustration of the two types of edges in a directed butterfly
-network.
-The node $(l,z)$ is shown as the bit string $z$ with a square around the
-$l$th bit.
-``Down'' edges increment $l$, leaving $z$ unchanged,
-while ``down-right'' edges increment $l$ and invert the $l$th bit of $z$.
-\label{fig:butterfly}
-}
-\end{figure}
-
-The structure of the butterfly network lends itself to a well-known
-(unipath) routing algorithm,
-which we later extend to the multipath case.
-The unipath algorithm first follows a down or down-right edge at every step,
-increasing the level $l$ by 1 and cycling through the
-indices of the place-within-level.
-If the current node's place-within-level matches the destination node's at
-index $l$,
-a down edge is chosen and the place-within-level does not change.
-Otherwise, a down-right edge is chosen and the $l$th component of the
-place-within-level is flipped,
-after which it matches the destination.
-After $m$ iterations of this, all levels have been visited
-and the place-within-level matches that of the destination.
-Simply following down (or up) edges will then increment (decrement) the
-level until the destination node is reached.
-
-\subsection{Multipath Routing Algorithm}
-
-We now present a routing algorithm to construct multiple independent paths
-between two nodes in a butterfly network.
-Independence requires that routes between two nodes $v$ and $w$ have no nodes
-in common, unless those nodes are less than distance $h$ from either $v$ or $w$.
-Informally, the algorithm achieves this by using the first $h$ hops to
-reach as many distinct nodes as possible,
-then routing from each of those to a distinct intermediary node,
-and finally routing from the intermediary to the destination.
-By ensuring that the place-within-level differs for all paths
-(outside of the trusted neighborhoods of $v$ and $w$)
-the algorithm guaranatees independence.
-As with the unipath routing algorithm,
-each of the multiple paths proceed using down and down-right edges,
-cycling through levels one at a time.
-However, we cycle through the levels twice, once to route from $v$ to a
-particular path's intermediary node,
-and again to route from the intermediary to $w$.
-In the first cycle, path independence is guaranteed by ensuring that
-all paths differ in the first $h$ bits of the place-within-level.
-Similarly, in the second cycle,
-independence is guaranteed by ensuring that all paths differ in the
-$h$ bits of the place-within-level preceding the destination index
-(see Figure \ref{fig:route-overview}).
-
-\afterpage{
-\begin{table}[h!]
-\caption{Butterfly Multipath Routing Variables\label{tab:routing}}
-{
-\begin{center}
-\begin{tabular}{|l|l|}
-\hline
-NAME & VARIABLE \\\hline
-butterfly dimension & $m \in \mathbb{Z}_+$ \\\hline
-node level & $l \in \mathbb{Z} : 0 \leq l < m$ \\\hline
-node place within level & $z \in \mathbb{Z}_2^m$ \\\hline
-trust radius & $h \in \mathbb{Z} : 1 \leq h \leq \lfloor m/2 \rfloor$ \\\hline
-path index & $s \in \mathbb{Z}_2^h$ \\\hline
-\end{tabular}
-\end{center}
-}
-\end{table}
-
-\begin{figure}[t!]
-\begin{center}
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-                        ]
-\foreach \i in {0\ldots,{\ldots}0\ldots,{\ldots}0\ldots,{\ldots}0}% <-- content of nodes
-    \node[X] {\i};
-\draw[<->] ([yshift=1.5mm] A-1.north east) -- node[above=0.25mm] {$h$} ([yshift=1.5mm] A-1.north west);
-\draw[<->] ([yshift=1.5mm] A-2.north east) -- node[above=0.25mm] {$l_w - 2h$} ([yshift=1.5mm] A-2.north west);
-\draw[<->] ([yshift=1.5mm] A-3.north east) -- node[above=0.25mm] {$h$} ([yshift=1.5mm] A-3.north west);
-\draw[<->] ([yshift=1.5mm] A-4.north east) -- node[above=0.25mm] {$m - l_w$} ([yshift=1.5mm] A-4.north west);
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {start} ( A-1.west);
-\node (B1) [inner sep=1pt,above=of A-1.north,above=5ex] {\underline{A}};
-\node (B2) [inner sep=1pt,above=of A-2.north,above=5ex] {\underline{B}};
-\node (B3) [inner sep=1pt,above=of A-3.north,above=5ex] {\underline{C}};
-\node (B4) [inner sep=1pt,above=of A-4.north,above=5ex] {\underline{D}};
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick},
-                        ]
-\node[Y] {$s$};
-\foreach \i in {{\ldots}0\ldots,{\ldots}0\ldots,{\ldots}0}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {1.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick},
-                        ]
-\foreach \i in {$s$}% <-- content of nodes
-    \node[X] {\i};
-\node[Y] {{\ldots}1\ldots};
-\foreach \i in {{\ldots}0\ldots,{\ldots}0}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {2.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick},
-                        ]
-\foreach \i in {$s$,{\ldots}1\ldots}% <-- content of nodes
-    \node[X] {\i};
-\node[Y] {$\tilde{s}$};
-\foreach \i in {{\ldots}0}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {3.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {anchor=base, rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick},
-                        ]
-\foreach \i in {$s$,{\ldots}1\ldots,$\tilde{s}$}% <-- content of nodes
-    \node[X] {\i};
-\node[Y] {$z_{w,D}$};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {4.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {anchor=base, rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick},
-                        ]
-\node[Y] {$z_{w,A}$};
-\foreach \i in {{\ldots}1\ldots,$\tilde{s}$,$z_{w,D}$}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {5.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {anchor=base, rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick},
-                        ]
-\foreach \i in {$z_{w,A}$}% <-- content of nodes
-    \node[X] {\i};
-\node[Y] {$z_{w,B}$};
-\foreach \i in {$\tilde{s}$,$z_{w,D}$}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {6.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {anchor=base, rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick},
-                        ]
-\foreach \i in {$z_{w,A}$,$z_{w,B}$}% <-- content of nodes
-    \node[X] {\i};
-\node[Y] {$z_{w,C}$};
-\foreach \i in {$z_{w,D}$}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {7.} ( A-1.west);
-\end{tikzpicture}
-\end{center}
-\caption{
-Progression of place-within-level $z$ as the multipath routing algorithm
-cycles through the levels of the butterfly network.
-\label{fig:route-overview}
-}
-\begin{center}
-\includegraphics{fig-routing.pdf}
-\end{center}
-\caption{
-An example of one path as constructed by the proposed multipath
-routing algorithm.
-The path is shown for $s = 10_2$
-and $w = (6, 0110111_2)$.
-\label{fig:routing}
-}
-\end{figure}
-}
-
-We now begin the formal specification of our multipath routing scheme for the
-wraparound butterfly network.
-Utilizing vertex transitivity, we label the source node as
-$(l^{(0)}, z^{(0)}) = (0, 0)$ and denote the destination node as $w = (l_w, z_w)$,
-without loss of generality.
-
-Let $s$ be an $h$-bit binary string with $s_i$ denoting the bit at index $i$.
-There are $2^h$ such strings.
-Let $v_s^{(t)} = (l^{(t)}, z^{(t)})$ be the node at position $t$
-in the path parameterized by $s$.
-For convenience,
-we will omit the subscript $s$ when it is obvious from context.
-We define three distinct partitions of $m$-bit binary strings.
-Let $Q_{v^{(0)}}$ be the set of $m$-bit
-strings in which the bits at
-all indices $h \leq i < l_w - h$ match those of $z^{(0)}$,
-and let $\overline{Q_{v^{(0)}}}$ be its complement.
-Note that $Q_{v^{(0)}}$ is trivially all $m$-bit strings if $l_w < 2h$.
-Let $R_s$ be the set of $m$-bit strings with the lowest $h$
-bits all matching the bits of $s$,
-and let $\overline{R_s}$ be its complement.
-Let $S_s$ be the set of $m$-bit strings with the $h$ bits
-preceding index $l_w$ all matching the bits of $\tilde{s}$,
-where $\tilde{s}$ is a cyclic permutation of $s$:
-\beq
-\tilde{s}_i &=& s_{(i + l_w) \text{ mod } h},
-\eeq
-and let $\overline{S_s}$ be its complement.
-We will make use of the fact that:
-\beq
-s \neq s^\prime &\implies&
-S_s \cap S_{s^\prime} = R_s \cap R_{s^\prime} = \emptyset.
-\eeq
-
-Routes are constructed in 7 stages.
-The network topology dictates that $l^{(t+1)} = l^{(t)} + 1$ (mod $m$),
-so we let $l = t$ (mod $m$).
-and that $z^{(t+1)}$ is equal to $z^{(t)}$ with or without the bit in index
-$l^{(t)}$ inverted, depending on whether the down or down-right edge was
-taken at step $t$.
-\begin{description}
-\item[Stage 1: ($0 \leq t < h$)]{
-Down or down-right edges
-are chosen such that the $t$th bit of $z^{(t+1)}$ is equal to the $t$th bit
-of $s$.
-Throughout Stage 1, all nodes are within the sender's trusted neighborhood.
-Throughout Stage 1, $z^{(t)} \in Q_{v^{(0)}}$.
-At the end of Stage 1, $z^{(h)} \in S_s$, and $z^{(t)}$ will remain so until the level cycles to $0$ at $t = m$.
-}
-\item[Stage 2: ($h \leq t < l_w - h$)]{
-Edges are chosen to make the $t$th bit of
-$z^{(t+1)}$ the inverse of the $t$th bit of $z^{(0)}$.
-Note that this stage does not occur when $l_w < 2h$.
-If this stage occurs, then $z^{(t)} \in \overline{Q_{v^{(0)}}}$ until these
-levels are reached again in stage 6.
-}
-\item[Stage 3: ($l_w - h \leq t < l_w$)]{
-The bits of $z^{(t)}$ are chosen to match $\tilde{s}$,
-such that after the stage is complete, $z^{(t)} \in R_s$.
-}
-\item[Stage 4: ($l_w \leq t < m$)]{
-Paths are chosen such that the $t$th bit of $z^{(t+1)}$ matches that of the
-destination node $z_w$.
-This stage will not occur if $l_w > m - h$.
-}
-\item[Stage 5: ($m \leq t < m + h$)]{
-There are two cases.
-If $2h < l_w < m - h$,
-then there is no overlap between the indices defining $R_s$ and $S_s$.
-In this case, the first $h$ bits of $z^{(t)}$ are set to
-match $z_w$.
-Otherwise there is some overlap between the indices defining $R_s$ and
-$S_s$.
-In this case, the each of the first $h$ bits of $z^{(t)}$ is either kept the
-same if $l_w - h \leq l < l_w$, or set to the corresponding bit of $z_w$
-otherwise.
-In this stage and after, $z^{(t)}$ is no longer guaranteed to be in $R_s$.
-However, $z^{(t)}$ remains in $S_s$ during and after this stage.
-}
-\item[Stage 6: ($m + h \leq t < m + l_w - h$)]{
-In this stage, edges are chosen to set the bits of $z^{(t)}$ to their
-corresponding value in $z_w$.
-$z^{(t)} \in \overline{Q_{v^{(0)}}}$ throughout this stage,
-but not afterwards.
-}
-\item[Stage 7: ($m + l_w - h \leq t < m + l_w$)] {
-The $h$ bits of $z^{(t)}$
-preceding index $l_w$ are set to match $z_w$.
-All nodes in this stage are within $h$ hops of $w$ and thus in its trusted
-neighborhood.
-After this stage, $v^{(m + l_w)} = w$ and routing is complete.
-}
-\end{description}
-
-\begin{theorem}
-Given an $m$-bit wraparound butterfly network ($m \geq 0$),
-and a trust radius $h \geq 1$,
-for all node pairs $(v, w)$,
-if there exists a node $u$ that belongs to
-two distinct routes $v^{(\cdot)}_s$ and $v^{(\cdot)}_{s^\prime}$
-($s \neq s^\prime$),
-it implies that $u \in T_h(v) \cup T_h(w)$.
-\end{theorem}
-\begin{proof}
-Nodes from two paths can only coincide if their levels are the same.
-Nodes which share a level must either be in the same stage, or 4 stages
-apart.
-Let ($a$,$a^\prime$) denote a pair of sub-paths corresponding to stage $a$ of
-one path and stage $a^\prime$ of another.
-Excluding paths that intersect in their trusted neighborhoods, (1,1) and (7,7),
-we have reduced the list of possible intersections to the following cases:
-(2,2), (3,3), (4,4), (5,5), (6,6), (1,5), (2,6), and (3,7).
-Nodes in stages 2--4 belong to $R_s$ so cannot overlap with any stage 2--4
-nodes from another path, eliminating (2,2), (3,3), and (4,4).
-Similarly, nodes in stages 4--6 belong to a unique $S_s$,
-eliminating (5,5) and (6,6).
-Nodes in stage 1 belong to $Q_{v^{(0)}}$ while those in stage 5 belong in
-its complement, eliminating (1,5).
-Similarly, for all $l$ in stage 2, $z^{(l)}$ is equal to $z^{(0)}$,
-while in stage 6, $z^{(l)}$ is the inverse, eliminating (2,6)
-This leaves only (3,7), a collision which can occur only for only one path
-(with $s$ matching the first $h$ bits of $z_w$), and which enters the trusted
-neighborhood in stage 3.
-For this single path, we can proceed directly from stage 2 to stage 7,
-eliminating the last possible collision.
-\end{proof}
-
-Thus, assuming the partial trust model with trust transitive
-for $h$ hops, we can construct $2^h$ paths on a wraparound butterfly topology
-which do not intersect outside the trusted neighborhoods of the source and
-destination.
-Note that the node sequence $v_s^{(t)}$ can be calculated entirely
-from the source $v$, destination $w$, and path parameter $s$,
-meaning that with this information nodes are able to determine which neighbor
-to route a given message copy to.
-Furthermore, the existence of $2^h$ paths places a lower bound on the
-effective redundancy $\delta_h$,
-showing that the decentralized, redundant, structured networks such as the
-butterfly can have a very low probability of failure when faced with
-adversarial faults, even from a very powerful attacker.
-
-\section{Discussion}
-\label{sec-discussion}
-
-While decentralized protocols have received much attention for their potential
-fault tolerance applications,
-a better understanding of how network structure contributes to fault tolerance
-is necessary.
-We have proposed a network-based scheme for adversarial fault tolerance on the
-butterfly topology,
-utilizing a novel {\em concurrent multipath routing} algorithm.
-We have also demonstrated how {\em partial trust transitivity},
-in addition to being more realistic than infinite transitivity,
-enables the analysis of the relationship between trust, network structure,
-and fault tolerance.
-
-Fault-tolerant network infrastructures have many direct applications.
-Areas such as cryptocurrency
-\cite{mazieres_stellar_2015,nakamoto_bitcoin:_2008,king_ppcoin:_2012},
-secure multiparty computation
-\cite{yao_protocols_1982,chaum_multiparty_1988,goldreich_how_1987},
-and wireless sensor networks
-\cite{khiani_comparative_2013}
-have immediate need for scalable, fault-tolerant infrastructures.
-Many Internet services (e.g., email, social networks, cloud storage)
-are still highly centralized and vulnerable to technical and
-non-technical (i.e., coercive) attacks.
-Fault tolerance using {\em both}
-decentralized protocols and
-decentralized network structures
-is one promising approach to securing these services.
-
-We have focused specifically on adversarial faults that block or
-change messages (e.g., censorship).
-We have not specifically addressed {\em confidentiality}:
-how to keep those messages private.
-Existing cryptographic techniques for confidentiality are relatively
-mature compared to those for tolerating censorship.
-In fact, an attacker might use targeted attacks to achieve censorship
-as a fallback when surveillance isn't feasible.
-However, the techniques presented in this paper
-are entirely compatible with, and in some cases could enhance, existing
-confidentiality techniques.
-For example, {\em man-in-the-middle} attacks exploit a privileged
-network position to attack otherwise secure cryptography,
-suggesting that structural approaches can complement cryptographic ones.
-
-While our present proposal is specific to the butterfly topology,
-the multipath fault tolerance scheme could
-be applied to any network that has both sufficient redundancy and a
-routing algorithm to discover independent paths.
-For general networks, finding all such paths is NP-hard,
-but efficient, suboptimal algorithms exist
-\cite{reiter_resilient_1998}.
-However, our approach is best suited for structured networks.
-While the ability to control network topology is only achievable in some
-applications, those applications are nonetheless significant
-(e.g. overlay networks \cite{lua_survey_2005, korzun_structured_2013},
-formal organizations \cite{mohr_explaining_1982},
-government-regulated cellular networks \cite{walker_mass_2012},
-and call tree notification systems \cite{nickerson_thinking_2010}).
-Furthermore, when high levels of adversarial fault tolerance are required,
-designers of new sociotechnical systems might choose to
-incorporate such networks.
-For example,
-a coalition of groups supporting free expression could use our results
-to construct a censorship-resistant communication network
-by cultivating the necessary trust relationships.
-
-Our work suggests several directions for future work towards developing
-practical, attack-tolerant communication infrastructure.
-The development of new multipath routing algorithms on other structured
-networks could achieve higher levels of redundancy.
-It is also desirable to identify dynamics that give
-rise to structured networks,
-and to evaluate whether our results can be generalized to unstructured or
-approximately structured networks.
-Finally, these results could be implemented to address specific
-applications, e.g., secure messaging, domain name resolution, or
-anonymous web browsing.
-
-\section{Conclusion}
-\label{sec-conclusion}
-
-We have presented a novel multipath routing algorithm for the butterfly
-topology,
-as well as a scheme for using this algorithm to achieve a high level of
-attack-tolerance in a communication network,
-even when no single path is entirely trusted.
-Under this scheme, the probability of an adversary causing an undetectable error
-decreases exponentially with the network's effective redundancy.
-The effective redundancy, in the case of the butterfly topology,
-grows exponentially with the trust radius.
-Furthermore, a small increase in the number of messages sent can compensate
-for a large increase in the number of messages compromised by an adversary.
-We have also demonstrated how the assumption of partial trust
-transitivity can enable a formal analysis of the effects of network structure
-on attack-tolerance.
-These results are directly applicable to systems in which the link
-structure can be imposed by the designer,
-and more generally, provide a formalism that can be used more to evaluate
-the role of network structure, trust transitivity, and effective redundancy
-on adversarial fault tolerance.
-
-%ACKNOWLEDGMENTS are optional
-\section{Acknowledgments}
-
-%
-% The following two commands are all you need in the
-% initial runs of your .tex file to
-% produce the bibliography for the citations in your paper.
-\bibliographystyle{abbrv}
-\bibliography{paper}  % sigproc.bib is the name of the Bibliography in this case
-% You must have a proper ".bib" file
-%  and remember to run:
-% latex bibtex latex latex
-% to resolve all references
-%
-% ACM needs 'a single self-contained file'!
-%
-%
-\end{document}

### 2017-03-01 websci-2017/source/paper.tex
Update algorithm and proof.

@@
 \item{We prove that the number of independent paths between two nodes
-in a wraparound butterfly network with partial trust transitivity
+in a wrap-around butterfly network with partial trust transitivity
 increases exponentially with the trust radius;
@@
 trusted by $u$,
-but if compromised can entirely isolate $u$ from the rest of the network.
+and if compromised can entirely isolate $u$ from the rest of the network.
 These trust assumptions imply that when Alice sends a message to Bob,
@@
 
-We now propose a method for quantifying the effective redundancy of a network
-as a function of its structure and level of trust transitivity.
-For the purpose of analyzing adversarial fault tolerance,
-this method allows us to model each pair of nodes
-in a complex communication network
-as being directly connected by a dedicated {\em virtual channel}
-that provides some level of redundancy.
-
-Having defined the assumptions of the partial trust model,
-we now quantify the {\em effective redundancy} between Alice and Bob
+Our goal is to achieve fault tolerance through redundancy.
+In the network setting, redundnacy is achieved using
+{\em independent paths} \cite{reiter_resilient_1998},
+which no common points of failure.
+Typically, it is assumed that paths must be disjoint in order to be
+independent.
+However, under the partial trust assumption, two non-disjoint paths
+can still be independent as long as their intersection contains only trusted
+nodes,
+greatly increasing the level of redundancy available.
+Under the partial trust assumption, the available redundancy thus depends on
+both the network structure and the level of trust.
+
+We now quantify the {\em effective redundancy} between Alice and Bob
 when trust radius $h$ is assumed.
@@
 an indication that the network is decentralized.
-The redundancy of the entire graph can be characterized by the minimum over
+The efrective redundancy of the entire graph can be characterized by the minimum over
 all vertex pairs:
@@
 The more quickly $\delta_h$ grows with $h$,
-the better a network is at creating redundancy from trust.
+the better a network is at leveraging trust transitivity to create redundancy.
 Thus, the scaling of $\delta_h$ can be used to quantify a network's ability
@@
 ({\em concurrent multipath routing}).
+Having found the effective redundancy between two nodes,
+we can simplify our model, replacing independent paths through the
+complex network with direct channels between the endpoints.
 We model our sender (Alice) and receiver (Bob) as
-communicating over a direct link with $\delta_h$ redundant channels.
+communicating over $\delta_h$ direct and redundant virtual channels.
 The partial trust model allows us to make this simplifying assumption
@@
 what is the probability that an adversary (Mal) causes an undetectable
-error?
-
-Let us first consider the case when
+error after inducing a fixed number of faults?
+
+Let us first consider the scenario in which
 Alice sends a message copy over each available channel.
+We can also assume that each message includes the number of messages sent,
+the full list of channels used, etc.,
+making that information available to Bob.
 When Bob receives the messages, there are several possibilities.
 If some of the messages are missing
-(Alice can include the number of messages as part of the message)
 or if some of the messages disagree,
 Bob knows that some of the messages were either blocked or altered,
-and he has successfully detected an error.
-Bob has not accepted any compromised information,
-and can request retransmission, so no failure has occurred.
-If Bob receives all of the messages, and they all agree,
-he accepts the message.
-If the messages agree, there are two possible cases.
+and he has successfully tolerated the fault(s).
+Bob can then take any of several actions:
+1. request retransmission;
+2. end receipts so Alice knows which paths have been compromised;
+or 3. attempt error correction using majority voting.
+If instead, Bob finds that all the messages are present and agree,
+there are two possible cases.
 The first case is that Mal has not compromised any of the messages,
@@
 redundant paths actually utilized $\alpha$ and compromised $\beta$,
-it decreases exponentially with the effective redundancy $\delta_h$.
+it decreases exponentially with the effective redundancy $\delta_h$
+(which we will later see increases exponentially in $h$ in the
+butterfly topology).
 This result is significant because $\delta_h$
@@
 The structure of the butterfly network is highly constrained,
-making it most suitable for applications where network structure can be
-designed or dictated.
+making it most suitable for applications where portions of the
+network structure can be designed or dictated.
 Examples of such networks include:
@@
 and call tree notification systems \cite{nickerson_thinking_2010}.
-For critical security applications in particular,
-the novel level of adversarial fault tolerance offered by this scheme
-is highly desirable,
-even if it requires new approaches to designing communication
-and trust networks.
+However, when attack-tolerance is desired,
+it will always require control over network structure in order to
+eliminate single points of failure.
+The regular structure of the butterfly is not a limitation of our approach,
+but rather a reflection of the inherent difficulty of attack-tolerance.
 Lastly, we note that the partial trust model and multipath fault tolerance
@@
 topology or routing algorithm,
-and are not necessarily limited to such applications.
+and our choice of the butterfly topology is only one of many possible choices.
 
@@
 \item[Vertex-transitivity:]
-Because the wraparound butterfly is vertex transitive,
+Because the wrap-around butterfly is vertex transitive,
 it is maximally decentralized;
@@
 \item[Sparsity:]
-With a constant degree of 4, the wraparound butterfly is extremely sparse,
+With a constant degree of 4, the wrap-around butterfly is extremely sparse,
 and can scale indefinitely without node degree becoming a limitation;
@@
 while ``down-right'' edges increment $l$ and invert the $l$th bit of $z$.
+In the wrap-around variant, the nodes with maximum $l$ have down and down-right
+edges to the nodes with $l=0$.
 \label{fig:butterfly}
@@
 
-We now present a routing algorithm to construct multiple independent paths
-between two nodes in a butterfly network.
-Independence requires that routes between two nodes $v$ and $w$ have no nodes
-in common, unless those nodes are less than distance $h$ from either $v$ or $w$.
-Informally, the algorithm achieves this by using the first $h$ hops to
-reach as many distinct nodes as possible,
-then routing from each of those to a distinct intermediary node,
-and finally routing from the intermediary to the destination.
-By ensuring that the place-within-level differs for all paths
-(outside of the trusted neighborhoods of $v$ and $w$)
-the algorithm guaranatees independence.
+We now present a routing algorithm to construct $2^h$ independent paths
+between two nodes in a butterfly network,
+where $h$ is the trust radius under the partial trust model.
+Informally, the algorithm sends each message to a distinct node on Alice's
+trust boundary, then to a distinct intermediate node in the untrusted region,
+then to a distinct node on Bob's trust boundary, and finally to Bob.
+The intermediate nodes are in a sense "far" from each other and ensure that
+no two paths overlap in the untrusted region.
+Each path can be parameterized by a single integer $s$, which identifies
+the specific node on Alice's trust boundary
+(or equivalentely a node on Bob's trust boundary, or the untrusted intermdiate).
+
+The algorithm guarantees paths are independent by ensuring that
+(outside the trusted neighborhoods)
+they only include
+nodes that match the path parameter $s$ at certain indexes in their
+place-within-level.
+Since each path has a unique parameter $s$,
+its set of untrusted nodes is disjoint from all other paths.
 As with the unipath routing algorithm,
-each of the multiple paths proceed using down and down-right edges,
+each of the multiple paths proceed from a source $v$ to a destination $u$
+using down and down-right edges,
 cycling through levels one at a time.
@@
 and again to route from the intermediary to $w$.
-In the first cycle, path independence is guaranteed by ensuring that
-all paths differ in the first $h$ bits of the place-within-level.
-Similarly, in the second cycle,
-independence is guaranteed by ensuring that all paths differ in the
+In the first cycle (stages 1--4), path independence is guaranteed by ensuring that
+all nodes match the path parameter $s$ in the first $h$ bits of the place-within-level.
+Similarly, in the second cycle (stages 5--7),
+independence is guaranteed by ensuring that all paths match $s$ in the
 $h$ bits of the place-within-level preceding the destination index
@@
 We now begin the formal specification of our multipath routing scheme for the
-wraparound butterfly network.
+wrap-around butterfly network.
 Utilizing vertex transitivity, we label the source node as
@@
 \begin{theorem}
-Given an $m$-bit wraparound butterfly network ($m \geq 0$),
-and a trust radius $h \geq 1$,
-for all node pairs $(v, w)$,
-if there exists a node $u$ that belongs to
-two distinct routes $v^{(\cdot)}_s$ and $v^{(\cdot)}_{s^\prime}$
-($s \neq s^\prime$),
-it implies that $u \in T_h(v) \cup T_h(w)$.
+Given an $m$-bit wrap-around butterfly network ($m > 1$),
+and a radius $h$ ($1 \leq h \leq \left\lfloor \frac{m}{2} \right\rfloor$),
+for all node pairs $(v, w)$ such that $d(v,w) \geq 2h$,
+there exist $2^h$ paths $v_s$ ($0 \le s < 2^h$) from
+$v$ to $w$ such that
+$s \neq s^\prime \implies v_s \cap v_{s^\prime} \subset T_h(u) \cup T_h(v)$.
 \end{theorem}
@@
 Thus, assuming the partial trust model with trust transitive
-for $h$ hops, we can construct $2^h$ paths on a wraparound butterfly topology
+for $h$ hops, we can construct $2^h$ paths on a wrap-around butterfly topology
 which do not intersect outside the trusted neighborhoods of the source and

### 2017-03-01 websci-2017/source/paper.tex
Update discussion conclusion abstract.

@@
 \begin{abstract}
-Networks with single points of failure are particularly susceptible to
-targeted attacks.
-In communication networks, these types of faults can leave users vulnerable to
+Targeted attacks against network infrastructure are notoriously difficult
+to withstand.
+Much of the existing work on network fault-tolerance focuses on random faults and
+does not apply to adversarial faults (attacks).
+In the case of communication networks,
+targeted attacks can leave users vulnerable to
 censorship and targeted surveillance,
-even when cryptography is utilized.
+even when cryptography is used.
 Centralized networks have single points of failure by definition,
-leading to a growing popularity in decentralized architectures and protocols.
-However, centralized network structure can arise even when protocols are
+leading to a growing popularity in decentralized architectures and protocols
+for greater fault-tolerance.
+However, centralized network {\em structure} can arise even when {\em protocols} are
 decentralized.
-Despite being based on decentralized protocols,
+Despite their decentralized protocols,
 the Internet and World-Wide Web have
 been shown both theoretically and historically to be highly susceptible to
-adversarial faults, in part due to emergent structural centralization.
-Existing network trust models, such as webs of trust, fail to adequately address
-network structure.
-We describe a novel, adversarial fault-tolerant,
-concurrent multipath routing algorithm for the decentralized
-butterfly network topology.
-We also develop a partial trust model that makes it possible to
-quantify the adversarial fault tolerance of a network,
-and which makes more realistic transitivity assumptions than webs of trust.
-When network topology can be dictated, these results can be used to create
-scalable, attack-tolerant infrastructures. More generally, our results provide
-a formalism for evaluating the effects of network structure on adversarial
-fault tolerance.
+attack, in part due to emergent structural centralization.
+When single points of failure exist,
+they are potentially vulnerable to non-technological
+(i.e., coercive) attacks,
+suggesting the importance of a structural approach to attack-tolerance.
+We show how the assumption of partial trust transitivity,
+while more realistic than the assumption used by webs of trust,
+can be used to quantify the effective redundancy of a network
+as a function of trust transitivity.
+We also prove that the effective redundancy of the wrap-around butterfly topology
+increases exponentially with trust transitivity and
+describe a novel concurrent multipath routing algorithm for constructing
+paths to utilize that redundancy.
+When network structure can be dictated
+our results can be used to create scalable,
+attack-tolerant infrastructures.
+More generally, our results provide a theoretical formalism for evaluating the effects
+of network structure on adversarial fault-tolerance.
 \end{abstract}
@@
 Decentralization}
-
-\vskip 3em
 
@@
 fault tolerance applications,
-a better understanding of how network structure contributes to fault tolerance
-is necessary.
-We have proposed a network-based scheme for adversarial fault tolerance on the
+centralized structures are always vulnerable to exploitation by non-technical means (i.e., coercion),
+and there is a need for a better understanding of the relationship between network
+structure and attack-tolerance.
+We have proposed a network-based scheme for {\em adversarial} fault tolerance on the
 butterfly topology,
-utilizing a novel {\em concurrent multipath routing} algorithm.
+utilizing a novel concurrent multipath routing algorithm.
 We have also demonstrated how {\em partial trust transitivity},
 in addition to being more realistic than infinite transitivity,
-enables the analysis of the relationship between trust, network structure,
-and fault tolerance.
+provides a theoretical foundation for
+quantitative analysis of the relationship between trust,
+network structure, and attack tolerance.
+
+Such attacks include many forms of censorship and surveillance,
+which have important social implications.
+We have already discussed two such cases:
+Pakistan's inadvertant censorship of YouTube
+\cite{hunter_pakistan_2008}
+and the FBI's surveillance-turned-censorship of Lavabit
+\cite{poulsen_edward_2013}
+The reader may wonder how our methods could be employed in scenarios
+such as large-scale state-sponsored censorship.
+Censorship-resistant infrastructure often replaces central servers
+(e.g., the router in the 2008 YouTube incident) with multiple servers across
+the world, synchronized through consensus protocols.
+The {\em directory authorities} used by the Tor
+\cite{dingledine_tor_2004} network are one example.
+However, the size of such authority networks is often limited by the number of
+trust relationships (degree) each node can maintain, and the inherent insecurity of
+extending transitive trust to an ever-larger network.
+Our work fills a much-needed gap by quantifying the connection between network-structure,
+trust transitivity,
+and attack-tolerance.
+We provide both a theoretical framework and specific example of how network structure
+can be engineered to leverage trust for a high level of attack-tolerance,
+without sacrificing scalability.
 
@@
 
-We have focused specifically on adversarial faults that block or
+We have focused primarily on adversarial faults that block or
 change messages (e.g., censorship).
-We have not specifically addressed {\em confidentiality}:
-how to keep those messages private.
-Existing cryptographic techniques for confidentiality are relatively
-mature compared to those for tolerating censorship.
-In fact, an attacker might use targeted attacks to achieve censorship
-as a fallback when surveillance isn't feasible.
+Existing cryptographic techniques for circumventing surveillance
+are relatively mature compared to those for tolerating censorship.
 However, the techniques presented in this paper
 are entirely compatible with, and in some cases could enhance, existing
-confidentiality techniques.
+anti-surveillance techniques.
 For example, {\em man-in-the-middle} attacks exploit a privileged
@@
 \cite{reiter_resilient_1998}.
-However, our approach is best suited for structured networks.
-While the ability to control network topology is only achievable in some
-applications, those applications are nonetheless significant
-(e.g. overlay networks \cite{lua_survey_2005, korzun_structured_2013},
-formal organizations \cite{mohr_explaining_1982},
-government-regulated cellular networks \cite{walker_mass_2012},
-and call tree notification systems \cite{nickerson_thinking_2010}).
-Furthermore, when high levels of adversarial fault tolerance are required,
-designers of new sociotechnical systems might choose to
-incorporate such networks.
-For example,
-a coalition of groups supporting free expression could use our results
-to construct a censorship-resistant communication network
-by cultivating the necessary trust relationships.
+However, we have argued that attack-tolerance requires the ability to
+influence network structure and reduce reilance on single points of failure.
+Our work is most applicable to cases where the need for attack-tolerance
+justifies investment in deliberate infrastructure.
+For example, a coalition of groups supporting free expression could use our work
+to construct a censorship-resistant communication network.
+In general, such groups would need to invest resources into vetting their neighbors
+to establish trust, but there are scenarios in which the attack-tolerance
+requirement would justify that investment.
+It is also worth noting that because faulty paths can be identified in our scheme,
+it may at times be appropriate to assume mutual trust and revoke that relationship
+when it is violated.
+Any entities dependent on the proper functioning of the network would have an incentive
+to resist attack in order to maintain their ability to participate.
 
@@
 
-We have presented a novel multipath routing algorithm for the butterfly
+We have presented a novel concurrent multipath routing algorithm for the butterfly
 topology,
-as well as a scheme for using this algorithm to achieve a high level of
-attack-tolerance in a communication network,
-even when no single path is entirely trusted.
+as well as a scheme for using this algorithm to construct a highly attack-tolerant
+virtual channel between any two nodes, even when no fully-trusted path exists
+between them.
 Under this scheme, the probability of an adversary causing an undetectable error
@@
 We have also demonstrated how the assumption of partial trust
-transitivity can enable a formal analysis of the effects of network structure
-on attack-tolerance.
+transitivity can enable a quantitative analysis of the
+relationships between network structure, trust, and attack-tolerance.
 These results are directly applicable to systems in which the link
 structure can be imposed by the designer,
-and more generally, provide a formalism that can be used more to evaluate
+and more generally, provide a theoretical foudnation that can be used more to evaluate
 the role of network structure, trust transitivity, and effective redundancy
-on adversarial fault tolerance.
+on attack-tolerance.
 

### 2017-03-01 websci-2017/source/paper.tex
tweaks.

@@
 %Conference
-\conferenceinfo{WWW '17}{April 3--7, 2017, Perth, Australia}
-
-\acmPrice{\$15.00}
+\conferenceinfo{WebSci '17}{June 26--8, 2017, Troy, NY, USA}
+
+%\acmPrice{\$15.00}
 
@@
 Targeted attacks against network infrastructure are notoriously difficult
-to withstand.
+to guard against.
 Much of the existing work on network fault-tolerance focuses on random faults and
@@
 targeted attacks can leave users vulnerable to
-censorship and targeted surveillance,
+censorship and surveillance,
 even when cryptography is used.
@@
 \keywords{
-Butterfly Network;
+Butterfly Topology;
 Fault Tolerance;
@@
 This paper presents several contributions towards advancing those goals.
+While our motivation comes from the Internet router network and
+World Wide WEb,
+our work is theoretical, focusing on abstract networks,
+and could potentially be applied to many different types of networks,
+whether made of physical wires, virtual tunnels,
+or other types of links.
 
@@
 Generally, faults in network paths can be correlated,
-preventing the appliation of standard fault tolerance techniques
+preventing the application of standard fault tolerance techniques
 \cite{avizienis_basic_2004, von_neumann_probabilistic_1956},
@@
 
-Previous work on evaluating incorporating network structure into
+Previous work on incorporating network structure into
 trust models has focused on authentication
@@
 copies along independent paths.
-In this section, we address the both requirements by proposing a novel routing
+In this section, we address both requirements by proposing a novel routing
 algorithm for constructing independent paths on the butterfly network topology.

### 2017-03-16 websci-2017/source/paper.tex
Feedback from Daniel.

@@
 to guard against.
+In the case of communication networks,
+such attacks can leave users vulnerable to
+censorship and surveillance,
+even when cryptography is used.
 Much of the existing work on network fault-tolerance focuses on random faults and
 does not apply to adversarial faults (attacks).
-In the case of communication networks,
-targeted attacks can leave users vulnerable to
-censorship and surveillance,
-even when cryptography is used.
 Centralized networks have single points of failure by definition,
@@
 The action was initiated by government order in Pakistan,
-and spread beyond Pakistan when a single misconfigured router allowed the false
+and spread beyond Pakistan when a router misconfiguration allowed the false
 routing information to propagate globally.
@@
 While our motivation comes from the Internet router network and
-World Wide WEb,
+World Wide Web,
 our work is theoretical, focusing on abstract networks,
@@
 (rather than disrupting the network as a whole).
+This assumption arises because scalable network architectures can be made
+large enough that an attacker must focus their resources in order to have
+a significant effect.
 Compromised nodes may behave incorrectly by blocking, altering,
@@
 on attack-tolerance,
-how to constructed trusted channels from redundant partially trusted paths,
-and propose a method for achieving a high-level of attack-tolerance on
+how to use local trust and redundancy to achieve greater attack-tolerance
+when no single path is fully trusted,
+and propose a novel routing algorithm for constructing such paths on
 the butterfly network topology.
@@
 
-The reader may wonder whether it is realistic or useful to assume control over
+It is important to note that the butterfly is a highly structured and constrained
+network topology,
+very different form those found in social networks and other
+self-organized networks.
+The reader may wonder whether it is realistic or useful to assume such control over
 the network structure.
@@
 those servers.
+However, there may be ways to achieve improved attack-tolerance from architectures
+more flexible than the butterfly, which is a potential area for future work.
 
@@
 fault tolerance.
-As an alternative to the NP-hard task of finding independent paths on arbitrary
-networks, we provide a routing algorithm to efficiently construct independent
-paths on a the butterfly topology.
+When network topology can be controlled, we sidestep the NP-hard problem of finding
+independent paths on arbitrary networks by using the mathematical structure of
+the butterfly topology to construct provably independent paths.
 
@@
 and does not rely on assumptions about existing social network structure.
-
-send redundant information
-over multiple paths when routing a message through a network,
 
@@
 {\em independent paths} \cite{reiter_resilient_1998},
-which no common points of failure.
+which have no common points of failure.
 Typically, it is assumed that paths must be disjoint in order to be
@@
 where $h$ is the trust radius under the partial trust model.
-Informally, the algorithm sends each message to a distinct node on Alice's
+Informally, Alice sends each message to a distinct node on her
 trust boundary, then to a distinct intermediate node in the untrusted region,
 then to a distinct node on Bob's trust boundary, and finally to Bob.
-The intermediate nodes are in a sense "far" from each other and ensure that
+The intermediate nodes are in a sense ``far'' from each other and ensure that
 no two paths overlap in the untrusted region.
@@
 the specific node on Alice's trust boundary
-(or equivalentely a node on Bob's trust boundary, or the untrusted intermdiate).
+(or equivalentely the node on Bob's trust boundary, or the untrusted intermdiate).
 
@@
 and again to route from the intermediary to $w$.
-In the first cycle (stages 1--4), path independence is guaranteed by ensuring that
+Each cycle is divided into stages,
+with different properties used to prove independence in different stages
+(see Figure \ref{fig:route-overview}).
+In the first cycle (stages 2--4), path independence is guaranteed by ensuring that
 all nodes match the path parameter $s$ in the first $h$ bits of the place-within-level.
-Similarly, in the second cycle (stages 5--7),
+Similarly, in the second cycle (stages 5--6),
 independence is guaranteed by ensuring that all paths match $s$ in the
-$h$ bits of the place-within-level preceding the destination index
-(see Figure \ref{fig:route-overview}).
+$h$ bits of the place-within-level preceding the destination index.
+A full example is illustrated in Figure \ref{fig:routing}.
 
@@
 
+\subsubsection{Algorithm Specification}
+
 We now begin the formal specification of our multipath routing scheme for the
 wrap-around butterfly network.
+For convenience, the relevant variables are summarized in Table \ref{tab:routing}.
 Utilizing vertex transitivity, we label the source node as
@@
 \end{description}
+
+\subsubsection{Proof of Path Independence}
 
@@
 The {\em directory authorities} used by the Tor
-\cite{dingledine_tor_2004} network are one example.
+\cite{dingledine_tor:_2004} network are one example.
 However, the size of such authority networks is often limited by the number of

### 2017-03-16 websci-2017/source/paper.tex
Final pass.

@@
 We show how the assumption of partial trust transitivity,
-while more realistic than the assumption used by webs of trust,
+while more realistic than the assumption underlying webs of trust,
 can be used to quantify the effective redundancy of a network
@@
 paths to utilize that redundancy.
-When network structure can be dictated
+When portions of network structure can be dictated
 our results can be used to create scalable,
@@
 Much of the world's infrastructure is networked: power grids,
-cellular networks, raods, and of course, the Internet.
+cellular networks, roads, and of course, the Internet.
 As with any critical infrastructure, the cost of failures can be
@@
 in order to protect their users from surveillance.
+Once again, the attack was successful due to a highly centralized
+architecture:
+SSL keys under control of a single entity, in a single legal jurisdiction.
 While originally intended as surveillance,
 this action effectively became an act of censorship.
-In this case, the attack was again successful due to a highly centralized
-architecture:
-SSL keys under control of a single entity, in a single legal jurisdiction.
-It is also important to note that there was no technological failure,
-the attack was achieved through coercion.
+It is also important to note that while Lavabit's cryptography worked as intended,
+the attack was still successful because the system was
+vulnerable to non-technical coercion.
 So we see that such vulnerabilities are not limited to any one system or protocol,
@@
 (rather than disrupting the network as a whole).
-This assumption arises because scalable network architectures can be made
+This assumption applies to scalable network architectures that can be made
 large enough that an attacker must focus their resources in order to have
@@
 those servers.
-However, there may be ways to achieve improved attack-tolerance from architectures
+Additionally, there may be ways to achieve improved attack-tolerance from architectures
 more flexible than the butterfly, which is a potential area for future work.
@@
 {\em adversarial} faults in {\em complex networks}.
-We do so by modelling redundant independent paths
+We do so by modeling redundant independent paths
 with partial trust transitivity as a single virtual channel,
@@
 algorithm for the butterfly network topology
-Section \ref{sec-discussion} discusses the our results.
+Section \ref{sec-discussion} discusses the results.
 And Section \ref{sec-conclusion} concludes.
@@
 \cite{reiter_resilient_1998}.
-Our work complements these by introducing the partial trust assumption and
-explicitly quantifying its relationship to network structure and adversarial
-fault tolerance.
+Our work complements these by introducing the partial trust assumption
+extending the focus beyond authentication.
 When network topology can be controlled, we sidestep the NP-hard problem of finding
@@
 More recently, peer-to-peer systems based on distributed hash tables have used
-structured ``overlay'' networks to map table keys to local TCP/IP routes
+structured {\em overlay networks} to map table keys to local TCP/IP routes
 \cite{lua_survey_2005,korzun_structured_2013}.
@@
 properties,
-which can be used to to guarantee attack-tolerance properties.
+which can be used to to improve attack-tolerance.
 
@@
 more favorably disposed to Alice's adversary, more likely to cooperate with that
-adversary, and less likely to take proactive security measures against that
+adversary, or less likely to take proactive security measures against that
 adversary.
@@
 and will be convenient for proving our results.
-We now proceed to define it formally.
+We now proceed to define our model formally.
 
@@
 \beq
-\delta_h(V) \equiv \min_{v,w \in V} \delta_{v,w,h}.
+\delta_h(G) \equiv \min_{v,w \in V} \delta_{v,w,h}.
 \eeq
@@
 We assume that Mal is capable of compromising $l \leq \delta_h$ channels.
-Having full knowledge of the network,
-Mal's best strategy is to identify a minimum node cut in the network
-and compromise nodes from that cut.
-With this strategy, each compromised node reduces effective redundancy by one,
-effectively compromising one of the virtual channels between Alice and Bob.
 Since Alice chooses channels randomly,
@@
 so Mal can do no better than also choosing randomly.
+We can also return to the full network setting by noting that each of the
+$\delta_h$ independent paths in the network can serve as independent channels
+between Alice and Bob.
+Mal's best strategy is now to identify a minimum node cut in the network
+and randomly compromise nodes from that cut.
+With this strategy, each compromised node reduces effective redundancy by one,
+equivalent to compromising one of the channels between Alice and Bob.
 If $k > l$, at least one message will get through uncompromised and all
@@
 and the strength of trust transitivity.
-Thus, the scheme can be effective, even when the number of copies
-sent $k$ is a small fraction of the effective redundancy.
+{\em Thus, the scheme can be effective, even when the number of copies
+sent $k$ is a small fraction of the effective redundancy}.
 In other words, this scheme exhibits a {\em stabilizing asymmetry}:
@@
 
-In order to derive the above results, we have assumed that Alice and all
+In order to derive the above results,
+we have assumed that Alice and all
 intermediary agents are able to identify specific,
@@
 Each cycle is divided into stages,
-with different properties used to prove independence in different stages
+with different properties used to prove independence at each stage
 (see Figure \ref{fig:route-overview}).
-In the first cycle (stages 2--4), path independence is guaranteed by ensuring that
+In the first cycle (stages 1--4), path independence is guaranteed by ensuring that
 all nodes match the path parameter $s$ in the first $h$ bits of the place-within-level.
-Similarly, in the second cycle (stages 5--6),
+Similarly, in the second cycle (stages 5--7),
 independence is guaranteed by ensuring that all paths match $s$ in the
@@
 and the FBI's surveillance-turned-censorship of Lavabit
-\cite{poulsen_edward_2013}
+\cite{poulsen_edward_2013}.
 The reader may wonder how our methods could be employed in scenarios
@@
 the world, synchronized through consensus protocols.
-The {\em directory authorities} used by the Tor
-\cite{dingledine_tor:_2004} network are one example.
+The {\em directory authorities} used by the Tor project
+\cite{dingledine_tor:_2004} are one example.
 However, the size of such authority networks is often limited by the number of
-trust relationships (degree) each node can maintain, and the inherent insecurity of
+trusted relationships (degree) each node can maintain, and the inherent insecurity of
 extending transitive trust to an ever-larger network.
@@
 It is also worth noting that because faulty paths can be identified in our scheme,
-it may at times be appropriate to assume mutual trust and revoke that relationship
-when it is violated.
+it may at times be appropriate to begin by assuming mutual trust and
+revoking that relationship if it is violated.
 Any entities dependent on the proper functioning of the network would have an incentive

### 2017-04-08 websci-2017/source/whitelabel.tex
Add preprint.

--- 2017-03-16 websci-2017/source/paper.tex
@@
-\documentclass{sig-alternate-05-2015}
-
+\documentclass[twocolumn]{article}
+
+\usepackage[top=0.75in,left=0.75in,right=0.75in,bottom=1in]{geometry}
+\setlength{\columnsep}{0.33in}
+
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{amsthm}
 \usepackage{epigraph}
@@
 
-% Copyright
-\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\setcopyright{usgov}
-%\setcopyright{usgovmixed}
-%\setcopyright{cagov}
-%\setcopyright{cagovmixed}
-
-% DOI
-%\doi{10.475/123_4}
-
-% ISBN
-%\isbn{123-4567-24-567/08/06}
-
-%Conference
-\conferenceinfo{WebSci '17}{June 26--8, 2017, Troy, NY, USA}
-
-%\acmPrice{\$15.00}
-
-%
-% --- Author Metadata here ---
-%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
-%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
-%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
-% --- End of Author Metadata ---
-
 \title{Towards Attack-Tolerant Networks: Concurrent Multipath Routing and the Butterfly Network}
-%
-% You need the command \numberofauthors to handle the 'placement
-% and alignment' of the authors beneath the title.
-%
-% For aesthetic reasons, we recommend 'three authors at a time'
-% i.e. three 'name/affiliation blocks' be placed beneath the title.
-%
-% NOTE: You are NOT restricted in how many 'rows' of
-% "name/affiliations" may appear. We just ask that you restrict
-% the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate'
-% we ask you to refrain from putting more than six authors
-% (two rows with three columns) beneath the article title.
-% More than six makes the first-page appear very cluttered indeed.
-%
-% Use the \alignauthor commands to handle the names
-% and affiliations for an 'aesthetic maximum' of six authors.
-% Add names, affiliations, addresses for
-% the seventh etc. author(s) as the argument for the
-% \additionalauthors command.
-% These 'additional authors' will be output/set for you
-% without further effort on your part as the last section in
-% the body of your article BEFORE References or any Appendices.
-
-\numberofauthors{2} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
-%
 \author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
-\alignauthor
 Edward L. Platt \\
-\affaddr{University of Michigan}\\
-\email{elplatt@umich.edu}
-\alignauthor
+University of Michigan\\
+elplatt@umich.edu
+\and
 Daniel M. Romero \\
-\affaddr{University of Michigan}\\
-\email{drom@umich.edu}
-}
-%       \affaddr{Institute for Clarity in Documentation}\\
-%       \affaddr{1932 Wallamaloo Lane}\\
-%       \affaddr{Wallamaloo, New Zealand}\\
-%       \email{trovato@corporation.com}
-% 2nd. author
-%\alignauthor
-%G.K.M. Tobin\titlenote{The secretary disavows
-%any knowledge of this author's actions.}\\
-%       \affaddr{Institute for Clarity in Documentation}\\
-%       \affaddr{P.O. Box 1212}\\
-%       \affaddr{Dublin, Ohio 43017-6221}\\
-%       \email{webmaster@marysville-ohio.com}
-% 3rd. author
+University of Michigan\\
+drom@umich.edu
+}
+\date{}
 
@@
 
-%
-% The code below should be generated by the tool at
-% http://dl.acm.org/ccs.cfm
-% Please copy and paste the code instead of the example below. 
-%
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10002978.10002986</concept_id>
-<concept_desc>Security and privacy~Formal methods and theory of security</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003033.10003034</concept_id>
-<concept_desc>Networks~Network architectures</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10011007.10010940.10010971.10010972.10010540</concept_id>
-<concept_desc>Software and its engineering~Peer-to-peer architectures</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Security and privacy~Formal methods and theory of security}
-\ccsdesc[500]{Networks~Network architectures}
-\ccsdesc[500]{Software and its engineering~Peer-to-peer architectures}
-
-%
-% End generated code
-%
-
-%
-%  Use this command to print the description
-%
-\printccsdesc
-
-% We no longer use \terms command
-%\terms{Theory}
-
-\keywords{
-Butterfly Topology;
-Fault Tolerance;
-Adversarial Faults;
-Multipath Routing;
-Censorship;
-Decentralization}
+\noindent{\bf Keywords:}
+Butterfly Topology,
+Fault Tolerance,
+Adversarial Faults,
+Multipath Routing,
+Censorship,
+Decentralization.
 
@@
 \section{Acknowledgments}
-
-%
-% The following two commands are all you need in the
-% initial runs of your .tex file to
-% produce the bibliography for the citations in your paper.
+The authors would like to thank Tony Garnock-Jones, A. Frederick Dudley, and
+Nathaniel Bezanson for helpful conversations.
+This research was partly supported by the National Science Foundation under Grant No. IIS-1617820.
+
 \bibliographystyle{abbrv}
 \bibliography{paper}  % sigproc.bib is the name of the Bibliography in this case
-% You must have a proper ".bib" file
-%  and remember to run:
-% latex bibtex latex latex
-% to resolve all references
-%
-% ACM needs 'a single self-contained file'!
-%
-%
 \end{document}

### 2017-10-23 websci-2017/source/whitelabel.tex
Add source for www2018

@@
 an indication that the network is decentralized.
-The efrective redundancy of the entire graph can be characterized by the minimum over
+The effective redundancy of the entire graph can be characterized by the minimum over
 all vertex pairs:
@@
 1. request retransmission;
-2. end receipts so Alice knows which paths have been compromised;
+2. send receipts so Alice knows which paths have been compromised;
 or 3. attempt error correction using majority voting.
@@
 structure can be imposed by the designer,
-and more generally, provide a theoretical foudnation that can be used more to evaluate
+and more generally, provide a theoretical foundation that can be used more to evaluate
 the role of network structure, trust transitivity, and effective redundancy

### 2017-10-23 www-2018/source/paper.tex
Add source for www2018

--- 2017-03-16 websci-2017/source/paper.tex
@@
-\documentclass{sig-alternate-05-2015}
-
+\documentclass[sigconf]{acmart}
+
+\usepackage{booktabs} % For formal tables
 \usepackage{epigraph}
@@
 
-%\usepackage{draftwatermark}
-%\SetWatermarkText{DRAFT}
-
-\begin{document}
-
 % Copyright
-\setcopyright{acmcopyright}
+%\setcopyright{none}
+%\setcopyright{acmcopyright}
 %\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
+\setcopyright{rightsretained}
 %\setcopyright{usgov}
@@
 % DOI
-%\doi{10.475/123_4}
+\acmDOI{10.475/123_4}
 
 % ISBN
-%\isbn{123-4567-24-567/08/06}
+\acmISBN{123-4567-24-567/08/06}
 
 %Conference
-\conferenceinfo{WebSci '17}{June 26--8, 2017, Troy, NY, USA}
-
-%\acmPrice{\$15.00}
-
-%
-% --- Author Metadata here ---
-%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
-%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
-%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
-% --- End of Author Metadata ---
-
-\title{Towards Attack-Tolerant Networks: Concurrent Multipath Routing and the Butterfly Network}
-%
-% You need the command \numberofauthors to handle the 'placement
-% and alignment' of the authors beneath the title.
-%
-% For aesthetic reasons, we recommend 'three authors at a time'
-% i.e. three 'name/affiliation blocks' be placed beneath the title.
-%
-% NOTE: You are NOT restricted in how many 'rows' of
-% "name/affiliations" may appear. We just ask that you restrict
-% the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate'
-% we ask you to refrain from putting more than six authors
-% (two rows with three columns) beneath the article title.
-% More than six makes the first-page appear very cluttered indeed.
-%
-% Use the \alignauthor commands to handle the names
-% and affiliations for an 'aesthetic maximum' of six authors.
-% Add names, affiliations, addresses for
-% the seventh etc. author(s) as the argument for the
-% \additionalauthors command.
-% These 'additional authors' will be output/set for you
-% without further effort on your part as the last section in
-% the body of your article BEFORE References or any Appendices.
-
-\numberofauthors{2} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
-%
-\author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
-\alignauthor
-Edward L. Platt \\
-\affaddr{University of Michigan}\\
+\acmConference[WEBCONF'18]{The Web Conference}{April 2018}{Lyon, France} 
+\acmYear{2018}
+\copyrightyear{2018}
+
+\acmArticle{4}
+\acmPrice{15.00}
+
+\begin{document}
+\title{
+    Towards Attack-Tolerant Networks:
+    Concurrent Multipath Routing and the Butterfly Network}
+
+\author{Edward L. Platt}
+\affiliation{%
+  \institution{University of Michigan\\School of Information}
+  \city{Ann Arbor} 
+  \state{Michigan} 
+}
 \email{elplatt@umich.edu}
-\alignauthor
-Daniel M. Romero \\
-\affaddr{University of Michigan}\\
+
+\author{Daniel M. Romero}
+\affiliation{%
+  \institution{University of Michigan\\School of Information}
+  \city{Ann Arbor} 
+  \state{Michigan} 
+}
 \email{drom@umich.edu}
-}
-%       \affaddr{Institute for Clarity in Documentation}\\
-%       \affaddr{1932 Wallamaloo Lane}\\
-%       \affaddr{Wallamaloo, New Zealand}\\
-%       \email{trovato@corporation.com}
-% 2nd. author
-%\alignauthor
-%G.K.M. Tobin\titlenote{The secretary disavows
-%any knowledge of this author's actions.}\\
-%       \affaddr{Institute for Clarity in Documentation}\\
-%       \affaddr{P.O. Box 1212}\\
-%       \affaddr{Dublin, Ohio 43017-6221}\\
-%       \email{webmaster@marysville-ohio.com}
-% 3rd. author
-
-\maketitle
+
+% The default list of authors is too long for headers.
+\renewcommand{\shortauthors}{B. Trovato et al.}
+
 \begin{abstract}
@@
 
-%
-% End generated code
-%
-
-%
-%  Use this command to print the description
-%
-\printccsdesc
-
-% We no longer use \terms command
-%\terms{Theory}
-
 \keywords{
-Butterfly Topology;
-Fault Tolerance;
-Adversarial Faults;
-Multipath Routing;
-Censorship;
+Adversarial Faults,
+Resilience,
+Butterfly Network,
+Fault Tolerance,
+Multipath Routing,
+Censorship,
 Decentralization}
+
+\maketitle
 
@@
 \section{Acknowledgments}
-
-%
-% The following two commands are all you need in the
-% initial runs of your .tex file to
-% produce the bibliography for the citations in your paper.
-\bibliographystyle{abbrv}
-\bibliography{paper}  % sigproc.bib is the name of the Bibliography in this case
-% You must have a proper ".bib" file
-%  and remember to run:
-% latex bibtex latex latex
-% to resolve all references
-%
-% ACM needs 'a single self-contained file'!
-%
-%
+The authors would like to thank Tony Garnock-Jones, A. Frederick Dudley, and
+Nathaniel Bezanson for helpful conversations.
+This research was partly supported by the National Science Foundation under Grant No. IIS-1617820.
+
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{paper}
+
 \end{document}

### 2017-10-26 www-2018/source/paper.tex
Update intro and abstract, move some sections.

@@
 \usetikzlibrary{arrows,chains,decorations.pathreplacing}
+\usepackage{pifont}% http://ctan.org/pkg/pifont
 
@@
 \newtheorem{theorem}{Theorem}
+
+\newcommand{\cmark}{\ding{51}}
+\newcommand{\xmark}{\ding{55}}
 
@@
 
-\author{Edward L. Platt}
-\affiliation{%
-  \institution{University of Michigan\\School of Information}
-  \city{Ann Arbor} 
-  \state{Michigan} 
-}
-\email{elplatt@umich.edu}
-
-\author{Daniel M. Romero}
-\affiliation{%
-  \institution{University of Michigan\\School of Information}
-  \city{Ann Arbor} 
-  \state{Michigan} 
-}
-\email{drom@umich.edu}
+%\author{Edward L. Platt}
+%\affiliation{%
+%  \institution{University of Michigan\\School of Information}
+%  \city{Ann Arbor} 
+%  \state{Michigan} 
+%}
+%\email{elplatt@umich.edu}
+
+%\author{Daniel M. Romero}
+%\affiliation{%
+%  \institution{University of Michigan\\School of Information}
+%  \city{Ann Arbor} 
+%  \state{Michigan} 
+%}
+%\email{drom@umich.edu}
+\author{Anonymized\\for\\blind review}
 
 % The default list of authors is too long for headers.
-\renewcommand{\shortauthors}{B. Trovato et al.}
+%\renewcommand{\shortauthors}{E.L. Platt and D.M. Romero.}
+\renewcommand{\shortauthors}{Anonymized for blind review}
 
 \begin{abstract}
+
+It is desirable for critical large-scale communication networks
+such as the internet to be resilient against any types of disruption
+\cite{sterbenz_resilience_2010}.
+Vulnerabilities in communication infrastructure can be used by
+state, corporate, and private actors to achieve censorship and surveillance,
+to the detriment of free expression and free association
+\cite{}.
+Self-organized networks such as the internet's router network typically have
+heavy-tailed degree distributions
+/cite{barabasi_scale-free_2009},
+making them resilent against random faults,
+but highly vulnerable to targeted attacks against high-centrality nodes
+\cite{albert_error_2000}.
+Cryptographic solutions fail to address the underlying topological problem,
+and remain vulnerable to man-in-the-middle attacks
+\cite{nayak_different_2010}
+and coercion
+\cite{gruel_caveat_2013}.
+
+1 Main result
+We present a novel concurrent multipath routing (CMR) algorithm for the wraparound
+butterfly network topology,
+as well as a highly attack-tolerant
+Structured Multipath Fault Tolerance (SMFT) architechture
+which incorporaes the butterfly CMR algorithm.
+
+2-3 Main result vs previous knowledge
+
+1-2 General context
+
+2-3 Broader perspective
+
 Targeted attacks against network infrastructure are notoriously difficult
@@
 suggesting the importance of a structural approach to attack-tolerance.
-We show how the assumption of partial trust transitivity,
+We show how the assumption of bounded trust transitivity,
 while more realistic than the assumption underlying webs of trust,
@@
 
+Is it possible for any large-scale communication network to resist targeted attacks?
+The internet was originally designed to withstand targeted (nuclear) attacks \cite{baran_distributed_1964},
+and the resilience of the internet has long been part of common wisdom \cite{elmer-dewitt_first_1993}.
+But 17 years after Albert et. al \cite{albert_error_2000}
+showed the internet's router network to be vulnerable to targeted attacks (vs. random faults),
+the fundamental problem remains unsolved.
+The ongoing vulnerability of the internet is evidenced by a long history of 
+censorship and surveillance incidents achieved by means of targeted attacks
+\cite{dainotti_analysis_2011}.
+In this paper, we present a theoretical infrastructure for networked communication,
+that tolerates not only random faults, but also targeted attacks.
+While not yet practical, our proposal is the first demonstration that
+attack-tolerant communication networks are even possible.
+
 Much of the world's infrastructure is networked: power grids,
@@
 We call this assumption
-{\em partial trust transitivity},
-and refer to such paths as {\em partially trusted}.
+{\em bounded trust transitivity}.
 
@@
 We begin by describing how fault tolerance techniques can be adapted and
-evaluated in a network setting with partial trust transitivity and
+evaluated in a network setting with bounded trust transitivity and
 adversarial faults.
@@
 We do so by modeling redundant independent paths
-with partial trust transitivity as a single virtual channel,
+with bounded trust transitivity as a single virtual channel,
 and show that the probability of detecting adversarial faults
@@
 }
-\item{We prove that the number of independent paths between two nodes
-in a wrap-around butterfly network with partial trust transitivity
-increases exponentially with the trust radius;
+\item{We prove that the number of $h$-internal vertex disjoint
+paths between two nodes in a wrap-around butterfly network
+increases exponentially with $h$;
 }
@@
 structured networks.
-Section \ref{sec-butterfly} describes the concurrent multipath routing
-algorithm for the butterfly network topology
+Section \ref{sec-butterfly} gives background on the butterfly network topology.
+Section \ref{sec-bf-routing} presents our concurrent multipath routing
+algorithm for the butterfly.
 Section \ref{sec-discussion} discusses the results.
@@
 Our work addresses both of these limitations by incorporating a more realistic
-assumption of partial transitivity.
+assumption of bounded trust transitivity.
 
@@
 \cite{reiter_resilient_1998}.
-Our work complements these by introducing the partial trust assumption
+Our work complements these by introducing the bounded trust transitivity assumption
 extending the focus beyond authentication.
@@
 
+\begin{table*}[t!]
+\caption{Comparison of attack-tolerant network communication architectures.\label{tab:compare}}
+\begin{center}
+\begin{tabular}{l|ccccc}
+                              & Decentralized & Trust-based & Bounded-trust & Topological & Point-to-point \\
+\hline
+Public Key Infrastructure     &               & \cmark      & \cmark        &             & \cmark     \\
+Web of Trust                  & \cmark        & \cmark      &               &             & \cmark     \\
+Freenet                       & \cmark        & \cmark      & \cmark        &             &            \\
+Federated Byzantine Agreement & \cmark        & \cmark      & \cmark        &             & \cmark     \\
+Blockchain                    & \cmark        &             & \cmark        &             &            \\
+Fiat-Saia                     & \cmark        & \cmark      & \cmark        & \cmark      &            \\
+Structured Multipath FT       & \cmark        & \cmark      & \cmark        & \cmark      & \cmark     \\
+\hline
+\end{tabular}
+\end{center}
+\end{table*}
+
 \section{Trust Networks and Fault Tolerance}
@@
 which are chosen strategically to maximize the likelihood of a failure.
-
-\subsection{Partial Trust Model}
-
-A central question in large-scale, secure communication is this:
-how can two parties communicate reliably and securely
-when no direct trusted link exists between them?
-The commonly-used web of trust approach
-\cite{zimmermann_official_1995,ferguson_practical_2003}
-extends trust infinitely transitively:
-to friends of friends, and friends of friends of friends, and so on.
-However, the assumption of infinitely transitive trust is unrealistic
-\cite{christianson_why_1997},
-and does not allow for the analysis of the effects of network structure.
-
-An alternative assumption might be that each hop away from Alice in
-in the network
-reduces the probability that a node can resist compromise exponentially.
-Such a situation could occur if nodes more distant from Alice are
-more favorably disposed to Alice's adversary, more likely to cooperate with that
-adversary, or less likely to take proactive security measures against that
-adversary.
-The above model can be further
-approximated by assuming that nodes up to some fixed number
-of hops cannot be compromised, and that those beyond can.
-This simplified version is still more realistic than infinite transitivity
-and will be convenient for proving our results.
-We now proceed to define our model formally.
-
-We define the {\em partial trust model} on
-an undirected graph $G = (V,E)$,
-although the model can easily be extended to directed multigraphs.
-Vertices representing commiunicating agents,
-and with edges representing mutually trusted communication links.
-Let $v \in V$ be an arbitrary sender (Alice)
-and $w \in V$ be an arbitrary receiver (Bob).
-We assume the presence of an adversary (Mal) who knows the
-full structure of the network,
-and who can compromise a fixed number of nodes,
-gaining complete control of their behavior.
-We also assume that Mal is an adversary of Alice and/or Bob specifically,
-rather than the network as a whole.
-So Mal can compromise any node except for those trusted by Alice or Bob.
-We define a {\em trust radius} $h$ such that nodes $u$ and
-$u^\prime$ trust each other if their distance is less than $h$.
-For a given node $u$,
-we call the set of trusted nodes its
-{\em trusted neighborhood} $T_h(u)$,
-and all nodes at exactly distance $h$ the
-{\em trust boundary} $B_h(u)$:
-\beq
-T_h(u) &=& \left\{ u^\prime \mid d(u,u^\prime) < h \right\} \\
-B_h(u) &=& \left\{ u^\prime \mid d(u,u^\prime) = h \right\}.
-\eeq
-The trust boundary $B_h$ plays an important role because these nodes are not
-trusted by $u$,
-and if compromised can entirely isolate $u$ from the rest of the network.
-These trust assumptions imply that when Alice sends a message to Bob,
-Mal can only cause faults in the set of nodes outside both of their trusted
-neighborhoods: $V \setminus \left(T_h(v) \cup T_h(w)\right)$.
-We refer to this set of nodes as the {\em untrusted region}.
-
-\subsection{Effective Redundancy}
-
-Our goal is to achieve fault tolerance through redundancy.
-In the network setting, redundnacy is achieved using
-{\em independent paths} \cite{reiter_resilient_1998},
-which have no common points of failure.
-Typically, it is assumed that paths must be disjoint in order to be
-independent.
-However, under the partial trust assumption, two non-disjoint paths
-can still be independent as long as their intersection contains only trusted
-nodes,
-greatly increasing the level of redundancy available.
-Under the partial trust assumption, the available redundancy thus depends on
-both the network structure and the level of trust.
-
-We now quantify the {\em effective redundancy} between Alice and Bob
-when trust radius $h$ is assumed.
-This quantity, $\delta_{v,w,h}$ is exactly the max-flow/min-cut of
-the graph after Alice's and Bob's trusted neighborhoods have been
-collapsed into single source/sink vertices.
-Each trust boundary forms a cut of the network and places an upper bound on the
-min-cut:
-\beq
-\delta_{v,w,h} \leq \min\left( \mid B_h(v) \mid, \mid B_h(w) \mid \right).
-\eeq
-Equality holds when there are no bottlenecks within the untrusted region,
-an indication that the network is decentralized.
-The effective redundancy of the entire graph can be characterized by the minimum over
-all vertex pairs:
-\beq
-\delta_h(G) \equiv \min_{v,w \in V} \delta_{v,w,h}.
-\eeq
-Thus, for any pair of nodes in the network, at least $\delta_h$ independent,
-redundant paths can be constructed between them.
-$\delta_h$ is a purely structural network property,
-and places an upper bound on the effectiveness of any
-redundancy-based fault tolerance scheme.
-The more quickly $\delta_h$ grows with $h$,
-the better a network is at leveraging trust transitivity to create redundancy.
-Thus, the scaling of $\delta_h$ can be used to quantify a network's ability
-to withstand targeted attacks,
-even when the exact trust radius $h$ is unknown.
-
-\begin{figure}
-\centerline{\includegraphics[width=3in,height=2.08in]{fig-partial-trust}}
-\caption{
-Illustration of a trusted communication network and the network properties
-used by the {\em partial trust model}.
-Edges represent mutually trusted communication links.
-The sender (Alice, $v$) and receiver (Bob, $w$) trust all nodes
-less than the {\em trust radius} $h$ hops away.
-These nodes form their {\em trusted neighborhoods} $T_h(v)$ and $T_h(w)$.
-We assume that all faults occur in the remaining nodes: the
-{\em untrusted region}.
-The untrusted nodes in contact with the trusted neighborhoods for the
-{\em trust boundaries} $B_h(v)$ and $B_h(w)$,
-which (in the absence of central bottlenecks) determine the
-{\em effective redundancy} $\delta_h$ provided by the network.
-Alice and Bob can be modelled as connected by a direct link with
-at least $\delta_h$ redundant channels.
-}
-\label{fig:trust-source}
-\end{figure}
 
@@
 
-\section{Multipath Butterfly Routing}
+\subsection{Partial Trust Model}
+
+A central question in large-scale, secure communication is this:
+how can two parties communicate reliably and securely
+when no direct trusted link exists between them?
+The commonly-used web of trust approach
+\cite{zimmermann_official_1995,ferguson_practical_2003}
+extends trust infinitely transitively:
+to friends of friends, and friends of friends of friends, and so on.
+However, the assumption of infinitely transitive trust is unrealistic
+\cite{christianson_why_1997},
+and does not allow for the analysis of the effects of network structure.
+
+An alternative assumption might be that each hop away from Alice in
+in the network
+reduces the probability that a node can resist compromise exponentially.
+Such a situation could occur if nodes more distant from Alice are
+more favorably disposed to Alice's adversary, more likely to cooperate with that
+adversary, or less likely to take proactive security measures against that
+adversary.
+The above model can be further
+approximated by assuming that nodes up to some fixed number
+of hops cannot be compromised, and that those beyond can.
+This simplified version is still more realistic than infinite transitivity
+and will be convenient for proving our results.
+We now proceed to define our model formally.
+
+We define the {\em partial trust model} on
+an undirected graph $G = (V,E)$,
+although the model can easily be extended to directed multigraphs.
+Vertices representing commiunicating agents,
+and with edges representing mutually trusted communication links.
+Let $v \in V$ be an arbitrary sender (Alice)
+and $w \in V$ be an arbitrary receiver (Bob).
+We assume the presence of an adversary (Mal) who knows the
+full structure of the network,
+and who can compromise a fixed number of nodes,
+gaining complete control of their behavior.
+We also assume that Mal is an adversary of Alice and/or Bob specifically,
+rather than the network as a whole.
+So Mal can compromise any node except for those trusted by Alice or Bob.
+We define a {\em trust radius} $h$ such that nodes $u$ and
+$u^\prime$ trust each other if their distance is less than $h$.
+For a given node $u$,
+we call the set of trusted nodes its
+{\em trusted neighborhood} $T_h(u)$,
+and all nodes at exactly distance $h$ the
+{\em trust boundary} $B_h(u)$:
+\beq
+T_h(u) &=& \left\{ u^\prime \mid d(u,u^\prime) < h \right\} \\
+B_h(u) &=& \left\{ u^\prime \mid d(u,u^\prime) = h \right\}.
+\eeq
+The trust boundary $B_h$ plays an important role because these nodes are not
+trusted by $u$,
+and if compromised can entirely isolate $u$ from the rest of the network.
+These trust assumptions imply that when Alice sends a message to Bob,
+Mal can only cause faults in the set of nodes outside both of their trusted
+neighborhoods: $V \setminus \left(T_h(v) \cup T_h(w)\right)$.
+We refer to this set of nodes as the {\em untrusted region}.
+
+\subsection{Effective Redundancy}
+
+Our goal is to achieve fault tolerance through redundancy.
+In the network setting, redundnacy is achieved using
+{\em independent paths} \cite{reiter_resilient_1998},
+which have no common points of failure.
+Typically, it is assumed that paths must be disjoint in order to be
+independent.
+However, under the bounded trust transitivity assumption, two non-disjoint paths
+can still be independent as long as their intersection contains only trusted
+nodes,
+greatly increasing the level of redundancy available.
+Under the bounded trust transitivity assumption, the available redundancy thus depends on
+both the network structure and the level of trust.
+
+We now quantify the {\em effective redundancy} between Alice and Bob
+when trust radius $h$ is assumed.
+This quantity, $\delta_{v,w,h}$ is exactly the max-flow/min-cut of
+the graph after Alice's and Bob's trusted neighborhoods have been
+collapsed into single source/sink vertices.
+Each trust boundary forms a cut of the network and places an upper bound on the
+min-cut:
+\beq
+\delta_{v,w,h} \leq \min\left( \mid B_h(v) \mid, \mid B_h(w) \mid \right).
+\eeq
+Equality holds when there are no bottlenecks within the untrusted region,
+an indication that the network is decentralized.
+The effective redundancy of the entire graph can be characterized by the minimum over
+all vertex pairs:
+\beq
+\delta_h(G) \equiv \min_{v,w \in V} \delta_{v,w,h}.
+\eeq
+Thus, for any pair of nodes in the network, at least $\delta_h$ independent,
+redundant paths can be constructed between them.
+$\delta_h$ is a purely structural network property,
+and places an upper bound on the effectiveness of any
+redundancy-based fault tolerance scheme.
+The more quickly $\delta_h$ grows with $h$,
+the better a network is at leveraging trust transitivity to create redundancy.
+Thus, the scaling of $\delta_h$ can be used to quantify a network's ability
+to withstand targeted attacks,
+even when the exact trust radius $h$ is unknown.
+
+\begin{figure}
+\centerline{\includegraphics[width=3in,height=2.08in]{fig-partial-trust}}
+\caption{
+Illustration of a trusted communication network and the network properties
+used by the {\em partial trust model}.
+Edges represent mutually trusted communication links.
+The sender (Alice, $v$) and receiver (Bob, $w$) trust all nodes
+less than the {\em trust radius} $h$ hops away.
+These nodes form their {\em trusted neighborhoods} $T_h(v)$ and $T_h(w)$.
+We assume that all faults occur in the remaining nodes: the
+{\em untrusted region}.
+The untrusted nodes in contact with the trusted neighborhoods for the
+{\em trust boundaries} $B_h(v)$ and $B_h(w)$,
+which (in the absence of central bottlenecks) determine the
+{\em effective redundancy} $\delta_h$ provided by the network.
+Alice and Bob can be modelled as connected by a direct link with
+at least $\delta_h$ redundant channels.
+}
+\label{fig:trust-source}
+\end{figure}
+
+\section{The Butterfly Network Topology}
 \label{sec-butterfly}
@@
 level until the destination node is reached.
+
+\section{Multipath Butterfly Routing}
+\label{sec-bf-route}
 
@@
 utilizing a novel concurrent multipath routing algorithm.
-We have also demonstrated how {\em partial trust transitivity},
+We have also demonstrated how {\em bounded trust transitivity},
 in addition to being more realistic than infinite transitivity,
@@
 for a large increase in the number of messages compromised by an adversary.
-We have also demonstrated how the assumption of partial trust
+We have also demonstrated how the assumption of bounded trust
 transitivity can enable a quantitative analysis of the
@@
 \section{Acknowledgments}
-The authors would like to thank Tony Garnock-Jones, A. Frederick Dudley, and
-Nathaniel Bezanson for helpful conversations.
-This research was partly supported by the National Science Foundation under Grant No. IIS-1617820.
+Anonymized for blind review
+%The authors would like to thank Tony Garnock-Jones, A. Frederick Dudley, and
+%Nathaniel Bezanson for helpful conversations.
+%This research was partly supported by the National Science Foundation under Grant No. IIS-1617820.
 

### 2017-10-26 www-2018/source/paper.tex
Finish draft of new abs.

@@
 \begin{abstract}
-
 It is desirable for critical large-scale communication networks
@@
 heavy-tailed degree distributions
-/cite{barabasi_scale-free_2009},
-making them resilent against random faults,
+\cite{barabasi_scale-free_2009},
+making them resilient against random faults,
 but highly vulnerable to targeted attacks against high-centrality nodes
@@
 and coercion
-\cite{gruel_caveat_2013}.
-
-1 Main result
+\cite{grewal_caveat_2013}.
+Coercion-resistant, topological approaches to attack tolerance are needed
+to address the current vulnerability of communications infrastructure to
+targeted censorship and surveillance.
 We present a novel concurrent multipath routing (CMR) algorithm for the wraparound
@@
 as well as a highly attack-tolerant
-Structured Multipath Fault Tolerance (SMFT) architechture
-which incorporaes the butterfly CMR algorithm.
-
-2-3 Main result vs previous knowledge
-
-1-2 General context
-
-2-3 Broader perspective
-
-Targeted attacks against network infrastructure are notoriously difficult
-to guard against.
-In the case of communication networks,
-such attacks can leave users vulnerable to
-censorship and surveillance,
-even when cryptography is used.
-Much of the existing work on network fault-tolerance focuses on random faults and
-does not apply to adversarial faults (attacks).
-Centralized networks have single points of failure by definition,
-leading to a growing popularity in decentralized architectures and protocols
-for greater fault-tolerance.
-However, centralized network {\em structure} can arise even when {\em protocols} are
-decentralized.
-Despite their decentralized protocols,
-the Internet and World-Wide Web have
-been shown both theoretically and historically to be highly susceptible to
-attack, in part due to emergent structural centralization.
-When single points of failure exist,
-they are potentially vulnerable to non-technological
-(i.e., coercive) attacks,
-suggesting the importance of a structural approach to attack-tolerance.
-We show how the assumption of bounded trust transitivity,
-while more realistic than the assumption underlying webs of trust,
-can be used to quantify the effective redundancy of a network
-as a function of trust transitivity.
-We also prove that the effective redundancy of the wrap-around butterfly topology
-increases exponentially with trust transitivity and
-describe a novel concurrent multipath routing algorithm for constructing
-paths to utilize that redundancy.
-When portions of network structure can be dictated
-our results can be used to create scalable,
-attack-tolerant infrastructures.
-More generally, our results provide a theoretical formalism for evaluating the effects
-of network structure on adversarial fault-tolerance.
+Structured Multipath Fault Tolerance (SMFT) architecture
+which incorporates the butterfly CMR algorithm.
+Our architecture is the first theoretical demonstration of a
+coercion-resistant point-to-point communication architecture
+that does not require infinitely transitive trust.
+We also present for modeling attack-tolerance as a function of
+trust transitivity,
+and show how the internet's router network could be rewired for greater
+attack-tolerance.
+We identify a previously unexplored relationship between
+network topology, trust transitivity, and attack-tolerance,
+and provide a framework for further exploration of this relationship.
+Our results are the first we're aware of to demonstrate
+that large-scale communication networks can be
+designed to tolerate targeted attacks,
+including coercion and other non-technical attacks.
+Furthermore, if attack-tolerance requirements are strict enough to justify
+a high level of control over network topology, our results could be applied
+directly.
+More broadly, we hope that this work will serve as a starting point for the
+development of additional topology-based attack-tolerant
+communication architectures to guard
+against the dangers of censorship and surveillance.
 \end{abstract}

### 2017-10-27 www-2018/source/paper.tex
Intro edits.

@@
 \begin{abstract}
-It is desirable for critical large-scale communication networks
-such as the internet to be resilient against any types of disruption
-\cite{sterbenz_resilience_2010}.
-Vulnerabilities in communication infrastructure can be used by
-state, corporate, and private actors to achieve censorship and surveillance,
-to the detriment of free expression and free association
-\cite{}.
-Self-organized networks such as the internet's router network typically have
-heavy-tailed degree distributions
-\cite{barabasi_scale-free_2009},
-making them resilient against random faults,
-but highly vulnerable to targeted attacks against high-centrality nodes
+It is crucial for large-scale communication networks such as the internet
+to be resilient against attacks\cite{sterbenz_resilience_2010}
+such as censorship and surveillance,
+which pose a threat to free expression and free association.
+Self-organized networks such as the internet’s router network typically have
+heavy-tailed degree distributions\cite{barabasi_scale-free_2009},
+making them highly vulnerable to targeted attacks against central nodes
 \cite{albert_error_2000}.
-Cryptographic solutions fail to address the underlying topological problem,
-and remain vulnerable to man-in-the-middle attacks
-\cite{nayak_different_2010}
-and coercion
-\cite{grewal_caveat_2013}.
-Coercion-resistant, topological approaches to attack tolerance are needed
-to address the current vulnerability of communications infrastructure to
-targeted censorship and surveillance.
-We present a novel concurrent multipath routing (CMR) algorithm for the wraparound
-butterfly network topology,
-as well as a highly attack-tolerant
-Structured Multipath Fault Tolerance (SMFT) architecture
-which incorporates the butterfly CMR algorithm.
-Our architecture is the first theoretical demonstration of a
-coercion-resistant point-to-point communication architecture
-that does not require infinitely transitive trust.
-We also present for modeling attack-tolerance as a function of
-trust transitivity,
-and show how the internet's router network could be rewired for greater
-attack-tolerance.
-We identify a previously unexplored relationship between
-network topology, trust transitivity, and attack-tolerance,
-and provide a framework for further exploration of this relationship.
-Our results are the first we're aware of to demonstrate
-that large-scale communication networks can be
-designed to tolerate targeted attacks,
-including coercion and other non-technical attacks.
-Furthermore, if attack-tolerance requirements are strict enough to justify
-a high level of control over network topology, our results could be applied
-directly.
+While cryptographic solutions exist,
+they fail to address the underlying topological problem,
+and remain vulnerable to
+man-in-the-middle attacks \cite{nayak_different_2010}
+and coercion \cite{grewal_caveat_2013}.
+Coercion-resistant, topological approaches to attack tolerance are needed to
+address the current vulnerability of communications infrastructure to
+censorship and surveillance.
+We present a novel concurrent multipath routing (CMR) algorithm for the
+wraparound butterfly network topology,
+as well as a highly attack-tolerant Structured Multipath Fault Tolerance (SMFT)
+architecture which incorporates the butterfly CMR algorithm. We also identify a
+previously unexplored relationship between network topology, trust transitivity,
+and attack-tolerance, and provide a framework for further exploration of this
+relationship. Our work is the first theoretical demonstration of a point-to-point
+communication network architecture that can resist coercion and other
+non-technical attacks, without requiring infinitely transitive trust.
+To address cases where the network structure cannot be fully controlled,
+we demonstrate how a snapshot of the internet’s router network can be partially
+rewired for greater attack-tolerance.
 More broadly, we hope that this work will serve as a starting point for the
-development of additional topology-based attack-tolerant
-communication architectures to guard
-against the dangers of censorship and surveillance.
+development of additional topology-based attack-tolerant communication
+architectures to guard against the dangers of censorship and surveillance.
 \end{abstract}
@@
 
-Is it possible for any large-scale communication network to resist targeted attacks?
-The internet was originally designed to withstand targeted (nuclear) attacks \cite{baran_distributed_1964},
-and the resilience of the internet has long been part of common wisdom \cite{elmer-dewitt_first_1993}.
+Is it possible for any large-scale communication network to resist targeted
+attacks?
+The internet was originally designed to withstand targeted (nuclear) attacks
+\cite{baran_distributed_1964},
+and the resilience of the internet has long been part of common wisdom
+\cite{elmer-dewitt_first_1993}.
 But 17 years after Albert et. al \cite{albert_error_2000}
-showed the internet's router network to be vulnerable to targeted attacks (vs. random faults),
-the fundamental problem remains unsolved.
+showed the internet's router network to be vulnerable to targeted attacks
+(vs. random faults), the fundamental problem remains unsolved.
 The ongoing vulnerability of the internet is evidenced by a long history of 
@@
 \cite{dainotti_analysis_2011}.
-In this paper, we present a theoretical infrastructure for networked communication,
-that tolerates not only random faults, but also targeted attacks.
-While not yet practical, our proposal is the first demonstration that
-attack-tolerant communication networks are even possible.
-
-Much of the world's infrastructure is networked: power grids,
-cellular networks, roads, and of course, the Internet.
-As with any critical infrastructure, the cost of failures can be
-immense, so methods for tolerating various kind of faults within networks are an
+In this paper, we present the first theoretical architecture
+for point-to-point networked communication,
+that tolerates not only random faults, but also targeted attacks,
+without relying on infinitely transitive trust
+\cite{christianson_why_1997}.
+
+Methods for tolerating various kind of faults within networks are an
 important and ongoing area of research
@@
 But despite longstanding common wisdom \cite{elmer-dewitt_first_1993},
-both theoretical results and recent events (described below)
-have demonstrated that the Internet can be surprisingly vulnerable to attack.
-Decentralization remains a promising approach towards
-building resilient networks,
-but there is a need to better understand the relationship between
-decentralized network structure and adversarial fault tolerance.
-
-The Internet's vulnerability to censorship and other targeted attacks
+both theoretical results and recent events
+have demonstrated that the internet is surprisingly vulnerable to attack.
+
+Analysis of the internet's router network has shown that while it
+is remarkably resilient against random faults,
+it is highly susceptible to adversarial faults \cite{albert_error_2000}.
+These results have been attributed to the heavy-tailed
+degree distribution of the Internet's router network
+\cite{barabasi_emergence_1999,barabasi_scale-free_2009}.
+Random failures are highly likely to affect only low-degree nodes,
+thus having little effect.
+However, adversarial faults target the few high-degree nodes,
+and therefore remove a large number of edges with each fault.
+So while the {\em protocols} of the Internet are decentralized,
+the {\em network structure} is somewhat centralized. 
+In other words, the protocols of the Internet do not {\em require}
+centralization, but centralization may still emerge from the sociotechnical
+processes that create its network structure.
+
+The internet's vulnerability to censorship and other targeted attacks
 has been demonstrated by several recent events.
@@
 the ability of an attacker to succeed without even trying only highlights
-the Internet's vulnerability to adversarial faults.
+the internet's vulnerability to adversarial faults.
 
@@
 
-Analysis of the Internet's router network has shown that while it
-is remarkably resilient against random faults,
-it is highly susceptible to adversarial faults \cite{albert_error_2000}.
-These results have been attributed to the scale-free structure of the Internet's
-router network
-\cite{barabasi_emergence_1999,barabasi_scale-free_2009}.
-In scale-free networks and other networks with heavy-tail degree distributions,
-random failures are highly likely to
-affect only low-degree nodes, thus having
-little effect.
-However, Adversarial faults target the few high-degree nodes,
-and therefore remove a large number of edges with each fault.
-So while the {\em protocols} of the Internet are decentralized,
-the {\em network structure} is somewhat centralized. 
-In other words, the protocols of the Internet do not {\em require}
-centralization, but centralization may still emerge from the sociotechnical
-processes that create its network structure.
-
-With strong theoretical and historical evidence that centralized
-network structure can create vulnerabilities,
-methods for analyzing structural vulnerabilities and for designing
-fault-tolerant networks are needed.
+Coercion-resistant, topological approaches to attack tolerance are needed to
+address the current vulnerability of communications infrastructure
+to censorship and surveillance.
 This paper presents several contributions towards advancing those goals.
-While our motivation comes from the Internet router network and
-World Wide Web,
-our work is theoretical, focusing on abstract networks,
+While our motivation by the network of internet routers,
+our work is theoretical,
 and could potentially be applied to many different types of networks,
@@
 
-We consider a setting in which a source node (Alice)
-in a network attempts to route a message
-to a target node (Bob) by forwarding it through the links of the network.
-A ``link'' in this context could represent any kind of connection
-(e.g., physical cables, encrypted channels).
-We assume that some nodes in the network may be compromised by an attacker
-(Mal).
-We also assume that Mal is an adversary of Alice specifically and targets
-nodes strategically with the goal of interfering with Alice's communications
-(rather than disrupting the network as a whole).
-This assumption applies to scalable network architectures that can be made
-large enough that an attacker must focus their resources in order to have
-a significant effect.
-Compromised nodes may behave incorrectly by blocking, altering,
-or incorrectly routing messages.
-We assume that Mal has full knowledge of the network structure, but has
-limited resources and thus can only compromise a fixed number of nodes.
-
-We also assume that nodes {\em trust} their immediate neighbors.
-So Mal is unable to compromise Alice's node, or her direct neighbors.
-In the commonly used {\em web of trust} approach
-\cite{zimmermann_official_1995,ferguson_practical_2003},
-we would extend that trust transitively to the entire network.
-However, we make a weaker and more realistic assumption:
-that trust is extended transitively to nodes within a fixed
-number of hops.
-So only the beginning and end of a path between Alice and Bob is trusted
-against interference from Mal.
-We call this assumption
-{\em bounded trust transitivity}.
+We consider a setting in which a source node attempts to route a message
+to a target node, while an adversary attempts to block or intercept the message
+by compromising a number of intermediate nodes.
+We also assume that the source and destination nodes trust their neighbors
+i.e., that the adversary is unable to compromise them,
+as in the web of trust approach
+\cite{zimmermann_official_1995,ferguson_practical_2003}.
+Unlike the web of trust approach, which assumes that trust is infinitely
+transitive,
+we assume {\em bounded trust transitivity}.
 
@@
 We have already seen that whenever a single point of failure exists in the network,
-there is a potential for an attacker to exploit it through coercion,
-without needing to compromise the technology.
+there is a potential for an attacker to exploit it through coercion.
 So, {\em attack-tolerance cannot be achieved without the ability to influence

### 2017-10-30 www-2018/source/paper.tex
Updates to multipath fault tolerance.

@@
 This paper presents several contributions towards advancing those goals.
-While our motivation by the network of internet routers,
-our work is theoretical,
-and could potentially be applied to many different types of networks,
+While our work is motivated by the network of internet routers,
+the results are theoretical,
+and can potentially be applied to many different types of networks,
 whether made of physical wires, virtual tunnels,
@@
 we show how to evaluate the influence of network structure
-on attack-tolerance,
-how to use local trust and redundancy to achieve greater attack-tolerance
-when no single path is fully trusted,
-and propose a novel routing algorithm for constructing such paths on
-the butterfly network topology.
+on attack-tolerance.
+We next present a structured multipath fault tolerance (SMFT) to extend
+standard fault tolerance techniques to the case of adversarial faults in
+networks
+\cite{avizienis_basic_2004, von_neumann_probabilistic_1956}.
+The SMFT scheme requires the existence of a
+concurrent multipath routing (CMR) algorithm
+\cite{zin_survey_2015, qadir_exploiting_2015, khiani_comparative_2013},
+to takes advantage of the independence of faults
+along {\em independent paths}.
+We also present a novel CMR algorithm for the butterfly network topology.
 The butterfly topology is popular in parallel processing
@@
 those servers.
-Additionally, there may be ways to achieve improved attack-tolerance from architectures
-more flexible than the butterfly, which is a potential area for future work.
-
-We begin by describing how fault tolerance techniques can be adapted and
-evaluated in a network setting with bounded trust transitivity and
-adversarial faults.
-Generally, faults in network paths can be correlated,
-preventing the application of standard fault tolerance techniques
-\cite{avizienis_basic_2004, von_neumann_probabilistic_1956},
-which assume independent faults.
-By constructing {\em independent paths},
-which have no untrusted nodes in common,
-we show how to model communication across a complex network in the presence
-of correlated adversarial faults as communication across redundant
-simple channels with random errors.
-Redundant messages can be sent across these channels in parallel,
-a technique known as {\em concurrent multipath routing}
-\cite{zin_survey_2015, qadir_exploiting_2015, khiani_comparative_2013},
-and used for fault tolerance.
-The receiver can then use the redundant messages to detect and/or correct
-errors.
-We formally evaluate the
-effects of network structure on attack-tolerance and show that the probability
-of an undetected error decreases exponentially with the number of
-independent paths between source and destination,
-even when no individual path is entirely trusted.
-
-We also propose a novel concurrent multipath routing algorithm for the butterfly
-topology.
-The algorithm constructs independent paths,
-which can be combined with the fault-tolerant concurrent multipath routing
-scheme above to
-achieve a high level of adversarial fault tolerance on the butterfly topology.
 
@@
 \item{
-We propose a novel method for extending standard fault tolerance techniques to
+We propose a novel structured multipath fault tolerance (SMFT) scheme
+for extending standard fault tolerance techniques to
 {\em adversarial} faults in {\em complex networks}.
-We do so by modeling redundant independent paths
-with bounded trust transitivity as a single virtual channel,
+We do so by modeling redundant {\em independent paths}
+with {\em bounded trust transitivity} as a single virtual channel,
 and show that the probability of detecting adversarial faults
@@
 paths between two nodes in a wrap-around butterfly network
-increases exponentially with $h$;
-}
-\item{We present a scalable, efficient, and attack-tolerant concurrent
-multipath routing algorithm on the butterfly network topology.}
+increases exponentially with $h$,
+and thus that the number of independent paths in a network
+increases exponentially with the level of trust transitivity;
+}
+\item{We present a scalable and efficient concurrent
+multipath routing (CMR) algorithm on the butterfly network topology,
+which can be combined with SMFT to achieve a high level of attack tolerance.}
 \end{itemize}
@@
 Section \ref{sec-butterfly} gives background on the butterfly network topology.
-Section \ref{sec-bf-routing} presents our concurrent multipath routing
+Section \ref{sec-bf-route} presents our concurrent multipath routing
 algorithm for the butterfly.
@@
 
-There has been considerable work on trust in network security.
-Both centralized and decentralized approaches are commonly used to create
-trust infrastructures.
+There has been considerable work on trust-based attack-tolerance techniques
+in network security, both centralized and decentralized.
 Centralized approaches such as {\em public key infrastructure} (PKI)
 suffer from a number of vulnerabilities
-\cite{ellison_ten_2000},
-which stem largely from the single points of failure inherent to
+\cite{ellison_ten_2000}, including vulnerability to coercion,
+which stems largely from the single points of failure inherent to
 centralization.
@@
 Trust is then extended to new individuals if they are vouched for by a
-currently-trusted individual,
-making it possible to quickly establish a large group of trusted nodes.
-However, web of trust's assumption of infinite transitivity is unrealistic
-\cite{christianson_why_1997},
-and does not distinguish between paths of different lengths.
-Our work addresses both of these limitations by incorporating a more realistic
-assumption of bounded trust transitivity.
-
-Previous work on incorporating network structure into
-trust models has focused on authentication
-protocols, showing that independent paths can reduce an adversary's ability
+currently-trusted individual.
+Infinite trust transitivity is helpful for establishing a large group of
+trusted nodes, but unfortunately unrealistic
+\cite{christianson_why_1997}.
+Additionally, the web of trust approach
+does not distinguish between paths of different lengths.
+Our work addresses both of these limitations by requiring only
+bounded trust transitivity.
+
+Previous work applying network topology to attack tolerance has
+focused on authentication,
+showing that independent paths can reduce an adversary's ability
 to impersonate a target
@@
 \cite{reiter_resilient_1998}.
-Our work complements these by introducing the bounded trust transitivity assumption
-extending the focus beyond authentication.
+Our work complements these results by extending our focus beyond authentication,
+to communication.
 When network topology can be controlled, we sidestep the NP-hard problem of finding
@@
 some circumstances, but are limited to small networks due to poor scalability.
-Proof-of-work \cite{dwork_pricing_1993,nakamoto_bitcoin:_2008}
-and proof-of-stake \cite{king_ppcoin:_2012}
-provide better scalability,
-but are wasteful of computational and energy resources.
+Proof-of-work \cite{dwork_pricing_1993,nakamoto_bitcoin:_2008} (blockchain)
+systems provide better scalability,
+but are wasteful of computational and energy resources,
+and do not take advantage of trusted relationships.
 Federated Byzantine Agreement (FBA) \cite{mazieres_stellar_2015}
 is scalable, allows for flexible trust,
-and is highly fault-tolerant on networks meeting a set of requirements.
-However, FBA does not provide a method for evaluating the
-fault tolerance properties of different network structures
+and is highly fault-tolerant on networks meeting specific requirements.
+However, FBA does not provide a method for constructing networks to meet
+those requirements,
 or for calculating the failure probabilities within a particular network.
 
-There are relatively few attack-tolerance schemes
-that focus on network structure,
-compared to more popular cryptographic approaches
-\cite{ferguson_practical_2003}.
 All existing attack-tolerant networks we are aware of are content-addressable
-networks: data is routed to and from storage nodes rather than between sender
-and receiver.
+networks (CANs) in which data is stored and retrieved based on key values,
+rather than point-to-point networks, in which data is communicated between
+two parties.
 Fiat and Saia described a scheme that combines the butterfly topology
@@
 as data travels from its origin to the redundant storage locations.
-Unlike the above content-addressable networks, our proposal is purely network based
+Unlike the above content-addressable networks, our architecture is purely network based
 and does not require nodes to store data indefinitely.
-Our proposal also improves on the scalablity of Fiat and Saia's work,
-and does not rely on assumptions about existing social network structure.
+Our architecture also improves on the scalablity of the Fiat-Saia network,
+and makes requirements about network topology explicit.
 
@@
 with the network edges constrained by sensors' physical location.
-We have found only few examples in the existing literature of applications of
-concurrent multipath routing to {\em adversarial} fault tolerance,
+We have found very few examples of CMR applied to {\em adversarial}
+fault tolerance in the existing literature,
 and all have focused on ad-hoc wireless sensor networks, without attention
@@
 
+Our proposed architecture is differentiated from existing systems by several
+properties (Table \ref{tab:compare}).
+Decentralized architectures are more resistent to coercion
+\cite{grewal_caveat_2013}
+and man-in-the-middle attacks \cite{nayak_different_2010}.
+Trust-based systems are more sustainable than proof-of-work.
+Bounded-trust systems do not require the unrealistic assumption of infinite
+trust transitivity.
+Topological approaches address the root cause of vulnerability
+in heavy-tail networks,
+rather than relying on technology that can be side-stepped through coercion.
+Point-to-point communication allows two individuals to exchange messages
+without requiring large amounts of indefinite data storage on intermediate
+nodes.
+
 \begin{table*}[t!]
@@
 Federated Byzantine Agreement & \cmark        & \cmark      & \cmark        &             & \cmark     \\
-Blockchain                    & \cmark        &             & \cmark        &             &            \\
+Proof of Work/Blockchain      & \cmark        &             & \cmark        &             &            \\
 Fiat-Saia                     & \cmark        & \cmark      & \cmark        & \cmark      &            \\
@@
 We will make use of standard fault tolerance terminology, summarized here.
-A {\em fault} is said to occur when one component
+A {\em fault} is occurs when one component
 of a system behaves incorrectly (e.g., a routing node blocks or
@@
 The result of that fault (e.g., a recipient receiving conflicting messages)
-is called an {\em error} state.
-If the error is undetected or corrected to the wrong value, the system is
-said to have experienced a {\em failure} (e.g., an altered message is
+is an {\em error} state.
+If the error is undetected or corrected to the wrong value,
+the system has experienced a {\em failure} (e.g., an altered message is
 accepted as authentic).
@@
 
-Once we have determined a network's effective redundancy,
-we can apply redundancy-based fault tolerance techniques,
-by sending multiple copies of a message
-({\em concurrent multipath routing}).
-Having found the effective redundancy between two nodes,
-we can simplify our model, replacing independent paths through the
-complex network with direct channels between the endpoints.
-We model our sender (Alice) and receiver (Bob) as
-communicating over $\delta_h$ direct and redundant virtual channels.
-The partial trust model allows us to make this simplifying assumption
-for analyzing a fault tolerance scheme,
-but implementing such a scheme will require a method for constructing
-specific network paths.
-We will return to the question of constructing paths in the next section.
+Standard fault tolerance methods use redundancy to detect and correct
+statistically indpendent faults.
+In complex networks however, faults can be correlated when,
+for example, two messages pass through the same faulty node.
+For now, let us assume our sender (Alice) and receiver (Bob) are connected
+by $\delta_h$ direct channels, with indpendent errors.
+We will return to the question of constructing these channels in subsequent
+sections.
 For now, we concern ourselves with the question:
@@
 what is the probability that an adversary (Mal) causes an undetectable
-error after inducing a fixed number of faults?
+error after causing faults in a fixed number of channels?
 
@@
 so Mal can do no better than also choosing randomly.
-We can also return to the full network setting by noting that each of the
-$\delta_h$ independent paths in the network can serve as independent channels
-between Alice and Bob.
-Mal's best strategy is now to identify a minimum node cut in the network
-and randomly compromise nodes from that cut.
-With this strategy, each compromised node reduces effective redundancy by one,
-equivalent to compromising one of the channels between Alice and Bob.
 If $k > l$, at least one message will get through uncompromised and all
@@
 depends on the fractions of
-redundant paths actually utilized $\alpha$ and compromised $\beta$,
-it decreases exponentially with the effective redundancy $\delta_h$
-(which we will later see increases exponentially in $h$ in the
-butterfly topology).
-This result is significant because $\delta_h$
+channels actually utilized $\alpha$ and compromised $\beta$,
+it decreases exponentially with $\delta_h$.
+This result is significant because,
+as we will soon show, $\delta_h$
 depends only on the network structure
 and the strength of trust transitivity.
-{\em Thus, the scheme can be effective, even when the number of copies
-sent $k$ is a small fraction of the effective redundancy}.
+{\em Thus, the scheme can be effective, even when the number of channels used
+$k$ is a small fraction of the channels available}
 In other words, this scheme exhibits a {\em stabilizing asymmetry}:
@@
 adversaries,
-as long as the network structure provides large $\delta_h$.
-
-In order to derive the above results,
-we have assumed that Alice and all
-intermediary agents are able to identify specific,
-independent network paths that achieve the effective redundancy $\delta_h$.
-We now proceed to describe a routing algorithm for doing so in the special case
-of the butterfly network topology.
+as long as the network provides large $\delta_h$.
 
@@
 
-\subsection{Partial Trust Model}
-
-A central question in large-scale, secure communication is this:
-how can two parties communicate reliably and securely
-when no direct trusted link exists between them?
-The commonly-used web of trust approach
-\cite{zimmermann_official_1995,ferguson_practical_2003}
+\subsection{Bounded Trust Model}
+
+So far, we have assumed that Alice and Bob have access to some number
+$\delta_h$ of channels with statistically indpendent faults.
+However, in real communication architectures, direct links between all
+pairs of individuals are intractable and messages must be routed through
+a number of intermediate nodes.
+In an adversarial setting,
+the existence of intermediate nodes introduces two problems:
+1. intermediate nodes may be compromised by the adversary and
+2. faults on paths are no longer uncorrelated: two paths may pass through the
+same compromised node.
+We show how to how to resolve these problems using a combination of network
+structure and bounded trust transitivity.
+
+Trust-based approaches to secure communication assume that some parties
+cannot be compromised.
+One common approach, the web of trust
+\cite{zimmermann_official_1995,ferguson_practical_2003},
 extends trust infinitely transitively:
-to friends of friends, and friends of friends of friends, and so on.
+Alice trusts her friends, as well as her friends' friends, and so on.
 However, the assumption of infinitely transitive trust is unrealistic
-\cite{christianson_why_1997},
-and does not allow for the analysis of the effects of network structure.
+\cite{christianson_why_1997}.
+Furthermore, infinite trust transitivity does not allow for analysis of
+network strucutre; all that matters is whether some path exists,
+rather than the number of paths and their structure.
 
@@
 The above model can be further
-approximated by assuming that nodes up to some fixed number
+simplified by assuming that nodes up to some fixed number
 of hops cannot be compromised, and that those beyond can.
@@
 
-We define the {\em partial trust model} on
+We define the {\em bounded trust model} on
 an undirected graph $G = (V,E)$,
 although the model can easily be extended to directed multigraphs.
-Vertices representing commiunicating agents,
-and with edges representing mutually trusted communication links.
-Let $v \in V$ be an arbitrary sender (Alice)
-and $w \in V$ be an arbitrary receiver (Bob).
-We assume the presence of an adversary (Mal) who knows the
+Vertices representing commiunicating parties,
+with edges representing mutually trusted communication links.
+We define a {\em trust radius} $h$ such that nodes $v$ and
+$w$ trust each other if their distance is less than $h$.
+For a given node $v$,
+we call the set of trusted nodes its
+{\em trusted neighborhood} $T_h(v)$,
+and all nodes at exactly distance $h$ the
+{\em trust boundary} $B_h(v)$:
+\beq
+T_h(v) &=& \left\{ w \mid d(v,w) < h \right\} \\
+B_h(v) &=& \left\{ w \mid d(v,w) = h \right\}.
+\eeq
+The trust boundary $B_h$ plays an important role because these nodes are not
+trusted by $u$,
+and if compromised can entirely isolate $v$ from the rest of the network.
+
+Now let $s \in V$ be an arbitrary sender
+and $t \in V$ be an arbitrary receiver.
+We assume the presence of an adversary who knows the
 full structure of the network,
@@
 gaining complete control of their behavior.
-We also assume that Mal is an adversary of Alice and/or Bob specifically,
-rather than the network as a whole.
-So Mal can compromise any node except for those trusted by Alice or Bob.
-We define a {\em trust radius} $h$ such that nodes $u$ and
-$u^\prime$ trust each other if their distance is less than $h$.
-For a given node $u$,
-we call the set of trusted nodes its
-{\em trusted neighborhood} $T_h(u)$,
-and all nodes at exactly distance $h$ the
-{\em trust boundary} $B_h(u)$:
-\beq
-T_h(u) &=& \left\{ u^\prime \mid d(u,u^\prime) < h \right\} \\
-B_h(u) &=& \left\{ u^\prime \mid d(u,u^\prime) = h \right\}.
-\eeq
-The trust boundary $B_h$ plays an important role because these nodes are not
-trusted by $u$,
-and if compromised can entirely isolate $u$ from the rest of the network.
-These trust assumptions imply that when Alice sends a message to Bob,
-Mal can only cause faults in the set of nodes outside both of their trusted
-neighborhoods: $V \setminus \left(T_h(v) \cup T_h(w)\right)$.
+We also assume that the adversary is specifically targeting communicaiton
+between $s$ and $t$ and
+can compromise any node except for those trusted by $s$ or $t$.
+Under these trust assumptions, adversarial faults can only occur
+outside the trusted neighborhoods of $s$ and $t$:
+$V \setminus \left(T_h(v) \cup T_h(w)\right)$.
 We refer to this set of nodes as the {\em untrusted region}.
@@
 
-In previous sections, we showed that reliable communication across a network
-can be achieved even when any single message path might be compromised by
-an adversary,
-provided the network has sufficient redundancy,
-and provided the sender and intermediaries know how to route message
-copies along independent paths.
-In this section, we address both requirements by proposing a novel routing
-algorithm for constructing independent paths on the butterfly network topology.
-This architecture and routing algorithm achieve an
-effective redundancy that increases exponentially with the trust radius,
-allowing a very high level of adversarial fault tolerance.
-
+In order to implement structured multipath fault tolerance,
+we need a strucutred network topology with high effective redundancy.
+In this paper, we apply SMFT to the butterfly network topology
+\cite{kshemkalyani_distributed_2008}.
 The structure of the butterfly network is highly constrained,
@@
 but rather a reflection of the inherent difficulty of attack-tolerance.
-Lastly, we note that the partial trust model and multipath fault tolerance
-schemes of the previous section do not rely on any particular network
-topology or routing algorithm,
-and our choice of the butterfly topology is only one of many possible choices.
+To address the case when the network cannot be fully controlled,
+we show how partially rewiring a snapshot of the internet's router
+network can greatly increase it's effective redundancy and
+attack-tolerance properties.
 
@@
 While several variations on the butterfly network exist,
-we utilize the wrap-around butterfly.
-We denote the $m$-dimensional, directed wrap-around butterfly as a
-graph $\wbf(m)$:
+we utilize the $m$-dimensional, directed wrap-around butterfly
+(Figure \ref{fig:bf-route}),
+denoted $\wbf(m)$:
 \beq
@@
 Multiple paths exist between any two nodes.
-Specifically, we will prove below that the number of independent paths between two
-nodes increases exponentially with the trust
-radius $h$.
+Specifically, we will prove below that the number of
+$h$-internally vertex disjoint paths between two
+nodes increases exponentially with $h$.
 \end{description}
+
+\begin{figure}
+\begin{center}
+\includegraphics[height=3in,angle=90]{fig-bf-route.pdf}
+\end{center}
+\caption{
+A 3-dimensional wraparound butterfly network.
+Note that the rightmost nodes are the same nodes as the leftmost,
+drawn twice for visual clarity.
+The highlighted nodes and edges show the path from node (0,000)
+to node (1,101).
+\label{fig:bf-route}
+}
+\end{figure}
 
@@
 The structure of the butterfly network lends itself to a well-known
-(unipath) routing algorithm,
+(unipath) routing algorithm (Figure \ref{fig:bf-route}),
 which we later extend to the multipath case.
@@
 
+\subsection{Butterfly Rewiring}
+
+\begin{figure*}[h!]
+\centerline{\includegraphics[width=7in,height=1.75in]{fig-mincut}}
+\caption{
+Simulation of targeted attacks against a snapshot of the internet's router
+network with a fraction of the edges rewired into a partial butterfly configuration.
+The effective redundancy is shown for several values of trust transitivity $h$.
+For $h>0$, the original network has higher effective redundancy up to a crossover point,
+after which the rewired network performs better.
+}
+\label{fig:mincut}
+\end{figure*}
+
+Even when a butterfly topology cannot be implemented perfectly,
+it can still increase the attack tolerance properties of a network.
+Here, we simulate targeted attacks against a snapshot of the internet's
+router network on January 2, 2000
+\cite{leskovec_graphs_2005}.
+However, we equate a subset of the nodes with the nodes of a 9-dimensional
+butterfly network and rewire a fraction of the edges to match the butterfly
+topology.
+At each step of the simulation, betweenness centrality is recalculated and the
+most central node is removed.
+
+Our simulations show improved resistance to fragmentation and higher
+effective redundancy when even a fraction of edges have been rewired to
+match the butterfly topology.
+While the original router network fragments when about 1\% of the nodes
+have been removed (Figure \ref{fig:diameter}),
+this number increases to 2\%
+with only 10% of the edges rewired to match the butterfly.
+With 90\% of the edges rewired,
+the network remains unfragmented beyond the failure of the 8\% most
+central nodes.
+The effective redundancy for varius values of trust transitivity $h$
+are shown in Figure \ref{fig:mincut}.
+The effective redundancy is calculated by collapsing nodes and edges within
+$h$ hops of source and destination into single nodes and finding the mincut
+between them,
+averaging over 150 source-destination pairs.
+For all but $h=0$,
+the original network has higher redundancy for a small fraction of
+node failures,
+while the rewired network has higher redundancy beyond a crossover point.
+We interpret these results to suggest that even when a small number of
+highly central nodes have been removed from a heavy-tailed network,
+most nodes are still able to take advantage of the remaining hubs.
+As larger hubs continue to be removed, the connectivity of the network
+decreases until the crossover point,
+at which point the rewired network offers higher effective redundancy.
+
+\begin{figure}
+\centerline{\includegraphics[width=3in,height=2.25in]{fig-diameter}}
+\caption{
+Simulation of targeted attacks against a snapshot of the internet's router
+network with a fraction of the edges rewired into a partial butterfly configuration.
+The original network fragments at about 1\%.
+With 90\% of the edges rewired,
+the network remains unfragmented beyond the failure of the 8\% most
+central nodes.
+}
+\label{fig:diameter}
+\end{figure}
+
 \section{Multipath Butterfly Routing}
 \label{sec-bf-route}
-
-\subsection{Multipath Routing Algorithm}
 
@@
 
-\subsubsection{Algorithm Specification}
+\subsection{Algorithm Specification}
 
@@
 
-\subsubsection{Proof of Path Independence}
+\subsection{Proof of Path Independence}
 

### 2017-10-31 www-2018/source/paper.tex
Update discussion and conclusion.

@@
 }
-\item{We prove that the number of $h$-internal vertex disjoint
+\item{We prove that the number of $h$-internally vertex disjoint
 paths between two nodes in a wrap-around butterfly network
@@
 For now, let us assume our sender (Alice) and receiver (Bob) are connected
-by $\delta_h$ direct channels, with indpendent errors.
+by $\delta$ direct channels, with indpendent errors.
 We will return to the question of constructing these channels in subsequent
@@
 For now, we concern ourselves with the question:
-given that the network provides $\delta_h$ redundant channels between
+given that the network provides $\delta$ redundant channels between
 Alice and Bob,
@@
 In a more sophisticated multipath fault tolerance scheme,
-Alice randomly chooses $k \leq \delta_h$ channels and sends a copy of
+Alice randomly chooses $k \leq \delta$ channels and sends a copy of
 her message on each.
-We assume that Mal is capable of compromising $l \leq \delta_h$ channels.
+We assume that Mal is capable of compromising $l \leq \delta$ channels.
 Since Alice chooses channels randomly,
@@
 \label{eq:pf}
-p_f &=& \frac{l!(\delta_h-k)!}{\delta_h!(l-k)!}.
+p_f &=& \frac{l!(\delta-k)!}{\delta!(l-k)!}.
 \eeq
-Letting $k=\alpha \delta_h$ and $l=\beta \delta_h$, then applying Stirling's
+Letting $k=\alpha \delta$ and $l=\beta \delta$, then applying Stirling's
 approximation gives:
@@
     (1-\alpha)
-\right]^{\delta_h}.
+\right]^{\delta}.
 \end{eqnarray}
@@
 channels actually utilized $\alpha$ and compromised $\beta$,
-it decreases exponentially with $\delta_h$.
+it decreases exponentially with $\delta$.
 This result is significant because,
-as we will soon show, $\delta_h$
+as we will soon show, $\delta$
 depends only on the network structure
@@
 {\em Thus, the scheme can be effective, even when the number of channels used
-$k$ is a small fraction of the channels available}
-In other words, this scheme exhibits a {\em stabilizing asymmetry}:
+$k$ is a small fraction of the channels available}.
+In other words, this scheme exhibits a {\em stabilizing asymmetry:}
 senders can tolerate attacks from significantly more powerful
 adversaries,
-as long as the network provides large $\delta_h$.
+as long as the network provides large $\delta$.
 
@@
 The probability of an undetectable error as a function of the number of
-message copies and the number of adversarial faults.
+redundant channels and the number of adversarial faults.
 }
@@
 So far, we have assumed that Alice and Bob have access to some number
-$\delta_h$ of channels with statistically indpendent faults.
+$\delta$ of channels with statistically indpendent faults.
 However, in real communication architectures, direct links between all
-pairs of individuals are intractable and messages must be routed through
+pairs of individuals not possible and messages must be routed through
 a number of intermediate nodes.
@@
 1. intermediate nodes may be compromised by the adversary and
-2. faults on paths are no longer uncorrelated: two paths may pass through the
-same compromised node.
+2. faults on paths are no longer statistically independent:
+two paths may pass through the same compromised node.
 We show how to how to resolve these problems using a combination of network
@@
 \cite{christianson_why_1997}.
-Furthermore, infinite trust transitivity does not allow for analysis of
-network strucutre; all that matters is whether some path exists,
-rather than the number of paths and their structure.
+Furthermore, infinite trust transitivity obscures the importance of network structure,
+as it depends only on whether some path exists, not the number or quality of paths.
 
 An alternative assumption might be that each hop away from Alice in
-in the network
-reduces the probability that a node can resist compromise exponentially.
+in the network reduces the probability that a node can resist compromise
+by a multiplicatve constant.
 Such a situation could occur if nodes more distant from Alice are
@@
 adversary.
-The above model can be further
-simplified by assuming that nodes up to some fixed number
-of hops cannot be compromised, and that those beyond can.
+An even simpler version assumes that nodes up to some
+fixed number of hops cannot be compromised, and that those beyond can.
 This simplified version is still more realistic than infinite transitivity
@@
 We refer to this set of nodes as the {\em untrusted region}.
+We now show how it is possible to communicate reliably,
+even when all available paths go through the untrusted region.
 
@@
 
-Our goal is to achieve fault tolerance through redundancy.
-In the network setting, redundnacy is achieved using
+Our approach is to achieve fault tolerance through redundancy.
+To do so, we must use only
 {\em independent paths} \cite{reiter_resilient_1998},
 which have no common points of failure.
-Typically, it is assumed that paths must be disjoint in order to be
-independent.
-However, under the bounded trust transitivity assumption, two non-disjoint paths
-can still be independent as long as their intersection contains only trusted
-nodes,
-greatly increasing the level of redundancy available.
-Under the bounded trust transitivity assumption, the available redundancy thus depends on
-both the network structure and the level of trust.
-
-We now quantify the {\em effective redundancy} between Alice and Bob
-when trust radius $h$ is assumed.
-This quantity, $\delta_{v,w,h}$ is exactly the max-flow/min-cut of
-the graph after Alice's and Bob's trusted neighborhoods have been
-collapsed into single source/sink vertices.
-Each trust boundary forms a cut of the network and places an upper bound on the
+Typically, it is assumed that in order to be independent,
+paths must be internally vertex disjoint, i.e.,
+have no nodes in common except the endpoints.
+However, under the bounded trust model,
+intersecting paths can still be independent if their
+intersection contains only trusted nodes.
+We define two paths with common endpoints
+to be {\em $h$-internally vertex disjoint}
+if all common vertices are less than distance $h$ from
+one of the endpoints.
+This condition holds if and only if two paths are indpendent
+under the bounded trust model with radius $h$.
+
+When trust radius $h$ is assumed,
+the number of $h$-internally vertex disjoint paths between
+two nodes $s$ and $t$ represents the number of
+channels that can be constructed between them having
+statistically indpendent faults.
+We thus refer to this quantity as the 
+{\em effective redundancy} $\delta_{s,t,h}$.
+The effective redundancy can also be interpreted as the
+max-flow/min-cut of a graph after each trusted neighborhood has been
+collapsed into a single vertex.
+The trust boundaries form a cut of the network and place an upper bound on the
 min-cut:
 \beq
-\delta_{v,w,h} \leq \min\left( \mid B_h(v) \mid, \mid B_h(w) \mid \right).
+\delta_{v,w,h} \leq \min\left( \mid B_h(s) \mid, \mid B_h(t) \mid \right).
 \eeq
@@
 an indication that the network is decentralized.
-The effective redundancy of the entire graph can be characterized by the minimum over
-all vertex pairs:
+The effective redundancy of the entire graph can be characterized by the minimum
+over all vertex pairs:
 \beq
-\delta_h(G) \equiv \min_{v,w \in V} \delta_{v,w,h}.
+\delta_h(G) \equiv \min_{s,t \in V} \delta_{s,t,h}.
 \eeq
@@
 redundant paths can be constructed between them.
-$\delta_h$ is a purely structural network property,
-and places an upper bound on the effectiveness of any
-redundancy-based fault tolerance scheme.
 The more quickly $\delta_h$ grows with $h$,
@@
 Illustration of a trusted communication network and the network properties
-used by the {\em partial trust model}.
+used by the {\em bounded trust model}.
 Edges represent mutually trusted communication links.
-The sender (Alice, $v$) and receiver (Bob, $w$) trust all nodes
+The sender (Alice, $s$) and receiver (Bob, $t$) trust all nodes
 less than the {\em trust radius} $h$ hops away.
@@
 {\em effective redundancy} $\delta_h$ provided by the network.
-Alice and Bob can be modelled as connected by a direct link with
-at least $\delta_h$ redundant channels.
+Alice and Bob can achieve the same level of attack-tolerance
+as if they were directly connected by $\delta_h$ redundant channels.
 }
@@
 \end{figure}
+
+\subsection{Structured Multipath Fault Tolerance}
+
+Finding a maximal set of independent paths for an abitrary network is NP hard
+\cite{reiter_resilient_1998},
+posing a challenge for multipath fault tolerance.
+We propose side-stepping this problem by using structured networks,
+for which independent paths can be generated efficiently.
+We call this approach {\em structured multipath fault tolerance} (SMFT),
+and now proceed to show how it is implemented on the butterfly
+network topology.
 
@@
 \draw[<->] ([yshift=1.5mm] A-4.north east) -- node[above=0.25mm] {$m - l_w$} ([yshift=1.5mm] A-4.north west);
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {start} ( A-1.west);
-\node (B1) [inner sep=1pt,above=of A-1.north,above=5ex] {\underline{A}};
-\node (B2) [inner sep=1pt,above=of A-2.north,above=5ex] {\underline{B}};
-\node (B3) [inner sep=1pt,above=of A-3.north,above=5ex] {\underline{C}};
-\node (B4) [inner sep=1pt,above=of A-4.north,above=5ex] {\underline{D}};
+\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {\underline{Stage}} ( A-1.west);
 \end{tikzpicture}
@@
                 minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick},
+                outer sep=0pt, on chain, thick, fill=blue!20!white},
                         ]
@@
 Given an $m$-bit wrap-around butterfly network ($m > 1$),
-and a radius $h$ ($1 \leq h \leq \left\lfloor \frac{m}{2} \right\rfloor$),
+and an integer $h$ ($1 \leq h \leq \left\lfloor \frac{m}{2} \right\rfloor$),
 for all node pairs $(v, w)$ such that $d(v,w) \geq 2h$,
-there exist $2^h$ paths $v_s$ ($0 \le s < 2^h$) from
+there exist at least $2^h$ $h$-internally vertex disjoint
+paths $v_s$ ($0 \le s < 2^h$) from
 $v$ to $w$ such that
@@
 
-While decentralized protocols have received much attention for their potential
-fault tolerance applications,
-centralized structures are always vulnerable to exploitation by non-technical means (i.e., coercion),
-and there is a need for a better understanding of the relationship between network
-structure and attack-tolerance.
-We have proposed a network-based scheme for {\em adversarial} fault tolerance on the
-butterfly topology,
-utilizing a novel concurrent multipath routing algorithm.
-We have also demonstrated how {\em bounded trust transitivity},
-in addition to being more realistic than infinite transitivity,
-provides a theoretical foundation for
-quantitative analysis of the relationship between trust,
-network structure, and attack tolerance.
-
-Such attacks include many forms of censorship and surveillance,
-which have important social implications.
+Our work has been motivated by the vulnerability of current communications
+infrastructure to surveillance and censorship,
+which are often achieved by coercive targeted attacks against central nodes.
 We have already discussed two such cases:
@@
 The reader may wonder how our methods could be employed in scenarios
-such as large-scale state-sponsored censorship.
+such as large-scale state-sponsored censorship
+\cite{xu_internet_2011}.
 Censorship-resistant infrastructure often replaces central servers
@@
 \cite{dingledine_tor:_2004} are one example.
-However, the size of such authority networks is often limited by the number of
+However, the size of such networks is limited by the number of
 trusted relationships (degree) each node can maintain, and the inherent insecurity of
 extending transitive trust to an ever-larger network.
-Our work fills a much-needed gap by quantifying the connection between network-structure,
-trust transitivity,
-and attack-tolerance.
-We provide both a theoretical framework and specific example of how network structure
+Our work provides both a theoretical framework
+and a specific example of how network structure
 can be engineered to leverage trust for a high level of attack-tolerance,
@@
 
-Fault-tolerant network infrastructures have many direct applications.
-Areas such as cryptocurrency
-\cite{mazieres_stellar_2015,nakamoto_bitcoin:_2008,king_ppcoin:_2012},
-secure multiparty computation
-\cite{yao_protocols_1982,chaum_multiparty_1988,goldreich_how_1987},
-and wireless sensor networks
-\cite{khiani_comparative_2013}
-have immediate need for scalable, fault-tolerant infrastructures.
-Many Internet services (e.g., email, social networks, cloud storage)
-are still highly centralized and vulnerable to technical and
-non-technical (i.e., coercive) attacks.
-Fault tolerance using {\em both}
-decentralized protocols and
-decentralized network structures
-is one promising approach to securing these services.
-
 We have focused primarily on adversarial faults that block or
-change messages (e.g., censorship).
-Existing cryptographic techniques for circumventing surveillance
-are relatively mature compared to those for tolerating censorship.
-However, the techniques presented in this paper
-are entirely compatible with, and in some cases could enhance, existing
-anti-surveillance techniques.
-For example, {\em man-in-the-middle} attacks exploit a privileged
-network position to attack otherwise secure cryptography,
-suggesting that structural approaches can complement cryptographic ones.
-
-While our present proposal is specific to the butterfly topology,
-the multipath fault tolerance scheme could
-be applied to any network that has both sufficient redundancy and a
-routing algorithm to discover independent paths.
-For general networks, finding all such paths is NP-hard,
-but efficient, suboptimal algorithms exist
-\cite{reiter_resilient_1998}.
-However, we have argued that attack-tolerance requires the ability to
-influence network structure and reduce reilance on single points of failure.
-Our work is most applicable to cases where the need for attack-tolerance
-justifies investment in deliberate infrastructure.
-For example, a coalition of groups supporting free expression could use our work
-to construct a censorship-resistant communication network.
-In general, such groups would need to invest resources into vetting their neighbors
-to establish trust, but there are scenarios in which the attack-tolerance
-requirement would justify that investment.
-It is also worth noting that because faulty paths can be identified in our scheme,
-it may at times be appropriate to begin by assuming mutual trust and
-revoking that relationship if it is violated.
-Any entities dependent on the proper functioning of the network would have an incentive
-to resist attack in order to maintain their ability to participate.
-
-Our work suggests several directions for future work towards developing
-practical, attack-tolerant communication infrastructure.
-The development of new multipath routing algorithms on other structured
-networks could achieve higher levels of redundancy.
-It is also desirable to identify dynamics that give
-rise to structured networks,
-and to evaluate whether our results can be generalized to unstructured or
-approximately structured networks.
-Finally, these results could be implemented to address specific
-applications, e.g., secure messaging, domain name resolution, or
-anonymous web browsing.
+change messages (censorship) but our work is also relevant to
+surveillance.
+While cryptographic anti-surveillance techniques exist,
+they remain vulnerable to man-in-the-middle attacks,
+in which an intermediate node masquerades as the destination.
+Such attacks can be detected if the original message reaches the true
+destination unaltered,
+which SMFT can help to ensure.
+
+In its current form, our work has several limitations.
+Most obviously, it requires complete control over the network structure.
+However, we have shown that even partially control over network structure can
+improve attack tolerance propeties.
+Still a more flexible network structure is desirable.
+There is also the question of how to construct such a network without a centeral
+authority.
+This limitation may not be as severe as it seems,
+due to the nested structure of the butterfly network.
+We conjecture that smaller independently-formed networks could be
+merged into a single larger network without central coordination.
+
+In addition to addressing the above limitations,
+we see several potential directions for future work.
+The development of new structured networks or multipath routing algorithms
+could achieve higher levels of redundancy and attack-tolerance.
+It is also desirable to examine how changes to social dynamics could shift
+self-organized networks towards a more decentralized structure.
+Finally, our results could be implemented to address specific applications, e.g.,
+secure messaging, domain name resolution, or anonymous web browsing.
 
@@
 
-We have presented a novel concurrent multipath routing algorithm for the butterfly
-topology,
-as well as a scheme for using this algorithm to construct a highly attack-tolerant
-virtual channel between any two nodes, even when no fully-trusted path exists
-between them.
-Under this scheme, the probability of an adversary causing an undetectable error
+Coercion-resistant, topological approaches to attack tolerance are needed to
+address the current vulnerability of communications infrastructure to
+censorship and surveillance.
+We have presented a novel concurrent multipath routing (CMR)
+algorithm for the butterfly network,
+as well as a structured multipath fault tolerance (SMFT) scheme,
+which can be combined to create a coercion-resistant,
+attack-tolerant point-to-point communication architecture.
+We have also shown how assuming bounded trust
+transitivity can enable a quantitative analysis of the
+relationships between network structure, trust, and attack-tolerance.
+In our architecture,
+the probability of an adversary causing an undetectable error
 decreases exponentially with the network's effective redundancy.
 The effective redundancy, in the case of the butterfly topology,
-grows exponentially with the trust radius.
+grows exponentially with the radius of trust transitivity.
 Furthermore, a small increase in the number of messages sent can compensate
 for a large increase in the number of messages compromised by an adversary.
-We have also demonstrated how the assumption of bounded trust
-transitivity can enable a quantitative analysis of the
-relationships between network structure, trust, and attack-tolerance.
 These results are directly applicable to systems in which the link
-structure can be imposed by the designer,
-and more generally, provide a theoretical foundation that can be used more to evaluate
-the role of network structure, trust transitivity, and effective redundancy
-on attack-tolerance.
+structure can be imposed by the designer.
+Even when network structure cannot be perfectly controlled,
+we have shown that partially rewiring a snapshot of the internet's router network
+can greatly increase its attack-tolerance properties.
+We believe that this work provides a foundation for the
+development of additional topology-based communication
+architectures to guard against technical and coercive adversarial attacks,
+including censorship and surveillance.
 

### 2017-11-01 www-2018/source/paper.tex
Final revisions.

@@
 on attack-tolerance.
-We next present a structured multipath fault tolerance (SMFT) to extend
+We next present a structured multipath fault tolerance (SMFT) scheme to extend
 standard fault tolerance techniques to the case of adversarial faults in
@@
 there is a potential for an attacker to exploit it through coercion.
-So, {\em attack-tolerance cannot be achieved without the ability to influence
+So we claim that {\em attack-tolerance cannot be achieved without the ability to influence
 network structure}.
@@
 {\em adversarial} faults in {\em complex networks}.
-We do so by modeling redundant {\em independent paths}
-with {\em bounded trust transitivity} as a single virtual channel,
-and show that the probability of detecting adversarial faults
-approaches 1 exponentially with the number of paths;
+We do so by showing that the probability of detecting adversarial faults
+approaches 1 exponentially with the number of {\em independent paths},
+and that with $h$-degree {\em bounded trust transitivity},
+all $h$-internally vertex disjoint paths are independent.
 }
@@
 paths between two nodes in a wrap-around butterfly network
-increases exponentially with $h$,
-and thus that the number of independent paths in a network
-increases exponentially with the level of trust transitivity;
-}
-\item{We present a scalable and efficient concurrent
-multipath routing (CMR) algorithm on the butterfly network topology,
-which can be combined with SMFT to achieve a high level of attack tolerance.}
+is at least $2^h$,
+and present a scalable and efficient concurrent multipath routing (CMR) algorithm
+to find these paths,
+which can be combined with SMFT to achieve a high level of attack-tolerance.
+}
+\item{We show that rewiring a snapshot of the internet's router network with
+edges from a butterfly network allows it to tolerate a higher number of failures
+without fragmenting, and increases the effective redundancy in the presence
+of a large number of adversarial faults.}
 \end{itemize}
@@
 Section \ref{sec-bf-route} presents our concurrent multipath routing
-algorithm for the butterfly.
+algorithm for the butterfly network.
 Section \ref{sec-discussion} discusses the results.
@@
 Standard fault tolerance methods use redundancy to detect and correct
-statistically indpendent faults.
+statistically independent faults.
 In complex networks however, faults can be correlated when,
@@
 However, in real communication architectures, direct links between all
-pairs of individuals not possible and messages must be routed through
+pairs of individuals are not possible and messages must be routed through
 a number of intermediate nodes.
@@
 and call tree notification systems \cite{nickerson_thinking_2010}.
-However, when attack-tolerance is desired,
-it will always require control over network structure in order to
-eliminate single points of failure.
-The regular structure of the butterfly is not a limitation of our approach,
-but rather a reflection of the inherent difficulty of attack-tolerance.
+More flexible architectures may be possible,
+but attack-tolerance will always require some level of influence over
+network structure in order to limit single points of failure.
 To address the case when the network cannot be fully controlled,
@@
 router network on January 2, 2000
-\cite{leskovec_graphs_2005}.
-However, we equate a subset of the nodes with the nodes of a 9-dimensional
-butterfly network and rewire a fraction of the edges to match the butterfly
-topology.
+\cite{leskovec_graphs_2005}, having 6493 nodes and 13914 edges.
 At each step of the simulation, betweenness centrality is recalculated and the
 most central node is removed.
+We also simulate attack on several rewired networks.
+We 1. generate edges corresponding to a 9-dimensional butterfly network between
+the 4608 highest-degree router nodes,
+2. choose a fraction $f$ of those edges at random,
+3. add those edges to the router network, and
+4. remove an equal number of the original edges at random.
 
@@
 this number increases to 2\%
-with only 10% of the edges rewired to match the butterfly.
-With 90\% of the edges rewired,
+with only 10\% of the butterfly edges present.
+With 90\% of the butterfly edges present
 the network remains unfragmented beyond the failure of the 8\% most
@@
 averaging over 150 source-destination pairs.
-For all but $h=0$,
-the original network has higher redundancy for a small fraction of
-node failures,
+For $h=0$, the rewired version has strictly higher redundancy.
+For $h>0$,
+the original network has higher redundancy when the number of failures is small,
 while the rewired network has higher redundancy beyond a crossover point.
@@
 The original network fragments at about 1\%.
-With 90\% of the edges rewired,
+With 90\% of the butterfly edges present,
 the network remains unfragmented beyond the failure of the 8\% most

### 2018-05-18 plosone/attack.tex
Add plosone draft.

--- 2017-11-01 www-2018/source/paper.tex
@@
-\documentclass[sigconf]{acmart}
+% Template for PLoS
+% Version 3.5 March 2018
+%
+% % % % % % % % % % % % % % % % % % % % % %
+%
+% -- IMPORTANT NOTE
+%
+% This template contains comments intended 
+% to minimize problems and delays during our production 
+% process. Please follow the template instructions
+% whenever possible.
+%
+% % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Once your paper is accepted for publication, 
+% PLEASE REMOVE ALL TRACKED CHANGES in this file 
+% and leave only the final text of your manuscript. 
+% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
+% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
+%
+%
+% There are no restrictions on package use within the LaTeX files except that 
+% no packages listed in the template may be deleted.
+%
+% Please do not include colors or graphics in the text.
+%
+% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
+%
+% % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- FIGURES AND TABLES
+%
+% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
+%
+% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
+% - Figures should be uploaded separately from your manuscript file. 
+% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
+% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
+% For figure citations, please use "Fig" instead of "Figure".
+% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
+%
+% Tables should be cell-based and may not contain:
+% - spacing/line breaks within cells to alter layout or alignment
+% - do not nest tabular environments (no tabular environments within tabular environments)
+% - no graphics or colored text (cell background color/shading OK)
+% See http://journals.plos.org/plosone/s/tables for table guidelines.
+%
+% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
+%
+% IMPORTANT
+% Below are a few tips to help format your equations and other special characters according to our specifications.
+For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
+%
+% For inline equations, please be sure to include all portions of an equation in the math environment.
+For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
+%
+% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
+%
+% Please add line breaks to long display equations when possible in order to fit size of the column. 
+%
+% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
+%
+% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
+%
+% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
+%
+% % % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Please contact latex@plos.org with any questions.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+
+\documentclass[10pt,letterpaper]{article}
+\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
+
+% amsmath and amssymb packages, useful for mathematical formulas and symbols
+\usepackage{amsmath,amssymb}
+
+% Use adjustwidth environment to exceed column width (see example table in text)
+\usepackage{changepage}
+
+% Use Unicode characters when possible
+\usepackage[utf8x]{inputenc}
+
+% textcomp package and marvosym package for additional characters
+\usepackage{textcomp,marvosym}
+
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
+
+% Use nameref to cite supporting information files (see Supporting Information section for more info)
+\usepackage{nameref,hyperref}
+
+% line numbers
+\usepackage[right]{lineno}
+
+% ligatures disabled
+\usepackage{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
+
+% color can be used to apply background shading to table cells only
+\usepackage[table]{xcolor}
+
+% array package and thick rules for tables
+\usepackage{array}
+
+% create "+" rule type for thick vertical lines
+\newcolumntype{+}{!{\vrule width 2pt}}
+
+% create \thickcline for thick horizontal lines of variable length
+\newlength\savedwidth
+\newcommand\thickcline[1]{%
+  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+  \cline{#1}%
+  \noalign{\vskip\arrayrulewidth}%
+  \noalign{\global\arrayrulewidth\savedwidth}%
+}
+
+% \thickhline command for thick horizontal lines that span the table
+\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+\hline
+\noalign{\global\arrayrulewidth\savedwidth}}
+
+% Remove comment for double spacing
+\usepackage{setspace} 
+\doublespacing
+
+% Text layout
+\raggedright
+\setlength{\parindent}{0.5cm}
+\textwidth 5.25in 
+\textheight 8.75in
+
+% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
+% Captions will be left justified
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
+\renewcommand{\figurename}{Fig}
+
+% Use the PLoS provided BiBTeX style
+\bibliographystyle{plos2015}
+
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
+
+% Header and Footer with logo
+\usepackage{lastpage,fancyhdr,graphicx}
+\usepackage{epstopdf}
+%\pagestyle{myheadings}
+\pagestyle{fancy}
+\fancyhf{}
+%\setlength{\headheight}{27.023pt}
+%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+\rfoot{\thepage/\pageref{LastPage}}
+\renewcommand{\headrulewidth}{0pt}
+\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+\fancyheadoffset[L]{2.25in}
+\fancyfootoffset[L]{2.25in}
+\lfoot{\today}
 
@@
 \usepackage{epigraph}
-\usepackage[utf8]{inputenc}
 \usepackage{afterpage}
@@
 \usetikzlibrary{arrows,chains,decorations.pathreplacing}
-\usepackage{pifont}% http://ctan.org/pkg/pifont
+\usepackage{pifont} % http://ctan.org/pkg/pifont
+\usepackage{amsthm}
+
+%% Include all macros below
 
@@
 \newtheorem{theorem}{Theorem}
-
 \newcommand{\cmark}{\ding{51}}
@@
 
-% Copyright
-%\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-\setcopyright{rightsretained}
-%\setcopyright{usgov}
-%\setcopyright{usgovmixed}
-%\setcopyright{cagov}
-%\setcopyright{cagovmixed}
-
-% DOI
-\acmDOI{10.475/123_4}
-
-% ISBN
-\acmISBN{123-4567-24-567/08/06}
-
-%Conference
-\acmConference[WEBCONF'18]{The Web Conference}{April 2018}{Lyon, France} 
-\acmYear{2018}
-\copyrightyear{2018}
-
-\acmArticle{4}
-\acmPrice{15.00}
+\newcommand{\lorem}{{\bf LOREM}}
+\newcommand{\ipsum}{{\bf IPSUM}}
+
+%% END MACROS SECTION
 
 \begin{document}
-\title{
-    Towards Attack-Tolerant Networks:
-    Concurrent Multipath Routing and the Butterfly Network}
-
-%\author{Edward L. Platt}
-%\affiliation{%
-%  \institution{University of Michigan\\School of Information}
-%  \city{Ann Arbor} 
-%  \state{Michigan} 
-%}
-%\email{elplatt@umich.edu}
-
-%\author{Daniel M. Romero}
-%\affiliation{%
-%  \institution{University of Michigan\\School of Information}
-%  \city{Ann Arbor} 
-%  \state{Michigan} 
-%}
-%\email{drom@umich.edu}
-\author{Anonymized\\for\\blind review}
-
-% The default list of authors is too long for headers.
-%\renewcommand{\shortauthors}{E.L. Platt and D.M. Romero.}
-\renewcommand{\shortauthors}{Anonymized for blind review}
-
-\begin{abstract}
+\vspace*{0.2in}
+
+% Title must be 250 characters or less.
+\begin{flushleft}
+{\Large
+\textbf\newline{Towards attack tolerant networks: concurrent multipath routing and the butterfly network} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+}
+\newline
+% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
+\\
+Edward L. Platt\textsuperscript{1,*},
+Daniel M. Romero\textsuperscript{1}
+\\
+\bigskip
+\textbf{1} School of Information, University of Michigan, Ann Arbor, Michigan, USA
+\\
+\bigskip
+
+% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
+% 
+% Remove or comment out the author notes below if they aren't used.
+%
+% Primary Equal Contribution Note
+
+% Additional Equal Contribution Note
+% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
+
+% Current address notes
+%\textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
+% \textcurrency b Insert second current address 
+% \textcurrency c Insert third current address
+
+% Deceased author note
+%\dag Deceased
+
+% Group/Consortium Author Note
+%\textpilcrow Membership list can be found in the Acknowledgments section.
+
+% Use the asterisk to denote corresponding authorship and provide email address in note below.
+* elplatt@umich.edu
+
+\end{flushleft}
+% Please keep the abstract below 300 words
+\section*{Abstract}
 It is crucial for large-scale communication networks such as the internet
@@
 More broadly, we hope that this work will serve as a starting point for the
-development of additional topology-based attack-tolerant communication
+evelopment of additional topology-based attack-tolerant communication
 architectures to guard against the dangers of censorship and surveillance.
-\end{abstract}
-
-%
-% The code below should be generated by the tool at
-% http://dl.acm.org/ccs.cfm
-% Please copy and paste the code instead of the example below. 
-%
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10002978.10002986</concept_id>
-<concept_desc>Security and privacy~Formal methods and theory of security</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003033.10003034</concept_id>
-<concept_desc>Networks~Network architectures</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10011007.10010940.10010971.10010972.10010540</concept_id>
-<concept_desc>Software and its engineering~Peer-to-peer architectures</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Security and privacy~Formal methods and theory of security}
-\ccsdesc[500]{Networks~Network architectures}
-\ccsdesc[500]{Software and its engineering~Peer-to-peer architectures}
-
-\keywords{
-Adversarial Faults,
-Resilience,
-Butterfly Network,
-Fault Tolerance,
-Multipath Routing,
-Censorship,
-Decentralization}
-
-\maketitle
-
-\section{Introduction}
+
+% Please keep the Author Summary between 150 and 200 words
+% Use first person. PLOS ONE authors please skip this step. 
+% Author Summary not valid for PLOS ONE submissions.   
+%\section*{Author summary}
+%Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Curabitur eget porta erat.
+Morbi consectetur est vel gravida pretium.
+Suspendisse ut dui eu ante cursus gravida non sed sem.
+Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
+Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
+Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
+Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
+Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
+
+\linenumbers
+
+% Use "Eq" instead of "Equation" for equation citations.
+\section*{Introduction}
 
@@
 While our work is motivated by the network of internet routers,
-the results are theoretical,
-and can potentially be applied to many different types of networks,
-whether made of physical wires, virtual tunnels,
-or other types of links.
+the results are topological,
+and can potentially be applied to many different types of networks.
 
@@
 
-\section{Background and Related Work}
+\section*{Background and Related Work}
 \label{sec-related}
@@
 content-addressable network \cite{fiat_censorship_2002},
-although this scheme does not scale well and is impractical due to a
-high level of data replication.
+although this scheme requires high levels of data replication and indefinite
+storage.
 Perhaps the most mature structural solution is the Freenet collaboration
@@
 a single path.
-The special case of {\em concurrent} multipath routing uses mutliple paths
+The special case of {\em concurrent} multipath routing uses multiple paths
 simultaneously.
@@
 wireless sensor networks (WSNs),
-and have thus focused on ad hoc, unstructured networks, often having a central
+and have thus focused on ad-hoc, unstructured networks, often having a central
 base station.
@@
 to the role of network structure.
-The alogorithm we present for the butterfly topology complements existing work
-by addressing cases where links are not constrained by physical
-distance,
-and where network structure can be engineered for greater attack-tolerance.
 
@@
 properties (Table \ref{tab:compare}).
-Decentralized architectures are more resistent to coercion
+Decentralized architectures are more resistant to coercion
 \cite{grewal_caveat_2013}
@@
 \begin{center}
-\begin{tabular}{l|ccccc}
-                              & Decentralized & Trust-based & Bounded-trust & Topological & Point-to-point \\
+\begin{tabular}{l|p{0.6in}p{0.6in}p{0.6in}p{0.6in}p{0.6in}}
+                              & Decentra-lized & Trust-based & Bounded-trust & Topo-logical & Point-to-point \\
 \hline
-Public Key Infrastructure     &               & \cmark      & \cmark        &             & \cmark     \\
+PKI            &               & \cmark      & \cmark        &             & \cmark     \\
 Web of Trust   & \cmark        & \cmark      &               &             & \cmark     \\
 Freenet        & \cmark        & \cmark      & \cmark        &             &            \\
-Federated Byzantine Agreement & \cmark        & \cmark      & \cmark        &             & \cmark     \\
-Proof of Work/Blockchain      & \cmark        &             & \cmark        &             &            \\
+FBA            & \cmark        & \cmark      & \cmark        &             & \cmark     \\
+Proof of Work  & \cmark        &             & \cmark        &             &            \\
 Fiat-Saia      & \cmark        & \cmark      & \cmark        & \cmark      &            \\
-Structured Multipath FT       & \cmark        & \cmark      & \cmark        & \cmark      & \cmark     \\
+SMFT           & \cmark        & \cmark      & \cmark        & \cmark      & \cmark     \\
 \hline
@@
 
-\section{Trust Networks and Fault Tolerance}
+\section*{Trust Networks and Fault Tolerance}
 \label{sec-ft}
@@
 
-\subsection{Multipath Fault Tolerance}
+\subsection*{Multipath Fault Tolerance}
 
@@
 For now, let us assume our sender (Alice) and receiver (Bob) are connected
-by $\delta$ direct channels, with indpendent errors.
+by $\delta$ direct channels, with independent errors.
 We will return to the question of constructing these channels in subsequent
@@
 Bob can then take any of several actions:
-1. request retransmission;
+1. request re-transmission;
 2. send receipts so Alice knows which paths have been compromised;
@@
 
-Figure \ref{fig:pfail} shows the value of $p_f$
+Fig. \ref{fig:pfail} shows the value of $p_f$
 as a function of $k$ and $l$.
-Equation (\ref{eq:pf_approx}) shows that while $p_f$
+Eq. (\ref{eq:pf_approx}) shows that while $p_f$
 depends on the fractions of
@@
 
-\begin{figure}
+\begin{figure}[!h]
 \centerline{\includegraphics{fig-perror}}
@@
 
-\subsection{Bounded Trust Model}
+\subsection*{Bounded Trust Model}
 
 So far, we have assumed that Alice and Bob have access to some number
-$\delta$ of channels with statistically indpendent faults.
+$\delta$ of channels with statistically independent faults.
 However, in real communication architectures, direct links between all
@@
 in the network reduces the probability that a node can resist compromise
-by a multiplicatve constant.
+by a multiplicative constant.
 Such a situation could occur if nodes more distant from Alice are
@@
 although the model can easily be extended to directed multigraphs.
-Vertices representing commiunicating parties,
+Vertices representing communicating parties,
 with edges representing mutually trusted communication links.
@@
 gaining complete control of their behavior.
-We also assume that the adversary is specifically targeting communicaiton
+We also assume that the adversary is specifically targeting communication
 between $s$ and $t$ and
@@
 
-\subsection{Effective Redundancy}
+\subsection*{Effective Redundancy}
 
@@
 one of the endpoints.
-This condition holds if and only if two paths are indpendent
+This condition holds if and only if two paths are independent
 under the bounded trust model with radius $h$.
@@
 channels that can be constructed between them having
-statistically indpendent faults.
+statistically independent faults.
 We thus refer to this quantity as the 
@@
 \begin{figure}
-\centerline{\includegraphics[width=3in,height=2.08in]{fig-partial-trust}}
+\centerline{\includegraphics[width=4in,height=2.78in]{fig-partial-trust.pdf}}
 \caption{
@@
 
-\subsection{Structured Multipath Fault Tolerance}
-
-Finding a maximal set of independent paths for an abitrary network is NP hard
+\subsection*{Structured Multipath Fault Tolerance}
+
+Finding a maximal set of independent paths for an arbitrary network is NP hard
 \cite{reiter_resilient_1998},
@@
 
-\section{The Butterfly Network Topology}
+\section*{The Butterfly Network Topology}
 \label{sec-butterfly}
@@
 In order to implement structured multipath fault tolerance,
-we need a strucutred network topology with high effective redundancy.
+we need a structured network topology with high effective redundancy.
 In this paper, we apply SMFT to the butterfly network topology
@@
 
-\subsection{Butterfly Network Topology}
+\subsection*{Butterfly Network Topology}
 
@@
 we utilize the $m$-dimensional, directed wrap-around butterfly
-(Figure \ref{fig:bf-route}),
+(Fig. \ref{fig:bf-route}),
 denoted $\wbf(m)$:
@@
 where $\mathbb{Z}_m$ is the set of integers modulo $m$,
-$\oplus$ represents componentwise addition modulo 2,
+$\oplus$ represents component-wise addition modulo 2,
 and $1_l$ is a vector with a 1 in index $l$ and 0 elsewhere.
@@
 There are two types of edges: down, and down-right
-(shown in Figure \ref{fig:butterfly}).
+(shown in Fig. \ref{fig:butterfly}).
 Down edges ($E_\downarrow$) connect nodes sharing the same $z$ value
@@
 but one having the place-within-level equal to $z$ with the $l$th bit inverted.
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[height=3in]{fig-bf-route}
+\end{center}
+\caption{
+A 3-dimensional wraparound butterfly network.
+Note that the rightmost nodes are the same nodes as the leftmost,
+drawn twice for visual clarity.
+The highlighted nodes and edges show the path from node (0,000)
+to node (1,101).
+\label{fig:bf-route}
+}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics{fig-butterfly}
+\end{center}
+\caption{
+Schematic illustration of the two types of edges in a directed butterfly
+network.
+The node $(l,z)$ is shown as the bit string $z$ with a square around the
+$l$th bit.
+``Down'' edges increment $l$, leaving $z$ unchanged,
+while ``down-right'' edges increment $l$ and invert the $l$th bit of $z$.
+In the wrap-around variant, the nodes with maximum $l$ have down and down-right
+edges to the nodes with $l=0$.
+\label{fig:butterfly}
+}
+\end{figure}
 
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[height=3in,angle=90]{fig-bf-route.pdf}
-\end{center}
-\caption{
-A 3-dimensional wraparound butterfly network.
-Note that the rightmost nodes are the same nodes as the leftmost,
-drawn twice for visual clarity.
-The highlighted nodes and edges show the path from node (0,000)
-to node (1,101).
-\label{fig:bf-route}
-}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics{fig-butterfly.pdf}
-\end{center}
-\caption{
-Schematic illustration of the two types of edges in a directed butterfly
-network.
-The node $(l,z)$ is shown as the bit string $z$ with a square around the
-$l$th bit.
-``Down'' edges increment $l$, leaving $z$ unchanged,
-while ``down-right'' edges increment $l$ and invert the $l$th bit of $z$.
-In the wrap-around variant, the nodes with maximum $l$ have down and down-right
-edges to the nodes with $l=0$.
-\label{fig:butterfly}
-}
-\end{figure}
-
 The structure of the butterfly network lends itself to a well-known
-(unipath) routing algorithm (Figure \ref{fig:bf-route}),
+(unipath) routing algorithm (Fig. \ref{fig:bf-route}),
 which we later extend to the multipath case.
@@
 
-\subsection{Butterfly Rewiring}
-
-\begin{figure*}[h!]
-\centerline{\includegraphics[width=7in,height=1.75in]{fig-mincut}}
-\caption{
-Simulation of targeted attacks against a snapshot of the internet's router
-network with a fraction of the edges rewired into a partial butterfly configuration.
-The effective redundancy is shown for several values of trust transitivity $h$.
-For $h>0$, the original network has higher effective redundancy up to a crossover point,
-after which the rewired network performs better.
-}
-\label{fig:mincut}
-\end{figure*}
+\subsection*{Butterfly Rewiring}
 
@@
 While the original router network fragments when about 1\% of the nodes
-have been removed (Figure \ref{fig:diameter}),
+have been removed (Fig. \ref{fig:diameter}),
 this number increases to 2\%
@@
 central nodes.
-The effective redundancy for varius values of trust transitivity $h$
-are shown in Figure \ref{fig:mincut}.
+The effective redundancy for various values of trust transitivity $h$
+are shown in Fig. \ref{fig:mincut}.
 The effective redundancy is calculated by collapsing nodes and edges within
-$h$ hops of source and destination into single nodes and finding the mincut
+$h$ hops of source and destination into single nodes and finding the min-cut
 between them,
@@
 \begin{figure}
-\centerline{\includegraphics[width=3in,height=2.25in]{fig-diameter}}
+\centerline{\includegraphics{fig-diameter}}
 \caption{
@@
 
-\section{Multipath Butterfly Routing}
+\begin{figure*}[h!]
+\centerline{\includegraphics[width=5in,height=5in]{fig-mincut}}
+\caption{
+Simulation of targeted attacks against a snapshot of the internet's router
+network with a fraction of the edges rewired into a partial butterfly configuration.
+The effective redundancy is shown for several values of trust transitivity $h$.
+For $h>0$, the original network has higher effective redundancy up to a crossover point,
+after which the rewired network performs better.
+}
+\label{fig:mincut}
+\end{figure*}
+
+\section*{Multipath Butterfly Routing}
 \label{sec-bf-route}
@@
 the specific node on Alice's trust boundary
-(or equivalentely the node on Bob's trust boundary, or the untrusted intermdiate).
+(or equivalently the node on Bob's trust boundary, or the untrusted intermediate).
 
@@
 with different properties used to prove independence at each stage
-(see Figure \ref{fig:route-overview}).
+(see Fig. \ref{fig:route-overview}).
 In the first cycle (stages 1--4), path independence is guaranteed by ensuring that
@@
 $h$ bits of the place-within-level preceding the destination index.
-A full example is illustrated in Figure \ref{fig:routing}.
-
-\afterpage{
+A full example is illustrated in Fig. \ref{fig:routing}.
+
 \begin{table}[h!]
 \caption{Butterfly Multipath Routing Variables\label{tab:routing}}
-{
 \begin{center}
@@
 \end{center}
-}
 \end{table}
 
-\begin{figure}[t!]
+\begin{figure}[h!]
 \begin{center}
@@
 }
+\end{figure}
+
+\begin{figure}[!h]
 \begin{center}
@@
 \end{figure}
-}
-
-\subsection{Algorithm Specification}
+
+\subsection*{Algorithm Specification}
 
@@
 
-\subsection{Proof of Path Independence}
+\subsection*{Proof of Path Independence}
 
@@
 
-\section{Discussion}
+\section*{Discussion}
 \label{sec-discussion}
@@
 However, we have shown that even partially control over network structure can
-improve attack tolerance propeties.
+improve attack tolerance properties.
 Still a more flexible network structure is desirable.
-There is also the question of how to construct such a network without a centeral
+There is also the question of how to construct such a network without a central
 authority.
@@
 
-\section{Conclusion}
+\section*{Conclusion}
 \label{sec-conclusion}
@@
 
-%ACKNOWLEDGMENTS are optional
-\section{Acknowledgments}
-Anonymized for blind review
-%The authors would like to thank Tony Garnock-Jones, A. Frederick Dudley, and
-%Nathaniel Bezanson for helpful conversations.
-%This research was partly supported by the National Science Foundation under Grant No. IIS-1617820.
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{paper}
+\section*{Acknowledgments}
+The authors would like to thank Tony Garnock-Jones, A. Frederick Dudley, and
+Nathaniel Bezanson for helpful conversations.
+This research was partly supported by the National Science Foundation under
+Grant No. IIS-1617820.
+
+\nolinenumbers
+
+%\bibliography{references}
+
+% Either type in your references using
+% \begin{thebibliography}{}
+% \bibitem{}
+% Text
+% \end{thebibliography}
+%
+% or
+%
+% Compile your BiBTeX database using our plos2015.bst
+% style file and paste the contents of your .bbl file
+% here. See http://journals.plos.org/plosone/s/latex for 
+% step-by-step instructions.
+% 
+\begin{thebibliography}{10}
+
+\bibitem{sterbenz_resilience_2010}
+Sterbenz JP, Hutchison D, Çetinkaya EK, Jabbar A, Rohrer JP, Schöller M,
+  et~al.
+\newblock Resilience and survivability in communication networks: {Strategies},
+  principles, and survey of disciplines.
+\newblock Computer Networks. 2010;54(8):1245--1265.
+
+\bibitem{barabasi_scale-free_2009}
+Barabási AL, {others}.
+\newblock Scale-free networks: a decade and beyond.
+\newblock Science. 2009;325(5939):412.
+
+\bibitem{albert_error_2000}
+Albert R, Jeong H, Barabási AL.
+\newblock Error and attack tolerance of complex networks.
+\newblock Nature. 2000;406(6794):378--382.
+
+\bibitem{nayak_different_2010}
+Nayak GN, Samaddar SG.
+\newblock Different flavours of man-in-the-middle attack, consequences and
+  feasible solutions.
+\newblock In: Computer {Science} and {Information} {Technology} ({ICCSIT}),
+  2010 3rd {IEEE} {International} {Conference} on. vol.~5. IEEE; 2010. p.
+  491--495.
+
+\bibitem{grewal_caveat_2013}
+Grewal GS, Ryan MD, Bursuc S, Ryan PY.
+\newblock Caveat coercitor: {Coercion}-evidence in electronic voting.
+\newblock In: Security and {Privacy} ({SP}), 2013 {IEEE} {Symposium} on. IEEE;
+  2013. p. 367--381.
+
+\bibitem{elmer-dewitt_first_1993}
+Elmer-Dewitt P, Jackson D.
+\newblock First nation in cyberspace.
+\newblock Time. 1993;6:62--64.
+
+\bibitem{baran_distributed_1964}
+Baran P, {others}.
+\newblock On distributed communications.
+\newblock Volumes I-XI, RAND Corporation Research Documents, August. 1964; p.
+  637--648.
+
+\bibitem{dainotti_analysis_2011}
+Dainotti A, Squarcella C, Aben E, Claffy KC, Chiesa M, Russo M, et~al.
+\newblock Analysis of country-wide internet outages caused by censorship.
+\newblock In: Proceedings of the 2011 {ACM} {SIGCOMM} conference on {Internet}
+  measurement conference. ACM; 2011. p. 1--18.
+
+\bibitem{christianson_why_1997}
+Christianson B, Harbison WS.
+\newblock Why isn't trust transitive?
+\newblock In: Security protocols. Springer; 1997. p. 171--176.
+
+\bibitem{zin_survey_2015}
+Zin SM, Anuar NB, Kiah MLMM, Ahmedy I.
+\newblock Survey of secure multipath routing protocols for {WSNs}.
+\newblock J Netw Comput Appl. 2015;55:123--153.
+
+\bibitem{barabasi_emergence_1999}
+Barabási AL, Albert R.
+\newblock Emergence of scaling in random networks.
+\newblock Science. 1999;286(5439):509--512.
+
+\bibitem{hunter_pakistan_2008}
+Hunter P.
+\newblock Pakistan {YouTube} block exposes fundamental internet security
+  weakness: {Concern} that pakistani action affected youtube access elsewhere
+  in world.
+\newblock Computer Fraud \& Security. 2008;2008(4):10--11.
+
+\bibitem{poulsen_edward_2013}
+Poulsen K.
+\newblock Edward {Snowden}’s e-mail provider defied {FBI} demands to turn
+  over crypto keys, documents show.
+\newblock WIRED. 2013;.
+
+\bibitem{zimmermann_official_1995}
+Zimmermann PR.
+\newblock The official {PGP} user's guide.
+\newblock MIT press; 1995.
+
+\bibitem{ferguson_practical_2003}
+Ferguson N, Schneier B.
+\newblock Practical cryptography.
+\newblock New York: Wiley; 2003.
+
+\bibitem{avizienis_basic_2004}
+Avizienis A, Laprie JC, Randell B, Landwehr C.
+\newblock Basic concepts and taxonomy of dependable and secure computing.
+\newblock IEEE T Depend Secure. 2004;1(1):11--33.
+
+\bibitem{von_neumann_probabilistic_1956}
+Von~Neumann J.
+\newblock Probabilistic logics and the synthesis of reliable organisms from
+  unreliable components.
+\newblock Automata studies. 1956;34:43--98.
+
+\bibitem{qadir_exploiting_2015}
+Qadir J, Ali A, Yau KLA, Sathiaseelan A, Crowcroft J.
+\newblock Exploiting the power of multiplicity: a holistic survey of
+  network-layer multipath.
+\newblock IEEE Comm Surv Tut. 2015;17(4):2176--2213.
+
+\bibitem{khiani_comparative_2013}
+Khiani SR, Dethe C, Thakare V.
+\newblock Comparative {Analysis} of {Multipath} {Routing} {Techniques} and
+  {Design} of {Secure} {Energy} {Aware} {Routing} {Algorithm} for {Wireless}
+  {Sensor} {Network}.
+\newblock IJACR. 2013;3(3):374.
+
+\bibitem{kshemkalyani_distributed_2008}
+Kshemkalyani AD, Singhal M.
+\newblock Distributed computing: principles, algorithms, and systems.
+\newblock Cambridge University Press; 2008.
+
+\bibitem{lua_survey_2005}
+Lua EK, Crowcroft J, Pias M, Sharma R, Lim S.
+\newblock A survey and comparison of peer-to-peer overlay network schemes.
+\newblock IEEE Commun Surv Tut. 2005;7(2):72--93.
+
+\bibitem{korzun_structured_2013}
+Korzun D, Gurtov A.
+\newblock Structured peer-to-peer systems: fundamentals of hierarchical
+  organization, routing, scaling, and security.
+\newblock New York, NY: Springer; 2013.
+
+\bibitem{mohr_explaining_1982}
+Mohr LB.
+\newblock Explaining organizational behavior.
+\newblock Jossey-Bass; 1982.
+
+\bibitem{walker_mass_2012}
+Walker DC.
+\newblock Mass notification and crisis communications: {Planning},
+  preparedness, and systems.
+\newblock CRC Press; 2012.
+
+\bibitem{nickerson_thinking_2010}
+Nickerson JV, Tversky B, Corter JE, Yu L, Mason D.
+\newblock Thinking with networks.
+\newblock In: {CogSci}. vol.~36; 2010.
+
+\bibitem{ellison_ten_2000}
+Ellison C, Schneier B.
+\newblock Ten risks of {PKI}: {What} you're not being told about public key
+  infrastructure.
+\newblock Comput Secur J. 2000;16(1):1--7.
+
+\bibitem{levien_attack-resistant_2009}
+Levien R.
+\newblock Attack-resistant trust metrics.
+\newblock In: Computing with {Social} {Trust}. Springer; 2009. p. 121--132.
+
+\bibitem{reiter_resilient_1998}
+Reiter MK, Stubblebine SG.
+\newblock Resilient authentication using path independence.
+\newblock IEEE T Comput. 1998;47(12):1351--1362.
+
+\bibitem{lamport_byzantine_1982}
+Lamport L, Shostak R, Pease M.
+\newblock The {Byzantine} generals problem.
+\newblock TOPLAS. 1982;4(3):382--401.
+
+\bibitem{castro_practical_1999}
+Castro M, Liskov B, {others}.
+\newblock Practical {Byzantine} fault tolerance.
+\newblock In: {OSDI}. vol.~99; 1999. p. 173--186.
+
+\bibitem{dwork_pricing_1993}
+Dwork C, Naor M.
+\newblock Pricing via processing or combatting junk mail.
+\newblock In: Advances in {Cryptology}. Springer; 1993. p. 139--147.
+
+\bibitem{nakamoto_bitcoin:_2008}
+Nakamoto S.
+\newblock Bitcoin: {A} peer-to-peer electronic cash system.
+\newblock bitcoinorg. 2008; p.~28.
+
+\bibitem{mazieres_stellar_2015}
+Mazières D.
+\newblock Stellar {Consensus} {Protocol}: {A} {Federated} {Model} for
+  {Internet}-level {Consensus}; 2015.
+
+\bibitem{fiat_censorship_2002}
+Fiat A, Saia J.
+\newblock Censorship resistant peer-to-peer content addressable networks.
+\newblock In: {SIAM} {SODA}. ACM; 2002. p. 94--103.
+
+\bibitem{clarke_freenet:_2001}
+Clarke I, Sandberg O, Wiley B, Hong TW.
+\newblock Freenet: {A} distributed anonymous information storage and retrieval
+  system.
+\newblock In: Designing {Privacy} {Enhancing} {Technologies}. Springer; 2001.
+  p. 46--66.
+
+\bibitem{shamir_how_1979}
+Shamir A.
+\newblock How to share a secret.
+\newblock Communications of the ACM. 1979;22(11):612--613.
+
+\bibitem{blakley_safeguarding_1979}
+Blakley GR.
+\newblock Safeguarding cryptographic keys.
+\newblock P Natl Comp Conf. 1979;48:313--317.
+
+\bibitem{zhang_using_2002}
+Zhang H, Goel A, Govindan R.
+\newblock Using the small-world model to improve freenet performance.
+\newblock In: {INFOCOM}. vol.~3. IEEE; 2002. p. 1228--1237.
+
+\bibitem{kleinberg_small-world_2000}
+Kleinberg J.
+\newblock The small-world phenomenon: {An} algorithmic perspective.
+\newblock In: {STOC}. ACM; 2000. p. 163--170.
+
+\bibitem{alrajeh_secure_2013}
+Alrajeh NA, Alabed MS, Elwahiby MS.
+\newblock Secure ant-based routing protocol for wireless sensor network.
+\newblock Int J Distrib Sens N. 2013;2013.
+
+\bibitem{kohno_improvement_2012}
+Kohno E, Okazaki T, Takeuchi M, Ohta T, Kakuda Y, Aida M.
+\newblock Improvement of assurance including security for wireless sensor
+  networks using dispersed data transmission.
+\newblock J Comp Sys Sci. 2012;78(6):1703--1715.
+
+\bibitem{khalil_unmask:_2010}
+Khalil I, Bagchi S, Rotaru CN, Shroff NB.
+\newblock {UnMask}: {Utilizing} neighbor monitoring for attack mitigation in
+  multihop wireless sensor networks.
+\newblock Ad Hoc Networks. 2010;8(2):148--164.
+
+\bibitem{lou_h-spread:_2006}
+Lou W, Kwon Y.
+\newblock H-{SPREAD}: a hybrid multipath scheme for secure and reliable data
+  collection in wireless sensor networks.
+\newblock IEEE T Veh Technol. 2006;55(4):1320--1330.
+
+\bibitem{liu_secure_2012}
+Liu A, Zheng Z, Zhang C, Chen Z, Shen X.
+\newblock Secure and energy-efficient disjoint multipath routing for {WSNs}.
+\newblock IEEE T Veh Technol. 2012;61(7):3255--3265.
+
+\bibitem{leskovec_graphs_2005}
+Leskovec J, Kleinberg J, Faloutsos C.
+\newblock Graphs over time: densification laws, shrinking diameters and
+  possible explanations.
+\newblock In: Proceedings of the eleventh {ACM} {SIGKDD} international
+  conference on {Knowledge} discovery in data mining. ACM; 2005. p. 177--187.
+
+\bibitem{xu_internet_2011}
+Xu X, Mao ZM, Halderman JA.
+\newblock Internet censorship in {China}: {Where} does the filtering occur?
+\newblock In: International {Conference} on {Passive} and {Active} {Network}
+  {Measurement}. Springer; 2011. p. 133--142.
+
+\bibitem{dingledine_tor:_2004}
+Dingledine R, Mathewson N, Syverson P.
+\newblock Tor: {The} second-generation onion router.
+\newblock DTIC Document; 2004.
+
+\end{thebibliography}
 

### 2018-05-18 www-2018/source/paper.tex
Add old www source files.

@@
 rewired for greater attack-tolerance.
-More broadly, we hope that this work will serve as a starting point for the
-development of additional topology-based attack-tolerant communication
-architectures to guard against the dangers of censorship and surveillance.
+%More broadly, we hope that this work will serve as a starting point for the
+%development of additional topology-based attack-tolerant communication
+%architectures to guard against the dangers of censorship and surveillance.
 \end{abstract}
@@
 So far, we have assumed that Alice and Bob have access to some number
-$\delta$ of channels with statistically indpendent faults.
+$\delta$ of channels with statistically independent faults.
 However, in real communication architectures, direct links between all
@@
 
-Finding a maximal set of independent paths for an abitrary network is NP hard
+Finding a maximal set of independent paths for an arbitrary network is NP hard
 \cite{reiter_resilient_1998},
@@
 central nodes.
-The effective redundancy for varius values of trust transitivity $h$
+The effective redundancy for various values of trust transitivity $h$
 are shown in Figure \ref{fig:mincut}.
 The effective redundancy is calculated by collapsing nodes and edges within
-$h$ hops of source and destination into single nodes and finding the mincut
+$h$ hops of source and destination into single nodes and finding the min-cut
 between them,

### 2018-05-21 plosone/attack.tex
Revisions to intro based on www feedback.

@@
 \cite{elmer-dewitt_first_1993}.
-But 17 years after Albert et. al \cite{albert_error_2000}
-showed the internet's router network to be vulnerable to targeted attacks
-(vs. random faults), the fundamental problem remains unsolved.
+But 18 years after Albert et. al \cite{albert_error_2000}
+showed that the topology of the internet's router network
+makes it vulnerable to targeted attacks
+(vs. random faults), the fundamental problem of attack-tolerant network topology
+remains unsolved.
+Attack-tolerant topologies are desirable not only for the router network,
+but for any physical or virtual network where a compromised node puts
+communication at risk.
+For example, the network of verified keys in the public key infrastructure
+underlying secure http \cite{ellison_ten_2000},
+or the network of DNS nameservers.
 The ongoing vulnerability of the internet is evidenced by a long history of 
@@
 \cite{dainotti_analysis_2011}.
-In this paper, we present the first theoretical architecture
-for point-to-point networked communication,
-that tolerates not only random faults, but also targeted attacks,
+In this paper, we present the first theoretical network topology
+supporting attack-toloerant, point-to-point networked communication,
 without relying on infinitely transitive trust
@@
 the information flowing through them.
+While cryptograpy can provide some protection against surveillance, it is
+vulnerable to {\em man-in-the-middle} attacks \cite{nayak_different_2010},
+and cannot overcome censorship when communication is blocked.
+In this paper, we instead consider a topological approach.
 The Internet's decentralized design was motivated
@@
 The action (known as a {\em black hole attack}) was intended to censor YouTube
-within Pakistan only, but resulted in a worldwide cascading failure.
-The action was initiated by government order in Pakistan,
-and spread beyond Pakistan when a router misconfiguration allowed the false
-routing information to propagate globally.
-While the government order and router misconfiguration initiated the outage,
-it was the structure of the Internet's router network that allowed a fault in a
-single router to propagate globally.
-And while the action was not an intentional attack against the global Internet,
+within Pakistan only, but resulted in a worldwide cascading failure
+when a router misconfiguration allowed the false routing information to
+propagate outside of Pakistan.
+This incident exemplifies the type of attack requiring a topological approach.
+First, the attack was {\em non-technological} (a government order),
+allowing the attacker to bypass any cryptographic or technology-based defenses.
+Second, the attack originated at a {\em single point of failure}
+(a misconfigured router).
+Third, the behavior of the compromised component (the router) cascaded
+through a {\em network} (the network of internet routers) because the correct
+behavior of other components depended on the correct behavior
+of the single point of failure.
+And while the action was not an intentional attack against the global internet,
 the ability of an attacker to succeed without even trying only highlights
@@
 
-In 2013, the Texas-based email provider Lavabit was ordered to disclose
+As another example, in 2013,
+the Texas-based email provider Lavabit was ordered to disclose
 their private SSL keys to the FBI \cite{poulsen_edward_2013}.
@@
 in order to protect their users from surveillance.
-Once again, the attack was successful due to a highly centralized
-architecture:
-SSL keys under control of a single entity, in a single legal jurisdiction.
+Once again, the attack was non-technical.
+And again, the attack was on a single point of failure:
+Lavabit's web server and that server's TLS/SSL keys.
+In this case, the affected network was the
+internet's public key infrastructure.
+With the private keys, an attacker would be able to intercept and
+surveil traffic because the issuing certificate authority (and any
+users trusting that authority) would incorrectly trust that they were
+communicating with Lavabit.
 While originally intended as surveillance,
 this action effectively became an act of censorship.
-It is also important to note that while Lavabit's cryptography worked as intended,
-the attack was still successful because the system was
-vulnerable to non-technical coercion.
 So we see that such vulnerabilities are not limited to any one system or protocol,
@@
 
-Coercion-resistant, topological approaches to attack tolerance are needed to
-address the current vulnerability of communications infrastructure
-to censorship and surveillance.
-This paper presents several contributions towards advancing those goals.
-While our work is motivated by the network of internet routers,
-the results are topological,
-and can potentially be applied to many different types of networks.
+This paper addresses the need for a theoretial understanding of network and
+redundancy-based approaches to attack tolerance.
+Our primary result is theoretical: an algorithm for constructing highly redundant
+paths in a particular network topology.
+While we motivate the need for such network topologies using examples such as
+the internet router network and webs of trust,
+we do not propoose that our algorithm as a pracitcal solution for any of these
+examples.
+Rather, our result is a demonstration that such topological approaches to
+attack-tolerance are theoretically possible.
 
@@
 by compromising a number of intermediate nodes.
-We also assume that the source and destination nodes trust their neighbors
-i.e., that the adversary is unable to compromise them,
+We also assume that edges represent {\em direct trust},
+i.e., the belief that an adversary is unlikely to compromise a neighbor,
 as in the web of trust approach
-\cite{zimmermann_official_1995,ferguson_practical_2003}.
-Unlike the web of trust approach, which assumes that trust is infinitely
-transitive,
-we assume {\em bounded trust transitivity}.
+\cite{zimmermann_official_1995,richters_trust_2011}.
+The web of trust approach requires infinite trust transitivity in that
+for one node to trust another,
+at least one path of direct trusted edges must exist between them.
+The assumption of infinite transitivity is unrealistic
+\cite{christianson_why_1997},
+so we instead make a stricter assumption: {\em bounded trust transitivity},
+that trust transitivity can only be extended over a finite number of edges.
+Even with this more restrictive assumption, we show how trust can be established
+between two nodes even when no path of directly trusted edges exists between them.
 
@@
 The well-known and widely-used {\em web of trust} approach
-\cite{zimmermann_official_1995,ferguson_practical_2003}
+\cite{zimmermann_official_1995,richters_trust_2011}
 is a decentralized alternative.
@@
 One common approach, the web of trust
-\cite{zimmermann_official_1995,ferguson_practical_2003},
+\cite{zimmermann_official_1995,richters_trust_2011},
 extends trust infinitely transitively:
@@
 \begin{center}
-\includegraphics{fig-routing.pdf}
+\includegraphics{fig-routing}
 \end{center}

### 2018-05-21 plosone/attack.tex
Change description of bounded trust.

@@
 \cite{zimmermann_official_1995,richters_trust_2011}.
-The web of trust approach requires infinite trust transitivity in that
-for one node to trust another,
-at least one path of direct trusted edges must exist between them.
+The web of trust approach typically assumes infinite trust transitivity in that
+it is possible for one node to have some level of trust for another as long
+as they are connected by a path of directly trusted edges,
+regardless of the length of that path.
 The assumption of infinite transitivity is unrealistic

### 2018-05-22 plosone/attack.tex
More revisions based on www feedback.

@@
 the network structure.
-We have already seen that whenever a single point of failure exists in the network,
-there is a potential for an attacker to exploit it through coercion.
-So we claim that {\em attack-tolerance cannot be achieved without the ability to influence
+Regardless of implementation difficulty,
+we argue that {\em attack-tolerance cannot be achieved without the ability to influence
 network structure}.
-Luckily, there are scenarios in which network topology can be dictated.
-Examples include overlay networks
+Targeted attacks, by definition, target single points of failure.
+Attack tolerance can thus be achieved in two ways: 1. preventing failure at
+those points, or 2. preventing the existence of central points.
+Because individual points are always vulnerable to coercive, non-technological
+attacks, the former method is insufficient.
+We must instead rely on some control over topology to prevent the existence of
+single points of failure.
+The difficulty of achieving control over network topology can be mitigated by
+a number of approaches.
+Attack-tolerant networks might be sub-components of
+larger, less-constrained systems.
+For example, a single centralized server might be replaced by a distributed
+network of servers,
+each with different ownership, physical location, and legal jurisdiction,
+without placing any unrealistic constraints on the clients connecting to
+those servers.
+When attack tolerant topologies are nested (as in the case of the butterfly
+topology), multiple independent sub-components could be merged into larger
+ones over time.
+Real-world examples of structured neworks include: overlay networks
 \cite{lua_survey_2005, korzun_structured_2013},
@@
 {\em when the need for attack-tolerance is high enough to warrant investment
-in infrastructure, networks can be engineered and maintained as infrastructure}.
-It is also worth noting that attack-tolerant networks may be sub-components of
-larger, less-constrained systems.
-For example, a single server might be replaced by a distributed network of servers,
-each with different ownership, physical location, and legal jurisdiction,
-without placing any unrealistic constraints on the clients connecting to
-those servers.
+in infrastructure, topology can be engineered and maintained as infrastructure}.
 
@@
 {\em adversarial} faults in {\em complex networks}.
-We do so by showing that the probability of detecting adversarial faults
-approaches 1 exponentially with the number of {\em independent paths},
-and that with $h$-degree {\em bounded trust transitivity},
-all $h$-internally vertex disjoint paths are independent.
+Assuming $h$-degree {\em bounded trust transitivity},
+We show that the probability of detecting adversarial faults
+$h$-internally vertex disjoint paths.
 }
@@
 }
-\item{We show that rewiring a snapshot of the internet's router network with
-edges from a butterfly network allows it to tolerate a higher number of failures
+\item{We show that rewiring a the edges of the internet's router network to
+resemble a butterfly network allows it to tolerate a higher number of failures
 without fragmenting, and increases the effective redundancy in the presence
@@
 In a web of trust,
-individuals choose who they trust initially.
-Trust is then extended to new individuals if they are vouched for by a
-currently-trusted individual.
+individuals can have {\em direct trust},
+as well as {\em indirect trust} for those trusted by someone directly trusted.
+Typically, this transitivity is extended to any number of hops,
+sometimes reducing trust by a multiplicative factor at each hop.
 Infinite trust transitivity is helpful for establishing a large group of
@@
 \cite{christianson_why_1997}.
-Additionally, the web of trust approach
-does not distinguish between paths of different lengths.
-Our work addresses both of these limitations by requiring only
+Our work addresses this limitation by assuming only
 bounded trust transitivity.
@@
 senders can tolerate attacks from significantly more powerful
-adversaries,
-as long as the network provides large $\delta$.
+adversaries.
+Furthermore, this scheme requires only a small increase in network traffic
+as long as the network provides a large number of independent paths $\delta$.
 
@@
 redundant channels and the number of adversarial faults.
+A small increase in the number of utilized paths (network traffic)
+can compensate for a large increase in attacker power.
 }

### 2019-01-11 plosone/attack-figs.tex
Add plosone revisions.

--- 2018-05-22 plosone/attack.tex
@@
 the internet router network and webs of trust,
-we do not propoose that our algorithm as a pracitcal solution for any of these
+we do not propose that our algorithm as a pracitcal solution for any of these
 examples.
 Rather, our result is a demonstration that such topological approaches to
-attack-tolerance are theoretically possible.
+attack-tolerance are theoretically possible and suggest the importance
+of further theoretical and applied work.
 
@@
 One common approach, the web of trust
-\cite{zimmermann_official_1995,richters_trust_2011},
-extends trust infinitely transitively:
-Alice trusts her friends, as well as her friends' friends, and so on.
+\cite{zimmermann_official_1995,richters_trust_2011}.
+Alice has {\em direct trust} for some number of nodes.
+Alice has {\em indirect trust} for nodes separated by two or more hops of
+direct trust.
+Typically, this transitive trust is extended indefinitely,
+sometimes reduced by a multiplicative factor at each step.
 However, the assumption of infinitely transitive trust is unrealistic
@@
 as it depends only on whether some path exists, not the number or quality of paths.
-
-An alternative assumption might be that each hop away from Alice in
-in the network reduces the probability that a node can resist compromise
-by a multiplicative constant.
-Such a situation could occur if nodes more distant from Alice are
-more favorably disposed to Alice's adversary, more likely to cooperate with that
-adversary, or less likely to take proactive security measures against that
-adversary.
-An even simpler version assumes that nodes up to some
-fixed number of hops cannot be compromised, and that those beyond can.
-This simplified version is still more realistic than infinite transitivity
-and will be convenient for proving our results.
-We now proceed to define our model formally.
+We adopt a simpler, yet more realistic {\em bounded trust} assumption:
+that nodes up to some fixed number of hops cannot be compromised,
+and that those beyond can.
+This assumption will be convenient for proving our results,
+which we now proceed to define formally.
 
@@
 \cite{kshemkalyani_distributed_2008}.
-The structure of the butterfly network is highly constrained,
-making it most suitable for applications where portions of the
-network structure can be designed or dictated.
-Examples of such networks include:
-overlay networks \cite{lua_survey_2005, korzun_structured_2013},
-formal organizations \cite{mohr_explaining_1982},
-government-regulated cellular networks \cite{walker_mass_2012},
-and call tree notification systems \cite{nickerson_thinking_2010}.
+The butterfly network is highly structured,
+making it most suitable for applications where portions of the network structure
+can be controlled or influenced.
+The butterfly network is also recursive, with larger versions composed out of
+multiple smaller versions,
+making it possible for independently organized
+attack-tolerant networks to merge into larger ones over time.
 More flexible architectures may be possible,
@@
 network can greatly increase it's effective redundancy and
-attack-tolerance properties.
+attack-tolerance properties, without requiring additional edges.
 
@@
 For any two nodes, the length of the shortest path between them is
-$O(\log N)$, where N is the number of nodes in the network;
+$O(\log N)$, where N is the number of nodes in the network
+(corresponding to low-latency in real-world terms);
 \item[Sparsity:]
@@
 We also simulate attack on several rewired networks.
+The rewiring process alters the network structure to resemble a butterfly
+topology, without adding any additional edges.
 We 1. generate edges corresponding to a 9-dimensional butterfly network between
@@
 network with a fraction of the edges rewired into a partial butterfly configuration.
-The original network fragments at about 1\%.
-With 90\% of the butterfly edges present,
+The original network fragments when the top 1\% of nodes are removed.
+With only 10\% of the butterfly edges present,
+this value doubles to 2\%.
+With 90\% of the edges rewired,
 the network remains unfragmented beyond the failure of the 8\% most
 central nodes.
+The rewiring scheme does not require adding any additional edges.
 }
@@
 \begin{center}
-\begin{tabular}{|l|l|}
-\hline
-NAME & VARIABLE \\\hline
-butterfly dimension & $m \in \mathbb{Z}_+$ \\\hline
-node level & $l \in \mathbb{Z} : 0 \leq l < m$ \\\hline
-node place within level & $z \in \mathbb{Z}_2^m$ \\\hline
-trust radius & $h \in \mathbb{Z} : 1 \leq h \leq \lfloor m/2 \rfloor$ \\\hline
-path index & $s \in \mathbb{Z}_2^h$ \\\hline
+\begin{tabular}{ll}
+Name & Variable \\\hline
+butterfly dimension & $m \in \mathbb{Z}_+$ \\
+node level & $l \in \mathbb{Z} : 0 \leq l < m$ \\
+node place within level & $z \in \mathbb{Z}_2^m$ \\
+trust radius & $h \in \mathbb{Z} : 1 \leq h \leq \lfloor m/2 \rfloor$ \\
+path index & $s \in \mathbb{Z}_2^h$ \\
 \end{tabular}
@@
 \cite{hunter_pakistan_2008}
-and the FBI's surveillance-turned-censorship of Lavabit
+and the FBI's surveillance and censorship of Lavabit
 \cite{poulsen_edward_2013}.
+
 The reader may wonder how our methods could be employed in scenarios
@@
 merged into a single larger network without central coordination.
+When nodes and edges exist in geographic space (as in cables connecting internet
+routers) this scheme might require connections between very distant locations.
+Such connections are extremely expensive to construct and maintain,
+although many exist today in the form of redundant internet backbones and
+undersea cables. An alternative solution might involve satellite links,
+which connect distant geographic points much more easily.
+While turning our network theoretical results into practical applications
+will require considerable additional work, we believe that work is inevitably
+necessary in order to create attack-tolerant networks.
 
@@
 grows exponentially with the radius of trust transitivity.
-Furthermore, a small increase in the number of messages sent can compensate
+Furthermore, a small increase in the number of messages sent
+(traffic volume) can compensate
 for a large increase in the number of messages compromised by an adversary.
-These results are directly applicable to systems in which the link
-structure can be imposed by the designer.
+These results require some control over the structure of a network,
+or some portion of the network.
 Even when network structure cannot be perfectly controlled,
@@
 
-\bibitem{ferguson_practical_2003}
-Ferguson N, Schneier B.
-\newblock Practical cryptography.
-\newblock New York: Wiley; 2003.
+\bibitem{richters_trust_2011}
+Richters O, Peixoto TP.
+\newblock Trust transitivity in social networks.
+\newblock PloS one. 2011;6(4):e18384.
 

### 2021-04-27 figures.tex
Update on Overleaf.

\begin{figure}
% discrete_generated_nk/archive_bg16/2020-07-24 122519 9741a16
    \centering
    \includegraphics[width=3.25in,height=2.75in]{fig-performance-bg16.eps}
    \caption{Mean final scores for social learning on a rugged landscape with neighbor sampling (100 runs of 300 learning-step iterations). We use the same task parameters as \cite{barkoczi_social_2016} and reproduce their results: the best-neighbor strategy leads to inefficient networks (e.g., lattice) outperforming efficient ones (e.g., complete), while the conform strategy leads to efficient networks outperforming inefficient ones. We find the same result holds in the Network Deliberation setting: the best-neighbor strategy leads to the inefficient long-path group assignment outperforming random group assignment, while the conform strategy leads to the efficient random-group assignment outperforming long-path.}
    \label{fig:result-bg16}
\end{figure}

\begin{figure}
% discrete_generated_nk/archive_hybrid/2020-07-24 153547 9741a16
    \centering
    \includegraphics[width=3.25in,height=2.75in]{fig-performance-hybrid.eps}
    \caption{Mean final scores for social learning on a rugged landscape using hybrid social-indiviual learning and neighbor sampling (100 runs of 300 learning-step iterations). Performance of the conform strategy is similar to Fig. \ref{fig:result-bg16}, while the best-neighbor and local-majority strategies show improved performance across all networks.}
    \label{fig:result-hybrid}
\end{figure}

\begin{figure}
% discrete_generated_nk/archive_hybrid/2020-07-25 003254 9741a16
    \centering
    \includegraphics[width=3.25in,height=2.75in]{fig-performance-nosample.eps}
    \caption{Mean final scores for social learning on a rugged landscape using hybrid social-individual learning without neighbor sampling. Average node degree is 4. While the best-neighbor strategy shows comparable performance across all networks, the conform and local majority strategies perform significantly better in both network deliberation settings than in the single-group networks, with conform/random-group matching the best performance of any single-group setting. As with the sampling case, we find that under the best-neighbor strategy, long-path group assignment outperforms random group assignmnt, while under the conform strategy, random group assignment provides the best performance. The local-majority strategy performs similarly in both of the network deliberation settings, landing between the other two strategies.}
    \label{fig:result-nosample}
\end{figure}

\begin{figure}
% discrete_generated_nk/archive_hybrid/2020-08-14 160620 d441939
    \centering
    \includegraphics[width=3.25in,height=2.75in]{fig-performance-complex.eps}
    \caption{Mean final scores for social learning on a highly rugged landscape ($K=12$) using hybrid social-individual learning without neighbor sampling. Average node degree is 4. Results are similar to Fig. \ref{fig:result-nosample}, but local majority now outperforms conformity for both ND networks. Neither Best Neighbor nor Conform strategy can create a solution outside the basins of attraction already discovered, while Local Majority can. As the number of local maxima becomes much larger than the number of agents, the generative nature of the Local Majority strategy allows it to discover basins of attraction for new local maxima, increasing the probability of finding the global maximum.}
    \label{fig:result-complex}
\end{figure}

\subsection{Noisy Estimation}

\begin{figure}
% discrete_generated/archive/2020-10-07 112527 d441939
    \centering
    \includegraphics[width=3.25in,height=2.75in]{fig-performance-estimation-nosample.eps}
    \caption{Mean final scores for social learning on a noisy estimation task (1000 runs of 50 learning step iterations). The task consists of estimating a 7-bit target state from noisy observations. Each bit of each agent's observation has a 40\% chance of being incorrect. Average node degree is 4 (excluding complete network) and no sampling is used. The best-neighbor strategy performs well across all networks and the conform stragey performs poorly across all networks. However, the local majority strategy performs well on the complete network and for network deliberation with random group assignments.}
    \label{fig:result-nosample}
\end{figure}

### 2021-04-27 supplement.tex
Update on Overleaf.

\documentclass[twocolumn,10pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{times}
\usepackage{geometry}
\usepackage[ruled,vlined]{algorithm2e}

\usepackage[switch]{lineno}
\linenumbers

\newtheorem{definition}{Definition}
\newtheorem{property}{Property}
\newtheorem{claim}{Claim}

\DeclareMathOperator{\unif}{unif}
\DeclareMathOperator{\sample}{sample}
\DeclareMathOperator{\setcount}{count}
\DeclareMathOperator{\mode}{mode}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

\usepackage{geometry}
\geometry{left=1.5cm,right=1.5cm,top=2cm,bottom=2cm}
\setlength{\columnsep}{0.5cm}

\title{Supplementary Information: \\
Small interlocking groups improve mass deliberation\\in the presence of strong social influence}
\author{
%Edward L. Platt\\
%University of Michigan\\
%elplatt@umich.edu
%and
%Herminio Bodon\\
%Northwestern University\\
%HerminioBodon2020\\@u.northwestern.edu
%\and
%Daniel M. Romero\\
%University of Michigan\\
%drom@umich.edu
Authors Redacted \\
for \\
Double-Blind Peer-Review
}
\date{\today}

\begin{document}

\maketitle

\section{Simulation Procedure}
We begin with a network $(V,E_t)$. The vertices $V$ correspond to agents. The edges $E_t$ allow agents to exchange information with their immediate neighbors at time $t$.
Agents collaborate to find a solution $s$ from a space of solutions $\mathcal{S}$ that maximizes an objective function $Q(s)$.
We use binary strings of length $d$ as our solution space: $s \in \mathbb{Z}^d$.
At any one time $t$, each agent $v$ has exactly one preferred solution $s_{v,t}$.
We generate a tunably rugged objective function $Q(s)$ using an NK-Model \cite{kauffman_towards_1987} (see Section \ref{subsec:task}) with N=15, K=6, exp=8.

The simulation begins at time $t=0$ by generating a set of initial solutions $s_{v,0}$ for the agents. These are generated randomly, with each possible solution having equal probability.
The simulation proceeds iteratively.
At time $t$ each agent applies one of several {\em learning strategies} (see Section \ref{sec:learning}), to determine its preferred solution at time $t+1$.
Learning strategies can rely on the agent's own solution, the solutions of its neighbors, and potentially additional information shared by its neighbors.
Learning strategies may also incorporate information about the objective function, modeling an agents' ability to evaluate the quality of solutions (see Section \ref{sec:capability}).
To allow for ties, learning strategies produce a set of solutions rather than a single solution.
In our simulations, we choose winners at random in the case of a tie.
We allow the simulation to proceed for 300 iterations, which we have found sufficient to guarantee convergence with our chosen parameters.

\section{Network Topologies}

We use networks to represent constraints on who talks to whom. Each agent is represented by a vertex, and two agents are able to interact when their vertices are connected by an edge. In this paper we are concerned with network deliberation, which takes place on interlock networks. Interlocks are composed of small cliques (i.e., pods) with some vertices belonging to multiple cliques. For the interlock networks in this paper, each vertex belongs to exactly one pod at a time, with pod membership periodically reassigned to achieve multiple membership. As a control, we also consider networks representing conventional deliberation.

\subsection{Conventional Deliberation}
We model the communication network structure of conventional deliberation using two static networks: small-world networks \cite{watts_collective_1998}, and preferential attachment networks \cite{barabasi_emergence_1999}. In typical deliberative settings, from deliberative assemblies to online forum threads, the contributions of any member are potentially visible to all other members.
The network of {\em potential} communication can be modelled by a complete graph.
However, a complete graph may not be the best model for the {\em actual} communication that takes place in such a network. In a real-world deliberation, some participants are more influential than others \cite{goel_structure_2012, shaw_laboratories_2014}, and many communications can be missed or ignored by any particular participant.

Human social networks often exhibit large clustering, and short path lengths.
To model such dynamics, we use Watts-Strogatz small-world networks  \cite{watts_collective_1998}.
These networks model, for example, when participants interact mostly with their strong ties, but occasionally with a weak tie. Social networks have also been found to exhibit skewed degree distributions with long tails \cite{barabasi_emergence_1999}.
We model these networks using the Barabási-Albert preferential attachment model.
These networks model settings where a small number of individuals produce a disproportionate share of the communication due to factors such as social capital, expertise, or confidence.

\subsection{Network Deliberation: Interlock Networks}
Both of our network deliberation conditions use networks built from interlocking pods.
These pods are small complete graphs of roughly equal size.
At any given time, each agent can communicate with all other agents in exactly one pod.
Periodically, all agents are simultaneously reassigned to new pods.
These reassignments create an interlocking structure, allowing information to flow through the entire group.
This structure can be represented as a time-dependent or directed network.
The properties of an interlock network depend on the particular method for assigning agents to pods.
Crucially, interlocking structure is created when agents are placed with others they have not yet interacted with.
We thus require that pod assignment schemes satisfy the following mixing property:
\begin{property}
\label{prop:mixing}
For any agent $v$, and any two assignment stages $t$, $t^\prime$, the probability that at least one neighbor of $v$ at stage $t$ is also a neighbor at stage $t^\prime$ approaches 0 as $|V|$ grows.
\end{property}

\subsubsection{Random-Pod Assignment}
The {\em random-pod} assignment method (Algorithm \ref{alg:random}) simply assigns agents to groups at random. Pseudocode for this method is shown in Algorithm \ref{alg:random}.

\begin{algorithm}
\SetAlgoLined
\DontPrintSemicolon
\SetKwFunction{removeRandom}{removeRandom}
\SetKwFunction{insert}{insert}
\SetKwFunction{emp}{empty}
\SetKw{algnot}{not}
\KwData{A vertex list $V$, the pod size $M \in \mathbb{Z}$.}
\KwResult{A partition of the vertices.}
    $N \longleftarrow \lceil \frac{|V|}{M} \rceil$ \;
    $P \longleftarrow$ List of $N$ empty sets \;
    \For{$i \in 1, \ldots, M$}{
        \ForEach{$s \in P$}{
            \If{\algnot $V.\emp()$}{
                $v \longleftarrow V.\removeRandom()$ \;
                $s.\insert(v)$ \;
            }
        }
    }
    \Return{$P$}
\caption{Random-Pod Assignment}
\label{alg:random}
\end{algorithm}

\begin{claim}
Random-pod assignment satisfies Property \ref{prop:mixing}.
\end{claim}

\begin{proof}
Let $v$ be a vertex of graph $G=(V,E_t)$ with a set of $M$ neighbors at stage $t$ denoted $N_t(v)$.
The probability that the $k$th neighbor chosen at time $t^\prime \neq t$ belongs to $N_t(v)$, given that the first $k - 1$ did not, is:
\begin{align*}
p_\text{repeat}(k)
&= \frac{M - 1}{|V| - k} \\
&\leq \frac{M - 1}{|V| - M}.
\end{align*}
The probability that none of the $M-1$ neighbors chosen at time $t^\prime$ belong to $N(v)$ thus satisfies:
\begin{align*}
p_\text{mixed}
&\geq
\left( 1 - \frac{M - 1}{|V| - M} \right)^{M-1} \\
\lim_{|V|\rightarrow\infty} p_\text{mixed}
&= 1
&\qedhere
\end{align*}
\end{proof}

Random pod assignment also creates structurally efficient networks with short geodesic distances, as shown in Figure \ref{fig:broadcast}.
There has been evidence both for
\cite{lazer_network_2007, derex_partial_2016, mason_propagation_2008, barkoczi_social_2016}
and against
\cite{mason_collaborative_2012, barkoczi_social_2016}
the effectiveness of structurally efficient networks in social learning.

\begin{figure}
    \centering
    \includegraphics[width=3.33in]{fig-broadcast.eps}
    \caption{Mean time necessary for a signal broadcast from one node to reach the entire network. We use mean broadcast time is used as a measure of geodesic length for time-varying networks.}
    \label{fig:broadcast}
\end{figure}

\subsubsection{Long-Path Pod Assignment}

To study the effects of path length, we require an alternative pod assignment method that produces long paths.
However, we must still retain Property \ref{prop:mixing} such that it is rare for two agents to repeatedly share the same pod.
This property ensures the creation of many new edges at each stage, which is difficult--but not impossible--to reconcile with long average path length.
We now present a {\em long-path} pod assignment method, which meets both of the above goals.

We begin with a high-level overview.
Agents are assigned an integer position on a 1-dimensional circular lattice.
By preferring short-distance links on this lattice, we maintain long path lengths.
We also partition agents according to the remainder of their position, modulo some prime, i.e., their residue class.
By limiting links to agents in the same residue class, and using a unique prime at each stage, we ensure that it is rare for multiple agents to share a pod twice in a row.
Specifically, for stages using primes $p$ and $q$, two agents will share a pod for both stages when their positions are equal modulo $pq$. Pseudocode for long-path assignment is shown in Algorithm \ref{alg:long}.

\begin{algorithm}
\SetAlgoLined
\DontPrintSemicolon
\SetKw{algnot}{not}
\SetKwFunction{removeRandom}{removeRandom}
\SetKwFunction{moduli}{moduli}
\SetKwFunction{append}{append}
\SetKwFunction{enqueue}{enqueue}
\SetKwFunction{dequeue}{dequeue}
\SetKwFunction{concat}{concatenate}
\KwData{A vertex list $V$, the pod size $M \in \mathbb{Z}$, the stage $t \geq 0 \in \mathbb{Z}$, a list of co-prime integers $\moduli$.}
\KwResult{A partition of the vertices.}
    $N \longleftarrow \lceil \frac{|V|}{M} \rceil$ \;
    $P \longleftarrow$ List of $N$ empty sets \;
    \eIf{$t = 0$}{
        // Place all vertices in same residue class \;
        $p \longleftarrow 1$ \;
    }{
        // Choose integer to define residue class \;
        $p \longleftarrow \moduli[t]$ \;
    }
    // Assign vertices to residue classes \;
    $R \longleftarrow$ List of $p$ empty queues \;     
    \For{$z \in 0, \ldots, |V|-1$}{
        $R[z \bmod p].\enqueue(V[z])$ \;
    }
    // Divide each residue class into pods \;
    $P \longleftarrow $ Empty list of lists. \;
    \For{$r \in R$}{
        $N_r \longleftarrow \lceil \frac{|r|}{M} \rceil$ \;
        $P_r \longleftarrow$ List of $N_r$ empty lists \;
        \For{$i \in 1, \ldots, M$}{
            \For {$j \in 0, \ldots, N_r - 1$}{
                \If{ \algnot $r.\emp()$}{
                    $v \leftarrow r.\dequeue()$ \;
                    $P_r[j].\append(v)$ \;
                }
            }
        }
        $P.\concat(P_r)$
    }
    \Return{$P$}
\caption{Long-Path Assignment}
\label{alg:long}
\end{algorithm}

\begin{claim}
When the pod size $M$ is less than the modulus $p_t$ for all stages $t$, long-path pod assignment satisfies Property \ref{prop:mixing}.
\end{claim}

\begin{proof}
Let $p$ and $q$ be co-prime integers used as moduli for stages $t$ and $t^\prime$ respectively.
Let $z(x)$ denote the integer position of vertex $x$.
Let $v$ be a vertex, and let us assume vertex $w$ belongs to the same pod as $v$ at both time $t$ and $t^\prime$.
This assumption implies:
\begin{align*}
    z(v) &\equiv z(w) \pmod{p} \\
    z(v) &\equiv z(w) \pmod{q} \\
    \implies z(v) - z(w) &\equiv 0 \pmod{p} \\
    &\equiv 0 \pmod{q} \\
    &= npq & \text{for some integer n}.
\end{align*}
By the definition of the long-path algorithm:
\begin{align*}
    |z(v)-z(w)| & \leq (M - 1)p \\
    \implies |n|pq & \leq (M - 1)p \\
    |n|q & \leq (M - 1).
\end{align*}
By assumption, $q \geq M$, so the above can only be satisfied by $n=0$,
which in turn implies $v=w$. \qedhere
\end{proof}

To see that the long-path procedure produces long geodesics, note that the pods of size $M$ connect nodes at most $(M - 1)p_i$ apart in position, preventing the creation of any shortcut edges, as long as $p_i$ is small compared to $|V|$.
Numerical simulations confirm that the long-path algorithm produces geodesics larger than random pod assignment (Figure \ref{fig:broadcast}).
The number of agents may not be an exact multiple of $M$, so the final pods may be truncated. However, the network will be a sub-graph of a network that does have a multiple of $M$ nodes, so the structure will not be fundamentally changed, and these edge effects should become negligible as the number of agents increases.

\section{Learning Strategies}
\label{sec:learning}
The choice of network structure defines which agents can exchange information, but not how those agents act on the information received. An agent's actions based on available information are determined by that agent's {\em learning strategy}. Learning strategies can be divided into {\em individual learning strategies} which use only information directly observable by the agent, and {\em social learning strategies} which use information communicated by neighboring agents.

Generally, a learning strategy may incorporate both social and individual components, but we will discus each individually before approaching hybrid strategies. The learning strategies available to an agent also depend on the agent's ability to evaluate the quality of candidate solutions, which we refer to as its {\em capabilities}, which we now explore in further detail.

\subsection{Agent Capabilities}
\label{sec:capability}
In human social learning, the ability of individuals to evaluate the quality of a solution can vary with factors such as expertise and task type. When comparing learning strategies, it will be helpful to classify them according to the capabilities they require of agents. We formalize these capabilities as oracle functions, which accept one or more task solutions as input and reveal whether the supplied state(s) satisfy some particular property.

One of the strongest capabilities an agent might posses, is to compare two arbitrary solutions to determine which yields a higher value of the objective function. We represent this capability using the {\em arbitrary comparison} oracle.
\begin{definition}
The arbitrary comparison oracle $\mathcal{O^>}(s_1, s_2)$ is given by:
\begin{eqnarray}
\mathcal{O}^>(s_1, s_2) &\equiv&
Q(s_1) > Q(s_2)
\end{eqnarray}
\end{definition}

Under a weaker assumption, agents might be ``experts'' on their current solution and able to compare that particular solution to any other. This ability is represented by a {\em single-solution comparison} oracle.
\begin{definition}
The single-solution comparison oracle for solution $s$ is given by:
\begin{eqnarray}
\mathcal{O}^>_s(\tilde{s}) &\equiv&
Q(\tilde{s}) > Q(s)
\end{eqnarray}
\end{definition}

In many contexts, it is reasonable to assume that agents can explore the effects of small changes to their current solution. When the solutions are binary strings, such variations can be modeled by flipping a single bit of the solution string. We formalize this capability through the {\em mutation comparison} oracle.
\begin{definition}
The mutation comparison oracle for solution $s$ is given by:
\begin{eqnarray}
\mathcal{O}^\oplus_s(i) &\equiv&
Q(s \oplus e_i) > Q(s),
\end{eqnarray}
where $\oplus$ represents addition mod 2 and $e_i$ represents the binary string with a single 1 at index $i$.
\end{definition}

\subsection{Mutation-Based Individual Learning}
In real-world collaborations, individuals sometimes work independently, even when communication is available. Independent work might be motivated by practical concerns (such as distributing labor) or by social dynamics. We model individual learning using single-bit mutations, following the examples of
\cite{lazer_network_2007} and \cite{barkoczi_social_2016}. Agents constructs a candidate solution by flipping a single bit of their solution at an index selected uniformly at random. If the candidate solution yields a higher value of the objective function, the agent adopts it as the new solution. As this strategy requires comparing solutions that differ by at most one bit, an oracle at least as powerful as the mutation comparison oracle is required.

\begin{definition}
The mutation individual learning strategy $\mathcal{L}_{\text{I}}(v)$ is defined as:
\begin{eqnarray}
i &\sim& \unif(1,d) \nonumber\\
\mathcal{L}_{\text{I}}(v) &\equiv&
\begin{cases}
\{s(v) \oplus e_i\} & \mbox{if $\mathcal{O}^{\oplus}_{s(v)}(i)$}
\\
\{s(v)\} & \mbox{otherwise}.
\end{cases}
\end{eqnarray}
\end{definition}

\subsection{Best-Neighbor}
Among social strategies, the straightforward greedy approach results in the {\em best-neighbor} strategy, which has been widely used in prior work \cite{lazer_network_2007, mason_propagation_2008, barkoczi_social_2016}. In this strategy, an agent simply compares the solutions of all agents in its neighborhood and adopts the best. While straightforward, this strategy makes a strong assumption about agent capabilities: access to the arbitrary comparisons.
\begin{definition}
The Best-Neighbor strategy $\mathcal{L}_{BN}(v)$ is defined as:
\begin{eqnarray}
\mathcal{L}_{BN}(v)
&\equiv&
\{ \, s \!\in\! S(v)
\mid
\forall \tilde{s} \!\in\! S(v) \, \
\lnot \, \mathcal{O}^{>}(\tilde{s}, s),
\end{eqnarray}
where $S(v)$ is the multiset of solutions of $v$'s neighbors.
\end{definition}

\subsection{Confident-Neighbor}
This paper introduces {\em confident-neighbor}, an alternative to the best-neighbor strategy which relies only on the single comparison oracle and which reduces to best-neighbor for interlocking pod networks. The confident neighbor strategy also represents a moderate level of social loafing \cite{karau_social_1993}, in which agents do not actively seek to improve their solution, but passively adopt better solutions if they are presented.

Confident neighbor proceeds in two stages. In the first stage, agents determine if their current solution is at least as good as all others in their neighborhood. If so, we call the agent {\em confident}. In the second stage, confident agents broadcast their solution to all of their neighbors. Non-confident agents choose randomly between any broadcast solutions they receive, or keep their original solution if they receive none.

\begin{definition}
The confident-neighbor strategy $\mathcal{L}_{CN}$ is defined as:
\begin{eqnarray}
C(v) &\equiv& \{
\, s(w)  \nonumber \\
&& \,
\mid
\forall w \!\in\! N(v) \, \forall u \!\in\! N(w) \,
\lnot \, \mathcal{O}^{>}_{s(w)}(s(u)) \,
\}
\\
\mathcal{L}_{\text{CN}}(v)
&\equiv& 
\begin{cases}
C(v) & \text{if } C(v) \neq \varnothing \\
\{ \, s(v) \, \}& \text{otherwise,}
\end{cases}
\end{eqnarray}
where $N(v)$ is the set of vertices neighboring $v$ and $s(v)$ is the current candidate solution for agent $v$.
\end{definition}

Confident-neighbor differs from best-neighbor because of two subtle but important considerations. First, an agent's neighbors need not be adjacent to each other. Second, an agent's neighbors can be adjacent to others outside that agent's neighborhood. As a result, an agent might receive zero, one, or many broadcasts. The exception is when agents belong to a clique (e.g., in network deliberation), in which case agents receive exactly one broadcast for each maximal solution in the clique, thus yielding the same results as best-neighbor. Confident-neighbor is thus more appropriate than best-neighbor for comparing network deliberation to single group deliberation, as the same information is utilized in both cases.

\subsection{Conform}
In some contexts, agents might rely on information other than solution quality in their learning strategies.
Agents might do so out of necessity if they are not able to compare arbitrary solutions.
Alternatively, agents might ignore solution quality due to social dynamics, e.g. social loafing \cite{karau_social_1993}. In these contexts, agents might instead evaluate solutions based on popularity, which produces the {\em conform} learning strategy.
When using the conform strategy, agents count the number of times each solution appears among their neighbors, and adopt the most popular.
\begin{definition}
The conform strategy $\mathcal{L}_{\text{C}}(v)$ is defined as:
\begin{eqnarray}
\mathcal{L}_{\text{C}}(v) &=&
\mode(S(v)),
\end{eqnarray}
where $S(v)$ is the multiset of candidate solutions for all vertices neighboring $v$, and $\mode()$ returns a set containing the mode or modes of a multiset.
\end{definition}
Note that the conform strategy does not depend on the objective function or any oracles, meaning it incorporates no new information about the quality of the solutions.

% Exclude bitwise majority from paper
\iffalse

\subsection{Bitwise Majority}
Instead of evaluating entire solutions on their popularity, it is also possible to evaluate the popularity of each component of a solution, i.e., each bit in the bit string \cite{platt_network_2018}.
We refer to this strategy as {\em bitwise majority} social learning.
As with previous strategies, bitwise majority can result in ties for individual bits, which result in ties for the resulting full solution.

\begin{definition}
The bitwise majority strategy $\mathcal{L}_{bitwise}$ is defined as:
\begin{eqnarray}
S_i &=& \bigcup_{x \in S} \{x_i\} \\
\mathcal{L}_{bitwise}(S)
&=& \mode(S_1) \times \mode(S_2) \ldots \times \mode(S_d),
\end{eqnarray}
where $x_i$ is the $i$th element of the binary string $x$,
$S_i$ is a multiset, and $\mode(S)$ returns a set containing the mode or modes of $S$, and $\times$ is the Cartesian product.
\end{definition}
As with conform, the bitwise majority strategy does not incorporate any new information about solution quality.
Bitwise majority also has the notable property of being able to produce novel solutions that vary significantly from any previously seen.

% End of bitwise majority
\fi

\subsection{Combining Learning Strategies}\label{subsubsec:combine}
Models of social learning often combine truly social strategies with individual learning \cite{lazer_network_2007, barkoczi_social_2016,gomez_clustering_2019}.
We find that the method used to combine social and individual strategies can result in a notable difference in outcome. As a robustness check, we consider two cases for all of our analyses:
\begin{description}
\iffalse
\item[Serial]{
Agents first perform individual learning to produce an intermediate solution, and then apply social learning to the intermediate solutions. This method does not rely on any of the solution comparison oracles.}
\fi
\item[Parallel]{
Agents apply both social and individual learning strategies to their initial solution to produce two competing intermediate solutions, then adopt the better of the two. This method relies on the arbitrary comparison oracle.}
\item[Fallback]{
Agents first apply social learning to the current solution, but "fall back" to individual learning if the result is not an improvement on the original solution \cite{lazer_network_2007, barkoczi_social_2016,gomez_clustering_2019}. Fallback relies on the single comparison oracle and could be motivated by limited agent capability or social loafing.}
\end{description}

\iffalse
We must also consider {\em criticality} \cite{barkoczi_social_2016, rendell_rogersparadox_2010}.
After combining individual and social learning according to one of the above methods, non-critical agents immediately adopt the result.
Critical agents, on the other hand, compare the quality of the new solution to their previous solution, only adopting the new one if it provides an improvement.
Note that critical behavior relies on information about solution quality and requires single-solution comparison capability, or stronger.
\fi

\section{NK Model}
\label{subsec:task}
The NK Model \cite{kauffman_towards_1987} is an optimization problem particularly well-suited for modelling complex tasks. The model is used to create a fitness function $Q(s)$ over some discrete space $\mathcal{S}$, typically binary strings of a fixed length.
The model is parameterized by two variables.
The first, $N$, is the dimension of the solution space, i.e., the length of each binary string.
The second parameter, $K$, determines the ``ruggedness'' of the fitness function, i.e., the number of local maxima.
In effect, the $K$ parameter allows the complexity of the optimization problem to be tuned.
The ability to tune task complexity makes the NK model well-suited for studying the role of complexity in various settings.
The construction of an NK fitness function from a set of parameters is a stochastic process.
So particular values of $N$ and $K$ define a class of fitness functions, which can be sampled to produce a specific fitness function.

We now show the construction of the NK model fitness function.
A class of NK model fitness functions can be defined by a tuple of integer parameters
$(N, K)$ such that $N > 0$ and $0 \leq K < N$.
We begin by defining $N$ fitness contributions functions:
\begin{eqnarray}
q_i &:& \mathbb{Z}^{K+1} \rightarrow [0, 1].
\end{eqnarray}
The value of each $q_i(x)$ is chosen uniformly at random in the range $[0, 1]$ at the time the function is defined.
We also define $N$ projection operators $P_i$ which select $K+1$ bits from a length-$N$ bit string.
Each $P_i$ selects the bit at index $i$ and $K$ other indices, chosen uniformly at random at the time $P_i$ is constructed.
For a solution $s$, the $i$th fitness contribution is evaluated on the $K + 1$ bits selected by the $i$th projection: $q_i(P_i(s))$.
The value of the fitness function is the mean of all fitness contributions:
\begin{eqnarray}
Q(s) &=& \frac{1}{N}\sum_{i=1}^N q_i(P_i(s)).
\end{eqnarray}

The parameter $K$ alters the ruggedness of the fitness function by controlling the interdependence between the $q_i$.
When $K = 0$, each $q_i$ depends on a single unique index of $s$,
allowing the $q_i$ to be optimized independently.
In this case, $Q(s)$ has a single maximum: the global maximum.
However, for $K > 0$, two things happen: it becomes possible for each $q_i$ to have multiple local maxima, and some of the $q_i$ become coupled due to dependence on the same indices of $s$.
The result is that as $K$ increases, both the number of local maxima of $Q(s)$ \cite{weinberger_local_1991} and the difficulty of simultaneously optimizing the $q_i$ increase.
In other words, $Q$ becomes more rugged, and more complex.

The distribution of local maximum values is asymptotically normal for large $K$ \cite{weinberger_local_1991}.
When a skewed distribution is preferred, it is common to exponentiate the value of $Q(s)$ as a final step \cite{lazer_network_2007, barkoczi_social_2016, gomez_clustering_2019}.

\section{Supplemental Results}

Figures \ref{fig:results-frac-parallel} and \ref{fig:results-frac-fallback} show pairwise comparisons of solution quality between strategy/network settings. Figures \ref{fig:results-bn-parallel-dist}--\ref{fig:results-conf-fallback-dist} show the complete solution quality distributions for all settings.
Tables \ref{tab:t-instrat-fallback}--\ref{tab:t-innet-parallel} show the t-values and Bonferroni-corrected p-values for comparisons across networks and strategies.

\begin{figure}
    \label{fig:results-frac-parallel}
    \centering
    \includegraphics[width=3.33in]{fig-result-frac-parallel.eps}
\caption{Fraction of simulations in which the row condition outperforms the column condition. Results are shown for parallel individual learning. In the case of a tie, weight is divided evenly.}
\end{figure}

\begin{figure}
    \label{fig:results-frac-fallback}
    \centering
    \includegraphics[width=3.33in]{fig-result-frac-parallel.eps}
\caption{Fraction of simulations in which the row condition outperforms the column condition. Results are shown for fallback individual learning. In the case of a tie, weight is divided evenly.}
\end{figure}

\begin{figure}
    \label{fig:results-bn-parallel-dist}
    \centering
    \includegraphics[width=3.33in]{results-bn-parallel-dist.eps}
\caption{Solution quality distribution for best-neighbor strategy with parallel individual learning.}
\end{figure}
\begin{figure}
    \label{fig:results-cn-parallel-dist}
    \centering
    \includegraphics[width=3.33in]{results-cn-parallel-dist.eps}
\caption{Solution quality distribution for confident-neighbor strategy with parallel individual learning.}
\end{figure}
\begin{figure}
    \label{fig:results-conf-parallel-dist}
    \centering
    \includegraphics[width=3.33in]{results-conf-parallel-dist.eps}
\caption{Solution quality distribution for the conform strategy with parallel individual learning.}
\end{figure}

\begin{figure}
    \label{fig:results-bn-fallback-dist}
    \centering
    \includegraphics[width=3.33in]{results-bn-fallback-dist.eps}
\caption{Solution quality distribution for best-neighbor strategy with fallback individual learning.}
\end{figure}
\begin{figure}
    \label{fig:results-cn-fallback-dist}
    \centering
    \includegraphics[width=3.33in]{results-cn-fallback-dist.eps}
\caption{Solution quality distribution for confident-neighbor strategy with fallback individual learning.}
\end{figure}
\begin{figure}
    \label{fig:results-conf-fallback-dist}
    \centering
    \includegraphics[width=3.33in]{results-conf-fallback-dist.eps}
\caption{Solution quality distribution for the conform strategy with fallback individual learning.}
\end{figure}

\begin{table*}[]
    \label{tab:t-instrat-fallback}
    \centering
    \begin{tabular}{l|ll|ccc}
        Strategy & Net A & Net B & \overline{B - A} & t & p* \\
        \hline
Best Neighbor&Pref. Attach.&Small World&-8.70e-02&-1.33e+01&1.27e-35\\
Best Neighbor&Pref. Attach.&Long Path&-5.98e-02&-8.62e+00&1.51e-15\\
Best Neighbor&Pref. Attach.&Random Pod&-1.15e-02&-1.83e+00&4.01e+00\\
Best Neighbor&Small World&Pref. Attach.&8.70e-02&1.33e+01&1.27e-35\\
Best Neighbor&Small World&Long Path&2.72e-02&4.53e+00&3.88e-04\\
Best Neighbor&Small World&Random Pod&7.55e-02&1.18e+01&1.66e-28\\
Best Neighbor&Long Path&Pref. Attach.&5.98e-02&8.62e+00&1.51e-15\\
Best Neighbor&Long Path&Small World&-2.72e-02&-4.53e+00&3.88e-04\\
Best Neighbor&Long Path&Random Pod&4.83e-02&7.34e+00&2.69e-11\\
Best Neighbor&Random Pod&Pref. Attach.&1.15e-02&1.83e+00&4.01e+00\\
Best Neighbor&Random Pod&Small World&-7.55e-02&-1.18e+01&1.66e-28\\
Best Neighbor&Random Pod&Long Path&-4.83e-02&-7.34e+00&2.69e-11\\
\hline
Confident Neighbor&Pref. Attach.&Small World&-7.52e-02&-1.13e+01&2.94e-26\\
Confident Neighbor&Pref. Attach.&Long Path&2.39e-03&3.64e-01&4.30e+01\\
Confident Neighbor&Pref. Attach.&Random Pod&5.80e-02&8.88e+00&1.86e-16\\
Confident Neighbor&Small World&Pref. Attach.&7.52e-02&1.13e+01&2.94e-26\\
Confident Neighbor&Small World&Long Path&7.76e-02&1.21e+01&1.08e-29\\
Confident Neighbor&Small World&Random Pod&1.33e-01&1.92e+01&1.36e-68\\
Confident Neighbor&Long Path&Pref. Attach.&-2.39e-03&-3.64e-01&4.30e+01\\
Confident Neighbor&Long Path&Small World&-7.76e-02&-1.21e+01&1.08e-29\\
Confident Neighbor&Long Path&Random Pod&5.56e-02&8.50e+00&4.00e-15\\
Confident Neighbor&Random Pod&Pref. Attach.&-5.80e-02&-8.88e+00&1.86e-16\\
Confident Neighbor&Random Pod&Small World&-1.33e-01&-1.92e+01&1.36e-68\\
Confident Neighbor&Random Pod&Long Path&-5.56e-02&-8.50e+00&4.00e-15\\
\hline
Conform&Pref. Attach.&Small World&3.60e-02&6.78e+00&1.24e-09\\
Conform&Pref. Attach.&Long Path&-1.63e-01&-2.22e+01&2.25e-87\\
Conform&Pref. Attach.&Random Pod&-2.28e-01&-3.19e+01&9.08e-153\\
Conform&Small World&Pref. Attach.&-3.60e-02&-6.78e+00&1.24e-09\\
Conform&Small World&Long Path&-1.99e-01&-3.17e+01&1.59e-151\\
Conform&Small World&Random Pod&-2.64e-01&-4.43e+01&2.20e-236\\
Conform&Long Path&Pref. Attach.&1.63e-01&2.22e+01&2.25e-87\\
Conform&Long Path&Small World&1.99e-01&3.17e+01&1.59e-151\\
Conform&Long Path&Random Pod&-6.48e-02&-8.31e+00&1.79e-14\\
Conform&Random Pod&Pref. Attach.&2.28e-01&3.19e+01&9.08e-153\\
Conform&Random Pod&Small World&2.64e-01&4.43e+01&2.20e-236\\
Conform&Random Pod&Long Path&6.48e-02&8.31e+00&1.79e-14\\
\hline
    \end{tabular}
    \caption{Two-tailed paired t-values across networks, within strategies. Results are for fallback individual learning. For all rows, dof=999 and $p^*$ is the Bonferroni-corrected p-value with m=60.}
    \label{tab:my_label}
\end{table*}

\begin{table*}[]
    \label{tab:t-innet-fallback}
    \centering
    \begin{tabular}{l|ll|ccc}
        Network & Strategy A & Strategy B & \overline{B - A} & t & p* \\
        \hline
Pref. Attach.&Best Neighbor&Confident Neighbor&-6.43e-02&-9.39e+00&2.33e-18\\
Pref. Attach.&Best Neighbor&Conform&8.96e-02&1.19e+01&6.83e-29\\
Pref. Attach.&Confident Neighbor&Best Neighbor&6.43e-02&9.39e+00&2.33e-18\\
Pref. Attach.&Confident Neighbor&Conform&1.54e-01&2.06e+01&7.57e-77\\
Pref. Attach.&Conform&Best Neighbor&-8.96e-02&-1.19e+01&6.83e-29\\
Pref. Attach.&Conform&Confident Neighbor&-1.54e-01&-2.06e+01&7.57e-77\\
\hline
Small World&Best Neighbor&Confident Neighbor&-5.26e-02&-8.39e+00&9.70e-15\\
Small World&Best Neighbor&Conform&2.13e-01&3.52e+01&4.46e-175\\
Small World&Confident Neighbor&Best Neighbor&5.26e-02&8.39e+00&9.70e-15\\
Small World&Confident Neighbor&Conform&2.65e-01&4.45e+01&1.24e-237\\
Small World&Conform&Best Neighbor&-2.13e-01&-3.52e+01&4.46e-175\\
Small World&Conform&Confident Neighbor&-2.65e-01&-4.45e+01&1.24e-237\\
\hline
Long Path&Best Neighbor&Confident Neighbor&-2.12e-03&-3.42e-01&4.39e+01\\
Long Path&Best Neighbor&Conform&-1.37e-02&-1.71e+00&5.20e+00\\
Long Path&Confident Neighbor&Best Neighbor&2.12e-03&3.42e-01&4.39e+01\\
Long Path&Confident Neighbor&Conform&-1.16e-02&-1.44e+00&8.96e+00\\
Long Path&Conform&Best Neighbor&1.37e-02&1.71e+00&5.20e+00\\
Long Path&Conform&Confident Neighbor&1.16e-02&1.44e+00&8.96e+00\\
\hline
Random Pod&Best Neighbor&Confident Neighbor&5.18e-03&8.40e-01&2.41e+01\\
Random Pod&Best Neighbor&Conform&-1.27e-01&-1.64e+01&1.07e-51\\
Random Pod&Confident Neighbor&Best Neighbor&-5.18e-03&-8.40e-01&2.41e+01\\
Random Pod&Confident Neighbor&Conform&-1.32e-01&-1.67e+01&1.39e-53\\
Random Pod&Conform&Best Neighbor&1.27e-01&1.64e+01&1.07e-51\\
Random Pod&Conform&Confident Neighbor&1.32e-01&1.67e+01&1.39e-53\\
\hline
    \end{tabular}
    \caption{Two-tailed paired t-values across strategies, within networks. Results are for fallback individual learning. For all rows, dof=999 and $p^*$ is the Bonferroni-corrected p-value with m=60.}
    \label{tab:my_label}
\end{table*}

\begin{table*}[]
    \label{tab:t-instrat-parallel}
    \centering
    \begin{tabular}{l|ll|ccc}
        Strategy & Net A & Net B & \overline{B - A} & t & p* \\
    \hline
Best Neighbor&Pref. Attach.&Small World&-6.75e-02&-1.13e+01&2.48e-26\\
Best Neighbor&Pref. Attach.&Long Path&-4.24e-02&-6.78e+00&1.25e-09\\
Best Neighbor&Pref. Attach.&Random Pod&-8.14e-03&-1.32e+00&1.12e+01\\
Best Neighbor&Small World&Pref. Attach.&6.75e-02&1.13e+01&2.48e-26\\
Best Neighbor&Small World&Long Path&2.50e-02&4.25e+00&1.42e-03\\
Best Neighbor&Small World&Random Pod&5.93e-02&9.38e+00&2.54e-18\\
Best Neighbor&Long Path&Pref. Attach.&4.24e-02&6.78e+00&1.25e-09\\
Best Neighbor&Long Path&Small World&-2.50e-02&-4.25e+00&1.42e-03\\
Best Neighbor&Long Path&Random Pod&3.43e-02&5.36e+00&6.14e-06\\
Best Neighbor&Random Pod&Pref. Attach.&8.14e-03&1.32e+00&1.12e+01\\
Best Neighbor&Random Pod&Small World&-5.93e-02&-9.38e+00&2.54e-18\\
Best Neighbor&Random Pod&Long Path&-3.43e-02&-5.36e+00&6.14e-06\\
\hline
Confident Neighbor&Pref. Attach.&Small World&-4.49e-02&-7.20e+00&7.03e-11\\
Confident Neighbor&Pref. Attach.&Long Path&-5.03e-03&-7.65e-01&2.67e+01\\
Confident Neighbor&Pref. Attach.&Random Pod&2.61e-02&4.02e+00&3.74e-03\\
Confident Neighbor&Small World&Pref. Attach.&4.49e-02&7.20e+00&7.03e-11\\
Confident Neighbor&Small World&Long Path&3.99e-02&6.96e+00&3.80e-10\\
Confident Neighbor&Small World&Random Pod&7.10e-02&1.18e+01&1.98e-28\\
Confident Neighbor&Long Path&Pref. Attach.&5.03e-03&7.65e-01&2.67e+01\\
Confident Neighbor&Long Path&Small World&-3.99e-02&-6.96e+00&3.80e-10\\
Confident Neighbor&Long Path&Random Pod&3.11e-02&4.89e+00&7.03e-05\\
Confident Neighbor&Random Pod&Pref. Attach.&-2.61e-02&-4.02e+00&3.74e-03\\
Confident Neighbor&Random Pod&Small World&-7.10e-02&-1.18e+01&1.98e-28\\
Confident Neighbor&Random Pod&Long Path&-3.11e-02&-4.89e+00&7.03e-05\\
\hline
Conform&Pref. Attach.&Small World&3.67e-02&6.85e+00&7.73e-10\\
Conform&Pref. Attach.&Long Path&-1.73e-01&-2.40e+01&7.10e-99\\
Conform&Pref. Attach.&Random Pod&-2.77e-01&-4.08e+01&6.29e-213\\
Conform&Small World&Pref. Attach.&-3.67e-02&-6.85e+00&7.73e-10\\
Conform&Small World&Long Path&-2.09e-01&-3.46e+01&3.46e-171\\
Conform&Small World&Random Pod&-3.14e-01&-5.68e+01&1.63e-313\\
Conform&Long Path&Pref. Attach.&1.73e-01&2.40e+01&7.10e-99\\
Conform&Long Path&Small World&2.09e-01&3.46e+01&3.46e-171\\
Conform&Long Path&Random Pod&-1.04e-01&-1.43e+01&8.55e-41\\
Conform&Random Pod&Pref. Attach.&2.77e-01&4.08e+01&6.29e-213\\
Conform&Random Pod&Small World&3.14e-01&5.68e+01&1.63e-313\\
Conform&Random Pod&Long Path&1.04e-01&1.43e+01&8.55e-41\\
    \hline
    \end{tabular}
    \caption{Two-tailed paired t-values across networks, within strategies. Results are for parallel individual learning. For all rows, dof=999 and $p^*$ is the Bonferroni-corrected p-value with m=60.}
    \label{tab:my_label}
\end{table*}

\begin{table*}[]
    \label{tab:t-innet-parallel}
    \centering
    \begin{tabular}{l|ll|ccc}
        Network & Strategy A & Strategy B & \overline{B - A} & t & p* \\
    \hline
Pref. Attach.&Best Neighbor&Confident Neighbor&-3.78e-02&-5.94e+00&2.39e-07\\
Pref. Attach.&Best Neighbor&Conform&2.00e-01&2.77e+01&3.28e-124\\
Pref. Attach.&Confident Neighbor&Best Neighbor&3.78e-02&5.94e+00&2.39e-07\\
Pref. Attach.&Confident Neighbor&Conform&2.38e-01&3.40e+01&1.67e-167\\
Pref. Attach.&Conform&Best Neighbor&-2.00e-01&-2.77e+01&3.28e-124\\
Pref. Attach.&Conform&Confident Neighbor&-2.38e-01&-3.40e+01&1.67e-167\\
\hline
Small World&Best Neighbor&Confident Neighbor&-1.52e-02&-2.74e+00&3.80e-01\\
Small World&Best Neighbor&Conform&3.05e-01&5.30e+01&1.25e-290\\
Small World&Confident Neighbor&Best Neighbor&1.52e-02&2.74e+00&3.80e-01\\
Small World&Confident Neighbor&Conform&3.20e-01&5.89e+01&0.00e+00\\
Small World&Conform&Best Neighbor&-3.05e-01&-5.30e+01&1.25e-290\\
Small World&Conform&Confident Neighbor&-3.20e-01&-5.89e+01&0.00e+00\\
\hline
Long Path&Best Neighbor&Confident Neighbor&-3.88e-04&-6.64e-02&5.68e+01\\
Long Path&Best Neighbor&Conform&7.02e-02&9.34e+00&3.64e-18\\
Long Path&Confident Neighbor&Best Neighbor&3.88e-04&6.64e-02&5.68e+01\\
Long Path&Confident Neighbor&Conform&7.06e-02&9.15e+00&1.83e-17\\
Long Path&Conform&Best Neighbor&-7.02e-02&-9.34e+00&3.64e-18\\
Long Path&Conform&Confident Neighbor&-7.06e-02&-9.15e+00&1.83e-17\\
\hline
Random Pod&Best Neighbor&Confident Neighbor&-3.59e-03&-5.46e-01&3.51e+01\\
Random Pod&Best Neighbor&Conform&-6.84e-02&-9.54e+00&6.11e-19\\
Random Pod&Confident Neighbor&Best Neighbor&3.59e-03&5.46e-01&3.51e+01\\
Random Pod&Confident Neighbor&Conform&-6.48e-02&-8.96e+00&9.43e-17\\
Random Pod&Conform&Best Neighbor&6.84e-02&9.54e+00&6.11e-19\\
Random Pod&Conform&Confident Neighbor&6.48e-02&8.96e+00&9.43e-17\\
    \hline
    \end{tabular}
    \caption{Two-tailed paired t-values across strategies, within networks. Results are for parallel individual learning. For all rows, dof=999 and $p^*$ is the Bonferroni-corrected p-value with m=60.}
    \label{tab:my_label}
\end{table*}

\bibliography{references}
\bibliographystyle{plain}

\end{document}

### 2021-09-14 NetDelibSim/main.tex
Move NetDelibSim to subdirectory.

\documentclass[twocolumn,10pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{times}
\usepackage{geometry}

\usepackage[switch]{lineno}
\linenumbers

\newtheorem{definition}{Definition}
\newtheorem{property}{Property}
\newtheorem{claim}{Claim}

\DeclareMathOperator{\unif}{unif}
\DeclareMathOperator{\sample}{sample}
\DeclareMathOperator{\setcount}{count}
\DeclareMathOperator{\mode}{mode}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

\usepackage{geometry}
\geometry{left=1.5cm,right=1.5cm,top=2cm,bottom=2cm}
\setlength{\columnsep}{0.5cm}

\title{Small interlocking groups improve mass deliberation\\in the presence of strong social influence}
\author{
%Edward L. Platt\\
%University of Michigan\\
%elplatt@umich.edu
%and
%Herminio Bodon\\
%Northwestern University\\
%HerminioBodon2020\\@u.northwestern.edu
%\and
%Daniel M. Romero\\
%University of Michigan\\
%drom@umich.edu
Authors Redacted \\
for \\
Double-Blind Peer-Review
}
\date{\today}

\begin{document}

\maketitle
\begin{abstract}
Deliberation, a form of collective problem-solving, is a key component of democracy, social movements, and online peer-production.
The network structure of interpersonal communication plays an important role in collective problem-solving, particularly in deliberation, which can become prohibitively complex and time-intensive at large scales.
Network properties such as structural efficiency and degree distribution have been shown to influence the speed and quality of collective problem-solving, in combination with the behavioral strategies employed by individuals.
However, many successful examples of mass deliberation differ in structure from common models of collective problem-solving by exhibiting networks of small interlocking groups.
In this paper, we present an agent-based model of deliberation in networks of small interlocking groups and compare performance on these networks with conventional network structures.
We find that networks of small interlocking groups improve solution quality when problem-solvers exhibit strong social influence.
This effect is compatible with, but distinct from, the previously observed benefit of network efficiency in the presence of strong social influence.
Our findings suggest a possible mechanism contributing to the success of existing mass deliberative projects as well as a principle for the design of new projects.
By contributing to more effective deliberation at larger scales, we hope this work will contribute to democratizing the governance of large sociotechnical systems.

\end{abstract}

\section{Introduction}

Deliberation enables the generation, aggregation, and synthesis of diverse knowledge \cite{dewey_creative_1940, anderson_epistemology_2006, ackerman_deliberation_2002, ostrom_collective_2000},
as well as the early identification and resolution of conflict \cite{gentry_consensus_1982}.
As a participatory process, deliberation builds trust among participants, increases the perception of fairness, and incentivises cooperation with group decisions
\cite{ostrom_collective_2000, bowles_endogenous_1998}.
As a dynamic process, deliberation alters private preferences through discourse \cite{habermas_structural_1991}, sidestepping fundamental limitations of voting, e.g., the Condorcet paradox \cite{condorcet_essay_1785, brandt_computational_2012} and Arrow's impossibility theorem \cite{arrow_social_2012, brandt_computational_2012}.
Despite these benefits, deliberation becomes prohibitively difficult in larger groups, due to the increased time and effort needed to reach decisions \cite{fishkin_voice_1997, gentry_consensus_1982} and to the emergence of power inequalities \cite{freeman_tyranny_1972, boehm_egalitarian_1993}.
Yet examples of successful mass deliberation do exist, including Wikipedia \cite{giles_internet_2005, keegan_evolution_2017}, free and open source software projects \cite{benkler_coases_2002}, grassroots protest and crisis-response networks \cite{gonzalez-bailon_networked_2016, brugh_combining_2019} and self-managed organizations \cite{laloux_reinventing_2014}.
The commonalities between such projects provide insight into overcoming the challenges of mass deliberation.
Specifically, we evaluate a model of mass deliberation inspired by common network features observed in successful projects.

Communication in large collaborations is necessarily restricted when group size exceeds individual members' capacity for communication. The structure of the resulting communication network is an important factor in the success of collaborative tasks \cite{kearns_experiments_2012, mason_collaborative_2012, mason_propagation_2008}, such as deliberation. Among successful mass deliberative projects, communication networks often exhibits a common structure: small interlocking groups \cite{benkler_coases_2002, gonzalez-bailon_networked_2016, brugh_combining_2019, laloux_reinventing_2014}. A similar structure has been observed in the interlocking directorates of corporate boards \cite{karau_social_1993} and the interlocking publics of the public sphere \cite{habermas_structural_1991}, as well as utilized for collective design in network rotation \cite{salehi_hive_2018}.
These interlocking groups have been given various names: committees, working groups, projects, modules, affinity groups, circles, teams, cores, nodes, zones, cells, or syndicates.
We shall use the term {\em pods} to encompass all groups exhibiting two key characteristics: small enough to enable deliberation, and interlocking to enable information diffusion.
We shall use the term {\em network deliberation} to refer to mass deliberation using an interlocking pod structure.
Communication network structure also influences the success of collaborations by controlling the speed of information diffusion \cite{lazer_network_2007,mason_propagation_2008,mason_collaborative_2012,barkoczi_social_2016, gomez_clustering_2019,derex_partial_2016,kearns_experiments_2012}. Structurally efficient networks enable fast diffusion and favor exploitation of known information over exploration of the unknown \cite{lazer_network_2007, barkoczi_social_2016}. In the context of network deliberation, structural efficiency is determined by how individuals are assigned to pods.
We will focus on two such methods: one efficient, one inefficient.

Collective tasks such as deliberation are influenced not only by communication structure, but also by individual behavior. Individuals weigh social influence against their personal knowledge and expertise  \cite{boyd_culture_1988}. 
Even in a collective setting, individuals sometimes work independently to search for new and innovative ideas \cite{lazer_network_2007, barkoczi_social_2016}, to utilize their unique information and perspectives \cite{hong_interpreted_2009}, and to critically evaluate the ideas of others \cite{rendell_rogersparadox_2010}.
Alternatively, individuals can defer to others in order to exploit and propagate good ideas \cite{barkoczi_social_2016, boyd_culture_1988}, to conserve their own resources through free-riding and social loafing \cite{karau_social_1993}, or due to a high level of trust \cite{salehi_hive_2018} or social influence \cite{banerjee_simple_1992, smith_pathological_2000}. Individual behavior interacts with network structure to vary the level of exploration of new ideas versus the exploitation of known information \cite{barkoczi_social_2016}. While the focus of network deliberation is structural, the effect of network structure must be understood in the context of individual behavior and social dynamics.

We model network deliberation using a social learning framework to address two questions. How effective is network deliberation compared to conventional single-group deliberation? And how do network efficiency and individual behavior influence the outcome of network deliberation? We present the following three contributions.
First, network deliberation out-performs conventional deliberation in both solution speed and quality when agents rely heavily on social learning. This suggests that network deliberation is a promising deliberation design for groups with a high degree of social influence. Second, network deliberation performs best on structurally efficient networks when agents exhibit conformist behavior, while it performs best on structurally inefficient networks when agents greedily maximize solution quality, consistent with findings for conventional networks \cite{barkoczi_social_2016}.
Third, the naive greedy strategy can be modified to improve performance on conventional networks despite using strictly less information. In the case of network deliberation, the naive and modified greedy strategies are equivalent.

Deliberation can be modeled as a form of collective problem-solving and social learning in which agents hold private beliefs, share information with their neighbors, and update their beliefs based on a combination of individual analysis and social influence.
Following existing models of collective problem-solving \cite{lazer_network_2007, barkoczi_social_2016, gomez_clustering_2019}, our model comprises: 1. a collective task, 2. a network, and 3. a learning strategy.
Agents maintain a candidate solution to the task and iteratively apply their learning strategy to the available information, yielding an updated candidate.

As their collective task, agents seek an optimal point on a rugged fitness landscape.
This landscape is generated using the tuneably-rugged NK model \cite{kauffman_towards_1987, weinberger_local_1991} (see Methods for more details). In this case, task solutions take the form of binary strings.

To model network deliberation, we construct interlocking networks in stages. At each stage, agents are partitioned into small {\em pods} and connected to all other agents in the same pod.
Our choice of pod size is informed by real-world The upper limit of effective small group size has been estimated variously as five \cite{freeman_tyranny_1972}, five--eight\cite{lohman_designing_2000}, and eight \cite{miflin_small_2004}.
We choose a pod size of 5.
Different methods for assigning agents to pods produce networks with different properties.
Interlocking structure is created when agents are placed with others they have not yet interacted with.
We thus require that pod assignment methods satisfy the following mixing property:
\begin{property}
\label{prop:mixing}
For any agent $v$, and any two assignment stages $t$, $t^\prime$, the probability that at least one neighbor of $v$ at stage $t$ is also a neighbor at stage $t^\prime$ approaches 0 as $|V|$ grows.
\end{property}
We study two such methods: a random assignment method that produces structurally efficient networks, and a long-path method that produces inefficient networks. For comparison, we also simulate conventional deliberation. While the potential communication links in conventional deliberation form a fully connected network, actual communication is better modeled by preferential attachment \cite{barabasi_emergence_1999} or small-world \cite{watts_collective_1998} networks (see Methods).

An agent's learning strategy models both informational constraints and behavioral dynamics \cite{lazer_network_2007, barkoczi_social_2016} such as social influence and social loafing \cite{karau_social_1993}.
We consider three social learning strategies.
The {\em best-neighbor} strategy is the naive greedy strategy: agents simply adopt the neighboring solution with the highest value. This strategy requires that agents have the ability to compare the quality of two arbitrary solutions, a strong assumption representing either skilled agents, or a simple task.
The {\em conform} strategy is pure social influence. Agents adopt the most common solution among their neighbors.
This strategy requires no knowledge of solution quality, modeling contexts where agents are unable to compare solution quality, or where agents choose not to, as in social loafing \cite{karau_social_1993}.
We also introduce the {\em confident-neighbor} strategy, a variant of best-neighbor in which agents announce when they have the best solution in their neighborhood. This variant makes weaker assumptions about agent ability (see Methods for more detail) but is equivalent to best-neighbor when network deliberation is used.
This strategy allows us to compare low-social influence and high-social influence strategies under equivalent assumptions of agent ability.

Between each stage of social learning, agents may perform individual learning. Following existing literature \cite{lazer_network_2007, barkoczi_social_2016}, we model individual learning by mutating a single bit of an agent's solution, and keeping it if the mutation improves solution quality.
A complete learning strategy must also specify how social and individual learning are integrated. Existing literature \cite{lazer_network_2007, barkoczi_social_2016} applies individual learning only when social learning fails to produce an improved solution. We refer to this method as {\em fallback} individual learning. Fallback can be also be interpreted as a model of social loafing \cite{karau_social_1993}. For completeness, we include analysis for a second method in which agents first perform social and individual learning in {\em parallel} and then choose the better of the two learning outputs (See Methods for more detail).

We find that network deliberation identifies higher quality solutions than conventional deliberation when agents use the conform strategy, while requiring less time to converge. Within network deliberation, we find that the structurally efficient random pod network outperforms the structurally inefficient long-path network when agents use the conform strategy. However, when either best-neighbor or confident-neighbor is used, the inefficient network is preferable, consistent with findings for conventional networks \cite{barkoczi_social_2016}. We also find that the confident-neighbor strategy matches or outperforms best-neighbor across all networks, despite using strictly less information about solution quality.

\begin{figure*}[]
    \label{fig:results}
    \centering
    \begin{minipage}{3.33in}
        \centering
        \includegraphics[width=3.33in]{fig-perf-both.eps}
    \end{minipage}%
    \begin{minipage}{1.875in}
    \centering
    \includegraphics[width=1.875in]{fig-converge-both.eps}
    \end{minipage}
\caption{
Solution quality and convergence time for agent-based simulations of different learning strategies. Error bars represent 95\% confidence interval. Results for network deliberation are shown shaded.
}
\end{figure*}

\section{Results}
Here we present the results of 1000 runs of an agent-based model of deliberation. In each run, an NK model is generated (N=15, K=6) and the model is run for 300 iterations for each network/strategy combination (see Methods for more details). The results of these simulations are shown in Figure \ref{fig:results}.

\subsection{Network vs. Conventional Deliberation}
We are particularly interested in comparing the performance of network deliberation to that of conventional single-group deliberation. When the conform strategy is used, we find that both of the network deliberation conditions yield better performance than conventional networks (Figure \ref{fig:results-conform-diff}). Analysis of the solution quality distributions (Figure \ref{fig:results-conform-dist}) shows two distinct effects. First, network deliberation frequently finds maximal solutions while conventional deliberation does not. Second, even when non-maximal solutions are found, network deliberation finds higher-quality solutions. This benefit is robust across both parallel and fallback individual learning. The effect size is dramatic. Conventional networks combined with the conform strategy show the lowest performance of all conditions, while random pod network deliberation combined with the conform strategy shows higher performance than nearly all other conditions. The exception being small world network combined with the best/confident-neighbor strategy, which shows equal performance within our statistical resolution. 
Often, increased performance comes at the expense of speed as more time is devoted to exploring novel solutions. However, we find that network deliberation improves both performance and speed when the conform strategy is used, suggesting that interlocking network structure can enable more efficient exploration.

\begin{figure}
    \label{fig:results-conform-diff}
    \centering
    \includegraphics[width=3.33in]{fig-results-netdelib-diff.eps}
\caption{Difference in solution quality between network deliberation and conventional deliberation. Plots for Best/Confident-Neighbor show results for long-path and small-world networks, while the plot for conform shows results for random-pod and preferential-attachment networks. Results are shown for parallel individual learning.}
\end{figure}

\begin{figure}
    \label{fig:results-conform-dist}
    \centering
    \includegraphics[width=3.33in]{fig-results-conform-dist.eps}
\caption{Distributions of solution quality for social learning on a preferential-attachment network and random-pod network deliberation. Results are shown for parallel individual learning.}
\end{figure}

The conform strategy models contexts in which individuals rely heavily on social influence. Such contexts can arise from pro-conformity social norms, high levels of trust, social loafing, and limited information. As a form of rational ignorance, conformist behavior can conserve individual resources, but poses the risk of forming information cascades and propagating misinformation and inferior or harmful innovations \cite{banerjee_simple_1992}. Our simulations show that network deliberation can greatly improve the outcome of deliberation in the presence of strong social influence. This insight suggests one possible mechanism behind the success of mass deliberative projects, as well as the potential for network-based interventions in mass deliberation.

\subsection{Structural Efficiency}

The speed at which information can move through a communication network, i.e., it's structural efficiency, has been found to influence the success of collective problem-solving. We might expect the same to be true in network deliberation. In network deliberation, network structure is created through group membership. We compare the results of two group assignment methods: a random pod assignment yielding an efficient network, and a long-path method yielding a less efficient network (Figure \ref{fig:results-netdelib-efficiency}). We find that for greedy strategies (best-neighbor and confident-neighbor), the inefficient network yields the better results. The results are opposite for the conform strategy, with the efficient network yielding better results. These results are consistent with those observed for conventional single-group networks \cite{barkoczi_social_2016}, but to the best of our knowledge, they had not been explored for interlocking network like the ones we use to model network deliberation.

\begin{figure}
    \label{fig:results-netdelib-efficiency}
    \centering
    \includegraphics[width=3.33in]{fig-results-netdelib-efficiency.eps}
\caption{Difference in solution quality between random-pod (efficient) and long-path (inefficient) network. Results are shown for parallel individual learning.}
\end{figure}

The role of structural efficiency in collective tasks can be partially attributed to diversity \cite{lazer_network_2007, hong_groups_2004}. Greedy strategies such as best/confident-neighbor result in agents adopting the highest quality solution they have seen so far and discarding all others. Such strategies quickly reduce the number of solutions present in the population, quickly identifying a local maximum at the cost of diversity. As a result, agents duplicate efforts during individual learning, exploring mutations of the same solution. Inefficient networks can mitigate this effect by slowing the spread of high-quality solutions and allowing time for the exploration of new and potentially better solutions. Our results show exactly this, with long-path pod assignment improving the performance of greedy algorithms. The improved performance of the conform strategy in random pod assignment can be attributed to the converse of the above. When social influence is strong, efficient networks can interrupt information cascades by introducing new information from far across the network.
Our results are thus generally consistent with existing theory on structural efficiency and collective problem-solving.
However, we note an interesting deviation: under the conform strategy, long-path network deliberation out-performs the two conventional networks, both of which are more efficient.

\subsection{Confident-Neighbor Strategy}

In general, the best-neighbor strategy requires strong assumptions about agents' ability to evaluate solutions. In the special case of network deliberation, the assumptions can be weakened by applying a variant of the strategy, which has not been previously studied. The variant, which we call {\em confident-neighbor}, is equivalent to best-neighbor for network deliberation, but not in general (see Methods for more detail). Confident-neighbor thus provides a comparison between network deliberation and conventional single-group deliberation under weaker informational assumptions than best-neighbor. Surprisingly, the weaker assumptions of confident-neighbor lead to higher performance on the preferential attachment network, and similar performance on the-small world network (within our statistical resolution). Figure \ref{fig:results-confident} shows histograms of the quality difference between confident-neighbor and best-neighbor for both conventional networks. In the majority of cases, both strategies find comparable solutions. However, we find that when these strategies do find different solutions, those found by confident-neighbor are more often better, at least for the preferential-attachment network.

\begin{figure}
    \label{fig:results-confident}
    \centering
    \includegraphics[width=3.33in]{fig-results-confident.eps}
\caption{Histograms of the difference in solution quality between confident-neighbor and best-neighbor for both conventional networks.
Results are shown for parallel individual learning.}
\end{figure}

As with conform, the confident-neighbor strategy can be used when available information is insufficient to use best-neighbor. Unlike conform, confident-neighbor models weak social influence. The improved performance of confident-neighbor over best-neighbor, despite using less information, can be attributed at least in part to diversity and exploration. Confident neighbor takes significantly longer to converge, suggesting it maintains a diversity of solutions and allows more time to explore new solutions. Our results suggest several interventions with the potential to improve conventional single-group deliberation. In the presence of weak social influence, reducing the amount of information available can counter-intuitively increase performance. Conversely, when information is already limited, performance might be increased by weakening social influence. However, while confident-neighbor can increase the success of conventional deliberation, it performs comparably to network deliberation using the conform strategy, suggesting that when social influence is strong, network deliberation remains preferable.

\section{Discussion}

We have proposed and evaluated a model of mass deliberation based on the observed interlocking network structure of successful mass-deliberative projects.
Our central finding suggests that network deliberation can improve deliberative output in the presence of strong social influence (i.e., the conform strategy).
Individuals rely on social influence for a number of reasons.
Strong social influence can stem from social factors such as trust or social loafing.
Alternatively, social influence can be used to supplement individual skill when it is insufficient for a given task.
We see that a variety of contexts can lead to reliance on social influence.
Empirical evidence suggests the presence of strong social influence in real-world collaborations such as Wikipedia \cite{platt_network_2018} and in lab studies of human collaboration \cite{mason_collaborative_2012, barkoczi_social_2016}.
In the case of weak social influence (i.e., the  best-neighbor and confident-neighbor strategies), we do not observe a significant difference in output quality between network deliberation and conventional single-group deliberation. In all cases, network deliberation performs better than single-group deliberation or comparatively well. 
Our results suggest that network deliberation is both a contributing mechanism to the success of mass-deliberative projects as well as a useful tool for the design of large-scale sociotechnical systems and interventions for existing systems.

How does network deliberation improve collaborative output? In part, the high performance of random-pod network deliberation can be attributed to the high structural efficiency of the resulting network. However, structural efficiency cannot be the complete mechanism. Long-path network deliberation is extremely inefficient, but still outperforms conventional networks in the presence of strong social influence. The additional mechanisms at play remain an open question.

In order to distinguish between the effects of social influence and individual ability, we have introduced the confident-neighbor strategy, which (unlike best-neighbor) assumes the same agent capabilities as the conform strategy (in a critical learning setting).
While not central to our study of network deliberation, our findings regarding the 
confident-neighbor strategy have important implications for the study of social learning in general. 
Our findings show that just as reducing network efficiency can improve solution quality by raising diversity, reducing the available information can improve the performance of greedy learning strategies. Our investigation of confident-neighbor stems from a careful formal analysis of the information required by each strategy (See Methods). The surprisingly high performance of confident-neighbor shows the importance of isolating social influence from agent ability.

Our work is limited by a number of assumptions, which require further investigation to achieve greater generalizability. We assume a static objective function, while real-world objectives can change with changing external conditions as well as changing individual preferences. We also assume a homogeneous strategy across all agents, which raises questions of how different strategies interact within the same collaboration. We have also assumed that agents are truthful and do not form coalitions outside of the existing network, both assumptions which can be violated in real-world settings.
We have also assumed that agents have access to the full solution string, and that agents incorporate all bits of the string into their quality estimates. Extremely complex real-world tasks might be better-modeled by a collection of sub-tasks, with agents having access to only a subset of the solution string.
Finally, we have limited our study to learning strategies that explore new solutions solely through small mutations of existing solutions. Other potential strategies might generate entirely new solutions, for example, by combining parts of multiple previous solutions. Such generative strategies would add the ability to explore a wider range of solutions. While our work extends the understanding of social learning and collaboration, many questions still remain open for exploration.

Effective mass deliberation is key to realizing the democratizing power of the internet. Our findings suggest that the presence of small and interlocking pods can contribute to the success of mass deliberation, particularly when the topic is complex and social influence is strong. These findings have implications for a range of deliberative contexts: participatory government and budgeting, worker-owned cooperatives, grassroots social movements, and the governance of decentralized systems (e.g., cryptocurrency). By enabling better deliberation at larger scales, we hope this work will contribute to democratizing the governance of large sociotechnical systems, and empowering the individuals impacted by those systems.

\section{Methods}

In deliberation, individuals update their beliefs and preferences through interaction with other participants, making it a form of social learning. We thus simulate deliberation using a model of social learning closely resembling those of other authors \cite{
lazer_network_2007,
mason_propagation_2008,
barkoczi_social_2016,
gomez_clustering_2019}.
We model communication between agents using a time-dependent network $(V,E_t)$. The vertices $V$ represent agents. The edges $E_t$ allow agents to exchange information with their immediate neighbors at time $t$.
Agents collaborate to complete a collective task.
Specifically, agents seek to identify a bit string $s$ that maximizes an objective function $Q(s)$ generated using the NK model \cite{kauffman_towards_1987, weinberger_local_1991}.
The NK model is a ``tuanbly rugged'' objective function parameterized by two variables: the number of bits in the input string (N), and the number of other bits each bit interacts with (K).
These parameters can be used to tune problem size and complexity, respectively. The resulting objective function maps N-bit strings to real numbers in $[0,1]$. The resulting values can be exponentiated to modify the skewness of the solution value distribution. A full description of our model can be found in the Supplementary Information.
In keeping with other authors \cite{lazer_network_2007, barkoczi_social_2016}, we use N=15, K=6, and an exponent of 8 throughout this paper. In keeping with convention, we normalize fitness values to the global maximum.
In our simulations each agent maintains exactly one preferred solution (initially random). The simulation proceeds iteratively, with agents updating their preferred solution once per iteration based on observation of their neighbors' solutions. Throughout this paper, we simulate 300 iterations to ensure convergence. We present results for 1000 randomly generated NK model objectives, with each network/strategy combination simulated once per objective function.

\subsection{Networks}

We carry out simulations for four different network topologies. In addition to the well-known preferential-attachment \cite{barabasi_emergence_1999} and small-world \cite{watts_collective_1998} networks, we construct two types of networks to model network deliberation.
Both types exhibit interlocking structure and satisfy Property \ref{prop:mixing} (proof in Supplementary Information).
These two network deliberation conditions differ in how agents are assigned to pods and in the structural efficiency of the resulting network.
At each iteration, existing edges are replaced with cliques corresponding to the pods. An edge is created between two agents if and only if they share a pod.
For {\em random-pod} assignment, agents are assigned to pods at random.
This assignment method produces short geodesic paths and efficient network structure (see Supplementary Information).
For {\em long-path} assignment, agents are first assigned an integer label. At each stage, a new prime integer is chosen and agents are divided into classes based on the remainder of their label modulo the prime.
Each class is then divided into size-M pods by selecting the lowest M labels, then the next-lowest M, and so on.
The prime residue classes guarantee interlocking structure, while the grouping of similar labels prevents the creation of long-distance ``shortcut'' edges, producing an inefficient network (see Supplementary Information).
All networks are constructed with 100 vertices/agents and a mean degree of 4 (pod size of 5).

\subsection{Learning Strategies}
We simulate three learning strategies. Each strategy combines an individual strategy and a social strategy. The individual strategy is the same for all conditions. Agents chose a random bit of their current solution and invert that bit. If the resulting solution is an improvement, it is kept, otherwise the original is kept.
We simulate two methods of combining individual and social strategies. In the {\em fallback} condition, individual learning is only applied when social learning fails to improve the solution.
In the {\em parallel} condition, individual and social strategies are applied to the original solution in parallel and the better of the two is chosen. In all cases, a final check is performed, following \cite{barkoczi_social_2016}: new solutions are only accepted if they improve on the original solution. Descriptions of the social learning strategies follow (formal definitions can be found in the Supplementary Information).

The {\em best-neighbor} strategy closely follows that used by other authors \cite{lazer_network_2007, barkoczi_social_2016}. To apply this strategy, an agent searches the solutions of all immediate neighbors, and chooses the solution maximizing the payoff. In the event of a tie, one of the maximal solutions is chosen at random, with probability proportional to the number of neighbors exhibiting that solution.

The {\em conform} strategy also follows previous authors \cite{barkoczi_social_2016}. To apply this strategy, agents search the solutions of all immediate neighbors, counting the number of times each unique solution appears. The solution with the highest count is adopted. In the event of a tie, the original solution is maintained.

The {\em confident-neighbor} strategy is a novel variant of best-neighbor. To apply this strategy, agents search the solutions of all immediate neighbors to determine if any are superior to their own. If no superior solutions are found, the agent is {\em confident}. The confident agent maintains its original solution and broadcasts that solution to all neighbors. Alternatively, if superior solutions are found, the agent waits for broadcasts from confident neighbors. Agents may receive zero, one, or many such broadcasts. If no broadcasts are received, the agent maintains its original solution. If one or more broadcasts are received, one of them is chosen at random and that solution is adopted. It is worth noting that this strategy is equivalent to best-neighbor when the network is composed of one or more cliques, as in network deliberation.

\subsection{Statistical Methods}
All comparisons are made using two-tailed paired t-tests. Pairs of observations correspond to the same instance of the NK-model objective function. Significance values have been corrected for multiple comparisons using the Bonferroni correction. A full table of t-values and corrected p-values is available in the Supplemental Information.

\section{Data Availability}
Agent-based modeling code and data are available at
[redacted for double-blind peer-review].
%https://github.com/elplatt/Exp-Net-Delib.

\section{Acknowledgements}
[redacted for double-blind peer-review].
% We thank Grant Schoenebeck for feedback on study design and early drafts. We thank Salem Elzway for feedback on early drafts. HB's work on this project was supported by the UMSI REMS program. This work was partly supported by the Air Force Office of Scientific Research under Award No. FA9550-19-1-0029.

\section{Author Contributions}
[redacted for double-blind peer-review].
%ELP and DMR contributed to study conception, and data interpretation. ELP and HB contributed to development of simulation software and data collection. ELP drafted the manuscript.

\section{Competing Interests}
The authors declare no competing interests.

\bibliography{references}
\bibliographystyle{plain}

\end{document}

### 2021-09-14 AttackTolerance/decentralization-2019/abstract.tex
Move AttackTolerance to subfolder.

% This is the template for submission of abstracts to NetSci 2017 in Indianapolis, IN.
% It is modified from NetSci 2016.
% The editor of the booklet reserves the right to modify your submission.

%% To process this file run LaTeX2e

%%********DO NOT EDIT****************
\documentclass[12pt]{article}
\usepackage{mathptmx}
\usepackage{graphicx}
\pagestyle{empty}

\setlength\topmargin{0pt}
\addtolength\topmargin{-\headheight}
\addtolength\topmargin{-\headsep}
\setlength\oddsidemargin{0pt}
\setlength\textwidth{\paperwidth}
\addtolength\textwidth{-2in}
\setlength\textheight{\paperheight}
\addtolength\textheight{-2in}
\usepackage{layout}

\renewcommand{\title}[1]{\noindent\textbf{#1}\bigskip\\}
\renewcommand{\author}[1]{\noindent #1\bigskip\\}
%%***********************************

\begin{document}

%**********USER DEFINED**************
%Enter title here
\title{Towards Attack-Tolerant Networks: Multipath Fault Tolerance}
%Enter author(s) and address here
\author{Edward L. Platt$^1$ and Daniel M. Romero$^1$\bigskip\\
{\small
1. University of Michigan, Ann Arbor, MI, USA\\
}
}
%Enter abstract here
Centralized systems are susceptible to
targeted attacks against central points,
suggesting the importance of decentralization in attack-tolerant systems.
While many techniques exist for tolerating random faults,
better techniques for tolerating {\em adversarial faults} such as targeted attacks are needed.
In communication networks, targeted attack can leave users vulnerable to
censorship and surveillance of messages at central points.
Even when encryption is used, it can be bypassed by coercion, e.g., subpoenas.
While decentralized {\em protocols} have become a popular approach to attack-tolerance, centralized network {\em
structure} can arise even when protocols are
decentralized.
Despite their decentralized protocols,
the internet and World-Wide Web have
been shown both theoretically and historically to be highly susceptible to
adversarial faults [1], in part due to emergent structural centralization.
In this work, we present
1. A fault tolerace scheme for networked communication having a failure probability that decreses exponentially with the number of available independent paths;
2. A routing algorithm for constructing such indpendent paths in the butterfly network topology [2]; and
3. A bounded-transitivity trust model that makes it possible to quantify the adversarial fault tolerance of a network.
Our work is the first theoretical demonstration of a point-to-point communication network architecture able to tolerate coercion and other
targeted attacks, without requiring infinitely transitive trust.
We also evaluate the fault tolerance of an imperfectly implemented butterfly topology by partially rewiring a snapshot of the internet's router network.
We find that rewiring only 10\% of the edges allows the network to withstand faults in the most central 2\% of nodes,
increasing to 8\% of nodes for 90\% rewiring.
Our results show that it is possible, in principle, to create highly attack-tolerant communication systems that can withstand targeted and coercive attacks, but only when it is possible to influence network structure.
\begin{figure}[!h]
\begin{center}
\includegraphics[scale=0.75]{fig-perror}
\end{center}
\caption{
Failure probability decreases rapidly with the number of
message copies, and increases slowly with the number of faults.
Messages are difficult to attack and easy to defend.
}
\end{figure}

\noindent[1] R. Albert et al. Error and attack tolerance of complex networks. {\em Nature,} 406(6794), 2000.
\\
\noindent[2] A. D. Kshemkalyani and M. Singhal. {\em Distributed Computing}. Cambridge U. Press, 2008.

% Place the abstract of your talk/poster here, 250 Words maximum.
% Mathematical formulae may be set in LaTeX, but do NOT use the
% bibliography environment -- if you must have references, use an 
% enumerated list.
%
% Your abstract (plus one figure) should not exceed one page.
%
% Check carefully.

%************************************
\end{document}

### 2021-09-14 AttackTolerance/plosone/r2/Manuscript.tex
Move AttackTolerance to subfolder.

--- 2019-01-11 plosone/attack-figs.tex
@@
 \usepackage{amsthm}
+\usepackage{float}
 
@@
 This paper addresses the need for a theoretial understanding of network and
-redundancy-based approaches to attack tolerance.
+redundancy-based approaches to attack-tolerance.
+For the purpose of this paper, {\em attack-tolerance} refers to the ability of a pair of
+communicating nodes to detect,
+with high probability,
+when a message has been blocked or altered
+in the presence of adversarial faults.
 Our primary result is theoretical: an algorithm for constructing highly redundant
@@
 Regardless of implementation difficulty,
-we argue that {\em attack-tolerance cannot be achieved without the ability to influence
-network structure}.
+we argue that attack-tolerance cannot be achieved without the ability to influence
+network structure.
 Targeted attacks, by definition, target single points of failure.
@@
 In general,
-{\em when the need for attack-tolerance is high enough to warrant investment
-in infrastructure, topology can be engineered and maintained as infrastructure}.
+when the need for attack-tolerance is high enough to warrant investment
+in infrastructure, topology can be engineered and maintained as infrastructure.
 
@@
 \item{We prove that the number of $h$-internally vertex disjoint
-paths between two nodes in a wrap-around butterfly network
-is at least $2^h$,
+paths between two nodes in a directed wrap-around butterfly network
+is exactly $2^h$,
 and present a scalable and efficient concurrent multipath routing (CMR) algorithm
@@
 
-\begin{figure}[!h]
-\centerline{\includegraphics{fig-perror}}
+\begin{figure}[H]
+%\centerline{\includegraphics{fig-perror}}
 \caption{
@@
 
-We define the {\em bounded trust model} on
+We define the {\em bounded trust model} (Fig. \ref{fig:trust-source}) on
 an undirected graph $G = (V,E)$,
@@
 and if compromised can entirely isolate $v$ from the rest of the network.
+
+\begin{figure}[H]
+%\centerline{\includegraphics[width=4in,height=2.78in]{fig-partial-trust.pdf}}
+\caption{
+Illustration of a trusted communication network and the network properties
+used by the {\em bounded trust model}.
+Edges represent mutually trusted communication links.
+The sender (Alice, $s$) and receiver (Bob, $t$) trust all nodes
+less than the {\em trust radius} $h$ hops away.
+These nodes form their {\em trusted neighborhoods} $T_h(v)$ and $T_h(w)$.
+We assume that all faults occur in the remaining nodes: the
+{\em untrusted region}.
+The untrusted nodes in contact with the trusted neighborhoods for the
+{\em trust boundaries} $B_h(v)$ and $B_h(w)$,
+which (in the absence of central bottlenecks) determine the
+{\em effective redundancy} $\delta_h$ provided by the network.
+Alice and Bob can achieve the same level of attack-tolerance
+as if they were directly connected by $\delta_h$ redundant channels.
+}
+\label{fig:trust-source}
+\end{figure}
 
@@
 even when the exact trust radius $h$ is unknown.
-
-\begin{figure}
-\centerline{\includegraphics[width=4in,height=2.78in]{fig-partial-trust.pdf}}
-\caption{
-Illustration of a trusted communication network and the network properties
-used by the {\em bounded trust model}.
-Edges represent mutually trusted communication links.
-The sender (Alice, $s$) and receiver (Bob, $t$) trust all nodes
-less than the {\em trust radius} $h$ hops away.
-These nodes form their {\em trusted neighborhoods} $T_h(v)$ and $T_h(w)$.
-We assume that all faults occur in the remaining nodes: the
-{\em untrusted region}.
-The untrusted nodes in contact with the trusted neighborhoods for the
-{\em trust boundaries} $B_h(v)$ and $B_h(w)$,
-which (in the absence of central bottlenecks) determine the
-{\em effective redundancy} $\delta_h$ provided by the network.
-Alice and Bob can achieve the same level of attack-tolerance
-as if they were directly connected by $\delta_h$ redundant channels.
-}
-\label{fig:trust-source}
-\end{figure}
 
@@
 
-\begin{figure}[!h]
+\begin{figure}[H]
 \begin{center}
-\includegraphics[height=3in]{fig-bf-route}
+%\includegraphics[height=3in]{fig-bf-route}
 \end{center}
 \caption{
-A 3-dimensional wraparound butterfly network.
+A 3-dimensional wrap-around butterfly network.
 Note that the rightmost nodes are the same nodes as the leftmost,
@@
 \begin{center}
-\includegraphics{fig-butterfly}
+%\includegraphics{fig-butterfly}
 \end{center}
@@
 
-\begin{figure}
-\centerline{\includegraphics{fig-diameter}}
+\begin{figure}[H]
+%\centerline{\includegraphics{fig-diameter}}
 \caption{
@@
 
-\begin{figure*}[h!]
-\centerline{\includegraphics[width=5in,height=5in]{fig-mincut}}
+\begin{figure}[H]
+%\centerline{\includegraphics[width=5in,height=5in]{fig-mincut}}
 \caption{
@@
 \label{fig:mincut}
-\end{figure*}
+\end{figure}
 
@@
 
+\begin{figure}[H]
+\caption{
+Progression of place-within-level $z$ as the multipath routing algorithm
+cycles through the levels of the butterfly network.
+\label{fig:route-overview}
+}
+\end{figure}
+
+\begin{figure}[H]
+\begin{center}
+%\includegraphics{fig-routing}
+\end{center}
+\caption{
+An example of one path as constructed by the proposed multipath
+routing algorithm.
+The path is shown for $s = 10_2$
+and $w = (6, 0110111_2)$.
+\label{fig:routing}
+}
+\end{figure}
+
 \begin{table}[h!]
@@
 \end{table}
-
-\begin{figure}[h!]
-\begin{center}
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-                        ]
-\foreach \i in {0\ldots,{\ldots}0\ldots,{\ldots}0\ldots,{\ldots}0}% <-- content of nodes
-    \node[X] {\i};
-\draw[<->] ([yshift=1.5mm] A-1.north east) -- node[above=0.25mm] {$h$} ([yshift=1.5mm] A-1.north west);
-\draw[<->] ([yshift=1.5mm] A-2.north east) -- node[above=0.25mm] {$l_w - 2h$} ([yshift=1.5mm] A-2.north west);
-\draw[<->] ([yshift=1.5mm] A-3.north east) -- node[above=0.25mm] {$h$} ([yshift=1.5mm] A-3.north west);
-\draw[<->] ([yshift=1.5mm] A-4.north east) -- node[above=0.25mm] {$m - l_w$} ([yshift=1.5mm] A-4.north west);
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {\underline{Stage}} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick, fill=blue!20!white},
-                        ]
-\node[Y] {$s$};
-\foreach \i in {{\ldots}0\ldots,{\ldots}0\ldots,{\ldots}0}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {1.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick, fill=blue!20!white},
-                        ]
-\foreach \i in {$s$}% <-- content of nodes
-    \node[X] {\i};
-\node[Y] {{\ldots}1\ldots};
-\foreach \i in {{\ldots}0\ldots,{\ldots}0}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {2.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick, fill=blue!20!white},
-                        ]
-\foreach \i in {$s$,{\ldots}1\ldots}% <-- content of nodes
-    \node[X] {\i};
-\node[Y] {$\tilde{s}$};
-\foreach \i in {{\ldots}0}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {3.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {anchor=base, rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick, fill=blue!20!white},
-                        ]
-\foreach \i in {$s$,{\ldots}1\ldots,$\tilde{s}$}% <-- content of nodes
-    \node[X] {\i};
-\node[Y] {$z_{w,D}$};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {4.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {anchor=base, rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick, fill=blue!20!white},
-                        ]
-\node[Y] {$z_{w,A}$};
-\foreach \i in {{\ldots}1\ldots,$\tilde{s}$,$z_{w,D}$}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {5.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {anchor=base, rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick, fill=blue!20!white},
-                        ]
-\foreach \i in {$z_{w,A}$}% <-- content of nodes
-    \node[X] {\i};
-\node[Y] {$z_{w,B}$};
-\foreach \i in {$\tilde{s}$,$z_{w,D}$}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {6.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {anchor=base, rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick, fill=blue!20!white},
-                        ]
-\foreach \i in {$z_{w,A}$,$z_{w,B}$}% <-- content of nodes
-    \node[X] {\i};
-\node[Y] {$z_{w,C}$};
-\foreach \i in {$z_{w,D}$}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {7.} ( A-1.west);
-\end{tikzpicture}
-\end{center}
-\caption{
-Progression of place-within-level $z$ as the multipath routing algorithm
-cycles through the levels of the butterfly network.
-\label{fig:route-overview}
-}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics{fig-routing}
-\end{center}
-\caption{
-An example of one path as constructed by the proposed multipath
-routing algorithm.
-The path is shown for $s = 10_2$
-and $w = (6, 0110111_2)$.
-\label{fig:routing}
-}
-\end{figure}
 
@@
 \end{description}
+
+\subsection*{Running Time}
+At each node, the choice of edge is made using $O(1)$ lookups of the endpoint
+values at index $t$ and possibly $O(1)$ lookups of the value at a particular index of $s$.
+For an $m$-dimensional butterfly network, the algorithm chooses edges at most $2m = O(m)$ times.
+In an $m$-dimensional butterfly, there are $N = m2^m$ nodes; $m = O(\log N)$.
+The total running time to calculate a single path is thus $O(k log N)$.
+Sending a message with redundancy $k$ thus requires $O(k \log N)$ time.
 
@@
 to route a given message copy to.
-Furthermore, the existence of $2^h$ paths places a lower bound on the
-effective redundancy $\delta_h$,
-showing that the decentralized, redundant, structured networks such as the
+Each node in a directed wrap-around butterfly network has an out-degree of 2,
+placing an upper bound of $2^h$ on the effective redundancy,
+which we have just shown the above algorithm achieves, so the bound is tight.
+Thus, decentralized, redundant, structured networks such as the
 butterfly can have a very low probability of failure when faced with
@@
 Nathaniel Bezanson for helpful conversations.
-This research was partly supported by the National Science Foundation under
-Grant No. IIS-1617820.
+The authors would also like to thank the reviewers for their helpful comments.
 

### 2021-09-14 AttackTolerance/preprint/source/paper.tex
Move AttackTolerance to subfolder.

--- 2019-01-11 plosone/attack-figs.tex
@@
-% Template for PLoS
-% Version 3.5 March 2018
-%
-% % % % % % % % % % % % % % % % % % % % % %
-%
-% -- IMPORTANT NOTE
-%
-% This template contains comments intended 
-% to minimize problems and delays during our production 
-% process. Please follow the template instructions
-% whenever possible.
-%
-% % % % % % % % % % % % % % % % % % % % % % % 
-%
-% Once your paper is accepted for publication, 
-% PLEASE REMOVE ALL TRACKED CHANGES in this file 
-% and leave only the final text of your manuscript. 
-% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
-% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
-%
-%
-% There are no restrictions on package use within the LaTeX files except that 
-% no packages listed in the template may be deleted.
-%
-% Please do not include colors or graphics in the text.
-%
-% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
-%
-% % % % % % % % % % % % % % % % % % % % % % %
-%
-% -- FIGURES AND TABLES
-%
-% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
-%
-% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
-% - Figures should be uploaded separately from your manuscript file. 
-% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
-% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
-% For figure citations, please use "Fig" instead of "Figure".
-% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
-%
-% Tables should be cell-based and may not contain:
-% - spacing/line breaks within cells to alter layout or alignment
-% - do not nest tabular environments (no tabular environments within tabular environments)
-% - no graphics or colored text (cell background color/shading OK)
-% See http://journals.plos.org/plosone/s/tables for table guidelines.
-%
-% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
-%
-% % % % % % % % % % % % % % % % % % % % % % % %
-%
-% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
-%
-% IMPORTANT
-% Below are a few tips to help format your equations and other special characters according to our specifications.
-For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
-%
-% For inline equations, please be sure to include all portions of an equation in the math environment.
-For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
-%
-% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
-%
-% Please add line breaks to long display equations when possible in order to fit size of the column. 
-%
-% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
-%
-% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
-%
-% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
-%
-% % % % % % % % % % % % % % % % % % % % % % % % 
-%
-% Please contact latex@plos.org with any questions.
-%
-% % % % % % % % % % % % % % % % % % % % % % % %
-
-\documentclass[10pt,letterpaper]{article}
-\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
-
-% amsmath and amssymb packages, useful for mathematical formulas and symbols
-\usepackage{amsmath,amssymb}
-
-% Use adjustwidth environment to exceed column width (see example table in text)
-\usepackage{changepage}
-
-% Use Unicode characters when possible
-\usepackage[utf8x]{inputenc}
-
-% textcomp package and marvosym package for additional characters
-\usepackage{textcomp,marvosym}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-
-% Use nameref to cite supporting information files (see Supporting Information section for more info)
-\usepackage{nameref,hyperref}
-
-% line numbers
-\usepackage[right]{lineno}
-
-% ligatures disabled
-\usepackage{microtype}
-\DisableLigatures[f]{encoding = *, family = * }
-
-% color can be used to apply background shading to table cells only
-\usepackage[table]{xcolor}
-
-% array package and thick rules for tables
-\usepackage{array}
-
-% create "+" rule type for thick vertical lines
-\newcolumntype{+}{!{\vrule width 2pt}}
-
-% create \thickcline for thick horizontal lines of variable length
-\newlength\savedwidth
-\newcommand\thickcline[1]{%
-  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
-  \cline{#1}%
-  \noalign{\vskip\arrayrulewidth}%
-  \noalign{\global\arrayrulewidth\savedwidth}%
-}
-
-% \thickhline command for thick horizontal lines that span the table
-\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
-\hline
-\noalign{\global\arrayrulewidth\savedwidth}}
-
-% Remove comment for double spacing
-\usepackage{setspace} 
-\doublespacing
-
-% Text layout
-\raggedright
-\setlength{\parindent}{0.5cm}
-\textwidth 5.25in 
-\textheight 8.75in
-
-% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
-% Captions will be left justified
-\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
-\renewcommand{\figurename}{Fig}
-
-% Use the PLoS provided BiBTeX style
-\bibliographystyle{plos2015}
-
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-% Header and Footer with logo
-\usepackage{lastpage,fancyhdr,graphicx}
-\usepackage{epstopdf}
-%\pagestyle{myheadings}
-\pagestyle{fancy}
-\fancyhf{}
-%\setlength{\headheight}{27.023pt}
-%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
-\rfoot{\thepage/\pageref{LastPage}}
-\renewcommand{\headrulewidth}{0pt}
-\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
-\fancyheadoffset[L]{2.25in}
-\fancyfootoffset[L]{2.25in}
-\lfoot{\today}
-
-\usepackage{booktabs} % For formal tables
+\documentclass{sig-alternate-05-2015}
+
+\usepackage[utf8]{inputenc}
+\usepackage{afterpage}
 \usepackage{epigraph}
-\usepackage{afterpage}
 \usepackage{tikz}
@@
 \usepackage{pifont} % http://ctan.org/pkg/pifont
-\usepackage{amsthm}
-
-%% Include all macros below
+\usepackage{float}
 
@@
 
-\newcommand{\lorem}{{\bf LOREM}}
-\newcommand{\ipsum}{{\bf IPSUM}}
-
-%% END MACROS SECTION
-
 \begin{document}
-\vspace*{0.2in}
-
-% Title must be 250 characters or less.
-\begin{flushleft}
-{\Large
-\textbf\newline{Towards attack tolerant networks: concurrent multipath routing and the butterfly network} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
-}
-\newline
-% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
-\\
-Edward L. Platt\textsuperscript{1,*},
-Daniel M. Romero\textsuperscript{1}
-\\
-\bigskip
-\textbf{1} School of Information, University of Michigan, Ann Arbor, Michigan, USA
-\\
-\bigskip
-
-% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
+
+% Copyright
+%\setcopyright{acmcopyright}
+%\setcopyright{acmlicensed}
+%\setcopyright{rightsretained}
+%\setcopyright{usgov}
+%\setcopyright{usgovmixed}
+%\setcopyright{cagov}
+%\setcopyright{cagovmixed}
+
+% DOI
+\doi{}
+
+% ISBN
+\isbn{}
+
+%Conference
+\conferenceinfo{Pre-print}{Currently under peer review.}
+
 %
-% Remove or comment out the author notes below if they aren't used.
+% --- Author Metadata here ---
+%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
+%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
+%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
+% --- End of Author Metadata ---
+
+\title{Towards Attack-Tolerant Networks: Concurrent Multipath Routing and the Butterfly Network}
 %
-% Primary Equal Contribution Note
-
-% Additional Equal Contribution Note
-% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
-
-% Current address notes
-%\textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
-% \textcurrency b Insert second current address 
-% \textcurrency c Insert third current address
-
-% Deceased author note
-%\dag Deceased
-
-% Group/Consortium Author Note
-%\textpilcrow Membership list can be found in the Acknowledgments section.
-
-% Use the asterisk to denote corresponding authorship and provide email address in note below.
-* elplatt@umich.edu
-
-\end{flushleft}
-% Please keep the abstract below 300 words
-\section*{Abstract}
+% You need the command \numberofauthors to handle the 'placement
+% and alignment' of the authors beneath the title.
+%
+% For aesthetic reasons, we recommend 'three authors at a time'
+% i.e. three 'name/affiliation blocks' be placed beneath the title.
+%
+% NOTE: You are NOT restricted in how many 'rows' of
+% "name/affiliations" may appear. We just ask that you restrict
+% the number of 'columns' to three.
+%
+% Because of the available 'opening page real-estate'
+% we ask you to refrain from putting more than six authors
+% (two rows with three columns) beneath the article title.
+% More than six makes the first-page appear very cluttered indeed.
+%
+% Use the \alignauthor commands to handle the names
+% and affiliations for an 'aesthetic maximum' of six authors.
+% Add names, affiliations, addresses for
+% the seventh etc. author(s) as the argument for the
+% \additionalauthors command.
+% These 'additional authors' will be output/set for you
+% without further effort on your part as the last section in
+% the body of your article BEFORE References or any Appendices.
+
+\numberofauthors{2} %  in this sample file, there are a *total*
+% of EIGHT authors. SIX appear on the 'first-page' (for formatting
+% reasons) and the remaining two appear in the \additionalauthors section.
+%
+\author{
+% You can go ahead and credit any number of authors here,
+% e.g. one 'row of three' or two rows (consisting of one row of three
+% and a second row of one, two or three).
+%
+% The command \alignauthor (no curly braces needed) should
+% precede each author name, affiliation/snail-mail address and
+% e-mail address. Additionally, tag each line of
+% affiliation/address with \affaddr, and tag the
+% e-mail address with \email.
+%
+% 1st. author
+\alignauthor
+Edward L. Platt \\
+\affaddr{University of Michigan}\\
+\email{elplatt@umich.edu}
+\alignauthor
+Daniel M. Romero \\
+\affaddr{University of Michigan}\\
+\email{drom@umich.edu}
+}
+%       \affaddr{Institute for Clarity in Documentation}\\
+%       \affaddr{1932 Wallamaloo Lane}\\
+%       \affaddr{Wallamaloo, New Zealand}\\
+%       \email{trovato@corporation.com}
+% 2nd. author
+%\alignauthor
+%G.K.M. Tobin\titlenote{The secretary disavows
+%any knowledge of this author's actions.}\\
+%       \affaddr{Institute for Clarity in Documentation}\\
+%       \affaddr{P.O. Box 1212}\\
+%       \affaddr{Dublin, Ohio 43017-6221}\\
+%       \email{webmaster@marysville-ohio.com}
+% 3rd. author
+
+\maketitle
+\begin{abstract}
 It is crucial for large-scale communication networks such as the internet
@@
 architectures to guard against the dangers of censorship and surveillance.
-
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLOS ONE authors please skip this step. 
-% Author Summary not valid for PLOS ONE submissions.   
-%\section*{Author summary}
-%Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Curabitur eget porta erat.
-Morbi consectetur est vel gravida pretium.
-Suspendisse ut dui eu ante cursus gravida non sed sem.
-Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
-Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
-Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
-Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
-Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
-
-\linenumbers
-
-% Use "Eq" instead of "Equation" for equation citations.
+\end{abstract}
+
 \section*{Introduction}
@@
 
-\begin{figure}[!h]
+\begin{figure}[H]
 \centerline{\includegraphics{fig-perror}}
@@
 and if compromised can entirely isolate $v$ from the rest of the network.
+
+\begin{figure}[H]
+\centerline{\includegraphics[width=4in,height=2.78in]{fig-partial-trust.pdf}}
+\caption{
+Illustration of a trusted communication network and the network properties
+used by the {\em bounded trust model}.
+Edges represent mutually trusted communication links.
+The sender (Alice, $s$) and receiver (Bob, $t$) trust all nodes
+less than the {\em trust radius} $h$ hops away.
+These nodes form their {\em trusted neighborhoods} $T_h(v)$ and $T_h(w)$.
+We assume that all faults occur in the remaining nodes: the
+{\em untrusted region}.
+The untrusted nodes in contact with the trusted neighborhoods for the
+{\em trust boundaries} $B_h(v)$ and $B_h(w)$,
+which (in the absence of central bottlenecks) determine the
+{\em effective redundancy} $\delta_h$ provided by the network.
+Alice and Bob can achieve the same level of attack-tolerance
+as if they were directly connected by $\delta_h$ redundant channels.
+}
+\label{fig:trust-source}
+\end{figure}
 
@@
 
-\begin{figure}[!h]
+\begin{figure}[H]
 \begin{center}
@@
 \caption{
-A 3-dimensional wraparound butterfly network.
+A 3-dimensional wrap-around butterfly network.
 Note that the rightmost nodes are the same nodes as the leftmost,
@@
 
-\begin{figure}
+\begin{figure}[H]
 \centerline{\includegraphics{fig-diameter}}
@@
 
-\begin{figure*}[h!]
+\begin{figure}[H]
 \centerline{\includegraphics[width=5in,height=5in]{fig-mincut}}
@@
 
-\begin{table}[h!]
-\caption{Butterfly Multipath Routing Variables\label{tab:routing}}
-\begin{center}
-\begin{tabular}{ll}
-Name & Variable \\\hline
-butterfly dimension & $m \in \mathbb{Z}_+$ \\
-node level & $l \in \mathbb{Z} : 0 \leq l < m$ \\
-node place within level & $z \in \mathbb{Z}_2^m$ \\
-trust radius & $h \in \mathbb{Z} : 1 \leq h \leq \lfloor m/2 \rfloor$ \\
-path index & $s \in \mathbb{Z}_2^h$ \\
-\end{tabular}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
+\begin{figure}[H]
 \begin{center}
@@
 \end{figure}
+
+\begin{table}[h!]
+\caption{Butterfly Multipath Routing Variables\label{tab:routing}}
+\begin{center}
+\begin{tabular}{ll}
+Name & Variable \\\hline
+butterfly dimension & $m \in \mathbb{Z}_+$ \\
+node level & $l \in \mathbb{Z} : 0 \leq l < m$ \\
+node place within level & $z \in \mathbb{Z}_2^m$ \\
+trust radius & $h \in \mathbb{Z} : 1 \leq h \leq \lfloor m/2 \rfloor$ \\
+path index & $s \in \mathbb{Z}_2^h$ \\
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 Nathaniel Bezanson for helpful conversations.
-This research was partly supported by the National Science Foundation under
-Grant No. IIS-1617820.
-
-\nolinenumbers
-
-%\bibliography{references}
-
-% Either type in your references using
-% \begin{thebibliography}{}
-% \bibitem{}
-% Text
-% \end{thebibliography}
+This research was partly supported by the National Science Foundation under Grant No. IIS-1617820.
+
 %
-% or
+% The following two commands are all you need in the
+% initial runs of your .tex file to
+% produce the bibliography for the citations in your paper.
+\bibliographystyle{abbrv}
+\bibliography{paper}  % sigproc.bib is the name of the Bibliography in this case
+% You must have a proper ".bib" file
+%  and remember to run:
+% latex bibtex latex latex
+% to resolve all references
 %
-% Compile your BiBTeX database using our plos2015.bst
-% style file and paste the contents of your .bbl file
-% here. See http://journals.plos.org/plosone/s/latex for 
-% step-by-step instructions.
+% ACM needs 'a single self-contained file'!
 %
-\begin{thebibliography}{10}
-
-\bibitem{sterbenz_resilience_2010}
-Sterbenz JP, Hutchison D, Çetinkaya EK, Jabbar A, Rohrer JP, Schöller M,
-  et~al.
-\newblock Resilience and survivability in communication networks: {Strategies},
-  principles, and survey of disciplines.
-\newblock Computer Networks. 2010;54(8):1245--1265.
-
-\bibitem{barabasi_scale-free_2009}
-Barabási AL, {others}.
-\newblock Scale-free networks: a decade and beyond.
-\newblock Science. 2009;325(5939):412.
-
-\bibitem{albert_error_2000}
-Albert R, Jeong H, Barabási AL.
-\newblock Error and attack tolerance of complex networks.
-\newblock Nature. 2000;406(6794):378--382.
-
-\bibitem{nayak_different_2010}
-Nayak GN, Samaddar SG.
-\newblock Different flavours of man-in-the-middle attack, consequences and
-  feasible solutions.
-\newblock In: Computer {Science} and {Information} {Technology} ({ICCSIT}),
-  2010 3rd {IEEE} {International} {Conference} on. vol.~5. IEEE; 2010. p.
-  491--495.
-
-\bibitem{grewal_caveat_2013}
-Grewal GS, Ryan MD, Bursuc S, Ryan PY.
-\newblock Caveat coercitor: {Coercion}-evidence in electronic voting.
-\newblock In: Security and {Privacy} ({SP}), 2013 {IEEE} {Symposium} on. IEEE;
-  2013. p. 367--381.
-
-\bibitem{elmer-dewitt_first_1993}
-Elmer-Dewitt P, Jackson D.
-\newblock First nation in cyberspace.
-\newblock Time. 1993;6:62--64.
-
-\bibitem{baran_distributed_1964}
-Baran P, {others}.
-\newblock On distributed communications.
-\newblock Volumes I-XI, RAND Corporation Research Documents, August. 1964; p.
-  637--648.
-
-\bibitem{dainotti_analysis_2011}
-Dainotti A, Squarcella C, Aben E, Claffy KC, Chiesa M, Russo M, et~al.
-\newblock Analysis of country-wide internet outages caused by censorship.
-\newblock In: Proceedings of the 2011 {ACM} {SIGCOMM} conference on {Internet}
-  measurement conference. ACM; 2011. p. 1--18.
-
-\bibitem{christianson_why_1997}
-Christianson B, Harbison WS.
-\newblock Why isn't trust transitive?
-\newblock In: Security protocols. Springer; 1997. p. 171--176.
-
-\bibitem{zin_survey_2015}
-Zin SM, Anuar NB, Kiah MLMM, Ahmedy I.
-\newblock Survey of secure multipath routing protocols for {WSNs}.
-\newblock J Netw Comput Appl. 2015;55:123--153.
-
-\bibitem{barabasi_emergence_1999}
-Barabási AL, Albert R.
-\newblock Emergence of scaling in random networks.
-\newblock Science. 1999;286(5439):509--512.
-
-\bibitem{hunter_pakistan_2008}
-Hunter P.
-\newblock Pakistan {YouTube} block exposes fundamental internet security
-  weakness: {Concern} that pakistani action affected youtube access elsewhere
-  in world.
-\newblock Computer Fraud \& Security. 2008;2008(4):10--11.
-
-\bibitem{poulsen_edward_2013}
-Poulsen K.
-\newblock Edward {Snowden}’s e-mail provider defied {FBI} demands to turn
-  over crypto keys, documents show.
-\newblock WIRED. 2013;.
-
-\bibitem{zimmermann_official_1995}
-Zimmermann PR.
-\newblock The official {PGP} user's guide.
-\newblock MIT press; 1995.
-
-\bibitem{richters_trust_2011}
-Richters O, Peixoto TP.
-\newblock Trust transitivity in social networks.
-\newblock PloS one. 2011;6(4):e18384.
-
-\bibitem{avizienis_basic_2004}
-Avizienis A, Laprie JC, Randell B, Landwehr C.
-\newblock Basic concepts and taxonomy of dependable and secure computing.
-\newblock IEEE T Depend Secure. 2004;1(1):11--33.
-
-\bibitem{von_neumann_probabilistic_1956}
-Von~Neumann J.
-\newblock Probabilistic logics and the synthesis of reliable organisms from
-  unreliable components.
-\newblock Automata studies. 1956;34:43--98.
-
-\bibitem{qadir_exploiting_2015}
-Qadir J, Ali A, Yau KLA, Sathiaseelan A, Crowcroft J.
-\newblock Exploiting the power of multiplicity: a holistic survey of
-  network-layer multipath.
-\newblock IEEE Comm Surv Tut. 2015;17(4):2176--2213.
-
-\bibitem{khiani_comparative_2013}
-Khiani SR, Dethe C, Thakare V.
-\newblock Comparative {Analysis} of {Multipath} {Routing} {Techniques} and
-  {Design} of {Secure} {Energy} {Aware} {Routing} {Algorithm} for {Wireless}
-  {Sensor} {Network}.
-\newblock IJACR. 2013;3(3):374.
-
-\bibitem{kshemkalyani_distributed_2008}
-Kshemkalyani AD, Singhal M.
-\newblock Distributed computing: principles, algorithms, and systems.
-\newblock Cambridge University Press; 2008.
-
-\bibitem{lua_survey_2005}
-Lua EK, Crowcroft J, Pias M, Sharma R, Lim S.
-\newblock A survey and comparison of peer-to-peer overlay network schemes.
-\newblock IEEE Commun Surv Tut. 2005;7(2):72--93.
-
-\bibitem{korzun_structured_2013}
-Korzun D, Gurtov A.
-\newblock Structured peer-to-peer systems: fundamentals of hierarchical
-  organization, routing, scaling, and security.
-\newblock New York, NY: Springer; 2013.
-
-\bibitem{mohr_explaining_1982}
-Mohr LB.
-\newblock Explaining organizational behavior.
-\newblock Jossey-Bass; 1982.
-
-\bibitem{walker_mass_2012}
-Walker DC.
-\newblock Mass notification and crisis communications: {Planning},
-  preparedness, and systems.
-\newblock CRC Press; 2012.
-
-\bibitem{nickerson_thinking_2010}
-Nickerson JV, Tversky B, Corter JE, Yu L, Mason D.
-\newblock Thinking with networks.
-\newblock In: {CogSci}. vol.~36; 2010.
-
-\bibitem{ellison_ten_2000}
-Ellison C, Schneier B.
-\newblock Ten risks of {PKI}: {What} you're not being told about public key
-  infrastructure.
-\newblock Comput Secur J. 2000;16(1):1--7.
-
-\bibitem{levien_attack-resistant_2009}
-Levien R.
-\newblock Attack-resistant trust metrics.
-\newblock In: Computing with {Social} {Trust}. Springer; 2009. p. 121--132.
-
-\bibitem{reiter_resilient_1998}
-Reiter MK, Stubblebine SG.
-\newblock Resilient authentication using path independence.
-\newblock IEEE T Comput. 1998;47(12):1351--1362.
-
-\bibitem{lamport_byzantine_1982}
-Lamport L, Shostak R, Pease M.
-\newblock The {Byzantine} generals problem.
-\newblock TOPLAS. 1982;4(3):382--401.
-
-\bibitem{castro_practical_1999}
-Castro M, Liskov B, {others}.
-\newblock Practical {Byzantine} fault tolerance.
-\newblock In: {OSDI}. vol.~99; 1999. p. 173--186.
-
-\bibitem{dwork_pricing_1993}
-Dwork C, Naor M.
-\newblock Pricing via processing or combatting junk mail.
-\newblock In: Advances in {Cryptology}. Springer; 1993. p. 139--147.
-
-\bibitem{nakamoto_bitcoin:_2008}
-Nakamoto S.
-\newblock Bitcoin: {A} peer-to-peer electronic cash system.
-\newblock bitcoinorg. 2008; p.~28.
-
-\bibitem{mazieres_stellar_2015}
-Mazières D.
-\newblock Stellar {Consensus} {Protocol}: {A} {Federated} {Model} for
-  {Internet}-level {Consensus}; 2015.
-
-\bibitem{fiat_censorship_2002}
-Fiat A, Saia J.
-\newblock Censorship resistant peer-to-peer content addressable networks.
-\newblock In: {SIAM} {SODA}. ACM; 2002. p. 94--103.
-
-\bibitem{clarke_freenet:_2001}
-Clarke I, Sandberg O, Wiley B, Hong TW.
-\newblock Freenet: {A} distributed anonymous information storage and retrieval
-  system.
-\newblock In: Designing {Privacy} {Enhancing} {Technologies}. Springer; 2001.
-  p. 46--66.
-
-\bibitem{shamir_how_1979}
-Shamir A.
-\newblock How to share a secret.
-\newblock Communications of the ACM. 1979;22(11):612--613.
-
-\bibitem{blakley_safeguarding_1979}
-Blakley GR.
-\newblock Safeguarding cryptographic keys.
-\newblock P Natl Comp Conf. 1979;48:313--317.
-
-\bibitem{zhang_using_2002}
-Zhang H, Goel A, Govindan R.
-\newblock Using the small-world model to improve freenet performance.
-\newblock In: {INFOCOM}. vol.~3. IEEE; 2002. p. 1228--1237.
-
-\bibitem{kleinberg_small-world_2000}
-Kleinberg J.
-\newblock The small-world phenomenon: {An} algorithmic perspective.
-\newblock In: {STOC}. ACM; 2000. p. 163--170.
-
-\bibitem{alrajeh_secure_2013}
-Alrajeh NA, Alabed MS, Elwahiby MS.
-\newblock Secure ant-based routing protocol for wireless sensor network.
-\newblock Int J Distrib Sens N. 2013;2013.
-
-\bibitem{kohno_improvement_2012}
-Kohno E, Okazaki T, Takeuchi M, Ohta T, Kakuda Y, Aida M.
-\newblock Improvement of assurance including security for wireless sensor
-  networks using dispersed data transmission.
-\newblock J Comp Sys Sci. 2012;78(6):1703--1715.
-
-\bibitem{khalil_unmask:_2010}
-Khalil I, Bagchi S, Rotaru CN, Shroff NB.
-\newblock {UnMask}: {Utilizing} neighbor monitoring for attack mitigation in
-  multihop wireless sensor networks.
-\newblock Ad Hoc Networks. 2010;8(2):148--164.
-
-\bibitem{lou_h-spread:_2006}
-Lou W, Kwon Y.
-\newblock H-{SPREAD}: a hybrid multipath scheme for secure and reliable data
-  collection in wireless sensor networks.
-\newblock IEEE T Veh Technol. 2006;55(4):1320--1330.
-
-\bibitem{liu_secure_2012}
-Liu A, Zheng Z, Zhang C, Chen Z, Shen X.
-\newblock Secure and energy-efficient disjoint multipath routing for {WSNs}.
-\newblock IEEE T Veh Technol. 2012;61(7):3255--3265.
-
-\bibitem{leskovec_graphs_2005}
-Leskovec J, Kleinberg J, Faloutsos C.
-\newblock Graphs over time: densification laws, shrinking diameters and
-  possible explanations.
-\newblock In: Proceedings of the eleventh {ACM} {SIGKDD} international
-  conference on {Knowledge} discovery in data mining. ACM; 2005. p. 177--187.
-
-\bibitem{xu_internet_2011}
-Xu X, Mao ZM, Halderman JA.
-\newblock Internet censorship in {China}: {Where} does the filtering occur?
-\newblock In: International {Conference} on {Passive} and {Active} {Network}
-  {Measurement}. Springer; 2011. p. 133--142.
-
-\bibitem{dingledine_tor:_2004}
-Dingledine R, Mathewson N, Syverson P.
-\newblock Tor: {The} second-generation onion router.
-\newblock DTIC Document; 2004.
-
-\end{thebibliography}
-
+%
 \end{document}

### 2021-09-14 AttackTolerance/preprint/source/preprint.tex
Move AttackTolerance to subfolder.

--- 2021-09-14 AttackTolerance/preprint/source/paper.tex
@@
-\documentclass{sig-alternate-05-2015}
-
+\documentclass[twocolumn]{article}
+
+\usepackage[top=0.75in,left=0.75in,right=0.75in,bottom=1in]{geometry}
+\setlength{\columnsep}{0.33in}
+
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{amsthm}
+\usepackage{epigraph}
 \usepackage[utf8]{inputenc}
 \usepackage{afterpage}
-\usepackage{epigraph}
 \usepackage{tikz}
@@
 \usepackage{pifont} % http://ctan.org/pkg/pifont
+\usepackage{amsthm}
 \usepackage{float}
@@
 
+%\usepackage{draftwatermark}
+%\SetWatermarkText{DRAFT}
+
 \begin{document}
 
-% Copyright
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\setcopyright{usgov}
-%\setcopyright{usgovmixed}
-%\setcopyright{cagov}
-%\setcopyright{cagovmixed}
-
-% DOI
-\doi{}
-
-% ISBN
-\isbn{}
-
-%Conference
-\conferenceinfo{Pre-print}{Currently under peer review.}
-
-%
-% --- Author Metadata here ---
-%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
-%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
-%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
-% --- End of Author Metadata ---
-
 \title{Towards Attack-Tolerant Networks: Concurrent Multipath Routing and the Butterfly Network}
-%
-% You need the command \numberofauthors to handle the 'placement
-% and alignment' of the authors beneath the title.
-%
-% For aesthetic reasons, we recommend 'three authors at a time'
-% i.e. three 'name/affiliation blocks' be placed beneath the title.
-%
-% NOTE: You are NOT restricted in how many 'rows' of
-% "name/affiliations" may appear. We just ask that you restrict
-% the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate'
-% we ask you to refrain from putting more than six authors
-% (two rows with three columns) beneath the article title.
-% More than six makes the first-page appear very cluttered indeed.
-%
-% Use the \alignauthor commands to handle the names
-% and affiliations for an 'aesthetic maximum' of six authors.
-% Add names, affiliations, addresses for
-% the seventh etc. author(s) as the argument for the
-% \additionalauthors command.
-% These 'additional authors' will be output/set for you
-% without further effort on your part as the last section in
-% the body of your article BEFORE References or any Appendices.
-
-\numberofauthors{2} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
-%
 \author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
-\alignauthor
 Edward L. Platt \\
-\affaddr{University of Michigan}\\
-\email{elplatt@umich.edu}
-\alignauthor
+University of Michigan\\
+elplatt@umich.edu
+\and
 Daniel M. Romero \\
-\affaddr{University of Michigan}\\
-\email{drom@umich.edu}
-}
-%       \affaddr{Institute for Clarity in Documentation}\\
-%       \affaddr{1932 Wallamaloo Lane}\\
-%       \affaddr{Wallamaloo, New Zealand}\\
-%       \email{trovato@corporation.com}
-% 2nd. author
-%\alignauthor
-%G.K.M. Tobin\titlenote{The secretary disavows
-%any knowledge of this author's actions.}\\
-%       \affaddr{Institute for Clarity in Documentation}\\
-%       \affaddr{P.O. Box 1212}\\
-%       \affaddr{Dublin, Ohio 43017-6221}\\
-%       \email{webmaster@marysville-ohio.com}
-% 3rd. author
+University of Michigan\\
+drom@umich.edu
+}
+\date{}
 
@@
 \end{abstract}
+
+\noindent{\bf Keywords:}
+Butterfly Topology,
+Fault Tolerance,
+Adversarial Faults,
+Multipath Routing,
+Censorship,
+Decentralization.
 
@@
 
-\begin{figure}[H]
-\centerline{\includegraphics{fig-perror}}
+\begin{figure}[ht]
+\centerline{\includegraphics[width=3in]{fig-perror}}
 \caption{
@@
 
-\begin{figure}[H]
-\centerline{\includegraphics[width=4in,height=2.78in]{fig-partial-trust.pdf}}
+\begin{figure}[ht]
+\centerline{\includegraphics[width=3in]{fig-partial-trust.pdf}}
 \caption{
@@
 
-\begin{figure}[H]
+\begin{figure}[ht]
 \begin{center}
-\includegraphics[height=3in]{fig-bf-route}
+\includegraphics[width=2.5in]{fig-bf-route}
 \end{center}
@@
 
-\begin{figure}[H]
-\centerline{\includegraphics{fig-diameter}}
+\begin{figure}[ht]
+\centerline{\includegraphics[width=3in]{fig-diameter}}
 \caption{
@@
 
-\begin{figure}[H]
-\centerline{\includegraphics[width=5in,height=5in]{fig-mincut}}
+\begin{figure}[ht]
+\centerline{\includegraphics[width=3in]{fig-mincut}}
 \caption{
@@
 
-\begin{figure}[H]
+\begin{figure}[ht]
 \begin{center}
@@
 Nathaniel Bezanson for helpful conversations.
-This research was partly supported by the National Science Foundation under Grant No. IIS-1617820.
-
-%
-% The following two commands are all you need in the
-% initial runs of your .tex file to
-% produce the bibliography for the citations in your paper.
+This research was partly supported by the National Science Foundation unde;r Grant No. IIS-1617820.
+
 \bibliographystyle{abbrv}
-\bibliography{paper}  % sigproc.bib is the name of the Bibliography in this case
-% You must have a proper ".bib" file
-%  and remember to run:
-% latex bibtex latex latex
-% to resolve all references
-%
-% ACM needs 'a single self-contained file'!
-%
-%
+\bibliography{preprint}  % sigproc.bib is the name of the Bibliography in this case
 \end{document}

### 2021-09-14 AttackTolerance/websci-2017/source/paper.tex
Move AttackTolerance to subfolder.

--- 2017-03-16 websci-2017/source/paper.tex
@@
 \section{Acknowledgments}
+The authors would like to thank Tony Garnock-Jones, A. Frederick Dudley, and
+Nathaniel Bezanson for helpful conversations.
+This research was partly supported by the National Science Foundation under Grant No. IIS-1617820.
 

### 2021-09-14 AttackTolerance/foci-2018/foci.tex

--- 2019-01-11 plosone/attack-figs.tex
@@
-% Template for PLoS
-% Version 3.5 March 2018
-%
-% % % % % % % % % % % % % % % % % % % % % %
-%
-% -- IMPORTANT NOTE
-%
-% This template contains comments intended 
-% to minimize problems and delays during our production 
-% process. Please follow the template instructions
-% whenever possible.
-%
-% % % % % % % % % % % % % % % % % % % % % % % 
-%
-% Once your paper is accepted for publication, 
-% PLEASE REMOVE ALL TRACKED CHANGES in this file 
-% and leave only the final text of your manuscript. 
-% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
-% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
-%
-%
-% There are no restrictions on package use within the LaTeX files except that 
-% no packages listed in the template may be deleted.
-%
-% Please do not include colors or graphics in the text.
-%
-% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
-%
-% % % % % % % % % % % % % % % % % % % % % % %
-%
-% -- FIGURES AND TABLES
-%
-% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
-%
-% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
-% - Figures should be uploaded separately from your manuscript file. 
-% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
-% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
-% For figure citations, please use "Fig" instead of "Figure".
-% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
-%
-% Tables should be cell-based and may not contain:
-% - spacing/line breaks within cells to alter layout or alignment
-% - do not nest tabular environments (no tabular environments within tabular environments)
-% - no graphics or colored text (cell background color/shading OK)
-% See http://journals.plos.org/plosone/s/tables for table guidelines.
-%
-% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
-%
-% % % % % % % % % % % % % % % % % % % % % % % %
-%
-% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
-%
-% IMPORTANT
-% Below are a few tips to help format your equations and other special characters according to our specifications.
-For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
-%
-% For inline equations, please be sure to include all portions of an equation in the math environment.
-For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
-%
-% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
-%
-% Please add line breaks to long display equations when possible in order to fit size of the column. 
-%
-% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
-%
-% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
-%
-% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
-%
-% % % % % % % % % % % % % % % % % % % % % % % % 
-%
-% Please contact latex@plos.org with any questions.
-%
-% % % % % % % % % % % % % % % % % % % % % % % %
-
-\documentclass[10pt,letterpaper]{article}
-\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
-
-% amsmath and amssymb packages, useful for mathematical formulas and symbols
-\usepackage{amsmath,amssymb}
-
-% Use adjustwidth environment to exceed column width (see example table in text)
-\usepackage{changepage}
-
-% Use Unicode characters when possible
-\usepackage[utf8x]{inputenc}
-
-% textcomp package and marvosym package for additional characters
-\usepackage{textcomp,marvosym}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-
-% Use nameref to cite supporting information files (see Supporting Information section for more info)
-\usepackage{nameref,hyperref}
-
-% line numbers
-\usepackage[right]{lineno}
-
-% ligatures disabled
-\usepackage{microtype}
-\DisableLigatures[f]{encoding = *, family = * }
-
-% color can be used to apply background shading to table cells only
-\usepackage[table]{xcolor}
-
-% array package and thick rules for tables
-\usepackage{array}
-
-% create "+" rule type for thick vertical lines
-\newcolumntype{+}{!{\vrule width 2pt}}
-
-% create \thickcline for thick horizontal lines of variable length
-\newlength\savedwidth
-\newcommand\thickcline[1]{%
-  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
-  \cline{#1}%
-  \noalign{\vskip\arrayrulewidth}%
-  \noalign{\global\arrayrulewidth\savedwidth}%
-}
-
-% \thickhline command for thick horizontal lines that span the table
-\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
-\hline
-\noalign{\global\arrayrulewidth\savedwidth}}
-
-% Remove comment for double spacing
-\usepackage{setspace} 
-\doublespacing
-
-% Text layout
-\raggedright
-\setlength{\parindent}{0.5cm}
-\textwidth 5.25in 
-\textheight 8.75in
-
-% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
-% Captions will be left justified
-\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
-\renewcommand{\figurename}{Fig}
-
-% Use the PLoS provided BiBTeX style
-\bibliographystyle{plos2015}
-
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-% Header and Footer with logo
-\usepackage{lastpage,fancyhdr,graphicx}
-\usepackage{epstopdf}
-%\pagestyle{myheadings}
-\pagestyle{fancy}
-\fancyhf{}
-%\setlength{\headheight}{27.023pt}
-%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
-\rfoot{\thepage/\pageref{LastPage}}
-\renewcommand{\headrulewidth}{0pt}
-\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
-\fancyheadoffset[L]{2.25in}
-\fancyfootoffset[L]{2.25in}
-\lfoot{\today}
+% TEMPLATE for Usenix papers, specifically to meet requirements of
+%  USENIX '05
+% originally a template for producing IEEE-format articles using LaTeX.
+%   written by Matthew Ward, CS Department, Worcester Polytechnic Institute.
+% adapted by David Beazley for his excellent SWIG paper in Proceedings,
+%   Tcl 96
+% turned into a smartass generic template by De Clarke, with thanks to
+%   both the above pioneers
+% use at your own risk.  Complaints to /dev/null.
+% make it two column with no page numbering, default is 10 point
+
+% Munged by Fred Douglis <douglis@research.att.com> 10/97 to separate
+% the .sty file from the LaTeX source template, so that people can
+% more easily include the .sty file into an existing document.  Also
+% changed to more closely follow the style guidelines as represented
+% by the Word sample file.
+
+% Note that since 2010, USENIX does not require endnotes. If you want
+% foot of page notes, don't include the endnotes package in the 
+% usepackage command, below.
+
+% This version uses the latex2e styles, not the very ancient 2.09 stuff.
+\documentclass[letterpaper,twocolumn,10pt]{article}
+\usepackage{usenix,epsfig,endnotes}
 
@@
 \usepackage{pifont} % http://ctan.org/pkg/pifont
+\usepackage{amsmath}
+\usepackage{amsfonts}
 \usepackage{amsthm}
-
-%% Include all macros below
 
@@
 
-\newcommand{\lorem}{{\bf LOREM}}
-\newcommand{\ipsum}{{\bf IPSUM}}
-
-%% END MACROS SECTION
-
 \begin{document}
-\vspace*{0.2in}
-
-% Title must be 250 characters or less.
-\begin{flushleft}
-{\Large
-\textbf\newline{Towards attack tolerant networks: concurrent multipath routing and the butterfly network} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
-}
-\newline
-% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
-\\
-Edward L. Platt\textsuperscript{1,*},
-Daniel M. Romero\textsuperscript{1}
-\\
-\bigskip
-\textbf{1} School of Information, University of Michigan, Ann Arbor, Michigan, USA
-\\
-\bigskip
-
-% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
-% 
-% Remove or comment out the author notes below if they aren't used.
-%
-% Primary Equal Contribution Note
-
-% Additional Equal Contribution Note
-% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
-
-% Current address notes
-%\textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
-% \textcurrency b Insert second current address 
-% \textcurrency c Insert third current address
-
-% Deceased author note
-%\dag Deceased
-
-% Group/Consortium Author Note
-%\textpilcrow Membership list can be found in the Acknowledgments section.
-
-% Use the asterisk to denote corresponding authorship and provide email address in note below.
-* elplatt@umich.edu
-
-\end{flushleft}
-% Please keep the abstract below 300 words
-\section*{Abstract}
+
+%don't want date printed
+\date{}
+
+%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
+\title{\Large \bf Towards Attack-Tolerant Networks: Concurrent mutlipath routing and the butterfly network}
+
+%for single author (just remove % characters)
+\author{
+{\rm Anonymous}\\
+For Blind Review
+%\author{
+%{\rm Edward L.\ Platt}\\
+%University of Michigan School of Information
+%\and
+%{\rm Daniel M.\ Romero}\\
+%University of Michigan School of Information
+% copy the following lines to add more authors
+% \and
+% {\rm Name}\\
+%Name Institution
+} % end author
+
+\maketitle
+
+% Use the following at camera-ready time to suppress page numbers.
+% Comment it out when you first submit the paper for review.
+\thispagestyle{empty}
+
+\subsection*{Abstract}
 It is crucial for large-scale communication networks such as the internet
-to be resilient against attacks\cite{sterbenz_resilience_2010}
+to be resilient against attacks \cite{sterbenz_resilience_2010}
 such as censorship and surveillance,
@@
 Self-organized networks such as the internet’s router network typically have
-heavy-tailed degree distributions\cite{barabasi_scale-free_2009},
+heavy-tailed degree distributions \cite{barabasi_scale-free_2009},
 making them highly vulnerable to targeted attacks against central nodes
@@
 as well as a highly attack-tolerant Structured Multipath Fault Tolerance (SMFT)
-architecture which incorporates the butterfly CMR algorithm. We also identify a
-previously unexplored relationship between network topology, trust transitivity,
-and attack-tolerance, and provide a framework for further exploration of this
-relationship. Our work is the first theoretical demonstration of a point-to-point
+architecture which incorporates the butterfly CMR algorithm.
+Our work is the first theoretical demonstration of a point-to-point
 communication network architecture that can resist coercion and other
-non-technical attacks, without requiring infinitely transitive trust.
+non-technical attacks, without requiring transitive trust.
 To address cases where the network structure cannot be fully controlled,
-we demonstrate how a snapshot of the internet’s router network can be partially
+we demonstrate how a snapshot of the internet's router network can be partially
 rewired for greater attack-tolerance.
-More broadly, we hope that this work will serve as a starting point for the
-evelopment of additional topology-based attack-tolerant communication
-architectures to guard against the dangers of censorship and surveillance.
-
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLOS ONE authors please skip this step. 
-% Author Summary not valid for PLOS ONE submissions.   
-%\section*{Author summary}
-%Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Curabitur eget porta erat.
-Morbi consectetur est vel gravida pretium.
-Suspendisse ut dui eu ante cursus gravida non sed sem.
-Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
-Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
-Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
-Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
-Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
-
-\linenumbers
-
-% Use "Eq" instead of "Equation" for equation citations.
-\section*{Introduction}
+
+\section{Introduction}
 
@@
 \cite{elmer-dewitt_first_1993}.
-But 18 years after Albert et. al \cite{albert_error_2000}
-showed that the topology of the internet's router network
-makes it vulnerable to targeted attacks
-(vs. random faults), the fundamental problem of attack-tolerant network topology
+But 18 years after Albert et al.\ \cite{albert_error_2000}
+showed the internet's router network
+is vulnerable to targeted attacks,
+the fundamental problem of attack-tolerant network topology
 remains unsolved.
-Attack-tolerant topologies are desirable not only for the router network,
-but for any physical or virtual network where a compromised node puts
+Attack-tolerant topologies are desirable
+for any physical or virtual network where a compromised node puts
 communication at risk.
@@
 supporting attack-toloerant, point-to-point networked communication,
-without relying on infinitely transitive trust
+without relying on transitive trust
 \cite{christianson_why_1997}.
@@
 
-As another example, in 2013,
+Similarly, in 2013,
 the Texas-based email provider Lavabit was ordered to disclose
 their private SSL keys to the FBI \cite{poulsen_edward_2013}.
-Rather than complying,
-Lavabit ceased operations
+Lavabit instead chose to cease operations
 in order to protect their users from surveillance.
@@
 And again, the attack was on a single point of failure:
-Lavabit's web server and that server's TLS/SSL keys.
+Lavabit's web server and that server's SSL keys.
 In this case, the affected network was the
@@
 With the private keys, an attacker would be able to intercept and
-surveil traffic because the issuing certificate authority (and any
-users trusting that authority) would incorrectly trust that they were
+surveil traffic because the users would incorrectly trust that they were
 communicating with Lavabit.
-While originally intended as surveillance,
-this action effectively became an act of censorship.
-So we see that such vulnerabilities are not limited to any one system or protocol,
+So we see that such vulnerabilities are not limited to any one system,
 but result from centralized structure itself.
@@
 Our primary result is theoretical: an algorithm for constructing highly redundant
-paths in a particular network topology.
-While we motivate the need for such network topologies using examples such as
-the internet router network and webs of trust,
-we do not propose that our algorithm as a pracitcal solution for any of these
-examples.
-Rather, our result is a demonstration that such topological approaches to
-attack-tolerance are theoretically possible and suggest the importance
+paths in a particular network topology,
+suggesting the importance
 of further theoretical and applied work.
 
-We consider a setting in which a source node attempts to route a message
-to a target node, while an adversary attempts to block or intercept the message
-by compromising a number of intermediate nodes.
-We also assume that edges represent {\em direct trust},
-i.e., the belief that an adversary is unlikely to compromise a neighbor,
-as in the web of trust approach
-\cite{zimmermann_official_1995,richters_trust_2011}.
-The web of trust approach typically assumes infinite trust transitivity in that
-it is possible for one node to have some level of trust for another as long
-as they are connected by a path of directly trusted edges,
-regardless of the length of that path.
-The assumption of infinite transitivity is unrealistic
-\cite{christianson_why_1997},
-so we instead make a stricter assumption: {\em bounded trust transitivity},
-that trust transitivity can only be extended over a finite number of edges.
-Even with this more restrictive assumption, we show how trust can be established
-between two nodes even when no path of directly trusted edges exists between them.
-
-Under the above assumptions,
-we show how to evaluate the influence of network structure
-on attack-tolerance.
-We next present a structured multipath fault tolerance (SMFT) scheme to extend
-standard fault tolerance techniques to the case of adversarial faults in
-networks
-\cite{avizienis_basic_2004, von_neumann_probabilistic_1956}.
-The SMFT scheme requires the existence of a
-concurrent multipath routing (CMR) algorithm
-\cite{zin_survey_2015, qadir_exploiting_2015, khiani_comparative_2013},
-to takes advantage of the independence of faults
-along {\em independent paths}.
-We also present a novel CMR algorithm for the butterfly network topology.
-The butterfly topology is popular in parallel processing
-\cite{kshemkalyani_distributed_2008} and
-peer-to-peer \cite{lua_survey_2005, korzun_structured_2013}
-applications, due to its regular structure, low degree, and high connectivity.
-
-It is important to note that the butterfly is a highly structured and constrained
-network topology,
-very different form those found in social networks and other
-self-organized networks.
-The reader may wonder whether it is realistic or useful to assume such control over
-the network structure.
-Regardless of implementation difficulty,
-we argue that {\em attack-tolerance cannot be achieved without the ability to influence
-network structure}.
-Targeted attacks, by definition, target single points of failure.
-Attack tolerance can thus be achieved in two ways: 1. preventing failure at
-those points, or 2. preventing the existence of central points.
-Because individual points are always vulnerable to coercive, non-technological
-attacks, the former method is insufficient.
-We must instead rely on some control over topology to prevent the existence of
-single points of failure.
-The difficulty of achieving control over network topology can be mitigated by
-a number of approaches.
-Attack-tolerant networks might be sub-components of
-larger, less-constrained systems.
-For example, a single centralized server might be replaced by a distributed
-network of servers,
-each with different ownership, physical location, and legal jurisdiction,
-without placing any unrealistic constraints on the clients connecting to
-those servers.
-When attack tolerant topologies are nested (as in the case of the butterfly
-topology), multiple independent sub-components could be merged into larger
-ones over time.
-Real-world examples of structured neworks include: overlay networks
-\cite{lua_survey_2005, korzun_structured_2013},
-formal organizations \cite{mohr_explaining_1982},
-government-regulated cellular networks \cite{walker_mass_2012},
-and call tree notification systems \cite{nickerson_thinking_2010}.
-In general,
-{\em when the need for attack-tolerance is high enough to warrant investment
-in infrastructure, topology can be engineered and maintained as infrastructure}.
-
-Our main contributions are:
-\begin{itemize}
-\item{
-We propose a novel structured multipath fault tolerance (SMFT) scheme
-for extending standard fault tolerance techniques to
-{\em adversarial} faults in {\em complex networks}.
-Assuming $h$-degree {\em bounded trust transitivity},
-We show that the probability of detecting adversarial faults
-$h$-internally vertex disjoint paths.
-}
-\item{We prove that the number of $h$-internally vertex disjoint
-paths between two nodes in a wrap-around butterfly network
+We make two main contributions.
+We prove that the number of 
+$h$-internally vertex disjoint
+paths between two nodes in
+a wrap-around butterfly network
 is at least $2^h$,
 and present a scalable and efficient concurrent multipath routing (CMR) algorithm
-to find these paths,
-which can be combined with SMFT to achieve a high level of attack-tolerance.
-}
-\item{We show that rewiring a the edges of the internet's router network to
+to find these paths.
+This result is combined with structured multipath fault-tolerance (SMFT)
+to achieve a high level of attack-tolerance.
+We also show that rewiring a the edges of the internet's router network to
 resemble a butterfly network allows it to tolerate a higher number of failures
 without fragmenting, and increases the effective redundancy in the presence
-of a large number of adversarial faults.}
-\end{itemize}
-
-This paper is organized as follows.
-Section \ref{sec-related} reviews background and related work.
-Section \ref{sec-ft} describes adversarial fault tolerance on
-structured networks.
-Section \ref{sec-butterfly} gives background on the butterfly network topology.
-Section \ref{sec-bf-route} presents our concurrent multipath routing
-algorithm for the butterfly network.
-Section \ref{sec-discussion} discusses the results.
-And Section \ref{sec-conclusion} concludes.
-
-\section*{Background and Related Work}
+of a large number of adversarial faults.
+
+\section{Background and Related Work}
 \label{sec-related}
 
-There has been considerable work on trust-based attack-tolerance techniques
-in network security, both centralized and decentralized.
-Centralized approaches such as {\em public key infrastructure} (PKI)
+Centralized approaches to attack-tolerance
+such as {\em public key infrastructure} (PKI)
 suffer from a number of vulnerabilities
@@
 centralization.
-The well-known and widely-used {\em web of trust} approach
+The {\em web of trust} is a decentralized alternative
 \cite{zimmermann_official_1995,richters_trust_2011}
-is a decentralized alternative.
-In a web of trust,
-individuals can have {\em direct trust},
-as well as {\em indirect trust} for those trusted by someone directly trusted.
-Typically, this transitivity is extended to any number of hops,
-sometimes reducing trust by a multiplicative factor at each hop.
-Infinite trust transitivity is helpful for establishing a large group of
-trusted nodes, but unfortunately unrealistic
+but depends on the unrealistic assumption of trust transitivity
 \cite{christianson_why_1997}.
-Our work addresses this limitation by assuming only
-bounded trust transitivity.
-
-Previous work applying network topology to attack tolerance has
-focused on authentication,
-showing that independent paths can reduce an adversary's ability
-to impersonate a target
-\cite{levien_attack-resistant_2009}.
-Other work has shown that identifying independent paths in arbitrary networks
-is NP-hard and provided approximation algorithms
-\cite{reiter_resilient_1998}.
-Our work complements these results by extending our focus beyond authentication,
-to communication.
-When network topology can be controlled, we sidestep the NP-hard problem of finding
-independent paths on arbitrary networks by using the mathematical structure of
-the butterfly topology to construct provably independent paths.
-
-Many distributed consensus protocols (such as those used by cryptocurrencies)
+
+Many distributed consensus protocols (e.g., cryptocurrencies)
 are designed to tolerate arbitrary or adversarial faults.
@@
 However, FBA does not provide a method for constructing networks to meet
-those requirements,
-or for calculating the failure probabilities within a particular network.
+those requirements.
 
@@
 and does not require nodes to store data indefinitely.
-Our architecture also improves on the scalablity of the Fiat-Saia network,
-and makes requirements about network topology explicit.
 
@@
 Our proposed architecture is differentiated from existing systems by several
-properties (Table \ref{tab:compare}).
+properties.
 Decentralized architectures are more resistant to coercion
@@
 and man-in-the-middle attacks \cite{nayak_different_2010}.
-Trust-based systems are more sustainable than proof-of-work.
-Bounded-trust systems do not require the unrealistic assumption of infinite
-trust transitivity.
 Topological approaches address the root cause of vulnerability
@@
 rather than relying on technology that can be side-stepped through coercion.
+The topological approach is also more sustainable than proof-of-work systems.
 Point-to-point communication allows two individuals to exchange messages
@@
 
-\begin{table*}[t!]
-\caption{Comparison of attack-tolerant network communication architectures.\label{tab:compare}}
-\begin{center}
-\begin{tabular}{l|p{0.6in}p{0.6in}p{0.6in}p{0.6in}p{0.6in}}
-                              & Decentra-lized & Trust-based & Bounded-trust & Topo-logical & Point-to-point \\
-\hline
-PKI            &               & \cmark      & \cmark        &             & \cmark     \\
-Web of Trust   & \cmark        & \cmark      &               &             & \cmark     \\
-Freenet        & \cmark        & \cmark      & \cmark        &             &            \\
-FBA            & \cmark        & \cmark      & \cmark        &             & \cmark     \\
-Proof of Work  & \cmark        &             & \cmark        &             &            \\
-Fiat-Saia      & \cmark        & \cmark      & \cmark        & \cmark      &            \\
-SMFT           & \cmark        & \cmark      & \cmark        & \cmark      & \cmark     \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-
-\section*{Trust Networks and Fault Tolerance}
+%\begin{table*}[t!]
+%\caption{Comparison of attack-tolerant network communication architectures.\label{tab:compare}}
+%\begin{center}
+%\begin{tabular}{l|p{0.6in}p{0.6in}p{0.6in}p{0.6in}p{0.6in}}
+%                              & Decentra-lized & Trust-based & Bounded-trust & Topo-logical & Point-to-point \\
+%\hline
+%PKI            &               & \cmark      & \cmark        &             & \cmark     \\
+%Web of Trust   & \cmark        & \cmark      &               &             & \cmark     \\
+%Freenet        & \cmark        & \cmark      & \cmark        &             &            \\
+%FBA            & \cmark        & \cmark      & \cmark        &             & \cmark     \\
+%Proof of Work  & \cmark        &             & \cmark        &             &            \\
+%Fiat-Saia      & \cmark        & \cmark      & \cmark        & \cmark      &            \\
+%SMFT           & \cmark        & \cmark      & \cmark        & \cmark      & \cmark     \\
+%\hline
+%\end{tabular}
+%\end{center}
+%\end{table*}
+
+\section{Networks and Fault Tolerance}
 \label{sec-ft}
-
-Within the field of {\em fault tolerance},
-many techniques have been developed for building reliable systems
-out of unreliable components
-\cite{avizienis_basic_2004, von_neumann_probabilistic_1956}.
-We will make use of standard fault tolerance terminology, summarized here.
-A {\em fault} is occurs when one component
-of a system behaves incorrectly (e.g., a routing node blocks or
-alters a message).
-The result of that fault (e.g., a recipient receiving conflicting messages)
-is an {\em error} state.
-If the error is undetected or corrected to the wrong value,
-the system has experienced a {\em failure} (e.g., an altered message is
-accepted as authentic).
-Note that when an error is detected but cannot be corrected,
-the system has still tolerated the fault because it has not accepted an error
-state.
-We are concerned in particular with {\em adversarial faults},
-which are chosen strategically to maximize the likelihood of a failure.
-
-\subsection*{Multipath Fault Tolerance}
-
-Standard fault tolerance methods use redundancy to detect and correct
+\subsection{Multipath Fault Tolerance}
+
+Standard fault tolerance methods
+\cite{avizienis_basic_2004, von_neumann_probabilistic_1956}
+use redundancy to detect and correct
 statistically independent faults.
@@
 For now, let us assume our sender (Alice) and receiver (Bob) are connected
-by $\delta$ direct channels, with independent errors.
-We will return to the question of constructing these channels in subsequent
-sections.
-For now, we concern ourselves with the question:
-given that the network provides $\delta$ redundant channels between
-Alice and Bob,
-what is the probability that an adversary (Mal) causes an undetectable
-error after causing faults in a fixed number of channels?
-
-Let us first consider the scenario in which
-Alice sends a message copy over each available channel.
-We can also assume that each message includes the number of messages sent,
-the full list of channels used, etc.,
-making that information available to Bob.
-When Bob receives the messages, there are several possibilities.
-If some of the messages are missing
-or if some of the messages disagree,
-Bob knows that some of the messages were either blocked or altered,
-and he has successfully tolerated the fault(s).
-Bob can then take any of several actions:
-1. request re-transmission;
-2. send receipts so Alice knows which paths have been compromised;
-or 3. attempt error correction using majority voting.
-If instead, Bob finds that all the messages are present and agree,
-there are two possible cases.
-The first case is that Mal has not compromised any of the messages,
-and Bob has correctly accepted them, so no failure has occurred.
-The second case is that Mal has compromised {\em all} of the messages,
-so Bob has accepted an erroneous message and a failure has occurred.
-In the present scenario,
-whether a failure occurs depends only on whether Mal has the resources to
-compromise all of the channels.
-In a more realistic scenario,
-both Alice and Mal have limited resources and are not able to use or
-compromise all available channels.
-
-In a more sophisticated multipath fault tolerance scheme,
-Alice randomly chooses $k \leq \delta$ channels and sends a copy of
-her message on each.
-We assume that Mal is capable of compromising $l \leq \delta$ channels.
-Since Alice chooses channels randomly,
-all channels are equally likely to contain a message,
-so Mal can do no better than also choosing randomly.
-If $k > l$, at least one message will get through uncompromised and all
-errors are detectable.
-Otherwise, the probability of Mal producing an undetectable error is
-the probability that all of Alice's chosen channels are compromised:
-\beq
-\label{eq:pf}
-p_f &=& \frac{l!(\delta-k)!}{\delta!(l-k)!}.
-\eeq
-Letting $k=\alpha \delta$ and $l=\beta \delta$, then applying Stirling's
-approximation gives:
+by $\delta$ {\em independent paths} such that no two paths contain the same
+faulty node.
+Let $k = \alpha\delta$ be the number of messages Alice can send and
+$l = \beta\delta$ be the number of nodes Mal can compromise.
+If Alice choses paths at random and Mal compromises paths at random,
+the probability of blocking or altering all of Alice's messages can be
+shown to be approximately:
 \begin{eqnarray}
@@
 depends on the fractions of
-channels actually utilized $\alpha$ and compromised $\beta$,
+paths actually utilized $\alpha$ and compromised $\beta$,
 it decreases exponentially with $\delta$.
-This result is significant because,
-as we will soon show, $\delta$
-depends only on the network structure
-and the strength of trust transitivity.
-{\em Thus, the scheme can be effective, even when the number of channels used
+This result is significant because
+$\delta$ depends only on the network structure.
+{\em Thus, the scheme can be effective, even when the number of paths used
 $k$ is a small fraction of the channels available}.
@@
 
-\begin{figure}[!h]
-\centerline{\includegraphics{fig-perror}}
+\begin{figure}[h]
+\centerline{\includegraphics{fig-perror-25}}
 \caption{
@@
 
-\subsection*{Bounded Trust Model}
-
-So far, we have assumed that Alice and Bob have access to some number
-$\delta$ of channels with statistically independent faults.
-However, in real communication architectures, direct links between all
-pairs of individuals are not possible and messages must be routed through
-a number of intermediate nodes.
-In an adversarial setting,
-the existence of intermediate nodes introduces two problems:
-1. intermediate nodes may be compromised by the adversary and
-2. faults on paths are no longer statistically independent:
-two paths may pass through the same compromised node.
-We show how to how to resolve these problems using a combination of network
-structure and bounded trust transitivity.
-
-Trust-based approaches to secure communication assume that some parties
-cannot be compromised.
-One common approach, the web of trust
-\cite{zimmermann_official_1995,richters_trust_2011}.
-Alice has {\em direct trust} for some number of nodes.
-Alice has {\em indirect trust} for nodes separated by two or more hops of
-direct trust.
-Typically, this transitive trust is extended indefinitely,
-sometimes reduced by a multiplicative factor at each step.
-However, the assumption of infinitely transitive trust is unrealistic
-\cite{christianson_why_1997}.
-Furthermore, infinite trust transitivity obscures the importance of network structure,
-as it depends only on whether some path exists, not the number or quality of paths.
-We adopt a simpler, yet more realistic {\em bounded trust} assumption:
-that nodes up to some fixed number of hops cannot be compromised,
-and that those beyond can.
-This assumption will be convenient for proving our results,
-which we now proceed to define formally.
-
-We define the {\em bounded trust model} on
-an undirected graph $G = (V,E)$,
-although the model can easily be extended to directed multigraphs.
-Vertices representing communicating parties,
-with edges representing mutually trusted communication links.
-We define a {\em trust radius} $h$ such that nodes $v$ and
-$w$ trust each other if their distance is less than $h$.
-For a given node $v$,
-we call the set of trusted nodes its
-{\em trusted neighborhood} $T_h(v)$,
-and all nodes at exactly distance $h$ the
-{\em trust boundary} $B_h(v)$:
-\beq
-T_h(v) &=& \left\{ w \mid d(v,w) < h \right\} \\
-B_h(v) &=& \left\{ w \mid d(v,w) = h \right\}.
-\eeq
-The trust boundary $B_h$ plays an important role because these nodes are not
-trusted by $u$,
-and if compromised can entirely isolate $v$ from the rest of the network.
-
-Now let $s \in V$ be an arbitrary sender
-and $t \in V$ be an arbitrary receiver.
-We assume the presence of an adversary who knows the
-full structure of the network,
-and who can compromise a fixed number of nodes,
-gaining complete control of their behavior.
-We also assume that the adversary is specifically targeting communication
-between $s$ and $t$ and
-can compromise any node except for those trusted by $s$ or $t$.
-Under these trust assumptions, adversarial faults can only occur
-outside the trusted neighborhoods of $s$ and $t$:
-$V \setminus \left(T_h(v) \cup T_h(w)\right)$.
-We refer to this set of nodes as the {\em untrusted region}.
-We now show how it is possible to communicate reliably,
-even when all available paths go through the untrusted region.
-
-\subsection*{Effective Redundancy}
-
-Our approach is to achieve fault tolerance through redundancy.
-To do so, we must use only
-{\em independent paths} \cite{reiter_resilient_1998},
-which have no common points of failure.
-Typically, it is assumed that in order to be independent,
-paths must be internally vertex disjoint, i.e.,
-have no nodes in common except the endpoints.
-However, under the bounded trust model,
-intersecting paths can still be independent if their
-intersection contains only trusted nodes.
-We define two paths with common endpoints
-to be {\em $h$-internally vertex disjoint}
-if all common vertices are less than distance $h$ from
-one of the endpoints.
-This condition holds if and only if two paths are independent
-under the bounded trust model with radius $h$.
-
-When trust radius $h$ is assumed,
-the number of $h$-internally vertex disjoint paths between
-two nodes $s$ and $t$ represents the number of
-channels that can be constructed between them having
-statistically independent faults.
-We thus refer to this quantity as the 
-{\em effective redundancy} $\delta_{s,t,h}$.
-The effective redundancy can also be interpreted as the
-max-flow/min-cut of a graph after each trusted neighborhood has been
-collapsed into a single vertex.
-The trust boundaries form a cut of the network and place an upper bound on the
-min-cut:
-\beq
-\delta_{v,w,h} \leq \min\left( \mid B_h(s) \mid, \mid B_h(t) \mid \right).
-\eeq
-Equality holds when there are no bottlenecks within the untrusted region,
-an indication that the network is decentralized.
-The effective redundancy of the entire graph can be characterized by the minimum
-over all vertex pairs:
+\subsection{Effective Redundancy}
+
+The effect of an attack on communication between two nodes
+depends on how many of the paths between those nodes pass through
+the a compromised node.
+The number of such compromised paths depends on both the network topology
+and the position of the endpoint nodes relative to the compromised node.
+If the compromised node is one of the endpoints,
+then all paths from that node are compromised.
+When the endpoints are farther from the compromised node,
+the paths from the endpoints branch out into many
+paths before one of those passes through the compromised node.
+
+The network topology determines how many redundant paths exist
+less than a given distance $h$ away from each node.
+With more paths, any single compromised node has a smaller effect.
+For a given node pair $(s, t)$,
+We call this number of paths $\delta_{h,s,t}$ the
+{\em pairwise effective redundancy}.
+For the entire network, the {\em effective redundancy} is the minimum over all
+vertex pairs:
 \beq
@@
 \eeq
-Thus, for any pair of nodes in the network, at least $\delta_h$ independent,
-redundant paths can be constructed between them.
-The more quickly $\delta_h$ grows with $h$,
-the better a network is at leveraging trust transitivity to create redundancy.
-Thus, the scaling of $\delta_h$ can be used to quantify a network's ability
-to withstand targeted attacks,
-even when the exact trust radius $h$ is unknown.
-
-\begin{figure}
-\centerline{\includegraphics[width=4in,height=2.78in]{fig-partial-trust.pdf}}
-\caption{
-Illustration of a trusted communication network and the network properties
-used by the {\em bounded trust model}.
-Edges represent mutually trusted communication links.
-The sender (Alice, $s$) and receiver (Bob, $t$) trust all nodes
-less than the {\em trust radius} $h$ hops away.
-These nodes form their {\em trusted neighborhoods} $T_h(v)$ and $T_h(w)$.
-We assume that all faults occur in the remaining nodes: the
-{\em untrusted region}.
-The untrusted nodes in contact with the trusted neighborhoods for the
-{\em trust boundaries} $B_h(v)$ and $B_h(w)$,
-which (in the absence of central bottlenecks) determine the
-{\em effective redundancy} $\delta_h$ provided by the network.
-Alice and Bob can achieve the same level of attack-tolerance
-as if they were directly connected by $\delta_h$ redundant channels.
-}
-\label{fig:trust-source}
-\end{figure}
-
-\subsection*{Structured Multipath Fault Tolerance}
+
+In graph theoretical terms, the effective redundancy at distance $h$ is
+the minimum number of $h${\em-internally vertex-disjoint} paths between any
+node pair.
+A higher effective redudncancy means that,
+for a given distance $h$ between an endpoint and a compromised node,
+the more redundant paths are available to route around that attack.
+In the absence of centralized bottlenecks,
+the effective redundancy is limited by the endpoint
+closest to an attack.
+
+\subsection{Structured Multipath Fault Tolerance}
 
@@
 
-\section*{The Butterfly Network Topology}
+\section{The Butterfly Network Topology}
 \label{sec-butterfly}
@@
 \cite{kshemkalyani_distributed_2008}.
-The butterfly network is highly structured,
-making it most suitable for applications where portions of the network structure
-can be controlled or influenced.
-The butterfly network is also recursive, with larger versions composed out of
+The butterfly network is recursive, with larger versions composed out of
 multiple smaller versions,
-making it possible for independently organized
-attack-tolerant networks to merge into larger ones over time.
-More flexible architectures may be possible,
-but attack-tolerance will always require some level of influence over
-network structure in order to limit single points of failure.
+suggesting large attack-tolerant networks could be constructed by merging smaller ones.
 To address the case when the network cannot be fully controlled,
@@
 
-\subsection*{Butterfly Network Topology}
+\subsection{Butterfly Network Topology}
 
@@
 While several variations on the butterfly network exist,
-we utilize the $m$-dimensional, directed wrap-around butterfly
-(Fig. \ref{fig:bf-route}),
+we utilize the $m$-dimensional, directed wrap-around butterfly,
 denoted $\wbf(m)$:
@@
 known as {\em the place-within-level}.
-There are two types of edges: down, and down-right
-(shown in Fig. \ref{fig:butterfly}).
+There are two types of edges: down, and down-right.
 Down edges ($E_\downarrow$) connect nodes sharing the same $z$ value
@@
 
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[height=3in]{fig-bf-route}
-\end{center}
-\caption{
-A 3-dimensional wraparound butterfly network.
-Note that the rightmost nodes are the same nodes as the leftmost,
-drawn twice for visual clarity.
-The highlighted nodes and edges show the path from node (0,000)
-to node (1,101).
-\label{fig:bf-route}
-}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics{fig-butterfly}
-\end{center}
-\caption{
-Schematic illustration of the two types of edges in a directed butterfly
-network.
-The node $(l,z)$ is shown as the bit string $z$ with a square around the
-$l$th bit.
-``Down'' edges increment $l$, leaving $z$ unchanged,
-while ``down-right'' edges increment $l$ and invert the $l$th bit of $z$.
-In the wrap-around variant, the nodes with maximum $l$ have down and down-right
-edges to the nodes with $l=0$.
-\label{fig:butterfly}
-}
-\end{figure}
+%\begin{figure}[h]
+%\begin{center}
+%\includegraphics[width=2.66in]{fig-bf-route}
+%\end{center}
+%\caption{
+%A 3-dimensional wraparound butterfly network.
+%Note that the rightmost nodes are the same nodes as the leftmost,
+%drawn twice for visual clarity.
+%The highlighted nodes and edges show the path from node (0,000)
+%to node (1,101).
+%\label{fig:bf-route}
+%}
+%\end{figure}
+
+%\begin{figure}[h]
+%\begin{center}
+%\includegraphics{fig-butterfly}
+%\end{center}
+%\caption{
+%Schematic illustration of the two types of edges in a directed butterfly
+%network.
+%The node $(l,z)$ is shown as the bit string $z$ with a square around the
+%$l$th bit.
+%``Down'' edges increment $l$, leaving $z$ unchanged,
+%while ``down-right'' edges increment $l$ and invert the $l$th bit of $z$.
+%In the wrap-around variant, the nodes with maximum $l$ have down and down-right
+%edges to the nodes with $l=0$.
+%\label{fig:butterfly}
+%}
+%\end{figure}
 
@@
 Specifically, we will prove below that the number of
-$h$-internally vertex disjoint paths between two
+$h$-internally vertex-disjoint paths between two
 nodes increases exponentially with $h$.
@@
 The structure of the butterfly network lends itself to a well-known
-(unipath) routing algorithm (Fig. \ref{fig:bf-route}),
+(unipath) routing algorithm,
 which we later extend to the multipath case.
@@
 
-\subsection*{Butterfly Rewiring}
+\subsection{Butterfly Rewiring}
 
@@
 central nodes.
-The effective redundancy for various values of trust transitivity $h$
-are shown in Fig. \ref{fig:mincut}.
-The effective redundancy is calculated by collapsing nodes and edges within
-$h$ hops of source and destination into single nodes and finding the min-cut
-between them,
-averaging over 150 source-destination pairs.
-For $h=0$, the rewired version has strictly higher redundancy.
-For $h>0$,
-the original network has higher redundancy when the number of failures is small,
-while the rewired network has higher redundancy beyond a crossover point.
-We interpret these results to suggest that even when a small number of
-highly central nodes have been removed from a heavy-tailed network,
-most nodes are still able to take advantage of the remaining hubs.
-As larger hubs continue to be removed, the connectivity of the network
-decreases until the crossover point,
-at which point the rewired network offers higher effective redundancy.
 
 \begin{figure}
-\centerline{\includegraphics{fig-diameter}}
+\begin{center}
+\includegraphics[width=2.5in]{fig-diameter}
 \caption{
@@
 this value doubles to 2\%.
-With 90\% of the edges rewired,
-the network remains unfragmented beyond the failure of the 8\% most
-central nodes.
-The rewiring scheme does not require adding any additional edges.
 }
+\end{center}
 \label{fig:diameter}
@@
 
-\begin{figure*}[h!]
-\centerline{\includegraphics[width=5in,height=5in]{fig-mincut}}
-\caption{
-Simulation of targeted attacks against a snapshot of the internet's router
-network with a fraction of the edges rewired into a partial butterfly configuration.
-The effective redundancy is shown for several values of trust transitivity $h$.
-For $h>0$, the original network has higher effective redundancy up to a crossover point,
-after which the rewired network performs better.
-}
-\label{fig:mincut}
-\end{figure*}
-
-\section*{Multipath Butterfly Routing}
+%\begin{figure*}[h]
+%\centerline{\includegraphics[width=6in]{fig-mincut}}
+%\caption{
+%Simulation of targeted attacks against a snapshot of the internet's router
+%network with a fraction of the edges rewired into a partial butterfly configuration.
+%The effective redundancy is shown for several values of trust transitivity $h$.
+%For $h>0$, the original network has higher effective redundancy up to a crossover point,
+%after which the rewired network performs better.
+%}
+%\label{fig:mincut}
+%\end{figure*}
+
+\section{Multipath Butterfly Routing}
 \label{sec-bf-route}
 
-We now present a routing algorithm to construct $2^h$ independent paths
+We now present a routing algorithm to construct $2^h$
+$h${\em-internally vertex-disjoint} paths
 between two nodes in a butterfly network,
-where $h$ is the trust radius under the partial trust model.
-Informally, Alice sends each message to a distinct node on her
-trust boundary, then to a distinct intermediate node in the untrusted region,
-then to a distinct node on Bob's trust boundary, and finally to Bob.
+where $h$ is minimum distance from an endpoint to a compromised node.
+Informally, Alice sends each message $h$ hops,
+then to a distinct intermediate node,
+then to a node $h$ hops from Bob, and finally to Bob.
 The intermediate nodes are in a sense ``far'' from each other and ensure that
-no two paths overlap in the untrusted region.
-Each path can be parameterized by a single integer $s$, which identifies
-the specific node on Alice's trust boundary
-(or equivalently the node on Bob's trust boundary, or the untrusted intermediate).
-
-The algorithm guarantees paths are independent by ensuring that
-(outside the trusted neighborhoods)
-they only include
-nodes that match the path parameter $s$ at certain indexes in their
-place-within-level.
-Since each path has a unique parameter $s$,
-its set of untrusted nodes is disjoint from all other paths.
-As with the unipath routing algorithm,
-each of the multiple paths proceed from a source $v$ to a destination $u$
-using down and down-right edges,
-cycling through levels one at a time.
-However, we cycle through the levels twice, once to route from $v$ to a
-particular path's intermediary node,
-and again to route from the intermediary to $w$.
-Each cycle is divided into stages,
-with different properties used to prove independence at each stage
-(see Fig. \ref{fig:route-overview}).
-In the first cycle (stages 1--4), path independence is guaranteed by ensuring that
-all nodes match the path parameter $s$ in the first $h$ bits of the place-within-level.
-Similarly, in the second cycle (stages 5--7),
-independence is guaranteed by ensuring that all paths match $s$ in the
-$h$ bits of the place-within-level preceding the destination index.
-A full example is illustrated in Fig. \ref{fig:routing}.
-
-\begin{table}[h!]
-\caption{Butterfly Multipath Routing Variables\label{tab:routing}}
-\begin{center}
-\begin{tabular}{ll}
-Name & Variable \\\hline
-butterfly dimension & $m \in \mathbb{Z}_+$ \\
-node level & $l \in \mathbb{Z} : 0 \leq l < m$ \\
-node place within level & $z \in \mathbb{Z}_2^m$ \\
-trust radius & $h \in \mathbb{Z} : 1 \leq h \leq \lfloor m/2 \rfloor$ \\
-path index & $s \in \mathbb{Z}_2^h$ \\
-\end{tabular}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-\begin{center}
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-                        ]
-\foreach \i in {0\ldots,{\ldots}0\ldots,{\ldots}0\ldots,{\ldots}0}% <-- content of nodes
-    \node[X] {\i};
-\draw[<->] ([yshift=1.5mm] A-1.north east) -- node[above=0.25mm] {$h$} ([yshift=1.5mm] A-1.north west);
-\draw[<->] ([yshift=1.5mm] A-2.north east) -- node[above=0.25mm] {$l_w - 2h$} ([yshift=1.5mm] A-2.north west);
-\draw[<->] ([yshift=1.5mm] A-3.north east) -- node[above=0.25mm] {$h$} ([yshift=1.5mm] A-3.north west);
-\draw[<->] ([yshift=1.5mm] A-4.north east) -- node[above=0.25mm] {$m - l_w$} ([yshift=1.5mm] A-4.north west);
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {\underline{Stage}} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick, fill=blue!20!white},
-                        ]
-\node[Y] {$s$};
-\foreach \i in {{\ldots}0\ldots,{\ldots}0\ldots,{\ldots}0}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {1.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick, fill=blue!20!white},
-                        ]
-\foreach \i in {$s$}% <-- content of nodes
-    \node[X] {\i};
-\node[Y] {{\ldots}1\ldots};
-\foreach \i in {{\ldots}0\ldots,{\ldots}0}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {2.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick, fill=blue!20!white},
-                        ]
-\foreach \i in {$s$,{\ldots}1\ldots}% <-- content of nodes
-    \node[X] {\i};
-\node[Y] {$\tilde{s}$};
-\foreach \i in {{\ldots}0}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {3.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {anchor=base, rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick, fill=blue!20!white},
-                        ]
-\foreach \i in {$s$,{\ldots}1\ldots,$\tilde{s}$}% <-- content of nodes
-    \node[X] {\i};
-\node[Y] {$z_{w,D}$};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {4.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {anchor=base, rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick, fill=blue!20!white},
-                        ]
-\node[Y] {$z_{w,A}$};
-\foreach \i in {{\ldots}1\ldots,$\tilde{s}$,$z_{w,D}$}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {5.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {anchor=base, rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick, fill=blue!20!white},
-                        ]
-\foreach \i in {$z_{w,A}$}% <-- content of nodes
-    \node[X] {\i};
-\node[Y] {$z_{w,B}$};
-\foreach \i in {$\tilde{s}$,$z_{w,D}$}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {6.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {anchor=base, rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick, fill=blue!20!white},
-                        ]
-\foreach \i in {$z_{w,A}$,$z_{w,B}$}% <-- content of nodes
-    \node[X] {\i};
-\node[Y] {$z_{w,C}$};
-\foreach \i in {$z_{w,D}$}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {7.} ( A-1.west);
-\end{tikzpicture}
-\end{center}
-\caption{
-Progression of place-within-level $z$ as the multipath routing algorithm
-cycles through the levels of the butterfly network.
-\label{fig:route-overview}
-}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics{fig-routing}
-\end{center}
-\caption{
-An example of one path as constructed by the proposed multipath
-routing algorithm.
-The path is shown for $s = 10_2$
-and $w = (6, 0110111_2)$.
-\label{fig:routing}
-}
-\end{figure}
-
-\subsection*{Algorithm Specification}
+no two paths overlap.
+Each path can be parameterized by a single integer $s$.
+
+%\begin{table}[h]
+%\caption{Butterfly Multipath Routing Variables\label{tab:routing}}
+%\begin{center}
+%\begin{tabular}{ll}
+%Name & Variable \\\hline
+%butterfly dimension & $m \in \mathbb{Z}_+$ \\
+%node level & $l \in \mathbb{Z} : 0 \leq l < m$ \\
+%node place within level & $z \in \mathbb{Z}_2^m$ \\
+%trust radius & $h \in \mathbb{Z} : 1 \leq h \leq \lfloor m/2 \rfloor$ \\
+%path index & $s \in \mathbb{Z}_2^h$ \\
+%\end{tabular}
+%\end{center}
+%\end{table}
+
+%\begin{figure}[h]
+%\begin{center}
+%\includegraphics{fig-routing}
+%\end{center}
+%\caption{
+%An example of one path as constructed by the proposed multipath
+%routing algorithm.
+%The path is shown for $s = 10_2$
+%and $w = (6, 0110111_2)$.
+%\label{fig:routing}
+%}
+%\end{figure}
+
+\subsection{Algorithm Specification}
 
@@
 wrap-around butterfly network.
-For convenience, the relevant variables are summarized in Table \ref{tab:routing}.
 Utilizing vertex transitivity, we label the source node as
@@
 
-\subsection*{Proof of Path Independence}
+\subsection{Proof of Path Independence}
 
@@
 
-Thus, assuming the partial trust model with trust transitive
-for $h$ hops, we can construct $2^h$ paths on a wrap-around butterfly topology
-which do not intersect outside the trusted neighborhoods of the source and
-destination.
-Note that the node sequence $v_s^{(t)}$ can be calculated entirely
-from the source $v$, destination $w$, and path parameter $s$,
-meaning that with this information nodes are able to determine which neighbor
-to route a given message copy to.
-Furthermore, the existence of $2^h$ paths places a lower bound on the
-effective redundancy $\delta_h$,
-showing that the decentralized, redundant, structured networks such as the
-butterfly can have a very low probability of failure when faced with
-adversarial faults, even from a very powerful attacker.
-
-\section*{Discussion}
+\section{Discussion}
 \label{sec-discussion}
 
-Our work has been motivated by the vulnerability of current communications
-infrastructure to surveillance and censorship,
-which are often achieved by coercive targeted attacks against central nodes.
-We have already discussed two such cases:
-Pakistan's inadvertant censorship of YouTube
-\cite{hunter_pakistan_2008}
-and the FBI's surveillance and censorship of Lavabit
-\cite{poulsen_edward_2013}.
-
-The reader may wonder how our methods could be employed in scenarios
-such as large-scale state-sponsored censorship
-\cite{xu_internet_2011}.
-Censorship-resistant infrastructure often replaces central servers
-(e.g., the router in the 2008 YouTube incident) with multiple servers across
-the world, synchronized through consensus protocols.
-The {\em directory authorities} used by the Tor project
-\cite{dingledine_tor:_2004} are one example.
-However, the size of such networks is limited by the number of
-trusted relationships (degree) each node can maintain, and the inherent insecurity of
-extending transitive trust to an ever-larger network.
-Our work provides both a theoretical framework
-and a specific example of how network structure
-can be engineered to leverage trust for a high level of attack-tolerance,
-without sacrificing scalability.
-
-We have focused primarily on adversarial faults that block or
-change messages (censorship) but our work is also relevant to
-surveillance.
-While cryptographic anti-surveillance techniques exist,
-they remain vulnerable to man-in-the-middle attacks,
-in which an intermediate node masquerades as the destination.
-Such attacks can be detected if the original message reaches the true
-destination unaltered,
-which SMFT can help to ensure.
-
 In its current form, our work has several limitations.
-Most obviously, it requires complete control over the network structure.
-However, we have shown that even partially control over network structure can
+Most obviously, it requires control over the network structure.
+However, we have shown that even partial control over network structure can
 improve attack tolerance properties.
-Still a more flexible network structure is desirable.
 There is also the question of how to construct such a network without a central
 authority.
-This limitation may not be as severe as it seems,
-due to the nested structure of the butterfly network.
 We conjecture that smaller independently-formed networks could be
 merged into a single larger network without central coordination.
-When nodes and edges exist in geographic space (as in cables connecting internet
-routers) this scheme might require connections between very distant locations.
+When nodes are tied to geographic locations,
+the butterfly topology would require connections between very distant locations.
 Such connections are extremely expensive to construct and maintain,
-although many exist today in the form of redundant internet backbones and
-undersea cables. An alternative solution might involve satellite links,
+although they do exist in the form of internet backbones.
+An alternative solution might involve satellite links,
 which connect distant geographic points much more easily.
@@
 
-In addition to addressing the above limitations,
-we see several potential directions for future work.
-The development of new structured networks or multipath routing algorithms
-could achieve higher levels of redundancy and attack-tolerance.
-It is also desirable to examine how changes to social dynamics could shift
-self-organized networks towards a more decentralized structure.
-Finally, our results could be implemented to address specific applications, e.g.,
-secure messaging, domain name resolution, or anonymous web browsing.
-
-\section*{Conclusion}
+\section{Conclusion}
 \label{sec-conclusion}
@@
 attack-tolerant point-to-point communication architecture.
-We have also shown how assuming bounded trust
-transitivity can enable a quantitative analysis of the
-relationships between network structure, trust, and attack-tolerance.
-In our architecture,
-the probability of an adversary causing an undetectable error
-decreases exponentially with the network's effective redundancy.
-The effective redundancy, in the case of the butterfly topology,
-grows exponentially with the radius of trust transitivity.
-Furthermore, a small increase in the number of messages sent
-(traffic volume) can compensate
-for a large increase in the number of messages compromised by an adversary.
-These results require some control over the structure of a network,
-or some portion of the network.
 Even when network structure cannot be perfectly controlled,
@@
 We believe that this work provides a foundation for the
-development of additional topology-based communication
-architectures to guard against technical and coercive adversarial attacks,
+development of topology-based 
+architectures to guard against adversarial attacks,
 including censorship and surveillance.
 
-\section*{Acknowledgments}
-The authors would like to thank Tony Garnock-Jones, A. Frederick Dudley, and
-Nathaniel Bezanson for helpful conversations.
-This research was partly supported by the National Science Foundation under
-Grant No. IIS-1617820.
-
-\nolinenumbers
-
-%\bibliography{references}
-
-% Either type in your references using
-% \begin{thebibliography}{}
-% \bibitem{}
-% Text
-% \end{thebibliography}
-%
-% or
-%
-% Compile your BiBTeX database using our plos2015.bst
-% style file and paste the contents of your .bbl file
-% here. See http://journals.plos.org/plosone/s/latex for 
-% step-by-step instructions.
-% 
-\begin{thebibliography}{10}
-
-\bibitem{sterbenz_resilience_2010}
-Sterbenz JP, Hutchison D, Çetinkaya EK, Jabbar A, Rohrer JP, Schöller M,
-  et~al.
-\newblock Resilience and survivability in communication networks: {Strategies},
-  principles, and survey of disciplines.
-\newblock Computer Networks. 2010;54(8):1245--1265.
-
-\bibitem{barabasi_scale-free_2009}
-Barabási AL, {others}.
-\newblock Scale-free networks: a decade and beyond.
-\newblock Science. 2009;325(5939):412.
-
-\bibitem{albert_error_2000}
-Albert R, Jeong H, Barabási AL.
-\newblock Error and attack tolerance of complex networks.
-\newblock Nature. 2000;406(6794):378--382.
-
-\bibitem{nayak_different_2010}
-Nayak GN, Samaddar SG.
-\newblock Different flavours of man-in-the-middle attack, consequences and
-  feasible solutions.
-\newblock In: Computer {Science} and {Information} {Technology} ({ICCSIT}),
-  2010 3rd {IEEE} {International} {Conference} on. vol.~5. IEEE; 2010. p.
-  491--495.
-
-\bibitem{grewal_caveat_2013}
-Grewal GS, Ryan MD, Bursuc S, Ryan PY.
-\newblock Caveat coercitor: {Coercion}-evidence in electronic voting.
-\newblock In: Security and {Privacy} ({SP}), 2013 {IEEE} {Symposium} on. IEEE;
-  2013. p. 367--381.
-
-\bibitem{elmer-dewitt_first_1993}
-Elmer-Dewitt P, Jackson D.
-\newblock First nation in cyberspace.
-\newblock Time. 1993;6:62--64.
-
-\bibitem{baran_distributed_1964}
-Baran P, {others}.
-\newblock On distributed communications.
-\newblock Volumes I-XI, RAND Corporation Research Documents, August. 1964; p.
-  637--648.
-
-\bibitem{dainotti_analysis_2011}
-Dainotti A, Squarcella C, Aben E, Claffy KC, Chiesa M, Russo M, et~al.
-\newblock Analysis of country-wide internet outages caused by censorship.
-\newblock In: Proceedings of the 2011 {ACM} {SIGCOMM} conference on {Internet}
-  measurement conference. ACM; 2011. p. 1--18.
-
-\bibitem{christianson_why_1997}
-Christianson B, Harbison WS.
-\newblock Why isn't trust transitive?
-\newblock In: Security protocols. Springer; 1997. p. 171--176.
-
-\bibitem{zin_survey_2015}
-Zin SM, Anuar NB, Kiah MLMM, Ahmedy I.
-\newblock Survey of secure multipath routing protocols for {WSNs}.
-\newblock J Netw Comput Appl. 2015;55:123--153.
-
-\bibitem{barabasi_emergence_1999}
-Barabási AL, Albert R.
-\newblock Emergence of scaling in random networks.
-\newblock Science. 1999;286(5439):509--512.
-
-\bibitem{hunter_pakistan_2008}
-Hunter P.
-\newblock Pakistan {YouTube} block exposes fundamental internet security
-  weakness: {Concern} that pakistani action affected youtube access elsewhere
-  in world.
-\newblock Computer Fraud \& Security. 2008;2008(4):10--11.
-
-\bibitem{poulsen_edward_2013}
-Poulsen K.
-\newblock Edward {Snowden}’s e-mail provider defied {FBI} demands to turn
-  over crypto keys, documents show.
-\newblock WIRED. 2013;.
-
-\bibitem{zimmermann_official_1995}
-Zimmermann PR.
-\newblock The official {PGP} user's guide.
-\newblock MIT press; 1995.
-
-\bibitem{richters_trust_2011}
-Richters O, Peixoto TP.
-\newblock Trust transitivity in social networks.
-\newblock PloS one. 2011;6(4):e18384.
-
-\bibitem{avizienis_basic_2004}
-Avizienis A, Laprie JC, Randell B, Landwehr C.
-\newblock Basic concepts and taxonomy of dependable and secure computing.
-\newblock IEEE T Depend Secure. 2004;1(1):11--33.
-
-\bibitem{von_neumann_probabilistic_1956}
-Von~Neumann J.
-\newblock Probabilistic logics and the synthesis of reliable organisms from
-  unreliable components.
-\newblock Automata studies. 1956;34:43--98.
-
-\bibitem{qadir_exploiting_2015}
-Qadir J, Ali A, Yau KLA, Sathiaseelan A, Crowcroft J.
-\newblock Exploiting the power of multiplicity: a holistic survey of
-  network-layer multipath.
-\newblock IEEE Comm Surv Tut. 2015;17(4):2176--2213.
-
-\bibitem{khiani_comparative_2013}
-Khiani SR, Dethe C, Thakare V.
-\newblock Comparative {Analysis} of {Multipath} {Routing} {Techniques} and
-  {Design} of {Secure} {Energy} {Aware} {Routing} {Algorithm} for {Wireless}
-  {Sensor} {Network}.
-\newblock IJACR. 2013;3(3):374.
-
-\bibitem{kshemkalyani_distributed_2008}
-Kshemkalyani AD, Singhal M.
-\newblock Distributed computing: principles, algorithms, and systems.
-\newblock Cambridge University Press; 2008.
-
-\bibitem{lua_survey_2005}
-Lua EK, Crowcroft J, Pias M, Sharma R, Lim S.
-\newblock A survey and comparison of peer-to-peer overlay network schemes.
-\newblock IEEE Commun Surv Tut. 2005;7(2):72--93.
-
-\bibitem{korzun_structured_2013}
-Korzun D, Gurtov A.
-\newblock Structured peer-to-peer systems: fundamentals of hierarchical
-  organization, routing, scaling, and security.
-\newblock New York, NY: Springer; 2013.
-
-\bibitem{mohr_explaining_1982}
-Mohr LB.
-\newblock Explaining organizational behavior.
-\newblock Jossey-Bass; 1982.
-
-\bibitem{walker_mass_2012}
-Walker DC.
-\newblock Mass notification and crisis communications: {Planning},
-  preparedness, and systems.
-\newblock CRC Press; 2012.
-
-\bibitem{nickerson_thinking_2010}
-Nickerson JV, Tversky B, Corter JE, Yu L, Mason D.
-\newblock Thinking with networks.
-\newblock In: {CogSci}. vol.~36; 2010.
-
-\bibitem{ellison_ten_2000}
-Ellison C, Schneier B.
-\newblock Ten risks of {PKI}: {What} you're not being told about public key
-  infrastructure.
-\newblock Comput Secur J. 2000;16(1):1--7.
-
-\bibitem{levien_attack-resistant_2009}
-Levien R.
-\newblock Attack-resistant trust metrics.
-\newblock In: Computing with {Social} {Trust}. Springer; 2009. p. 121--132.
-
-\bibitem{reiter_resilient_1998}
-Reiter MK, Stubblebine SG.
-\newblock Resilient authentication using path independence.
-\newblock IEEE T Comput. 1998;47(12):1351--1362.
-
-\bibitem{lamport_byzantine_1982}
-Lamport L, Shostak R, Pease M.
-\newblock The {Byzantine} generals problem.
-\newblock TOPLAS. 1982;4(3):382--401.
-
-\bibitem{castro_practical_1999}
-Castro M, Liskov B, {others}.
-\newblock Practical {Byzantine} fault tolerance.
-\newblock In: {OSDI}. vol.~99; 1999. p. 173--186.
-
-\bibitem{dwork_pricing_1993}
-Dwork C, Naor M.
-\newblock Pricing via processing or combatting junk mail.
-\newblock In: Advances in {Cryptology}. Springer; 1993. p. 139--147.
-
-\bibitem{nakamoto_bitcoin:_2008}
-Nakamoto S.
-\newblock Bitcoin: {A} peer-to-peer electronic cash system.
-\newblock bitcoinorg. 2008; p.~28.
-
-\bibitem{mazieres_stellar_2015}
-Mazières D.
-\newblock Stellar {Consensus} {Protocol}: {A} {Federated} {Model} for
-  {Internet}-level {Consensus}; 2015.
-
-\bibitem{fiat_censorship_2002}
-Fiat A, Saia J.
-\newblock Censorship resistant peer-to-peer content addressable networks.
-\newblock In: {SIAM} {SODA}. ACM; 2002. p. 94--103.
-
-\bibitem{clarke_freenet:_2001}
-Clarke I, Sandberg O, Wiley B, Hong TW.
-\newblock Freenet: {A} distributed anonymous information storage and retrieval
-  system.
-\newblock In: Designing {Privacy} {Enhancing} {Technologies}. Springer; 2001.
-  p. 46--66.
-
-\bibitem{shamir_how_1979}
-Shamir A.
-\newblock How to share a secret.
-\newblock Communications of the ACM. 1979;22(11):612--613.
-
-\bibitem{blakley_safeguarding_1979}
-Blakley GR.
-\newblock Safeguarding cryptographic keys.
-\newblock P Natl Comp Conf. 1979;48:313--317.
-
-\bibitem{zhang_using_2002}
-Zhang H, Goel A, Govindan R.
-\newblock Using the small-world model to improve freenet performance.
-\newblock In: {INFOCOM}. vol.~3. IEEE; 2002. p. 1228--1237.
-
-\bibitem{kleinberg_small-world_2000}
-Kleinberg J.
-\newblock The small-world phenomenon: {An} algorithmic perspective.
-\newblock In: {STOC}. ACM; 2000. p. 163--170.
-
-\bibitem{alrajeh_secure_2013}
-Alrajeh NA, Alabed MS, Elwahiby MS.
-\newblock Secure ant-based routing protocol for wireless sensor network.
-\newblock Int J Distrib Sens N. 2013;2013.
-
-\bibitem{kohno_improvement_2012}
-Kohno E, Okazaki T, Takeuchi M, Ohta T, Kakuda Y, Aida M.
-\newblock Improvement of assurance including security for wireless sensor
-  networks using dispersed data transmission.
-\newblock J Comp Sys Sci. 2012;78(6):1703--1715.
-
-\bibitem{khalil_unmask:_2010}
-Khalil I, Bagchi S, Rotaru CN, Shroff NB.
-\newblock {UnMask}: {Utilizing} neighbor monitoring for attack mitigation in
-  multihop wireless sensor networks.
-\newblock Ad Hoc Networks. 2010;8(2):148--164.
-
-\bibitem{lou_h-spread:_2006}
-Lou W, Kwon Y.
-\newblock H-{SPREAD}: a hybrid multipath scheme for secure and reliable data
-  collection in wireless sensor networks.
-\newblock IEEE T Veh Technol. 2006;55(4):1320--1330.
-
-\bibitem{liu_secure_2012}
-Liu A, Zheng Z, Zhang C, Chen Z, Shen X.
-\newblock Secure and energy-efficient disjoint multipath routing for {WSNs}.
-\newblock IEEE T Veh Technol. 2012;61(7):3255--3265.
-
-\bibitem{leskovec_graphs_2005}
-Leskovec J, Kleinberg J, Faloutsos C.
-\newblock Graphs over time: densification laws, shrinking diameters and
-  possible explanations.
-\newblock In: Proceedings of the eleventh {ACM} {SIGKDD} international
-  conference on {Knowledge} discovery in data mining. ACM; 2005. p. 177--187.
-
-\bibitem{xu_internet_2011}
-Xu X, Mao ZM, Halderman JA.
-\newblock Internet censorship in {China}: {Where} does the filtering occur?
-\newblock In: International {Conference} on {Passive} and {Active} {Network}
-  {Measurement}. Springer; 2011. p. 133--142.
-
-\bibitem{dingledine_tor:_2004}
-Dingledine R, Mathewson N, Syverson P.
-\newblock Tor: {The} second-generation onion router.
-\newblock DTIC Document; 2004.
-
-\end{thebibliography}
+\section{Acknowledgments}
+
+Anonymized for blind review.
+%The authors would like to thank Tony Garnock-Jones, A. Frederick Dudley, and
+%Nathaniel Bezanson for helpful conversations.
+%This research was partly supported by the National Science Foundation under
+%Grant No. IIS-1617820.
+
+{\footnotesize \bibliographystyle{acm}
+\bibliography{references}}
+
+%\theendnotes
 

### 2021-09-14 AttackTolerance/netsci-2017/abstract.tex

--- 2021-09-14 AttackTolerance/decentralization-2019/abstract.tex
@@
 %Enter abstract here
-Centralized systems are susceptible to
-targeted attacks against central points,
-suggesting the importance of decentralization in attack-tolerant systems.
-While many techniques exist for tolerating random faults,
-better techniques for tolerating {\em adversarial faults} such as targeted attacks are needed.
-In communication networks, targeted attack can leave users vulnerable to
-censorship and surveillance of messages at central points.
-Even when encryption is used, it can be bypassed by coercion, e.g., subpoenas.
-While decentralized {\em protocols} have become a popular approach to attack-tolerance, centralized network {\em
-structure} can arise even when protocols are
+Networks with single points of failure are particularly susceptible to
+targeted attacks.
+In communication networks, these types of faults can leave users vulnerable to
+censorship and targeted surveillance,
+even when cryptography is utilized.
+Centralized networks have single points of failure by definition,
+leading to a growing popularity in decentralized architectures and protocols.
+However, centralized network structure can arise even when protocols are
 decentralized.
-Despite their decentralized protocols,
-the internet and World-Wide Web have
+Despite being based on decentralized protocols,
+the Internet and World-Wide Web have
 been shown both theoretically and historically to be highly susceptible to
 adversarial faults [1], in part due to emergent structural centralization.
-In this work, we present
-1. A fault tolerace scheme for networked communication having a failure probability that decreses exponentially with the number of available independent paths;
-2. A routing algorithm for constructing such indpendent paths in the butterfly network topology [2]; and
-3. A bounded-transitivity trust model that makes it possible to quantify the adversarial fault tolerance of a network.
-Our work is the first theoretical demonstration of a point-to-point communication network architecture able to tolerate coercion and other
-targeted attacks, without requiring infinitely transitive trust.
-We also evaluate the fault tolerance of an imperfectly implemented butterfly topology by partially rewiring a snapshot of the internet's router network.
-We find that rewiring only 10\% of the edges allows the network to withstand faults in the most central 2\% of nodes,
-increasing to 8\% of nodes for 90\% rewiring.
-Our results show that it is possible, in principle, to create highly attack-tolerant communication systems that can withstand targeted and coercive attacks, but only when it is possible to influence network structure.
+Existing network trust models, such as webs of trust [2], fail to adequately address
+network structure.
+We describe a novel, adversarial fault-tolerant,
+concurrent multipath routing algorithm for the decentralized
+butterfly network topology [3].
+We also develop a partial trust model that makes it possible to
+quantify the adversarial fault tolerance of a network,
+and which makes more realistic transitivity assumptions than webs of trust.
+When network topology can be dictated, these results can be used to create
+scalable, attack-tolerant infrastructures. More generally, our results provide
+a formalism for evaluating the effects of network structure on adversarial
+fault tolerance.
+
+\bigskip
+\noindent[1] R. Albert et al. Error and attack tolerance of complex networks. {\em Nature,} 406(6794), 2000.
+\\
+\noindent[2] P. R. Zimmermann. {\em The Official PGP User's Guide.} MIT Press, 1995.
+\\
+\noindent[3] A. D. Kshemkalyani and M. Singhal. {\em Distributed Computing: Principles, Algorithms, and Systems}. Cambridge University Press, 2008.
+\\
+
 \begin{figure}[!h]
@@
 \caption{
-Failure probability decreases rapidly with the number of
+In multipath fault tolerance,
+the failure probability decreases rapidly with the number of
 message copies, and increases slowly with the number of faults.
@@
 \end{figure}
-
-\noindent[1] R. Albert et al. Error and attack tolerance of complex networks. {\em Nature,} 406(6794), 2000.
-\\
-\noindent[2] A. D. Kshemkalyani and M. Singhal. {\em Distributed Computing}. Cambridge U. Press, 2008.
-
 % Place the abstract of your talk/poster here, 250 Words maximum.

### 2021-09-14 AttackTolerance/plosone/attack-old.tex

--- 2018-05-21 plosone/attack.tex
@@
 \cite{elmer-dewitt_first_1993}.
-But 18 years after Albert et. al \cite{albert_error_2000}
-showed that the topology of the internet's router network
-makes it vulnerable to targeted attacks
-(vs. random faults), the fundamental problem of attack-tolerant network topology
-remains unsolved.
-Attack-tolerant topologies are desirable not only for the router network,
-but for any physical or virtual network where a compromised node puts
-communication at risk.
-For example, the network of verified keys in the public key infrastructure
-underlying secure http \cite{ellison_ten_2000},
-or the network of DNS nameservers.
+But 17 years after Albert et. al \cite{albert_error_2000}
+showed the internet's router network to be vulnerable to targeted attacks
+(vs. random faults), the fundamental problem remains unsolved.
 The ongoing vulnerability of the internet is evidenced by a long history of 
@@
 \cite{dainotti_analysis_2011}.
-In this paper, we present the first theoretical network topology
-supporting attack-toloerant, point-to-point networked communication,
+In this paper, we present the first theoretical architecture
+for point-to-point networked communication,
+that tolerates not only random faults, but also targeted attacks,
 without relying on infinitely transitive trust
@@
 the information flowing through them.
-While cryptograpy can provide some protection against surveillance, it is
-vulnerable to {\em man-in-the-middle} attacks \cite{nayak_different_2010},
-and cannot overcome censorship when communication is blocked.
-In this paper, we instead consider a topological approach.
 The Internet's decentralized design was motivated
@@
 The action (known as a {\em black hole attack}) was intended to censor YouTube
-within Pakistan only, but resulted in a worldwide cascading failure
-when a router misconfiguration allowed the false routing information to
-propagate outside of Pakistan.
-This incident exemplifies the type of attack requiring a topological approach.
-First, the attack was {\em non-technological} (a government order),
-allowing the attacker to bypass any cryptographic or technology-based defenses.
-Second, the attack originated at a {\em single point of failure}
-(a misconfigured router).
-Third, the behavior of the compromised component (the router) cascaded
-through a {\em network} (the network of internet routers) because the correct
-behavior of other components depended on the correct behavior
-of the single point of failure.
-And while the action was not an intentional attack against the global internet,
+within Pakistan only, but resulted in a worldwide cascading failure.
+The action was initiated by government order in Pakistan,
+and spread beyond Pakistan when a router misconfiguration allowed the false
+routing information to propagate globally.
+While the government order and router misconfiguration initiated the outage,
+it was the structure of the Internet's router network that allowed a fault in a
+single router to propagate globally.
+And while the action was not an intentional attack against the global Internet,
 the ability of an attacker to succeed without even trying only highlights
@@
 
-As another example, in 2013,
-the Texas-based email provider Lavabit was ordered to disclose
+In 2013, the Texas-based email provider Lavabit was ordered to disclose
 their private SSL keys to the FBI \cite{poulsen_edward_2013}.
@@
 in order to protect their users from surveillance.
-Once again, the attack was non-technical.
-And again, the attack was on a single point of failure:
-Lavabit's web server and that server's TLS/SSL keys.
-In this case, the affected network was the
-internet's public key infrastructure.
-With the private keys, an attacker would be able to intercept and
-surveil traffic because the issuing certificate authority (and any
-users trusting that authority) would incorrectly trust that they were
-communicating with Lavabit.
+Once again, the attack was successful due to a highly centralized
+architecture:
+SSL keys under control of a single entity, in a single legal jurisdiction.
 While originally intended as surveillance,
 this action effectively became an act of censorship.
+It is also important to note that while Lavabit's cryptography worked as intended,
+the attack was still successful because the system was
+vulnerable to non-technical coercion.
 So we see that such vulnerabilities are not limited to any one system or protocol,
@@
 
-This paper addresses the need for a theoretial understanding of network and
-redundancy-based approaches to attack tolerance.
-Our primary result is theoretical: an algorithm for constructing highly redundant
-paths in a particular network topology.
-While we motivate the need for such network topologies using examples such as
-the internet router network and webs of trust,
-we do not propoose that our algorithm as a pracitcal solution for any of these
-examples.
-Rather, our result is a demonstration that such topological approaches to
-attack-tolerance are theoretically possible.
+Coercion-resistant, topological approaches to attack tolerance are needed to
+address the current vulnerability of communications infrastructure
+to censorship and surveillance.
+This paper presents several contributions towards advancing those goals.
+While our work is motivated by the network of internet routers,
+the results are topological,
+and can potentially be applied to many different types of networks.
 
@@
 by compromising a number of intermediate nodes.
-We also assume that edges represent {\em direct trust},
-i.e., the belief that an adversary is unlikely to compromise a neighbor,
+We also assume that the source and destination nodes trust their neighbors
+i.e., that the adversary is unable to compromise them,
 as in the web of trust approach
-\cite{zimmermann_official_1995,richters_trust_2011}.
-The web of trust approach requires infinite trust transitivity in that
-for one node to trust another,
-at least one path of direct trusted edges must exist between them.
-The assumption of infinite transitivity is unrealistic
-\cite{christianson_why_1997},
-so we instead make a stricter assumption: {\em bounded trust transitivity},
-that trust transitivity can only be extended over a finite number of edges.
-Even with this more restrictive assumption, we show how trust can be established
-between two nodes even when no path of directly trusted edges exists between them.
+\cite{zimmermann_official_1995,ferguson_practical_2003}.
+Unlike the web of trust approach, which assumes that trust is infinitely
+transitive,
+we assume {\em bounded trust transitivity}.
 
@@
 The well-known and widely-used {\em web of trust} approach
-\cite{zimmermann_official_1995,richters_trust_2011}
+\cite{zimmermann_official_1995,ferguson_practical_2003}
 is a decentralized alternative.
@@
 One common approach, the web of trust
-\cite{zimmermann_official_1995,richters_trust_2011},
+\cite{zimmermann_official_1995,ferguson_practical_2003},
 extends trust infinitely transitively:

### 2021-09-14 AttackTolerance/plosone/r1/NetFT-plosone-r1.tex

--- 2019-01-11 plosone/attack-figs.tex
@@
 \begin{figure}[!h]
-\centerline{\includegraphics{fig-perror}}
+%\centerline{\includegraphics{fig-perror}}
 \caption{
@@
 \begin{figure}
-\centerline{\includegraphics[width=4in,height=2.78in]{fig-partial-trust.pdf}}
+%\centerline{\includegraphics[width=4in,height=2.78in]{fig-partial-trust.pdf}}
 \caption{
@@
 \begin{center}
-\includegraphics[height=3in]{fig-bf-route}
+%\includegraphics[height=3in]{fig-bf-route}
 \end{center}
@@
 \begin{figure}
-\centerline{\includegraphics{fig-diameter}}
+%\centerline{\includegraphics{fig-diameter}}
 \caption{
@@
 \begin{figure*}[h!]
-\centerline{\includegraphics[width=5in,height=5in]{fig-mincut}}
+%\centerline{\includegraphics[width=5in,height=5in]{fig-mincut}}
 \caption{
@@
 \begin{figure}[h!]
-\begin{center}
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-                        ]
-\foreach \i in {0\ldots,{\ldots}0\ldots,{\ldots}0\ldots,{\ldots}0}% <-- content of nodes
-    \node[X] {\i};
-\draw[<->] ([yshift=1.5mm] A-1.north east) -- node[above=0.25mm] {$h$} ([yshift=1.5mm] A-1.north west);
-\draw[<->] ([yshift=1.5mm] A-2.north east) -- node[above=0.25mm] {$l_w - 2h$} ([yshift=1.5mm] A-2.north west);
-\draw[<->] ([yshift=1.5mm] A-3.north east) -- node[above=0.25mm] {$h$} ([yshift=1.5mm] A-3.north west);
-\draw[<->] ([yshift=1.5mm] A-4.north east) -- node[above=0.25mm] {$m - l_w$} ([yshift=1.5mm] A-4.north west);
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {\underline{Stage}} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick, fill=blue!20!white},
-                        ]
-\node[Y] {$s$};
-\foreach \i in {{\ldots}0\ldots,{\ldots}0\ldots,{\ldots}0}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {1.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick, fill=blue!20!white},
-                        ]
-\foreach \i in {$s$}% <-- content of nodes
-    \node[X] {\i};
-\node[Y] {{\ldots}1\ldots};
-\foreach \i in {{\ldots}0\ldots,{\ldots}0}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {2.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick, fill=blue!20!white},
-                        ]
-\foreach \i in {$s$,{\ldots}1\ldots}% <-- content of nodes
-    \node[X] {\i};
-\node[Y] {$\tilde{s}$};
-\foreach \i in {{\ldots}0}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {3.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {anchor=base, rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick, fill=blue!20!white},
-                        ]
-\foreach \i in {$s$,{\ldots}1\ldots,$\tilde{s}$}% <-- content of nodes
-    \node[X] {\i};
-\node[Y] {$z_{w,D}$};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {4.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {anchor=base, rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick, fill=blue!20!white},
-                        ]
-\node[Y] {$z_{w,A}$};
-\foreach \i in {{\ldots}1\ldots,$\tilde{s}$,$z_{w,D}$}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {5.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {anchor=base, rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick, fill=blue!20!white},
-                        ]
-\foreach \i in {$z_{w,A}$}% <-- content of nodes
-    \node[X] {\i};
-\node[Y] {$z_{w,B}$};
-\foreach \i in {$\tilde{s}$,$z_{w,D}$}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {6.} ( A-1.west);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}[
-node distance=0pt,
- start chain = A going right,
-    X/.style = {anchor=base, rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain},
-    Y/.style = {rectangle, draw,% styles of nodes in string (chain)
-                minimum width=10ex, minimum height=3ex,
-                outer sep=0pt, on chain, thick, fill=blue!20!white},
-                        ]
-\foreach \i in {$z_{w,A}$,$z_{w,B}$}% <-- content of nodes
-    \node[X] {\i};
-\node[Y] {$z_{w,C}$};
-\foreach \i in {$z_{w,D}$}% <-- content of nodes
-    \node[X] {\i};
-\draw ( A-1.west) -- node[left=5ex,minimum width=10ex] {7.} ( A-1.west);
-\end{tikzpicture}
-\end{center}
 \caption{
@@
 \begin{center}
-\includegraphics{fig-routing}
+%\includegraphics{fig-routing}
 \end{center}

### 2021-09-14 AttackTolerance/plosone/r4/Manuscript.tex

--- 2021-09-14 AttackTolerance/plosone/r2/Manuscript.tex
@@
 \usepackage{booktabs} % For formal tables
-\usepackage{epigraph}
 \usepackage{afterpage}
@@
 It is crucial for large-scale communication networks such as the internet
-to be resilient against attacks\cite{sterbenz_resilience_2010}
+to be resilient against attacks
 such as censorship and surveillance,
@@
 Self-organized networks such as the internet’s router network typically have
-heavy-tailed degree distributions\cite{barabasi_scale-free_2009},
-making them highly vulnerable to targeted attacks against central nodes
-\cite{albert_error_2000}.
+heavy-tailed degree distributions,
+making them highly vulnerable to targeted attacks against central nodes.
 While cryptographic solutions exist,
@@
 and remain vulnerable to
-man-in-the-middle attacks \cite{nayak_different_2010}
-and coercion \cite{grewal_caveat_2013}.
+man-in-the-middle attacks
+and coercion.
 Coercion-resistant, topological approaches to attack tolerance are needed to
@@
 \section*{Introduction}
-
-\epigraph{The Net interprets censorship as damage and routes around it.}{\textit{--John Gilmore} \cite{elmer-dewitt_first_1993}}
 
@@
 
+\bibitem{baran_distributed_1964}
+Baran P, {others}.
+\newblock On distributed communications.
+\newblock Volumes I-XI, RAND Corporation Research Documents, August. 1964; p.
+  637--648.
+
+\bibitem{elmer-dewitt_first_1993}
+Elmer-Dewitt P, Jackson D.
+\newblock First nation in cyberspace.
+\newblock Time. 1993;6:62--64.
+
+\bibitem{albert_error_2000}
+Albert R, Jeong H, Barabási AL.
+\newblock Error and attack tolerance of complex networks.
+\newblock Nature. 2000;406(6794):378--382.
+
+\bibitem{ellison_ten_2000}
+Ellison C, Schneier B.
+\newblock Ten risks of {PKI}: {What} you're not being told about public key
+  infrastructure.
+\newblock Comput Secur J. 2000;16(1):1--7.
+
+\bibitem{dainotti_analysis_2011}
+Dainotti A, Squarcella C, Aben E, Claffy KC, Chiesa M, Russo M, et~al.
+\newblock Analysis of country-wide internet outages caused by censorship.
+\newblock In: Proceedings of the 2011 {ACM} {SIGCOMM} conference on {Internet}
+  measurement conference. ACM; 2011. p. 1--18.
+
+\bibitem{christianson_why_1997}
+Christianson B, Harbison WS.
+\newblock Why isn't trust transitive?
+\newblock In: Security protocols. Springer; 1997. p. 171--176.
+
+\bibitem{zin_survey_2015}
+Zin SM, Anuar NB, Kiah MLMM, Ahmedy I.
+\newblock Survey of secure multipath routing protocols for {WSNs}.
+\newblock J Netw Comput Appl. 2015;55:123--153.
+
 \bibitem{sterbenz_resilience_2010}
@@
 \newblock Computer Networks. 2010;54(8):1245--1265.
-
-\bibitem{barabasi_scale-free_2009}
-Barabási AL, {others}.
-\newblock Scale-free networks: a decade and beyond.
-\newblock Science. 2009;325(5939):412.
-
-\bibitem{albert_error_2000}
-Albert R, Jeong H, Barabási AL.
-\newblock Error and attack tolerance of complex networks.
-\newblock Nature. 2000;406(6794):378--382.
 
@@
 
-\bibitem{grewal_caveat_2013}
-Grewal GS, Ryan MD, Bursuc S, Ryan PY.
-\newblock Caveat coercitor: {Coercion}-evidence in electronic voting.
-\newblock In: Security and {Privacy} ({SP}), 2013 {IEEE} {Symposium} on. IEEE;
-  2013. p. 367--381.
-
-\bibitem{elmer-dewitt_first_1993}
-Elmer-Dewitt P, Jackson D.
-\newblock First nation in cyberspace.
-\newblock Time. 1993;6:62--64.
-
-\bibitem{baran_distributed_1964}
-Baran P, {others}.
-\newblock On distributed communications.
-\newblock Volumes I-XI, RAND Corporation Research Documents, August. 1964; p.
-  637--648.
-
-\bibitem{dainotti_analysis_2011}
-Dainotti A, Squarcella C, Aben E, Claffy KC, Chiesa M, Russo M, et~al.
-\newblock Analysis of country-wide internet outages caused by censorship.
-\newblock In: Proceedings of the 2011 {ACM} {SIGCOMM} conference on {Internet}
-  measurement conference. ACM; 2011. p. 1--18.
-
-\bibitem{christianson_why_1997}
-Christianson B, Harbison WS.
-\newblock Why isn't trust transitive?
-\newblock In: Security protocols. Springer; 1997. p. 171--176.
-
-\bibitem{zin_survey_2015}
-Zin SM, Anuar NB, Kiah MLMM, Ahmedy I.
-\newblock Survey of secure multipath routing protocols for {WSNs}.
-\newblock J Netw Comput Appl. 2015;55:123--153.
+\bibitem{barabasi_scale-free_2009}
+Barabási AL, {others}.
+\newblock Scale-free networks: a decade and beyond.
+\newblock Science. 2009;325(5939):412.
 
@@
 
-\bibitem{ellison_ten_2000}
-Ellison C, Schneier B.
-\newblock Ten risks of {PKI}: {What} you're not being told about public key
-  infrastructure.
-\newblock Comput Secur J. 2000;16(1):1--7.
-
 \bibitem{levien_attack-resistant_2009}
@@
 
+\bibitem{grewal_caveat_2013}
+Grewal GS, Ryan MD, Bursuc S, Ryan PY.
+\newblock Caveat coercitor: {Coercion}-evidence in electronic voting.
+\newblock In: Security and {Privacy} ({SP}), 2013 {IEEE} {Symposium} on. IEEE;
+  2013. p. 367--381.
+
 \bibitem{leskovec_graphs_2005}

### 2021-09-15 ch-intro.tex
Create dissertation proposal main file.

The internet has enabled collaborations at a scale never before possible,
but the best practices for organizing such large collaborations are still not clear.
Wikipedia is a visible and successful example of such a collaboration which might offer
insight into what makes large-scale, decentralized collaborations successful.
We analyze the relationship between the structural properties of WikiProject coeditor networks
and the performance and efficiency of those projects.
We confirm the existence of an overall performance-efficiency trade-off,
while observing that some projects are higher than others in both performance
and efficiency,
suggesting the existence factors correlating positively with both.
Namely, we find an association between low-degree coeditor networks
and both high performance and high efficiency.
We also confirm results seen in previous numerical and small-scale lab studies:
higher performance with less skewed node distributions,
and higher performance with shorter path lengths.
We use agent-based models to explore possible mechanisms for
degree-dependent performance and efficiency.
We present a novel local-majority learning strategy designed to satisfy properties
of real-world collaborations.
The local-majority strategy as well as a localized conformity-based strategy
both show degree-dependent performance and efficiency,
but in opposite directions,
suggesting that these factors depend on both network structure and learning strategy.
Our results suggest 
possible benefits to decentralized collaborations made of smaller,
more tightly-knit teams,
and that these benefits may be modulated by the particular learning strategies
in use.

### 2021-09-15 ch-wp-prod-perf.tex
Create dissertation proposal main file.

\epigraph
{The problem with Wikipedia is that it only works in practice. In theory, it's a total disaster.}
{---Gareth Owen \cite{elsharbaty_editing_2016} }

The internet has enabled collaborations at a global scale.
Wikipedia, a free encyclopedia that invites anyone to edit articles,
is one of the most successful and visible examples of such a collaboration.
Organizing groups without top-down control is notoriously
difficult
\cite{freeman_tyranny_1972},
and yet Wikipedia, with millions of self-organized editors,
has produced a high-quality encyclopedia \cite{giles_internet_2005,keegan_evolution_2017}.
A better theoretical understanding of projects like Wikipedia is highly desirable as it could
help inform the design of new collaborative projects.
We focus on one aspect of a large-scale decentralized collaboration:
its network structure \cite{newman_structure_2003}.
How does Wikipedia's non-hierarchical structure relate to its success?

We look at WikiProjects on the English-language Wikipedia.
WikiProjects are collections of thematically related articles,
each with their own standards and norms.
When measuring the quality of collaborative projects,
there are at least two distinct measures to consider.
The first measure is short-term:
how effective a unit of work is at improving
the collaboration's output,
which we call {\em efficiency}.
The other measure is long-term:
the highest quality typically reached by an output,
which we call {\em performance}.
These two terms are often used interchangeably,
but we find it fruitful to distinguish between the two.
We find that Wikipedia exhibits an overall trade-off between performance and efficiency.
However, some WikiProjects surpass others in both efficiency and performance,
suggesting the existence of factors that correlate positively with both.

Our study focuses on the coeditor networks of each WikiProject:
which editors have edited at least one article in common?
These relationships represent the possible flow of information.
We focus specifically on mean degree, degree skewness, and path length.
High-degree editors have more collaborators,
which can increase diversity and access to information at the possible
expense of higher coordination costs
\cite{hong_groups_2004,golub_naive_2010}.
Highly skewed degree distributions can amplify the biases of high-degree
editors while reducing the need for explicit coordination
\cite{kearns_experiments_2012}.
Networks with shorter path lengths allow information to travel more quickly
at the possible expense of less local diversity
\cite{mason_propagation_2008,barkoczi_social_2016}.

In addition to our empirical study,
we use agent-based modeling to examine the consequences of specific
assumptions on networked collaboration.
We model individual behavior using a {\em social learning strategy} that
assumes agents 1. can only access a fraction of the model's state,
2. interact with others who share their concerns,
and 3. integrate their preferences into a single state.
Our model is the first we are aware of to incorporate these assumptions,
which are present across many real-world collaborations,
including Wikipedia.
\pagebreak

Our main findings are:
\begin{itemize}
\setlength\itemsep{0pt}
\item Despite an overall performance/efficiency trade-off,
WikiProjects with low-degree coeditor networks tend to have both higher performance and higher efficiency;
\item Short paths are associated with higher performance, consistent with a conformity-based learning strategy;
\item Structural inequality, as measured by degree skewness, is associated with lower performance;
\item Our agent-based model
shows that the efficiency and performance of collaborations can
depend on network degree,
and that the direction of that dependence varies with social learning strategy.
\end{itemize}

Our findings shed light on the importance of network structure for successful
collaboration.
These findings might be informative for future interventions that
recommend tasks based on how they will
influence network structure,
or for interventions that seek to encourage behaviors
complementary to existing network structure.

\section{Background and Related Work}
\label{sec:background}
The present paper investigates the relationship between social networks
and collaboration outcomes.
This connection has been explored by a number of theoretical,
numerical, and small-scale lab studies in the field of
{\em social learning}.
We contribute to this literature with a large-scale, empirical field study.
In much of the existing literature,
degree distribution correlates with outcome measures.
But aside from the naive Bayes case, it is unknown whether
the correlation is explained best by degree or by another structural property, such as characteristic path length.
In the empirical networks we study,
unlike artificial networks,
the structural properties vary independently,
making it easier to isolate individual
network properties that correlate with outcome variables.

{\bf Social Learning.} In {\em networked social learning},
agents are represented by nodes on a network
and can interact only with their neighbors.
Social learning tasks can be divided into cases where agents have {\em generated signals}
(independently noisy estimates of a true value)
and those where agents have {\em interpreted signals}
(solutions based on different selections of available data)
\cite{hong_interpreted_2009}.
The behavior of individual agents is described by their
{\em social learning strategy}.
For generated signals,
a naive Bayesian approach converges to the truth
when all agents have the same degree,
while the speed of convergence depends on the {\em spectral gap}
between the two largest eigenvalues of the network's interaction matrix
\cite{degroot_reaching_1974,golub_naive_2010}.
Complex social learning tasks can also be modeled as the problem
of maximizing an objective function with many local maxima,
referred to as a {\em rugged landscape}
\cite{lazer_network_2007,mason_propagation_2008,mason_collaborative_2012,grim_scientific_2013,barkoczi_social_2016}.
Numerical simulations have shown that efficient networks
(those with short paths between nodes)
can result in faster convergence at the cost of a less optimal solution,
due to less time for exploration
\cite{mason_propagation_2008,grim_scientific_2013}.
However, when conformity-based social learning strategies are used,
efficient networks can sometimes find more optimal solutions than
inefficient ones \cite{barkoczi_social_2016}.
Using an agent-based model, Hong and Page \cite{hong_groups_2004} found that
diverse groups can outperform groups composed of the best individual
problem-solvers.

{\bf Lab experiments.} Lab-based experiments on networked collaboration
suggest a complex interaction between network topology and other factors.
While groups of networked human subjects perform very well on
difficult graph-coloring tasks, the best performing network architectures
(e.g., fully-connected vs. small-world) vary
from task to task \cite{kearns_experiments_2012}.
The same studies found that while human subjects tend to perform well on
many networks, they perform worst on self-organized
networks, possibly due to higher structural inequality
(degree skewness).
Similarly, some network topologies are able to reach faster decisions in the
presence of more information, while others show the opposite effect
\cite{kearns_experimental_2006}.
Based on lab experiments, Fowler and Christakis \cite{fowler_cooperative_2010}
suggest that individual decisions towards altruism are conditional on their
neighbor's behavior and ``contagious'' up to three degrees away.
Later experiments by Suri and Watts \cite{suri_cooperation_2011} confirmed the
existence of conditional altruism,
but concluded that altruism
influences only first-degree neighbors.

{\bf Digital Communities.}
Research on digital communities has also examined the role of 
diversity and inequality
in collaborative work and decision-making.
In sociology, research has focused on the relationship between
network structure and social capital.
Powerful individuals are often ``brokers''
who act as exclusive intermediaries between disconnected portions of the
social network \cite{silverman_patronage_1965}.
Similarly, successful innovation in organizations often occurs in ``structural
holes'' between groups \cite{granovetter_strength_1973}.

For Wikipedia specifically,
Robert and Romero \cite{robert_crowd_2015} found that
larger group sizes yield higher article ratings
when the groups are diverse and experienced. Kittur and Kraut found that different types of coordination have a complex
effect on the quality of Wikipedia articles \cite{kittur_harnessing_2008}.
Both explicit and implicit coordination result in higher quality articles,
with explicit coordination being especially central in the early life of an
article.
Shaw and Hill \cite{shaw_laboratories_2014}
found that behavior in online wiki communities is consistent
with the ``iron law of oligarchy,''
which states that
earlier members of a group will, over time, gain disproportionate
decision-making power and act increasingly out of self-interest rather than
the good of the group \cite{michels_political_1999}.
Similarly, Halfaker et al. \cite{halfaker_rise_2013} attributed decreasing
participation on Wikipedia to poor retention of new users.
Looking specifically at Wikipedia policies determined by editor consensus,
Keegan and Fiesler \cite{keegan_evolution_2017} found a trend
from flexible rule-making towards less flexible maintenance and deliberation.
Using content analysis,
Morgan et al. \cite{morgan_project_2013} found WikiProjects to be
more loosely organized than traditional teams.

Across the broad range of work discussed above,
a few key themes emerge.
Both the efficiency and the performance of a collaboration are important
considerations and vary depending on both network structure and type of task
\cite{kearns_experiments_2012}.
While generated signal models of social learning predict no relationship
between the two
\cite{golub_naive_2010},
contagion-style innovation models predict a trade-off
\cite{mason_collaborative_2012,barkoczi_social_2016}.
Such a trade-off has been observed in simulations and lab experiments on
collaboration
\cite{kearns_experiments_2012,grim_scientific_2013}.

\section{WikiProjects}
\label{sec:wp}

Many articles on Wikipedia belong to one or more WikiProjects.
WikiProjects are groups of thematically-related articles
(e.g., articles related to Philosophy).
Information about an article's associated WikiProjects
can be viewed on that
article's talk page (Figure \ref{fig:knitting}).
Each WikiProject has its own page and talk page,
containing information about conventions within the project
as well as discussions about individual articles.
WikiProjects are thus distinct communities, with distinct norms and processes.
These communities are the fundamental units of analysis in this paper.

One of the main roles of a WikiProject is to evaluate the quality of its articles.
Quality assessments are made through consensus-based deliberation on the WikiProject
talk page.
Within a WikiProject,
assessments are typically made using the following {\em assessment classes}
(in order of increasing quality):
Stub, Start, C, B, A.
Different WikiProjects can assign different quality assessments to the same
article.
Differences between quality assessments could reflect different quality standards,
different grading systems, different responsiveness to changes in an article, etc.

In addition to the above assessment classes, articles on Wikipedia can be tagged as
``good article'' (GA) or ``featured article'' (FA) quality.
FA and GA determinations are made using a Wikipedia-wide consensus,
independently of WikiProject-based evaluations.
FA articles are ``the best articles Wikipedia has to offer''
\cite{wikipedia_contributors_wikipedia:featured_2018}.
GA articles meet ``a core set of editorial standards`` but are ``not featured article quality''
\cite{wikipedia_contributors_wikipedia:good_2017}.
When an article is assigned GA or FA status,
WikiProject quality assessments are often updated to reflect that status.
For example, the article {\em Mewtwo} was assessed as GA status on October 5,
2009 and shortly afterwards
its quality assessment was changed from B to GA within both
{\em WikiProject Pok\'emon} and
{\em WikiProject Video Games}.
This example also illustrates a quirk of conventions on Wikipedia:
very often, articles pass to GA or FA directly from B, skipping A.
The majority of WikiProjects rarely use the A class quality assessment.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=3in]{fig/WPProdPerf/fig-knitting.png}
\caption{
From Wikipedia {\em Knitting} talk page.
Two WikiProjects have assessed the article as B-class quality.
\label{fig:knitting}
}
\end{center}
\end{figure}

\subsection{Data}

Our analysis combines multiple data
sets from the English-language Wikipedia \cite{platt_english_2018}.
For information about edit history, we used a publicly-available data set containing
metadata (time, article id, user) about all edits from July 12, 2006 to December 2, 2015.
We used a custom script to scrape article quality assessments from logs produced
by WP 1.0 Bot for 2279 unique WikiProjects
between May 4, 2006 and December 2, 2015.
Finally, we used a publicly-available database dump of page events (including rename events)
to reconstruct the article id for each title mentioned in the assessment logs.

\subsection{Efficiency and Performance}

When individuals collaborate to solve a problem,
there are many ways to gauge their success.
One possibility is {\em efficiency}:
how quickly they find a solution.
Another is {\em performance}:
how good their solution is.
Evidence from numerical simulations
\cite{lazer_network_2007,mason_propagation_2008,mason_collaborative_2012,grim_scientific_2013,barkoczi_social_2016},
lab studies \cite{kearns_experiments_2012},
and field observations \cite{gentry_consensus_1982}
all suggest a trade-off between efficiency and performance.
While common, this trade-off is not absolute,
suggesting it is sometimes possible to simultaneously increase
performance and efficiency.
The identification of factors associated with both higher efficiency
and higher performance has obvious practical importance.
In this paper, we focus on how network structure
relates to efficiency and performance within WikiProjects.

For a WikiProject, efficiency quantifies how much participants can raise the
assessed quality of an article for a fixed amount of work.
We measure work by the number of revisions made.
Quality assessments are made through consensus of the project participants themselves.
Different projects can have different standards and practices for assessing article quality,
so the efficiency is not a measure of how quickly some objective measure of quality improves,
but rather of how quickly the project participants can reach consensus on the improvements that
need to be made and make those improvements.
Because our definition relies on assessment transitions,
we must define efficiency variables for
each of the project-level quality assessments: A, B, and C.
For a particular grade $G$,
we desire our definition of efficiency to meet the following conditions:
\begin{itemize}
\setlength\itemsep{0pt}
\item{Strictly increasing in the number of articles reaching grade $G$ (with revision count fixed);}
\item{Strictly decreasing in the number of revisions (with transition count fixed);}
\item{Independent of WikiProject size: not affected by adding an article having the same efficiency.}
\end{itemize}

We now define an efficiency measure which meets the above criteria.
Let $T(W,G)$ be the set of article assessment transitions from below grade $G$
to grade $G$ or higher in project $W$.
Let $N(W,G)$ be the number of articles in project $W$ which ever transition
from below grade $G$ to grade $G$ (or higher).
Given a transition $t$,
let $r(t)$ be the number of revisions to the article
since its previous grade transition,
and let $g(t)$ be the number of grade levels crossed bt $t$.
We quantify the efficiency $E(W,G)$ as the inverse of the mean number of revisions
per transition:
\beq
E(W,G)
&=&
\left[
\frac{1}{
N(W,G)
}
\sum_{t \in T(W,G)} \frac{r(t)}{g(t)}
\right]^{-1},
\eeq
where the $g(t)$ term accounts for assessments that raise article quality by
several grades by
dividing the revisions evenly between all grade levels achieved.

For performance, we wish to quantify how good articles tend to be when they reach a stable state.
Measuring performance is difficult for two reasons:
there is no objective measure of article quality available,
and articles are always changing, making it difficult to know which articles should be considered
complete or stable.
We use an extremely simple performance measure that gives surprisingly consistent results.
In addition to per-project quality assessment, articles can be given ``featured article'' or
``good article'' status.
The criteria for these statuses are consistent across all of Wikipedia,
and any editor can participate in the discussion and decision to award good or featured
status.
In other words, the good and featured statuses are less subjective than per-project assessments.

Our performance measure $P(W)$ is defined simply as
the percentage of articles in project $W$ which have reached
good or featured status:
\beq
P(W) &=& \frac{f(W) + g(W)}{n(W)},
\eeq
where $f(W)$ and $g(W)$ are the numbered of featured and good articles respectively,
and $n(W)$ is the total number of articles.

\subsection{Coeditor Networks}

We would like to determine how the social network structure of
Wikipedia---the pattern of who interacts with whom---relates to
efficiency and performance.
There are several types of interactions we could focus on,
including:
coediting, user talk messages, and talk page replies.
We choose to focus on coediting: when two editors have made changes to the
same article or talk page.
While editors can communicate directly through user talk messages,
the number of such messages is small compared to the number of edits to article
and talk pages.
We also could have considered direct replies between editors on article talk
pages, but these replies are typically seen (and intended to be seen)
by everyone reading the talk page,
and are part of larger conversations.
When an editor views a page,
they are potentially viewing content from and interactions
between all editors who came before them,
motivating our choice to focus on the social network structure of
coeditors.

The {\em coeditor network} of a WikiProject consists of nodes representing editors
and edges connecting pairs of editors who have edited the same article.
The edges are directed, with the direction representing
{\em plausible information flow};
an edge from Alice to Bob exists if Alice edited an article and then Bob edited the same article at
a later time. Note that edges can exist in both directions. 
We make the simplifying assumption of unit weight for all edges.
We focus on three structural properties:
degree, characteristic path length, and min-cut.
Degree and characteristic path length have been shown to correlate with
performance and efficiency in some social learning settings
\cite{golub_naive_2010,mason_propagation_2008,grim_scientific_2013},
while min-cut can be interpreted as a measure of decentralization,
common feature of peer-produced communities such as Wikipedia
\cite{benkler_wealth_2006}.

The degree distribution is the simplest network property we analyze.
The in-degree (out-degree) of a node is the number of edges to (from) that node.
Taking the average of either in-degree or out-degree gives the same value:
the {\em mean degree} of the network.
In our context, the mean degree represents, on average,
how many other editors each editor has collaborated with.
We also consider the {\em skewness} of the in-degree and out-degree distributions.
A large positive degree skewness for a WikiProject coeditor network
implies that a small number of editors have a very large number of collaborators,
while a small positive value implies that the editors having the most collaborators
don't have many more than a typical editor.

We also calculate the characteristic path length for each WikiProject coeditor network.
The {\em distance} from node $s$ to node $t$ is the distance of the shortest path
from $s$ to $t$.
The {\em characteristic path length} (or just {\em path length})
is the mean distance between all editor pairs,
excluding unconnected pairs.
To account for unconnected nodes, we also measure the {\em connected fraction}:
the fraction of ordered node pairs with a directed path from source to sink.
The path length represents how quickly information can move through the network.
Networks with longer paths require more interactions for information to propagate,
which has been shown to reduce efficiency in some settings
\cite{mason_propagation_2008,barkoczi_social_2016}.

Our final network measure quantifies the connectivity of a project's coeditor
network using min-cut size.
The minimum $st$-cut between nodes $s$ and $t$ is the set of edges that must beremoved
for no path exists from $s$ to $t$.
The minimum cut (min-cut) of a graph is the smallest minimum $st$-cut over all node pairs $st$. 
The size of the graph min-cut quantifies the connectivity of a graph,
but only incorporates information about edges lying on paths crossing the min-cut.
Instead, we use the mean size of all minimum $st$-cuts, which we refer to as the
{\em mean min-cut}.
This measure quantifies the number of redundant paths information can take through the network.
Networks with higher redundancy are more resilient to errors on one path \cite{albert_error_2000}
and allow innovations to propagate through complex contagion,
in which innovations are only adopted after multiple exposures \cite{centola_complex_2007}.

The mean path and min-cut are computationally intensive,
requiring distance and minimum $st$-cut calculations for all node pairs.
For larger projects, these calculations are impractical and we thus employed
sampling to determine mean path length and mean min-cut.
For mean path length, source nodes were sampled, and path length was calculated to all destination nodes
from each of these.
For min-cut, node pairs were sampled.
In both cases, stratification was used to ensure the same number of nodes were were sampled from each of
12 node degree quantiles.
We estimated the error due to sampling by determining true values for a medium-sized project,
and calculating error as a function of sample-size.
Sample sizes were chosen such that relative error was below 10\%.
Even with sampling, however, it was impractical to calculate these properties for the largest projects,
so we exclude the 183 largest projects from the analysis.

\begin{figure}[t!]
\centering
\includegraphics[width=2.5in,height=2.5in]{fig/WPProdPerf/fig-eff-perf-hist.pdf}
\caption{
Histograms of WikiProject efficiency and performance.
Both measures are highly right-skewed, but form unimodal distributions
with low skewness after log transformation.
\label{fig:eff-perf-hist}
}
\end{figure}

\subsection{Empirical Results}

We find that both efficiency and performance are highly right-skewed,
with a small number of projects having values much higher than the average.
After log-transforming the values, both the efficiency and the performance have
a unimodal distribution with low skew (see Figure \ref{fig:eff-perf-hist}).
Our findings confirm the trade-off between performance and efficiency
observed in many other settings (Figure \ref{fig:eff-perf}).
However, when looking at specific projects, some are higher in both performance
and efficiency,
suggesting the existence of factors which correlate positively with both.

We also find that mean min-cut is highly correlated with degree ($r=0.980$, $p<0.001$),
so we exclude min-cut from regression models to prevent collinearity.
The high correlation between mean degree and min-cut implies that,
in most cases,
the minimum $st$-cut is simply the set of edges from $s$
or the set of edges to $t$.
The rarity of non-trivial min-cuts suggests that WikiProject coeditor
networks have very few central
bottlenecks and are thus highly decentralized.

\begin{figure}[t!]
\centering
\includegraphics[width=2.5in,height=1.67in]{fig/WPProdPerf/fig-perf-eff.pdf}
\caption{
WikiProject performance is anticorrelated with B-level efficiency,
with Pearson r of -0.12.
Results are similar for other grade levels.
On average, highly efficient WikiProjects are under-performing,
but when looking at specific WikiProjects,
some are higher than others in both performance and efficiency.
\label{fig:eff-perf}
}
\end{figure}

%\begin{figure}
%\centering
%\includegraphics[width=2.33in,height=2.33in]{fig/WPProdPerf/fig-degree-mincut.pdf}
%\caption{
%Min-cut of WikiProject coeditor networks vs. their mean degree.
%\label{fig:degree-mincut}
%}
%\end{figure}

\begin{table}
\small
\begin{tabular}{lllllllll}
&Perf$^\dagger$&A-Eff$^\dagger$&B-Eff$^\dagger$&C-Eff$^\dagger$\\
\hline
Mean degree$^\dagger$&-0.7$^{***}$&-0.8$^{***}$&-0.6$^{***}$&-0.3$^{*}$\\
Out degree skew$^\dagger$&-0.4$^{***}$&-0.5$^{**}$&-0.3$^{*}$&-0.06\\
Mean path length$^\dagger$&-0.33$^{***}$&-0.09&-0.05&-0.09\\
C-Efficiency$^\dagger$&-0.08$^{*}$&\+---&\+---&\+---\\
Connected frac.&0.01&0.09$^{*}$&0.15$^{***}$&0.06\\
Talk fraction$^\dagger$&0&-0.02&-0.03&0.01\\
Mean similarity$^\dagger$&0.06$^{**}$&-0.03&0.01&0.02\\
Mean editors/art.$^\dagger$&0.3$^{**}$&0.3&0.2$^{*}$&0.09\\
Article count$^\dagger$&-0.4&0.7$^{*}$&0.8$^{**}$&0.7$^{**}$\\
Editor count$^\dagger$&0.4&0.9$^{**}$&0.8$^{**}$&0.5$^{*}$\\
Revision count$^\dagger$&0.6$^{*}$&-1$^{**}$&-1.1$^{***}$&-1$^{***}$\\
First assessment&0.05&0.11$^{**}$&0.31$^{***}$&0.43$^{***}$\\
Mean article age&-0.03&-0.04&-0.01&-0.05$^{*}$\\
\hline
N&1179&966&1260&1415\\
R$^2_{adj}$&0.37&0.17&0.30&0.43\\
\hline
\end{tabular}
\begin{tablenotes}
\item $\dagger$ Log-transformed. * $p < 0.05$. ** $p < 0.01$. *** $p < 0.001$.
\end{tablenotes}
\caption{Standardized coefficients for OLS models.
\label{tab:model}
}
\end{table}

To study the relationship between network structure, efficiency, and performance,
we model the performance and efficiency of WikiProjects using ordinary least-squares linear regression.
Each WikiProject is a single observation.
The models include each project's coeditor network properties as independent variables.
We also include the following project-level variables to control for confounding factors.
\begin{description}
\setlength\itemsep{0pt}
\item[C-efficiency]
(performance only).
Quantifies how quickly a WikiProject improves articles.
Efficiencies for different grades are highly correlated,
so we include only one.
\item[Connected fraction.]
Fraction of coeditor pairs connected by a path.
\item[Talk fraction.] Fraction of total revisions made to talk pages.
\item[Mean similarity.] Mean Jaccard similarity (by article) with other WikiProjects; a measure of topical complexity.
\item[Mean editors/article.] Mean number of editors collaborating on each article in a WikiProject.
\item[Article count.] Total number of articles in the WikiProject.
\item[Editor count.] Total number of editors working on articles within a WikiProject.
\item[Revision count.] Total number of revisions to articles in a WikiProject.
\item[First assessment.] Timestamp of first assessment; a measure of how long a WikiProject has been active.
\item[Mean article age.] Mean age of articles within a WikiProject.
\end{description}

Our models are summarized in Table \ref{tab:model}.
Min-cut is excluded from all models to avoid collinearity,
as it is highly correlated with degree.
In-degree and out-degree skewness were also highly correlated,
so we only include out-degree skewness
(results are similar for in-degree skewness).
Heavy-tailed variables are log-transformed.
To test the robustness of our results,
we also computed models using cube root instead of logarithmic transformations,
and using only top- and high-importance articles.
The results were qualitatively similar results for all variables,
except for degree-skewness, which had an inconsistent sign across models.

We see that B-efficiency and C-efficiency have very similar models, but that A-efficiency behaves
differently in its dependence on degree skewness and connectivity.
The different behavior of A-efficiency is likely explained by the
observation that the A-Class quality is infrequently used in practice.
The A-Class quality level is usually passed when an article reaches
good or featured article status,
which follow deifferent a consensus process from other ratings.

The negative dependence of performance on C-efficiency suggests there is generally a trade-off between
performance and efficiency.
However, low degree is correlated with both higher efficiency and higher performance,
suggesting that it is possible to improve both simultaneously.
Much of the existing numerical work on networked social learning focuses on path length rather than degree,
so we explore this result further using simulations in the next section.

For path length, we find that longer lengths correspond to lower performance, contrary to the conjecture
that longer path lengths allow more exploration \cite{mason_propagation_2008}
but consistent with a conformity-based social learning strategy \cite{barkoczi_social_2016}.

We also observe that high degree skewness is correlated with lower performance and lower A-efficiency,
suggesting that articles in projects with decentralized coeditor networks reach featured or good status
more efficiently, and reach higher quality ratings in general.

\section{Agent-Based Model}
\label{sec:sim}

\begin{table*}[!ht]
\small
\centering
\begin{tabular}{lccccc}
Name          & Social stage & Individual stage & Limited concern & Unknown objective & Single truth \\
\hline
Best+I & Best neighbor   & Global & & & \\
Conf+I & Conformity      & Global & & \Checkmark & \\
Best+LI & Best neighbor  & Local  & \Checkmark & & \\
Conf+LI & Conformity     & Local  & \Checkmark & \Checkmark & \\
LMaj+LI & Local majority & Local  & \Checkmark & \Checkmark & \Checkmark \\
\hline
\end{tabular}
\caption{
Definitions and properties of social learning strategies.
Each consists of a social stage and an individual stage.
Individual stages use hill-climbing based on either the global state,
or the agent's local concern.
\label{tab:strat}
}
\end{table*}

In addition to our empirical study,
we use a simple agent-based model of collaboration to better understand
the relationship between node degree, efficiency, and performance.
Numerical models allow us to determine the effect of changing a single
variable (e.g., network structure, learning strategy),
which is impractical in the empirical setting.
It is important to note that the goal of our model is not to simulate
all the intricacies of Wikipedia or any other specific platform.
Rather, our goal is to determine whether the correlations we observe
between degree and outcome variables
on Wikipedia can be reproduced in a more general setting.

Past work in the field of social learning typically models collaboration
as an optimization problem:
finding a state of the world which maximizes some objective function
\cite{lazer_network_2007,mason_propagation_2008,mason_collaborative_2012,barkoczi_social_2016}.
Wikipedia itself can be regarded as an optimization problem.
On Wikipedia, editors are generally seeking to improve the quality of articles
and have some personal preference over possible states of an article.
When editors do not agree on the optimal state of an article,
the conflict is resolved through a consensus-based deliberation.
This consensus process can be regarded as a {\em social choice function}
\cite{arrow_social_2012,brandt_computational_2012}
which maps individual preferences to community preferences.
Wikipedia can thus be thought of as a group of editors with individual preferences
for article states,
collaborating to optimize articles according to community preferences.
Note that these community preferences do not assume the existence of any
ground truth, other than the preferences themselves.

To simulate collaboration, we need a model problem for collaborators to solve.
Following existing literature on social learning,
we use the NK model \cite{kauffman_towards_1987}
to create NP-hard, nonlinear optimization problems.
The NK model produces an objective function with a
{\em rugged landscape}, i.e., many local optima.
The ruggedness of the model can be tuned through the parameters $N$
(the dimensionality of the solution space)
and $K$ (the level of inderdependence between dimensions).
Formally, the NK model produces an objective function $F$ mapping a binary string $S$ of length
$N$ to a real value in $[0,1]$.
Model state is divided into $N$ {\em loci}, with locus $i$ having a binary state $S_i$
and a value $f_i(S)$ dependent on its own state
and on the state of $K$ random other loci.
The functions $f_i(S)$ are created by selecting a random value in $[0,1]$ for
each possible state of locus $i$ and its $K$ neighbors.
The value of the model $F(S)$ is the mean of all locus values $f_i(S)$.
In our simulations,
agents iteratively search for a bit string $S$ that maximizes $F(S)$.

In a typical social learning model,
a set of agents each maintain an estimate of the optimal state
and iteratively update that estimate based on information available from other agents,
according to some {\em learning strategy}.
In networked social learning,
agents are associated with the nodes of a network and share information
only with their neighbors.
We define efficiency and performance in terms of the solution values for each time step
(averaged over many trials).
We define the performance to be the mean solution value after the process has converged,
while the efficiency is the reciprocal of the number of steps required to converge.
We measure the time to convergence as the number of steps required to reach 99\% of the maximum
mean solution value.

Without additional constraints,
the above model is missing
several key properties of real-world collaborations.
In designing our agent-based model, we paid attention to the
following properties.
\begin{description}
\setlength\itemsep{0pt}
\item[Limited concern.] Agents are concerned only with a subset of the entire
state when making decisions and determining preferences.
(On Wikipedia, editors typically interact with a small subset of the articles.)
\item[Concern-based network.] Agents interact with other agents who share a
common concern over some subset of the state.
(On Wikipedia, editors interact with others who share interests in the same articles.)
\item[Unknown objective.] Agents rank states in order of preference,
but do not have access to the objective function.
(On Wikipedia, there is no ground truth measure of quality.)
\item[Single source of truth.] At any given time, the system is in a single state
and agent preferences are based on local modifications to that state.
(At any point in time, there is only one current version of Wikipedia.)
\end{description}

\subsection{Concern-Based Networks}

On Wikipedia, editors interact by editing articles and talk pages.
Thus, the editors who interact with each other are exactly those who care about the same content.
Rather than using arbitrary networks,
we devise a network structure inspired by the above observation.
We do so by associating agents with particular loci in the NK model.
We also wish to study the effect of varying network degree,
which we achieve through a rewiring process described below.

Our concern-based networks are generated directly from the structure of the NK model.
The value of each NK locus depends on its own state and the state of $K$ other loci.
For each locus, we define an agent and assign these $K+1$ loci as its concern.
Next, an agent-agent co-affiliation network is created by connecting two agents if they share at least
one locus in their concerns.
This process is analogous to our construction of WikiProject coeditor networks.

To create a tunable degree, we duplicate each agent and its concern,
then randomly rewire a fraction of agent concerns before creating the agent-agent network.
With no rewiring, the duplication process creates a high overlap between agent concerns.
This overlap results in redundant links to a small number of agents,
rather than unique links to a large number of agents,
and therefore to an agent-agent network with small average degree.
By randomly rewiring the agent concerns, the redundancy is reduced
and the average degree of the agent-agent network is increased.

\subsection{Networked Learning Strategies}

Learning strategies determine how agents update their preferences based on
available information \cite{barkoczi_social_2016}.
Agents can engage in individual learning
by applying a hill-climbing algorithm to their current solution.
In each iteration, one bit of the NK solution string is flipped to maximize the solution value.
If no change improves the value, the original solution is kept.
The above strategy relies only on rankings of states,
satisfying the unknown objective assumption.
However, it relies on information about the entire state,
violating the limited concern assumption.
In order to satisfy this assumption,
we also define a local variant
in which only a subset of bits in the NK solution string are considered.
This variant reflects a more realistic style of collaboration,
in which individual agents focus on sub-problems.

In social learning,
agents can also incorporate information from
other agents they are connected to by an edge.
While individual learning always converges to the local maximum relative to the starting point,
social learning strategies allow agents to ``jump'' to drastically different solutions with higher local maxima.
In our model, we use both the conformity and best-neighbor strategies from \cite{barkoczi_social_2016}.
In the {\em best-neighbor} strategy, each agent compares its solution to a sample of neighbors, and chooses the solution with the highest value.
In order to compare solutions between neighbors,
the exact value of the objective function must be known for each solution,
so this strategy does not satisfy the unknown objective assumption
or the limited concern assumption.
In the {\em conformity} strategy, agents simply choose the most common solution among their neighbors
(ties are broken uniformly at random).
This strategy does not rely on solution value at all, so clearly satisfies the
unknown objective and limited concern assumptions.
In both cases, a single iteration of individual learning is performed after each social learning
iteration.
Because each agent maintains a separate estimate of the solution,
neither strategy satisfies the single source of truth assumption.

\subsection{Local Majority Strategy}

To satisfy the single source of truth assumption,
we introduce a new strategy: {\em local majority}.
In local majority, agents all begin with the same starting state and apply individual learning to their concern to generate
possible improvements to the solution.
Next, a new solution is constructed by considering each locus of the NK solution
individually.
Every agent concerned with a locus votes for its state based on their preferred
new solution and the majority state is chosen.
The result of this process is that all agents integrate their solutions into
a single state,
which forms the basis for the next iteration.
This strategy more realistically reflects collaborations like Wikipedia:
at any given time, a Wikipedia article has a single state, determined by consensus,
but editors may have differing opinions on how to improve that article.

\subsection{Simulation results}

We simulated 100 trials for rewiring values of 0.0, 0.167, 0.333, 0.5, 0.667, 0.833, and 1.0.
For each trial we generated an NK model with $N=250$ and $K=7$,
generated a concern-based network, and ran each social learning strategy
(Table \ref{tab:strat}) for 300 iterations.
For conformity and best-neighbor strategy, we used a sample size of 3, following
\cite{barkoczi_social_2016}.
We confirmed that all trials converged to their maximum value before reaching
the last iteration.
Networks had mean degree 116.6 with 1.3 standard deviation,
and mean path length of 1.766 with 0.0027 standard deviation.
The coefficient of variation for degree is approximately 10\%,
while only 1\% for mean path length, confirming that the rewiring process has a stronger influence
on degree than on path length.

\begin{figure}[t!]
\centering
\includegraphics[width=3in,height=1.5in]{fig/WPProdPerf/fig-val-iter.pdf}
\caption{
Mean agent solution value over time, averaged over 100 trials.
Strategies are defined in Table \ref{tab:strat}.
\label{fig:val-iter}
}
\end{figure}

\begin{table}
\small
\centering
\begin{tabular}{lll}
Strategy & Performance & Efficiency\\
\hline
Best+I  & 0.722 $\pm$ 0.001 & 0.0221 $\pm$ 0.0003 \\
Conf+I  & 0.721 $\pm$ 0.001 & 0.0174 $\pm$ 0.0002 \\
Best+LI & 0.726 $\pm$ 0.001 & 0.0131 $\pm$ 0.0002 \\
Conf+LI & 0.586 $\pm$ 0.001 & 0.030 $\pm$ 0.001 \\
LMaj+LI & 0.729 $\pm$ 0.001 & 0.046 $\pm$ 0.002 \\
\hline
\end{tabular}
\caption{
Simulated Performance and Efficiency.
Results shown for 100 trials with P[rewire] = 0.
Strategies are defined in Table \ref{tab:strat}.
Local strategies are less efficient than their non-local counterparts.
Local best-neighbor out-performs global,
while local conformity is the worst performer in all cases.
The local majority strategy is both most efficient and most performant.
\label{tab:sim-eff-perf}
}
\end{table}

\begin{table}
\small
\centering
\begin{tabular}{lll}
Strategy & Perf. Std. Coeff. & Eff. Std. Coeff.\\
\hline
Best+I  & -4.2$\times{10^{-5}}$ & \+4.1$\times{10^{-5}}$ \\
Conf+I  & \+2.7$\times{10^{-5}}$ & \+9.4$\times{10^{-5}}$ \\
Best+LI & -9.6$\times{10^{-4\;**}}$ & \+7.7$\times{10^{-5}}$ \\
Conf+LI & -1.5$\times{10^{-3\;***}}$ & \+8.7$\times{10^{-5\;*}}$ \\
LMaj+LI & \+1.2$\times{10^{-4\;**}}$ & -0.038$^{***}$ \\
\hline
\end{tabular}
\begin{tablenotes}
\item \centering * $p < 0.05$. ** $p < 0.01$. *** $p < 0.001$.
\end{tablenotes}
\caption{
Degree regression coefficients for simulations.
\label{tab:sim-perf}
}
\end{table}

\begin{figure*}
\includegraphics[width=6.67in,height=1.33in]{fig/WPProdPerf/fig-deg-eff.pdf}
\includegraphics[width=6.67in,height=1.33in]{fig/WPProdPerf/fig-deg-perf.pdf}
\caption{
Efficiency and Performance of social learning strategies vs. mean network degree.
Each point represents a single trial of 300 iterations.
Strategies are defined in Table \ref{tab:strat}.
The local best-neighbor strategy shows
decreased performance at high degree,
with no significant change in efficiency.
Local conformity shows decreased performance and increased efficiency at high
degree.
Local majority shows the opposite behavior:
increased performance and decreased efficiency at high degree,
with the efficiency showing the largest effect size of all strategies.
\label{fig:deg-eff-perf}
}
\end{figure*}

Figure \ref{fig:val-iter} shows how agents' solutions improve after repeated applications of
different learning strategies and rewiring values.
Each curve represents an average over 100 trials, each with 250 agents.
The mean performance and efficiency are reported in Table \ref{tab:sim-eff-perf}.
For all rewiring values, local strategies are less efficient and more performant
than their non-local counterparts.
For the best-neighbor strategy, local outperforms global.
Local conformity performs notably worse than all others.
Local majority is both more efficient and more performant than others,
with its performance increasing with higher rewiring.
This implies that,
at least in a simple collaboration model,
performance and efficiency can be simultaneously increased.
Furthermore, performance and efficiency are potentially affected by both the choice of
learning strategy and the average degree of the agents' social network.

The effects of degree on performance and efficiency are shown in Figure \ref{fig:deg-eff-perf} and
Table \ref{tab:sim-perf}.
For non-local versions of both conformity and best-neighbor strategy,
there is no significant effect of degree on performance or efficiency.
The local best-neighbor strategy shows reduced performance with increasing degree,
but no change in efficiency.
Local conformity and local majority show opposite behavior as degree increases:
with local conformity gaining efficiency at the expense of performance,
while local majority increases in performance and decreases in efficiency.
The largest effect size is achieved for efficiency in the local majority simulation,
which is consistent with the efficiency behavior observed in WikiProjects.
However, the performance behavior for local majority is opposite that observed on
Wikipedia.
These agent-based models confirm that network degree has the potential to influence
the performance and efficiency of collaborations.
Furthermore, this influence can be drastically different depending on the strategies
used by collaborators.

\section{Discussion}
\label{sec:discuss}

While existing research into the role of network structure in collaboration
has focused on numerical simulations and lab experiments,
analysis of large real-world systems is an important next step.
Our empirical analysis contributes several findings towards a better
understanding of large, decentralized, real-world collaboration.
We observe several results consistent with previous work:
a trade-off between performance and efficiency
\cite{mason_propagation_2008,grim_scientific_2013},
higher performance for shorter path lengths in a conformity setting
\cite{barkoczi_social_2016},
and a reduction in performance with increased structural inequality
\cite{kearns_experiments_2012}.
By using real-world networks, we were also able to analyze network properties independently.
While most existing work has focused on the importance of path length,
our findings suggest degree distribution may be just as, or more, important.
The association of low degree with both high performance and high efficiency is compelling,
as it sidesteps the usual trade-off between performance and efficiency.
In low-degree networks, agents have more repeated interactions with smaller groups of collaborators, suggesting that small team sizes could be beneficial for large collaborations.
Similarly, the observation that performance is higher in projects with less structural inequality suggests that,
if the challenges of egalitarian organizing are overcome,
decentralized collaborations may produce better outcomes than those with centralized, top-down structures.

Our agent-based models offer a several insights.
We observe degree-dependent performance and efficiency
for both local conformity and local majority strategies.
However, these two strategies have opposite degree dependence,
suggesting that different strategies may be preferable for high-degree and low-degree
networks.
Our local majority strategy,
designed to satisfy several properties found in real-world collaborations,
shows the strongest effects on performance and efficiency as network degree changes.
For the local majority strategy,
the relationship between degree and efficiency is consistent with our
empirical observations on Wikipedia,
suggesting one possible mechanism underlying that efficiency dependence.
However, the performance dependence of this strategy
is opposite that observed on Wikipedia,
suggesting that either the local majority strategy
is incompatible with actual behavior on Wikipedia or
that other factors outweigh the contribution of mean degree.

Our work has several limitations.
Our empirical analysis is purely correlative and cannot be used to draw
conclusions about the causal influence of network structure on collaboration.
However, the consistency of our results with other lab-based and numerical studies
suggests that the causal link is worthy of further study.
Similarly, our study focuses entirely on a single online community,
and while the results are suggestive, they do not necessarily generalize.
We have focused on structure, ignoring content-related variables.
For simplicity, we have assumed unweighted edges and
measured work by revision counts rather than bytes changed.

Our work suggests several directions for future work.
Is the correlation between network structure, performance, and efficiency causal?
A time-dependent analysis of our data could offer insight.
Are similar relationships observed in other large-scale collaborations?
Does varying degree independently of path length influence
performance and efficiency in a controlled lab setting?

A better understanding of the relationship between network structure and collaboration
outcomes has practical applications.
Online communities using recommender systems could make recommendations guided by desirable network properties.
Similarly, network structure could be used to identify under-performing
groups in need of an intervention.
The relationship between network structure and learning strategy suggests that
behaviors interact with network structure,
which could be used to encourage behaviors complementary to existing network
structure.

\section{Conclusion}
\label{sec:conclusion}

In this paper, we have described the relationship between the structural properties of WikiProject
coeditor networks, their performance, and their efficiency.
As in other studies, we see a trade-off between performance and efficiency.
However, 
some properties, such as low degree, are associated with
both higher performance and higher efficiency.
We also find that the correlations between path length and performance are consistent with a conformity-based
social learning strategy, but not a greedy best-neighbor strategy.
We observe improved performance in more decentralized projects,
as has been seen in small-scale lab experiments.
We have also proposed a novel local majority learning strategy that is more realistic,
more efficient, and higher performance than existing strategies.
While most previous social learning simulations focus on path length,
we observe degree-dependent performance and efficiency
in both the local majority strategy
and a localized version of the conformity strategy.
We find that the direction of that dependence varies with the
specific strategy being used.
While additional work is needed to determine causal relationships
and the generalizability of our results,
we have shown evidence that several phenomena predicted by numerical
and small-scale lab experiments are present in a large,
real-world collaboration.
Our results suggest that the success of large-scale collaborations may be aided
by greater decentralization,
consensus or conformity-based decision-making,
and more tightly-knit collaborations between smaller teams.

### 2021-09-15 proposal.tex
Create dissertation proposal main file.

\documentclass{report}

\usepackage{apacite}
\usepackage{bbding}
\usepackage{courier}
\usepackage{epigraph}
\usepackage{graphicx}
\usepackage{helvet}
\usepackage[normal]{threeparttable}
\usepackage{natbib}
\usepackage{times}
\usepackage{titlesec}
\usepackage{url}

\newcommand{\beq}{\begin{eqnarray}}
\newcommand{\eeq}{\end{eqnarray}}
\newcommand{\+}{\phantom{-}}

\author{Edward L. Platt}
\title{Network Deliberation: The role of network structure in large-scale, internet-enabled, participatory decision-making}

\begin{document}

\maketitle

\begin{abstract}
As groups grow in size, they gain access to additional resources,
creating opportunities for collective intelligence and collective action.
However, at very large scales, group decision-making becomes prohibitively slow and difficult
to coordinate.
Traditional solutions include representative decision-making and/or a shift from deliberation
to voting.
Both approaches sacrifice desirable properties.
Representative decision-making loses the potential benefits of collective intelligence and
introduces hierarchies that may place the interests of specific individuals ahead of the
interests of the group.
Voting sacrifices generativity: allowing a choice between predefined options,
without allowing for improvement to those options.
Arrow’s impossibility theorem also fundamentally limits the fairness of voting.
This project proposes Networked Deliberation as a potential means of large-scale decision-making.
In Networked Deliberation, members of a large group are repeatedly partitioned into small
deliberative pods.
Overlap between pods at different stages enables group-wide diffusion of information and
preferences.
Different methods for assigning members to pods result in different network topologies.
In partnership with large community groups, this project will use a web-based platform to
evaluate Network Deliberation.
For a variety of network topologies, the speed and quality of preference convergence will be
evaluated using quantitative information from ranked-choice polls throughout the process.
Deliberator satisfaction and experience will also be evaluated using qualitative analysis of
deliberative text and surveys.
This work seeks to improve the ability of very large groups to quickly develop a consensus,
enabling a more effective use of shared resources to achieve common goals.
\end{abstract}

\tableofcontents

\chapter{Introduction}
\include{ch-intro.tex}

\chapter{Network structure, productivity, and performance in WikiProjects}
\include{ch-wp-prod-perf.tex}

\chapter{Small interlocking groups improve mass deliberation in the presence
of strong social influence}

\chapter{Experimental evaluation of Network Deliberaiton}

\chapter*{Acknowledgements}
Thanks to Daniel M. Romero.

Scott E. Page,
Tawanna Dillahunt,
Ceren Budak,
for their time, guidance, and helpful feedback.

Jane Im,

Danielle Livneh,
Karthik Ramanathan
for help collecting WikiProjects data.

Yan Chen,
Tanya Rosenblat,

Grant Schoenbeck,
Kentaro Toyama,
Charles R. Severance

Veronica Falandino

Jeffrey W. Lockhart,
Danaja Maldeniya,
Lia Bozarth,
Christopher Quarles,
Chanda Phelan,
Padma Chirumamilla,
Earnest Wheeler,
Sam Carton,
Elliott Brannon,
Tawfiq Ammari,
2015 UMSI PhD chohort
Jean Hardy,
Cindy Lin,

the attendees of the May 25, 2017 MIT Center for Civic Media lab meeting
the Berkman-Klein Center Cooperation Working Group
the 2019 CSCW Workshop on Team and Group Diversity
for helpful feedback;

This research was partly supported by
the National Science Foundation under Grant No. IIS-1617820.

Andy Brosius,
Persephone Hernandez-Vogt,

\nocite{*}
\bibliographystyle{apacite}
\bibliography{prelim}

\end{document}

### 2021-09-19 NetDelibSim/main.tex
Move main text of ABM paper into proposal.

@@
 \maketitle
-\begin{abstract}
-Deliberation, a form of collective problem-solving, is a key component of democracy, social movements, and online peer-production.
-The network structure of interpersonal communication plays an important role in collective problem-solving, particularly in deliberation, which can become prohibitively complex and time-intensive at large scales.
-Network properties such as structural efficiency and degree distribution have been shown to influence the speed and quality of collective problem-solving, in combination with the behavioral strategies employed by individuals.
-However, many successful examples of mass deliberation differ in structure from common models of collective problem-solving by exhibiting networks of small interlocking groups.
-In this paper, we present an agent-based model of deliberation in networks of small interlocking groups and compare performance on these networks with conventional network structures.
-We find that networks of small interlocking groups improve solution quality when problem-solvers exhibit strong social influence.
-This effect is compatible with, but distinct from, the previously observed benefit of network efficiency in the presence of strong social influence.
-Our findings suggest a possible mechanism contributing to the success of existing mass deliberative projects as well as a principle for the design of new projects.
-By contributing to more effective deliberation at larger scales, we hope this work will contribute to democratizing the governance of large sociotechnical systems.
-
-\end{abstract}
-
-\section{Introduction}
-
-Deliberation enables the generation, aggregation, and synthesis of diverse knowledge \cite{dewey_creative_1940, anderson_epistemology_2006, ackerman_deliberation_2002, ostrom_collective_2000},
-as well as the early identification and resolution of conflict \cite{gentry_consensus_1982}.
-As a participatory process, deliberation builds trust among participants, increases the perception of fairness, and incentivises cooperation with group decisions
-\cite{ostrom_collective_2000, bowles_endogenous_1998}.
-As a dynamic process, deliberation alters private preferences through discourse \cite{habermas_structural_1991}, sidestepping fundamental limitations of voting, e.g., the Condorcet paradox \cite{condorcet_essay_1785, brandt_computational_2012} and Arrow's impossibility theorem \cite{arrow_social_2012, brandt_computational_2012}.
-Despite these benefits, deliberation becomes prohibitively difficult in larger groups, due to the increased time and effort needed to reach decisions \cite{fishkin_voice_1997, gentry_consensus_1982} and to the emergence of power inequalities \cite{freeman_tyranny_1972, boehm_egalitarian_1993}.
-Yet examples of successful mass deliberation do exist, including Wikipedia \cite{giles_internet_2005, keegan_evolution_2017}, free and open source software projects \cite{benkler_coases_2002}, grassroots protest and crisis-response networks \cite{gonzalez-bailon_networked_2016, brugh_combining_2019} and self-managed organizations \cite{laloux_reinventing_2014}.
-The commonalities between such projects provide insight into overcoming the challenges of mass deliberation.
-Specifically, we evaluate a model of mass deliberation inspired by common network features observed in successful projects.
-
-Communication in large collaborations is necessarily restricted when group size exceeds individual members' capacity for communication.
-The structure of the resulting communication network is an important factor in the success of collaborative tasks \cite{kearns_experiments_2012, mason_collaborative_2012, mason_propagation_2008}, such as deliberation.
-Among successful mass deliberative projects, communication networks often exhibits a common structure: small interlocking groups \cite{benkler_coases_2002, gonzalez-bailon_networked_2016, brugh_combining_2019, laloux_reinventing_2014}.
-A similar structure has been observed in the interlocking directorates of corporate boards \cite{karau_social_1993} and the interlocking publics of the public sphere \cite{habermas_structural_1991}, as well as utilized for collective design in network rotation \cite{salehi_hive_2018}.
-These interlocking groups have been given various names: committees, working groups, projects, modules, affinity groups, circles, teams, cores, nodes, zones, cells, or syndicates.
-We shall use the term {\em pods} to encompass all groups exhibiting two key characteristics: small enough to enable deliberation, and interlocking to enable information diffusion.
-We shall use the term {\em network deliberation} to refer to mass deliberation using an interlocking pod structure.
-Communication network structure also influences the success of collaborations by controlling the speed of information diffusion \cite{lazer_network_2007,mason_propagation_2008,mason_collaborative_2012,barkoczi_social_2016, gomez_clustering_2019,derex_partial_2016,kearns_experiments_2012}.
-Structurally efficient networks enable fast diffusion and favor exploitation of known information over exploration of the unknown \cite{lazer_network_2007, barkoczi_social_2016}.
-In the context of network deliberation, structural efficiency is determined by how individuals are assigned to pods.
-We will focus on two such methods: one efficient, one inefficient.
-
-Collective tasks such as deliberation are influenced not only by communication structure, but also by individual behavior.
-Individuals weigh social influence against their personal knowledge and expertise  \cite{boyd_culture_1988}.
-Even in a collective setting, individuals sometimes work independently to search for new and innovative ideas \cite{lazer_network_2007, barkoczi_social_2016}, to utilize their unique information and perspectives \cite{hong_interpreted_2009}, and to critically evaluate the ideas of others \cite{rendell_rogersparadox_2010}.
-Alternatively, individuals can defer to others in order to exploit and propagate good ideas \cite{barkoczi_social_2016, boyd_culture_1988}, to conserve their own resources through free-riding and social loafing \cite{karau_social_1993}, or due to a high level of trust \cite{salehi_hive_2018} or social influence \cite{banerjee_simple_1992, smith_pathological_2000}.
-Individual behavior interacts with network structure to vary the level of exploration of new ideas versus the exploitation of known information \cite{barkoczi_social_2016}.
-While the focus of network deliberation is structural, the effect of network structure must be understood in the context of individual behavior and social dynamics.
-
-We model network deliberation using a social learning framework to address two questions.
-How effective is network deliberation compared to conventional single-group deliberation?
-And how do network efficiency and individual behavior influence the outcome of network deliberation?
-We present the following three contributions.
-First, network deliberation out-performs conventional deliberation in both solution speed and quality when agents rely heavily on social learning.
-This suggests that network deliberation is a promising deliberation design for groups with a high degree of social influence.
-Second, network deliberation performs best on structurally efficient networks when agents exhibit conformist behavior, while it performs best on structurally inefficient networks when agents greedily maximize solution quality, consistent with findings for conventional networks \cite{barkoczi_social_2016}.
-Third, the naive greedy strategy can be modified to improve performance on conventional networks despite using strictly less information.
-In the case of network deliberation, the naive and modified greedy strategies are equivalent.
-
-Deliberation can be modeled as a form of collective problem-solving and social learning in which agents hold private beliefs, share information with their neighbors, and update their beliefs based on a combination of individual analysis and social influence.
-Following existing models of collective problem-solving \cite{lazer_network_2007, barkoczi_social_2016, gomez_clustering_2019}, our model comprises: 1.
-a collective task, 2.
-a network, and 3.
-a learning strategy.
-Agents maintain a candidate solution to the task and iteratively apply their learning strategy to the available information, yielding an updated candidate.
-
-As their collective task, agents seek an optimal point on a rugged fitness landscape.
-This landscape is generated using the tuneably-rugged NK model \cite{kauffman_towards_1987, weinberger_local_1991} (see Methods for more details).
-In this case, task solutions take the form of binary strings.
-
-To model network deliberation, we construct interlocking networks in stages. At each stage, agents are partitioned into small {\em pods} and connected to all other agents in the same pod.
-Our choice of pod size is informed by real-world The upper limit of effective small group size has been estimated variously as five \cite{freeman_tyranny_1972}, five--eight\cite{lohman_designing_2000}, and eight \cite{miflin_small_2004}.
-We choose a pod size of 5.
-Different methods for assigning agents to pods produce networks with different properties.
-Interlocking structure is created when agents are placed with others they have not yet interacted with.
-We thus require that pod assignment methods satisfy the following mixing property:
-\begin{property}
-\label{prop:mixing}
-For any agent $v$, and any two assignment stages $t$, $t^\prime$, the probability that at least one neighbor of $v$ at stage $t$ is also a neighbor at stage $t^\prime$ approaches 0 as $|V|$ grows.
-\end{property}
-We study two such methods: a random assignment method that produces structurally efficient networks, and a long-path method that produces inefficient networks.
-For comparison, we also simulate conventional deliberation.
-While the potential communication links in conventional deliberation form a fully connected network, actual communication is better modeled by preferential attachment \cite{barabasi_emergence_1999} or small-world \cite{watts_collective_1998} networks (see Methods).
-
-An agent's learning strategy models both informational constraints and behavioral dynamics \cite{lazer_network_2007, barkoczi_social_2016} such as social influence and social loafing \cite{karau_social_1993}.
-We consider three social learning strategies.
-The {\em best-neighbor} strategy is the naive greedy strategy: agents simply adopt the neighboring solution with the highest value.
-This strategy requires that agents have the ability to compare the quality of two arbitrary solutions, a strong assumption representing either skilled agents, or a simple task.
-The {\em conform} strategy is pure social influence. Agents adopt the most common solution among their neighbors.
-This strategy requires no knowledge of solution quality, modeling contexts where agents are unable to compare solution quality, or where agents choose not to, as in social loafing \cite{karau_social_1993}.
-We also introduce the {\em confident-neighbor} strategy, a variant of best-neighbor in which agents announce when they have the best solution in their neighborhood.
-This variant makes weaker assumptions about agent ability (see Methods for more detail) but is equivalent to best-neighbor when network deliberation is used.
-This strategy allows us to compare low-social influence and high-social influence strategies under equivalent assumptions of agent ability.
-
-Between each stage of social learning, agents may perform individual learning.
-Following existing literature \cite{lazer_network_2007, barkoczi_social_2016}, we model individual learning by mutating a single bit of an agent's solution, and keeping it if the mutation improves solution quality.
-A complete learning strategy must also specify how social and individual learning are integrated.
-Existing literature \cite{lazer_network_2007, barkoczi_social_2016} applies individual learning only when social learning fails to produce an improved solution.
-We refer to this method as {\em fallback} individual learning.
-Fallback can be also be interpreted as a model of social loafing \cite{karau_social_1993}.
-For completeness, we include analysis for a second method in which agents first perform social and individual learning in {\em parallel} and then choose the better of the two learning outputs (See Methods for more detail).
-
-We find that network deliberation identifies higher quality solutions than conventional deliberation when agents use the conform strategy, while requiring less time to converge.
-Within network deliberation, we find that the structurally efficient random pod network outperforms the structurally inefficient long-path network when agents use the conform strategy.
-However, when either best-neighbor or confident-neighbor is used, the inefficient network is preferable, consistent with findings for conventional networks \cite{barkoczi_social_2016}.
-We also find that the confident-neighbor strategy matches or outperforms best-neighbor across all networks, despite using strictly less information about solution quality.
-
-\begin{figure*}[]
-    \label{fig:results}
-    \centering
-    \begin{minipage}{3.33in}
-        \centering
-        \includegraphics[width=3.33in]{fig-perf-both.eps}
-    \end{minipage}%
-    \begin{minipage}{1.875in}
-    \centering
-    \includegraphics[width=1.875in]{fig-converge-both.eps}
-    \end{minipage}
-\caption{
-Solution quality and convergence time for agent-based simulations of different learning strategies. Error bars represent 95\% confidence interval. Results for network deliberation are shown shaded.
-}
-\end{figure*}
-
-\section{Results}
-Here we present the results of 1000 runs of an agent-based model of deliberation.
-In each run, an NK model is generated (N=15, K=6) and the model is run for 300 iterations for each network/strategy combination (see Methods for more details).
-The results of these simulations are shown in Figure \ref{fig:results}.
-
-\subsection{Network vs. Conventional Deliberation}
-We are particularly interested in comparing the performance of network deliberation to that of conventional single-group deliberation.
-When the conform strategy is used, we find that both of the network deliberation conditions yield better performance than conventional networks (Figure \ref{fig:results-conform-diff}).
-Analysis of the solution quality distributions (Figure \ref{fig:results-conform-dist}) shows two distinct effects.
-First, network deliberation frequently finds maximal solutions while conventional deliberation does not.
-Second, even when non-maximal solutions are found, network deliberation finds higher-quality solutions.
-This benefit is robust across both parallel and fallback individual learning.
-The effect size is dramatic.
-Conventional networks combined with the conform strategy show the lowest performance of all conditions, while random pod network deliberation combined with the conform strategy shows higher performance than nearly all other conditions.
-The exception being small world network combined with the best/confident-neighbor strategy, which shows equal performance within our statistical resolution.
-Often, increased performance comes at the expense of speed as more time is devoted to exploring novel solutions.
-However, we find that network deliberation improves both performance and speed when the conform strategy is used, suggesting that interlocking network structure can enable more efficient exploration.
-
-\begin{figure}
-    \label{fig:results-conform-diff}
-    \centering
-    \includegraphics[width=3.33in]{fig-results-netdelib-diff.eps}
-\caption{Difference in solution quality between network deliberation and conventional deliberation.
-Plots for Best/Confident-Neighbor show results for long-path and small-world networks, while the plot for conform shows results for random-pod and preferential-attachment networks.
-Results are shown for parallel individual learning.}
-\end{figure}
-
-\begin{figure}
-    \label{fig:results-conform-dist}
-    \centering
-    \includegraphics[width=3.33in]{fig-results-conform-dist.eps}
-\caption{Distributions of solution quality for social learning on a preferential-attachment network and random-pod network deliberation. Results are shown for parallel individual learning.}
-\end{figure}
-
-The conform strategy models contexts in which individuals rely heavily on social influence.
-Such contexts can arise from pro-conformity social norms, high levels of trust, social loafing, and limited information.
-As a form of rational ignorance, conformist behavior can conserve individual resources, but poses the risk of forming information cascades and propagating misinformation and inferior or harmful innovations \cite{banerjee_simple_1992}.
-Our simulations show that network deliberation can greatly improve the outcome of deliberation in the presence of strong social influence.
-This insight suggests one possible mechanism behind the success of mass deliberative projects, as well as the potential for network-based interventions in mass deliberation.
-
-\subsection{Structural Efficiency}
-
-The speed at which information can move through a communication network, i.e., it's structural efficiency, has been found to influence the success of collective problem-solving.
-We might expect the same to be true in network deliberation.
-In network deliberation, network structure is created through group membership.
-We compare the results of two group assignment methods: a random pod assignment yielding an efficient network, and a long-path method yielding a less efficient network (Figure \ref{fig:results-netdelib-efficiency}).
-We find that for greedy strategies (best-neighbor and confident-neighbor), the inefficient network yields the better results.
-The results are opposite for the conform strategy, with the efficient network yielding better results.
-These results are consistent with those observed for conventional single-group networks \cite{barkoczi_social_2016}, but to the best of our knowledge, they had not been explored for interlocking network like the ones we use to model network deliberation.
-
-\begin{figure}
-    \label{fig:results-netdelib-efficiency}
-    \centering
-    \includegraphics[width=3.33in]{fig-results-netdelib-efficiency.eps}
-\caption{Difference in solution quality between random-pod (efficient) and long-path (inefficient) network. Results are shown for parallel individual learning.}
-\end{figure}
-
-The role of structural efficiency in collective tasks can be partially attributed to diversity \cite{lazer_network_2007, hong_groups_2004}.
-Greedy strategies such as best/confident-neighbor result in agents adopting the highest quality solution they have seen so far and discarding all others.
-Such strategies quickly reduce the number of solutions present in the population, quickly identifying a local maximum at the cost of diversity.
-As a result, agents duplicate efforts during individual learning, exploring mutations of the same solution.
-Inefficient networks can mitigate this effect by slowing the spread of high-quality solutions and allowing time for the exploration of new and potentially better solutions.
-Our results show exactly this, with long-path pod assignment improving the performance of greedy algorithms.
-The improved performance of the conform strategy in random pod assignment can be attributed to the converse of the above.
-When social influence is strong, efficient networks can interrupt information cascades by introducing new information from far across the network.
-Our results are thus generally consistent with existing theory on structural efficiency and collective problem-solving.
-However, we note an interesting deviation: under the conform strategy, long-path network deliberation out-performs the two conventional networks, both of which are more efficient.
-
-\subsection{Confident-Neighbor Strategy}
-
-In general, the best-neighbor strategy requires strong assumptions about agents' ability to evaluate solutions.
-In the special case of network deliberation, the assumptions can be weakened by applying a variant of the strategy, which has not been previously studied.
-The variant, which we call {\em confident-neighbor}, is equivalent to best-neighbor for network deliberation, but not in general (see Methods for more detail).
-Confident-neighbor thus provides a comparison between network deliberation and conventional single-group deliberation under weaker informational assumptions than best-neighbor.
-Surprisingly, the weaker assumptions of confident-neighbor lead to higher performance on the preferential attachment network, and similar performance on the-small world network (within our statistical resolution).
-Figure \ref{fig:results-confident} shows histograms of the quality difference between confident-neighbor and best-neighbor for both conventional networks.
-In the majority of cases, both strategies find comparable solutions.
-However, we find that when these strategies do find different solutions, those found by confident-neighbor are more often better, at least for the preferential-attachment network.
-
-\begin{figure}
-    \label{fig:results-confident}
-    \centering
-    \includegraphics[width=3.33in]{fig-results-confident.eps}
-\caption{Histograms of the difference in solution quality between confident-neighbor and best-neighbor for both conventional networks.
-Results are shown for parallel individual learning.}
-\end{figure}
-
-As with conform, the confident-neighbor strategy can be used when available information is insufficient to use best-neighbor.
-Unlike conform, confident-neighbor models weak social influence.
-The improved performance of confident-neighbor over best-neighbor, despite using less information, can be attributed at least in part to diversity and exploration.
-Confident neighbor takes significantly longer to converge, suggesting it maintains a diversity of solutions and allows more time to explore new solutions.
-Our results suggest several interventions with the potential to improve conventional single-group deliberation.
-In the presence of weak social influence, reducing the amount of information available can counter-intuitively increase performance.
-Conversely, when information is already limited, performance might be increased by weakening social influence.
-However, while confident-neighbor can increase the success of conventional deliberation, it performs comparably to network deliberation using the conform strategy, suggesting that when social influence is strong, network deliberation remains preferable.
-
-\section{Discussion}
-
-We have proposed and evaluated a model of mass deliberation based on the observed interlocking network structure of successful mass-deliberative projects.
-Our central finding suggests that network deliberation can improve deliberative output in the presence of strong social influence (i.e., the conform strategy).
-Individuals rely on social influence for a number of reasons.
-Strong social influence can stem from social factors such as trust or social loafing.
-Alternatively, social influence can be used to supplement individual skill when it is insufficient for a given task.
-We see that a variety of contexts can lead to reliance on social influence.
-Empirical evidence suggests the presence of strong social influence in real-world collaborations such as Wikipedia \cite{platt_network_2018} and in lab studies of human collaboration \cite{mason_collaborative_2012, barkoczi_social_2016}.
-In the case of weak social influence (i.e., the  best-neighbor and confident-neighbor strategies), we do not observe a significant difference in output quality between network deliberation and conventional single-group deliberation.
-In all cases, network deliberation performs better than single-group deliberation or comparatively well.
-Our results suggest that network deliberation is both a contributing mechanism to the success of mass-deliberative projects as well as a useful tool for the design of large-scale sociotechnical systems and interventions for existing systems.
-
-How does network deliberation improve collaborative output?
-In part, the high performance of random-pod network deliberation can be attributed to the high structural efficiency of the resulting network.
-However, structural efficiency cannot be the complete mechanism.
-Long-path network deliberation is extremely inefficient, but still outperforms conventional networks in the presence of strong social influence.
-The additional mechanisms at play remain an open question.
-
-In order to distinguish between the effects of social influence and individual ability, we have introduced the confident-neighbor strategy, which (unlike best-neighbor) assumes the same agent capabilities as the conform strategy (in a critical learning setting).
-While not central to our study of network deliberation, our findings regarding the 
-confident-neighbor strategy have important implications for the study of social learning in general. 
-Our findings show that just as reducing network efficiency can improve solution quality by raising diversity, reducing the available information can improve the performance of greedy learning strategies.
-Our investigation of confident-neighbor stems from a careful formal analysis of the information required by each strategy (See Methods).
-The surprisingly high performance of confident-neighbor shows the importance of isolating social influence from agent ability.
-
-Our work is limited by a number of assumptions, which require further investigation to achieve greater generalizability.
-We assume a static objective function, while real-world objectives can change with changing external conditions as well as changing individual preferences.
-We also assume a homogeneous strategy across all agents, which raises questions of how different strategies interact within the same collaboration.
-We have also assumed that agents are truthful and do not form coalitions outside of the existing network, both assumptions which can be violated in real-world settings.
-We have also assumed that agents have access to the full solution string, and that agents incorporate all bits of the string into their quality estimates.
-Extremely complex real-world tasks might be better-modeled by a collection of sub-tasks, with agents having access to only a subset of the solution string.
-Finally, we have limited our study to learning strategies that explore new solutions solely through small mutations of existing solutions.
-Other potential strategies might generate entirely new solutions, for example, by combining parts of multiple previous solutions.
-Such generative strategies would add the ability to explore a wider range of solutions.
-While our work extends the understanding of social learning and collaboration, many questions still remain open for exploration.
-
-Effective mass deliberation is key to realizing the democratizing power of the internet.
-Our findings suggest that the presence of small and interlocking pods can contribute to the success of mass deliberation, particularly when the topic is complex and social influence is strong.
-These findings have implications for a range of deliberative contexts: participatory government and budgeting, worker-owned cooperatives, grassroots social movements, and the governance of decentralized systems (e.g., cryptocurrency).
-By enabling better deliberation at larger scales, we hope this work will contribute to democratizing the governance of large sociotechnical systems, and empowering the individuals impacted by those systems.
-
-\section{Methods}
-
-In deliberation, individuals update their beliefs and preferences through interaction with other participants, making it a form of social learning.
-We thus simulate deliberation using a model of social learning closely resembling those of other authors \cite{
-lazer_network_2007,
-mason_propagation_2008,
-barkoczi_social_2016,
-gomez_clustering_2019}.
-We model communication between agents using a time-dependent network $(V,E_t)$.
-The vertices $V$ represent agents.
-The edges $E_t$ allow agents to exchange information with their immediate neighbors at time $t$.
-Agents collaborate to complete a collective task.
-Specifically, agents seek to identify a bit string $s$ that maximizes an objective function $Q(s)$ generated using the NK model \cite{kauffman_towards_1987, weinberger_local_1991}.
-The NK model is a ``tuanbly rugged'' objective function parameterized by two variables: the number of bits in the input string (N), and the number of other bits each bit interacts with (K).
-These parameters can be used to tune problem size and complexity, respectively.
-The resulting objective function maps N-bit strings to real numbers in $[0,1]$.
-The resulting values can be exponentiated to modify the skewness of the solution value distribution.
-A full description of our model can be found in the Supplementary Information.
-In keeping with other authors \cite{lazer_network_2007, barkoczi_social_2016}, we use N=15, K=6, and an exponent of 8 throughout this paper.
-In keeping with convention, we normalize fitness values to the global maximum.
-In our simulations each agent maintains exactly one preferred solution (initially random).
-The simulation proceeds iteratively, with agents updating their preferred solution once per iteration based on observation of their neighbors' solutions.
-Throughout this paper, we simulate 300 iterations to ensure convergence.
-We present results for 1000 randomly generated NK model objectives, with each network/strategy combination simulated once per objective function.
-
-\subsection{Networks}
-
-We carry out simulations for four different network topologies.
-In addition to the well-known preferential-attachment \cite{barabasi_emergence_1999} and small-world \cite{watts_collective_1998} networks, we construct two types of networks to model network deliberation.
-Both types exhibit interlocking structure and satisfy Property \ref{prop:mixing} (proof in Supplementary Information).
-These two network deliberation conditions differ in how agents are assigned to pods and in the structural efficiency of the resulting network.
-At each iteration, existing edges are replaced with cliques corresponding to the pods. An edge is created between two agents if and only if they share a pod.
-For {\em random-pod} assignment, agents are assigned to pods at random.
-This assignment method produces short geodesic paths and efficient network structure (see Supplementary Information).
-For {\em long-path} assignment, agents are first assigned an integer label.
-At each stage, a new prime integer is chosen and agents are divided into classes based on the remainder of their label modulo the prime.
-Each class is then divided into size-M pods by selecting the lowest M labels, then the next-lowest M, and so on.
-The prime residue classes guarantee interlocking structure, while the grouping of similar labels prevents the creation of long-distance ``shortcut'' edges, producing an inefficient network (see Supplementary Information).
-All networks are constructed with 100 vertices/agents and a mean degree of 4 (pod size of 5).
-
-\subsection{Learning Strategies}
-We simulate three learning strategies.
-Each strategy combines an individual strategy and a social strategy.
-The individual strategy is the same for all conditions.
-Agents chose a random bit of their current solution and invert that bit.
-If the resulting solution is an improvement, it is kept, otherwise the original is kept.
-We simulate two methods of combining individual and social strategies. In the {\em fallback} condition, individual learning is only applied when social learning fails to improve the solution.
-In the {\em parallel} condition, individual and social strategies are applied to the original solution in parallel and the better of the two is chosen.
-In all cases, a final check is performed, following \cite{barkoczi_social_2016}: new solutions are only accepted if they improve on the original solution.
-Descriptions of the social learning strategies follow (formal definitions can be found in the Supplementary Information).
-
-The {\em best-neighbor} strategy closely follows that used by other authors \cite{lazer_network_2007, barkoczi_social_2016}.
-To apply this strategy, an agent searches the solutions of all immediate neighbors, and chooses the solution maximizing the payoff.
-In the event of a tie, one of the maximal solutions is chosen at random, with probability proportional to the number of neighbors exhibiting that solution.
-
-The {\em conform} strategy also follows previous authors \cite{barkoczi_social_2016}.
-To apply this strategy, agents search the solutions of all immediate neighbors, counting the number of times each unique solution appears.
-The solution with the highest count is adopted.
-In the event of a tie, the original solution is maintained.
-
-The {\em confident-neighbor} strategy is a novel variant of best-neighbor.
-To apply this strategy, agents search the solutions of all immediate neighbors to determine if any are superior to their own.
-If no superior solutions are found, the agent is {\em confident}.
-The confident agent maintains its original solution and broadcasts that solution to all neighbors.
-Alternatively, if superior solutions are found, the agent waits for broadcasts from confident neighbors.
-Agents may receive zero, one, or many such broadcasts.
-If no broadcasts are received, the agent maintains its original solution.
-If one or more broadcasts are received, one of them is chosen at random and that solution is adopted.
-It is worth noting that this strategy is equivalent to best-neighbor when the network is composed of one or more cliques, as in network deliberation.
-
-\subsection{Statistical Methods}
-All comparisons are made using two-tailed paired t-tests.
-Pairs of observations correspond to the same instance of the NK-model objective function.
-Significance values have been corrected for multiple comparisons using the Bonferroni correction.
-A full table of t-values and corrected p-values is available in the Supplemental Information.
-
-\section{Data Availability}
-Agent-based modeling code and data are available at
-[redacted for double-blind peer-review].
-%https://github.com/elplatt/Exp-Net-Delib.
-
-\section{Acknowledgements}
-[redacted for double-blind peer-review].
-% We thank Grant Schoenebeck for feedback on study design and early drafts.
-We thank Salem Elzway for feedback on early drafts.
-HB's work on this project was supported by the UMSI REMS program.
-This work was partly supported by the Air Force Office of Scientific Research under Award No.
-FA9550-19-1-0029.
-
-\section{Author Contributions}
-[redacted for double-blind peer-review].
-%ELP and DMR contributed to study conception, and data interpretation. ELP and HB contributed to development of simulation software and data collection. ELP drafted the manuscript.
-
-\section{Competing Interests}
-The authors declare no competing interests.
 

### 2021-09-19 acknowledgements.tex
Move main text of ABM paper into proposal.

--- 2021-09-15 proposal.tex
@@
-\documentclass{report}
-
-\usepackage{apacite}
-\usepackage{bbding}
-\usepackage{courier}
-\usepackage{epigraph}
-\usepackage{graphicx}
-\usepackage{helvet}
-\usepackage[normal]{threeparttable}
-\usepackage{natbib}
-\usepackage{times}
-\usepackage{titlesec}
-\usepackage{url}
-
-\newcommand{\beq}{\begin{eqnarray}}
-\newcommand{\eeq}{\end{eqnarray}}
-\newcommand{\+}{\phantom{-}}
-
-\author{Edward L. Platt}
-\title{Network Deliberation: The role of network structure in large-scale, internet-enabled, participatory decision-making}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-As groups grow in size, they gain access to additional resources,
-creating opportunities for collective intelligence and collective action.
-However, at very large scales, group decision-making becomes prohibitively slow and difficult
-to coordinate.
-Traditional solutions include representative decision-making and/or a shift from deliberation
-to voting.
-Both approaches sacrifice desirable properties.
-Representative decision-making loses the potential benefits of collective intelligence and
-introduces hierarchies that may place the interests of specific individuals ahead of the
-interests of the group.
-Voting sacrifices generativity: allowing a choice between predefined options,
-without allowing for improvement to those options.
-Arrow’s impossibility theorem also fundamentally limits the fairness of voting.
-This project proposes Networked Deliberation as a potential means of large-scale decision-making.
-In Networked Deliberation, members of a large group are repeatedly partitioned into small
-deliberative pods.
-Overlap between pods at different stages enables group-wide diffusion of information and
-preferences.
-Different methods for assigning members to pods result in different network topologies.
-In partnership with large community groups, this project will use a web-based platform to
-evaluate Network Deliberation.
-For a variety of network topologies, the speed and quality of preference convergence will be
-evaluated using quantitative information from ranked-choice polls throughout the process.
-Deliberator satisfaction and experience will also be evaluated using qualitative analysis of
-deliberative text and surveys.
-This work seeks to improve the ability of very large groups to quickly develop a consensus,
-enabling a more effective use of shared resources to achieve common goals.
-\end{abstract}
-
-\tableofcontents
-
-\chapter{Introduction}
-\include{ch-intro.tex}
-
-\chapter{Network structure, productivity, and performance in WikiProjects}
-\include{ch-wp-prod-perf.tex}
-
-\chapter{Small interlocking groups improve mass deliberation in the presence
-of strong social influence}
-
-\chapter{Experimental evaluation of Network Deliberaiton}
-
-\chapter*{Acknowledgements}
 Thanks to Daniel M. Romero.
@@
 
+Herminio Bodon,
+
 Danielle Livneh,
@@
 Tanya Rosenblat,
+Patrick Grim,
 
@@
 Kentaro Toyama,
-Charles R. Severance
+Charles R. Severance,
 
-Veronica Falandino
+Veronica Falandino,
+Linda Wood,
+Mita Gibson,
+Denise Bailey,
+Liz Ratzloff,
+Salem Elzway,
 
@@
 This research was partly supported by
-the National Science Foundation under Grant No. IIS-1617820.
+the National Science Foundation under Grant No. IIS-1617820,
+Air Force Office of Scientific Research under Award No. FA9550-19-1-0029,
+the UMSI REMS program,
 
@@
 
-\nocite{*}
-\bibliographystyle{apacite}
-\bibliography{prelim}
+Anthony Garnock-Jones,
+Elaine Lee,
 
-\end{document}
+i3 Detroit,
+
+Rahul Bhargava,
+Ethan Zuckerman,
+Sasha Costanza-Chock,
+Willow Idlewild,
+Charlie DeTar,
+
+Gerald J. Sussman

### 2021-09-19 ch-abm.tex
Move main text of ABM paper into proposal.

\section{Introduction}

Deliberation enables the generation, aggregation, and synthesis of diverse knowledge \cite{dewey_creative_1940, anderson_epistemology_2006, ackerman_deliberation_2002, ostrom_collective_2000},
as well as the early identification and resolution of conflict \cite{gentry_consensus_1982}.
As a participatory process, deliberation builds trust among participants, increases the perception of fairness, and incentivises cooperation with group decisions
\cite{ostrom_collective_2000, bowles_endogenous_1998}.
As a dynamic process, deliberation alters private preferences through discourse \cite{habermas_structural_1991}, sidestepping fundamental limitations of voting, e.g., the Condorcet paradox \cite{condorcet_essay_1785, brandt_computational_2012} and Arrow's impossibility theorem \cite{arrow_social_2012, brandt_computational_2012}.
Despite these benefits, deliberation becomes prohibitively difficult in larger groups, due to the increased time and effort needed to reach decisions \cite{fishkin_voice_1997, gentry_consensus_1982} and to the emergence of power inequalities \cite{freeman_tyranny_1972, boehm_egalitarian_1993}.
Yet examples of successful mass deliberation do exist, including Wikipedia \cite{giles_internet_2005, keegan_evolution_2017}, free and open source software projects \cite{benkler_coases_2002}, grassroots protest and crisis-response networks \cite{gonzalez-bailon_networked_2016, brugh_combining_2019} and self-managed organizations \cite{laloux_reinventing_2014}.
The commonalities between such projects provide insight into overcoming the challenges of mass deliberation.
Specifically, we evaluate a model of mass deliberation inspired by common network features observed in successful projects.

Communication in large collaborations is necessarily restricted when group size exceeds individual members' capacity for communication. The structure of the resulting communication network is an important factor in the success of collaborative tasks \cite{kearns_experiments_2012, mason_collaborative_2012, mason_propagation_2008}, such as deliberation. Among successful mass deliberative projects, communication networks often exhibits a common structure: small interlocking groups \cite{benkler_coases_2002, gonzalez-bailon_networked_2016, brugh_combining_2019, laloux_reinventing_2014}. A similar structure has been observed in the interlocking directorates of corporate boards \cite{karau_social_1993} and the interlocking publics of the public sphere \cite{habermas_structural_1991}, as well as utilized for collective design in network rotation \cite{salehi_hive_2018}.
These interlocking groups have been given various names: committees, working groups, projects, modules, affinity groups, circles, teams, cores, nodes, zones, cells, or syndicates.
We shall use the term {\em pods} to encompass all groups exhibiting two key characteristics: small enough to enable deliberation, and interlocking to enable information diffusion.
We shall use the term {\em network deliberation} to refer to mass deliberation using an interlocking pod structure.
Communication network structure also influences the success of collaborations by controlling the speed of information diffusion \cite{lazer_network_2007,mason_propagation_2008,mason_collaborative_2012,barkoczi_social_2016, gomez_clustering_2019,derex_partial_2016,kearns_experiments_2012}. Structurally efficient networks enable fast diffusion and favor exploitation of known information over exploration of the unknown \cite{lazer_network_2007, barkoczi_social_2016}. In the context of network deliberation, structural efficiency is determined by how individuals are assigned to pods.
We will focus on two such methods: one efficient, one inefficient.

Collective tasks such as deliberation are influenced not only by communication structure, but also by individual behavior. Individuals weigh social influence against their personal knowledge and expertise  \cite{boyd_culture_1988}. 
Even in a collective setting, individuals sometimes work independently to search for new and innovative ideas \cite{lazer_network_2007, barkoczi_social_2016}, to utilize their unique information and perspectives \cite{hong_interpreted_2009}, and to critically evaluate the ideas of others \cite{rendell_rogersparadox_2010}.
Alternatively, individuals can defer to others in order to exploit and propagate good ideas \cite{barkoczi_social_2016, boyd_culture_1988}, to conserve their own resources through free-riding and social loafing \cite{karau_social_1993}, or due to a high level of trust \cite{salehi_hive_2018} or social influence \cite{banerjee_simple_1992, smith_pathological_2000}. Individual behavior interacts with network structure to vary the level of exploration of new ideas versus the exploitation of known information \cite{barkoczi_social_2016}. While the focus of network deliberation is structural, the effect of network structure must be understood in the context of individual behavior and social dynamics.

We model network deliberation using a social learning framework to address two questions. How effective is network deliberation compared to conventional single-group deliberation? And how do network efficiency and individual behavior influence the outcome of network deliberation? We present the following three contributions.
First, network deliberation out-performs conventional deliberation in both solution speed and quality when agents rely heavily on social learning. This suggests that network deliberation is a promising deliberation design for groups with a high degree of social influence. Second, network deliberation performs best on structurally efficient networks when agents exhibit conformist behavior, while it performs best on structurally inefficient networks when agents greedily maximize solution quality, consistent with findings for conventional networks \cite{barkoczi_social_2016}.
Third, the naive greedy strategy can be modified to improve performance on conventional networks despite using strictly less information. In the case of network deliberation, the naive and modified greedy strategies are equivalent.

\section{Methods}

Deliberation can be modeled as a form of collective problem-solving and social learning in which agents hold private beliefs, share information with their neighbors, and update their beliefs based on a combination of individual analysis and social influence.
Following existing models of collective problem-solving \cite{lazer_network_2007, barkoczi_social_2016, gomez_clustering_2019}, our model comprises: 1. a collective task, 2. a network, and 3. a learning strategy.
Agents maintain a candidate solution to the task and iteratively apply their learning strategy to the available information, yielding an updated candidate.

As their collective task, agents seek an optimal point on a rugged fitness landscape.
This landscape is generated using the tuneably-rugged NK model \cite{kauffman_towards_1987, weinberger_local_1991} (see Methods for more details). In this case, task solutions take the form of binary strings.

To model network deliberation, we construct interlocking networks in stages. At each stage, agents are partitioned into small {\em pods} and connected to all other agents in the same pod.
Our choice of pod size is informed by real-world The upper limit of effective small group size has been estimated variously as five \cite{freeman_tyranny_1972}, five--eight\cite{lohman_designing_2000}, and eight \cite{miflin_small_2004}.
We choose a pod size of 5.
Different methods for assigning agents to pods produce networks with different properties.
Interlocking structure is created when agents are placed with others they have not yet interacted with.
We thus require that pod assignment methods satisfy the following mixing property:
\begin{property}
\label{prop:mixing}
For any agent $v$, and any two assignment stages $t$, $t^\prime$, the probability that at least one neighbor of $v$ at stage $t$ is also a neighbor at stage $t^\prime$ approaches 0 as $|V|$ grows.
\end{property}
We study two such methods: a random assignment method that produces structurally efficient networks, and a long-path method that produces inefficient networks. For comparison, we also simulate conventional deliberation. While the potential communication links in conventional deliberation form a fully connected network, actual communication is better modeled by preferential attachment \cite{barabasi_emergence_1999} or small-world \cite{watts_collective_1998} networks (see Methods).

An agent's learning strategy models both informational constraints and behavioral dynamics \cite{lazer_network_2007, barkoczi_social_2016} such as social influence and social loafing \cite{karau_social_1993}.
We consider three social learning strategies.
The {\em best-neighbor} strategy is the naive greedy strategy: agents simply adopt the neighboring solution with the highest value. This strategy requires that agents have the ability to compare the quality of two arbitrary solutions, a strong assumption representing either skilled agents, or a simple task.
The {\em conform} strategy is pure social influence. Agents adopt the most common solution among their neighbors.
This strategy requires no knowledge of solution quality, modeling contexts where agents are unable to compare solution quality, or where agents choose not to, as in social loafing \cite{karau_social_1993}.
We also introduce the {\em confident-neighbor} strategy, a variant of best-neighbor in which agents announce when they have the best solution in their neighborhood. This variant makes weaker assumptions about agent ability (see Methods for more detail) but is equivalent to best-neighbor when network deliberation is used.
This strategy allows us to compare low-social influence and high-social influence strategies under equivalent assumptions of agent ability.

Between each stage of social learning, agents may perform individual learning. Following existing literature \cite{lazer_network_2007, barkoczi_social_2016}, we model individual learning by mutating a single bit of an agent's solution, and keeping it if the mutation improves solution quality.
A complete learning strategy must also specify how social and individual learning are integrated. Existing literature \cite{lazer_network_2007, barkoczi_social_2016} applies individual learning only when social learning fails to produce an improved solution. We refer to this method as {\em fallback} individual learning. Fallback can be also be interpreted as a model of social loafing \cite{karau_social_1993}. For completeness, we include analysis for a second method in which agents first perform social and individual learning in {\em parallel} and then choose the better of the two learning outputs (See Methods for more detail).

In deliberation, individuals update their beliefs and preferences through interaction with other participants, making it a form of social learning. We thus simulate deliberation using a model of social learning closely resembling those of other authors \cite{
lazer_network_2007,
mason_propagation_2008,
barkoczi_social_2016,
gomez_clustering_2019}.
We model communication between agents using a time-dependent network $(V,E_t)$. The vertices $V$ represent agents. The edges $E_t$ allow agents to exchange information with their immediate neighbors at time $t$.
Agents collaborate to complete a collective task.
Specifically, agents seek to identify a bit string $s$ that maximizes an objective function $Q(s)$ generated using the NK model \cite{kauffman_towards_1987, weinberger_local_1991}.
The NK model is a ``tuanbly rugged'' objective function parameterized by two variables: the number of bits in the input string (N), and the number of other bits each bit interacts with (K).
These parameters can be used to tune problem size and complexity, respectively. The resulting objective function maps N-bit strings to real numbers in $[0,1]$. The resulting values can be exponentiated to modify the skewness of the solution value distribution. A full description of our model can be found in the Supplementary Information.
In keeping with other authors \cite{lazer_network_2007, barkoczi_social_2016}, we use N=15, K=6, and an exponent of 8 throughout this paper. In keeping with convention, we normalize fitness values to the global maximum.
In our simulations each agent maintains exactly one preferred solution (initially random). The simulation proceeds iteratively, with agents updating their preferred solution once per iteration based on observation of their neighbors' solutions. Throughout this paper, we simulate 300 iterations to ensure convergence. We present results for 1000 randomly generated NK model objectives, with each network/strategy combination simulated once per objective function.

\subsection{Networks}

We carry out simulations for four different network topologies. In addition to the well-known preferential-attachment \cite{barabasi_emergence_1999} and small-world \cite{watts_collective_1998} networks, we construct two types of networks to model network deliberation.
Both types exhibit interlocking structure and satisfy Property \ref{prop:mixing} (proof in Supplementary Information).
These two network deliberation conditions differ in how agents are assigned to pods and in the structural efficiency of the resulting network.
At each iteration, existing edges are replaced with cliques corresponding to the pods. An edge is created between two agents if and only if they share a pod.
For {\em random-pod} assignment, agents are assigned to pods at random.
This assignment method produces short geodesic paths and efficient network structure (see Supplementary Information).
For {\em long-path} assignment, agents are first assigned an integer label. At each stage, a new prime integer is chosen and agents are divided into classes based on the remainder of their label modulo the prime.
Each class is then divided into size-M pods by selecting the lowest M labels, then the next-lowest M, and so on.
The prime residue classes guarantee interlocking structure, while the grouping of similar labels prevents the creation of long-distance ``shortcut'' edges, producing an inefficient network (see Supplementary Information).
All networks are constructed with 100 vertices/agents and a mean degree of 4 (pod size of 5).

\subsection{Learning Strategies}
We simulate three learning strategies. Each strategy combines an individual strategy and a social strategy. The individual strategy is the same for all conditions. Agents chose a random bit of their current solution and invert that bit. If the resulting solution is an improvement, it is kept, otherwise the original is kept.
We simulate two methods of combining individual and social strategies. In the {\em fallback} condition, individual learning is only applied when social learning fails to improve the solution.
In the {\em parallel} condition, individual and social strategies are applied to the original solution in parallel and the better of the two is chosen. In all cases, a final check is performed, following \cite{barkoczi_social_2016}: new solutions are only accepted if they improve on the original solution. Descriptions of the social learning strategies follow (formal definitions can be found in the Supplementary Information).

The {\em best-neighbor} strategy closely follows that used by other authors \cite{lazer_network_2007, barkoczi_social_2016}. To apply this strategy, an agent searches the solutions of all immediate neighbors, and chooses the solution maximizing the payoff. In the event of a tie, one of the maximal solutions is chosen at random, with probability proportional to the number of neighbors exhibiting that solution.

The {\em conform} strategy also follows previous authors \cite{barkoczi_social_2016}. To apply this strategy, agents search the solutions of all immediate neighbors, counting the number of times each unique solution appears. The solution with the highest count is adopted. In the event of a tie, the original solution is maintained.

The {\em confident-neighbor} strategy is a novel variant of best-neighbor. To apply this strategy, agents search the solutions of all immediate neighbors to determine if any are superior to their own. If no superior solutions are found, the agent is {\em confident}. The confident agent maintains its original solution and broadcasts that solution to all neighbors. Alternatively, if superior solutions are found, the agent waits for broadcasts from confident neighbors. Agents may receive zero, one, or many such broadcasts. If no broadcasts are received, the agent maintains its original solution. If one or more broadcasts are received, one of them is chosen at random and that solution is adopted. It is worth noting that this strategy is equivalent to best-neighbor when the network is composed of one or more cliques, as in network deliberation.

\subsection{Statistical Methods}
All comparisons are made using two-tailed paired t-tests. Pairs of observations correspond to the same instance of the NK-model objective function. Significance values have been corrected for multiple comparisons using the Bonferroni correction. A full table of t-values and corrected p-values is available in the Supplemental Information.

\section{Results}

Here we present the results of 1000 runs of an agent-based model of deliberation.
In each run, an NK model is generated (N=15, K=6) and the model is run for 300
iterations for each network/strategy combination (see Methods for more details).
The results of these simulations are shown in Figure \ref{fig:results}.

We find that network deliberation identifies higher quality solutions than
conventional deliberation when agents use the conform strategy,
while requiring less time to converge. Within network deliberation,
we find that the structurally efficient random pod network outperforms the
structurally inefficient long-path network when agents use the conform strategy.
However, when either best-neighbor or confident-neighbor is used,
the inefficient network is preferable, consistent with findings for conventional
networks \cite{barkoczi_social_2016}. We also find that the confident-neighbor
strategy matches or outperforms best-neighbor across all networks,
despite using strictly less information about solution quality.

\begin{figure*}[]
    \label{fig:results}
    \centering
    \begin{minipage}{3.33in}
        \centering
        \includegraphics[width=3.33in]{fig/NetDelibABM/fig-perf-both.eps}
    \end{minipage}%
    \begin{minipage}{1.875in}
    \centering
    \includegraphics[width=1.875in]{fig/NetDelibABM/fig-converge-both.eps}
    \end{minipage}
\caption{
Solution quality and convergence time for agent-based simulations of different learning strategies. Error bars represent 95\% confidence interval. Results for network deliberation are shown shaded.
}
\end{figure*}

\subsection{Network vs. Conventional Deliberation}
We are particularly interested in comparing the performance of network deliberation to that of conventional single-group deliberation. When the conform strategy is used, we find that both of the network deliberation conditions yield better performance than conventional networks (Figure \ref{fig:results-conform-diff}). Analysis of the solution quality distributions (Figure \ref{fig:results-conform-dist}) shows two distinct effects. First, network deliberation frequently finds maximal solutions while conventional deliberation does not. Second, even when non-maximal solutions are found, network deliberation finds higher-quality solutions. This benefit is robust across both parallel and fallback individual learning. The effect size is dramatic. Conventional networks combined with the conform strategy show the lowest performance of all conditions, while random pod network deliberation combined with the conform strategy shows higher performance than nearly all other conditions. The exception being small world network combined with the best/confident-neighbor strategy, which shows equal performance within our statistical resolution. 
Often, increased performance comes at the expense of speed as more time is devoted to exploring novel solutions. However, we find that network deliberation improves both performance and speed when the conform strategy is used, suggesting that interlocking network structure can enable more efficient exploration.

\begin{figure}
    \label{fig:results-conform-diff}
    \centering
    \includegraphics[width=3.33in]{fig/NetDelibABM/fig-results-netdelib-diff.eps}
\caption{Difference in solution quality between network deliberation and conventional deliberation. Plots for Best/Confident-Neighbor show results for long-path and small-world networks, while the plot for conform shows results for random-pod and preferential-attachment networks. Results are shown for parallel individual learning.}
\end{figure}

\begin{figure}
    \label{fig:results-conform-dist}
    \centering
    \includegraphics[width=3.33in]{fig/NetDelibABM/fig-results-conform-dist.eps}
\caption{Distributions of solution quality for social learning on a preferential-attachment network and random-pod network deliberation. Results are shown for parallel individual learning.}
\end{figure}

The conform strategy models contexts in which individuals rely heavily on social influence. Such contexts can arise from pro-conformity social norms, high levels of trust, social loafing, and limited information. As a form of rational ignorance, conformist behavior can conserve individual resources, but poses the risk of forming information cascades and propagating misinformation and inferior or harmful innovations \cite{banerjee_simple_1992}. Our simulations show that network deliberation can greatly improve the outcome of deliberation in the presence of strong social influence. This insight suggests one possible mechanism behind the success of mass deliberative projects, as well as the potential for network-based interventions in mass deliberation.

\subsection{Structural Efficiency}

The speed at which information can move through a communication network, i.e., it's structural efficiency, has been found to influence the success of collective problem-solving. We might expect the same to be true in network deliberation. In network deliberation, network structure is created through group membership. We compare the results of two group assignment methods: a random pod assignment yielding an efficient network, and a long-path method yielding a less efficient network (Figure \ref{fig:results-netdelib-efficiency}). We find that for greedy strategies (best-neighbor and confident-neighbor), the inefficient network yields the better results. The results are opposite for the conform strategy, with the efficient network yielding better results. These results are consistent with those observed for conventional single-group networks \cite{barkoczi_social_2016}, but to the best of our knowledge, they had not been explored for interlocking network like the ones we use to model network deliberation.

\begin{figure}
    \label{fig:results-netdelib-efficiency}
    \centering
    \includegraphics[width=3.33in]{fig/NetDelibABM/fig-results-netdelib-efficiency.eps}
\caption{Difference in solution quality between random-pod (efficient) and long-path (inefficient) network. Results are shown for parallel individual learning.}
\end{figure}

The role of structural efficiency in collective tasks can be partially attributed to diversity \cite{lazer_network_2007, hong_groups_2004}. Greedy strategies such as best/confident-neighbor result in agents adopting the highest quality solution they have seen so far and discarding all others. Such strategies quickly reduce the number of solutions present in the population, quickly identifying a local maximum at the cost of diversity. As a result, agents duplicate efforts during individual learning, exploring mutations of the same solution. Inefficient networks can mitigate this effect by slowing the spread of high-quality solutions and allowing time for the exploration of new and potentially better solutions. Our results show exactly this, with long-path pod assignment improving the performance of greedy algorithms. The improved performance of the conform strategy in random pod assignment can be attributed to the converse of the above. When social influence is strong, efficient networks can interrupt information cascades by introducing new information from far across the network.
Our results are thus generally consistent with existing theory on structural efficiency and collective problem-solving.
However, we note an interesting deviation: under the conform strategy, long-path network deliberation out-performs the two conventional networks, both of which are more efficient.

\subsection{Confident-Neighbor Strategy}

In general, the best-neighbor strategy requires strong assumptions about agents' ability to evaluate solutions. In the special case of network deliberation, the assumptions can be weakened by applying a variant of the strategy, which has not been previously studied. The variant, which we call {\em confident-neighbor}, is equivalent to best-neighbor for network deliberation, but not in general (see Methods for more detail). Confident-neighbor thus provides a comparison between network deliberation and conventional single-group deliberation under weaker informational assumptions than best-neighbor. Surprisingly, the weaker assumptions of confident-neighbor lead to higher performance on the preferential attachment network, and similar performance on the-small world network (within our statistical resolution). Figure \ref{fig:results-confident} shows histograms of the quality difference between confident-neighbor and best-neighbor for both conventional networks. In the majority of cases, both strategies find comparable solutions. However, we find that when these strategies do find different solutions, those found by confident-neighbor are more often better, at least for the preferential-attachment network.

\begin{figure}
    \label{fig:results-confident}
    \centering
    \includegraphics[width=3.33in]{fig/NetDelibABM/fig-results-confident.eps}
\caption{Histograms of the difference in solution quality between confident-neighbor and best-neighbor for both conventional networks.
Results are shown for parallel individual learning.}
\end{figure}

As with conform, the confident-neighbor strategy can be used when available information is insufficient to use best-neighbor. Unlike conform, confident-neighbor models weak social influence. The improved performance of confident-neighbor over best-neighbor, despite using less information, can be attributed at least in part to diversity and exploration. Confident neighbor takes significantly longer to converge, suggesting it maintains a diversity of solutions and allows more time to explore new solutions. Our results suggest several interventions with the potential to improve conventional single-group deliberation. In the presence of weak social influence, reducing the amount of information available can counter-intuitively increase performance. Conversely, when information is already limited, performance might be increased by weakening social influence. However, while confident-neighbor can increase the success of conventional deliberation, it performs comparably to network deliberation using the conform strategy, suggesting that when social influence is strong, network deliberation remains preferable.

\section{Discussion}

We have proposed and evaluated a model of mass deliberation based on the observed interlocking network structure of successful mass-deliberative projects.
Our central finding suggests that network deliberation can improve deliberative output in the presence of strong social influence (i.e., the conform strategy).
Individuals rely on social influence for a number of reasons.
Strong social influence can stem from social factors such as trust or social loafing.
Alternatively, social influence can be used to supplement individual skill when it is insufficient for a given task.
We see that a variety of contexts can lead to reliance on social influence.
Empirical evidence suggests the presence of strong social influence in real-world collaborations such as Wikipedia \cite{platt_network_2018} and in lab studies of human collaboration \cite{mason_collaborative_2012, barkoczi_social_2016}.
In the case of weak social influence (i.e., the  best-neighbor and confident-neighbor strategies), we do not observe a significant difference in output quality between network deliberation and conventional single-group deliberation. In all cases, network deliberation performs better than single-group deliberation or comparatively well. 
Our results suggest that network deliberation is both a contributing mechanism to the success of mass-deliberative projects as well as a useful tool for the design of large-scale sociotechnical systems and interventions for existing systems.

How does network deliberation improve collaborative output? In part, the high performance of random-pod network deliberation can be attributed to the high structural efficiency of the resulting network. However, structural efficiency cannot be the complete mechanism. Long-path network deliberation is extremely inefficient, but still outperforms conventional networks in the presence of strong social influence. The additional mechanisms at play remain an open question.

In order to distinguish between the effects of social influence and individual ability, we have introduced the confident-neighbor strategy, which (unlike best-neighbor) assumes the same agent capabilities as the conform strategy (in a critical learning setting).
While not central to our study of network deliberation, our findings regarding the 
confident-neighbor strategy have important implications for the study of social learning in general. 
Our findings show that just as reducing network efficiency can improve solution quality by raising diversity, reducing the available information can improve the performance of greedy learning strategies. Our investigation of confident-neighbor stems from a careful formal analysis of the information required by each strategy (See Methods). The surprisingly high performance of confident-neighbor shows the importance of isolating social influence from agent ability.

Our work is limited by a number of assumptions, which require further investigation to achieve greater generalizability. We assume a static objective function, while real-world objectives can change with changing external conditions as well as changing individual preferences. We also assume a homogeneous strategy across all agents, which raises questions of how different strategies interact within the same collaboration. We have also assumed that agents are truthful and do not form coalitions outside of the existing network, both assumptions which can be violated in real-world settings.
We have also assumed that agents have access to the full solution string, and that agents incorporate all bits of the string into their quality estimates. Extremely complex real-world tasks might be better-modeled by a collection of sub-tasks, with agents having access to only a subset of the solution string.
Finally, we have limited our study to learning strategies that explore new solutions solely through small mutations of existing solutions. Other potential strategies might generate entirely new solutions, for example, by combining parts of multiple previous solutions. Such generative strategies would add the ability to explore a wider range of solutions. While our work extends the understanding of social learning and collaboration, many questions still remain open for exploration.

Effective mass deliberation is key to realizing the democratizing power of the internet. Our findings suggest that the presence of small and interlocking pods can contribute to the success of mass deliberation, particularly when the topic is complex and social influence is strong. These findings have implications for a range of deliberative contexts: participatory government and budgeting, worker-owned cooperatives, grassroots social movements, and the governance of decentralized systems (e.g., cryptocurrency). By enabling better deliberation at larger scales, we hope this work will contribute to democratizing the governance of large sociotechnical systems, and empowering the individuals impacted by those systems.

### 2021-09-19 ch-intro.tex
Move main text of ABM paper into proposal.

@@
 in use.
+
+Deliberation, a form of collective problem-solving, is a key component of
+democracy, social movements, and online peer-production.
+The network structure of interpersonal communication plays an important role in
+collective problem-solving, particularly in deliberation, which can become
+prohibitively complex and time-intensive at large scales.
+Network properties such as structural efficiency and degree distribution have
+been shown to influence the speed and quality of collective problem-solving,
+in combination with the behavioral strategies employed by individuals.
+However, many successful examples of mass deliberation differ in structure from
+common models of collective problem-solving by exhibiting networks of small
+interlocking groups.
+In this paper, we present an agent-based model of deliberation in networks of
+small interlocking groups and compare performance on these networks with
+conventional network structures.
+We find that networks of small interlocking groups improve solution quality
+when problem-solvers exhibit strong social influence.
+This effect is compatible with, but distinct from, the previously observed
+benefit of network efficiency in the presence of strong social influence.
+Our findings suggest a possible mechanism contributing to the success of
+existing mass deliberative projects as well as a principle for the design of
+new projects.
+By contributing to more effective deliberation at larger scales,
+we hope this work will contribute to democratizing the governance of large
+sociotechnical systems.

### 2021-09-19 proposal.tex
Add bibliography.
Move main text of ABM paper into proposal.

@@
 
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{amsthm}
 \usepackage{apacite}
@@
 \usepackage{helvet}
+\usepackage[utf8]{inputenc}
+\usepackage[switch]{lineno}
 \usepackage[normal]{threeparttable}
-\usepackage{natbib}
-\usepackage{times}
 \usepackage{titlesec}
@@
 \newcommand{\+}{\phantom{-}}
+
+\newtheorem{definition}{Definition}
+\newtheorem{property}{Property}
+\newtheorem{claim}{Claim}
+
+\DeclareMathOperator{\unif}{unif}
+\DeclareMathOperator{\sample}{sample}
+\DeclareMathOperator{\setcount}{count}
+\DeclareMathOperator{\mode}{mode}
+\DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator*{\argmin}{arg\,min}
+
+\linenumbers
 
@@
 \begin{abstract}
-As groups grow in size, they gain access to additional resources,
-creating opportunities for collective intelligence and collective action.
-However, at very large scales, group decision-making becomes prohibitively slow and difficult
-to coordinate.
-Traditional solutions include representative decision-making and/or a shift from deliberation
-to voting.
-Both approaches sacrifice desirable properties.
-Representative decision-making loses the potential benefits of collective intelligence and
-introduces hierarchies that may place the interests of specific individuals ahead of the
-interests of the group.
-Voting sacrifices generativity: allowing a choice between predefined options,
-without allowing for improvement to those options.
-Arrow’s impossibility theorem also fundamentally limits the fairness of voting.
-This project proposes Networked Deliberation as a potential means of large-scale decision-making.
-In Networked Deliberation, members of a large group are repeatedly partitioned into small
-deliberative pods.
-Overlap between pods at different stages enables group-wide diffusion of information and
-preferences.
-Different methods for assigning members to pods result in different network topologies.
-In partnership with large community groups, this project will use a web-based platform to
-evaluate Network Deliberation.
-For a variety of network topologies, the speed and quality of preference convergence will be
-evaluated using quantitative information from ranked-choice polls throughout the process.
-Deliberator satisfaction and experience will also be evaluated using qualitative analysis of
-deliberative text and surveys.
-This work seeks to improve the ability of very large groups to quickly develop a consensus,
-enabling a more effective use of shared resources to achieve common goals.
+\include{abstract.tex}
 \end{abstract}
@@
 of strong social influence}
+\include{ch-abm.tex}
 
@@
 \chapter*{Acknowledgements}
-Thanks to Daniel M. Romero.
-
-Scott E. Page,
-Tawanna Dillahunt,
-Ceren Budak,
-for their time, guidance, and helpful feedback.
-
-Jane Im,
-
-Danielle Livneh,
-Karthik Ramanathan
-for help collecting WikiProjects data.
-
-Yan Chen,
-Tanya Rosenblat,
-
-Grant Schoenbeck,
-Kentaro Toyama,
-Charles R. Severance
-
-Veronica Falandino
-
-Jeffrey W. Lockhart,
-Danaja Maldeniya,
-Lia Bozarth,
-Christopher Quarles,
-Chanda Phelan,
-Padma Chirumamilla,
-Earnest Wheeler,
-Sam Carton,
-Elliott Brannon,
-Tawfiq Ammari,
-2015 UMSI PhD chohort
-Jean Hardy,
-Cindy Lin,
-
-the attendees of the May 25, 2017 MIT Center for Civic Media lab meeting
-the Berkman-Klein Center Cooperation Working Group
-the 2019 CSCW Workshop on Team and Group Diversity
-for helpful feedback;
-
-This research was partly supported by
-the National Science Foundation under Grant No. IIS-1617820.
-
-Andy Brosius,
-Persephone Hernandez-Vogt,
+%\include{acknoledgements.tex}
 
@@
 \bibliographystyle{apacite}
-\bibliography{prelim}
+\bibliography{proposal}
 

### 2021-09-19 NetDelibSim/supplement.tex
Merge network subsection of ABM paper.

--- 2021-04-27 supplement.tex
@@
 
-The simulation begins at time $t=0$ by generating a set of initial solutions $s_{v,0}$ for the agents. These are generated randomly, with each possible solution having equal probability.
-The simulation proceeds iteratively.
-At time $t$ each agent applies one of several {\em learning strategies} (see Section \ref{sec:learning}), to determine its preferred solution at time $t+1$.
-Learning strategies can rely on the agent's own solution, the solutions of its neighbors, and potentially additional information shared by its neighbors.
-Learning strategies may also incorporate information about the objective function, modeling an agents' ability to evaluate the quality of solutions (see Section \ref{sec:capability}).
-To allow for ties, learning strategies produce a set of solutions rather than a single solution.
-In our simulations, we choose winners at random in the case of a tie.
-We allow the simulation to proceed for 300 iterations, which we have found sufficient to guarantee convergence with our chosen parameters.
-
 \section{Network Topologies}
@@
 As a control, we also consider networks representing conventional deliberation.
-
-\subsection{Conventional Deliberation}
-We model the communication network structure of conventional deliberation using two static networks: small-world networks \cite{watts_collective_1998}, and preferential attachment networks \cite{barabasi_emergence_1999}.
-In typical deliberative settings, from deliberative assemblies to online forum threads, the contributions of any member are potentially visible to all other members.
-The network of {\em potential} communication can be modelled by a complete graph.
-However, a complete graph may not be the best model for the {\em actual} communication that takes place in such a network.
-In a real-world deliberation, some participants are more influential than others \cite{goel_structure_2012, shaw_laboratories_2014}, and many communications can be missed or ignored by any particular participant.
-
-Human social networks often exhibit large clustering, and short path lengths.
-To model such dynamics, we use Watts-Strogatz small-world networks  \cite{watts_collective_1998}.
-These networks model, for example, when participants interact mostly with their strong ties, but occasionally with a weak tie.
-Social networks have also been found to exhibit skewed degree distributions with long tails \cite{barabasi_emergence_1999}.
-We model these networks using the Barabási-Albert preferential attachment model.
-These networks model settings where a small number of individuals produce a disproportionate share of the communication due to factors such as social capital, expertise, or confidence.
-
-\subsection{Network Deliberation: Interlock Networks}
-Both of our network deliberation conditions use networks built from interlocking pods.
-These pods are small complete graphs of roughly equal size.
-At any given time, each agent can communicate with all other agents in exactly one pod.
-Periodically, all agents are simultaneously reassigned to new pods.
-These reassignments create an interlocking structure, allowing information to flow through the entire group.
-This structure can be represented as a time-dependent or directed network.
-The properties of an interlock network depend on the particular method for assigning agents to pods.
-Crucially, interlocking structure is created when agents are placed with others they have not yet interacted with.
-We thus require that pod assignment schemes satisfy the following mixing property:
-\begin{property}
-\label{prop:mixing}
-For any agent $v$, and any two assignment stages $t$, $t^\prime$, the probability that at least one neighbor of $v$ at stage $t$ is also a neighbor at stage $t^\prime$ approaches 0 as $|V|$ grows.
-\end{property}
-
-\subsubsection{Random-Pod Assignment}
-The {\em random-pod} assignment method (Algorithm \ref{alg:random}) simply assigns agents to groups at random. Pseudocode for this method is shown in Algorithm \ref{alg:random}.
-
-\begin{algorithm}
-\SetAlgoLined
-\DontPrintSemicolon
-\SetKwFunction{removeRandom}{removeRandom}
-\SetKwFunction{insert}{insert}
-\SetKwFunction{emp}{empty}
-\SetKw{algnot}{not}
-\KwData{A vertex list $V$, the pod size $M \in \mathbb{Z}$.}
-\KwResult{A partition of the vertices.}
-    $N \longleftarrow \lceil \frac{|V|}{M} \rceil$ \;
-    $P \longleftarrow$ List of $N$ empty sets \;
-    \For{$i \in 1, \ldots, M$}{
-        \ForEach{$s \in P$}{
-            \If{\algnot $V.\emp()$}{
-                $v \longleftarrow V.\removeRandom()$ \;
-                $s.\insert(v)$ \;
-            }
-        }
-    }
-    \Return{$P$}
-\caption{Random-Pod Assignment}
-\label{alg:random}
-\end{algorithm}
-
-\begin{claim}
-Random-pod assignment satisfies Property \ref{prop:mixing}.
-\end{claim}
-
-\begin{proof}
-Let $v$ be a vertex of graph $G=(V,E_t)$ with a set of $M$ neighbors at stage $t$ denoted $N_t(v)$.
-The probability that the $k$th neighbor chosen at time $t^\prime \neq t$ belongs to $N_t(v)$, given that the first $k - 1$ did not, is:
-\begin{align*}
-p_\text{repeat}(k)
-&= \frac{M - 1}{|V| - k} \\
-&\leq \frac{M - 1}{|V| - M}.
-\end{align*}
-The probability that none of the $M-1$ neighbors chosen at time $t^\prime$ belong to $N(v)$ thus satisfies:
-\begin{align*}
-p_\text{mixed}
-&\geq
-\left( 1 - \frac{M - 1}{|V| - M} \right)^{M-1} \\
-\lim_{|V|\rightarrow\infty} p_\text{mixed}
-&= 1
-&\qedhere
-\end{align*}
-\end{proof}
-
-Random pod assignment also creates structurally efficient networks with short geodesic distances, as shown in Figure \ref{fig:broadcast}.
-There has been evidence both for
-\cite{lazer_network_2007, derex_partial_2016, mason_propagation_2008, barkoczi_social_2016}
-and against
-\cite{mason_collaborative_2012, barkoczi_social_2016}
-the effectiveness of structurally efficient networks in social learning.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=3.33in]{fig-broadcast.eps}
-    \caption{Mean time necessary for a signal broadcast from one node to reach the entire network. We use mean broadcast time is used as a measure of geodesic length for time-varying networks.}
-    \label{fig:broadcast}
-\end{figure}
-
-\subsubsection{Long-Path Pod Assignment}
-
-To study the effects of path length, we require an alternative pod assignment method that produces long paths.
-However, we must still retain Property \ref{prop:mixing} such that it is rare for two agents to repeatedly share the same pod.
-This property ensures the creation of many new edges at each stage, which is difficult--but not impossible--to reconcile with long average path length.
-We now present a {\em long-path} pod assignment method, which meets both of the above goals.
-
-We begin with a high-level overview.
-Agents are assigned an integer position on a 1-dimensional circular lattice.
-By preferring short-distance links on this lattice, we maintain long path lengths.
-We also partition agents according to the remainder of their position, modulo some prime, i.e., their residue class.
-By limiting links to agents in the same residue class, and using a unique prime at each stage, we ensure that it is rare for multiple agents to share a pod twice in a row.
-Specifically, for stages using primes $p$ and $q$, two agents will share a pod for both stages when their positions are equal modulo $pq$.
-Pseudocode for long-path assignment is shown in Algorithm \ref{alg:long}.
-
-\begin{algorithm}
-\SetAlgoLined
-\DontPrintSemicolon
-\SetKw{algnot}{not}
-\SetKwFunction{removeRandom}{removeRandom}
-\SetKwFunction{moduli}{moduli}
-\SetKwFunction{append}{append}
-\SetKwFunction{enqueue}{enqueue}
-\SetKwFunction{dequeue}{dequeue}
-\SetKwFunction{concat}{concatenate}
-\KwData{A vertex list $V$, the pod size $M \in \mathbb{Z}$, the stage $t \geq 0 \in \mathbb{Z}$, a list of co-prime integers $\moduli$.}
-\KwResult{A partition of the vertices.}
-    $N \longleftarrow \lceil \frac{|V|}{M} \rceil$ \;
-    $P \longleftarrow$ List of $N$ empty sets \;
-    \eIf{$t = 0$}{
-        // Place all vertices in same residue class \;
-        $p \longleftarrow 1$ \;
-    }{
-        // Choose integer to define residue class \;
-        $p \longleftarrow \moduli[t]$ \;
-    }
-    // Assign vertices to residue classes \;
-    $R \longleftarrow$ List of $p$ empty queues \;     
-    \For{$z \in 0, \ldots, |V|-1$}{
-        $R[z \bmod p].\enqueue(V[z])$ \;
-    }
-    // Divide each residue class into pods \;
-    $P \longleftarrow $ Empty list of lists. \;
-    \For{$r \in R$}{
-        $N_r \longleftarrow \lceil \frac{|r|}{M} \rceil$ \;
-        $P_r \longleftarrow$ List of $N_r$ empty lists \;
-        \For{$i \in 1, \ldots, M$}{
-            \For {$j \in 0, \ldots, N_r - 1$}{
-                \If{ \algnot $r.\emp()$}{
-                    $v \leftarrow r.\dequeue()$ \;
-                    $P_r[j].\append(v)$ \;
-                }
-            }
-        }
-        $P.\concat(P_r)$
-    }
-    \Return{$P$}
-\caption{Long-Path Assignment}
-\label{alg:long}
-\end{algorithm}
-
-\begin{claim}
-When the pod size $M$ is less than the modulus $p_t$ for all stages $t$, long-path pod assignment satisfies Property \ref{prop:mixing}.
-\end{claim}
-
-\begin{proof}
-Let $p$ and $q$ be co-prime integers used as moduli for stages $t$ and $t^\prime$ respectively.
-Let $z(x)$ denote the integer position of vertex $x$.
-Let $v$ be a vertex, and let us assume vertex $w$ belongs to the same pod as $v$ at both time $t$ and $t^\prime$.
-This assumption implies:
-\begin{align*}
-    z(v) &\equiv z(w) \pmod{p} \\
-    z(v) &\equiv z(w) \pmod{q} \\
-    \implies z(v) - z(w) &\equiv 0 \pmod{p} \\
-    &\equiv 0 \pmod{q} \\
-    &= npq & \text{for some integer n}.
-\end{align*}
-By the definition of the long-path algorithm:
-\begin{align*}
-    |z(v)-z(w)| & \leq (M - 1)p \\
-    \implies |n|pq & \leq (M - 1)p \\
-    |n|q & \leq (M - 1).
-\end{align*}
-By assumption, $q \geq M$, so the above can only be satisfied by $n=0$,
-which in turn implies $v=w$. \qedhere
-\end{proof}
-
-To see that the long-path procedure produces long geodesics, note that the pods of size $M$ connect nodes at most $(M - 1)p_i$ apart in position, preventing the creation of any shortcut edges, as long as $p_i$ is small compared to $|V|$.
-Numerical simulations confirm that the long-path algorithm produces geodesics larger than random pod assignment (Figure \ref{fig:broadcast}).
-The number of agents may not be an exact multiple of $M$, so the final pods may be truncated.
-However, the network will be a sub-graph of a network that does have a multiple of $M$ nodes, so the structure will not be fundamentally changed, and these edge effects should become negligible as the number of agents increases.
 

### 2021-09-19 ch-abm.tex
Merge network subsection of ABM paper.

@@
 
+\subsection{Overview}
+
 As their collective task, agents seek an optimal point on a rugged fitness landscape.
-This landscape is generated using the tuneably-rugged NK model \cite{kauffman_towards_1987, weinberger_local_1991} (see Methods for more details).
-In this case, task solutions take the form of binary strings.
+This landscape is generated using the NK model \cite{kauffman_towards_1987, weinberger_local_1991} (see Appendix \ref{apx-nk}).
+The NK model is a ``tuanbly rugged'' objective function parameterized by two variables:
+the number of bits in the input string ($N$),
+and the number bits used to compute each contribution to the objective function ($K+1$).
+These parameters can be used to tune problem size and complexity, respectively.
+The resulting objective function maps $N$-bit strings to real numbers in $[0,1]$.
 
@@
 We also introduce the {\em confident-neighbor} strategy, a variant of best-neighbor in which agents announce when they have the best solution in their neighborhood.
-This variant makes weaker assumptions about agent ability (see Methods for more detail) but is equivalent to best-neighbor when network deliberation is used.
+This variant makes weaker assumptions about agent ability but is equivalent to best-neighbor when network deliberation is used.
 This strategy allows us to compare low-social influence and high-social influence strategies under equivalent assumptions of agent ability.
@@
 Fallback can be also be interpreted as a model of social loafing \cite{karau_social_1993}.
-For completeness, we include analysis for a second method in which agents first perform social and individual learning in {\em parallel} and then choose the better of the two learning outputs (See Methods for more detail).
+For completeness, we include analysis for a second method in which agents first perform social and individual learning in {\em parallel} and then choose the better of the two learning outputs.
 
-In deliberation, individuals update their beliefs and preferences through interaction with other participants, making it a form of social learning.
-We thus simulate deliberation using a model of social learning closely resembling those of other authors \cite{
-lazer_network_2007,
-mason_propagation_2008,
-barkoczi_social_2016,
-gomez_clustering_2019}.
+\subsection{Agent-Based Model}
+
+We use an agent-based simualation to model deliberation between individuals.
 We model communication between agents using a time-dependent network $(V,E_t)$.
-The vertices $V$ represent agents.
+The vertices $V$ correspond to the agents (individuals).
 The edges $E_t$ allow agents to exchange information with their immediate neighbors at time $t$.
-Agents collaborate to complete a collective task.
-Specifically, agents seek to identify a bit string $s$ that maximizes an objective function $Q(s)$ generated using the NK model \cite{kauffman_towards_1987, weinberger_local_1991}.
-The NK model is a ``tuanbly rugged'' objective function parameterized by two variables: the number of bits in the input string (N), and the number of other bits each bit interacts with (K).
-These parameters can be used to tune problem size and complexity, respectively.
-The resulting objective function maps N-bit strings to real numbers in $[0,1]$.
-The resulting values can be exponentiated to modify the skewness of the solution value distribution.
-A full description of our model can be found in the Supplementary Information.
-In keeping with other authors \cite{lazer_network_2007, barkoczi_social_2016}, we use N=15, K=6, and an exponent of 8 throughout this paper.
-In keeping with convention, we normalize fitness values to the global maximum.
-In our simulations each agent maintains exactly one preferred solution (initially random).
-The simulation proceeds iteratively, with agents updating their preferred solution once per iteration based on observation of their neighbors' solutions.
-Throughout this paper, we simulate 300 iterations to ensure convergence.
-We present results for 1000 randomly generated NK model objectives, with each network/strategy combination simulated once per objective function.
+Over the course of the simulation, agents seek to maximize some objective function $Q(s)$.
+We use binary strings of length $d$ as our solution space: $s \in \mathbb{Z}^d$.
+We generate the objective function $Q(s)$ using the NK model \cite{kauffman_towards_1987, weinberger_local_1991} (see Appendix \ref{apx-nk}).
+
+The simulation begins at time $t=0$ by generating a set of initial solutions $s_{v,0}$ for the agents.
+These are generated randomly, with each possible solution having equal probability.
+The simulation proceeds iteratively.
+At time $t$ each agent applies one of several {\em learning strategies}, to determine its preferred solution at time $t+1$.
+Learning strategies can rely on the agent's own solution, the solutions of its neighbors, and potentially additional information shared by its neighbors.
+Learning strategies may also incorporate information about the objective function, modeling an agents' ability to evaluate the quality of solutions.
+To allow for ties, learning strategies produce a set of solutions rather than a single solution.
+In our simulations, we choose winners at random in the case of a tie.
+We present results for 1000 randomly generated NK model objectives.
+with each network/strategy combination simulated once per objective function.
+In keeping with other authors \cite{barkoczi_social_2016, lazer_network_2007} we use $N=15$ and $K=6$.
+We allow each simulation to proceed for 300 iterations, which we have found sufficient to guarantee convergence with our chosen parameters.
 
@@
 In addition to the well-known preferential-attachment \cite{barabasi_emergence_1999} and small-world \cite{watts_collective_1998} networks, we construct two types of networks to model network deliberation.
-Both types exhibit interlocking structure and satisfy Property \ref{prop:mixing} (proof in Supplementary Information).
+Both types exhibit interlocking structure and satisfy Property \ref{prop:mixing}.
 These two network deliberation conditions differ in how agents are assigned to pods and in the structural efficiency of the resulting network.
 At each iteration, existing edges are replaced with cliques corresponding to the pods. An edge is created between two agents if and only if they share a pod.
-For {\em random-pod} assignment, agents are assigned to pods at random.
-This assignment method produces short geodesic paths and efficient network structure (see Supplementary Information).
-For {\em long-path} assignment, agents are first assigned an integer label.
-At each stage, a new prime integer is chosen and agents are divided into classes based on the remainder of their label modulo the prime.
-Each class is then divided into size-M pods by selecting the lowest M labels, then the next-lowest M, and so on.
-The prime residue classes guarantee interlocking structure, while the grouping of similar labels prevents the creation of long-distance ``shortcut'' edges, producing an inefficient network (see Supplementary Information).
+In {\em random-pod} assignment, agents are assigned to pods at random.
+This assignment method produces short geodesic paths and efficient network structure.
+In {\em long-path} assignment, agents are assigned using an algorithm which guarantees interlocking structure while preventing the creation of long-distance ``shortcut'' edges, producing a structurally inefficient network.
 All networks are constructed with 100 vertices/agents and a mean degree of 4 (pod size of 5).
+
+\subsection{Conventional Deliberation}
+We model the communication network structure of conventional deliberation using two static networks: small-world networks \cite{watts_collective_1998}, and preferential attachment networks \cite{barabasi_emergence_1999}.
+In typical deliberative settings, from deliberative assemblies to online forum threads, the contributions of any member are potentially visible to all other members.
+The network of {\em potential} communication can be modelled by a complete graph.
+However, a complete graph may not be the best model for the {\em actual} communication that takes place in such a network.
+In a real-world deliberation, some participants are more influential than others \cite{goel_structure_2012, shaw_laboratories_2014}, and many communications can be missed or ignored by any particular participant.
+
+Human social networks often exhibit large clustering, and short path lengths.
+To model such dynamics, we use Watts-Strogatz small-world networks  \cite{watts_collective_1998}.
+These networks model, for example, when participants interact mostly with their strong ties, but occasionally with a weak tie.
+Social networks have also been found to exhibit skewed degree distributions with long tails \cite{barabasi_emergence_1999}.
+We model these networks using the Barabási-Albert preferential attachment model.
+These networks model settings where a small number of individuals produce a disproportionate share of the communication due to factors such as social capital, expertise, or confidence.
+
+\subsection{Network Deliberation: Interlock Networks}
+Both of our network deliberation conditions use networks built from interlocking pods.
+These pods are small complete graphs of roughly equal size.
+At any given time, each agent can communicate with all other agents in exactly one pod.
+Periodically, all agents are simultaneously reassigned to new pods.
+These reassignments create an interlocking structure, allowing information to flow through the entire group.
+This structure can be represented as a time-dependent or directed network.
+The properties of an interlock network depend on the particular method for assigning agents to pods.
+Crucially, interlocking structure is created when agents are placed with others they have not yet interacted with.
+We thus require that pod assignment schemes satisfy the mixing property (Property \ref{prop:mixing}).
+
+\subsubsection{Random-Pod Assignment}
+The {\em random-pod} assignment method (Algorithm \ref{alg:random}) simply assigns agents to groups at random. Pseudocode for this method is shown in Algorithm \ref{alg:random}.
+
+\begin{algorithm}
+\SetAlgoLined
+\DontPrintSemicolon
+\SetKwFunction{removeRandom}{removeRandom}
+\SetKwFunction{insert}{insert}
+\SetKwFunction{emp}{empty}
+\SetKw{algnot}{not}
+\KwData{A vertex list $V$, the pod size $M \in \mathbb{Z}$.}
+\KwResult{A partition of the vertices.}
+    $N \longleftarrow \lceil \frac{|V|}{M} \rceil$ \;
+    $P \longleftarrow$ List of $N$ empty sets \;
+    \For{$i \in 1, \ldots, M$}{
+        \ForEach{$s \in P$}{
+            \If{\algnot $V.\emp()$}{
+                $v \longleftarrow V.\removeRandom()$ \;
+                $s.\insert(v)$ \;
+            }
+        }
+    }
+    \Return{$P$}
+\caption{Random-Pod Assignment}
+\label{alg:random}
+\end{algorithm}
+
+\begin{claim}
+Random-pod assignment satisfies Property \ref{prop:mixing}.
+\end{claim}
+
+\begin{proof}
+Let $v$ be a vertex of graph $G=(V,E_t)$ with a set of $M$ neighbors at stage $t$ denoted $N_t(v)$.
+The probability that the $k$th neighbor chosen at time $t^\prime \neq t$ belongs to $N_t(v)$, given that the first $k - 1$ did not, is:
+\begin{align*}
+p_\text{repeat}(k)
+&= \frac{M - 1}{|V| - k} \\
+&\leq \frac{M - 1}{|V| - M}.
+\end{align*}
+The probability that none of the $M-1$ neighbors chosen at time $t^\prime$ belong to $N(v)$ thus satisfies:
+\begin{align*}
+p_\text{mixed}
+&\geq
+\left( 1 - \frac{M - 1}{|V| - M} \right)^{M-1} \\
+\lim_{|V|\rightarrow\infty} p_\text{mixed}
+&= 1
+&\qedhere
+\end{align*}
+\end{proof}
+
+Random pod assignment also creates structurally efficient networks with short geodesic distances, as shown in Figure \ref{fig:broadcast}.
+There has been evidence both for
+\cite{lazer_network_2007, derex_partial_2016, mason_propagation_2008, barkoczi_social_2016}
+and against
+\cite{mason_collaborative_2012, barkoczi_social_2016}
+the effectiveness of structurally efficient networks in social learning.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=3.33in]{fig/NetDelibABM/fig-broadcast.eps}
+    \caption{Mean time necessary for a signal broadcast from one node to reach the entire network. We use mean broadcast time as a measure of geodesic length for time-varying networks.}
+    \label{fig:broadcast}
+\end{figure}
+
+\subsubsection{Long-Path Pod Assignment}
+
+To study the effects of path length, we require an alternative pod assignment method that produces long paths.
+However, we must still retain Property \ref{prop:mixing} such that it is rare for two agents to repeatedly share the same pod.
+This property ensures the creation of many new edges at each stage, which is difficult--but not impossible--to reconcile with long average path length.
+We now present a {\em long-path} pod assignment method, which meets both of the above goals.
+
+We begin with a high-level overview.
+Agents are assigned an integer position on a 1-dimensional circular lattice.
+By preferring short-distance links on this lattice, we maintain long path lengths.
+We also partition agents according to the remainder of their position, modulo some prime, i.e., their residue class.
+By limiting links to agents in the same residue class, and using a unique prime at each stage, we ensure that it is rare for multiple agents to share a pod twice in a row.
+Specifically, for stages using primes $p$ and $q$, two agents will share a pod for both stages when their positions are equal modulo $pq$.
+Pseudocode for long-path assignment is shown in Algorithm \ref{alg:long}.
+
+\begin{algorithm}
+\SetAlgoLined
+\DontPrintSemicolon
+\SetKw{algnot}{not}
+\SetKwFunction{removeRandom}{removeRandom}
+\SetKwFunction{moduli}{moduli}
+\SetKwFunction{append}{append}
+\SetKwFunction{enqueue}{enqueue}
+\SetKwFunction{dequeue}{dequeue}
+\SetKwFunction{concat}{concatenate}
+\KwData{A vertex list $V$, the pod size $M \in \mathbb{Z}$, the stage $t \geq 0 \in \mathbb{Z}$, a list of co-prime integers $\moduli$.}
+\KwResult{A partition of the vertices.}
+    $N \longleftarrow \lceil \frac{|V|}{M} \rceil$ \;
+    $P \longleftarrow$ List of $N$ empty sets \;
+    \eIf{$t = 0$}{
+        // Place all vertices in same residue class \;
+        $p \longleftarrow 1$ \;
+    }{
+        // Choose integer to define residue class \;
+        $p \longleftarrow \moduli[t]$ \;
+    }
+    // Assign vertices to residue classes \;
+    $R \longleftarrow$ List of $p$ empty queues \;     
+    \For{$z \in 0, \ldots, |V|-1$}{
+        $R[z \bmod p].\enqueue(V[z])$ \;
+    }
+    // Divide each residue class into pods \;
+    $P \longleftarrow $ Empty list of lists. \;
+    \For{$r \in R$}{
+        $N_r \longleftarrow \lceil \frac{|r|}{M} \rceil$ \;
+        $P_r \longleftarrow$ List of $N_r$ empty lists \;
+        \For{$i \in 1, \ldots, M$}{
+            \For {$j \in 0, \ldots, N_r - 1$}{
+                \If{ \algnot $r.\emp()$}{
+                    $v \leftarrow r.\dequeue()$ \;
+                    $P_r[j].\append(v)$ \;
+                }
+            }
+        }
+        $P.\concat(P_r)$
+    }
+    \Return{$P$}
+\caption{Long-Path Assignment}
+\label{alg:long}
+\end{algorithm}
+
+\begin{claim}
+When the pod size $M$ is less than the modulus $p_t$ for all stages $t$, long-path pod assignment satisfies Property \ref{prop:mixing}.
+\end{claim}
+
+\begin{proof}
+Let $p$ and $q$ be co-prime integers used as moduli for stages $t$ and $t^\prime$ respectively.
+Let $z(x)$ denote the integer position of vertex $x$.
+Let $v$ be a vertex, and let us assume vertex $w$ belongs to the same pod as $v$ at both time $t$ and $t^\prime$.
+This assumption implies:
+\begin{align*}
+    z(v) &\equiv z(w) \pmod{p} \\
+    z(v) &\equiv z(w) \pmod{q} \\
+    \implies z(v) - z(w) &\equiv 0 \pmod{p} \\
+    &\equiv 0 \pmod{q} \\
+    &= npq & \text{for some integer n}.
+\end{align*}
+By the definition of the long-path algorithm:
+\begin{align*}
+    |z(v)-z(w)| & \leq (M - 1)p \\
+    \implies |n|pq & \leq (M - 1)p \\
+    |n|q & \leq (M - 1).
+\end{align*}
+By assumption, $q \geq M$, so the above can only be satisfied by $n=0$,
+which in turn implies $v=w$. \qedhere
+\end{proof}
+
+To see that the long-path procedure produces long geodesics, note that the pods of size $M$ connect nodes at most $(M - 1)p_i$ apart in position, preventing the creation of any shortcut edges, as long as $p_i$ is small compared to $|V|$.
+Numerical simulations confirm that the long-path algorithm produces geodesics larger than random pod assignment (Figure \ref{fig:broadcast}).
+The number of agents may not be an exact multiple of $M$, so the final pods may be truncated.
+However, the network will be a sub-graph of a network that does have a multiple of $M$ nodes, so the structure will not be fundamentally changed, and these edge effects should become negligible as the number of agents increases.
 
@@
 \subsection{Statistical Methods}
-All comparisons are made using two-tailed paired t-tests.
-Pairs of observations correspond to the same instance of the NK-model objective function.
+All comparisons are made using two-tailed paired t-tests. Pairs of observations correspond to the same instance of the NK-model objective function.
 Significance values have been corrected for multiple comparisons using the Bonferroni correction.
-A full table of t-values and corrected p-values is available in the Supplemental Information.
 
@@
 Here we present the results of 1000 runs of an agent-based model of deliberation.
-In each run, an NK model is generated (N=15, K=6) and the model is run for 300
-iterations for each network/strategy combination (see Methods for more details).
+In each run, an NK model is generated (N=15, K=6) and the model is simulated for 300
+iterations for each network/strategy combination.
 The results of these simulations are shown in Figure \ref{fig:results}.
@@
 In the special case of network deliberation, the assumptions can be weakened by applying a variant of the strategy, which has not been previously studied.
-The variant, which we call {\em confident-neighbor}, is equivalent to best-neighbor for network deliberation, but not in general (see Methods for more detail).
+The variant, which we call {\em confident-neighbor}, is equivalent to best-neighbor for network deliberation, but not in general.
 Confident-neighbor thus provides a comparison between network deliberation and conventional single-group deliberation under weaker informational assumptions than best-neighbor.
@@
 Our findings show that just as reducing network efficiency can improve solution quality by raising diversity, reducing the available information can improve the performance of greedy learning strategies.
-Our investigation of confident-neighbor stems from a careful formal analysis of the information required by each strategy (See Methods).
+Our investigation of confident-neighbor stems from a careful formal analysis of the information required by each strategy.
 The surprisingly high performance of confident-neighbor shows the importance of isolating social influence from agent ability.

### 2021-09-20 NetDelibSim/supplement.tex
Merge ABM paper strategy section.

@@
 \label{sec:learning}
-The choice of network structure defines which agents can exchange information, but not how those agents act on the information received.
-An agent's actions based on available information are determined by that agent's {\em learning strategy}.
-Learning strategies can be divided into {\em individual learning strategies} which use only information directly observable by the agent, and {\em social learning strategies} which use information communicated by neighboring agents.
-
-Generally, a learning strategy may incorporate both social and individual components, but we will discus each individually before approaching hybrid strategies.
-The learning strategies available to an agent also depend on the agent's ability to evaluate the quality of candidate solutions, which we refer to as its {\em capabilities}, which we now explore in further detail.
-
-\subsection{Agent Capabilities}
-\label{sec:capability}
-In human social learning, the ability of individuals to evaluate the quality of a solution can vary with factors such as expertise and task type.
-When comparing learning strategies, it will be helpful to classify them according to the capabilities they require of agents.
-We formalize these capabilities as oracle functions, which accept one or more task solutions as input and reveal whether the supplied state(s) satisfy some particular property.
-
-One of the strongest capabilities an agent might posses, is to compare two arbitrary solutions to determine which yields a higher value of the objective function.
-We represent this capability using the {\em arbitrary comparison} oracle.
-\begin{definition}
-The arbitrary comparison oracle $\mathcal{O^>}(s_1, s_2)$ is given by:
-\begin{eqnarray}
-\mathcal{O}^>(s_1, s_2) &\equiv&
-Q(s_1) > Q(s_2)
-\end{eqnarray}
-\end{definition}
-
-Under a weaker assumption, agents might be ``experts'' on their current solution and able to compare that particular solution to any other.
-This ability is represented by a {\em single-solution comparison} oracle.
-\begin{definition}
-The single-solution comparison oracle for solution $s$ is given by:
-\begin{eqnarray}
-\mathcal{O}^>_s(\tilde{s}) &\equiv&
-Q(\tilde{s}) > Q(s)
-\end{eqnarray}
-\end{definition}
-
-In many contexts, it is reasonable to assume that agents can explore the effects of small changes to their current solution.
-When the solutions are binary strings, such variations can be modeled by flipping a single bit of the solution string.
-We formalize this capability through the {\em mutation comparison} oracle.
-\begin{definition}
-The mutation comparison oracle for solution $s$ is given by:
-\begin{eqnarray}
-\mathcal{O}^\oplus_s(i) &\equiv&
-Q(s \oplus e_i) > Q(s),
-\end{eqnarray}
-where $\oplus$ represents addition mod 2 and $e_i$ represents the binary string with a single 1 at index $i$.
-\end{definition}
-
-\subsection{Mutation-Based Individual Learning}
-In real-world collaborations, individuals sometimes work independently, even when communication is available.
-Independent work might be motivated by practical concerns (such as distributing labor) or by social dynamics.
-We model individual learning using single-bit mutations, following the examples of
-\cite{lazer_network_2007} and \cite{barkoczi_social_2016}.
-Agents constructs a candidate solution by flipping a single bit of their solution at an index selected uniformly at random.
-If the candidate solution yields a higher value of the objective function, the agent adopts it as the new solution.
-As this strategy requires comparing solutions that differ by at most one bit, an oracle at least as powerful as the mutation comparison oracle is required.
-
-\begin{definition}
-The mutation individual learning strategy $\mathcal{L}_{\text{I}}(v)$ is defined as:
-\begin{eqnarray}
-i &\sim& \unif(1,d) \nonumber\\
-\mathcal{L}_{\text{I}}(v) &\equiv&
-\begin{cases}
-\{s(v) \oplus e_i\} & \mbox{if $\mathcal{O}^{\oplus}_{s(v)}(i)$}
-\\
-\{s(v)\} & \mbox{otherwise}.
-\end{cases}
-\end{eqnarray}
-\end{definition}
-
-\subsection{Best-Neighbor}
-Among social strategies, the straightforward greedy approach results in the {\em best-neighbor} strategy, which has been widely used in prior work \cite{lazer_network_2007, mason_propagation_2008, barkoczi_social_2016}.
-In this strategy, an agent simply compares the solutions of all agents in its neighborhood and adopts the best.
-While straightforward, this strategy makes a strong assumption about agent capabilities: access to the arbitrary comparisons.
-\begin{definition}
-The Best-Neighbor strategy $\mathcal{L}_{BN}(v)$ is defined as:
-\begin{eqnarray}
-\mathcal{L}_{BN}(v)
-&\equiv&
-\{ \, s \!\in\! S(v)
-\mid
-\forall \tilde{s} \!\in\! S(v) \, \
-\lnot \, \mathcal{O}^{>}(\tilde{s}, s),
-\end{eqnarray}
-where $S(v)$ is the multiset of solutions of $v$'s neighbors.
-\end{definition}
-
-\subsection{Confident-Neighbor}
-This paper introduces {\em confident-neighbor}, an alternative to the best-neighbor strategy which relies only on the single comparison oracle and which reduces to best-neighbor for interlocking pod networks.
-The confident neighbor strategy also represents a moderate level of social loafing \cite{karau_social_1993}, in which agents do not actively seek to improve their solution, but passively adopt better solutions if they are presented.
-
-Confident neighbor proceeds in two stages.
-In the first stage, agents determine if their current solution is at least as good as all others in their neighborhood.
-If so, we call the agent {\em confident}.
-In the second stage, confident agents broadcast their solution to all of their neighbors.
-Non-confident agents choose randomly between any broadcast solutions they receive, or keep their original solution if they receive none.
-
-\begin{definition}
-The confident-neighbor strategy $\mathcal{L}_{CN}$ is defined as:
-\begin{eqnarray}
-C(v) &\equiv& \{
-\, s(w)  \nonumber \\
-&& \,
-\mid
-\forall w \!\in\! N(v) \, \forall u \!\in\! N(w) \,
-\lnot \, \mathcal{O}^{>}_{s(w)}(s(u)) \,
-\}
-\\
-\mathcal{L}_{\text{CN}}(v)
-&\equiv& 
-\begin{cases}
-C(v) & \text{if } C(v) \neq \varnothing \\
-\{ \, s(v) \, \}& \text{otherwise,}
-\end{cases}
-\end{eqnarray}
-where $N(v)$ is the set of vertices neighboring $v$ and $s(v)$ is the current candidate solution for agent $v$.
-\end{definition}
-
-Confident-neighbor differs from best-neighbor because of two subtle but important considerations.
-First, an agent's neighbors need not be adjacent to each other.
-Second, an agent's neighbors can be adjacent to others outside that agent's neighborhood.
-As a result, an agent might receive zero, one, or many broadcasts.
-The exception is when agents belong to a clique (e.g., in network deliberation), in which case agents receive exactly one broadcast for each maximal solution in the clique, thus yielding the same results as best-neighbor.
-Confident-neighbor is thus more appropriate than best-neighbor for comparing network deliberation to single group deliberation, as the same information is utilized in both cases.
-
-\subsection{Conform}
-In some contexts, agents might rely on information other than solution quality in their learning strategies.
-Agents might do so out of necessity if they are not able to compare arbitrary solutions.
-Alternatively, agents might ignore solution quality due to social dynamics, e.g.
-social loafing \cite{karau_social_1993}.
-In these contexts, agents might instead evaluate solutions based on popularity, which produces the {\em conform} learning strategy.
-When using the conform strategy, agents count the number of times each solution appears among their neighbors, and adopt the most popular.
-\begin{definition}
-The conform strategy $\mathcal{L}_{\text{C}}(v)$ is defined as:
-\begin{eqnarray}
-\mathcal{L}_{\text{C}}(v) &=&
-\mode(S(v)),
-\end{eqnarray}
-where $S(v)$ is the multiset of candidate solutions for all vertices neighboring $v$, and $\mode()$ returns a set containing the mode or modes of a multiset.
-\end{definition}
-Note that the conform strategy does not depend on the objective function or any oracles, meaning it incorporates no new information about the quality of the solutions.
 
@@
 % End of bitwise majority
-\fi
-
-\subsection{Combining Learning Strategies}\label{subsubsec:combine}
-Models of social learning often combine truly social strategies with individual learning \cite{lazer_network_2007, barkoczi_social_2016,gomez_clustering_2019}.
-We find that the method used to combine social and individual strategies can result in a notable difference in outcome. As a robustness check, we consider two cases for all of our analyses:
-\begin{description}
-\iffalse
-\item[Serial]{
-Agents first perform individual learning to produce an intermediate solution, and then apply social learning to the intermediate solutions.
-This method does not rely on any of the solution comparison oracles.}
-\fi
-\item[Parallel]{
-Agents apply both social and individual learning strategies to their initial solution to produce two competing intermediate solutions, then adopt the better of the two.
-This method relies on the arbitrary comparison oracle.}
-\item[Fallback]{
-Agents first apply social learning to the current solution, but "fall back" to individual learning if the result is not an improvement on the original solution \cite{lazer_network_2007, barkoczi_social_2016,gomez_clustering_2019}.
-Fallback relies on the single comparison oracle and could be motivated by limited agent capability or social loafing.}
-\end{description}
-
-\iffalse
-We must also consider {\em criticality} \cite{barkoczi_social_2016, rendell_rogersparadox_2010}.
-After combining individual and social learning according to one of the above methods, non-critical agents immediately adopt the result.
-Critical agents, on the other hand, compare the quality of the new solution to their previous solution, only adopting the new one if it provides an improvement.
-Note that critical behavior relies on information about solution quality and requires single-solution comparison capability, or stronger.
 \fi

### 2021-09-20 ch-abm.tex
Merge ABM paper strategy section.

@@
 \subsection{Learning Strategies}
-We simulate three learning strategies.
-Each strategy combines an individual strategy and a social strategy.
-The individual strategy is the same for all conditions.
-Agents chose a random bit of their current solution and invert that bit.
-If the resulting solution is an improvement, it is kept, otherwise the original is kept.
-We simulate two methods of combining individual and social strategies. In the {\em fallback} condition, individual learning is only applied when social learning fails to improve the solution.
-In the {\em parallel} condition, individual and social strategies are applied to the original solution in parallel and the better of the two is chosen.
-In all cases, a final check is performed, following \cite{barkoczi_social_2016}: new solutions are only accepted if they improve on the original solution.
-Descriptions of the social learning strategies follow (formal definitions can be found in the Supplementary Information).
+The choice of network structure defines which agents can exchange information, but not how those agents act on the information received.
+An agent's actions based on available information are determined by that agent's {\em learning strategy}.
+Learning strategies can be divided into {\em individual learning strategies} which use only information directly observable by the agent, and {\em social learning strategies} which use information communicated by neighboring agents.
+Generally, a learning strategy may incorporate both social and individual components.
+In this paper, we consider three separate strategies: {\em best-neighbor}, {\em confident-neighbor}, and {\em conform}.
+Each is combined with the same mutation-based individual learning strategy.
 
-The {\em best-neighbor} strategy closely follows that used by other authors \cite{lazer_network_2007, barkoczi_social_2016}.
-To apply this strategy, an agent searches the solutions of all immediate neighbors, and chooses the solution maximizing the payoff.
-In the event of a tie, one of the maximal solutions is chosen at random, with probability proportional to the number of neighbors exhibiting that solution.
+To understand why real-world deliberators might choose different strategies,
+it is necessary to realize that the social learning strategies available to an
+individual depend on that individual's ability to evaluate the quality of candidate solutions.
+Individuals with less information about solution quality, must rely
+more heavily on the recommendations and influence of their peers.
+We formalize an agent's ability to evaluate the quality of a candidate solution
+as its {\em capabilities}, which we now explore in further detail.
 
-The {\em conform} strategy also follows previous authors \cite{barkoczi_social_2016}.
-To apply this strategy, agents search the solutions of all immediate neighbors, counting the number of times each unique solution appears.
-The solution with the highest count is adopted.
-In the event of a tie, the original solution is maintained.
+\subsubsection{Agent Capabilities}
+\label{sec:capability}
+In human social learning, the ability of individuals to evaluate the quality of a solution can vary with factors such as expertise and task type.
+When comparing learning strategies, it will be helpful to classify them according to the capabilities they require of agents.
+We formalize these capabilities as oracle functions, which accept one or more task solutions as input and reveal whether the supplied state(s) satisfy some particular property.
 
-The {\em confident-neighbor} strategy is a novel variant of best-neighbor.
-To apply this strategy, agents search the solutions of all immediate neighbors to determine if any are superior to their own.
-If no superior solutions are found, the agent is {\em confident}.
-The confident agent maintains its original solution and broadcasts that solution to all neighbors.
-Alternatively, if superior solutions are found, the agent waits for broadcasts from confident neighbors.
-Agents may receive zero, one, or many such broadcasts.
-If no broadcasts are received, the agent maintains its original solution.
-If one or more broadcasts are received, one of them is chosen at random and that solution is adopted.
-It is worth noting that this strategy is equivalent to best-neighbor when the network is composed of one or more cliques, as in network deliberation.
+One of the strongest capabilities an agent might posses, is to compare two arbitrary solutions to determine which yields a higher value of the objective function.
+We represent this capability using the {\em arbitrary comparison} oracle.
+\begin{definition}
+The arbitrary comparison oracle $\mathcal{O^>}(s_1, s_2)$ is given by:
+\begin{eqnarray}
+\mathcal{O}^>(s_1, s_2) &\equiv&
+Q(s_1) > Q(s_2)
+\end{eqnarray}
+\end{definition}
+
+Under a weaker assumption, agents might be ``experts'' on their current solution and able to compare that particular solution to any other.
+This ability is represented by a {\em single-solution comparison} oracle.
+\begin{definition}
+The single-solution comparison oracle for solution $s$ is given by:
+\begin{eqnarray}
+\mathcal{O}^>_s(\tilde{s}) &\equiv&
+Q(\tilde{s}) > Q(s)
+\end{eqnarray}
+\end{definition}
+
+In many contexts, it is reasonable to assume that agents can explore the effects of small changes to their current solution.
+When the solutions are binary strings, such variations can be modeled by flipping a single bit of the solution string.
+We formalize this capability through the {\em mutation comparison} oracle.
+\begin{definition}
+The mutation comparison oracle for solution $s$ is given by:
+\begin{eqnarray}
+\mathcal{O}^\oplus_s(i) &\equiv&
+Q(s \oplus e_i) > Q(s),
+\end{eqnarray}
+where $\oplus$ represents addition mod 2 and $e_i$ represents the binary string with a single 1 at index $i$.
+\end{definition}
+
+\subsubsection{Mutation-Based Individual Learning}
+In real-world collaborations, individuals sometimes work independently, even when communication is available.
+Independent work might be motivated by practical concerns (such as distributing labor) or by social dynamics.
+We model individual learning using single-bit mutations, following the examples of
+\cite{lazer_network_2007} and \cite{barkoczi_social_2016}.
+Agents constructs a candidate solution by flipping a single bit of their solution at an index selected uniformly at random.
+If the candidate solution yields a higher value of the objective function, the agent adopts it as the new solution.
+As this strategy requires comparing solutions that differ by at most one bit, an oracle at least as powerful as the mutation comparison oracle is required.
+
+\begin{definition}
+The mutation individual learning strategy $\mathcal{L}_{\text{I}}(v)$ is defined as:
+\begin{eqnarray}
+i &\sim& \unif(1,d) \nonumber\\
+\mathcal{L}_{\text{I}}(v) &\equiv&
+\begin{cases}
+\{s(v) \oplus e_i\} & \mbox{if $\mathcal{O}^{\oplus}_{s(v)}(i)$}
+\\
+\{s(v)\} & \mbox{otherwise}.
+\end{cases}
+\end{eqnarray}
+\end{definition}
+
+\subsubsection{Best-Neighbor}
+Among social strategies, the straightforward greedy approach results in the {\em best-neighbor} strategy, which has been widely used in prior work \cite{lazer_network_2007, mason_propagation_2008, barkoczi_social_2016}.
+In this strategy, an agent simply compares the solutions of all agents in its neighborhood and adopts the best.
+While straightforward, this strategy makes a strong assumption about agent capabilities: access to the arbitrary comparisons.
+\begin{definition}
+The Best-Neighbor strategy $\mathcal{L}_{BN}(v)$ is defined as:
+\begin{eqnarray}
+\mathcal{L}_{BN}(v)
+&\equiv&
+\{ \, s \!\in\! S(v)
+\mid
+\forall \tilde{s} \!\in\! S(v) \, \
+\lnot \, \mathcal{O}^{>}(\tilde{s}, s),
+\end{eqnarray}
+where $S(v)$ is the multiset of solutions of $v$'s neighbors.
+\end{definition}
+
+\subsubsection{Confident-Neighbor}
+This paper introduces {\em confident-neighbor}, an alternative to the best-neighbor strategy which relies only on the single comparison oracle and which reduces to best-neighbor for interlocking pod networks.
+The confident neighbor strategy also represents a moderate level of social loafing \cite{karau_social_1993}, in which agents do not actively seek to improve their solution, but passively adopt better solutions if they are presented.
+
+Confident neighbor proceeds in two stages.
+In the first stage, agents determine if their current solution is at least as good as all others in their neighborhood.
+If so, we call the agent {\em confident}.
+In the second stage, confident agents broadcast their solution to all of their neighbors.
+Non-confident agents choose randomly between any broadcast solutions they receive, or keep their original solution if they receive none.
+
+\begin{definition}
+The confident-neighbor strategy $\mathcal{L}_{CN}$ is defined as:
+\begin{eqnarray}
+C(v) &\equiv& \{
+\, s(w)  \nonumber \\
+&& \,
+\mid
+\forall w \!\in\! N(v) \, \forall u \!\in\! N(w) \,
+\lnot \, \mathcal{O}^{>}_{s(w)}(s(u)) \,
+\}
+\\
+\mathcal{L}_{\text{CN}}(v)
+&\equiv& 
+\begin{cases}
+C(v) & \text{if } C(v) \neq \varnothing \\
+\{ \, s(v) \, \}& \text{otherwise,}
+\end{cases}
+\end{eqnarray}
+where $N(v)$ is the set of vertices neighboring $v$ and $s(v)$ is the current candidate solution for agent $v$.
+\end{definition}
+
+Confident-neighbor differs from best-neighbor because of two subtle but important considerations.
+First, an agent's neighbors need not be adjacent to each other.
+Second, an agent's neighbors can be adjacent to others outside that agent's neighborhood.
+As a result, an agent might receive zero, one, or many broadcasts.
+The exception is when agents belong to a clique (e.g., in network deliberation), in which case agents receive exactly one broadcast for each maximal solution in the clique, thus yielding the same results as best-neighbor.
+Confident-neighbor is thus more appropriate than best-neighbor for comparing network deliberation to single group deliberation, as the same information is utilized in both cases.
+
+\subsubsection{Conform}
+In some contexts, agents might rely on information other than solution quality in their learning strategies.
+Agents might do so out of necessity if they are not able to compare arbitrary solutions.
+Alternatively, agents might ignore solution quality due to social dynamics, e.g.
+social loafing \cite{karau_social_1993}.
+In these contexts, agents might instead evaluate solutions based on popularity, which produces the {\em conform} learning strategy.
+When using the conform strategy, agents count the number of times each solution appears among their neighbors, and adopt the most popular.
+\begin{definition}
+The conform strategy $\mathcal{L}_{\text{C}}(v)$ is defined as:
+\begin{eqnarray}
+\mathcal{L}_{\text{C}}(v) &=&
+\mode(S(v)),
+\end{eqnarray}
+where $S(v)$ is the multiset of candidate solutions for all vertices neighboring $v$, and $\mode()$ returns a set containing the mode or modes of a multiset.
+\end{definition}
+Note that the conform strategy does not depend on the objective function or any oracles, meaning it incorporates no new information about the quality of the solutions.
+
+\subsubsection{Combining Learning Strategies}\label{subsubsec:combine}
+Models of social learning often combine truly social strategies with individual learning \cite{lazer_network_2007,barkoczi_social_2016,gomez_clustering_2019}.
+We find that the method used to combine social and individual strategies can result in a notable difference in outcome.
+\begin{description}
+\item[Parallel]{
+Agents apply both social and individual learning strategies to their initial solution to produce two competing intermediate solutions, then adopt the better of the two.
+This method relies on the arbitrary comparison oracle.}
+\item[Fallback]{
+Agents first apply social learning to the current solution, but "fall back" to individual learning if the result is not an improvement on the original solution \cite{lazer_network_2007, barkoczi_social_2016,gomez_clustering_2019}.
+Fallback relies on the single comparison oracle and could be motivated by limited agent capability or social loafing.}
+\item[Serial]{
+Agents first perform individual learning to produce an intermediate solution, and then apply social learning to the intermediate solutions.
+This method does not rely on any of the solution comparison oracles.}
+\end{description}
+We follow existing literature by focusing on the fallback method.
+As a robustness check, we also consider the parallel method,
+which uses the same agent capabilities but is arguably more powerful.
+
+Combined social learning strategies may also differ in their
+{\em criticality} \cite{barkoczi_social_2016, rendell_rogersparadox_2010}.
+After combining individual and social learning according to one of the above methods, non-critical agents immediately adopt the result.
+Critical agents, on the other hand, compare the quality of the new solution to their previous solution, only adopting the new one if it provides an improvement.
+Note that critical behavior relies on information about solution quality and requires single-solution comparison capability, or stronger.
 

### 2021-09-20 NetDelibSim/supplement.tex
Add supplemental results.

@@
 
-Figures \ref{fig:results-frac-parallel} and \ref{fig:results-frac-fallback} show pairwise comparisons of solution quality between strategy/network settings.
-Figures \ref{fig:results-bn-parallel-dist}--\ref{fig:results-conf-fallback-dist} show the complete solution quality distributions for all settings.
-Tables \ref{tab:t-instrat-fallback}--\ref{tab:t-innet-parallel} show the t-values and Bonferroni-corrected p-values for comparisons across networks and strategies.
-
-\begin{figure}
-    \label{fig:results-frac-parallel}
-    \centering
-    \includegraphics[width=3.33in]{fig-result-frac-parallel.eps}
-\caption{Fraction of simulations in which the row condition outperforms the column condition. Results are shown for parallel individual learning. In the case of a tie, weight is divided evenly.}
-\end{figure}
-
-\begin{figure}
-    \label{fig:results-frac-fallback}
-    \centering
-    \includegraphics[width=3.33in]{fig-result-frac-parallel.eps}
-\caption{Fraction of simulations in which the row condition outperforms the column condition. Results are shown for fallback individual learning. In the case of a tie, weight is divided evenly.}
-\end{figure}
-
-\begin{figure}
-    \label{fig:results-bn-parallel-dist}
-    \centering
-    \includegraphics[width=3.33in]{results-bn-parallel-dist.eps}
-\caption{Solution quality distribution for best-neighbor strategy with parallel individual learning.}
-\end{figure}
-\begin{figure}
-    \label{fig:results-cn-parallel-dist}
-    \centering
-    \includegraphics[width=3.33in]{results-cn-parallel-dist.eps}
-\caption{Solution quality distribution for confident-neighbor strategy with parallel individual learning.}
-\end{figure}
-\begin{figure}
-    \label{fig:results-conf-parallel-dist}
-    \centering
-    \includegraphics[width=3.33in]{results-conf-parallel-dist.eps}
-\caption{Solution quality distribution for the conform strategy with parallel individual learning.}
-\end{figure}
-
-\begin{figure}
-    \label{fig:results-bn-fallback-dist}
-    \centering
-    \includegraphics[width=3.33in]{results-bn-fallback-dist.eps}
-\caption{Solution quality distribution for best-neighbor strategy with fallback individual learning.}
-\end{figure}
-\begin{figure}
-    \label{fig:results-cn-fallback-dist}
-    \centering
-    \includegraphics[width=3.33in]{results-cn-fallback-dist.eps}
-\caption{Solution quality distribution for confident-neighbor strategy with fallback individual learning.}
-\end{figure}
-\begin{figure}
-    \label{fig:results-conf-fallback-dist}
-    \centering
-    \includegraphics[width=3.33in]{results-conf-fallback-dist.eps}
-\caption{Solution quality distribution for the conform strategy with fallback individual learning.}
-\end{figure}
-
-\begin{table*}[]
-    \label{tab:t-instrat-fallback}
-    \centering
-    \begin{tabular}{l|ll|ccc}
-        Strategy & Net A & Net B & \overline{B - A} & t & p* \\
-        \hline
-Best Neighbor&Pref. Attach.&Small World&-8.70e-02&-1.33e+01&1.27e-35\\
-Best Neighbor&Pref. Attach.&Long Path&-5.98e-02&-8.62e+00&1.51e-15\\
-Best Neighbor&Pref. Attach.&Random Pod&-1.15e-02&-1.83e+00&4.01e+00\\
-Best Neighbor&Small World&Pref. Attach.&8.70e-02&1.33e+01&1.27e-35\\
-Best Neighbor&Small World&Long Path&2.72e-02&4.53e+00&3.88e-04\\
-Best Neighbor&Small World&Random Pod&7.55e-02&1.18e+01&1.66e-28\\
-Best Neighbor&Long Path&Pref. Attach.&5.98e-02&8.62e+00&1.51e-15\\
-Best Neighbor&Long Path&Small World&-2.72e-02&-4.53e+00&3.88e-04\\
-Best Neighbor&Long Path&Random Pod&4.83e-02&7.34e+00&2.69e-11\\
-Best Neighbor&Random Pod&Pref. Attach.&1.15e-02&1.83e+00&4.01e+00\\
-Best Neighbor&Random Pod&Small World&-7.55e-02&-1.18e+01&1.66e-28\\
-Best Neighbor&Random Pod&Long Path&-4.83e-02&-7.34e+00&2.69e-11\\
-\hline
-Confident Neighbor&Pref. Attach.&Small World&-7.52e-02&-1.13e+01&2.94e-26\\
-Confident Neighbor&Pref. Attach.&Long Path&2.39e-03&3.64e-01&4.30e+01\\
-Confident Neighbor&Pref. Attach.&Random Pod&5.80e-02&8.88e+00&1.86e-16\\
-Confident Neighbor&Small World&Pref. Attach.&7.52e-02&1.13e+01&2.94e-26\\
-Confident Neighbor&Small World&Long Path&7.76e-02&1.21e+01&1.08e-29\\
-Confident Neighbor&Small World&Random Pod&1.33e-01&1.92e+01&1.36e-68\\
-Confident Neighbor&Long Path&Pref. Attach.&-2.39e-03&-3.64e-01&4.30e+01\\
-Confident Neighbor&Long Path&Small World&-7.76e-02&-1.21e+01&1.08e-29\\
-Confident Neighbor&Long Path&Random Pod&5.56e-02&8.50e+00&4.00e-15\\
-Confident Neighbor&Random Pod&Pref. Attach.&-5.80e-02&-8.88e+00&1.86e-16\\
-Confident Neighbor&Random Pod&Small World&-1.33e-01&-1.92e+01&1.36e-68\\
-Confident Neighbor&Random Pod&Long Path&-5.56e-02&-8.50e+00&4.00e-15\\
-\hline
-Conform&Pref. Attach.&Small World&3.60e-02&6.78e+00&1.24e-09\\
-Conform&Pref. Attach.&Long Path&-1.63e-01&-2.22e+01&2.25e-87\\
-Conform&Pref. Attach.&Random Pod&-2.28e-01&-3.19e+01&9.08e-153\\
-Conform&Small World&Pref. Attach.&-3.60e-02&-6.78e+00&1.24e-09\\
-Conform&Small World&Long Path&-1.99e-01&-3.17e+01&1.59e-151\\
-Conform&Small World&Random Pod&-2.64e-01&-4.43e+01&2.20e-236\\
-Conform&Long Path&Pref. Attach.&1.63e-01&2.22e+01&2.25e-87\\
-Conform&Long Path&Small World&1.99e-01&3.17e+01&1.59e-151\\
-Conform&Long Path&Random Pod&-6.48e-02&-8.31e+00&1.79e-14\\
-Conform&Random Pod&Pref. Attach.&2.28e-01&3.19e+01&9.08e-153\\
-Conform&Random Pod&Small World&2.64e-01&4.43e+01&2.20e-236\\
-Conform&Random Pod&Long Path&6.48e-02&8.31e+00&1.79e-14\\
-\hline
-    \end{tabular}
-    \caption{Two-tailed paired t-values across networks, within strategies.
-Results are for fallback individual learning.
-For all rows, dof=999 and $p^*$ is the Bonferroni-corrected p-value with m=60.}
-    \label{tab:my_label}
-\end{table*}
-
-\begin{table*}[]
-    \label{tab:t-innet-fallback}
-    \centering
-    \begin{tabular}{l|ll|ccc}
-        Network & Strategy A & Strategy B & \overline{B - A} & t & p* \\
-        \hline
-Pref. Attach.&Best Neighbor&Confident Neighbor&-6.43e-02&-9.39e+00&2.33e-18\\
-Pref. Attach.&Best Neighbor&Conform&8.96e-02&1.19e+01&6.83e-29\\
-Pref. Attach.&Confident Neighbor&Best Neighbor&6.43e-02&9.39e+00&2.33e-18\\
-Pref. Attach.&Confident Neighbor&Conform&1.54e-01&2.06e+01&7.57e-77\\
-Pref. Attach.&Conform&Best Neighbor&-8.96e-02&-1.19e+01&6.83e-29\\
-Pref. Attach.&Conform&Confident Neighbor&-1.54e-01&-2.06e+01&7.57e-77\\
-\hline
-Small World&Best Neighbor&Confident Neighbor&-5.26e-02&-8.39e+00&9.70e-15\\
-Small World&Best Neighbor&Conform&2.13e-01&3.52e+01&4.46e-175\\
-Small World&Confident Neighbor&Best Neighbor&5.26e-02&8.39e+00&9.70e-15\\
-Small World&Confident Neighbor&Conform&2.65e-01&4.45e+01&1.24e-237\\
-Small World&Conform&Best Neighbor&-2.13e-01&-3.52e+01&4.46e-175\\
-Small World&Conform&Confident Neighbor&-2.65e-01&-4.45e+01&1.24e-237\\
-\hline
-Long Path&Best Neighbor&Confident Neighbor&-2.12e-03&-3.42e-01&4.39e+01\\
-Long Path&Best Neighbor&Conform&-1.37e-02&-1.71e+00&5.20e+00\\
-Long Path&Confident Neighbor&Best Neighbor&2.12e-03&3.42e-01&4.39e+01\\
-Long Path&Confident Neighbor&Conform&-1.16e-02&-1.44e+00&8.96e+00\\
-Long Path&Conform&Best Neighbor&1.37e-02&1.71e+00&5.20e+00\\
-Long Path&Conform&Confident Neighbor&1.16e-02&1.44e+00&8.96e+00\\
-\hline
-Random Pod&Best Neighbor&Confident Neighbor&5.18e-03&8.40e-01&2.41e+01\\
-Random Pod&Best Neighbor&Conform&-1.27e-01&-1.64e+01&1.07e-51\\
-Random Pod&Confident Neighbor&Best Neighbor&-5.18e-03&-8.40e-01&2.41e+01\\
-Random Pod&Confident Neighbor&Conform&-1.32e-01&-1.67e+01&1.39e-53\\
-Random Pod&Conform&Best Neighbor&1.27e-01&1.64e+01&1.07e-51\\
-Random Pod&Conform&Confident Neighbor&1.32e-01&1.67e+01&1.39e-53\\
-\hline
-    \end{tabular}
-    \caption{Two-tailed paired t-values across strategies, within networks.
-Results are for fallback individual learning.
-For all rows, dof=999 and $p^*$ is the Bonferroni-corrected p-value with m=60.}
-    \label{tab:my_label}
-\end{table*}
-
-\begin{table*}[]
-    \label{tab:t-instrat-parallel}
-    \centering
-    \begin{tabular}{l|ll|ccc}
-        Strategy & Net A & Net B & \overline{B - A} & t & p* \\
-    \hline
-Best Neighbor&Pref. Attach.&Small World&-6.75e-02&-1.13e+01&2.48e-26\\
-Best Neighbor&Pref. Attach.&Long Path&-4.24e-02&-6.78e+00&1.25e-09\\
-Best Neighbor&Pref. Attach.&Random Pod&-8.14e-03&-1.32e+00&1.12e+01\\
-Best Neighbor&Small World&Pref. Attach.&6.75e-02&1.13e+01&2.48e-26\\
-Best Neighbor&Small World&Long Path&2.50e-02&4.25e+00&1.42e-03\\
-Best Neighbor&Small World&Random Pod&5.93e-02&9.38e+00&2.54e-18\\
-Best Neighbor&Long Path&Pref. Attach.&4.24e-02&6.78e+00&1.25e-09\\
-Best Neighbor&Long Path&Small World&-2.50e-02&-4.25e+00&1.42e-03\\
-Best Neighbor&Long Path&Random Pod&3.43e-02&5.36e+00&6.14e-06\\
-Best Neighbor&Random Pod&Pref. Attach.&8.14e-03&1.32e+00&1.12e+01\\
-Best Neighbor&Random Pod&Small World&-5.93e-02&-9.38e+00&2.54e-18\\
-Best Neighbor&Random Pod&Long Path&-3.43e-02&-5.36e+00&6.14e-06\\
-\hline
-Confident Neighbor&Pref. Attach.&Small World&-4.49e-02&-7.20e+00&7.03e-11\\
-Confident Neighbor&Pref. Attach.&Long Path&-5.03e-03&-7.65e-01&2.67e+01\\
-Confident Neighbor&Pref. Attach.&Random Pod&2.61e-02&4.02e+00&3.74e-03\\
-Confident Neighbor&Small World&Pref. Attach.&4.49e-02&7.20e+00&7.03e-11\\
-Confident Neighbor&Small World&Long Path&3.99e-02&6.96e+00&3.80e-10\\
-Confident Neighbor&Small World&Random Pod&7.10e-02&1.18e+01&1.98e-28\\
-Confident Neighbor&Long Path&Pref. Attach.&5.03e-03&7.65e-01&2.67e+01\\
-Confident Neighbor&Long Path&Small World&-3.99e-02&-6.96e+00&3.80e-10\\
-Confident Neighbor&Long Path&Random Pod&3.11e-02&4.89e+00&7.03e-05\\
-Confident Neighbor&Random Pod&Pref. Attach.&-2.61e-02&-4.02e+00&3.74e-03\\
-Confident Neighbor&Random Pod&Small World&-7.10e-02&-1.18e+01&1.98e-28\\
-Confident Neighbor&Random Pod&Long Path&-3.11e-02&-4.89e+00&7.03e-05\\
-\hline
-Conform&Pref. Attach.&Small World&3.67e-02&6.85e+00&7.73e-10\\
-Conform&Pref. Attach.&Long Path&-1.73e-01&-2.40e+01&7.10e-99\\
-Conform&Pref. Attach.&Random Pod&-2.77e-01&-4.08e+01&6.29e-213\\
-Conform&Small World&Pref. Attach.&-3.67e-02&-6.85e+00&7.73e-10\\
-Conform&Small World&Long Path&-2.09e-01&-3.46e+01&3.46e-171\\
-Conform&Small World&Random Pod&-3.14e-01&-5.68e+01&1.63e-313\\
-Conform&Long Path&Pref. Attach.&1.73e-01&2.40e+01&7.10e-99\\
-Conform&Long Path&Small World&2.09e-01&3.46e+01&3.46e-171\\
-Conform&Long Path&Random Pod&-1.04e-01&-1.43e+01&8.55e-41\\
-Conform&Random Pod&Pref. Attach.&2.77e-01&4.08e+01&6.29e-213\\
-Conform&Random Pod&Small World&3.14e-01&5.68e+01&1.63e-313\\
-Conform&Random Pod&Long Path&1.04e-01&1.43e+01&8.55e-41\\
-    \hline
-    \end{tabular}
-    \caption{Two-tailed paired t-values across networks, within strategies.
-Results are for parallel individual learning.
-For all rows, dof=999 and $p^*$ is the Bonferroni-corrected p-value with m=60.}
-    \label{tab:my_label}
-\end{table*}
-
-\begin{table*}[]
-    \label{tab:t-innet-parallel}
-    \centering
-    \begin{tabular}{l|ll|ccc}
-        Network & Strategy A & Strategy B & \overline{B - A} & t & p* \\
-    \hline
-Pref. Attach.&Best Neighbor&Confident Neighbor&-3.78e-02&-5.94e+00&2.39e-07\\
-Pref. Attach.&Best Neighbor&Conform&2.00e-01&2.77e+01&3.28e-124\\
-Pref. Attach.&Confident Neighbor&Best Neighbor&3.78e-02&5.94e+00&2.39e-07\\
-Pref. Attach.&Confident Neighbor&Conform&2.38e-01&3.40e+01&1.67e-167\\
-Pref. Attach.&Conform&Best Neighbor&-2.00e-01&-2.77e+01&3.28e-124\\
-Pref. Attach.&Conform&Confident Neighbor&-2.38e-01&-3.40e+01&1.67e-167\\
-\hline
-Small World&Best Neighbor&Confident Neighbor&-1.52e-02&-2.74e+00&3.80e-01\\
-Small World&Best Neighbor&Conform&3.05e-01&5.30e+01&1.25e-290\\
-Small World&Confident Neighbor&Best Neighbor&1.52e-02&2.74e+00&3.80e-01\\
-Small World&Confident Neighbor&Conform&3.20e-01&5.89e+01&0.00e+00\\
-Small World&Conform&Best Neighbor&-3.05e-01&-5.30e+01&1.25e-290\\
-Small World&Conform&Confident Neighbor&-3.20e-01&-5.89e+01&0.00e+00\\
-\hline
-Long Path&Best Neighbor&Confident Neighbor&-3.88e-04&-6.64e-02&5.68e+01\\
-Long Path&Best Neighbor&Conform&7.02e-02&9.34e+00&3.64e-18\\
-Long Path&Confident Neighbor&Best Neighbor&3.88e-04&6.64e-02&5.68e+01\\
-Long Path&Confident Neighbor&Conform&7.06e-02&9.15e+00&1.83e-17\\
-Long Path&Conform&Best Neighbor&-7.02e-02&-9.34e+00&3.64e-18\\
-Long Path&Conform&Confident Neighbor&-7.06e-02&-9.15e+00&1.83e-17\\
-\hline
-Random Pod&Best Neighbor&Confident Neighbor&-3.59e-03&-5.46e-01&3.51e+01\\
-Random Pod&Best Neighbor&Conform&-6.84e-02&-9.54e+00&6.11e-19\\
-Random Pod&Confident Neighbor&Best Neighbor&3.59e-03&5.46e-01&3.51e+01\\
-Random Pod&Confident Neighbor&Conform&-6.48e-02&-8.96e+00&9.43e-17\\
-Random Pod&Conform&Best Neighbor&6.84e-02&9.54e+00&6.11e-19\\
-Random Pod&Conform&Confident Neighbor&6.48e-02&8.96e+00&9.43e-17\\
-    \hline
-    \end{tabular}
-    \caption{Two-tailed paired t-values across strategies, within networks.
-Results are for parallel individual learning.
-For all rows, dof=999 and $p^*$ is the Bonferroni-corrected p-value with m=60.}
-    \label{tab:my_label}
-\end{table*}
-
 \bibliography{references}

### 2021-09-20 ch-abm.tex
Add supplemental results.

@@
 The results of these simulations are shown in Figure \ref{fig:results}.
+Figures \ref{fig:results-frac-parallel} and \ref{fig:results-frac-fallback} show pairwise comparisons of solution quality between strategy/network settings.
+Figures \ref{fig:results-bn-parallel-dist}--\ref{fig:results-conf-fallback-dist} show the complete solution quality distributions for all settings.
+Tables \ref{tab:t-instrat-fallback}--\ref{tab:t-innet-parallel} show the t-values and Bonferroni-corrected p-values for comparisons across networks and strategies.
 
@@
 \end{figure*}
+
+\begin{figure}
+    \label{fig:results-frac-parallel}
+    \centering
+    \includegraphics[width=3.33in]{fig/NetDelibABM/fig-result-frac-parallel.eps}
+\caption{Fraction of simulations in which the row condition outperforms the column condition. Results are shown for parallel individual learning. In the case of a tie, weight is divided evenly.}
+\end{figure}
+
+\begin{figure}
+    \label{fig:results-frac-fallback}
+    \centering
+%    \includegraphics[width=3.33in]{fig/NetDelibABM/fig-result-frac-parallel.eps}
+\caption{Fraction of simulations in which the row condition outperforms the column condition. Results are shown for fallback individual learning. In the case of a tie, weight is divided evenly.}
+\end{figure}
+
+\begin{figure}
+    \label{fig:results-bn-parallel-dist}
+    \centering
+    \includegraphics[width=3.33in]{fig/NetDelibABM/results-bn-parallel-dist.eps}
+\caption{Solution quality distribution for best-neighbor strategy with parallel individual learning.}
+\end{figure}
+\begin{figure}
+    \label{fig:results-cn-parallel-dist}
+    \centering
+    \includegraphics[width=3.33in]{fig/NetDelibABM/results-cn-parallel-dist.eps}
+\caption{Solution quality distribution for confident-neighbor strategy with parallel individual learning.}
+\end{figure}
+\begin{figure}
+    \label{fig:results-conf-parallel-dist}
+    \centering
+    \includegraphics[width=3.33in]{fig/NetDelibABM/results-conf-parallel-dist.eps}
+\caption{Solution quality distribution for the conform strategy with parallel individual learning.}
+\end{figure}
+
+\begin{figure}
+    \label{fig:results-bn-fallback-dist}
+    \centering
+    \includegraphics[width=3.33in]{fig/NetDelibABM/results-bn-fallback-dist.eps}
+\caption{Solution quality distribution for best-neighbor strategy with fallback individual learning.}
+\end{figure}
+\begin{figure}
+    \label{fig:results-cn-fallback-dist}
+    \centering
+    \includegraphics[width=3.33in]{fig/NetDelibABM/results-cn-fallback-dist.eps}
+\caption{Solution quality distribution for confident-neighbor strategy with fallback individual learning.}
+\end{figure}
+\begin{figure}
+    \label{fig:results-conf-fallback-dist}
+    \centering
+    \includegraphics[width=3.33in]{fig/NetDelibABM/results-conf-fallback-dist.eps}
+\caption{Solution quality distribution for the conform strategy with fallback individual learning.}
+\end{figure}
+
+\begin{table*}[]
+    \label{tab:t-instrat-fallback}
+    \centering
+    \begin{tabular}{l|ll|ccc}
+        Strategy & Net A & Net B & $\overline{B - A}$ & t & p* \\
+        \hline
+Best Neighbor&Pref. Attach.&Small World&-8.70e-02&-1.33e+01&1.27e-35\\
+Best Neighbor&Pref. Attach.&Long Path&-5.98e-02&-8.62e+00&1.51e-15\\
+Best Neighbor&Pref. Attach.&Random Pod&-1.15e-02&-1.83e+00&4.01e+00\\
+Best Neighbor&Small World&Pref. Attach.&8.70e-02&1.33e+01&1.27e-35\\
+Best Neighbor&Small World&Long Path&2.72e-02&4.53e+00&3.88e-04\\
+Best Neighbor&Small World&Random Pod&7.55e-02&1.18e+01&1.66e-28\\
+Best Neighbor&Long Path&Pref. Attach.&5.98e-02&8.62e+00&1.51e-15\\
+Best Neighbor&Long Path&Small World&-2.72e-02&-4.53e+00&3.88e-04\\
+Best Neighbor&Long Path&Random Pod&4.83e-02&7.34e+00&2.69e-11\\
+Best Neighbor&Random Pod&Pref. Attach.&1.15e-02&1.83e+00&4.01e+00\\
+Best Neighbor&Random Pod&Small World&-7.55e-02&-1.18e+01&1.66e-28\\
+Best Neighbor&Random Pod&Long Path&-4.83e-02&-7.34e+00&2.69e-11\\
+\hline
+Confident Neighbor&Pref. Attach.&Small World&-7.52e-02&-1.13e+01&2.94e-26\\
+Confident Neighbor&Pref. Attach.&Long Path&2.39e-03&3.64e-01&4.30e+01\\
+Confident Neighbor&Pref. Attach.&Random Pod&5.80e-02&8.88e+00&1.86e-16\\
+Confident Neighbor&Small World&Pref. Attach.&7.52e-02&1.13e+01&2.94e-26\\
+Confident Neighbor&Small World&Long Path&7.76e-02&1.21e+01&1.08e-29\\
+Confident Neighbor&Small World&Random Pod&1.33e-01&1.92e+01&1.36e-68\\
+Confident Neighbor&Long Path&Pref. Attach.&-2.39e-03&-3.64e-01&4.30e+01\\
+Confident Neighbor&Long Path&Small World&-7.76e-02&-1.21e+01&1.08e-29\\
+Confident Neighbor&Long Path&Random Pod&5.56e-02&8.50e+00&4.00e-15\\
+Confident Neighbor&Random Pod&Pref. Attach.&-5.80e-02&-8.88e+00&1.86e-16\\
+Confident Neighbor&Random Pod&Small World&-1.33e-01&-1.92e+01&1.36e-68\\
+Confident Neighbor&Random Pod&Long Path&-5.56e-02&-8.50e+00&4.00e-15\\
+\hline
+Conform&Pref. Attach.&Small World&3.60e-02&6.78e+00&1.24e-09\\
+Conform&Pref. Attach.&Long Path&-1.63e-01&-2.22e+01&2.25e-87\\
+Conform&Pref. Attach.&Random Pod&-2.28e-01&-3.19e+01&9.08e-153\\
+Conform&Small World&Pref. Attach.&-3.60e-02&-6.78e+00&1.24e-09\\
+Conform&Small World&Long Path&-1.99e-01&-3.17e+01&1.59e-151\\
+Conform&Small World&Random Pod&-2.64e-01&-4.43e+01&2.20e-236\\
+Conform&Long Path&Pref. Attach.&1.63e-01&2.22e+01&2.25e-87\\
+Conform&Long Path&Small World&1.99e-01&3.17e+01&1.59e-151\\
+Conform&Long Path&Random Pod&-6.48e-02&-8.31e+00&1.79e-14\\
+Conform&Random Pod&Pref. Attach.&2.28e-01&3.19e+01&9.08e-153\\
+Conform&Random Pod&Small World&2.64e-01&4.43e+01&2.20e-236\\
+Conform&Random Pod&Long Path&6.48e-02&8.31e+00&1.79e-14\\
+\hline
+    \end{tabular}
+    \caption{Two-tailed paired t-values across networks, within strategies.
+Results are for fallback individual learning.
+For all rows, dof=999 and $p^*$ is the Bonferroni-corrected p-value with m=60.}
+\end{table*}
+
+\begin{table*}[]
+    \label{tab:t-innet-fallback}
+    \centering
+    \begin{tabular}{l|ll|ccc}
+        Network & Strategy A & Strategy B & $\overline{B - A}$ & t & p* \\
+        \hline
+Pref. Attach.&Best Neighbor&Confident Neighbor&-6.43e-02&-9.39e+00&2.33e-18\\
+Pref. Attach.&Best Neighbor&Conform&8.96e-02&1.19e+01&6.83e-29\\
+Pref. Attach.&Confident Neighbor&Best Neighbor&6.43e-02&9.39e+00&2.33e-18\\
+Pref. Attach.&Confident Neighbor&Conform&1.54e-01&2.06e+01&7.57e-77\\
+Pref. Attach.&Conform&Best Neighbor&-8.96e-02&-1.19e+01&6.83e-29\\
+Pref. Attach.&Conform&Confident Neighbor&-1.54e-01&-2.06e+01&7.57e-77\\
+\hline
+Small World&Best Neighbor&Confident Neighbor&-5.26e-02&-8.39e+00&9.70e-15\\
+Small World&Best Neighbor&Conform&2.13e-01&3.52e+01&4.46e-175\\
+Small World&Confident Neighbor&Best Neighbor&5.26e-02&8.39e+00&9.70e-15\\
+Small World&Confident Neighbor&Conform&2.65e-01&4.45e+01&1.24e-237\\
+Small World&Conform&Best Neighbor&-2.13e-01&-3.52e+01&4.46e-175\\
+Small World&Conform&Confident Neighbor&-2.65e-01&-4.45e+01&1.24e-237\\
+\hline
+Long Path&Best Neighbor&Confident Neighbor&-2.12e-03&-3.42e-01&4.39e+01\\
+Long Path&Best Neighbor&Conform&-1.37e-02&-1.71e+00&5.20e+00\\
+Long Path&Confident Neighbor&Best Neighbor&2.12e-03&3.42e-01&4.39e+01\\
+Long Path&Confident Neighbor&Conform&-1.16e-02&-1.44e+00&8.96e+00\\
+Long Path&Conform&Best Neighbor&1.37e-02&1.71e+00&5.20e+00\\
+Long Path&Conform&Confident Neighbor&1.16e-02&1.44e+00&8.96e+00\\
+\hline
+Random Pod&Best Neighbor&Confident Neighbor&5.18e-03&8.40e-01&2.41e+01\\
+Random Pod&Best Neighbor&Conform&-1.27e-01&-1.64e+01&1.07e-51\\
+Random Pod&Confident Neighbor&Best Neighbor&-5.18e-03&-8.40e-01&2.41e+01\\
+Random Pod&Confident Neighbor&Conform&-1.32e-01&-1.67e+01&1.39e-53\\
+Random Pod&Conform&Best Neighbor&1.27e-01&1.64e+01&1.07e-51\\
+Random Pod&Conform&Confident Neighbor&1.32e-01&1.67e+01&1.39e-53\\
+\hline
+    \end{tabular}
+    \caption{Two-tailed paired t-values across strategies, within networks.
+Results are for fallback individual learning.
+For all rows, dof=999 and $p^*$ is the Bonferroni-corrected p-value with m=60.}
+\end{table*}
+
+\begin{table*}[]
+    \label{tab:t-instrat-parallel}
+    \centering
+    \begin{tabular}{l|ll|ccc}
+        Strategy & Net A & Net B & $\overline{B - A}$ & t & p* \\
+    \hline
+Best Neighbor&Pref. Attach.&Small World&-6.75e-02&-1.13e+01&2.48e-26\\
+Best Neighbor&Pref. Attach.&Long Path&-4.24e-02&-6.78e+00&1.25e-09\\
+Best Neighbor&Pref. Attach.&Random Pod&-8.14e-03&-1.32e+00&1.12e+01\\
+Best Neighbor&Small World&Pref. Attach.&6.75e-02&1.13e+01&2.48e-26\\
+Best Neighbor&Small World&Long Path&2.50e-02&4.25e+00&1.42e-03\\
+Best Neighbor&Small World&Random Pod&5.93e-02&9.38e+00&2.54e-18\\
+Best Neighbor&Long Path&Pref. Attach.&4.24e-02&6.78e+00&1.25e-09\\
+Best Neighbor&Long Path&Small World&-2.50e-02&-4.25e+00&1.42e-03\\
+Best Neighbor&Long Path&Random Pod&3.43e-02&5.36e+00&6.14e-06\\
+Best Neighbor&Random Pod&Pref. Attach.&8.14e-03&1.32e+00&1.12e+01\\
+Best Neighbor&Random Pod&Small World&-5.93e-02&-9.38e+00&2.54e-18\\
+Best Neighbor&Random Pod&Long Path&-3.43e-02&-5.36e+00&6.14e-06\\
+\hline
+Confident Neighbor&Pref. Attach.&Small World&-4.49e-02&-7.20e+00&7.03e-11\\
+Confident Neighbor&Pref. Attach.&Long Path&-5.03e-03&-7.65e-01&2.67e+01\\
+Confident Neighbor&Pref. Attach.&Random Pod&2.61e-02&4.02e+00&3.74e-03\\
+Confident Neighbor&Small World&Pref. Attach.&4.49e-02&7.20e+00&7.03e-11\\
+Confident Neighbor&Small World&Long Path&3.99e-02&6.96e+00&3.80e-10\\
+Confident Neighbor&Small World&Random Pod&7.10e-02&1.18e+01&1.98e-28\\
+Confident Neighbor&Long Path&Pref. Attach.&5.03e-03&7.65e-01&2.67e+01\\
+Confident Neighbor&Long Path&Small World&-3.99e-02&-6.96e+00&3.80e-10\\
+Confident Neighbor&Long Path&Random Pod&3.11e-02&4.89e+00&7.03e-05\\
+Confident Neighbor&Random Pod&Pref. Attach.&-2.61e-02&-4.02e+00&3.74e-03\\
+Confident Neighbor&Random Pod&Small World&-7.10e-02&-1.18e+01&1.98e-28\\
+Confident Neighbor&Random Pod&Long Path&-3.11e-02&-4.89e+00&7.03e-05\\
+\hline
+Conform&Pref. Attach.&Small World&3.67e-02&6.85e+00&7.73e-10\\
+Conform&Pref. Attach.&Long Path&-1.73e-01&-2.40e+01&7.10e-99\\
+Conform&Pref. Attach.&Random Pod&-2.77e-01&-4.08e+01&6.29e-213\\
+Conform&Small World&Pref. Attach.&-3.67e-02&-6.85e+00&7.73e-10\\
+Conform&Small World&Long Path&-2.09e-01&-3.46e+01&3.46e-171\\
+Conform&Small World&Random Pod&-3.14e-01&-5.68e+01&1.63e-313\\
+Conform&Long Path&Pref. Attach.&1.73e-01&2.40e+01&7.10e-99\\
+Conform&Long Path&Small World&2.09e-01&3.46e+01&3.46e-171\\
+Conform&Long Path&Random Pod&-1.04e-01&-1.43e+01&8.55e-41\\
+Conform&Random Pod&Pref. Attach.&2.77e-01&4.08e+01&6.29e-213\\
+Conform&Random Pod&Small World&3.14e-01&5.68e+01&1.63e-313\\
+Conform&Random Pod&Long Path&1.04e-01&1.43e+01&8.55e-41\\
+    \hline
+    \end{tabular}
+    \caption{Two-tailed paired t-values across networks, within strategies.
+Results are for parallel individual learning.
+For all rows, dof=999 and $p^*$ is the Bonferroni-corrected p-value with m=60.}
+\end{table*}
+
+\begin{table*}[]
+    \label{tab:t-innet-parallel}
+    \centering
+    \begin{tabular}{l|ll|ccc}
+        Network & Strategy A & Strategy B & $\overline{B - A}$ & t & p* \\
+    \hline
+Pref. Attach.&Best Neighbor&Confident Neighbor&-3.78e-02&-5.94e+00&2.39e-07\\
+Pref. Attach.&Best Neighbor&Conform&2.00e-01&2.77e+01&3.28e-124\\
+Pref. Attach.&Confident Neighbor&Best Neighbor&3.78e-02&5.94e+00&2.39e-07\\
+Pref. Attach.&Confident Neighbor&Conform&2.38e-01&3.40e+01&1.67e-167\\
+Pref. Attach.&Conform&Best Neighbor&-2.00e-01&-2.77e+01&3.28e-124\\
+Pref. Attach.&Conform&Confident Neighbor&-2.38e-01&-3.40e+01&1.67e-167\\
+\hline
+Small World&Best Neighbor&Confident Neighbor&-1.52e-02&-2.74e+00&3.80e-01\\
+Small World&Best Neighbor&Conform&3.05e-01&5.30e+01&1.25e-290\\
+Small World&Confident Neighbor&Best Neighbor&1.52e-02&2.74e+00&3.80e-01\\
+Small World&Confident Neighbor&Conform&3.20e-01&5.89e+01&0.00e+00\\
+Small World&Conform&Best Neighbor&-3.05e-01&-5.30e+01&1.25e-290\\
+Small World&Conform&Confident Neighbor&-3.20e-01&-5.89e+01&0.00e+00\\
+\hline
+Long Path&Best Neighbor&Confident Neighbor&-3.88e-04&-6.64e-02&5.68e+01\\
+Long Path&Best Neighbor&Conform&7.02e-02&9.34e+00&3.64e-18\\
+Long Path&Confident Neighbor&Best Neighbor&3.88e-04&6.64e-02&5.68e+01\\
+Long Path&Confident Neighbor&Conform&7.06e-02&9.15e+00&1.83e-17\\
+Long Path&Conform&Best Neighbor&-7.02e-02&-9.34e+00&3.64e-18\\
+Long Path&Conform&Confident Neighbor&-7.06e-02&-9.15e+00&1.83e-17\\
+\hline
+Random Pod&Best Neighbor&Confident Neighbor&-3.59e-03&-5.46e-01&3.51e+01\\
+Random Pod&Best Neighbor&Conform&-6.84e-02&-9.54e+00&6.11e-19\\
+Random Pod&Confident Neighbor&Best Neighbor&3.59e-03&5.46e-01&3.51e+01\\
+Random Pod&Confident Neighbor&Conform&-6.48e-02&-8.96e+00&9.43e-17\\
+Random Pod&Conform&Best Neighbor&6.84e-02&9.54e+00&6.11e-19\\
+Random Pod&Conform&Confident Neighbor&6.48e-02&8.96e+00&9.43e-17\\
+    \hline
+    \end{tabular}
+    \caption{Two-tailed paired t-values across strategies, within networks.
+Results are for parallel individual learning.
+For all rows, dof=999 and $p^*$ is the Bonferroni-corrected p-value with m=60.}
+\end{table*}
 

### 2021-09-26 ch-experiment.tex
Add intro and part of experiment ch.

\section{Introduction}

\section{Experimental Design}
Overview

\subsection{Participant Recruitment and Incentives}
Participants will be recruited to discuss a policy issue.
The policy issue will be chosen to be relevant to many individuals,
and to be amenable to a predefined list of proposed solutions,
as well as to be relevant at the time the experiment is conducted.
Example policy questions include
``Where should the use of electric scooters be permitted?'' and
``Given the COVID-19 pandemic, how should schools in the US reopen in Fall 2021?.''
Multiple waves of deliberation may be held using different questions and participants,
depending factors such as participant attrition.

Participants will be recruited using a number of methods.
Participants will be recruited individually through posts on web forums
relevant to the policy question (e.g., reddit),
relevant facebook groups,
and relevant email lists.
Participant pools will also be utilized,
such as lists of previous participants who have expressed interest in future
studies,
and the UMSI Online Recruitment System for Economic Experiments.
Individual participants will also be recruited using promoted posts and
advertisements on Twitter, Facebook, etc.
Separately, participants may also be recruited through partnership with
community groups who wish to try network deliberation as part of an organizational
decision-making strategy.
The number of such collaborations will depend on the number of sufficiently-large
partner organizations that have decisions amenable to network deliberation
over the course of the experiment.

All participants will be paid incentives for participation in the deliberation
and completion of ranked-choice votes and pre/post-experiment surveys.
It is crucial that the incentive structure incentivize participation,
without incentivizing any particular deliberative strategy
(e.g., prefering a specific outcome, maintaining an original position, or
seeking agreement/disagreement).
Any such incentive structure could motivate participants to act based on their
preferences for incentive payments rather than their true preferences regarding
the policy question under consideration,
and disincentivize honest deliberation about the policy question.
As such, incentives will be paid as the same fixed amount to all participants.

\subsection{Experimental Protocol}

At the beginning of the experiment, participants will complete a short
demographic survey.
This survey will enable the identification of sample bias in age, gender, race,
and other demographic factors.

When participants are enrolled, they will be assigned to one of three treatment
groups: control, random-pod, long-path.
The control group will deliberate in a single large group.
The random-pod and long-path treatments will deliberate in a series of small
pods (network deliberation).
These treatments will vary only by the pod assignment method used:
either random-pod or long-path (see Chapter \ref{chap:abm}).

Deliberation will be divided into a predetermined number of {\em rounds}, each lasting
a fixed length of time.
Current plans are to use three stages of two days each,
although the specifics and timing of a policy quesiton may require altering these
numbers.
Before and after each stage, participants will be asked to rank the proposed
solutions to the policy question according to their individual preference.
During each round, participants will be shown a discussion prompt and will be
able to post and reply to other participants.
Example prompts are shown in Tabel \ref{tab:prompts}.
Participants will only be able to see or interact with posts from others in
their current pod.
(all control group participants will be able to interact with all others
at all times).
Participants will be able to view posts and comments that were visible to them
in previous rounds, but will no longer be able to interact with or reply to
those posts and comments.

\begin{table}
\center
\label{tab:prmpts}
\begin{tabular}{|p{0.3in}|p{3.6in}|}
\hline
Stage & Prompt \\
\hline
1 &
Which proposals do you prefer, and why?
If there are disagreements, try asking questions to understand the perspective
of others and to understand the causes of the disagreement.
\\
\hline
2 & In the previous round of discussion, what opinions and reasoning did you
observe in your group?
How much agreement was there?
Were there any disagreements or conflicts?
If so, what were the sources of conflict, and how might they be resolved?
\\
\hline
3 &
What seem to be the most popular opinions?
Do you agree with them?
Has your opinion changed over the course of the discussion?
\\
\hline
\end{tabular}
\end{table}

\subsection{Power Analysis}

\subsection{Quantifying Preferences, Conflict, and Agreement}

\section{Study Software}

\section{Pilot Studies}

### 2021-09-27 ch-wp-prod-perf.tex
Change terminology in ch2 from 'efficiency' to 'productivity'.

@@
 the collaboration's output,
-which we call {\em efficiency}.
+which we call {\em productivity}.
 The other measure is long-term:
@@
 but we find it fruitful to distinguish between the two.
-We find that Wikipedia exhibits an overall trade-off between performance and efficiency.
-However, some WikiProjects surpass others in both efficiency and performance,
+We find that Wikipedia exhibits an overall trade-off between performance and productivity.
+However, some WikiProjects surpass others in both productivity and performance,
 suggesting the existence of factors that correlate positively with both.
@@
 \setlength\itemsep{0pt}
-\item Despite an overall performance/efficiency trade-off,
-WikiProjects with low-degree coeditor networks tend to have both higher performance and higher efficiency;
+\item Despite an overall performance/productivity trade-off,
+WikiProjects with low-degree coeditor networks tend to have both higher performance and higher productivity;
 \item Short paths are associated with higher performance, consistent with a conformity-based learning strategy;
@@
 \item Our agent-based model
-shows that the efficiency and performance of collaborations can
+shows that the productivity and performance of collaborations can
 depend on network degree,
@@
 a few key themes emerge.
-Both the efficiency and the performance of a collaboration are important
+Both the productivity and the performance of a collaboration are important
 considerations and vary depending on both network structure and type of task
@@
 
-\subsection{Efficiency and Performance}
+\subsection{Productivity and Performance}
 
@@
 there are many ways to gauge their success.
-One possibility is {\em efficiency}:
+One possibility is {\em productivity}:
 how quickly they find a solution.
@@
 and field observations \cite{gentry_consensus_1982}
-all suggest a trade-off between efficiency and performance.
+all suggest a trade-off between productivity and performance.
 While common, this trade-off is not absolute,
 suggesting it is sometimes possible to simultaneously increase
-performance and efficiency.
-The identification of factors associated with both higher efficiency
+performance and productivity.
+The identification of factors associated with both higher productivity
 and higher performance has obvious practical importance.
 In this paper, we focus on how network structure
-relates to efficiency and performance within WikiProjects.
-
-For a WikiProject, efficiency quantifies how much participants can raise the
+relates to productivity and performance within WikiProjects.
+
+For a WikiProject, productivity quantifies how much participants can raise the
 assessed quality of an article for a fixed amount of work.
@@
 Different projects can have different standards and practices for assessing article quality,
-so the efficiency is not a measure of how quickly some objective measure of quality improves,
+so the productivity is not a measure of how quickly some objective measure of quality improves,
 but rather of how quickly the project participants can reach consensus on the improvements that
@@
 Because our definition relies on assessment transitions,
-we must define efficiency variables for
+we must define productivity variables for
 each of the project-level quality assessments: A, B, and C.
 For a particular grade $G$,
-we desire our definition of efficiency to meet the following conditions:
+we desire our definition of productivity to meet the following conditions:
 \begin{itemize}
@@
 \item{Strictly decreasing in the number of revisions (with transition count fixed);}
-\item{Independent of WikiProject size: not affected by adding an article having the same efficiency.}
+\item{Independent of WikiProject size: not affected by adding an article having the same productivity.}
 \end{itemize}
 
-We now define an efficiency measure which meets the above criteria.
+We now define an productivity measure which meets the above criteria.
 Let $T(W,G)$ be the set of article assessment transitions from below grade $G$
@@
 and let $g(t)$ be the number of grade levels crossed bt $t$.
-We quantify the efficiency $E(W,G)$ as the inverse of the mean number of revisions
+We quantify the productivity $E(W,G)$ as the inverse of the mean number of revisions
 per transition:
@@
 Wikipedia---the pattern of who interacts with whom---relates to
-efficiency and performance.
+productivity and performance.
 There are several types of interactions we could focus on,
@@
 Degree and characteristic path length have been shown to correlate with
-performance and efficiency in some social learning settings
+performance and productivity in some social learning settings
 \cite{golub_naive_2010,mason_propagation_2008,grim_scientific_2013},
@@
 Networks with longer paths require more interactions for information to propagate,
-which has been shown to reduce efficiency in some settings
+which has been shown to reduce productivity in some settings
 \cite{mason_propagation_2008,barkoczi_social_2016}.
@@
 \caption{
-Histograms of WikiProject efficiency and performance.
+Histograms of WikiProject productivity and performance.
 Both measures are highly right-skewed, but form unimodal distributions
@@
 
-We find that both efficiency and performance are highly right-skewed,
+We find that both productivity and performance are highly right-skewed,
 with a small number of projects having values much higher than the average.
-After log-transforming the values, both the efficiency and the performance have
+After log-transforming the values, both the productivity and the performance have
 a unimodal distribution with low skew (see Figure \ref{fig:eff-perf-hist}).
-Our findings confirm the trade-off between performance and efficiency
+Our findings confirm the trade-off between performance and productivity
 observed in many other settings (Figure \ref{fig:eff-perf}).
 However, when looking at specific projects, some are higher in both performance
-and efficiency,
+and productivity,
 suggesting the existence of factors which correlate positively with both.
@@
 \caption{
-WikiProject performance is anticorrelated with B-level efficiency,
+WikiProject performance is anticorrelated with B-level productivity,
 with Pearson r of -0.12.
 Results are similar for other grade levels.
-On average, highly efficient WikiProjects are under-performing,
+On average, highly productive WikiProjects are under-performing,
 but when looking at specific WikiProjects,
-some are higher than others in both performance and efficiency.
+some are higher than others in both performance and productivity.
 \label{fig:eff-perf}
@@
 Mean path length$^\dagger$&-0.33$^{***}$&-0.09&-0.05&-0.09\\
-C-Efficiency$^\dagger$&-0.08$^{*}$&\+---&\+---&\+---\\
+C-productivity$^\dagger$&-0.08$^{*}$&\+---&\+---&\+---\\
 Connected frac.&0.01&0.09$^{*}$&0.15$^{***}$&0.06\\
@@
 
-To study the relationship between network structure, efficiency, and performance,
-we model the performance and efficiency of WikiProjects using ordinary least-squares linear regression.
+To study the relationship between network structure, productivity, and performance,
+we model the performance and productivity of WikiProjects using ordinary least-squares linear regression.
 Each WikiProject is a single observation.
@@
 \setlength\itemsep{0pt}
-\item[C-efficiency]
+\item[C-productivity]
 (performance only).
@@
 
-We see that B-efficiency and C-efficiency have very similar models, but that A-efficiency behaves
+We see that B-productivity and C-productivity have very similar models, but that A-productivity behaves
 differently in its dependence on degree skewness and connectivity.
-The different behavior of A-efficiency is likely explained by the
+The different behavior of A-productivity is likely explained by the
 observation that the A-Class quality is infrequently used in practice.
@@
 
-The negative dependence of performance on C-efficiency suggests there is generally a trade-off between
-performance and efficiency.
-However, low degree is correlated with both higher efficiency and higher performance,
+The negative dependence of performance on C-productivity suggests there is generally a trade-off between
+performance and productivity.
+However, low degree is correlated with both higher productivity and higher performance,
 suggesting that it is possible to improve both simultaneously.
@@
 
-We also observe that high degree skewness is correlated with lower performance and lower A-efficiency,
-suggesting that articles in projects with decentralized coeditor networks reach featured or good status
-more efficiently, and reach higher quality ratings in general.
+We also observe that high degree skewness is correlated with lower performance and lower A-productivity,
+suggesting that projects with decentralized coeditor networks produce featured or good status
+articles more quickly, and reach higher quality ratings in general.
 
@@
 we use a simple agent-based model of collaboration to better understand
-the relationship between node degree, efficiency, and performance.
+the relationship between node degree, productivity, and performance.
 Numerical models allow us to determine the effect of changing a single
@@
 only with their neighbors.
-We define efficiency and performance in terms of the solution values for each time step
+We define productivity and performance in terms of the solution values for each time step
 (averaged over many trials).
 We define the performance to be the mean solution value after the process has converged,
-while the efficiency is the reciprocal of the number of steps required to converge.
+while the productivity is the reciprocal of the number of steps required to converge.
 We measure the time to convergence as the number of steps required to reach 99\% of the maximum
@@
 \begin{tabular}{lll}
-Strategy & Performance & Efficiency\\
+Strategy & Performance & productivity\\
 \hline
@@
 \caption{
-Simulated Performance and Efficiency.
+Simulated Performance and productivity.
 Results shown for 100 trials with P[rewire] = 0.
 Strategies are defined in Table \ref{tab:strat}.
-Local strategies are less efficient than their non-local counterparts.
+Local strategies are less productive than their non-local counterparts.
 Local best-neighbor out-performs global,
 while local conformity is the worst performer in all cases.
-The local majority strategy is both most efficient and most performant.
+The local majority strategy is both most productive and most performant.
 \label{tab:sim-eff-perf}
@@
 \caption{
-Efficiency and Performance of social learning strategies vs. mean network degree.
+productivity and Performance of social learning strategies vs. mean network degree.
 Each point represents a single trial of 300 iterations.
@@
 decreased performance at high degree,
-with no significant change in efficiency.
-Local conformity shows decreased performance and increased efficiency at high
+with no significant change in productivity.
+Local conformity shows decreased performance and increased productivity at high
 degree.
 Local majority shows the opposite behavior:
-increased performance and decreased efficiency at high degree,
-with the efficiency showing the largest effect size of all strategies.
+increased performance and decreased productivity at high degree,
+with the productivity showing the largest effect size of all strategies.
 \label{fig:deg-eff-perf}
@@
 Each curve represents an average over 100 trials, each with 250 agents.
-The mean performance and efficiency are reported in Table \ref{tab:sim-eff-perf}.
-For all rewiring values, local strategies are less efficient and more performant
+The mean performance and productivity are reported in Table \ref{tab:sim-eff-perf}.
+For all rewiring values, local strategies are less productive and more performant
 than their non-local counterparts.
@@
 Local conformity performs notably worse than all others.
-Local majority is both more efficient and more performant than others,
+Local majority is both more productive and more performant than others,
 with its performance increasing with higher rewiring.
@@
 at least in a simple collaboration model,
-performance and efficiency can be simultaneously increased.
-Furthermore, performance and efficiency are potentially affected by both the choice of
+performance and productivity can be simultaneously increased.
+Furthermore, performance and productivity are potentially affected by both the choice of
 learning strategy and the average degree of the agents' social network.
 
-The effects of degree on performance and efficiency are shown in Figure \ref{fig:deg-eff-perf} and
+The effects of degree on performance and productivity are shown in Figure \ref{fig:deg-eff-perf} and
 Table \ref{tab:sim-perf}.
 For non-local versions of both conformity and best-neighbor strategy,
-there is no significant effect of degree on performance or efficiency.
+there is no significant effect of degree on performance or productivity.
 The local best-neighbor strategy shows reduced performance with increasing degree,
-but no change in efficiency.
+but no change in productivity.
 Local conformity and local majority show opposite behavior as degree increases:
-with local conformity gaining efficiency at the expense of performance,
-while local majority increases in performance and decreases in efficiency.
-The largest effect size is achieved for efficiency in the local majority simulation,
-which is consistent with the efficiency behavior observed in WikiProjects.
+with local conformity gaining productivity at the expense of performance,
+while local majority increases in performance and decreases in productivity.
+The largest effect size is achieved for productivity in the local majority simulation,
+which is consistent with the productivity behavior observed in WikiProjects.
 However, the performance behavior for local majority is opposite that observed on
@@
 These agent-based models confirm that network degree has the potential to influence
-the performance and efficiency of collaborations.
+the performance and productivity of collaborations.
 Furthermore, this influence can be drastically different depending on the strategies
@@
 We observe several results consistent with previous work:
-a trade-off between performance and efficiency
+a trade-off between performance and productivity
 \cite{mason_propagation_2008,grim_scientific_2013},
@@
 our findings suggest degree distribution may be just as, or more, important.
-The association of low degree with both high performance and high efficiency is compelling,
-as it sidesteps the usual trade-off between performance and efficiency.
+The association of low degree with both high performance and high productivity is compelling,
+as it sidesteps the usual trade-off between performance and productivity.
 In low-degree networks, agents have more repeated interactions with smaller groups of collaborators, suggesting that small team sizes could be beneficial for large collaborations.
@@
 Our agent-based models offer a several insights.
-We observe degree-dependent performance and efficiency
+We observe degree-dependent performance and productivity
 for both local conformity and local majority strategies.
@@
 designed to satisfy several properties found in real-world collaborations,
-shows the strongest effects on performance and efficiency as network degree changes.
+shows the strongest effects on performance and productivity as network degree changes.
 For the local majority strategy,
-the relationship between degree and efficiency is consistent with our
+the relationship between degree and productivity is consistent with our
 empirical observations on Wikipedia,
-suggesting one possible mechanism underlying that efficiency dependence.
+suggesting one possible mechanism underlying that productivity dependence.
 However, the performance dependence of this strategy
@@
 Our work suggests several directions for future work.
-Is the correlation between network structure, performance, and efficiency causal?
+Is the correlation between network structure, performance, and productivity causal?
 A time-dependent analysis of our data could offer insight.
@@
 Does varying degree independently of path length influence
-performance and efficiency in a controlled lab setting?
+performance and productivity in a controlled lab setting?
 
@@
 In this paper, we have described the relationship between the structural properties of WikiProject
-coeditor networks, their performance, and their efficiency.
-As in other studies, we see a trade-off between performance and efficiency.
+coeditor networks, their performance, and their productivity.
+As in other studies, we see a trade-off between performance and productivity.
 However, 
 some properties, such as low degree, are associated with
-both higher performance and higher efficiency.
+both higher performance and higher productivity.
 We also find that the correlations between path length and performance are consistent with a conformity-based
@@
 We have also proposed a novel local majority learning strategy that is more realistic,
-more efficient, and higher performance than existing strategies.
+more productive, and higher performance than existing strategies.
 While most previous social learning simulations focus on path length,
-we observe degree-dependent performance and efficiency
+we observe degree-dependent performance and productivity
 in both the local majority strategy

### 2021-09-27 ch-experiment.tex
Add participant number analysis to ch4.

@@
 \subsection{Participant Recruitment and Incentives}
-Participants will be recruited to discuss a policy issue.
+Participants will be recruited to discuss a policy issue
+using a custom-built psudeonymous online platform.
 The policy issue will be chosen to be relevant to many individuals,
@@
 
+The minimum number of participants required is determined by a number of factors.
+First, the long-path pod assignment method results in a minimum number of pods,
+determined by the chosen parameters.
+Specifically, the pod assignments for each round beyond the first are determined
+by a unique prime number,
+and the number of pods is a multiple of that prime.
+For $T=3$ rounds, two primes are necessary and choosing the lowest two (2, 3)
+yields the lowest minimum number of pods: 3.
+For all pods to have at least 4 members, the minimum number of participants in
+the long-path treatment is 12.
+
+The two network deliberation treatments must also have meaningfully different
+structural properties,
+which also necessitates a minimum number of participants.
+Structural differences between these two networks become more pronounced
+with a greater number of participants.
+While properties such as the broadcast time (see Chapter \ref{chap:abm})
+and average geodesic length can be used to compare the structural efficiency of
+two networks,
+they encounter a problem for these particular networks
+when there is a small number of rounds:
+for many pairs of individuals, no path will exist.
+In practical terms,
+an idea proposed by one participant may not have a plausible
+path to reach some of the other participants by the end of the deliberation,
+even if that idea is repeatedly shared by all who encounter it.
+As an alternative, a form of k-connectivty can be used to measure
+structural efficiency.
+The k-connectivity is simply the fraction of possible paths that exist.
+In other words, if each participant broadcasts a message, and that message is
+repeated by all who encounter it, what is the average fraction of participants
+a broadcast will reach?
+Figure \ref{fig:kcon} shows the k-connectivity as a function of number of
+participants for both the long-path and random-pod deliberation networks.
+While there is no clear threshold for the necessary number of participants,
+choosing a k-connectivity of $0.5$ (half of possible paths exist) seems a
+reasonable heuristic.
+For the chosen parameters, 35 or more participants are necessary to produce
+a random-pod network with $>0.5$ k-connectivity and
+a long-path network with $<0.5$ k-connectivity.
+
+Combining the above considerations,
+an absolute minimum of 12 participants per treatment is necessary to conduct
+a wave of the experiment,
+and a minimum of about 35 participants per treatment group is necessary to
+observe differences between the network deliberation treatment conditions.
+
+\begin{figure}
+\label{fig:kcon}
+\includegraphics[width=5in]{fig/Experiment/fig-kcon.eps}
+\caption{The k-connectivity of long-path and random-pod network deliberation
+networks for $T=3$ stages and a pod size of $M=5$.}
+\end{figure}
+
 \subsection{Experimental Protocol}
@@
 Current plans are to use three stages of two days each,
-although the specifics and timing of a policy quesiton may require altering these
-numbers.
+although these numbers may need to be changed
+to ensure time timeline of the deliberation is consistent with any constraints
+imposed by specific policy questions.
 Before and after each stage, participants will be asked to rank the proposed
@@
 able to post and reply to other participants.
-Example prompts are shown in Tabel \ref{tab:prompts}.
+Example prompts are shown in Table \ref{tab:prompts}.
 Participants will only be able to see or interact with posts from others in
@@
 \center
-\label{tab:prmpts}
+\label{tab:prompts}
 \begin{tabular}{|p{0.3in}|p{3.6in}|}
@@
 
-\subsection{Power Analysis}
+A subset of participants will be randomly selected to participate in
+60-minute semi-structured interviews about their experience.
+These interviewed will be qualitatively coded using grounded theory.
+The interviews will provide information on the subjective experience of
+participants, as well as help identify common themes that may provide
+insight into the strengths and weaknesses of network deliberation.
+
+In summary, the experiment will result in several data sets:
+\begin{itemize}
+\item The ranked choice preferences of each participant before and after each round of deliberation.
+\item The text of the online deliberation.
+\item Post-deliberation survey responses.
+\item Qualitatively-coded interview transcripts.
+\end{itemize}
 
@@
 \section{Pilot Studies}
+Several pilot studies have already been conducted:
+one in a synchronous lab setting, and two using the asynchronous online
+software.
+In addition to providing an opportunity to test software,
+these pilot studies have provided crucial information for developing a
+feasible experimental design.
+
+The initial pilot study was conducted on December 21, 2018,
+before the custom online platform was developed.
+In this pilot, NUMBER participants engaged in a synchronous 60-minute deliberation.
+Participants sat at individual computer workstations, separated by dividers.
+Participants discussed the following question:
+``Where should the use of electric scooters be permitted?''
+Participants were placed in a single treatment group, and divided into 4
+pods of size 4--5.
+These pods were reassigned using the random-pod method for a total of three
+rounds.
+The deliberation took place in chat rooms running the free and open-source
+OTree online experiment software.
+
+The second pilot took place after the initial development of the custom online
+platform, from August 24, 2020 through August 32, 2020.
+
+The third pilot took place from August 23, 2021 through August 29, 2021.

### 2021-09-28 ch-experiment.tex
Revisions to experimental design.

@@
 \section{Experimental Design}
-Overview
+
+This section describes the experimental design of the proposed field experiment.
+In this experiment, participants will use a pseudonymous online platform
+to deliberate on a policy issue.
+The deliberation will take place asynchronously, over a period of several days.
+At several times throughout the deliberation,
+participants will complete a ranked-choice poll regarding the policy issue,
+allowing the evolution of their preferences to be tracked over the course of the
+experiment.
+Depending on opportunity and necessity, multiple waves of deliberation may
+be conducted, each with a distinct policy issue and set of participants.
+
+\subsection{Deliberative Topic}
+The policy issue chosen as the topic of deliberation is a crucial component
+of the experiment.
+The topic will be chosen carefully according to several criteria.
+The chosen policy issue must be:
+\begin{itemize}
+\item Relevant to the experimental population;
+\item Amenable to participants changing their preferences based on new information or reasoning;
+\item Amenable to a predefined list of proposed solutions;
+\item Sufficiently complex to have three or more proposed solutions;
+\item Timely, but unlikely to be influenced by current events during the deliberation.
+\end{itemize}
+Examples of policy issues meeting these criteria include:
+\begin{itemize}
+\item ``Where should the use of electric scooters be permitted?''
+\item ``Given the COVID-19 pandemic, how should schools in the US reopen in Fall 2021?''
+\end{itemize}
 
 \subsection{Participant Recruitment and Incentives}
-Participants will be recruited to discuss a policy issue
-using a custom-built psudeonymous online platform.
-The policy issue will be chosen to be relevant to many individuals,
-and to be amenable to a predefined list of proposed solutions,
-as well as to be relevant at the time the experiment is conducted.
-Example policy questions include
-``Where should the use of electric scooters be permitted?'' and
-``Given the COVID-19 pandemic, how should schools in the US reopen in Fall 2021?.''
-Multiple waves of deliberation may be held using different questions and participants,
-depending factors such as participant attrition.
-
-Participants will be recruited using a number of methods.
+Participants will be recruited using several methods.
 Participants will be recruited individually through posts on web forums
-relevant to the policy question (e.g., reddit),
-relevant facebook groups,
-and relevant email lists.
+relevant to the policy question (e.g., reddit, facebook groups)
+and to relevant email lists.
 Participant pools will also be utilized,
-such as lists of previous participants who have expressed interest in future
-studies,
+including lists of previous participants
+who have expressed interest in future studies
 and the UMSI Online Recruitment System for Economic Experiments.
@@
 decision-making strategy.
-The number of such collaborations will depend on the number of sufficiently-large
+The number of such collaborations will depend on the number of interested
 partner organizations that have decisions amenable to network deliberation
@@
 
-All participants will be paid incentives for participation in the deliberation
-and completion of ranked-choice votes and pre/post-experiment surveys.
-It is crucial that the incentive structure incentivize participation,
-without incentivizing any particular deliberative strategy
-(e.g., prefering a specific outcome, maintaining an original position, or
-seeking agreement/disagreement).
-Any such incentive structure could motivate participants to act based on their
-preferences for incentive payments rather than their true preferences regarding
-the policy question under consideration,
-and disincentivize honest deliberation about the policy question.
-As such, incentives will be paid as the same fixed amount to all participants.
+In order to incentivize participation and to compensate participants fairly
+for their time,
+participants who complete the deliberation and pre/post-experiment surveys
+will be compensated financially.
+Participants will be paid a predetermined fixed amount in order to
+minimize the effect of the compensation on their deliberative behavior and
+expressed preferences.
 
+\subsection{Experimental Protocol}
+
+Before the deliberation period begins,
+participants will complete a short demographic survey.
+This survey will enable the identification of sample bias in age, gender, race,
+and other demographic factors.
+
+Participants will be randomly assigned to one of three treatment/control groups
+at time of enrollment:
+\begin{description}
+\item[Control]{
+The control group will engage in conventional single large group deliberation.
+Posts and comments made by participants in the control group will be visible
+to all others in the control group for the entire duration of the deliberation.}
+\item[Network Deliberation (Efficient)]{
+Participants in this treatment group will be divided into small
+(4--5 participant) pods.
+Posts and comments made by participants in this treatment group will only be
+visible to others in their current pod.
+All participants will be assigned to a new pod at the beginning of each round
+of deliberation using the random-pod assignment method,
+producing a structurally efficient communication network.}
+\item[Network Deliberation (Inefficient)]{
+Participants in this treatment group will be divided into small
+(4--5 participant) pods.
+Posts and comments made by participants in this treatment group will only be
+visible to others in their current pod.
+All participants will be assigned to a new pod at the beginning of each round
+of deliberation using the long-path assignment method,
+producing a structurally inefficient communication network.}
+\end{description}
+
+The deliberation period will be divided into a predetermined number of {\em rounds},
+each lasting a fixed length of time.
+Present plans are to divide the deliberation period into three stages of two days
+each.
+However, these numbers may be adjusted to ensure time timeline of the
+deliberation is consistent with any constraints imposed by specific policy issues.
+Before and after each stage, participants will be asked to rank the proposed
+solutions to the policy question according to their individual preference.
+During each round, participants will be shown a discussion prompt and will be
+able to post a response to that prompt.
+Example prompts are shown in Table \ref{tab:prompts}.
+Participants will be able to view and comment on the posts of other participants
+assigned to the same pod.
+Participants will also be able to view posts and comments that were visible to them
+in previous rounds, but will no longer be able to interact with or reply to
+those posts and comments.
+
+\begin{table}
+\center
+\label{tab:prompts}
+\begin{tabular}{|p{0.3in}|p{3.6in}|}
+\hline
+Stage & Prompt \\
+\hline
+1 &
+Which proposals do you prefer, and why?
+If there are disagreements, try asking questions to understand the perspective
+of others and to understand the causes of the disagreement.
+\\
+\hline
+2 & In the previous round of discussion, what opinions and reasoning did you
+observe in your group?
+How much agreement was there?
+Were there any disagreements or conflicts?
+If so, what were the sources of conflict, and how might they be resolved?
+\\
+\hline
+3 &
+What seem to be the most popular opinions?
+Do you agree with them?
+Has your opinion changed over the course of the discussion?
+\\
+\hline
+\end{tabular}
+\caption{Example prompts for each round of deliberation.}
+\end{table}
+
+\section{Qualitative Analysis}
+In addition to quantitative data,
+qualitative data will also be collected in conjunction with
+the deliberation experiment.
+These data will be used to corroborate the results of quantitative
+analysis and to achieve a richer understanding of individual participants'
+experience of network deliberation,
+as well as to as well as identify common themes that may provide
+insight into the strengths and weaknesses of network deliberation.
+
+The text of participants' posts and comments throughout the deliberation
+will be kept for analysis.
+This text can potentially be used to identify deliberative behaviors that
+lead to changes in preferences.
+Similarly, this text can be used to track the diffusion of specific keywords,
+arguments, and influence through the communication network.
+
+A subset of participants will be randomly selected to participate in
+60-minute semi-structured interviews about their experience.
+The interview script can be found in Appendix \ref{apx:interview}.
+These interviewed will be transcribed and qualitatively coded using a
+grounded theory approach.
+The initial code book will be developed by open-coding pilot study transcripts
+and an initial subset of interview transcripts.
+
+\subsection{Network Topologies}
 The minimum number of participants required is determined by a number of factors.
@@
 
-\subsection{Experimental Protocol}
-
-At the beginning of the experiment, participants will complete a short
-demographic survey.
-This survey will enable the identification of sample bias in age, gender, race,
-and other demographic factors.
-
-When participants are enrolled, they will be assigned to one of three treatment
-groups: control, random-pod, long-path.
-The control group will deliberate in a single large group.
-The random-pod and long-path treatments will deliberate in a series of small
-pods (network deliberation).
-These treatments will vary only by the pod assignment method used:
-either random-pod or long-path (see Chapter \ref{chap:abm}).
-
-Deliberation will be divided into a predetermined number of {\em rounds}, each lasting
-a fixed length of time.
-Current plans are to use three stages of two days each,
-although these numbers may need to be changed
-to ensure time timeline of the deliberation is consistent with any constraints
-imposed by specific policy questions.
-Before and after each stage, participants will be asked to rank the proposed
-solutions to the policy question according to their individual preference.
-During each round, participants will be shown a discussion prompt and will be
-able to post and reply to other participants.
-Example prompts are shown in Table \ref{tab:prompts}.
-Participants will only be able to see or interact with posts from others in
-their current pod.
-(all control group participants will be able to interact with all others
-at all times).
-Participants will be able to view posts and comments that were visible to them
-in previous rounds, but will no longer be able to interact with or reply to
-those posts and comments.
-
-\begin{table}
-\center
-\label{tab:prompts}
-\begin{tabular}{|p{0.3in}|p{3.6in}|}
-\hline
-Stage & Prompt \\
-\hline
-1 &
-Which proposals do you prefer, and why?
-If there are disagreements, try asking questions to understand the perspective
-of others and to understand the causes of the disagreement.
-\\
-\hline
-2 & In the previous round of discussion, what opinions and reasoning did you
-observe in your group?
-How much agreement was there?
-Were there any disagreements or conflicts?
-If so, what were the sources of conflict, and how might they be resolved?
-\\
-\hline
-3 &
-What seem to be the most popular opinions?
-Do you agree with them?
-Has your opinion changed over the course of the discussion?
-\\
-\hline
-\end{tabular}
-\end{table}
-
-A subset of participants will be randomly selected to participate in
-60-minute semi-structured interviews about their experience.
-These interviewed will be qualitatively coded using grounded theory.
-The interviews will provide information on the subjective experience of
-participants, as well as help identify common themes that may provide
-insight into the strengths and weaknesses of network deliberation.
-
-In summary, the experiment will result in several data sets:
-\begin{itemize}
-\item The ranked choice preferences of each participant before and after each round of deliberation.
-\item The text of the online deliberation.
-\item Post-deliberation survey responses.
-\item Qualitatively-coded interview transcripts.
-\end{itemize}
-
 \subsection{Quantifying Preferences, Conflict, and Agreement}

### 2021-09-28 ch-experiment.tex
Add quantitative analysis to ch4.

@@
 \section{Introduction}
+
+\section{Study Software}
+
+\section{Pilot Studies}
+Several pilot studies have already been conducted:
+one in a synchronous lab setting, and two using the asynchronous online
+software.
+In addition to providing an opportunity to test software,
+these pilot studies have provided crucial information for developing a
+feasible experimental design.
+
+The initial pilot study was conducted on December 21, 2018,
+before the custom online platform was developed.
+In this pilot, NUMBER participants engaged in a synchronous 60-minute deliberation.
+Participants sat at individual computer workstations, separated by dividers.
+Participants discussed the following question:
+``Where should the use of electric scooters be permitted?''
+Participants were placed in a single treatment group, and divided into 4
+pods of size 4--5.
+These pods were reassigned using the random-pod method for a total of three
+rounds.
+The deliberation took place in chat rooms running the free and open-source
+OTree online experiment software.
+
+The second pilot took place after the initial development of the custom online
+platform, from August 24, 2020 through August 32, 2020.
+
+The third pilot took place from August 23, 2021 through August 29, 2021.
 
@@
 
-\section{Qualitative Analysis}
-In addition to quantitative data,
-qualitative data will also be collected in conjunction with
-the deliberation experiment.
-These data will be used to corroborate the results of quantitative
-analysis and to achieve a richer understanding of individual participants'
-experience of network deliberation,
-as well as to as well as identify common themes that may provide
-insight into the strengths and weaknesses of network deliberation.
+\section{Quantitative Analysis}
 
-The text of participants' posts and comments throughout the deliberation
-will be kept for analysis.
-This text can potentially be used to identify deliberative behaviors that
-lead to changes in preferences.
-Similarly, this text can be used to track the diffusion of specific keywords,
-arguments, and influence through the communication network.
+The primary quantitative data will be the results of the ranked-choice
+polls conducted before and after each stage of deliberation.
+These results represent individual participants' stated preferences at multiple
+time points.
+Standard tools from social choice theory can be applied to such rank-ordered
+preferences \cite{arrow_social_2012, brandt_computational_2012}.
+At each time, these preferences can be aggregated into a {\em preference profile}
+representing the collective preferences of all participants in a treatment group.
+Preference profiles can be used to aggregate individual preferences into a single
+ranking via a {\em social welfare function}.
+Similarly, preference profiles can be used to select a single winner via a
+{\em social choice function}.
+These social choice and social welfare functions represent voting methods.
+Different voting methods can potentially yield different results for the same
+preference profile.
+However, the more similar preferences are across individual participants,
+the more likely it is that different voting methods will yield the same winner
+or social ranking.
+For example, when one proposal beats all others in pairwise contests, it is known as the
+{\em Condorcet winner}.
+Many voting methods yield the Condorcet winner when it exists, but the existence
+of a Condorcet winner is not guaranteed.
 
-A subset of participants will be randomly selected to participate in
-60-minute semi-structured interviews about their experience.
-The interview script can be found in Appendix \ref{apx:interview}.
-These interviewed will be transcribed and qualitatively coded using a
-grounded theory approach.
-The initial code book will be developed by open-coding pilot study transcripts
-and an initial subset of interview transcripts.
+For the purposes of this dissertation, the preference profiles will be used
+to quantify agreement and conflict over the course of deliberation.
+Within social choice theory, agreement can be defined by a {\em consensus class},
+a set of preference profiles meeting one of several consensus criteria
+\cite{elkind_rationalizations_2016}.
+Examples include strong unanimity (of rank orders),
+unanimity (of winners),
+majority (existence of majority),
+Condorcet (existence of Condorcet winner),
+and transitivity (of social preference).
+On the most restrictive end, in strong unanimity and unanimity,
+all decision-makers prefer the same alternative.
+In contrast, transitivity only requires that the social preferences induced by
+individual preferences are transitive.
+Distance-rationalizable voting systems are defined by projecting the preference
+profile onto the nearest consensus profile according to a suitable distance.
+For example, Dodgson's method \cite{dodgeson_method_1876, brandt_computational_2012}
+defines a distance based on swapping adjacent entries in individual preference profiles.
+However, for most voting systems,
+finding the consensus profile that minimizes distance is NP-Hard
+\cite{elkind_rationalizations_2016}.
+Social choice theory also provides measures of distances between two individual
+preference rankings.
+These measures can be extended to create measures of agreement for entire
+preference profiles that can be calculated efficiently.
 
-\subsection{Network Topologies}
+One possibility is to measure the distance from strong unanimity.
+Two such methods will be used to analyze the results of the deliberation experiment:
+the mean Kendal tau metric and the mean Spearman correlation.
+The Kendal tau metric is the number fraction of pairwise contests that have
+different results between two profiles \cite{kendall_new_1938}}.
+By averaging this measure over all pairs of members, an overall measure of agreement can be calculated for the group.
+This measure does not incorporate information about win/loss margins,
+only whether an alternative wins or loses.
+To quantify the win/loss margin, the mean Spearman correlation will be used.
+The Spearman correlation is the linear coefficient of correlation between the
+rank orders of alternatives for two voters \cite{spearman_proof_1904}.
+As with Kendall tau, the Spearman correlation can be averaged over all pairs of
+members to determine a measure of agreement for the entire group.
+It must be noted that representing agreement with a single number will inevitably
+ignore important information.
+Neither of these measures distinguishes a small number of voters disagreeing
+on many comparisons from a large number of voters disagreeing on a few.
+The Spearman correlation is also unable to distinguish these from situations in
+which a small number of voters disagree by a very large margin.
+
+While unanimity might be preferred,
+it may sometimes be more realistic to focus on transitivity.
+This consensus class requires only that the social preferences between pairs
+of alternatives are transitive.
+Such preference profiles guarantee that a Condorcet winner will exist,
+bringing many different voting methods into agreement on which should win.
+
+The deviation from transitivity will be qunatified by counting the number
+of ranked-pair violations.
+When performing a vote by the Tideman ranked pair system,
+the relative order of two alternatives in the social preference is
+determined by the winner of a pairwise vote,
+except in the special case that contests with a higher margin have already
+constrained their order due to transitivity.
+In the latter case, the winner of the pairwise contest may have a lower
+position in the social preference.
+If this situation occurs, the social preference is intransitive and there may
+not be a Condorcet winner.
+Each such pair represents a contest violating the social ranking,
+so the fraction of such pairs
+can be used to quantify the strength of intransitivity in the group’s preferences.
+This measure has the property that a Condorcet winner is guaranteed when the value is 0.
+
+\section{Power Analysis}
 The minimum number of participants required is determined by a number of factors.
@@
 
-\subsection{Quantifying Preferences, Conflict, and Agreement}
+\section{Qualitative Analysis}
+In addition to quantitative data,
+qualitative data will also be collected in conjunction with
+the deliberation experiment.
+These data will be used to corroborate the results of quantitative
+analysis and to achieve a richer understanding of individual participants'
+experience of network deliberation,
+as well as to as well as identify common themes that may provide
+insight into the strengths and weaknesses of network deliberation.
 
-\section{Study Software}
+The text of participants' posts and comments throughout the deliberation
+will be kept for analysis.
+This text can potentially be used to identify deliberative behaviors that
+lead to changes in preferences.
+Similarly, this text can be used to track the diffusion of specific keywords,
+arguments, and influence through the communication network.
 
-\section{Pilot Studies}
-Several pilot studies have already been conducted:
-one in a synchronous lab setting, and two using the asynchronous online
-software.
-In addition to providing an opportunity to test software,
-these pilot studies have provided crucial information for developing a
-feasible experimental design.
-
-The initial pilot study was conducted on December 21, 2018,
-before the custom online platform was developed.
-In this pilot, NUMBER participants engaged in a synchronous 60-minute deliberation.
-Participants sat at individual computer workstations, separated by dividers.
-Participants discussed the following question:
-``Where should the use of electric scooters be permitted?''
-Participants were placed in a single treatment group, and divided into 4
-pods of size 4--5.
-These pods were reassigned using the random-pod method for a total of three
-rounds.
-The deliberation took place in chat rooms running the free and open-source
-OTree online experiment software.
-
-The second pilot took place after the initial development of the custom online
-platform, from August 24, 2020 through August 32, 2020.
-
-The third pilot took place from August 23, 2021 through August 29, 2021.
+A subset of participants will be randomly selected to participate in
+60-minute semi-structured interviews about their experience.
+The interview script can be found in Appendix \ref{apx:interview}.
+These interviewed will be transcribed and qualitatively coded using a
+grounded theory approach.
+The initial code book will be developed by open-coding pilot study transcripts
+and an initial subset of interview transcripts.

### 2021-09-28 ch-experiment.tex
Add pilots.

@@
 before the custom online platform was developed.
-In this pilot, NUMBER participants engaged in a synchronous 60-minute deliberation.
+In this pilot, 18 participants engaged in a synchronous 60-minute deliberation.
 Participants sat at individual computer workstations, separated by dividers.
@@
 The deliberation took place in chat rooms running the free and open-source
-OTree online experiment software.
+OTree online experiment software \cite{chen_oTree_2016}.
 
@@
 platform, from August 24, 2020 through August 32, 2020.
+In this pilot, 7 participants engaged in an asynchronous deliberation over the
+course of 7 days.
+Participants were assigned to a single treatment group using network deliberation
+with random-pod assignment.
+Pods were size 4, and there were 3 rounds (2, 2, and 3 days).
+Participants were researchers considering a week-long collaboration and
+deliberated on the following question:
+``What is your current preference for week of science topics?''
 
 The third pilot took place from August 23, 2021 through August 29, 2021.
+Participants deliberated on the following question:
+``How should schools in the US reopen for fall 2021?''
+Participants were recruited on reddit and Twitter.
+18 participants were enrolled (excluding several identified as fake accounts).
+These participants were divided between a conventional deliberation control
+group and a random-pod treatment group.
+The random-pod treatment group had two pods of size 5.
+Deliberation was divided into 3 rounds (2, 2, and 3 days).
+
+In these pilots, participants showed quantitative and qualitative evidence of
+changing their preferences over the course of deliberation.
+Here is an example of one such interaction:
+\begin{quote}
+{\bfseries Aspen Sunflower:} Even though my \#1 choice includes mandatory vaccinations, I know it's not possible due to school starting now AND MDHS can't mandate it yet.
+Based on previous stage, I think most seemed to want vaccinations but also wanted options.
+One option doesn't fit all students.
+
+{\bfseries Aspen Poppy:} Major gap in my understanding here!
+The schools do not mandate vaccines for students, the state does.
+For the 12-17 age group, the vaccine is authorized under emergency use, full FDA approval is pending.
+This strengthens my desire to see hybrid learning with smaller in person groups and vaccinated staff.
+\end{quote}
+
+While the pilots did not have enough participants to identify any differences
+in the evolution of preferences between different treatment groups,
+evidence that preferences do change through deliberation affirms the central
+assumption of this dissertation and further motivates the question of how
+communication network structure influences that evolution.
 
@@
 The Kendal tau metric is the number fraction of pairwise contests that have
-different results between two profiles \cite{kendall_new_1938}}.
+different results between two profiles \cite{kendall_new_1938}.
 By averaging this measure over all pairs of members, an overall measure of agreement can be calculated for the group.

### 2021-09-28 apx-nk.tex
Add appendices.

--- 2021-09-20 NetDelibSim/supplement.tex
@@
-\documentclass[twocolumn,10pt]{article}
-
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage{graphicx}
-\usepackage{times}
-\usepackage{geometry}
-\usepackage[ruled,vlined]{algorithm2e}
-
-\usepackage[switch]{lineno}
-\linenumbers
-
-\newtheorem{definition}{Definition}
-\newtheorem{property}{Property}
-\newtheorem{claim}{Claim}
-
-\DeclareMathOperator{\unif}{unif}
-\DeclareMathOperator{\sample}{sample}
-\DeclareMathOperator{\setcount}{count}
-\DeclareMathOperator{\mode}{mode}
-\DeclareMathOperator*{\argmax}{arg\,max}
-\DeclareMathOperator*{\argmin}{arg\,min}
-
-\usepackage{geometry}
-\geometry{left=1.5cm,right=1.5cm,top=2cm,bottom=2cm}
-\setlength{\columnsep}{0.5cm}
-
-\title{Supplementary Information: \\
-Small interlocking groups improve mass deliberation\\in the presence of strong social influence}
-\author{
-%Edward L. Platt\\
-%University of Michigan\\
-%elplatt@umich.edu
-%and
-%Herminio Bodon\\
-%Northwestern University\\
-%HerminioBodon2020\\@u.northwestern.edu
-%\and
-%Daniel M. Romero\\
-%University of Michigan\\
-%drom@umich.edu
-Authors Redacted \\
-for \\
-Double-Blind Peer-Review
-}
-\date{\today}
-
-\begin{document}
-
-\maketitle
-
-\section{Simulation Procedure}
-We begin with a network $(V,E_t)$. The vertices $V$ correspond to agents. The edges $E_t$ allow agents to exchange information with their immediate neighbors at time $t$.
-Agents collaborate to find a solution $s$ from a space of solutions $\mathcal{S}$ that maximizes an objective function $Q(s)$.
-We use binary strings of length $d$ as our solution space: $s \in \mathbb{Z}^d$.
-At any one time $t$, each agent $v$ has exactly one preferred solution $s_{v,t}$.
-We generate a tunably rugged objective function $Q(s)$ using an NK-Model \cite{kauffman_towards_1987} (see Section \ref{subsec:task}) with N=15, K=6, exp=8.
-
-\section{Network Topologies}
-
-We use networks to represent constraints on who talks to whom.
-Each agent is represented by a vertex, and two agents are able to interact when their vertices are connected by an edge.
-In this paper we are concerned with network deliberation, which takes place on interlock networks.
-Interlocks are composed of small cliques (i.e., pods) with some vertices belonging to multiple cliques.
-For the interlock networks in this paper, each vertex belongs to exactly one pod at a time, with pod membership periodically reassigned to achieve multiple membership.
-As a control, we also consider networks representing conventional deliberation.
-
-\section{Learning Strategies}
-\label{sec:learning}
-
-% Exclude bitwise majority from paper
-\iffalse
-
-\subsection{Bitwise Majority}
-Instead of evaluating entire solutions on their popularity, it is also possible to evaluate the popularity of each component of a solution, i.e., each bit in the bit string \cite{platt_network_2018}.
-We refer to this strategy as {\em bitwise majority} social learning.
-As with previous strategies, bitwise majority can result in ties for individual bits, which result in ties for the resulting full solution.
-
-\begin{definition}
-The bitwise majority strategy $\mathcal{L}_{bitwise}$ is defined as:
-\begin{eqnarray}
-S_i &=& \bigcup_{x \in S} \{x_i\} \\
-\mathcal{L}_{bitwise}(S)
-&=& \mode(S_1) \times \mode(S_2) \ldots \times \mode(S_d),
-\end{eqnarray}
-where $x_i$ is the $i$th element of the binary string $x$,
-$S_i$ is a multiset, and $\mode(S)$ returns a set containing the mode or modes of $S$, and $\times$ is the Cartesian product.
-\end{definition}
-As with conform, the bitwise majority strategy does not incorporate any new information about solution quality.
-Bitwise majority also has the notable property of being able to produce novel solutions that vary significantly from any previously seen.
-
-% End of bitwise majority
-\fi
-
-\section{NK Model}
-\label{subsec:task}
 The NK Model \cite{kauffman_towards_1987} is an optimization problem particularly well-suited for modelling complex tasks.
@@
 When a skewed distribution is preferred, it is common to exponentiate the value of $Q(s)$ as a final step \cite{lazer_network_2007, barkoczi_social_2016, gomez_clustering_2019}.
-
-\section{Supplemental Results}
-
-\bibliography{references}
-\bibliographystyle{plain}
-
-\end{document}

### 2021-09-28 NetDelibSim/supplement.tex
Move nk material.

@@
 \label{subsec:task}
-The NK Model \cite{kauffman_towards_1987} is an optimization problem particularly well-suited for modelling complex tasks.
-The model is used to create a fitness function $Q(s)$ over some discrete space $\mathcal{S}$, typically binary strings of a fixed length.
-The model is parameterized by two variables.
-The first, $N$, is the dimension of the solution space, i.e., the length of each binary string.
-The second parameter, $K$, determines the ``ruggedness'' of the fitness function, i.e., the number of local maxima.
-In effect, the $K$ parameter allows the complexity of the optimization problem to be tuned.
-The ability to tune task complexity makes the NK model well-suited for studying the role of complexity in various settings.
-The construction of an NK fitness function from a set of parameters is a stochastic process.
-So particular values of $N$ and $K$ define a class of fitness functions, which can be sampled to produce a specific fitness function.
-
-We now show the construction of the NK model fitness function.
-A class of NK model fitness functions can be defined by a tuple of integer parameters
-$(N, K)$ such that $N > 0$ and $0 \leq K < N$.
-We begin by defining $N$ fitness contributions functions:
-\begin{eqnarray}
-q_i &:& \mathbb{Z}^{K+1} \rightarrow [0, 1].
-\end{eqnarray}
-The value of each $q_i(x)$ is chosen uniformly at random in the range $[0, 1]$ at the time the function is defined.
-We also define $N$ projection operators $P_i$ which select $K+1$ bits from a length-$N$ bit string.
-Each $P_i$ selects the bit at index $i$ and $K$ other indices, chosen uniformly at random at the time $P_i$ is constructed.
-For a solution $s$, the $i$th fitness contribution is evaluated on the $K + 1$ bits selected by the $i$th projection: $q_i(P_i(s))$.
-The value of the fitness function is the mean of all fitness contributions:
-\begin{eqnarray}
-Q(s) &=& \frac{1}{N}\sum_{i=1}^N q_i(P_i(s)).
-\end{eqnarray}
-
-The parameter $K$ alters the ruggedness of the fitness function by controlling the interdependence between the $q_i$.
-When $K = 0$, each $q_i$ depends on a single unique index of $s$,
-allowing the $q_i$ to be optimized independently.
-In this case, $Q(s)$ has a single maximum: the global maximum.
-However, for $K > 0$, two things happen: it becomes possible for each $q_i$ to have multiple local maxima, and some of the $q_i$ become coupled due to dependence on the same indices of $s$.
-The result is that as $K$ increases, both the number of local maxima of $Q(s)$ \cite{weinberger_local_1991} and the difficulty of simultaneously optimizing the $q_i$ increase.
-In other words, $Q$ becomes more rugged, and more complex.
-
-The distribution of local maximum values is asymptotically normal for large $K$ \cite{weinberger_local_1991}.
-When a skewed distribution is preferred, it is common to exponentiate the value of $Q(s)$ as a final step \cite{lazer_network_2007, barkoczi_social_2016, gomez_clustering_2019}.
-
 \section{Supplemental Results}

### 2021-09-28 proposal.tex
Merge network subsection of ABM paper.
Add intro and part of experiment ch.
Change terminology in ch2 from 'efficiency' to 'productivity'.
Figures and title page.

@@
 
+\usepackage[ruled,vlined]{algorithm2e}
 \usepackage{amsmath}
@@
 \usepackage{apacite}
+\usepackage[toc,page]{appendix}
 \usepackage{bbding}
@@
 
+\renewcommand{\labelenumii}{\theenumii}
+\renewcommand{\theenumii}{\theenumi.\arabic{enumii}.}
+
 \newtheorem{definition}{Definition}
@@
 
-\linenumbers
+%\linenumbers
 
@@
 
-\maketitle
+\begin{titlepage}
+   \begin{center}
+       \vspace*{1cm}
 
-\begin{abstract}
-\include{abstract.tex}
-\end{abstract}
+       \textbf{\Large Network Deliberation: The role of network structure in large-scale, internet-enabled, participatory decision-making}
+
+       \vspace{1.5cm}
+
+       \textbf{Edward L. Platt}
+
+       \today
+
+       \vfill
+
+       Dissertation Proposal\\
+       Doctor of Philosophy \\
+       University of Michigan \\
+       School of Information
+
+       \vspace{0.8cm}
+
+   \end{center}
+    {\bfseries Doctoral Committee:}\\
+    Associate Professor Daniel M. Romero \\
+    Professor Scott E. Page \\
+    Associate Professor Tawanna Dillahunt \\
+    Assistant Professor Ceren Budak
+
+\end{titlepage}
+
+%\begin{abstract}
+%\include{abstract.tex}
+%\end{abstract}
 
@@
 
+\listoffigures
+\listoftables
+
 \chapter{Introduction}
-\include{ch-intro.tex}
+\label{chap:intro}
+%\include{ch-intro.tex}
 
 \chapter{Network structure, productivity, and performance in WikiProjects}
-\include{ch-wp-prod-perf.tex}
+\label{chap:wp-prod-perf}
+%\include{ch-wp-prod-perf.tex}
 
@@
 of strong social influence}
-\include{ch-abm.tex}
+\label{chap:abm}
+%\include{ch-abm.tex}
 
 \chapter{Experimental evaluation of Network Deliberaiton}
+\label{chap:experiment}
+\include{ch-experiment.tex}
 
-\chapter*{Acknowledgements}
+%\chapter*{Acknowledgements}
 %\include{acknoledgements.tex}
+
+\begin{appendices}
+
+\chapter{The NK-Model}
+\label{apx:nk}
+\include{apx-nk.tex}
+
+\chapter{Interview Script}
+\label{apx:interview}
+\include{apx-interview.tex}
+
+\end{appendices}
 

### 2022-05-02 Proposal/ch-abm.tex
Move proposal to subfolder.

--- 2021-09-20 ch-abm.tex
@@
 As their collective task, agents seek an optimal point on a rugged fitness landscape.
-This landscape is generated using the NK model \cite{kauffman_towards_1987, weinberger_local_1991} (see Appendix \ref{apx-nk}).
+This landscape is generated using the NK model \cite{kauffman_towards_1987, weinberger_local_1991} (see Appendix \ref{apx:nk}).
 The NK model is a ``tuanbly rugged'' objective function parameterized by two variables:
@@
 We use binary strings of length $d$ as our solution space: $s \in \mathbb{Z}^d$.
-We generate the objective function $Q(s)$ using the NK model \cite{kauffman_towards_1987, weinberger_local_1991} (see Appendix \ref{apx-nk}).
+We generate the objective function $Q(s)$ using the NK model \cite{kauffman_towards_1987, weinberger_local_1991} (see Appendix \ref{apx:nk}).
 

### 2022-05-02 Proposal/ch-discussion.tex
Move proposal to subfolder.

The assumption that coordination requires coercive hierarchy is so pervasive
that it is common to see the words ``hierarchy'' and ``organization'' used
interchangeably, as in the conventional ``org chart.''
Hierarchy is certainly a well-demonstrated tool for achieving coordination,
but that coordination comes with costs.
In terms of principles, hierarchical organizations rely on coercion and power imbalances,
sacrificing egalitarianism.
More practically, when information is distributed but decision-making is centralized,
that decision-making necessarily excludes potentially useful information.
The emergence of successful large-scale, internet-enabled, non-hierarchical collaborations
suggests that perhaps, at leaast in some cases,
neither the principle of egalitarianism nor the practical wisdom of the
crowd needs to be sacrificed to achieve effective coordination and collaboration.
The key question is: how do such large-scale participatory projects achieve success?

The work described and proposed here is motivated by and founded on empirical observations,
but it is also speculative in nature.
By employing agent-based models and experiment, it is possible to understand
not just how things are, but how things could be.
Rather than seeking to identify principles that apply broadly to existing
collaborations, this work starts from empirical observations of collaborations
that have done something unusual and attempts to identify robust principles
that might be applied to transform existing collaborations or create new ones.
The repeated appearance of small interlocking groups in successful large-scale
participatory collaborations suggests that this type of network structure might
be such a principle.
It may seem counterintuitive or even impossible that collaborations and organizations might deliberately
alter their interpersonal network structure.
But in fact, countless practices such as job interviews, letters of recommendation,
performance reviews, internships, mentorships, etc. all serve exactly this purpose;
to deliberately shape the social network within a collaboration.
By analogy, the task of deliberately designing a neighborhood might seem impossible,
but urban planners have made a science of it.

Between the completed chapters of this dissertation, a few commonalities emerge.
While it is conventional to describe the ``quality'' of a collaborive output,
both Chapter \ref{chap:wp-prod-perf} and Chapter{chap:abm} demonstrate that
quality can be multidimensional.
Specifically, Chapter \ref{chap:wp-prod-perf} shows that productivity and performance can
be anti-correlated in some contexts and not in others.
Similarly, Chapter \ref{chap:abm} identifies contexts where altering network structure
and social learning strategies can improve just performance, just productivity, or both.
Another common theme is that subtle differences in task or behavior can have dramatic
influences on output quality.
With regards to network deliberation in particular, Chapter \ref{chap:abm} suggests
that it is most effective in collaborations with a tendancy towards conformity.
Such conformity might occur because of social norms, or to compensate for sparse information.
Interestingly, the correlations between network properties and performance on Wikipedia
were consistent with a conformity-based social learning strategy.
Why network deliberation seems to improve conformity-based collaboration remains an
open question for future work.
And whether network deliberation has a similar effect in a controlled experimental
setting is yet to be seen.

While this work only scratches the surface of factors contributing to successful collaboration,
especially in the messiness of the real world, it identifies previously overlooked distinctions
that may help refine the study of participatory collaboration. Furthermore, starting from empirical
observation of large-scale participatory collaborations, this work has identified, refined, and formalized
network deliberation, and shown that, at least in agent based models, exhibits a distinct influence
on collaborative outcomes, independent of other network properties.
The ongoing work described in Chatper \ref{chap:experiment} will shed light on whether network deliberation
plays a causal role in enabling large-scale participatory collaborations.

### 2022-05-02 Proposal/abstract.tex
Move remainder of proposal to subfolder.

As groups grow in size, they gain access to additional resources,
creating opportunities for collective intelligence and collective action.
However, at very large scales, group decision-making becomes prohibitively slow and difficult
to coordinate.
Traditional solutions include representative decision-making and/or a shift from deliberation
to voting.
Both approaches sacrifice desirable properties.
Representative decision-making loses the potential benefits of collective intelligence and
introduces hierarchies that may place the interests of specific individuals ahead of the
interests of the group.
Voting sacrifices generativity: allowing a choice between predefined options,
without allowing for improvement to those options.
Arrow’s impossibility theorem also fundamentally limits the fairness of voting.
This project proposes Networked Deliberation as a potential means of large-scale decision-making.
In Networked Deliberation, members of a large group are repeatedly partitioned into small
deliberative pods.
Overlap between pods at different stages enables group-wide diffusion of information and
preferences.
Different methods for assigning members to pods result in different network topologies.
In partnership with large community groups, this project will use a web-based platform to
evaluate Network Deliberation.
For a variety of network topologies, the speed and quality of preference convergence will be
evaluated using quantitative information from ranked-choice polls throughout the process.
Deliberator satisfaction and experience will also be evaluated using qualitative analysis of
deliberative text and surveys.
This work seeks to improve the ability of very large groups to quickly develop a consensus,
enabling a more effective use of shared resources to achieve common goals.

### 2022-05-02 Proposal/apx-interview.tex
Move remainder of proposal to subfolder.

\section{Preamble}

Hi! Thank you so much for taking time out to talk to us today. We appreciate it. Today we are here to understand how you use the internet to discuss complex issues and your experience participating in our discussion study.

\section{Warm-Up}

\begin{enumerate}
\item How often do you discuss political and policy issues with others, in general?
\item Where and how do you discuss those political and policy issues?
\item Which online social networks, forums, email lists, etc. do you regularly participate in?
\item In general, what have your experiences discussing politics and policy online been like?
\item Do you have any memorable experiences discussing an issue online?
\end{enumerate}

\section{Main Questions}

\begin{enumerate}
\item Reminder of topic, number of stages, length of stages
\item Going into the deliberation, what was your goal?
\item Overall, how would you describe your experience?
\item Do any memorable moments stand out to you?
\item How did it compare to other group deliberations you have participated in? 
\item What was your opinion in at the beginning of the deliberation?
\item What types of opinions did you encounter from other participants?
\begin{enumerate}
\item (follow-up) Can you walk me through how your opinions evolved throughout the deliberation?
\end{enumerate}
\item How much consensus existed at the beginning of the deliberation?
\begin{enumerate}
\item (follow-up) Can you walk me through how the group consensus evolved throughout the deliberation? 
\end{enumerate}
\item Did you witness or experience disagreement throughout the study?
\begin{enumerate}
\item (follow-up) Can you describe how disagreements and conflicts were addressed during deliberation? 
\item (follow-up) What is your opinion of the way disagreements and conflicts were addressed? (If clarification needed: was the process pleasant or unpleasant? Was it effective?)
\end{enumerate}
\item Did you share all of your thoughts with your groups?
\begin{enumerate}
\item (If the participant answers no) What do you think were the reasons?
\end{enumerate}
\item How often did you read posts by other participants? 
\item How often did you read replies to posts by other participants?
\item How often did you reply to posts by other participants?
\item What was your experience like when replying to previous comments?
\end{enumerate}

\section{Debrief}

\section{Final Questions}

\begin{enumerate}
\item Do you have any feedback on how to improve the system?
\end{enumerate}

### 2022-05-02 Proposal/ch-experiment.tex
Move remainder of proposal to subfolder.

--- 2021-09-28 ch-experiment.tex
@@
 \section{Introduction}
+Networks of small interlocking pods appear as a common thread throughout
+successful large-scale collaborations.
+However, it remains unclear whether that network structure plays a causal role
+in the success of those collaborations,
+and whether the specific structural properties of that interlocking network matter.
+Observational study suggests a correlation between properties such as degree,
+structural efficiency, and structural inequality and the performance/productivity
+of collaborations \cite{platt_network_2018}.
+Numerical simulation suggests a causal relationship in a simplified theoretical
+setting.
+This chapter proposes an online experiment to bridge these two findings.
 
 \section{Study Software}
+
+In order to conduct studies of network deliberation online, I have modified the
+free and open-source Loomio platform \cite{jackson_open_2016}
+to implement network deliberation.
+Loomio is a widely-used online deliberation platform which provides both
+forum functionality (Figure \ref{fig:reply})
+and various forms of voting, including ranked-choice (Figure \ref{fig:vote}).
+Loomio is popular among grassroots organizations and worker-owned cooperatives
+(Loomio itself is a worker-owned cooperative).
+Loomio was chosen for its combination of built-in functionality,
+existing user base, and open-source extensibility.
+
+The modified platform implements both random-pod and long-path assignment methods
+and automatically reassigns participants when the stage is advanced.
+In addition to network deliberation features, I have made several modifications to implement
+the experimental protocol.
+Upon signing in for the first time,
+participants are randomly assigned to a treatment group.
+Participants are also assigned an alias which they will use throughout the
+deliberation (Figure \ref{fig:alias}).
+
+\begin{figure}
+\label{fig:reply}
+\center
+\frame{\includegraphics[width=5in]{fig/Experiment/fig-reply.png}}
+\caption{Participants can perform standard forum actions, such as posting and commenting.}
+\end{figure}
+
+\begin{figure}
+\label{fig:vote}
+\center
+\frame{\includegraphics[width=3in]{fig/Experiment/fig-vote.png}}
+\caption{Participants are presented with a ranked-choice voting screen between rounds of deliberation.}
+\end{figure}
+
+\begin{figure}
+\label{fig:alias}
+\center
+\frame{\includegraphics[width=3in]{fig/Experiment/fig-alias.png}}
+\caption{Participants are assigned an alias for the duration of the experiment.}
+\end{figure}
 
@@
 The deliberation took place in chat rooms running the free and open-source
-OTree online experiment software \cite{chen_oTree_2016}.
+OTree online experiment software \cite{chen_otreeopen-source_2016}.
 
@@
 profile onto the nearest consensus profile according to a suitable distance.
-For example, Dodgson's method \cite{dodgeson_method_1876, brandt_computational_2012}
+For example, Dodgson's method \cite{dodgson_method_1876, brandt_computational_2012}
 defines a distance based on swapping adjacent entries in individual preference profiles.

### 2022-05-02 UMich/chapters/ch01-intro.tex
Move proposal content into UMich template.

The ability of large groups to reach mutually agreeable decisions is key to
democratic governance, social movements, and peer production (e.g., Wikipedia).
Faced with the intractability of large-scale decision-making, traditional
systems have sacrificed one or more desirable properties, such as participation
(as in representative democracy), deliberation (as in voting), equality
(as in command hierarchies), and speed (as in formal consensus).
This dissertation examines the emerging role of internet-enabled collaborative
networks in overcoming these historical limitations.

In recent years, examples of large-scale collaborations have emerged that seem
to achieve the previously unachievable. Millions of volunteer Wikipedia editors
have created a high-quality encyclopedia without centralized leadership
\cite{keegan_evolution_2017, giles_internet_2005}.
The Free Software movement has produced the Linux kernel and GNU operating
system, which power much of the modern internet
\cite{coleman_coding_2013, benkler_coases_2002, raymond_cathedral_1999}.
Social movements such as the Arab Spring, Occupy, Black Lives Matter, and
Podemos have reshaped national governments and brought attention to deeply
entrenched social issues
\cite{tufekci_twitter_2017, gonzalez-bailon_networked_2016}.
The emergence of these large-scale decentralized collaborations has been
attributed to the fast, bidirectional,
and global communication enabled by the internet
\cite{tufekci_twitter_2017, benkler_coases_2002}.
A better understanding of how specifically such communication sidesteps
historical barriers to large-scale collaboration will contribute to more
effective policy as well as best practices for organizational design and
intervention.
This dissertation focuses on one particularly challenging aspect of
such collaborations: decision-making
I specificaly examine how,
when groups are too large for all members to participate in all discussions,
the course and outcome of the decision-making process is influenced by the
communication network structure: the shape of who talks to whom.

\section{Theoretical Framework}

This dissertation draws on ideas from network science, economics, and complex
systems.
While a topic as broad as decision-making can be studied from many perspectives,
these fields provide a minimal framework for studying how individual preferences
and behaviors interact with interpersonal communication networks to influence
group decisions.

The fundamental challenge of large-scale collective decision-making is how to
reconcile the conflicting preferences of individual group members.
This challenge has been studied formally in social choice theory,
a sub-field of economics.
Prior work in social choice theory has found, somewhat discouragingly,
that even when all individual preferences are known perfectly,
making a fair collective decision isn't always possible.
In some cases, the method of aggregating individual preferences
(i.e. the voting system) can influence the outcome.
Social choice theory focuses on understanding of these limitations,
such as the Condorcet Paradox
\cite{condorcet_essay_1785} and
Arrow's Impossibility Theorem \cite{arrow_social_2012}.
This dissertation finds hope in the transformative potential of the
deliberative process.
Social choice theory typically assumes fixed individual preferences.
Deliberation allows individuals to influence and change each other's
preferences,
which creates the potential to sidestep the historical limitations of
social choice theory.
When individual preferences are allowed to vary, it becomes possible for an
irreconcilable set of preferences to evolve into one with a clear winner.
So far, this possibility has received relatively little attention,
most likely due to the historical intractability of large-scale deliberative
decision-making.
This dissertation explores the potential of the internet to enable effective
deliberation in large collectives.
Such large-scale deliberation creates the potential for members to
resolve conflicting preferences and reach mutually acceptable decisions
without relying on coercive or hierarchical processes that might introduce
power imbalances or informational biases.

Network science provides the tools for analyzing the structure of interpersonal
networks.
Interpersonal interactions in large collaborations are necessarily structured:
when a group is too large for each individual to interact with all other individuals,
the question of ``who talks to whom?'' creates a network structure.
By modelling collaborative groups as a collection of abstract ``nodes''
connected by interpersonal communication links,
a group's communication structure can be studied in isolation.
Findings from network science suggest that social processes on networks can be
influenced by structural properties such as
the {\em degree}: the number of links a node has,
{\em geodesic distance} the number of links separating two nodes,
and {\em clustering}: how common it is for two linked nodes to share links with
a third \cite{boccaletti_complex_2006}.
While network structure is certainly not the only factor to influence collective
decision-making,
studying network structure in isolation provides a baseline for the
further study of social dynamics and other non-structural factors.
Network structure is also significant as a potential point of intervention in
cases where social dynamics may be difficult to influence.

This dissertation also incorporates theoretical and computational models from
social learning theory.
Social learning theory acknowledges that individuals rarely
learn or make decisions in isolation, but rather learn from and imitate others
in their social network
\cite{golub_naive_2010}.
Social learning theory
formalizes both the types of tasks collectives perform
\cite{hong_interpreted_2009}
and the behavioral strategies individuals might employ
\cite{lazer_network_2007, barkoczi_social_2016}.
These strategies range from pure imitation to critical evaluation,
depending on the circumstances being modeled.
Social learning models provide a baseline to compare empirical observations
against,
as well as a language and framework for contextualizing findings.

Throughout this dissertation,
I motivate and develop a novel theoretical framework,
which I call {\em network deliberation}.
My review of the literature identifies a common theme
among successful large-scale internet-enabled collaborations:
large collectives composed of interlocking smaller groups.
These groups have various names, including:
committees, working groups, teams, circles, cores, syndicates,
affinity groups, zones, and nodes.
As an abstraction of these small interlocking group, I will use the term "pods."
Network deliberation describes large-scale collective deliberation achieved
through interlocking pods.
As in the theories of interlocking directorates \cite{levine_study_1979},
interlocking publics \cite{habermas_structural_1991},
and network rotation \cite{salehi_hive_2018},
pods allow for beneficial small group dynamics,
while the overlap between pods enables diffusion of information and opinions
through the greater collective.
In network deliberation, the method of assigning individuals to pods
(whether deliberate or self-organized) can produce interpersonal networks with
varying structures.
The central question of this dissertation is how the structure of these
interlocking-pod networks influences the process and outcome of deliberation
in large collaborations.

\section{Methodology}

Studying collective behavior on the scale of hundreds, thousands, and millions
presents significant methodological challenges.
To address these challenges, this dissertation combines multiple methodologies,
including: observational study, agent-based modelling, and field experiment.
I use observational studies to reconstruct real-world collaborative networks
from the English-language Wikipedia and analyze the collaborative output of
those networks (Chapter \ref{chap:wp-prod-perf}).
Observational study has the benefits of scaling to millions of individuals
in a real-world environment.
However, observational studies typically cannot establish causal relationships,
only correlation.

To begin to address the causal relationship between network structure and
deliberative outcome, I use agent-based models
(Chatpers \ref{chap:wp-prod-perf} and \ref{chap:abm}).
Agent-based models are computational models of large systems composed of
many agents following simple behavioral rules.
In this case, agents represent individual collaborators,
and their behavior is determined by their preferences and their strategy for
incorporating information learned from their neighbors.
Agent-based models can establish causality,
and do so in group sizes limited only by available computing power.
As simplified models, however, their results cannot necessarily be generalized
to real-world scenarios.

To bridge the limitations of the above methods,
this dissertation will include a controlled field experiment evaluating the
effect of network deliberation in real-world collective decision-making
(Chapter \ref{chap:experiment}).

\section{Contributions}
This dissertation describes the contributions of three projects.
Chapter \ref{chap:wp-prod-perf} describes an observational and computational
study of WikiProjects on the English-language Wikipedia,
and reports the following contributions:
\begin{itemize}
\setlength\itemsep{0pt}
\item Despite an overall productivity/performance trade-off,
WikiProjects with low-degree coeditor networks tend to have both higher
productivity and higher performance;
\item Short geodesic lengths are associated with higher performance,
consistent with a conformity-based learning strategy;
\item Structural inequality, as measured by degree skewness,
is associated with lower performance;
\item The agent-based model shows that the productivity and performance of
collaborations can depend on network degree,
and that the direction of that dependence can depend on social learning strategy.
\end{itemize}

Chapter \ref{chap:abm} describes an agent-based model of network deliberation,
comparing performance across several network topologies and social learning
strategies.
The primary contributions are:
\begin{itemize}
\setlength\itemsep{0pt}
\item When agents conform to social influence,
Network deliberation identifies solutions of higher quality than
conventional deliberation,
while requiring less time to converge.
\item Within network deliberation conditions,
the structurally-efficient random pod network outperforms the
structurally-inefficient long-path network when agents conform to social influence.
However, when agents prefer their own judgement to social influence,
the inefficient network yields higher performance,
consistent with findings for conventional networks.
\cite{barkoczi_social_2016}.
\item A novel social learning strategy, confident-neighbor,
which outperforms or matches the conventional best-neighbor strategy across all
networks considered,
despite using strictly less information about solution quality.
\end{itemize}

Chapter \ref{chap:experiment} proposes and describes current progress on a
field experiment to study network deliberation in large-scale human collaborations.
The study design uses periodic ranked-choice polls to track individual preferences
throughout the course of a large-scale online deliberation.
By varying communication network structure and tracking the evolution of individual
preferences,
this experiment evaluate the ability of network deliberation to resolve conflict
and build consensus, relative to conventional single-group deliberation.

\section{Dissertation Progress and Timeline}
At the time of this proposal,
the analysis of English-language WikiProjects described in Chapter
\ref{chap:wp-prod-perf} has been completed and resulted in a published
paper \cite{platt_network_2018}.
The agent-based modelling project described in Chapter \ref{chap:abm}
has been completed and resulted in a finished manuscript.
I have completed custom software, obtained IRB exemption, and conducted
several pilot studies for the field experiment described in
Chapter \ref{chap:experiment}.
I expect to have data collected from at least one full-scale experiment
by the end of 2021, and to have the data analyzed by mid-February, 2022.
I plan to have my dissertation completed by mid-May, 2022.

### 2022-05-02 UMich/chapters/ch02-WPProdPerf.tex
Move proposal content into UMich template.

--- 2021-09-27 ch-wp-prod-perf.tex
@@
 including Wikipedia.
-\pagebreak
 
@@
 \begin{center}
-\includegraphics[width=3in]{fig/WPProdPerf/fig-knitting.png}
+\includegraphics[width=3in]{chapters/figures/WPProdPerf/fig-knitting.png}
 \caption{
@@
 \centering
-\includegraphics[width=2.5in,height=2.5in]{fig/WPProdPerf/fig-eff-perf-hist.pdf}
+\includegraphics[width=2.5in,height=2.5in]{chapters/figures/WPProdPerf/fig-eff-perf-hist.pdf}
 \caption{
@@
 \centering
-\includegraphics[width=2.5in,height=1.67in]{fig/WPProdPerf/fig-perf-eff.pdf}
+\includegraphics[width=3in]{chapters/figures/WPProdPerf/fig-perf-prod}
 \caption{
@@
 %\centering
-%\includegraphics[width=2.33in,height=2.33in]{fig/WPProdPerf/fig-degree-mincut.pdf}
+%\includegraphics[width=2.33in,height=2.33in]{chapters/figures/WPProdPerf/fig-degree-mincut.pdf}
 %\caption{
@@
 \centering
-\includegraphics[width=3in,height=1.5in]{fig/WPProdPerf/fig-val-iter.pdf}
+\includegraphics[width=3in,height=1.5in]{chapters/figures/WPProdPerf/fig-val-iter.pdf}
 \caption{
@@
 \begin{figure*}
-\includegraphics[width=6.67in,height=1.33in]{fig/WPProdPerf/fig-deg-eff.pdf}
-\includegraphics[width=6.67in,height=1.33in]{fig/WPProdPerf/fig-deg-perf.pdf}
+\includegraphics[width=6.67in,height=1.33in]{chapters/figures/WPProdPerf/fig-deg-eff.pdf}
+\includegraphics[width=6.67in,height=1.33in]{chapters/figures/WPProdPerf/fig-deg-perf.pdf}
 \caption{

### 2022-05-02 UMich/chapters/ch03-NetDelibABM.tex
Move proposal content into UMich template.

--- 2022-05-02 Proposal/ch-abm.tex
@@
     \centering
-    \includegraphics[width=3.33in]{fig/NetDelibABM/fig-broadcast.eps}
+    \includegraphics[width=3.33in]{chapters/figures/NetDelibABM/fig-broadcast.eps}
     \caption{Mean time necessary for a signal broadcast from one node to reach the entire network. We use mean broadcast time as a measure of geodesic length for time-varying networks.}
@@
         \centering
-        \includegraphics[width=3.33in]{fig/NetDelibABM/fig-perf-both.eps}
+        \includegraphics[width=3.33in]{chapters/figures/NetDelibABM/fig-perf-both.eps}
     \end{minipage}%
@@
     \centering
-    \includegraphics[width=1.875in]{fig/NetDelibABM/fig-converge-both.eps}
+    \includegraphics[width=1.875in]{chapters/figures/NetDelibABM/fig-converge-both.eps}
     \end{minipage}
@@
     \centering
-    \includegraphics[width=3.33in]{fig/NetDelibABM/fig-result-frac-parallel.eps}
+    \includegraphics[width=3.33in]{chapters/figures/NetDelibABM/fig-result-frac-parallel.eps}
 \caption{Fraction of simulations in which the row condition outperforms the column condition. Results are shown for parallel individual learning. In the case of a tie, weight is divided evenly.}
@@
     \centering
-%    \includegraphics[width=3.33in]{fig/NetDelibABM/fig-result-frac-parallel.eps}
+%    \includegraphics[width=3.33in]{chapters/figures/NetDelibABM/fig-result-frac-parallel.eps}
 \caption{Fraction of simulations in which the row condition outperforms the column condition. Results are shown for fallback individual learning. In the case of a tie, weight is divided evenly.}
@@
     \centering
-    \includegraphics[width=3.33in]{fig/NetDelibABM/results-bn-parallel-dist.eps}
+    \includegraphics[width=3.33in]{chapters/figures/NetDelibABM/results-bn-parallel-dist.eps}
 \caption{Solution quality distribution for best-neighbor strategy with parallel individual learning.}
@@
     \centering
-    \includegraphics[width=3.33in]{fig/NetDelibABM/results-cn-parallel-dist.eps}
+    \includegraphics[width=3.33in]{chapters/figures/NetDelibABM/results-cn-parallel-dist.eps}
 \caption{Solution quality distribution for confident-neighbor strategy with parallel individual learning.}
@@
     \centering
-    \includegraphics[width=3.33in]{fig/NetDelibABM/results-conf-parallel-dist.eps}
+    \includegraphics[width=3.33in]{chapters/figures/NetDelibABM/results-conf-parallel-dist.eps}
 \caption{Solution quality distribution for the conform strategy with parallel individual learning.}
@@
     \centering
-    \includegraphics[width=3.33in]{fig/NetDelibABM/results-bn-fallback-dist.eps}
+    \includegraphics[width=3.33in]{chapters/figures/NetDelibABM/results-bn-fallback-dist.eps}
 \caption{Solution quality distribution for best-neighbor strategy with fallback individual learning.}
@@
     \centering
-    \includegraphics[width=3.33in]{fig/NetDelibABM/results-cn-fallback-dist.eps}
+    \includegraphics[width=3.33in]{chapters/figures/NetDelibABM/results-cn-fallback-dist.eps}
 \caption{Solution quality distribution for confident-neighbor strategy with fallback individual learning.}
@@
     \centering
-    \includegraphics[width=3.33in]{fig/NetDelibABM/results-conf-fallback-dist.eps}
+    \includegraphics[width=3.33in]{chapters/figures/NetDelibABM/results-conf-fallback-dist.eps}
 \caption{Solution quality distribution for the conform strategy with fallback individual learning.}
@@
     \centering
-    \includegraphics[width=3.33in]{fig/NetDelibABM/fig-results-netdelib-diff.eps}
+    \includegraphics[width=3.33in]{chapters/figures/NetDelibABM/fig-results-netdelib-diff.eps}
 \caption{Difference in solution quality between network deliberation and conventional deliberation.
@@
     \centering
-    \includegraphics[width=3.33in]{fig/NetDelibABM/fig-results-conform-dist.eps}
+    \includegraphics[width=3.33in]{chapters/figures/NetDelibABM/fig-results-conform-dist.eps}
 \caption{Distributions of solution quality for social learning on a preferential-attachment network and random-pod network deliberation. Results are shown for parallel individual learning.}
@@
     \centering
-    \includegraphics[width=3.33in]{fig/NetDelibABM/fig-results-netdelib-efficiency.eps}
+    \includegraphics[width=3.33in]{chapters/figures/NetDelibABM/fig-results-netdelib-efficiency.eps}
 \caption{Difference in solution quality between random-pod (efficient) and long-path (inefficient) network. Results are shown for parallel individual learning.}
@@
     \centering
-    \includegraphics[width=3.33in]{fig/NetDelibABM/fig-results-confident.eps}
+    \includegraphics[width=3.33in]{chapters/figures/NetDelibABM/fig-results-confident.eps}
 \caption{Histograms of the difference in solution quality between confident-neighbor and best-neighbor for both conventional networks.

### 2022-05-02 UMich/chapters/ch04-NetDelibExp.tex
Move proposal content into UMich template.

--- 2022-05-02 Proposal/ch-experiment.tex
@@
 \center
-\frame{\includegraphics[width=5in]{fig/Experiment/fig-reply.png}}
+\frame{\includegraphics[width=5in]{chapters/figures/Experiment/fig-reply.png}}
 \caption{Participants can perform standard forum actions, such as posting and commenting.}
@@
 \center
-\frame{\includegraphics[width=3in]{fig/Experiment/fig-vote.png}}
+\frame{\includegraphics[width=3in]{chapters/figures/Experiment/fig-vote.png}}
 \caption{Participants are presented with a ranked-choice voting screen between rounds of deliberation.}
@@
 \center
-\frame{\includegraphics[width=3in]{fig/Experiment/fig-alias.png}}
+\frame{\includegraphics[width=3in]{chapters/figures/Experiment/fig-alias.png}}
 \caption{Participants are assigned an alias for the duration of the experiment.}
@@
 \label{fig:kcon}
-\includegraphics[width=5in]{fig/Experiment/fig-kcon.eps}
+\includegraphics[width=5in]{chapters/figures/Experiment/fig-kcon.eps}
 \caption{The k-connectivity of long-path and random-pod network deliberation

### 2022-05-05 UMich/chapters/ch04-NetDelibExp.tex
Import NetDelibExp and formatting fixes.

--- 2022-05-02 Proposal/ch-experiment.tex
@@
 Numerical simulation suggests a causal relationship in a simplified theoretical
-setting.
-This chapter proposes an online experiment to bridge these two findings.
+setting (Chapter \ref{chap:abm}).
+To bridge these two findings,
+this chapter presents the results of an online experiment studying the role of network deliberation in the evolution of group preferences.
 
-\section{Study Software}
+Collaboration and collective action allow groups of individuals to manage shared resources and combine efforts towards common goals, but also introduce the problem of governance and collective decision-making.
+While collective governance of common resources is a historically hard problem, Ostrom described principles seen in successful systems \cite{ostrom_collective_2000}.
+More recently, the proliferation of inexpensive point-to-point global communication via the internet has enabled a number of successful large-scale collaborations and movements, including free/open-source software \cite{benkler_coases_2002, coleman_coding_2012, raymond_cathedral_1999}, Wikipedia \cite{keegan_evolution_2017, benkler_coases_2002, giles_internet_2005}, and social movements \cite{tufekci_twitter_2017, gonzalez-bailon_networked_2016}.
 
-In order to conduct studies of network deliberation online, I have modified the
-free and open-source Loomio platform \cite{jackson_open_2016}
-to implement network deliberation.
-Loomio is a widely-used online deliberation platform which provides both
-forum functionality (Figure \ref{fig:reply})
-and various forms of voting, including ranked-choice (Figure \ref{fig:vote}).
-Loomio is popular among grassroots organizations and worker-owned cooperatives
-(Loomio itself is a worker-owned cooperative).
-Loomio was chosen for its combination of built-in functionality,
-existing user base, and open-source extensibility.
+Scholars of democracy commonly describe governance systems in terms of two axes: representation and argumentation \cite{ackerman_deliberation_2002, gonzalez-bailon_networked_2016}.
+Here, {\em representation} refers to how many people participate in decision-making, while {\em argumentation} refers to the amount of discourse between the decision-makers.
+While both can improve governance, they are difficult to achieve simultaneously \cite{gentry_consensus_1982, freeman_tyranny_1972} and tradeoffs can be made to favor one over the other, as in voting (high representation, low argumentation) and representative democracy (low representation, high argumentation).
 
-The modified platform implements both random-pod and long-path assignment methods
-and automatically reassigns participants when the stage is advanced.
-In addition to network deliberation features, I have made several modifications to implement
-the experimental protocol.
-Upon signing in for the first time,
-participants are randomly assigned to a treatment group.
-Participants are also assigned an alias which they will use throughout the
-deliberation (Figure \ref{fig:alias}).
+We focus on deliberation due to fundamental limitations of voting.
+Results from social choice theory showt that any ranked-choice voting system is subject to limitations such as the Condorcet Paradox \cite{condorcet_essay_1785} and Arrow's Impossibility Theorem \cite{arrow_social_2012}.
+Deliberation offers one way to avoid these limitations by changing individual preferences \cite{ackerman_deliberation_2002}.
+The success of deliberation at generating consensus has been attributed to its ability to identify and resolve conflict \cite{gentry_consensus_1982}.
+However, deliberation also carries risks, such as the posibility of pushing individuals towards extreme and polarized preferences \cite{schkade_what_2007}.
 
-\begin{figure}
-\label{fig:reply}
-\center
-\frame{\includegraphics[width=5in]{fig/Experiment/fig-reply.png}}
-\caption{Participants can perform standard forum actions, such as posting and commenting.}
-\end{figure}
+In any sufficiently large group, it becomes unfeasible for all members to interact with each other.
+Understanding collaboration in large groups thus requires attention to the network structure of who interacts with whom.
+The field of social learning has studied both simple estimation tasks \cite{degroot_reaching_1974, golub_naive_2010} and more complex rugged-landscape optimization tasks \cite{lazer_network_2007, barkoczi_social_2016}.
+Observational studies have been used to analyze the network structures of online collaborations \cite{gonzalez-bailon_networked_2016, platt_network_2018}.
+Lab and field studies have also been used to study real-world collaborations on a small scale \cite{salehi_hive_2018, kearns_experiments_2012}.
+As discussed in Chapter \ref{chap:abm}, the successful large-scale collaborations enabled by the internet often exhibit network sructures characterized by small, interlocking groups, reminiscent of interlocking directorates \cite{levine_study_1979} and interlocking publics \cite{habermas_structural_1991}.
+In Chapter \ref{chap:abm}, we describe agent-based simulations showing that such network deliberation can improve performance on complex tasks when individuals exhibit strong social influence.dsd
+Here, we expand on existing research by performing a controlled online experimental evaluation of network deliberation.
+We use periodic polling to track preference evolution, allowing causal inference in a real-world setting.
 
-\begin{figure}
-\label{fig:vote}
-\center
-\frame{\includegraphics[width=3in]{fig/Experiment/fig-vote.png}}
-\caption{Participants are presented with a ranked-choice voting screen between rounds of deliberation.}
-\end{figure}
+While many studies have examined the role of network structure in collaboration, there has so far been little attention to the interlocking pod structure of network deliberation.
+The closest we are aware of is the study of network rotation by Salehi \& Bernstein \cite{salehi_hive_2018}.
+The present experiment goes beyond existing work primarily by tracking preferences throughout the experiment and by the use of a conventional single-group control condition.
+We also describe multiple network deliberation topologies for achieving efficient or inefficient networks.
+For the present study, we focus on an efficient topology.
 
-\begin{figure}
-\label{fig:alias}
-\center
-\frame{\includegraphics[width=3in]{fig/Experiment/fig-alias.png}}
-\caption{Participants are assigned an alias for the duration of the experiment.}
-\end{figure}
+The experiment described here focuses on several issues.
+Based on observational and numerical findings that successful large-scale collaborations often exhibit network deliberation struture, we test the following hypothesis:
+\begin{description}
+\item[H1:] Network deliberation results in higher agreement among participants than single large-group deliberation.
+\end{description}
+We are also concerned with the dynamics of preference evolution under network deliberation, which may provide insight into possible mechanisms.
+One such mechanism proposed in consensus decision-making literature is the ability to identify and resolve conflict.
+We thus also seek to address two research questions:
+\begin{description}
+\item[RQ1:] How do preferences evolve throughout network deliberation?
+\item[RQ2:] How effective is network deliberation at identifying and resolving conflict?
+\end{description}
 
-\section{Pilot Studies}
-Several pilot studies have already been conducted:
-one in a synchronous lab setting, and two using the asynchronous online
-software.
-In addition to providing an opportunity to test software,
-these pilot studies have provided crucial information for developing a
-feasible experimental design.
-
-The initial pilot study was conducted on December 21, 2018,
-before the custom online platform was developed.
-In this pilot, 18 participants engaged in a synchronous 60-minute deliberation.
-Participants sat at individual computer workstations, separated by dividers.
-Participants discussed the following question:
-``Where should the use of electric scooters be permitted?''
-Participants were placed in a single treatment group, and divided into 4
-pods of size 4--5.
-These pods were reassigned using the random-pod method for a total of three
-rounds.
-The deliberation took place in chat rooms running the free and open-source
-OTree online experiment software \cite{chen_otreeopen-source_2016}.
-
-The second pilot took place after the initial development of the custom online
-platform, from August 24, 2020 through August 32, 2020.
-In this pilot, 7 participants engaged in an asynchronous deliberation over the
-course of 7 days.
-Participants were assigned to a single treatment group using network deliberation
-with random-pod assignment.
-Pods were size 4, and there were 3 rounds (2, 2, and 3 days).
-Participants were researchers considering a week-long collaboration and
-deliberated on the following question:
-``What is your current preference for week of science topics?''
-
-The third pilot took place from August 23, 2021 through August 29, 2021.
-Participants deliberated on the following question:
-``How should schools in the US reopen for fall 2021?''
-Participants were recruited on reddit and Twitter.
-18 participants were enrolled (excluding several identified as fake accounts).
-These participants were divided between a conventional deliberation control
-group and a random-pod treatment group.
-The random-pod treatment group had two pods of size 5.
-Deliberation was divided into 3 rounds (2, 2, and 3 days).
-
-In these pilots, participants showed quantitative and qualitative evidence of
-changing their preferences over the course of deliberation.
-Here is an example of one such interaction:
-\begin{quote}
-{\bfseries Aspen Sunflower:} Even though my \#1 choice includes mandatory vaccinations, I know it's not possible due to school starting now AND MDHS can't mandate it yet.
-Based on previous stage, I think most seemed to want vaccinations but also wanted options.
-One option doesn't fit all students.
-
-{\bfseries Aspen Poppy:} Major gap in my understanding here!
-The schools do not mandate vaccines for students, the state does.
-For the 12-17 age group, the vaccine is authorized under emergency use, full FDA approval is pending.
-This strengthens my desire to see hybrid learning with smaller in person groups and vaccinated staff.
-\end{quote}
-
-While the pilots did not have enough participants to identify any differences
-in the evolution of preferences between different treatment groups,
-evidence that preferences do change through deliberation affirms the central
-assumption of this dissertation and further motivates the question of how
-communication network structure influences that evolution.
-
-\section{Experimental Design}
-
-This section describes the experimental design of the proposed field experiment.
-In this experiment, participants will use a pseudonymous online platform
-to deliberate on a policy issue.
-The deliberation will take place asynchronously, over a period of several days.
-At several times throughout the deliberation,
-participants will complete a ranked-choice poll regarding the policy issue,
-allowing the evolution of their preferences to be tracked over the course of the
-experiment.
-Depending on opportunity and necessity, multiple waves of deliberation may
-be conducted, each with a distinct policy issue and set of participants.
-
-\subsection{Deliberative Topic}
-The policy issue chosen as the topic of deliberation is a crucial component
-of the experiment.
-The topic will be chosen carefully according to several criteria.
-The chosen policy issue must be:
+We find difference in preference evolution between network deliberation and conventional deliberation. Our primary contributions are:
 \begin{itemize}
-\item Relevant to the experimental population;
-\item Amenable to participants changing their preferences based on new information or reasoning;
-\item Amenable to a predefined list of proposed solutions;
-\item Sufficiently complex to have three or more proposed solutions;
-\item Timely, but unlikely to be influenced by current events during the deliberation.
-\end{itemize}
-Examples of policy issues meeting these criteria include:
-\begin{itemize}
-\item ``Where should the use of electric scooters be permitted?''
-\item ``Given the COVID-19 pandemic, how should schools in the US reopen in Fall 2021?''
+    \item We find experimental support for the hypothesis that network deliberation is better at facilitating agreement than conventional deliberation.
+    \item We find evidence suggesting network deliberation can provide protection against information cascades.
+    \item We find no evidence that network deliberation facilitates substantial conflict-resolution, but that it may provide protection against polarization.
 \end{itemize}
 
-\subsection{Participant Recruitment and Incentives}
-Participants will be recruited using several methods.
-Participants will be recruited individually through posts on web forums
-relevant to the policy question (e.g., reddit, facebook groups)
-and to relevant email lists.
-Participant pools will also be utilized,
-including lists of previous participants
-who have expressed interest in future studies
-and the UMSI Online Recruitment System for Economic Experiments.
-Individual participants will also be recruited using promoted posts and
-advertisements on Twitter, Facebook, etc.
-Separately, participants may also be recruited through partnership with
-community groups who wish to try network deliberation as part of an organizational
-decision-making strategy.
-The number of such collaborations will depend on the number of interested
-partner organizations that have decisions amenable to network deliberation
-over the course of the experiment.
+We present relevant theory in Section \ref{sec:theory}.
+We describe the experiment in Section \ref{sec:experiment} and present the results in Section \ref{sec:res}.
+We discuss the results in Section \ref{sec:discussion} and conclude in Section \ref{sec:ch04conclusion}.
 
-In order to incentivize participation and to compensate participants fairly
-for their time,
-participants who complete the deliberation and pre/post-experiment surveys
-will be compensated financially.
-Participants will be paid a predetermined fixed amount in order to
-minimize the effect of the compensation on their deliberative behavior and
-expressed preferences.
+\section{Theory}
+\label{sec:theory}
+\subsection{Social Choice Theory}
+Our analysis of preferences is foudned in the formalism of social choice theory \cite{arrow_social_2012, gaertner_primer_2009}.
+In social choice theory, individual's {\em preferences} are represented by a ranked ordering of {\em alternatives} (e.g., proposals, candidates).
+The set of preferences for all individuals in a group is called the {\em preference profile}.
+A {\em social welfare function} can be applied to the profile to generate a {\em social preference}: an ordering of alternatives representing the preference of the group as a whole.
+A {\em social choice function} represents a voting system: it selects a subset of winners from the available alternatives.
 
-\subsection{Experimental Protocol}
+\subsection{Quantifying Agreement}
 
-Before the deliberation period begins,
-participants will complete a short demographic survey.
-This survey will enable the identification of sample bias in age, gender, race,
-and other demographic factors.
+Social choice theory typically focuses on identifying and describing the limitations of specific voting systems.
+Instead, we build on social choice theory to quantify the level of agreement in a group, based in individual profiles.
 
-Participants will be randomly assigned to one of three treatment/control groups
-at time of enrollment:
-\begin{description}
-\item[Control]{
-The control group will engage in conventional single large group deliberation.
-Posts and comments made by participants in the control group will be visible
-to all others in the control group for the entire duration of the deliberation.}
-\item[Network Deliberation (Efficient)]{
-Participants in this treatment group will be divided into small
-(4--5 participant) pods.
-Posts and comments made by participants in this treatment group will only be
-visible to others in their current pod.
-All participants will be assigned to a new pod at the beginning of each round
-of deliberation using the random-pod assignment method,
-producing a structurally efficient communication network.}
-\item[Network Deliberation (Inefficient)]{
-Participants in this treatment group will be divided into small
-(4--5 participant) pods.
-Posts and comments made by participants in this treatment group will only be
-visible to others in their current pod.
-All participants will be assigned to a new pod at the beginning of each round
-of deliberation using the long-path assignment method,
-producing a structurally inefficient communication network.}
-\end{description}
-
-The deliberation period will be divided into a predetermined number of {\em rounds},
-each lasting a fixed length of time.
-Present plans are to divide the deliberation period into three stages of two days
-each.
-However, these numbers may be adjusted to ensure time timeline of the
-deliberation is consistent with any constraints imposed by specific policy issues.
-Before and after each stage, participants will be asked to rank the proposed
-solutions to the policy question according to their individual preference.
-During each round, participants will be shown a discussion prompt and will be
-able to post a response to that prompt.
-Example prompts are shown in Table \ref{tab:prompts}.
-Participants will be able to view and comment on the posts of other participants
-assigned to the same pod.
-Participants will also be able to view posts and comments that were visible to them
-in previous rounds, but will no longer be able to interact with or reply to
-those posts and comments.
-
-\begin{table}
-\center
-\label{tab:prompts}
-\begin{tabular}{|p{0.3in}|p{3.6in}|}
-\hline
-Stage & Prompt \\
-\hline
-1 &
-Which proposals do you prefer, and why?
-If there are disagreements, try asking questions to understand the perspective
-of others and to understand the causes of the disagreement.
-\\
-\hline
-2 & In the previous round of discussion, what opinions and reasoning did you
-observe in your group?
-How much agreement was there?
-Were there any disagreements or conflicts?
-If so, what were the sources of conflict, and how might they be resolved?
-\\
-\hline
-3 &
-What seem to be the most popular opinions?
-Do you agree with them?
-Has your opinion changed over the course of the discussion?
-\\
-\hline
-\end{tabular}
-\caption{Example prompts for each round of deliberation.}
-\end{table}
-
-\section{Quantitative Analysis}
-
-The primary quantitative data will be the results of the ranked-choice
-polls conducted before and after each stage of deliberation.
-These results represent individual participants' stated preferences at multiple
-time points.
-Standard tools from social choice theory can be applied to such rank-ordered
-preferences \cite{arrow_social_2012, brandt_computational_2012}.
-At each time, these preferences can be aggregated into a {\em preference profile}
-representing the collective preferences of all participants in a treatment group.
-Preference profiles can be used to aggregate individual preferences into a single
-ranking via a {\em social welfare function}.
-Similarly, preference profiles can be used to select a single winner via a
-{\em social choice function}.
-These social choice and social welfare functions represent voting methods.
-Different voting methods can potentially yield different results for the same
-preference profile.
-However, the more similar preferences are across individual participants,
-the more likely it is that different voting methods will yield the same winner
-or social ranking.
-For example, when one proposal beats all others in pairwise contests, it is known as the
-{\em Condorcet winner}.
-Many voting methods yield the Condorcet winner when it exists, but the existence
-of a Condorcet winner is not guaranteed.
-
-For the purposes of this dissertation, the preference profiles will be used
-to quantify agreement and conflict over the course of deliberation.
+We use ranked-choice polls to construct preference profiles which are used to quantify agreement and conflict over the course of deliberation.
 Within social choice theory, agreement can be defined by a {\em consensus class},
@@
 One possibility is to measure the distance from strong unanimity.
-Two such methods will be used to analyze the results of the deliberation experiment:
-the mean Kendal tau metric and the mean Spearman correlation.
-The Kendal tau metric is the number fraction of pairwise contests that have
-different results between two profiles \cite{kendall_new_1938}.
-By averaging this measure over all pairs of members, an overall measure of agreement can be calculated for the group.
-This measure does not incorporate information about win/loss margins,
-only whether an alternative wins or loses.
-To quantify the win/loss margin, the mean Spearman correlation will be used.
-The Spearman correlation is the linear coefficient of correlation between the
-rank orders of alternatives for two voters \cite{spearman_proof_1904}.
-As with Kendall tau, the Spearman correlation can be averaged over all pairs of
-members to determine a measure of agreement for the entire group.
-It must be noted that representing agreement with a single number will inevitably
-ignore important information.
-Neither of these measures distinguishes a small number of voters disagreeing
-on many comparisons from a large number of voters disagreeing on a few.
-The Spearman correlation is also unable to distinguish these from situations in
-which a small number of voters disagree by a very large margin.
+Any distance metric $\delta$ with a range of $[0,1]$ can be converted to a measure of correlation with range $[-1, 1]$ by taking $1 - 2\delta$.
+One such measure is based on the Kendall tau metric, and the corresponding Kendall tau correlation \cite{kendall_new_1938}.
+The Kendall tau metric is the fraction of pairwise contests that have different results between two profiles.
+By averaging the corresponding correlation over all pairs of members, an overall measure of agreement can be calculated for the group.
+While commonly used in social choice theory, the Kendall tau weights all contests equally, regardless of the ranks of the alteratives.
+In practical settings with a small number of winners, the contests between highly-ranked alternatives are more consequential than those of low-ranked alternatives;
+participants don't care if they disagree on the ordering of losing alternatives as long as they agree on which ones win and lose.
+Various weighted extensions of the Kendall metric have been proposed \cite{shieh_weighted_1998, can_weighted_2014},
+each with benefits and drawbacks.
+We propose our own based on metric structure, uniqueness, and efficiency of calculation, which we describe below.
 
-While unanimity might be preferred,
-it may sometimes be more realistic to focus on transitivity.
-This consensus class requires only that the social preferences between pairs
-of alternatives are transitive.
-Such preference profiles guarantee that a Condorcet winner will exist,
-bringing many different voting methods into agreement on which should win.
+\subsubsection{Weighted Crossing Distance}
+We propose the following weighted distance metric for preference rankings.
+Given a set of $M$ alternatives, there are $M-1$ ways to divide an ordering into two nonempty sets of adjacent alternatives, i.e., a high set and a low set.
+For a pair of preferences $p, q$, we define a length $M - 1$ {\em crossing vector} $v_m(p, q)$ to be equal to the number of alternatives in the highest $m$ positions of $p$ and not in the highest $m$ positions of $q$ (it is easily shown that this measure is symmetric in $p$ and $q$).
+Informally, $v_m$ is the number of alternatives that cross place $m$.
+Each element of $v_m$ is an integer in the range $[0, v_m^{\max}]$ dependent on the index:
+\begin{equation}
+    v_m^{\max} =
+    \begin{cases}
+    m & m \leq (M - 1) / 2 \\
+    M - m & (M - 1) / 2 < m \leq M - 1.
+    \end{cases}
+\end{equation}
+We construct the weighted crossing distance by taking a weighted sum over this vector.
+There are many potential choices for weights.
+We choose the weight of place $m$ to be equal to the highest total of all lower ranked places plus one.
+Thus the weight at index $i = M - m$ (noting that low rank corresponds to high $m$) is given by:
+\begin{equation}
+    w_i = 1 + \sum_{k=1}^{i-1} w_k v_k^{\max}.
+\end{equation}
 
-The deviation from transitivity will be qunatified by counting the number
-of ranked-pair violations.
-When performing a vote by the Tideman ranked pair system,
-the relative order of two alternatives in the social preference is
-determined by the winner of a pairwise vote,
-except in the special case that contests with a higher margin have already
-constrained their order due to transitivity.
-In the latter case, the winner of the pairwise contest may have a lower
-position in the social preference.
-If this situation occurs, the social preference is intransitive and there may
-not be a Condorcet winner.
-Each such pair represents a contest violating the social ranking,
-so the fraction of such pairs
-can be used to quantify the strength of intransitivity in the group’s preferences.
-This measure has the property that a Condorcet winner is guaranteed when the value is 0.
+\subsection{Quantifying Social Influence}
 
-\section{Power Analysis}
+We are particularly interested in the role of social influence in preference evolution and require a formalism to quantify and analzye social influence.
+If two individuals preferences become closer, can we say whether one influenced the other? And if so, can we say which influence which?
+We cannot answer either question with certainty from preferences alone, but we can construct a reasonable proxy based on some simplifying assumptions.
+Namely, we will construct measures of social influence based on the assumptions that:
+\begin{itemize}
+    \item The magnitude of an individual's change in preference increases with their susceptibility with social influence;
+    \item The magnitude of an individual's change in preference increases with the magnitude of the distance between their preference and the influencing preference.
+\end{itemize}
+Thus, if the distance between the preferences of individuals $A$ and $B$ decreases, and the preference of $B$ changed more than the preference of $A$, we assume that individual $A$ influenced individual $B$.
+We now describe two formal measures that use these assumptions to quantify how susceptible to influence and how potential influential an individual is.
+
+\subsubsection{Conformity and Hipness}
+
+Let $X$ be a set of alternatives (e.g., proposals or candidates)
+and let $\mathcal{L}$ be the set of strict total orders over $X$.
+We represent the preference of an individual $p$ at time $t$
+by a strict total order $R_p(t) \in \mathcal{L}$.
+Given $n$ individuals, we call the $n$-tuple of their preferences
+at time $t$ the profile $R(t)$.
+
+Let $d:\mathcal{L}\times\mathcal{L} \to \mathbb{R}$ be a
+dissimilarity function on preferences.
+We define the mean dissimilarity of the profile $R$ at time $t$ as:
+\begin{equation}
+    \frac{1}{n(n-1)} \sum_p \sum_{q \neq p} d\left(p_t, q_t\right),
+\end{equation}
+where $p_t$ is shorthand for $R_p(t)$.
+The above quantity can be interpreted as a measure of disagreement among a group of individuals.
+
+Let $t_i$ be an initial time, and $t_f$ be some later time.
+The negative change in mean dissimilarity is given by:
+\begin{equation}
+    \label{eqn:Delta}
+    \Delta =
+    \frac{-1}{n(n-1)} \sum_p \sum_{q \neq p}
+    \left[
+    d\left(p_f, q_f)\right)
+    - d\left(p_i, q_i\right)
+    \right].
+\end{equation}
+This quantity can be interpreted as the increase in agreement within the group between times $t_i$ and $t_f$,
+with negative values signifying a decrease in agreement.
+
+We wish to quantify whether a change in agreement/disagreement is due to individuals converging towards initially popular alternatives, or to the the group being persuaded towards initially unpopular alternatives.
+In the former case, we would expect the majority of individuals to change their preferences by a small amount, while a smaller number of individuals holding unpopular opinions make larger shifts towards popular preferences.
+In the latter case, we would expect the opposite:
+some individuals with unpopular preferences hold closely to those opinions, while a large number of individuals move away from the previous popular preferences.
+So we are interested not just in the total change in agreement/disagreement,
+but also in how that change is distributed across the individuals in the group.
+
+To apportion the change in agreement between individuals $p$ and $q$,
+we make the assumption that an individual's contribution to the change in agreement is proportional to the change in their own preference.
+Specifically, we define the conformity $c(p,q)$ of individual $p$ toward individual $q$ as:
+\begin{eqnarray}
+    d_i &=& d(p_i, q_i) \\
+    d_f &=& d(p_f, q_f) \\
+    |p| &=& d(p_i, p_f) \\
+    c(p,q) &=& -(d_f - d_i)\frac{|p|}{|p| + |q|}.
+\end{eqnarray}
+Note that the change in agreement between two individuals can be written in terms of the conformity:
+\begin{equation}
+    -(d_f - d_i) = c(p, q) + c(q, p).
+\end{equation}
+Similarly, the total change in agreement (\ref{eqn:Delta}) can be rewritten as a sum of conformity over all pairs of individuals:
+\begin{equation}
+    \Delta =
+    \sum_p \sum_{q \neq p} c(p, q).
+\end{equation}
+By performing only one of the sums above, the change in agreement $\Delta$ can be written as a sum of terms, each corresponding to one individual.
+The sum can be performed over either the first or second argument of $c$.
+If the sum is performed over the second argument of $c$, the result is the mean conformity of a particular individual $p$ toward all others:
+\begin{equation}
+    C(p) = \sum_q c(p, q).
+\end{equation}
+Alternatively, if the sum is performed over the first argument of $c$, the result is the mean conformity of all others toward $p$, which we call the hipness $H(p)$:
+\begin{equation}
+    H(p) = \sum_q c(q, p).
+\end{equation}
+Both $C(p)$ and $H(p)$ are equal to $\Delta$ when summed over all individuals, but they describe conceptually different attributes of an individual.
+The conformity quantifies how much an individual has increased agreement by changing their preferences to match others.
+The hipness quantifies how much an individual has increased agreement due to holding an unchanging preference near to others' final preferences.
+
+\subsection{Network Topologies}
+In Chapter \ref{chap:abm}, we describe two models of network deliberation structure: random-pod and long-path.
+In order to expect a measurable difference from the control condition and between each other, a minimum network size is required.
+We decribe these considerations below.
+For the present experiment, we opt for only a control and random-pod condition based on the avaialble number of participants.
+
+\subsubsection{Network Size}
 The minimum number of participants required is determined by a number of factors.
-First, the long-path pod assignment method results in a minimum number of pods,
+First, the long-path assignment method results in a minimum number of pods,
 determined by the chosen parameters.
@@
 and a minimum of about 35 participants per treatment group is necessary to
-observe differences between the network deliberation treatment conditions.
+observe differences between the random-pod and long-path conditions.
 
@@
 \label{fig:kcon}
-\includegraphics[width=5in]{fig/Experiment/fig-kcon.eps}
+\center
+\includegraphics[width=5in]{chapters/figures/NetDelibExp/fig-kcon}
 \caption{The k-connectivity of long-path and random-pod network deliberation
@@
 
-\section{Qualitative Analysis}
-In addition to quantitative data,
-qualitative data will also be collected in conjunction with
-the deliberation experiment.
-These data will be used to corroborate the results of quantitative
-analysis and to achieve a richer understanding of individual participants'
-experience of network deliberation,
-as well as to as well as identify common themes that may provide
-insight into the strengths and weaknesses of network deliberation.
+\section{Experimental Study}
+\label{sec:experiment}
 
-The text of participants' posts and comments throughout the deliberation
-will be kept for analysis.
-This text can potentially be used to identify deliberative behaviors that
-lead to changes in preferences.
-Similarly, this text can be used to track the diffusion of specific keywords,
-arguments, and influence through the communication network.
+\subsection{Experimental Design}
 
-A subset of participants will be randomly selected to participate in
-60-minute semi-structured interviews about their experience.
-The interview script can be found in Appendix \ref{apx:interview}.
-These interviewed will be transcribed and qualitatively coded using a
-grounded theory approach.
-The initial code book will be developed by open-coding pilot study transcripts
-and an initial subset of interview transcripts.
+Participants used a pseudonymous online platform (described in Section \ref{sec:exp-platform}) to deliberate on a policy issue.
+
+The deliberation took place asynchronously over a period of 7 days, from Monday, November 15, 2021 through Sunday, November 21, 2021.
+The deliberation was divided into 3 rounds of 2--3 days each.
+During each round, participants were shown a discussion prompt and were able to post a response to that prompt (Table \ref{tab:prompts}).
+Participants were able to view and comment on the posts of other participants assigned to the same pod.
+Participants were also able to view posts and comments that were visible to them in previous rounds, but could only interact or reply to posts and comments in the current round.
+Before and after each round, participants completed a ranked-choice poll regarding the deliberation topic, allowing the evolution of their preferences to be tracked over the course of the experiment.
+
+\begin{table}
+\center
+\label{tab:prompts}
+\begin{tabular}{|p{0.5in}|p{4.5in}|}
+\hline
+Stage & Prompt \\
+\hline
+1 &
+Which proposals do you prefer, and why?
+If there are disagreements, try asking questions to understand the perspective
+of others and to understand the causes of the disagreement.
+\\
+\hline
+2 & In the previous round of discussion, what opinions and reasoning did you
+observe in your group?
+How much agreement was there?
+Were there any disagreements or conflicts?
+If so, what were the sources of conflict, and how might they be resolved?
+\\
+\hline
+3 &
+What seem to be the most popular opinions?
+Do you agree with them?
+Has your opinion changed over the course of the discussion?
+\\
+\hline
+\end{tabular}
+\caption{Prompts shown to participants during each round of deliberation.}
+\end{table}
+
+Participants were recruited from students in an undergraduate university course and received extra credit for participating. 65 participants enrolled and completed at least one round of deliberation.
+Basic demographic information was collected and is shown in Table \ref{tab:demographic}.
+
+\begin{table}[]
+    \centering
+    \begin{tabular}{l l c c}
+\hline
+\multicolumn{2}{l}{Demographic} & Control & Random-Pod\\
+\hline
+Age & 18--24 & 30 & 31 \\
+& 25--29 & 1 & 0 \\
+& 30--34 & 0 & 1 \\
+& Non-disclosed & 2 & 0 \\
+\hline
+Gender & Man & 18 & 14 \\
+& Non-binary & 0 & 1 \\
+& Woman & 13 & 17 \\
+& Non-disclosed & 2 & 0 \\
+\hline
+Race & Asian & 14 & 16 \\
+& Black or African American & 1 & 1 \\
+& Hispanic & 1 & 0 \\
+& White & 13 & 11 \\
+& White, Asian & 1 & 4 \\
+& Non-disclosed & 3 & 0 \\
+\hline
+    \end{tabular}
+    \caption{Demographic statistics for experiment participants.}
+    \label{tab:demographic}
+\end{table}
+
+Participants were assigned to one of two conditions at the time of enrollment.
+Assignments were uniformly at random.
+Stratified sampling would be preferable in order to ensure even demographic distributions, but was not feasible due to software and time constraints.
+However, we manually confirmed that each condition had comparable demographics.
+The two conditions were:
+\begin{description}
+\item[Control]{
+Participants in the control condition engaged in conventional deliberation in a single large group.
+Posts and comments by participants in the control condition were visible
+to all others in the control condition for the duration of the deliberation.}
+\item[Random-Pod (Network Deliberation)]{
+Participants in this condition were divided into small pods ($\leq 5$ participants).
+Posts and comments made by participants in this condition were only visible to others in their current pod.
+Participants in this condition were assigned to a new pod at the beginning of each round of deliberation using the random-pod assignment method, producing a structurally efficient communication network.}
+\end{description}
+
+\subsection{Deliberation Topic}
+The policy issue chosen as the topic of deliberation is a crucial component
+of the experiment.
+The topic was chosen carefully according to several criteria:
+\begin{itemize}
+\item Relevant to the experimental population;
+\item Amenable to participants changing their preferences based on new information or reasoning;
+\item Amenable to a predefined list of proposed solutions;
+\item Sufficiently complex to have three or more proposed solutions;
+\item Timely, but unlikely to be influenced by current events during the deliberation.
+\end{itemize}
+
+Participants were students in a university undergraduate course and deliberated on the topic and format of one section of their final exam.
+As the outcome of the final poll determined the content of a real portion of their final exam, participants had an intrinsic incentive to advocate for their true preferences.
+The question and alternatives presented to participants were:
+
+\begin{displayquote}
+
+The 2021 [redacted] final exam will include a section worth up to 10\% based on the content covered during the first part of the semester.
+Which of the following options should be chosen for the topic and format of that section of the exam?
+
+\begin{description}
+\item[Prop. 1]{Open-ended with partial credit (2 questions, 5 points each) - Ch. 1-2: Intro to networks.}
+\item[Prop. 2]{Open-ended with partial credit (2 questions, 5 points each) - Ch. 3: Bridges, clustering, triadic closure, strong and weak ties, and the Strong Triadic Closure Property.}
+\item[Prop. 3]{Multiple choice with no partial credit (5 questions, 2 points each) - Ch. 4: Homophily, selection, social influence, and social-affiliation networks.}
+\item[Prop. 4]{Multiple choice with no partial credit (5 questions, 2 points each) - Ch. 5: Signed networks and balance theory.}
+\item[Prop. 5]{True/false with no partial credit (10 questions, 1 point each) - Ch. 6: Game theory, best responses, and Dominant Strategies, Nash equilibrium.}
+\item[Prop. 6]{True/false with no partial credit (10 questions, 1 point each) - Ch. 9:  Auctions.}
+\end{description}
+\end{displayquote}
+
+\subsection{Experimental Platform}
+\label{sec:exp-platform}
+
+We have modified the free and open-source Loomio platform \cite{jackson_open_2016} to implement network deliberation.
+Loomio is a widely-used online deliberation platform which provides both forum functionality (Figure \ref{fig:reply}) and various forms of voting, including ranked-choice (Figure \ref{fig:vote}).
+Loomio is popular among worker-owned cooperatives and grassroots organizations (Loomio itself is a worker-owned cooperative).
+Loomio was chosen for its combination of built-in functionality, existing user base, and open-source extensibility.
+
+The modified platform implements both random-pod and long-path assignment methods and automatically reassigns participants when the stage is advanced.
+In addition to network deliberation features, we have made several modifications used by the experiment.
+Upon signing in for the first time, participants are randomly assigned to a treatment condition.
+Participants are also assigned an alias which they will use throughout the deliberation (Figure \ref{fig:alias}).
+
+\begin{figure}
+\label{fig:reply}
+\center
+\frame{\includegraphics[width=5in]{chapters/figures/NetDelibExp/fig-reply.png}}
+\caption{Participants can perform standard forum actions, such as posting and commenting.}
+\end{figure}
+
+\begin{figure}
+\label{fig:vote}
+\center
+\frame{\includegraphics[width=5in]{chapters/figures/NetDelibExp/fig-vote.png}}
+\caption{Participants are presented with a ranked-choice voting screen between rounds of deliberation.}
+\end{figure}
+
+\begin{figure}
+\label{fig:alias}
+\center
+\frame{\includegraphics[width=3in]{chapters/figures/NetDelibExp/fig-alias.png}}
+\caption{Participants are assigned an alias for the duration of the experiment.}
+\end{figure}
+
+\section{Results}
+\label{sec:res}
+
+\subsection{Preferences and Outcome}
+\label{sec:res-outcome}
+
+We report the outcomes of several voting methods over the course of the deliberation in Table {\ref{tab:outcome} and show the evolution of first-choice vote distributions in Figure \ref{fig:firstplace}.
+The pre-deliberation poll shows similar initial preferences across the control and random-pod groups.
+A Condorcet winner exists in both groups, and in both cases that winner is Prop. 2.
+Prop 2. also claims a plurality of first-choice votes in both groups.
+The only notable difference is that in the control group, Prop.
+1 wins under the Borda Count method, suggesting that the contest between Props.
+1 and 2 are very close and that participants who did not list Prop.
+1 in first place still ranked it highly.
+
+\begin{table}[]
+    \centering
+    \begin{tabular}{c c c c c c c}
+\hline
+& \multicolumn{3}{c}{Control} & \multicolumn{3}{c}{Random-Pod} \\
+\cmidrule(r){1-1}
+\cmidrule(lr){2-4}
+\cmidrule(l){5-7}
+Round & Condorcet & Plurality & Borda &
+Condorcet & Plurality & Borda \\
+\cmidrule(r){1-1}
+\cmidrule(lr){2-4}
+\cmidrule(l){5-7}
+
+0 & prop2 & prop2 & prop1 & 
+prop2 & prop2 & prop2 \\
+
+1 & prop1 & prop1 & prop1 &
+prop2 & prop2 & prop2 \\
+
+2 & prop1 & prop1/prop2 & prop1 &
+prop2 & prop2 & prop2 \\
+
+3 & prop1 & prop1 & prop1 &
+prop2 & prop2 & prop2 \\
+\hline
+    \end{tabular}
+    \caption{Referendum outcome after each round of deliberation (initial=0).}
+    \label{tab:outcome}
+\end{table}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=3in]{chapters/figures/NetDelibExp/fig-counts.pdf}
+    \caption{Number of first-choice votes for each alternative over the course of the deliberation.}
+    \label{fig:firstplace}
+\end{figure}
+
+Over the course of the deliberation, the preference profiles of the control and random-pod groups show notable differences.
+In the control group, Prop. 1 becomes the Condorcet winner after round 1 and wins across all other voting methods.
+Prop 1. remains the winner in the control group throughout the rest of the deliberation, with one exception: Prop. 1 and Prop. 2 tie for plurality voting in round 2.
+In the random-pod group, Prop. 2 remains the winner across all voting methods throughout the entire deliberation.
+We compare participants' preferences before and after the deliberation in Figure \ref{fig:countmatrix}.
+The most striking feature is that a number of participants (4) in the control group who initially ranked Prop. 2 highest then went on to rank Prop 1. the highest in the final poll.
+These participants are a significant driver of the switch from Prop 2. to Prop 1. in the control group.
+While the random-pod group does show some participants switching away from Prop. 2, those participants shift to the less popular Props. 3 and 4 rather than Prop 1.
+In the sections that follow, we will further analyze the evolution of participant preferneces and the potential mechanisms driving them.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=4in]{chapters/figures/NetDelibExp/fig-count-matrix.pdf}
+    \caption{Histogram of participants' initial (row) and final (column) first-choice votes.}
+    \label{fig:countmatrix}
+\end{figure}
+
+\subsection{Activity}
+\label{sec:res-activity}
+
+Here we describe the trends we observe in participants' deliberation activity and the potential impact on participant preferences.
+Overall, the nubmer of comments per participant shows a heavy-tailed distribution (Figure \ref{fig:commentcounts}) with a mode of one comment per participant.
+The random-pod group shows less of a heavy-tail, with more participants posting a single comment per round and a smaller maximum count (2--3 comments vs. 3--4 for control).
+The random-pod group thus shows a more equal division of activity across participants.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=6in]{chapters/figures/NetDelibExp/fig-comment-counts.pdf}
+    \caption{Histogram of the number of posts made by participants during each round of deliberation.}
+    \label{fig:commentcounts}
+\end{figure}
+
+Figure \ref{fig:commentrounds} shows the comment counts for each round as well as the Shannon entropy of participant comment counts.
+The random-pod group shows a lower total number of comments and a higher entropy, again showing a more equal division of activity across participants.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=6in]{chapters/figures/NetDelibExp/fig-comment-rounds.pdf}
+    \caption{The total number of comments in each round (left) and the Shannon entropy of participant comment counts (right).}
+    \label{fig:commentrounds}
+\end{figure}
+
+The more equal distribution of comments in the random-pod group is consistent with our expectation that network deliberation promotes equal participation.
+The reduction in overall participation is also notable and suggests low participation as a potential consideration in network deliberation settings.
+
+\subsection{Deliberation Content}
+\label{sec:res-content}
+
+The content of participant comments gives additional insight into how preferences were infuenced by deliberation.
+We pay special attention to comments pertaining to the two most popular alternatives: Props. 1 and 2.
+Preferences for these alternatives evolved differently between the control and random-pod groups.
+
+Comments across both groups frequently advocated for Props. 1 and 2 without specifying a preference between the two:
+\begin{displayquote}
+P09: Opened ended questions allow for partial credit while multiple choice and true false has no partial credit.
+I feel comfortable with chapters 1, 2, and 3 [Props.
+1 \& 2] as they are the easiest.
+I am least comfortable with auctions [Prop.
+6].
+\end{displayquote}
+However, some participants advocated for one or the other:
+\begin{displayquote}
+P64: I think Intro to Networks [Prop.
+1] should be chosen.
+I like the open ended question format of this, as it allows for partial credit.
+In addition to this, since these were the topics that were taught in the beginning they provide a base of everything we have learned in the class and as a result I think this material would be very helpful for it to be covered on our final exam.
+\end{displayquote}
+\begin{displayquote}
+P10: I prefer the open-ended questions on chapter 3 [Prop.
+2] because I feel like many of these types of questions would have several requirements, so it is very possible to receive partial credit if you partially understand.
+The main reason I choose chapter 3 [Prop.
+2] over chapters 1/2 [Prop.
+1] is because I don't know what questions from chapters 1/2 [Prop.
+1] would look like, and I know how to prepare for chapter 3 [Prop.
+2].
+\end{displayquote}
+While in the minority, some participants advocated for other proposals:
+\begin{displayquote}
+P43: I think the topics covered in chapter 6 [Prop. 5] should definitely be on the final. I'm personally indifferent between true/false.
+\end{displayquote}
+
+\subsection{Preference Evolution}
+
+As shown in Table \ref{tab:outcome} and Figure \ref{fig:firstplace}, both groups initially preferred Prop.
+1, with the control group switching to Prop.
+2 after round 1 while the random-pod group maintained its initial preference.
+We now analyze participant preferences and their evolution in more detail.
+
+\subsubsection{Agreement}
+\label{sec:res-agreement}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=6in]{chapters/figures/NetDelibExp/fig-agreement-fillattr.pdf}
+    \caption{Treatment-level agreement over the course of the deliberation.
+Agreement is calculated as the mean pairwise correlation between participants using either Kendall corellation (left) or weighted crossing correlation (right).}
+    \label{fig:agreement}
+\end{figure}
+
+We report the intra-group agreement throughout the course of the deliberation in Figure \ref{fig:agreement}.
+We calculate agreement using both the Kendall correlation
+and the weighted crossing correlation.
+The Kendall correlation is commonly used in social choice and weights contributions from each rank equally.
+The weighted crossing correlation places a higher weight on
+higher-ranked preferences.
+In both cases, the avarage pairwise agreement follows similar a similar pattern.
+Both the control and random-pod groups show an increase in agreement after the first round of deliberation,
+followed by a decrease.
+In the control group, the decrase is more rapid and more extreme than in the random-pod group.
+Using weighted crossing agreement, the difference-in-differences between control and random-pod groups is 0.68, noting that weighted crossing ranges from -1 to +1.
+
+Our results show that in our experiment, the random-pod group 
+experienced a small but robust increase in agreement,
+wile the control group experienced a sharp increase followed by an even sharper decrease.
+These results support the hypothesis H1 that network deliberation contributes to increased agreement relative to conventional single-group deliberation.
+However, this effect is due more to a decrease in agreement
+within the control group than an increase in the random-pod group.
+After an initial increase in agreement, both groups showed an unexpected ``rebound'' effect.
+In the control group, this rebound resulted in a net reduction in agreement, suggesting the presence of negative social influence.
+
+\subsubsection{First-Choice Votes}
+\label{sec:res-firstchoice}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=6in]{chapters/figures/NetDelibExp/fig-round-matrix.pdf}
+    \caption{Histograms of participants' first-choice votes before and after each round.}
+    \label{fig:roundmatrix}
+\end{figure}
+
+By focusing on first-choice votes and on the most popular alternatives (Props.
+1 \& 2), we can offer some insight into the differences in preference evolution we observe between the control and random-pod groups.
+Figure \ref{fig:roundmatrix} shows the evolution of first-choice preferences over the course of each round of deliberation.
+Initially, the control and random-pod groups had similar distributions of first-chioce votes for Props. 1 \& 2, with both clearly favoring Prop. 2 (13:8 for control and 13:5 for random-pod).
+However, first-round comments regarding these two proposals differed substantially between the two groups.
+In the control group, a majority of comments (7:2) favored Prop. 1 while in the random-pod group, a majority favored Prop. 2 (9:2).
+Remarkably, the support expressed for Prop. 1 by the control group during round 1 was counter to the group's initial preference for Prop. 2.
+After that round however, a majority of control group participants reported a first-choice preference for Prop. 1 over Prop. 2 (15:9).
+In the random-pod group, Prop. 2 remained the most popular and received an even higher share of first-choice votes (16:2) after round 1.
+
+In both the control and random-pod groups, we see round 1 deliberation clearly favoring one alternative, which then gains a number of first-choice votes, suggesting social influence.
+This social influence is further evidenced by the increase in agreement observed for both groups over the course round 1 (Figure \ref{fig:agreement}).
+
+The control group varies from the random-pod group in that the round 1 advocacy and subsequent preference shifts are opposite to the original preferences.
+In other words, there is evidence of an information cascade flipping the control group's preference from Prop. 2 to Prop. 1 over the course of round 1.
+While many factors contribute to the formation of an information cascade, we suggest that network deliberation structure of the random-pod group may have reduced the liklihood of such a cascade in that group.
+The pod structure limits how many participants can be exposed to influence from any particular individual during a single round.
+Furthermore, the more even distribution of participation observed in the random-pod group provides participants with more accurate information on the preference distribution of ther peers.
+The discussion is less skewed towards highly-vocal participants.
+
+In rounds 2 and 3, the preferences for Props. 1 \& 2 continued to evolve differently for the control and random-pod groups (Figure \ref{fig:roundmatrix}).
+In both cases, participants partially reverted towards their initial preferences.
+In the control group, particularly during round 2, Prop 1. lost many of its round 1 gains to Props. 3--6, although it still remained the most popular first-choice vote,
+and Prop 2. retained a sizeable minority of first-choice votes.
+As a result, the control group exhibited a divergence of preferences after round 2, contributing to the drop in agreement seen in Figure \ref{fig:agreement}.
+In the random-pod group, many of the participants who had switched from Prop. 1 to Prop 2. after round 1 switched back,
+partially reverting the winning margin of Prop 2. towards its original value and contributing to a reduction in agreement.
+However, the observed drop in agreement values can't be explained by first-choice votes alone.
+The two groups show similar distributions of first-choice vote counts (although assigned to different alterantives) but a notably different level of overall agreement.
+
+\subsection{Social Influence}
+\label{sec:res-socialinfluence}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=6in]{chapters/figures/NetDelibExp/fig-hipness-conformity.pdf}
+    \caption{Conformity and hipness by participant, calculated using initial and final preferences. Values are sorted in descending order for ease of comparison.}
+    \label{fig:conformity}
+\end{figure}
+
+To better understand the role of social influence in the evolution of preferences, we report the conformity and hipness distributions in Figure \ref{fig:conformity}.
+Summing either of these quantities over all participants yields the overall change in mean agreement: -0.597 for control and +0.049 for random-pod.
+
+In the control group, the participants with negative conformity outnumber those with positive conformity:
+more participants moved away from others' preferences than toward them, resulting in a negative shift in agreement.
+In the random-pod network deliberation group, the number of participants with positive conformity is slightly larger than the number with negative conformity, leading to a slightly positive change in agreement.
+The difference between the two groups is most pronounced among small magnitude conformity participants: those who did not change their preferences much over the deliberation.
+
+Alternatively, we can analyze the change in agreement in terms of hipness.
+The negative change in agreement in the control group can be attributed primarily to a small number of participants with highly-negative hipness.
+These participants did not change their preferences much over the course
+of the deliberation,
+but other participants' preferences moved away from theirs considerably.
+Among participants in the random-pod network deliberation group,
+those with negative hipness have a notably smaller magnitude,
+resulting in more agreement.
+
+\subsubsection{Dynamics of Social Influence}
+\label{sec:res-dynamics}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=6in]{chapters/figures/NetDelibExp/fig-influence-bar.pdf}
+    \caption{Participant contributions to conformity and hipness at each round.}
+    \label{fig:influence-bar}
+\end{figure}
+
+Figure \ref{fig:influence-bar} shows the conformity and hipness contributions of individual participants after each round
+(relative to initial preferences).
+The sum of positive and negative contributions (equivalent to the mean difference in agreement) is overlaid.
+In the control group, conformity begins mostly positive and high-magnitude after the first round, but switches to mostly negative in subsequent rounds.
+In contrast, the random-pod group shows a combination of positive and negative conformity throughout the experiment, with positive conformity maintaining a small majority.
+These observations suggest that participants in the control group first conformed towards initially popular preferences, but shifted away from them after the second round of deliberation.
+While in the random-pod group, participants showed diverse changes in perference with some convergence towards initially-popular preferences.
+
+Analyzing hipness suggests a similar picture.
+The control group shows overwhelmingly positive hipness after round 1, with a switch to primarily negative hipness in subsequent stages.
+The random-pod group shows primarily positive hipness throughout most of the experiment, although both positive and negative values are small in magnitude.
+So in the control group, changes in preference are divided unevenly between participants.
+At first, these changes conform towards the more fixed participants, but after round 2 diverge away from them.
+In constrast, within the random-pod group, changes in preference are distributed more evenly and consistently converge towards the more fixed participants.
+
+\subsubsection{Interpretation of Social Influence}
+\label{sec:res-interpretation}
+
+Combining the above insights from conformity and hipness measures suggests the following.
+The negative shift in agreement in the control group is due
+to a majority of participants shifting their preferences away (negative conformity) from a samll number of stubborn participants (highly-negative hipness) without converging on an alternative (lack of positive conformity).
+Furthermore, this shift occurs after round 2, following an initial stage of high-conformity.
+The random-pod participants also exhibit stubborn participants, but fewer and with smaller magnitude,
+and the the majority of participants converge toward each other's preferences (positive conformity).
+
+Our findings show that the random-pod network deliberation group experienced a positive change in agreement relative to the conventional deliberation control group.
+The above analysis of individual preference shifts suggests that the decreased agreement in the control group is due to social influence leading a number of participants to shift from more popular preferences to a divergent set of less popular preferences.
+
+\subsection{Survey}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=6in]{chapters/figures/NetDelibExp/fig-survey.pdf}
+    \caption{Participant post-experiment survey responses on a 5-point Likert scale.}
+    \label{fig:survey}
+\end{figure}
+
+\begin{table}[]
+    \small
+    \centering
+    \begin{tabular}{l c c c}
+\hline
+Question & \multicolumn{2}{c}{Median} & p \\
+\cmidrule(lr){2-3}
+& Control & Random-Pod & \\
+\hline
+Q1: Participants always agreed with each other
+& 2 & 3 & 0.070 \\
+Q2: Participants reached agreement by the end of the discussion
+& 3 & 4 & 0.015 \\
+Q3: I changed my mind over the course of the discussion
+& 3 & 4 & 0.663 \\
+Q4: I was exposed to new information or perspectives
+& 4 & 4 & 0.068 \\
+Q5: I was comfortable expressing my opinions and ideas
+& 5 & 5 & 0.181 \\
+Q6: I enjoyed the discussion
+& 4 & 4 & 0.722 \\
+\hline
+    \end{tabular}
+    \caption{Median responses for post-experiment survey (5-point Likert scale).
+Reported p values are for a two-sided Mann-Whitney test.
+The Šidák correction places the threshold at $p < 0.0085$ for a FWER of less than $\alpha=0.05$.}
+    \label{tab:survey}
+\end{table}
+
+The results of the post-experiment survey are shown in Table \ref{tab:survey} and Figure \ref{fig:survey}.
+After correcting for multiple comparisons using the Šidák correction, the Mann-Whitney test finds no significant difference between the control and random-pod groups.
+However the observed differences (and lack of differences) in median suggest a starting point for future work.
+We expected the random-pod group to be better at identifying and resolving disagreements, which would result in a lower score for Q1 and a higher score for Q2.
+Instead, we observe a higher median score for both.
+Given that initial preference profiles were similar for the two groups, the lower perception of agreement in the control group could be driven by a number of mechanisms, such as higher visibility of disagreements or polarization over the course of the deliberation.
+The difference in perceptions of agreement across conventional and network deliberation could be a fruitful area for future research.
+
+\section{Discussion}
+\label{sec:discussion}
+
+We hypothesized (H1) that network deliberation tends to reach a higher level of agreement than conventional single-group deliberation.
+The observed agreement calculated from polls throughout the experiment experiment supports this hypothesis (Section \ref{sec:res-agreement}).
+However, our hypothesis was based on the assumption that network deliberation would more effectively identify and resolve disagreements,
+leading us to expect a large increase increase in agreement in the random-pod group.
+Instead, we observe a small increase of agreement within the random-pod group and a large decrease within the control group.
+So while we do see more of an improvement in agreement under network deliberation, our observations are more consistent with a protective effect against a disagreement-producing mechanism, than with the originally hypothesized agreement-producing mechanism.
+
+We also sought to understand how preferences evolve during network deliberation vs. conventional deliberation (RQ1).
+We focused our analysis on first-choice votes in particular (Section \ref{sec:res-firstchoice}).
+We also quantified the observed social influence throughout the experiment (Section \ref{sec:res-dynamics}).
+In both the control and random-pod group, we observed preferences shifting in favor of alternatives that received overwhelming support in round 1 deliberation comments.
+These comments were consistent with initial preferences in the random-pod group, but counter to initial preferences in the control group.
+An information cascade, in which participants conformed to a less popular opinion believing it to be more popular, is one likely mechanism.
+The skewed distribution of activity in the control group (Section \ref{sec:res-activity}), as well as the large number of negative-conformity participants (Section \ref{sec:res-socialinfluence}) are both consistent with an information cascade in the control group.
+In network deliberation, participants are limited to interacting with others in the same pod, creating a barrier to the spread of information cascades.
+Thus, the network deliberation structure may explain why round 1 comments in the random-pod group were consistent with initial preferences, i.e., did not show evidence of an information cascade.
+
+It is interesting to note that while the control group showed considerable positive conformity in round 1, most participants switched to negative conformity in later rounds, while in the random-pod group, positive conformity was relatively consistent.
+In other words, the preferences of the control group converged in round 1, but diverged in later rounds, resulting in the observed sharp drop in agreement (Figure \ref{fig:influence-bar}).
+As nothing changed in the control group between the first and subsequent rounds, it is interesting to consider why the conformity switched from mostly positive to mostly negative, e.g., why preferences diverged during round 2 after converging in round 1.
+Possible explanations include a ``rebound'' effect from an information cascade or a time-dependent effect only emerging in later rounds.
+
+Motivated by prior literature identifying conflict discovery and resolution as a benefit of participatory decision-making, we also considered the effect of network deliberation on conflict resolution (RQ2).
+Unexpectedly, we found that the random-pod group provided only a small increase in agreement, but that the control group actually produced disagreement by the end of the deliberation.
+Interestingly, the first-choice votes show a similar level of agreement in the two groups, so the control group disagreement is largely due to second-choice votes and below.
+One possible explanation is that participants disagreeing on first-place votes became more polarized against each other, lowering each other's top-choice in their own preferences, but additional research is necessary to test this hypothesis.
+
+We note that in this experiment, the skewed distribution of participation in the control group acted to amplify proponents of a minority preference, which later became the majority preference in a possible information cascade.
+However, a skewed distribution of participation could potentially have the opposite effect of further amplifying an already majority preferences and overshadowing minority preferences.
+A better understanding of the factors determining which preferences are amplified would contribute to an understanding of when network deliberation is preferable to conventional.
+
+We also note that the overall level of participation was lower in the random-pod group, very likely due to the smaller group size.
+The low level of activity could be one factor contributing to the lack of conflict identification in the random-pod group.
+It is thus desirable to better understand how factors including group size contribute to activity level.
+
+\section{Conclusion}
+\label{sec:ch04conclusion}
+
+We have reported the results of a controlled experiment comparing an efficient network deliberation network to conventional single-group deliberation.
+We found that network deliberation produced a more positive change in agreement consistent with our hypothesis.
+However, we found that this change was due primarily to a decrease in agreement in the conventional group, suggesting a mechanism that protects against disagreement rather than facilitating agreement.
+We find evidence of an information cascade in the conventional group strong enough to alter the Condorcet winner.
+Our findings suggest that protection against information cascades may be one mechanism underlying the success of network deliberation in large-scale collaborations.

### 2022-05-05 UMich/chapters/ch01-intro.tex
Import NetDelibExp and formatting fixes.
Update intro.

@@
 system, which power much of the modern internet
-\cite{coleman_coding_2013, benkler_coases_2002, raymond_cathedral_1999}.
+\cite{coleman_coding_2012, benkler_coases_2002, raymond_cathedral_1999}.
 Social movements such as the Arab Spring, Occupy, Black Lives Matter, and
@@
 To bridge the limitations of the above methods,
-this dissertation will include a controlled field experiment evaluating the
+this dissertation includes a controlled field experiment evaluating the
 effect of network deliberation in real-world collective decision-making
@@
 
-Chapter \ref{chap:experiment} proposes and describes current progress on a
-field experiment to study network deliberation in large-scale human collaborations.
+Chapter \ref{chap:experiment} describes an experimental study network deliberation in large-scale human collaborations.
 The study design uses periodic ranked-choice polls to track individual preferences
@@
 preferences,
-this experiment evaluate the ability of network deliberation to resolve conflict
+this experiment evaluates the ability of network deliberation to resolve conflict
 and build consensus, relative to conventional single-group deliberation.
-
-\section{Dissertation Progress and Timeline}
-At the time of this proposal,
-the analysis of English-language WikiProjects described in Chapter
-\ref{chap:wp-prod-perf} has been completed and resulted in a published
-paper \cite{platt_network_2018}.
-The agent-based modelling project described in Chapter \ref{chap:abm}
-has been completed and resulted in a finished manuscript.
-I have completed custom software, obtained IRB exemption, and conducted
-several pilot studies for the field experiment described in
-Chapter \ref{chap:experiment}.
-I expect to have data collected from at least one full-scale experiment
-by the end of 2021, and to have the data analyzed by mid-February, 2022.
-I plan to have my dissertation completed by mid-May, 2022.
+The primary contributions are:
+\begin{itemize}
+\setlength\itemsep{0pt}
+    \item Findings support the hypothesis that network deliberation is better at facilitating agreement than conventional deliberation.
+    \item Findings suggest that network deliberation can provide protection against information cascades.
+    \item Findings include no evidence that network deliberation facilitates substantial conflict-resolution, but indicate that it may provide protection against polarization.
+\end{itemize}

### 2022-05-05 UMich/chapters/ch05-discussion.tex
Update discussion.

\epigraph
{Socialism would take too many evenings.}
{---Attributed to Oscar Wilde by Michael Walzer \cite{walzer_day_1971} }
Large-scale participatory governance is, to many, an oxymoron.
The assumption that large-scale coordination requires coercive hierarchy is so pervasive that it is common to see the words ``hierarchy'' and ``organization'' used interchangeably, as in the conventional ``org chart.''
Control hierarchies \cite{crumley_heterarchy_1995}, including representative democracy, are indeed very effective at addressing some of the challenges of large-scale governance.
Faced with the historical difficulty of achieving both large-scale participation and deliberative discourse \cite{ackerman_deliberation_2002}, control hierarchies defer governance of large groups to a subset of members, enabling more deliberative discourse and more rapid decision-making than traditional consensus deliberations \cite{gentry_consensus_1982}.
Hierarchies can also serve to establish consistency and preserve institutions, as the priorities of elites shift from maximizing group welfare to sustaining their positions, as described by the iron law of oligarchy \cite{michels_political_1999, shaw_laboratories_2014}.
However, the benefits of control hierarchy come with costs.

In terms of principles, hierarchical organizations rely on coercion and power imbalances, sacrificing egalitarianism.
More practically, when information is distributed but decision-making is centralized,
that decision-making necessarily excludes potentially useful information.
The emergence of successful large-scale, internet-enabled, parrticipatory collaborations suggests that perhaps, at leaast in some cases, neither the principle of egalitarianism nor the practical wisdom of the crowd needs to be sacrificed to achieve effective coordination and collaboration.
The key question I have attempted to address is: how do such large-scale participatory projects achieve success?
I have specificaly focused on one challenge facing such collaborations: how can a group collaborate effectively when it is impractical for all members to communicate and interaction is restricted to a network.

In the preceeding chapters, I have approached the above question using observational study, numerical simulation, and experiment.
A combination of literature reivew and obersvational analysis of English-language Wikipedia (\ref{chap:wp-prod-perf}) identified a common theme among successful examples: small, tightly-knit groups, interconnected through mutual membership: network deliberation.
Subsequent work investigates simple models of network deliberation to better understand how and why it helps large groups effectively self-govern.
These models are not intended to capture all properties of real-world network deliberation, but rather to isolate and investigate particular factors, such as social influence and network efficiency.
Both numerical simulations (Chapter \ref{chap:abm}) and experiment (Chapter \ref{chap:experiment}) find evidence that network deliberation can improve the effectiveness of large deliberations relative to conventional single-group deliberation.
In particular, these studies both identify a protective effect against the negative consequences of social influence, such as information cascades.
As such, network deliberation may be especially beneficial in settings with a high level of social influence.

\section{Future Work: Deliberative Settings}
While the studies presented here confirm the observation that network deliberation sometimes contributes to the success of large-scale deliberation, there are many factors that can influence the deliberative process, and there remains much work to be done to understand how these factors interact with network deliberation.

\subsection{Language and Culture}
The observational and experimental studies presented focus on the English-lanaguage setting, as well as on subsets of the population who are comfortable with wiki technology and/or attend university.
It is very possible that cultural norms, level of comfort and familiarity with technology, and language barriers between group members could have important consequences for large-scale deliberation that interact with the benefits of network deliberaiton.

\subsection{Modes of Deliberation}
There are many ways to design a deliberative interaction, whether online or offline. Any of the following considerations could have consequences for the successful implementation of large-scale deliberation.
\begin{description}
	\item[Synchronous/Asynchronous:] Deliberative activity can vary substantially between synchronous and asynchonous settings. Synchronous settings allow more interactivity, but can potentially include/exclude members based on availability and schedule.
	\item[Simultaneous/Sequential:] The simplified models studied here assigned individuals to one pod at a time in order to constrain network structure and study the dynamics of preference evolution, but real-world collaborations may have individuals participating in multiple groups simultaneously.
	\item[One/Many Topics:] All pods might be focused on the same topic, or each might be focused on a different topic (or sub-topic).
	\item[Formal/Informal:] Deliberations can vary widely in their level of formality. While intentionally-planned formal deliberations may be impactful, spontaneous light-hearted deliberations could be just as impactful. It is interesting to consider how intentional network structure might be facilitated in spontaneous intereactions.
	\item[Public/Private:] Deliberations can take place in settings with a range of privacy. Perception of privacy, as well as comfort and familiarity with other participants may impact the effectiveness of deliberation.
\end{description}

\subsection{Facilitation and Moderation}
When all members are equal participants in a deliberation, facilitation and moderation are a challenge.
In the experiment reported in Chapter \ref{chap:experiment}, facilitation was achieved through discussion prompts.
Moderation was performed by an experimenter, but the discussion remained respectful and no moderation actions were required.
An active facilitator might contribute subtantially to the quality of a deliberation, but there are many potential ways to choose a facilitator from among a group of peers.
Similarly, moderation is incredibly important for productive deliberation on sensitive topics, and it is not immediately obvious how moderation might work in a network deliberation environment.

\subsection{Emergent Alternatives}
In the simple model of network deliberation presented here, the available alternatives were determined before the beginning of the deliberation.
However, the deliberative process can produce new ideas and these ideas are sometimes improvements on those previously available \cite{salganik_wiki_2015}.
Additional research is necessary to undertand the consequences of introducing new ideas into a network deliberation, as well as how and when to introduce those ideas to different pods for best effect.

\subsection{Preference Distribution}
The initial distribution of preferences could have substantial impact on the outcome of network deliberation.
Additonal research is needed to address the many possible scenarios.
For example, if a preference is held by a small enough fraction of participants, they are likely to be assigned to different pods and perceive themselves as the only one with that preference.
Network deliberation creates a high likelihood that such a participants will eventually
be assigned to a pod with others sharing their preference, but in strong social influence settings, they might switch their preferences before then.

\section{Applications of Network Deliberation}
The studies of network deliberation presented here have a double purpose.
The first purpose is to better understand the role of network structure in successful large-scale deliberations such as Wikipedia.
The second is to identify principles that can be used in the analysis and creation of new sociotechnical systems.
This work starts from empirical observations of collaborations that have done something unusual and attempts to identify robust principles that might be applied to transform existing collaborations or create new ones.
The repeated appearance of small interlocking groups in successful large-scale participatory collaborations suggests that network deliberation may be such a principle.
The applicability of network deliberation depends on the ability to exert influence over network structure.
It may seem counterintuitive or even impossible that collaborations and organizations might intentionally alter their interpersonal network structure.
But in fact, countless practices such as job interviews, letters of recommendation,
performance reviews, internships, mentorships, etc. all serve exactly this purpose;
to deliberately shape the social network within a collaboration.
By analogy, the task of deliberately designing a neighborhood might seem impossible,
but urban planners have made a science of it.

One possible application of network deliberation principles is in the design of organizational structures and procedures.
In fact, small interlocking pods are a feature found in self-managed organizations \cite{laloux_reinventing_2014}.
As the pracitices and structures of such organizations are codified and refined, a better understanding of network deliberation can provide a guide to consructing more effective organizations and institutions.

Similarly, network deliberation emerges naturally in technical settings such as working groups in large collaborations (e.g., WikiProjects), federated social networks (e.g., mastodon), and group chat platforms (e.g., discord).
The deliberation that takes place on these platforms is shaped in part by the protocols and user interaces of the software.
The principles of network deliberation have the potential to inform the design of such protocols and interfaces in a way that minimizes harm and promotes effective interactions.

\section{Conclusion}
Inexpensive, instantaneous, global, peer-to-peer communication has only been possible for a tiny sliver of recent human history.
Existing organizational and institutional structures have evolved over hundreds or thousands of years.
While the full implications of digital communication for human institutions can't be predicted, our institutions have already started adapting in exciting ways.
In particular, we have seen the emergence of very large collaborations with participatory governance, which would have seemed impossible before the internet.
In the research presented here, I have empirically identified one of the commonalities between these unprecedented projects, namely interlocking networks of small pods: network deliberation.
The agent-based simulation and experiment presented here study the properties of a simplified model of network deliberation, in order to better unerstand the principles and mechanisms at work.
This research contributes knowledge on the interaction of network deliberaiton with factors such as network efficiency, individual behavior, and in particiular, social influence.
It is my hope that this work will help enable large groups to collaborate on common goals in a manner that allows all members to contribute and allows all voices to be heard.

### 2022-05-05 UMich/chapters/ch05-discussion.tex
Edits.

@@
 This research contributes knowledge on the interaction of network deliberaiton with factors such as network efficiency, individual behavior, and in particiular, social influence.
-It is my hope that this work will help enable large groups to collaborate on common goals in a manner that allows all members to contribute and allows all voices to be heard.
+It is my hope that this work will help large groups collaborate on common goals in a manner that allows all members to contribute and allows all voices to be heard.

### 2022-05-05 UMich/front_materials/abstract.tex
Edits.

--- 2022-05-02 Proposal/abstract.tex
@@
-As groups grow in size, they gain access to additional resources,
-creating opportunities for collective intelligence and collective action.
-However, at very large scales, group decision-making becomes prohibitively slow and difficult
-to coordinate.
-Traditional solutions include representative decision-making and/or a shift from deliberation
-to voting.
+As groups grow in size, they gain access to additional resources, creating opportunities for collective intelligence and collective action.
+However, at very large scales, group decision-making becomes prohibitively slow and difficult to coordinate.
+Traditional solutions include representative decision-making and/or a shift from deliberation to voting.
 Both approaches sacrifice desirable properties.
-Representative decision-making loses the potential benefits of collective intelligence and
-introduces hierarchies that may place the interests of specific individuals ahead of the
-interests of the group.
-Voting sacrifices generativity: allowing a choice between predefined options,
-without allowing for improvement to those options.
+Representative decision-making loses the potential benefits of collective intelligence and introduces hierarchies that may place the interests of specific individuals ahead of the interests of the group.
+Voting sacrifices generativity: allowing a choice between predefined options, without allowing for improvement to those options.
 Arrow’s impossibility theorem also fundamentally limits the fairness of voting.
-This project proposes Networked Deliberation as a potential means of large-scale decision-making.
-In Networked Deliberation, members of a large group are repeatedly partitioned into small
-deliberative pods.
-Overlap between pods at different stages enables group-wide diffusion of information and
-preferences.
+This project proposes Networked Deliberation as a potential means of effective large-scale decision-making.
+In Networked Deliberation, members of a large group are repeatedly partitioned into small deliberative pods.
+Overlap between pods at different stages enables group-wide diffusion of information and preferences.
 Different methods for assigning members to pods result in different network topologies.
-In partnership with large community groups, this project will use a web-based platform to
-evaluate Network Deliberation.
-For a variety of network topologies, the speed and quality of preference convergence will be
-evaluated using quantitative information from ranked-choice polls throughout the process.
-Deliberator satisfaction and experience will also be evaluated using qualitative analysis of
-deliberative text and surveys.
-This work seeks to improve the ability of very large groups to quickly develop a consensus,
-enabling a more effective use of shared resources to achieve common goals.
+This work combines observational study, agent-based modeling, and experiment to evaluate and better understand network deliberation.
+An empirical observation of WikiProjects on the English-language Wikipeida identifies the network properties of the most effective projects.
+A simple agent-based model of Network Deliberation shows improvements over conventional deliberation in the presence of strong social influence.
+Finally, a controlled experiment studies network deliberation in a real world setting, tracking how individual preferences evolve, and finding evidence that network deliberation provides protection against negative consequences of social influence.
+This work seeks to improve the ability of very large groups to quickly develop a consensus, enabling a more effective use of shared resources to achieve common goals.

### 2022-05-05 UMich/front_materials/acknowledgements.tex
Edits.

Thanks to my advisor, Daniel M. Romero for his time, patience, and guidance.
Thanks to my committee:
Scott E. Page,
Tawanna Dillahunt,
Ceren Budak,
for their time, expertise and feedback.
Thank you to Hermino Bodon for collaboration on the agent-based modeling code in Chapter 3.
Thanks to Jane Im for collaboration on interviewing the participants in the experiment described in Chapter 4.
Thanks to Danielle Livneh, and Karthik Ramanathan for help collecting WikiProjects data.
Thanks to professors Nicole Ellison, Yan Chen, Tanya Rosenblat, and Patrick Grim for feedback on projects and proposals that led to the work in this dissertation.
Thanks to professors Grant Schoenbeck, Kentaro Toyama, and Charles R. Severance for feedback and encouragement on various projects.
Thanks to Veronica Falandino for making UMSI a kinder place.
Thanks to Linda Wood and Mita Gibson of the Center for the Study of Complex Systems always being willing to help.
Thanks to the members and staff of the Graduate Employees Organization, AFT Michigan 3550, especially Denise Bailey and Liz Ratzloff, for advocating for the wellbeing of Michigan graduate students.
Thanks to Avdi Grimm for consulting on some very difficult Ruby bugs.
Thanks to Salem Elzway for technical writing feedback.
Thanks to the many Michigan graduate students who have given me feedback and guidance on work and classes:
Jeffrey W. Lockhart,
Danaja Maldeniya,
Lia Bozarth,
Christopher Quarles,
Chanda Phelan,
Padma Chirumamilla,
Earnest Wheeler,
Sam Carton,
Elliott Brannon,
Tawfiq Ammari,
and the 2015 UMSI PhD chohort,
in particular Jean Hardy and Cindy Lin.
Thanks to attendees of the May 25, 2017 MIT Center for Civic Media lab meeting,
the Berkman-Klein Center Cooperation Working Group,
and the 2019 CSCW Workshop on Team and Group Diversity
for helpful feedback.
Thank you to
Rahul Bhargava,
Ethan Zuckerman,
Sasha Costanza-Chock,
Willow Idlewild,
Charlie DeTar,
Gerald J. Sussman,
Anthony Garnock-Jones,
and Elaine Lee
for guidance and encouragement in returning to academia.
Thanks to the members of i3 Detroit past and present for inspiring my interest in participatory governance.
Thanks to Persephone Hernandez-Vogt for their companionship and encouragement.
Thanks to Andy Brosius for sharing their life and family with me, and for their many years of patience and encouragement.
This research was partly supported by
the National Science Foundation under Grant No. IIS-1617820,
Air Force Office of Scientific Research under Award No. FA9550-19-1-0029,
and the UMSI REMS program,

### 2022-05-11 UMich/chapters/ch01-intro.tex
Update discussion.
Last edits before final draft.

@@
+\epigraph
+{Pulling together is the aim of despotism and tyranny. Free men pull in all kinds of directions.}
+{---Lord Vetinari in {\em The Truth}, Terry Pratchett}
+
 The ability of large groups to reach mutually agreeable decisions is key to
-democratic governance, social movements, and peer production (e.g., Wikipedia).
+democratic governance, organizational effectiveness, social movements, and peer production.
 Faced with the intractability of large-scale decision-making, traditional
@@
 \cite{tufekci_twitter_2017, gonzalez-bailon_networked_2016}.
+Participatory governance at such scales is unprecedented.
 The emergence of these large-scale decentralized collaborations has been
@@
 This dissertation focuses on one particularly challenging aspect of
-such collaborations: decision-making
+such collaborations: decision-making.
 I specificaly examine how,
@@
 communication network structure: the shape of who talks to whom.
+
+In this dissertation, I combine observational, agent-based modeling, and experimental methods to evaluate the role of network structure in large-scale participatory decision-making.
+Taken together, the findings in these studies suggests that small, tightly-knit pods, interlocked through the sharing of members can benefit such collaborations.
+This common structure, which I refer to as network deliberation, appears to preserve some benefits of small-scale deliberation, while allowing efficient diffusion of information through very large groups.
+Furthermore, I find evidence that network deliberation can mitigate some of the negative aspects of social influence.
+These findings provide insight into the success of existing large-scale collaborations as well as provide guidance for the design of new sociotechnical systems.
 
@@
 decision-making.
-This dissertation explores the potential of the internet to enable effective
-deliberation in large collectives.
+This dissertation explores the potential of effective large-scale deliberation, as enabled by the internet.
 Such large-scale deliberation creates the potential for members to
@@
 against,
-as well as a language and framework for contextualizing findings.
+as well as a language and framework for placing findings into the context of the greater social learning literature.
 
@@
 which I call {\em network deliberation}.
+Network deliberation is an analytical framework that can be applied both to the empirical observation of existing sociotechical systems,
+as well as to guide the creation and analysis of simplified models.
 My review of the literature identifies a common theme
@@
 affinity groups, zones, and nodes.
-As an abstraction of these small interlocking group, I will use the term "pods."
+As an abstraction of these small interlocking group, I will use the term ``pods.''
 Network deliberation describes large-scale collective deliberation achieved
@@
 varying structures.
-The central question of this dissertation is how the structure of these
-interlocking-pod networks influences the process and outcome of deliberation
-in large collaborations.
+The central questions of this dissertation are 1.
+does an interlocking-pod structure improve deliberation relative to conventional structures?
+and 2.
+how does the structure of these interlocking-pod networks influence the process and outcome of deliberation
+in large collaborations?
 
@@
 To address these challenges, this dissertation combines multiple methodologies,
-including: observational study, agent-based modelling, and field experiment.
+including: observational study, agent-based modelling, and field experiments.
 I use observational studies to reconstruct real-world collaborative networks
@@
 \item Despite an overall productivity/performance trade-off,
-WikiProjects with low-degree coeditor networks tend to have both higher
-productivity and higher performance;
-\item Short geodesic lengths are associated with higher performance,
-consistent with a conformity-based learning strategy;
-\item Structural inequality, as measured by degree skewness,
-is associated with lower performance;
-\item The agent-based model shows that the productivity and performance of
-collaborations can depend on network degree,
-and that the direction of that dependence can depend on social learning strategy.
+more tightly-knit WikiProjects tend to produce articles more quickly as well as produce higher-quality articles;
+\item The high performance of tightly-knit networks is consistent with individuals showing susceptibility to social influence;
+\item Unequal participation is associated with lower performance;
+\item The agent-based model shows that the relationship between quality and productivity can be influenced by both network structure and individual behavior.
 \end{itemize}
@@
 \setlength\itemsep{0pt}
-\item When agents conform to social influence,
+\item When agents are strongly susceptible to social influence,
 Network deliberation identifies solutions of higher quality than
@@
 while requiring less time to converge.
-\item Within network deliberation conditions,
-the structurally-efficient random pod network outperforms the
-structurally-inefficient long-path network when agents conform to social influence.
-However, when agents prefer their own judgement to social influence,
-the inefficient network yields higher performance,
-consistent with findings for conventional networks.
+\item Within network deliberation,
+agents who base behavior on social influence peroform best on networks with short paths.
+However, when agents act on their own judgement,
+they perform best on networks with longer paths,
+consistent with findings for conventional deliberation.
 \cite{barkoczi_social_2016}.
 \item A novel social learning strategy, confident-neighbor,
-which outperforms or matches the conventional best-neighbor strategy across all
-networks considered,
-despite using strictly less information about solution quality.
+which outperforms conventional strategies across all
+networks despite relying on strictly less information.
 \end{itemize}

### 2022-07-05 UMich/front_materials/acknowledgements.tex
Last edits before final draft.
Update acknowledgements and dedication.

@@
-Thanks to my advisor, Daniel M. Romero for his time, patience, and guidance.
+Thanks to my advisor, Daniel M. Romero for his time, expertise, guidance, and patience.
 Thanks to my committee:
@@
 Ceren Budak,
-for their time, expertise and feedback.
-Thank you to Hermino Bodon for collaboration on the agent-based modeling code in Chapter 3.
-Thanks to Jane Im for collaboration on interviewing the participants in the experiment described in Chapter 4.
-Thanks to Danielle Livneh, and Karthik Ramanathan for help collecting WikiProjects data.
-Thanks to professors Nicole Ellison, Yan Chen, Tanya Rosenblat, and Patrick Grim for feedback on projects and proposals that led to the work in this dissertation.
-Thanks to professors Grant Schoenbeck, Kentaro Toyama, and Charles R. Severance for feedback and encouragement on various projects.
+for their time, expertise, and feedback.
+
+Thanks to Danielle Livneh, and Karthik Ramanathan for help collecting the WikiProjects data used in Chapter \ref{chap:wp-prod-perf}.
+Thanks to Hermino Bodon for collaboration on the agent-based modeling code in Chapter \ref{chap:abm}.
+Thanks to Jane Im for collaboration on interviewing the participants in the experiment described in Chapter \ref{chap:experiment}.
+Thanks to Nicole Ellison, Yan Chen, Tanya Rosenblat, and Patrick Grim for feedback on projects and proposals that led to the work in this dissertation.
+Thanks to Grant Schoenbeck, Kentaro Toyama, and Charles R. Severance for feedback and insightful discussion.
+
+Chapter \ref{chap:wp-prod-perf} has been adapted from previously published work \cite{platt_network_2018}.
+
+This research was partly supported by
+the National Science Foundation under Grant No. IIS-1617820,
+Air Force Office of Scientific Research under Award No. FA9550-19-1-0029,
+and the UMSI REMS program.
+
 Thanks to Veronica Falandino for making UMSI a kinder place.
 Thanks to Linda Wood and Mita Gibson of the Center for the Study of Complex Systems always being willing to help. 
-Thanks to the members and staff of the Graduate Employees Organization, AFT Michigan 3550, especially Denise Bailey and Liz Ratzloff, for advocating for the wellbeing of Michigan graduate students.
-Thanks to Avdi Grimm for consulting on some very difficult Ruby bugs.
+Thanks to the members and staff of the Graduate Employees Organization, AFT Michigan 3550, especially Denise Bailey and Liz Ratzloff, for advocating for the well-being of Michigan graduate students.
+Thanks to Avdi Grimm for consulting on some very challenging Ruby bugs.
 Thanks to Salem Elzway for technical writing feedback.
-Thanks to the many Michigan graduate students who have given me feedback and guidance on work and classes:
+
+Thanks to many Michigan graduate students for feedback and insightful conversations, including:
 Jeffrey W. Lockhart,
@@
 for helpful feedback.
-Thank you to
+Thanks to
+Lorrie LeJeune,
 Rahul Bhargava,
@@
 Charlie DeTar,
+Erhardt Graeff,
+J. Nathan Matias,
 Gerald J. Sussman,
 Anthony Garnock-Jones,
+Brandon Willard,
+Laurel Ruhlen,
 and Elaine Lee
 for guidance and encouragement in returning to academia.
-Thanks to the members of i3 Detroit past and present for inspiring my interest in participatory governance.
+Thanks to the past and present members of i3 Detroit for inspiring my interest in participatory governance.
+
+Thanks to Caitlin Cassady and Justin Veenhuis for taking in strays.
 Thanks to Persephone Hernandez-Vogt for their companionship and encouragement.
 Thanks to Andy Brosius for sharing their life and family with me, and for their many years of patience and encouragement.
-This research was partly supported by
-the National Science Foundation under Grant No. IIS-1617820,
-Air Force Office of Scientific Research under Award No. FA9550-19-1-0029,
-and the UMSI REMS program,

### 2022-07-08 UMich/chapters/ch03-NetDelibABM.tex
Import NetDelibExp and formatting fixes.
Update ch3 based on feedback and additional results.

@@
 We shall use the term {\em network deliberation} to refer to mass deliberation using an interlocking pod structure.
-Communication network structure also influences the success of collaborations by controlling the speed of information diffusion \cite{lazer_network_2007,mason_propagation_2008,mason_collaborative_2012,barkoczi_social_2016, gomez_clustering_2019,derex_partial_2016,kearns_experiments_2012}.
-Structurally efficient networks enable fast diffusion and favor exploitation of known information over exploration of the unknown \cite{lazer_network_2007, barkoczi_social_2016}.
+Communication network structure also influences the success of collaborations by limiting the maximum speed of information diffusion \cite{lazer_network_2007,mason_propagation_2008,mason_collaborative_2012,barkoczi_social_2016, gomez_clustering_2019,derex_partial_2016,kearns_experiments_2012}.
+Structurally efficient networks,
+those with typically short paths between individuals,
+enable fast diffusion and favor exploitation of known information over exploration of the unknown \cite{lazer_network_2007, barkoczi_social_2016}.
 In the context of network deliberation, structural efficiency is determined by how individuals are assigned to pods.
@@
 
-We model network deliberation using a social learning framework to address two questions.
+We apply a formal social learning framework to model network deliberation.
+Specifically, we address two questions.
 How effective is network deliberation compared to conventional single-group deliberation?
@@
 To model network deliberation, we construct interlocking networks in stages. At each stage, agents are partitioned into small {\em pods} and connected to all other agents in the same pod.
-Our choice of pod size is informed by real-world The upper limit of effective small group size has been estimated variously as five \cite{freeman_tyranny_1972}, five--eight\cite{lohman_designing_2000}, and eight \cite{miflin_small_2004}.
+Our choice of pod size is informed by studies of real-world group dynamics.
+The upper limit of effective small group size has been estimated variously as five \cite{freeman_tyranny_1972}, five--eight\cite{lohman_designing_2000}, and eight \cite{miflin_small_2004}.
 We choose a pod size of 5.
@@
 We allow each simulation to proceed for 300 iterations, which we have found sufficient to guarantee convergence with our chosen parameters.
+For real-world comparison, on average, an active English-language Wikipedia editor makes about 45 edits per month \cite{noauthor_wikipediastatistics_2022}.
 
@@
 In this paper, we consider three separate strategies: {\em best-neighbor}, {\em confident-neighbor}, and {\em conform}.
-Each is combined with the same mutation-based individual learning strategy.
+Each is combined with the same individual learning strategy.
 
-To understand why real-world deliberators might choose different strategies,
-it is necessary to realize that the social learning strategies available to an
+The social learning strategies available to an
 individual depend on that individual's ability to evaluate the quality of candidate solutions.
@@
 more heavily on the recommendations and influence of their peers.
-We formalize an agent's ability to evaluate the quality of a candidate solution
-as its {\em capabilities}, which we now explore in further detail.
+We now develop a formalism describing an agent's ability to evaluate the quality of a candidate solution.
+This formalism provides precise language for distinguishing between behavioral and informational constraints on agent behavior.
 
@@
 When comparing learning strategies, it will be helpful to classify them according to the capabilities they require of agents.
-We formalize these capabilities as oracle functions, which accept one or more task solutions as input and reveal whether the supplied state(s) satisfy some particular property.
+We formalize these {\em capabilities} as oracle functions, which accept one or more task solutions as input and reveal whether the states satisfy some particular property.
 
@@
 
-Confident-neighbor differs from best-neighbor because of two subtle but important considerations.
+Confident-neighbor can produce a different candidate than best-neighbor as a result of two subtle but important considerations.
 First, an agent's neighbors need not be adjacent to each other.
@@
 The exception is when agents belong to a clique (e.g., in network deliberation), in which case agents receive exactly one broadcast for each maximal solution in the clique, thus yielding the same results as best-neighbor.
-Confident-neighbor is thus more appropriate than best-neighbor for comparing network deliberation to single group deliberation, as the same information is utilized in both cases.
+This equivalence is notable because confident-neighbor requires only single-state
+comparisons rather than the arbitrary state comparisons used by best-neighbor.
+Best-neighbor thus requires stronger capabilities to achieve the same results under conventional deliberation,
+while confident-neighbor requires the same set of capabilities under both conventional and network deliberation.
 
@@
 iterations for each network/strategy combination.
-The results of these simulations are shown in Figure \ref{fig:results}.
+The results of these simulations are shown in Figure \ref{fig:ch03results}.
 Figures \ref{fig:results-frac-parallel} and \ref{fig:results-frac-fallback} show pairwise comparisons of solution quality between strategy/network settings.
@@
 \begin{figure*}[]
-    \label{fig:results}
     \centering
@@
 	}
+    \label{fig:ch03results}
 \end{figure*}
@@
 \begin{figure}
-    \label{fig:results-frac-parallel}
     \centering
@@
 	\caption{Fraction of simulations in which the row condition outperforms the column condition. Results are shown for parallel individual learning. In the case of a tie, weight is divided evenly.}
+    \label{fig:results-frac-parallel}
 \end{figure}
@@
 \begin{figure}
+    \centering
+    \includegraphics[width=3.33in]{chapters/figures/NetDelibABM/fig-result-frac-parallel.eps}
+	\caption{Fraction of simulations in which the row condition outperforms the column condition. Results are shown for fallback individual learning. In the case of a tie, weight is divided evenly.}
     \label{fig:results-frac-fallback}
-    \centering
-%    \includegraphics[width=3.33in]{chapters/figures/NetDelibABM/fig-result-frac-parallel.eps}
-\caption{Fraction of simulations in which the row condition outperforms the column condition. Results are shown for fallback individual learning. In the case of a tie, weight is divided evenly.}
 \end{figure}
@@
 \begin{figure}
-    \label{fig:results-bn-parallel-dist}
     \centering
@@
 \end{figure}
+    \label{fig:results-bn-parallel-dist}
+
 \begin{figure}
-    \label{fig:results-cn-parallel-dist}
     \centering
@@
 	\caption{Solution quality distribution for confident-neighbor strategy with parallel individual learning.}
+    \label{fig:results-cn-parallel-dist}
 \end{figure}
+
 \begin{figure}
-    \label{fig:results-conf-parallel-dist}
     \centering
@@
 	\caption{Solution quality distribution for the conform strategy with parallel individual learning.}
+    \label{fig:results-conf-parallel-dist}
 \end{figure}
@@
 \begin{figure}
-    \label{fig:results-bn-fallback-dist}
     \centering
@@
 	\caption{Solution quality distribution for best-neighbor strategy with fallback individual learning.}
+    \label{fig:results-bn-fallback-dist}
 \end{figure}
+
 \begin{figure}
-    \label{fig:results-cn-fallback-dist}
     \centering
@@
 	\caption{Solution quality distribution for confident-neighbor strategy with fallback individual learning.}
+    \label{fig:results-cn-fallback-dist}
 \end{figure}
+
 \begin{figure}
-    \label{fig:results-conf-fallback-dist}
     \centering
@@
 	\caption{Solution quality distribution for the conform strategy with fallback individual learning.}
+    \label{fig:results-conf-fallback-dist}
 \end{figure}
@@
 \begin{table*}[]
-    \label{tab:t-instrat-fallback}
     \centering
@@
 For all rows, dof=999 and $p^*$ is the Bonferroni-corrected p-value with m=60.}
+    \label{tab:t-instrat-fallback}
 \end{table*}
@@
 \begin{table*}[]
-    \label{tab:t-innet-fallback}
     \centering
@@
 For all rows, dof=999 and $p^*$ is the Bonferroni-corrected p-value with m=60.}
+    \label{tab:t-innet-fallback}
 \end{table*}
@@
 \begin{table*}[]
-    \label{tab:t-instrat-parallel}
     \centering
@@
 For all rows, dof=999 and $p^*$ is the Bonferroni-corrected p-value with m=60.}
+    \label{tab:t-instrat-parallel}
 \end{table*}
@@
 \begin{table*}[]
-    \label{tab:t-innet-parallel}
     \centering
@@
 For all rows, dof=999 and $p^*$ is the Bonferroni-corrected p-value with m=60.}
+    \label{tab:t-innet-parallel}
 \end{table*}
@@
 We are particularly interested in comparing the performance of network deliberation to that of conventional single-group deliberation.
-When the conform strategy is used, we find that both of the network deliberation conditions yield better performance than conventional networks (Figure \ref{fig:results-conform-diff}).
-Analysis of the solution quality distributions (Figure \ref{fig:results-conform-dist}) shows two distinct effects.
-First, network deliberation frequently finds maximal solutions while conventional deliberation does not.
-Second, even when non-maximal solutions are found, network deliberation finds higher-quality solutions.
-This benefit is robust across both parallel and fallback individual learning.
-The effect size is dramatic.
-Conventional networks combined with the conform strategy show the lowest performance of all conditions, while random pod network deliberation combined with the conform strategy shows higher performance than nearly all other conditions.
-The exception being small world network combined with the best/confident-neighbor strategy, which shows equal performance within our statistical resolution.
-Often, increased performance comes at the expense of speed as more time is devoted to exploring novel solutions.
-However, we find that network deliberation improves both performance and speed when the conform strategy is used, suggesting that interlocking network structure can enable more efficient exploration.
+
+When the conform strategy is used, we find that both network deliberation conditions yield better performance than conventional networks (Figure \ref{fig:results-conform-diff}).
+The effect size is substantial.
+The solution quality distributions (Figure \ref{fig:results-conform-dist}) show that network deliberation finds maximal solutions more frequently than conventional.
+Furthermore, non-maximal solutions found by network deliberation are typically higher-quality than those found by conventional.
+The quality difference is due to a reduction in solution quality in conventional networks that is not seen for network deliberation.
+Agents' transitions between basins of attraction provide additional insight into the mechanism.
+In conventional deliberation, transitions to higher and lower quality basins are nearly equally common (Figure \ref{fig:conform-transition-basin}).
+For network deliberation, there is a small bias towards higher quality basins, particularly as time progresses.
+Similarly, network deliberation typically converges to a single basin, while multiple basins remain occupied for conventional deliberation (Figure \ref{fig:conform-basin}).
+So we see that in the conventional setting, agents converge quickly to sub-optimal basins, and become stuck.
+While in network deliberation, agents converge quickly but continue to explore and transition to preferable basins.
+Why does this behavior occur for the conform strategy? And how does network deliberation provide protection?
+Under the conform strategy, popular solutions can spread quickly within highly-clustered communities, even if those solutions are low-quality,
+resulting in an information cascade \cite{banerjee_simple_1992}.
+Furthermore, once a community is saturated with a single solution, agents mutually reinforce each other, preventing any from exploring alternative solutions.
+However, network deliberation provides cross-cutting connections to multiple pods which can outweigh an information cascade originating in a single pod.
+Network deliberation thus protects against detrimental effects of social influence, such as information cascades.
+
+In many problem-solving settings, improvements to quality can come at the cost of speed, as more time is devoted to exploring novel solutions.
+Such a reduction in speed can result in increased latency and reduced productivity.
+However, we find that network deliberation improves both performance and speed when the conform strategy is used, suggesting that interlocking network structure enables more efficient exploration.
+
+In contrast to the conform strategy, network deliberation performs comparably or slightly worse than conventional under the best-neighbor and confident-neighbor strategies.
+This finding suggests that network deliberation is most beneficial when individuals are relying on social influence rather thank solution quality, but can potentially be applied in either context.
 
 \begin{figure}
-    \label{fig:results-conform-diff}
     \centering
-    \includegraphics[width=3.33in]{chapters/figures/NetDelibABM/fig-results-netdelib-diff.eps}
+    \includegraphics[width=3in]{chapters/figures/NetDelibABM/fig-results-netdelib-diff.eps}
 	\caption{Difference in solution quality between network deliberation and conventional deliberation.
@@
 Results are shown for parallel individual learning.}
+    \label{fig:results-conform-diff}
 \end{figure}
@@
 \begin{figure}
+    \centering
+    \includegraphics[width=3in]{chapters/figures/NetDelibABM/fig-results-conform-dist.eps}
+	\caption{Distributions of solution quality for social learning on a preferential-attachment network and random-pod network deliberation. Results are shown for parallel individual learning.}
     \label{fig:results-conform-dist}
+\end{figure}
+
+\begin{figure}
 \centering
-    \includegraphics[width=3.33in]{chapters/figures/NetDelibABM/fig-results-conform-dist.eps}
-\caption{Distributions of solution quality for social learning on a preferential-attachment network and random-pod network deliberation. Results are shown for parallel individual learning.}
+\includegraphics[width=6in]{chapters/figures/NetDelibABM/fig-conform-transition-basin.eps}
+	\caption{Number of agent transitions to basins of attraction with higher/lower maxima at each time step.}
+    \label{fig:conform-transition-basin}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=3in]{chapters/figures/NetDelibABM/fig-conform-basin.eps}
+	\caption{Number of basins of attraction occupied by at least one agent at each time step.}
+    \label{fig:conform-basin}
 \end{figure}
@@
 As a form of rational ignorance, conformist behavior can conserve individual resources, but poses the risk of forming information cascades and propagating misinformation and inferior or harmful innovations \cite{banerjee_simple_1992}.
-Our simulations show that network deliberation can greatly improve the outcome of deliberation in the presence of strong social influence.
+Our results show that network deliberation can greatly improve the outcome of deliberation in the presence of strong social influence.
 This insight suggests one possible mechanism behind the success of mass deliberative projects, as well as the potential for network-based interventions in mass deliberation.
@@
 The results are opposite for the conform strategy, with the efficient network yielding better results.
-These results are consistent with those observed for conventional single-group networks \cite{barkoczi_social_2016}, but to the best of our knowledge, they had not been explored for interlocking network like the ones we use to model network deliberation.
+These results are consistent with those observed for conventional single-group networks \cite{barkoczi_social_2016}, but to the best of our knowledge, they have not been explored for interlocking network like the ones we use to model network deliberation.
 
 \begin{figure}
-    \label{fig:results-netdelib-efficiency}
     \centering
@@
 	\caption{Difference in solution quality between random-pod (efficient) and long-path (inefficient) network. Results are shown for parallel individual learning.}
+    \label{fig:results-netdelib-efficiency}
 \end{figure}
@@
 \begin{figure}
-    \label{fig:results-confident}
     \centering
@@
 	Results are shown for parallel individual learning.}
+    \label{fig:results-confident}
 \end{figure}
@@
 Empirical evidence suggests the presence of strong social influence in real-world collaborations such as Wikipedia \cite{platt_network_2018} and in lab studies of human collaboration \cite{mason_collaborative_2012, barkoczi_social_2016}.
-In the case of weak social influence (i.e., the  best-neighbor and confident-neighbor strategies), we do not observe a significant difference in output quality between network deliberation and conventional single-group deliberation.
+In the case of weak social influence (i.e., the  best-neighbor and confident-neighbor strategies), we do not observe a substantial difference in output quality between network deliberation and conventional single-group deliberation.
 In all cases, network deliberation performs better than single-group deliberation or comparatively well.
@@
 How does network deliberation improve collaborative output?
-In part, the high performance of random-pod network deliberation can be attributed to the high structural efficiency of the resulting network.
-However, structural efficiency cannot be the complete mechanism.
-Long-path network deliberation is extremely inefficient, but still outperforms conventional networks in the presence of strong social influence.
-The additional mechanisms at play remain an open question.
+We have found that network deliberation protects against the negative effects of strong social influence by disrupting information cascades.
+Rather than allowing individuals to converge to polarized communities with suboptimal solutions, the interlocking-pod structure of network deliberation allows agents to continue to explore and improve their solutions.
 
@@
 
-Our work is limited by a number of assumptions, which require further investigation to achieve greater generalizability.
+Our work is limited by several assumptions, which require further investigation to achieve greater generalizability.
 We assume a static objective function, while real-world objectives can change with changing external conditions as well as changing individual preferences.
@@
 
-Effective mass deliberation is key to realizing the democratizing power of the internet.
+Effective mass deliberation is key to realizing the potential of large-scale collective action and the democratizing potential of the internet.
 Our findings suggest that the presence of small and interlocking pods can contribute to the success of mass deliberation, particularly when the topic is complex and social influence is strong.

### 2022-08-01 UMich/chapters/ch02-WPProdPerf.tex
Import NetDelibExp and formatting fixes.
Update figures in ch02.
Revisions for ch02.

@@
 suggesting the existence of factors that correlate positively with both.
+We are particularly interested in factors that give insight into the success of Wikipedia
+as a large-scale collaboration, such as those relating to the network structure of
+interactions between editors.
 
@@
 
-We now define an productivity measure which meets the above criteria.
+We now define a productivity measure which meets the above criteria.
 Let $T(W,G)$ be the set of article assessment transitions from below grade $G$
@@
 don't have many more than a typical editor.
+For traditional organizations, maximum effective team size has been
+variously estimated between 5 and 8 individuals
+\cite{freeman_tyranny_1972, lohman_designing_2000, miflin_small_2004},
+while WikiProject mean editor degrees range from 0 to 1256, with a median of 133.
 
@@
 network using min-cut size.
-The minimum $st$-cut between nodes $s$ and $t$ is the set of edges that must beremoved
+The minimum $st$-cut between nodes $s$ and $t$ is the smallest set of edges that must be removed
 for no path exists from $s$ to $t$.
 The minimum cut (min-cut) of a graph is the smallest minimum $st$-cut over all node pairs $st$. 
-The size of the graph min-cut quantifies the connectivity of a graph,
+The size (number of edges) of the graph min-cut quantifies the connectivity of a graph,
 but only incorporates information about edges lying on paths crossing the min-cut.
@@
 \centering
-\includegraphics[width=2.5in,height=2.5in]{chapters/figures/WPProdPerf/fig-eff-perf-hist.pdf}
+\includegraphics[width=6in,]{chapters/figures/WPProdPerf/fig-eff-perf-hist.pdf}
 \caption{
@@
 We also include the following project-level variables to control for confounding factors.
+For project similarity, we use the Jaccard similarity \cite{jaccard1912distribution},
+which provides a meaningful index for comparing set membership.
 \begin{description}
@@
 \item[Talk fraction.] Fraction of total revisions made to talk pages.
-\item[Mean similarity.] Mean Jaccard similarity (by article) with other WikiProjects; a measure of topical complexity.
+\item[Mean similarity.] Mean Jaccard (by article) with other WikiProjects; a measure of topical complexity.
 \item[Mean editors/article.] Mean number of editors collaborating on each article in a WikiProject.
@@
 Min-cut is excluded from all models to avoid collinearity,
-as it is highly correlated with degree.
+as it is highly correlated (r=0.44) with degree.
+In many cases the min-cut is simply the removal of the highest-degree node.
 In-degree and out-degree skewness were also highly correlated,
@@
 The different behavior of A-productivity is likely explained by the
-observation that the A-Class quality is infrequently used in practice.
+observation that the A-Class quality is infrequently used in practice,
+with an average of 16.8 A-Class articles per project versus
+118.2 and 251.1 for B-Class and C-Class.
 The A-Class quality level is usually passed when an article reaches
 good or featured article status,
-which follow deifferent a consensus process from other ratings.
+which follow a different consensus process from other ratings.
 
@@
 \centering
-\includegraphics[width=3in,height=1.5in]{chapters/figures/WPProdPerf/fig-val-iter.pdf}
+\includegraphics[width=6in,height=3in]{chapters/figures/WPProdPerf/fig-val-iter.pdf}
 \caption{
@@
 \begin{tabular}{lll}
-Strategy & Performance & productivity\\
+Strategy & Performance & Productivity\\
 \hline
@@
 Strategies are defined in Table \ref{tab:strat}.
-Local strategies are less productive than their non-local counterparts.
 Local best-neighbor out-performs global,
@@
 The local majority strategy is both most productive and most performant.
+Local strategies are more productive for the conform strategy and less productive for the best-neighbor strategy.
 \label{tab:sim-eff-perf}
@@
 \begin{tabular}{lll}
-Strategy & Perf. Std. Coeff. & Eff. Std. Coeff.\\
+Strategy & Perf. Std. Coeff. & Prod. Std. Coeff.\\
 \hline
@@
 \caption{
-productivity and Performance of social learning strategies vs. mean network degree.
+Performance and productivity of social learning strategies vs. mean network degree.
 Each point represents a single trial of 300 iterations.
@@
 
-Our agent-based models offer a several insights.
+Our agent-based models offer several insights.
 We observe degree-dependent performance and productivity
@@
 \section{Conclusion}
-\label{sec:conclusion}
+\label{sec:ch02conclusion}
 

### 2022-08-06 UMich/chapters/ch04-NetDelibExp.tex
Last edits before final draft.
Chapter 4 revisions based on defense feedback.

@@
 We focus on deliberation due to fundamental limitations of voting.
-Results from social choice theory showt that any ranked-choice voting system is subject to limitations such as the Condorcet Paradox \cite{condorcet_essay_1785} and Arrow's Impossibility Theorem \cite{arrow_social_2012}.
+Results from social choice theory show that any ranked-choice voting system is subject to limitations such as the Condorcet Paradox \cite{condorcet_essay_1785} and Arrow's Impossibility Theorem \cite{arrow_social_2012}.
 Deliberation offers one way to avoid these limitations by changing individual preferences \cite{ackerman_deliberation_2002}.
@@
 As discussed in Chapter \ref{chap:abm}, the successful large-scale collaborations enabled by the internet often exhibit network sructures characterized by small, interlocking groups, reminiscent of interlocking directorates \cite{levine_study_1979} and interlocking publics \cite{habermas_structural_1991}.
-In Chapter \ref{chap:abm}, we describe agent-based simulations showing that such network deliberation can improve performance on complex tasks when individuals exhibit strong social influence.dsd
+In Chapter \ref{chap:abm}, we describe agent-based simulations showing that such network deliberation can improve performance on complex tasks when individuals exhibit strong social influence.
 Here, we expand on existing research by performing a controlled online experimental evaluation of network deliberation.
@@
 The closest we are aware of is the study of network rotation by Salehi \& Bernstein \cite{salehi_hive_2018}.
-The present experiment goes beyond existing work primarily by tracking preferences throughout the experiment and by the use of a conventional single-group control condition.
+The present experiment goes beyond existing work in several ways.
+Our primarity innovation is the tracking of preferences throughout the experiment,
+which allows the quantification of consensus and social influence.
+We also include a large single-group control condition.
+Through the use of a pseudonymous online platform, we isolate deliberative text and network structure from other social cues that
+might result in social influence, e.g., tone of voice, visual appearance.
+While these factors undoubtedly play an important role in the social dynamics of deliberation,
+we are focused specifically on the role of communication network structure.
 We also describe multiple network deliberation topologies for achieving efficient or inefficient networks.
 For the present study, we focus on an efficient topology.
+By tracking preference evolution, limiting confounding factors, and using a controlled experimental design,
+we can examine the dynamics and causality of the relationship between network structure and preference evolution.
+
+Observational and numerical findings have shown that successful large-scale collaborations often exhibit a structure
+of small, interlocking pods: {\em network deliberation}.
+This pod structure can be seen in a range of collaborations, including the
+articles of Wikipedia \cite{platt_network_2018},
+the projects and modules of the free/open-source software community \cite{benkler_coases_2002},
+and the nodes of Occupy Sandy \cite{brugh_combining_2019}.
+Numerical simulations (Chapter \ref{chap:abm}) demonstrate the effectiveness of the network
+deliberation structure in the presence of social influence.
 
@@
 \subsection{Social Choice Theory}
-Our analysis of preferences is foudned in the formalism of social choice theory \cite{arrow_social_2012, gaertner_primer_2009}.
+Our analysis of preferences is founded in the formalism of social choice theory \cite{arrow_social_2012, gaertner_primer_2009}.
 In social choice theory, individual's {\em preferences} are represented by a ranked ordering of {\em alternatives} (e.g., proposals, candidates).
@@
 \begin{equation}
+\label{eqn:distance-max}
     v_m^{\max} =
@@
 \begin{equation}
+\label{eqn:weight}
     w_i = 1 + \sum_{k=1}^{i-1} w_k v_k^{\max}.
@@
 
+For example, consider the following two preference orderings (in least-to-most preferred order):
+\begin{eqnarray}
+p &=& (C, B, E, F, A, D) \\
+q &=& (F, E, B, C, A, D).
+\end{eqnarray}
+Applying Eq. \ref{eqn:distance-max} and \ref{eqn:weight} 
+gives the maximum place-values $v^{\text{max}}$ and place weights $w_i$:
+\begin{eqnarray}
+v^{\text{max}} &=& (1, 2, 3, 2, 1) \\
+w &=& (1, 2, 6, 24, 72).
+\end{eqnarray}
+Partitioning both vectors at index 3 gives:
+\begin{eqnarray}
+p &\mapsto& {C, B, E}, {F, A, D} \\
+q &\mapsto& {F, E, B}, {C, A, D}.
+\end{eqnarray}
+Noting that these partitions differ by a single swap ($F \rightleftarrows C$)
+the corresponding element of the crossing vector $v_3$ is $1$.
+Applying the same procedure to each index gives the full crossing vector:
+\begin{eqnarray}
+v(p,q) &=& (1, 2, 1, 0, 0).
+\end{eqnarray}
+Taking the inner product with the weight vector gives the total crossing distance:
+\begin{eqnarray}
+d(p, q) &=& v(p,q) \cdot w \\
+&=& 11.
+\end{eqnarray}
+Normalizing by the maximum crossing distance $143$ gives a normalized distance of approximately $0.08$.
+For comparison, the normalized Kendall tau distance bewtween these two preferences is $0.40$,
+much larger despite agreeing on the highest two choices.
+
 \subsection{Quantifying Social Influence}
 
-We are particularly interested in the role of social influence in preference evolution and require a formalism to quantify and analzye social influence.
+Deliberation is fundamentally a social influence process in which
+individuals exchange statements with their peers in an effort to influence their preferences.
+We will thus benefit from a formalism to quantify and analzye social influence.
 If two individuals preferences become closer, can we say whether one influenced the other? And if so, can we say which influence which?
@@
 \begin{itemize}
-    \item The magnitude of an individual's change in preference increases with their susceptibility with social influence;
+    \item The magnitude of an individual's change in preference increases with their susceptibility to social influence;
     \item The magnitude of an individual's change in preference increases with the magnitude of the distance between their preference and the influencing preference.
 \end{itemize}
+The distance between two individuals' preferences is symmetric between the two individuals,
+so the latter of the above conditions does not speak to who influences whom.
+The converse of the former condition however suggests that the greater the change in an individual's preference,
+the more they have been influenced by other.
 Thus, if the distance between the preferences of individuals $A$ and $B$ decreases, and the preference of $B$ changed more than the preference of $A$, we assume that individual $A$ influenced individual $B$.
+On the other hand, if the distance increases, we refer to this as negative social influence, suggesting one individual has influenced another to adopt an opposing preference.
 We now describe two formal measures that use these assumptions to quantify how susceptible to influence and how potential influential an individual is.
@@
 
+Measuring only the change in agreement leaves some ambiguities.
+For instance, changes in agreement could simply be due to individuals converging towards initially popular preferences.
+Alternatively initially unpopular preferences might overtake others in popularity,
+which suggests unequal social influence, and might indicate the presence of an information cascade.
 We wish to quantify whether a change in agreement/disagreement is due to individuals converging towards initially popular alternatives, or to the the group being persuaded towards initially unpopular alternatives.
@@
 
+We divide the change in agreement between individuals $p$ and $q$ into two parts:
+that due to $p$'s influence on $q$ and vice versa. 
 To apportion the change in agreement between individuals $p$ and $q$,
-we make the assumption that an individual's contribution to the change in agreement is proportional to the change in their own preference.
+we make the assumption that the contribution due to $q$'s influence on $p$'s preferences is
+proportional to the change in $p$'s preference.
 Specifically, we define the conformity $c(p,q)$ of individual $p$ toward individual $q$ as:
@@
 \end{eqnarray}
+These values are illustrated schematically in Figure \ref{fig:conform-hip-diagram}.
 Note that the change in agreement between two individuals can be written in terms of the conformity:
@@
 \end{equation}
+Note that a high hipness does not imply that an individual has influenced others to adopt similar preferences,
+only that others have adopted such preferences.
 Both $C(p)$ and $H(p)$ are equal to $\Delta$ when summed over all individuals, but they describe conceptually different attributes of an individual.
@@
 The hipness quantifies how much an individual has increased agreement due to holding an unchanging preference near to others' final preferences.
+
+\begin{figure}
+\centering
+\includegraphics[width=3in]{chapters/figures/NetDelibExp/fig-conform-hip-diagram.pdf}
+\caption{Schematic illustration of the quantities used in calculating the social influence
+between $p$ and $q$. Initial preferences are shown as $p_i$ and $q_i$ with final preferences
+shown as $p_f$ and $q_f$. The change in agreement from initial to final is primarily due to
+$p$'s movement, so we consider $p$ to be high in conformity, while $q$ is high in hipness.}
+\label{fig:conform-hip-diagram}
+\end{figure}
 
@@
 \begin{figure}
-\label{fig:kcon}
 \center
@@
 networks for $T=3$ stages and a pod size of $M=5$.}
+\label{fig:kcon}
 \end{figure}
@@
 \center
-\label{tab:prompts}
 \begin{tabular}{|p{0.5in}|p{4.5in}|}
@@
 \caption{Prompts shown to participants during each round of deliberation.}
+\label{tab:prompts}
 \end{table}
@@
 Assignments were uniformly at random.
-Stratified sampling would be preferable in order to ensure even demographic distributions, but was not feasible due to software and time constraints.
-However, we manually confirmed that each condition had comparable demographics.
+We manually confirmed that each condition had comparable demographics.
 The two conditions were:
@@
 \end{description}
+
+\subsubsection{Ethical Considerations}
+This study was reviewed by the University of Michigan's IRB and classified as exempt.
+To protect participant privacy, all participants were assigned randomly-generated pseudonyms.
+The keys connecting pseudonyms to identifiable information were stored securely and only available
+to the researchers.
+Participants were provided with an email address to report any harassment, which was continuously
+monitored throughout the experiment.
+The deliberations were also regularly reviewed for any disruptive, hateful, or harassing activity.
 
@@
 \begin{description}
-\item[Prop. 1]{Open-ended with partial credit (2 questions, 5 points each) - Ch. 1-2: Intro to networks.}
-\item[Prop. 2]{Open-ended with partial credit (2 questions, 5 points each) - Ch. 3: Bridges, clustering, triadic closure, strong and weak ties, and the Strong Triadic Closure Property.}
-\item[Prop. 3]{Multiple choice with no partial credit (5 questions, 2 points each) - Ch. 4: Homophily, selection, social influence, and social-affiliation networks.}
-\item[Prop. 4]{Multiple choice with no partial credit (5 questions, 2 points each) - Ch. 5: Signed networks and balance theory.}
-\item[Prop. 5]{True/false with no partial credit (10 questions, 1 point each) - Ch. 6: Game theory, best responses, and Dominant Strategies, Nash equilibrium.}
-\item[Prop. 6]{True/false with no partial credit (10 questions, 1 point each) - Ch. 9:  Auctions.}
+\item[Prop. 1]{Open-ended, partial credit (2 questions, 5 points each) - Ch. 1-2}
+\item[Prop. 2]{Open-ended, partial credit (2 questions, 5 points each) - Ch. 3}
+\item[Prop. 3]{Multiple choice, no partial credit (5 questions, 2 points each) - Ch. 4}
+\item[Prop. 4]{Multiple choice, no partial credit (5 questions, 2 points each) - Ch. 5}
+\item[Prop. 5]{True/false, no partial credit (10 questions, 1 point each) - Ch. 6}
+\item[Prop. 6]{True/false, no partial credit (10 questions, 1 point each) - Ch. 9}
 \end{description}
@@
 \begin{figure}
-\label{fig:reply}
 \center
@@
 \caption{Participants can perform standard forum actions, such as posting and commenting.}
+\label{fig:reply}
 \end{figure}
@@
 \begin{figure}
-\label{fig:vote}
 \center
@@
 \caption{Participants are presented with a ranked-choice voting screen between rounds of deliberation.}
+\label{fig:vote}
 \end{figure}
@@
 \begin{figure}
-\label{fig:alias}
 \center
@@
 \caption{Participants are assigned an alias for the duration of the experiment.}
+\label{fig:alias}
 \end{figure}
@@
     \centering
-    \includegraphics[width=3in]{chapters/figures/NetDelibExp/fig-counts.pdf}
+    \raisebox{-\height}{\includegraphics[width=3in]{chapters/figures/NetDelibExp/fig-counts.pdf}}
+    \raisebox{-\height}{\includegraphics[width=1in]{chapters/figures/NetDelibExp/fig-counts-legend.pdf}}
     \caption{Number of first-choice votes for each alternative over the course of the deliberation.}
@@
 While the random-pod group does show some participants switching away from Prop. 2, those participants shift to the less popular Props. 3 and 4 rather than Prop 1.
-In the sections that follow, we will further analyze the evolution of participant preferneces and the potential mechanisms driving them.
+In the sections that follow, we will further analyze the evolution of participant preferences and the potential mechanisms driving them.
 
@@
 Here we describe the trends we observe in participants' deliberation activity and the potential impact on participant preferences.
-Overall, the nubmer of comments per participant shows a heavy-tailed distribution (Figure \ref{fig:commentcounts}) with a mode of one comment per participant.
+Overall, the number of comments per participant shows a heavy-tailed distribution (Figure \ref{fig:commentcounts}) with a mode of one comment per participant.
 The random-pod group shows less of a heavy-tail, with more participants posting a single comment per round and a smaller maximum count (2--3 comments vs. 3--4 for control).
@@
 
-Figure \ref{fig:commentrounds} shows the comment counts for each round as well as the Shannon entropy of participant comment counts.
+Figure \ref{fig:commentrounds} shows the comment counts for each round as well as the Shannon entropy of the distribution of comment counts over participants.
 The random-pod group shows a lower total number of comments and a higher entropy, again showing a more equal division of activity across participants.
@@
 In the control group, this rebound resulted in a net reduction in agreement, suggesting the presence of negative social influence.
+In other words, participants altered their preferences to more strongly oppose those they interacted with.
 
@@
 In both the control and random-pod groups, we see round 1 deliberation clearly favoring one alternative, which then gains a number of first-choice votes, suggesting social influence.
-This social influence is further evidenced by the increase in agreement observed for both groups over the course round 1 (Figure \ref{fig:agreement}).
+This social influence is further evidenced by the increase in agreement observed for both groups over the course of round 1 (Figure \ref{fig:agreement}).
 
@@
 The negative shift in agreement in the control group is due
-to a majority of participants shifting their preferences away (negative conformity) from a samll number of stubborn participants (highly-negative hipness) without converging on an alternative (lack of positive conformity).
+to a majority of participants shifting their preferences away (negative conformity) from a small number of stubborn participants (highly-negative hipness) without converging on an alternative (lack of positive conformity).
 Furthermore, this shift occurs after round 2, following an initial stage of high-conformity.
 The random-pod participants also exhibit stubborn participants, but fewer and with smaller magnitude,
-and the the majority of participants converge toward each other's preferences (positive conformity).
+and the majority of participants converge toward each other's preferences (positive conformity).
 
@@
 However, our hypothesis was based on the assumption that network deliberation would more effectively identify and resolve disagreements,
-leading us to expect a large increase increase in agreement in the random-pod group.
+leading us to expect a large increase in agreement in the random-pod group.
 Instead, we observe a small increase of agreement within the random-pod group and a large decrease within the control group.

### 2022-08-08 UMich/chapters/ch04-NetDelibExp.tex
More ch04 revisions.

@@
 On the other hand, if the distance increases, we refer to this as negative social influence, suggesting one individual has influenced another to adopt an opposing preference.
+Negative social influence has the potential to produce polarization when individuals with similar views
+remain similar to each other, but adjust their views to more strongly oppose the rest of the group.
 We now describe two formal measures that use these assumptions to quantify how susceptible to influence and how potential influential an individual is.
@@
 In other words, participants altered their preferences to more strongly oppose those they interacted with.
+Negative social influence is of interest as a potential driver of polarization.
 
@@
 So while we do see more of an improvement in agreement under network deliberation, our observations are more consistent with a protective effect against a disagreement-producing mechanism, than with the originally hypothesized agreement-producing mechanism.
+The participants began with a relatively high level of agreement and disagreements were minor, so a more controversial topic may be necessary to identify potential conflict-resolution mechanisms.
 
@@
 
+While our experiment included both a control and experimental condition, having only one of each is a potential limitation.
+To veryify the robustness of our results, future work incorporating multiple control and treatment groups is needed.
+
 \section{Conclusion}

### 2022-08-08 UMich/chapters/ch05-discussion.tex
Ch05 revisions.

@@
 
-In the preceeding chapters, I have approached the above question using observational study, numerical simulation, and experiment.
-A combination of literature reivew and obersvational analysis of English-language Wikipedia (\ref{chap:wp-prod-perf}) identified a common theme among successful examples: small, tightly-knit groups, interconnected through mutual membership: network deliberation.
-Subsequent work investigates simple models of network deliberation to better understand how and why it helps large groups effectively self-govern.
-These models are not intended to capture all properties of real-world network deliberation, but rather to isolate and investigate particular factors, such as social influence and network efficiency.
-Both numerical simulations (Chapter \ref{chap:abm}) and experiment (Chapter \ref{chap:experiment}) find evidence that network deliberation can improve the effectiveness of large deliberations relative to conventional single-group deliberation.
-In particular, these studies both identify a protective effect against the negative consequences of social influence, such as information cascades.
-As such, network deliberation may be especially beneficial in settings with a high level of social influence.
+To answer the above question, I have introduced, developed, and applied a model of
+large-scale group deliberation: network deliberation.
+As a theoretical model, Network deliberation encapsulates common structural pattern
+observed across successful large-scale participatory projects:
+a network structure built from small interlocking pods.
+As with all theoretical models, many parameters are left variable.
+In the case of network deliberation, these parameters include:
+pod size, pod assignment strategy, and the behavioral/social dynamics of deliberation.
+This model can thus be used to analyze how such parameters influence the success of large-scale
+participatory projects while providing a common theoretical framework for making comparisons
+between different studies.
+
+In the preceeding chapters, I have shown how the network deliberation model can be
+used in the analysis of observational studies, numerical simulation, and experiment.
+A combination of literature reivew and obersvational analysis of English-language Wikipedia
+(\ref{chap:wp-prod-perf}) identified a common theme among successful examples:
+small, tightly-knit groups, interconnected through mutual membership: network deliberation.
+In Wikipedia for example, articles and their talk pages act as pods,
+creating venues for small group interactions, with interlocks created by editors who contribute
+to multiple articles.
+Subsequent work investigates simple models of network deliberation to better understand
+how and why it helps large groups effectively self-govern.
+Both numerical simulations (Chapter \ref{chap:abm}) and experiment (Chapter \ref{chap:experiment})
+find evidence that network deliberation can improve the effectiveness of large deliberations
+relative to conventional single-group deliberation.
+In particular, these studies both identify a protective effect against the negative
+consequences of social influence, such as information cascades.
+As such, these studies of network deliberation suggests an interlocking pod structure may be especially
+beneficial in settings susceptible to information cascades, such as those with a high
+level of social influence.
 
 \section{Future Work: Deliberative Settings}
-While the studies presented here confirm the observation that network deliberation sometimes contributes to the success of large-scale deliberation, there are many factors that can influence the deliberative process, and there remains much work to be done to understand how these factors interact with network deliberation.
+While the studies presented here confirm the observation that network deliberation
+sometimes contributes to the success of large-scale deliberation,
+there are many factors that can influence the deliberative process.
+It is difficult to predict whether or how these factors interact with network deliberation
+and there remains much work to be done.
 
@@
 \subsection{Facilitation and Moderation}
-When all members are equal participants in a deliberation, facilitation and moderation are a challenge.
-In the experiment reported in Chapter \ref{chap:experiment}, facilitation was achieved through discussion prompts.
-Moderation was performed by an experimenter, but the discussion remained respectful and no moderation actions were required.
-An active facilitator might contribute subtantially to the quality of a deliberation, but there are many potential ways to choose a facilitator from among a group of peers.
-Similarly, moderation is incredibly important for productive deliberation on sensitive topics, and it is not immediately obvious how moderation might work in a network deliberation environment.
+When all members are equal participants in a deliberation, facilitation and moderation are
+a challenge.
+In the experiment reported in Chapter \ref{chap:experiment}, facilitation was achieved
+through discussion prompts.
+Moderation was performed by an experimenter, but the discussion remained respectful and no
+moderation actions were required.
+An active facilitator might contribute subtantially to the quality of a deliberation,
+but there are many potential ways to choose a facilitator from among a group of peers.
+Similarly, moderation is incredibly important for productive deliberation on sensitive
+topics, and it is not immediately obvious how moderation might work in a network
+deliberation environment.
+Possibilities might include ad-hoc moderators, or bot/AI-asisted moderation.
 
@@
 In the simple model of network deliberation presented here, the available alternatives were determined before the beginning of the deliberation.
-However, the deliberative process can produce new ideas and these ideas are sometimes improvements on those previously available \cite{salganik_wiki_2015}.
+However, the deliberative process can produce new ideas and these ideas are sometimes
+improvements on those previously available \cite{salganik_wiki_2015}.
 Additional research is necessary to undertand the consequences of introducing new ideas into a network deliberation, as well as how and when to introduce those ideas to different pods for best effect.
@@
 be assigned to a pod with others sharing their preference, but in strong social influence settings, they might switch their preferences before then.
+
+\subsection{Next Steps}
+The future work needed most immediately is replication of experimental studies of network
+deliberation in various contexts.
+Such replications can test the robustness and generalizability of the results reported here.
+It is also desirable to develop experimental platforms that can accommodate emergent alternatives
+that arise during the deliberation, as such alternatives are often preferred to initial proposals
+\cite{salganik_wiki_2015}.
+The question of the role of structural efficiency also remains.
+Our numerical studies found that the effects of varying structural efficiency were much smaller
+than those caused by adopting network vs. conventional deliberation,
+but we have only evaluated structurally efficient networks empirically.
+
+Large-scale empirical studies present practical challenges, such as recruiting and providing incentives.
+One possible approach for overcoming these challenges in future studies is the development
+of a publicly-available platform that can be adopted by organizations for practical use,
+while simultaneously providing empirical data for study.
 
@@
 
-One possible application of network deliberation principles is in the design of organizational structures and procedures.
-In fact, small interlocking pods are a feature found in self-managed organizations \cite{laloux_reinventing_2014}.
-As the pracitices and structures of such organizations are codified and refined, a better understanding of network deliberation can provide a guide to consructing more effective organizations and institutions.
+One possible application of network deliberation principles is in the design of
+organizational structures and procedures.
+In fact, small interlocking pods are a feature found in self-managed organizations
+\cite{laloux_reinventing_2014}.
+As the practices and structures of such organizations are codified and refined,
+a better understanding of network deliberation can provide a guide to constructing
+more effective organizations and institutions.
+Simply recognizing the effectiveness of network deliberation structure suggests dividing
+larger teams into pods and ensuring that individuals are active in multiple pods or
+periodically rotate through them.
+Similarly, our studies suggest there may be benefits to having multiple pods working on the
+same tasks in parallel, particularly in the early stages of ideation,
+when the exploration/exploitation tradeoff is focused more on exploration.
+Furthermore, as the conesequences of parameters such as pod size and assignment methods are
+better understood, this understanding may inform the structure and composition of teams.
 
-Similarly, network deliberation emerges naturally in technical settings such as working groups in large collaborations (e.g., WikiProjects), federated social networks (e.g., mastodon), and group chat platforms (e.g., discord).
-The deliberation that takes place on these platforms is shaped in part by the protocols and user interaces of the software.
-The principles of network deliberation have the potential to inform the design of such protocols and interfaces in a way that minimizes harm and promotes effective interactions.
+Similarly, network deliberation emerges naturally in technical settings such as working
+groups in large collaborations (e.g., WikiProjects), federated social networks
+(e.g., mastodon), and group chat platforms (e.g., discord).
+The deliberation that takes place on these platforms is shaped in part by the protocols
+and user interfaces of the software.
+The principles of network deliberation have the potential to inform the design of such
+protocols and interfaces in a way that minimizes harm and promotes effective interactions.
+The success of network deliberation favors protocols that allow individuals to participate
+in multiple contexts, e.g., forums, microblogging servers, chat rooms.
+Similarly, our findings suggest the importance of being able to subdivide larger contexts
+into smaller pods.
 

### 2022-08-09 UMich/chapters/ch01-intro.tex
Dissertation revisions to abstract and ch01.

@@
 democratic governance, organizational effectiveness, social movements, and peer production.
-Faced with the intractability of large-scale decision-making, traditional
+Faced with the intractability of large-scale participatory decision-making, traditional
 systems have sacrificed one or more desirable properties, such as participation
@@
 to achieve the previously unachievable. Millions of volunteer Wikipedia editors
-have created a high-quality encyclopedia without centralized leadership
+have created a high-quality encyclopedia without formal centralized leadership
 \cite{keegan_evolution_2017, giles_internet_2005}.
@@
 Social movements such as the Arab Spring, Occupy, Black Lives Matter, and
-Podemos have reshaped national governments and brought attention to deeply
-entrenched social issues
+Podemos have brought attention to deeply entrenched social issues,
+and in some cases, contriubted to political regime changes
 \cite{tufekci_twitter_2017, gonzalez-bailon_networked_2016}.
@@
 such collaborations: decision-making.
-I specificaly examine how,
+Specifically, I examine how
 when groups are too large for all members to participate in all discussions,
@@
 
-In this dissertation, I combine observational, agent-based modeling, and experimental methods to evaluate the role of network structure in large-scale participatory decision-making.
-Taken together, the findings in these studies suggests that small, tightly-knit pods, interlocked through the sharing of members can benefit such collaborations.
-This common structure, which I refer to as network deliberation, appears to preserve some benefits of small-scale deliberation, while allowing efficient diffusion of information through very large groups.
-Furthermore, I find evidence that network deliberation can mitigate some of the negative aspects of social influence.
-These findings provide insight into the success of existing large-scale collaborations as well as provide guidance for the design of new sociotechnical systems.
+In this dissertation, I combine observational, agent-based modeling, and experimental
+methods to evaluate the role of network structure in large-scale participatory decision-making.
+I focus on group sizes well above the estimated 5--8 individual maximum for effective small goups
+\cite{freeman_tyranny_1972, lohman_designing_2000, miflin_small_2004},
+ranging from 32--33 participants in the experimental study to millions of editors in the
+observational study of English-language Wikipedia WikiProjects.
+Taken together, the findings in these studies suggests that small, tightly-knit pods,
+interlocked through the sharing of members can benefit such collaborations.
+This common structure, which I refer to as network deliberation, appears to preserve some
+benefits of small-scale deliberation, while allowing efficient diffusion of information
+through large groups.
+Furthermore, I find evidence that network deliberation can mitigate some of the negative
+aspects of social influence.
+These findings provide insight into the success of existing large-scale collaborations
+as well as provide guidance for the design of new sociotechnical systems.
 
@@
 Arrow's Impossibility Theorem \cite{arrow_social_2012}.
+While social choice theory typically assumes fixed fundamental preferences,
+preferences can be instrumental in nature,
+varying with factors such as available knowledge, group identity, or perspective.
 This dissertation finds hope in the transformative potential of the
@@
 through the greater collective.
+The network deliberation framework abstracts and generalizes phenomena
+such as interlocking directorates and interlocking publics,
+allowing their commonalities to be studied from the perspective of network structure.
+The framework also provides concrete parameters such as pod size and pod assignment method,
+which can be used in quantitative analysis of models and experiments,
+and provides a common language for making comparisons across models and experiments.
+Network deliberation is conceptually similar to network rotation in that both describe
+large-scale deliberation through repeated small-group interactions.
+There are two key differences, one conceptual, and one relating to network structure.
+Conceptually, network rotation reassigns pod membership one individual at a time in order
+to optimize certain network properties,
+while network deliberation focuses on providing an analytical framework for understanding
+the impact of different pod assignment methods on network structure and deliberation outcome.
+The one-at-a-time reassignments of network rotation also creates pods with high overlap,
+while network deliberation is focused on simultaneous or sequential pods with little overlap.
+
 In network deliberation, the method of assigning individuals to pods
@@
 Chapter \ref{chap:wp-prod-perf} describes an observational and computational
-study of WikiProjects on the English-language Wikipedia,
-and reports the following contributions:
+study of WikiProjects on the English-language Wikipedia.
+This study examines one of the real-world collaborations that inspired the network deliberation
+framework and shows that variations in the network structure of such a collaboration can
+correlate with quality indicators.
+Specifically, Chapter \ref{chap:wp-prod-perf} reports the following contributions:
 \begin{itemize}
@@
 \end{itemize}
+
+The above contributions suggest that network deliberation structure may be an important factor in the
+success of existing large-scale, internet-enabled collaborations, and furthermore,
+that it may help guide the design and organization of future collaborations.
+Specifically, network deliberation appears to help moderate the effects of social influence,
+providing some protection against both information cascades due to positive social influence,
+and polarization due to negative social influence.
+In summary, the work presented in the following chapters provides a framework for the
+future study of large-scale participatory governance,
+and provides evidence that large-scale participatory governance can be both practical and effective.

### 2022-08-09 UMich/front_materials/abstract.tex
Dissertation revisions to abstract and ch01.

@@
 Different methods for assigning members to pods result in different network topologies.
-This work combines observational study, agent-based modeling, and experiment to evaluate and better understand network deliberation.
-An empirical observation of WikiProjects on the English-language Wikipeida identifies the network properties of the most effective projects.
+This work combines observational study, agent-based modeling, and an experiment to evaluate and better understand network deliberation.
+An empirical observation of WikiProjects on the English-language Wikipedia identifies the network properties of the most effective projects.
 A simple agent-based model of Network Deliberation shows improvements over conventional deliberation in the presence of strong social influence.
 Finally, a controlled experiment studies network deliberation in a real world setting, tracking how individual preferences evolve, and finding evidence that network deliberation provides protection against negative consequences of social influence.
-This work seeks to improve the ability of very large groups to quickly develop a consensus, enabling a more effective use of shared resources to achieve common goals.
+This work seeks to provide a framework that enables an understanding of
+of how very large groups can quickly develop a consensus, enabling a more effective use of shared resources to achieve common goals.

### 2022-08-19 UMich/chapters/ch03-NetDelibABM.tex
Final version accepted by UMich.
Fix typos.

@@
 We allow each simulation to proceed for 300 iterations, which we have found sufficient to guarantee convergence with our chosen parameters.
-For real-world comparison, on average, an active English-language Wikipedia editor makes about 45 edits per month \cite{noauthor_wikipediastatistics_2022}.
 
@@
 
-One of the strongest capabilities an agent might posses, is to compare two arbitrary solutions to determine which yields a higher value of the objective function.
+One of the strongest capabilities an agent might posess, is to compare two arbitrary solutions to determine which yields a higher value of the objective function.
 We represent this capability using the {\em arbitrary comparison} oracle.

### 2022-08-19 UMich/front_materials/abstract.tex
Fix typos.

@@
 This work seeks to provide a framework that enables an understanding of
-of how very large groups can quickly develop a consensus, enabling a more effective use of shared resources to achieve common goals.
+how very large groups can quickly develop a consensus, enabling a more effective use of shared resources to achieve common goals.

### 2023-02-16 UMich/front_materials/abstract.tex
Change Networked Deliberation to Network Deliberation.

@@
 Arrow’s impossibility theorem also fundamentally limits the fairness of voting.
-This project proposes Networked Deliberation as a potential means of effective large-scale decision-making.
-In Networked Deliberation, members of a large group are repeatedly partitioned into small deliberative pods.
+This project proposes Network Deliberation as a potential means of effective large-scale decision-making.
+In Network Deliberation, members of a large group are repeatedly partitioned into small deliberative pods.
 Overlap between pods at different stages enables group-wide diffusion of information and preferences.
